### 2022-11-30 report/report.tex
init repo

\documentclass[5p]{elsarticle}

\journal{TBA}

\bibliographystyle{elsarticle-num}
\biboptions{numbers,sort&compress}

% format hacks
\usepackage{libertine}
\usepackage{libertinust1math}
% \usepackage{geometry}
% \geometry{
%     top=30mm,
%     bottom=35mm,
% }

\usepackage{amsmath}
\usepackage{bbold}
\usepackage{graphicx}
\usepackage{eurosym}
\usepackage{mathtools}
\usepackage{url}
\usepackage{booktabs}
\usepackage{epstopdf}
\usepackage{xfrac}
\usepackage{tabularx}
\usepackage{bm}
\usepackage{subcaption}
\usepackage{blindtext}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{threeparttable}
\usepackage{pdflscape}
\usepackage[export]{adjustbox}
\usepackage[version=4]{mhchem}
\usepackage[colorlinks]{hyperref}
\usepackage[parfill]{parskip}
\usepackage[nameinlink,sort&compress,capitalise,noabbrev]{cleveref}
\usepackage[leftcaption,raggedright]{sidecap}
\usepackage[prependcaption,textsize=footnotesize]{todonotes}

\usepackage{siunitx}
\sisetup{
	range-units = single,
	per-mode = symbol
}
\DeclareSIUnit\year{a}
\DeclareSIUnit{\tco}{t_{\ce{CO2}}}
\DeclareSIUnit{\sieuro}{\mbox{\euro}}

\usepackage{lipsum}

\usepackage[resetlabels,labeled]{multibib}
\newcites{S}{Supplementary References}
\bibliographystyleS{elsarticle-num}

\graphicspath{
    {../results/graphics/},
}

% \usepackage[
% 	type={CC},
% 	modifier={by},
% 	version={4.0},
% ]{doclicense}

\newcommand{\abs}[1]{\left|#1\right|}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\set}[1]{\left\{#1\right\} }
\DeclareMathOperator*{\argmin}{\arg\!\min}
\newcommand{\R}{\mathbb{R} }
\newcommand{\B}{\mathbb{B} }
\newcommand{\N}{\mathbb{N} }
\newcommand{\co}{\ce{CO2}~}
\def\el{${}_{\textrm{el}}$}
\def\th{${}_{\textrm{th}}$}
\def\deg{${}^\circ$}

\begin{document}

\begin{frontmatter}

	\title{Competing networks: Co-optimizing Hydrogen and Carbon Management for a Climate-neutral German Energy System}

	\author[affil]{ Fabian Hofmann, Christoph Tries, Tom Brown } % \corref{correspondingauthor}
	\ead{ m.hofmann@tu-berlin.de }

	\address[affil]{ Department of Digital Transformation in Energy Systems, Institute of Energy Technology, Technische Universität Berlin, Fakultät III, Einsteinufer 25 (TA 8), 10587 Berlin, Germany }

	\begin{abstract}
		\input{sections/abstract.tex}
	\end{abstract}

	\begin{keyword}
		TODO
	\end{keyword}

	% \begin{graphicalabstract}
	% \end{graphicalabstract}

	\begin{highlights}
		\item A
		\item B
		\item C
	\end{highlights}

\end{frontmatter}

% \listoftodos[TODOs]

% \tableofcontents

\section{Introduction}
\label{sec:intro}

\input{sections/introduction.tex}

\section{Methods}
\label{sec:methods}

\input{sections/methods.tex}

\section{Results}
\label{sec:results}

\input{sections/results.tex}

\section{Discussion}
\label{sec:discussion}

\input{sections/discussion.tex}

\section{Conclusion}
\label{sec:conclusion}

\input{sections/conclusion.tex}

\section*{Acknowledgements}

% XY gratefully acknowledge funding from XY under grant no. 00000.

\section*{License}
% \doclicenseLongText
% \doclicenseIcon

\section*{Author Contributions}

% following https://casrai.org/credit/

\textbf{Author XY}:
Conceptualization --
Data curation --
Formal Analysis --
Funding acquisition --
Investigation --
Methodology --
Project administration --
Resources --
Software --
Supervision --
Validation --
Visualization --
Writing - original draft --
Writing - review \& editing

\section*{Data and Code Availability}

A dataset of the model results is available at \href{zenodoTBA}{doi:zenodoTBA}.
The code to reproduce the experiments is available at 
\href{https://github.com/https://github.com/FabianHofmann/co2-network.git}{github.com/https://github.com/FabianHofmann/co2-network.git}.

% tidy with https://flamingtempura.github.io/bibtex-tidy/
\addcontentsline{toc}{section}{References}
\renewcommand{\ttdefault}{\sfdefault}
\bibliography{references}

% supplementary information

\newpage

\makeatletter
\renewcommand \thesection{S\@arabic\c@section}
\renewcommand\thetable{S\@arabic\c@table}
\renewcommand \thefigure{S\@arabic\c@figure}
\makeatother
\renewcommand{\citenumfont}[1]{S#1}
\setcounter{equation}{0}
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{section}{0}

\section{Supplementary Information}
\label{sec:si}

\input{sections/supplementary.tex}

\addcontentsline{toc}{section}{Supplementary References}
\renewcommand{\ttdefault}{\sfdefault}
\bibliographyS{references}

\end{document}

### 2022-11-30 report/sections/abstract.tex
init repo

The transformation of the global energy system to reach climate-neutrality requires large investments into new technologies. \dots

### 2022-11-30 report/sections/introduction.tex
init repo

\textbf{Literature}

\begin{itemize}
    \item CCUS or CCTUS seem to be covered mainly by IAM community, not energy system modelling. 
    \item \cite{CaptureMapGetSitespecific} Recent Swiss study on CCTS
    \item \cite{weiProposedGlobalLayout2021} Chinese Nature study on global CCS potential: 92 GtCO$_2$ mitigated globally of which 59 Gt will be sequestered. All at cost of 0.12\% of global GDP   
    \item \cite{ToolsGreenTransition} Endravas Capture Map estimates 1.4 Gt capturing potential from point sources in Europe.
    \item SimCCS Optimization software for MIP problems with infrastructure decision \cite{middletonSimCCSOpensourceTool2020}
\end{itemize}

\textbf{Companies and projects}
\begin{itemize}
    \item Potential Carbon Management hubs are evolving Tree Energy, Carbfix, Equinor    
    \item Mission Innovation Hydrogen Valley Platform \cite{H2ValleysMissionInnovation} collects hydrogen flagship projects to facilitate 
    \item European Hydrogen Backbone \cite{gasforclimateEuropeanHydrogenBackbone2022} 
\end{itemize}

\textbf{Legislation}
\begin{itemize}
    \item German climate protection act \cite{KlimaschutzgesetzKlimaneutralitaetBis} targets climate neutrality by 2045, and 65\% CO$_2$ reduction by 2030. (762 Mt CO2-eq emitted in 2021; 438 Mt targeted for 2030) 
    \item Fit-for-55 package by EU targets 55\% reduction by 2030 in the EU
    \item REpowerEU \cite{REPowerEU} targets accelerated energy transition and security. A fast ramp up of a green Hydrogen system should support transition of hard-to-abate sectors. Domestic production should be 10 MT H$_2$ by 2030, same as import. Additional funding for Clean Hydrogen Partnership through Horizon Europe Programme. 
    \item Delegated Act for green H$_2$
    \item 
\end{itemize}

\textbf{Research Questions}

\begin{itemize}
    \item How to optimally supply cities far from renewable wind production centers in throughout the whole year?    
    \item Where is synergy and where competition between H$_2$ and CO$_2$ management?
    \item Where is one technology in advantage?
    \item DAC vs. point source CC?
    \item Where to draw the line between small and large scale CO$_2$ transport?
\end{itemize}

\textbf{Background Research}

\begin{itemize}
    \item Economy of scale? What are the learning curves to expect?
\end{itemize}

### 2022-11-30 report/sections/methods.tex
init repo

Energy system modelling in PyPSA-EUR-sec \cite{brownSynergiesSectorCoupling2018} with version

### 2022-11-30 playaround/plot-networks-map.py

@@ -1,0 +1,29 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Nov 24 10:14:48 2022
+
+@author: fabian
+"""
+
+import pypsa
+import matplotlib.pyplot as plt
+import cartopy.crs as ccrs
+
+
+n = pypsa.Network("/home/fabian/papers/co2-network/workflow/subworkflows/pypsa-eur-sec/results/your-run-name/postnetworks/elec_s_5_lv1.0__Co2L0-3H-T-H-B-I-A-solar+p3-dist1_2050.nc")
+
+stats = n.statistics()
+
+# %% Plot networks
+
+kinds = list(filter(lambda s: "pipeline" in s, n.links.carrier.unique()))
+kinds.append("DC")
+
+fig, axes = plt.subplots(len(kinds), subplot_kw={"projection", ccrs.EqualEarth()})
+
+for kind, ax in zip(kinds, axes):
+
+    links = n.
+
+    n.plot(ax=ax, )

### 2023-02-07 report/sections/abstract.tex
add plotting scripts

@@
-The transformation of the global energy system to reach climate-neutrality requires large investments into new technologies. \dots
+Enabling a carbon-neutral European economy requires internationally coordinated investments into all sectors of the energy system.
+In the light of hard-to-abate emissions from process industry, energy-intensive transport and backup heat or power sources, it is important to determine the extent to which a system-wide carbon management is needed.
+Different technologies such as direct air capture, carbon capture at point sources and carbon transport, utilization in hydrocarbon synthesis and storage are considered as valuable contributors to formulate a comprehensive carbon strategy and take the final step to climate-neutrality.
+While the literature has already showed the synergies of these technologies, the question of how to design the infrastructure at high resolution together with a CO$_2$ transport system has not been addressed so far.
+Moreover, there is a growing number of individually planned carbon management projects whereas the formulation of a European carbon strategy is still missing from the political agenda.
+In this paper, we present the first study of a cost-optimal design of the European energy system with a full representation of carbon management technologies including a CO$_2$ transport network, considering all sectors with high spatial resolution.
+In addition, the system is allowed to optimize the layout of renewable energy sources as well as energy transmission and storage systems based on hydrogen and methane.
+Assuming net-zero emissions, we show that carbon management technologies provide critical cost-beneficial flexibilities to the European energy system.
+In particular, we show that building a CO$_2$ network complements the built out of a hydrogen network and is cost-effective across different rates of carbon sequestration in the offshore regions (100 -- 1000 Mt/a).
+Combined with a hydrogen grid, the CO$_2$ grid leads to cost savings of about 6\%, as it allows industrial and densely populated inland sites to source hydrogen from coastal regions with favorable renewable energy potential and, in return, supply CO$_2$ that can be sequestered or utilized for synthetic hydrocarbons for petrochemical feedstocks, aviation and shipping.
+At high sequestration rates, direct air capture facilities complement carbon capture technologies at point sources, leading however to higher use of fossil energy carriers in the power and heating sector.
+The optimization identifies three main carbon sinks that dominate and determine the CO$_2$ network design, namely a large sequestration cluster in the North Sea and two clusters on the coasts of Portugal and Greece.
+The results highlight the potential of a comprehensive carbon management strategy as a basis for designing future carbon networks and identifying the most promising sites for carbon capture and storage.

### 2023-02-07 workflow/scripts/co2scenarios.py

@@ -1,0 +1,681 @@
+#!/usr/bin/env python
+# coding: utf-8
+
+import os
+import re
+import sys
+from itertools import product
+from pathlib import Path
+
+import cartopy.crs as ccrs
+import geopandas as gpd
+import matplotlib as mpl
+import matplotlib.cm as cm
+import matplotlib.patches as mpatches
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+from pandas import IndexSlice as idx
+import xarray as xr
+import yaml
+from matplotlib.lines import Line2D
+from matplotlib.patches import FancyArrowPatch
+from matplotlib.transforms import Bbox
+from vresutils.costdata import annuity
+
+
+PATH = Path("/home/fabian/papers/co2-network/workflow/subworkflows/pypsa-eur-sec")
+
+sys.path.append(
+    "/home/fabian/papers/co2-network/workflow/subworkflows/pypsa-eur-sec/scripts"
+)
+from plot_summary import rename_techs
+
+CLUSTERS = 181
+OUTPUT = "20230125-carbon-small/"
+
+MAIN_SCENARIOS = Path("20230125-carbon-small")
+
+plt.style.use(["bmh", "matplotlibrc"])
+
+with open(MAIN_SCENARIOS / "configs/config.yaml") as file:
+    config = yaml.safe_load(file)
+
+if not os.path.exists(OUTPUT):
+    os.makedirs(OUTPUT)
+
+
+def rename_techs_tyndp(tech):
+    tech = rename_techs(tech)
+    if "heat pump" in tech or "resistive heater" in tech:
+        return "power-to-heat"
+    elif tech in ["H2 Electrolysis"]:  # , "H2 liquefaction"]:
+        return "power-to-hydrogen"
+    elif "H2 pipeline" in tech:
+        return "H2 pipeline"
+    elif tech == "H2":
+        return "H2 storage"
+    elif tech in ["OCGT", "CHP", "gas boiler", "H2 Fuel Cell"]:
+        return "gas-to-power/heat"
+    # elif "solar" in tech:
+    #    return "solar"
+    elif tech in ["Fischer-Tropsch", "methanolisation"]:
+        return "power-to-liquid"
+    elif "offshore wind" in tech:
+        return "offshore wind"
+    elif "SMR" in tech:
+        return tech.replace("SMR", "steam methane reforming")
+    elif "DAC" in tech:
+        return "direct air capture"
+    elif "CC" in tech or "sequestration" in tech:
[…]

### 2023-02-07 workflow/scripts/common.py

@@ -1,0 +1,159 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Wed Jan 18 13:55:01 2023
+
+@author: fabian
+"""
+
+from pathlib import Path
+import warnings
+import pypsa
+from pypsa.descriptors import Dict
+from pypsa.components import components, component_attrs
+import pandas as pd
+import yaml
+import numpy as np
+from constants import SITES, NODE_COMPS
+
+root = Path("/home/fabian/papers/co2-network")
+pypsa_eur = root / "workflow/subworkflows/pypsa-eur"
+pypsa_eur_sec = root / "workflow/subworkflows/pypsa-eur-sec"
+override_dir = pypsa_eur_sec / "data/override_component_attrs"
+
+
+def import_network(path):
+    overrides = override_component_attrs(override_dir)
+    n = pypsa.Network(path, override_component_attrs=overrides)
+    sanitize_locations(n)
+    fill_missing_carriers(n)
+    modify_carrier_names(n)
+    add_carrier_nice_names(n)
+    add_colors(n)
+    if n.carriers.notnull().all().all() and (n.carriers != "").all().all():
+        warnings.warn("Some carriers have no color or nice_name")
+    n.carriers = n.carriers.sort_values(["color"])
+    return n
+
+
+def get_transmission_links(n):
+    # only choose transmission links
+    return n.links.bus0.map(n.buses.location) != n.links.bus1.map(n.buses.location)
+
+
+def get_generation_carriers(n):
+    carriers = []
+    for c in NODE_COMPS:
+        cars = n.df(c).carrier
+        if c == "Link":
+            cars = cars[~get_transmission_links(n)]
+        carriers.append(set(cars))
+    return list(set.union(*carriers))
+
+
+def assign_location(n):
+    for c in n.one_port_components | n.branch_components:
+        df = n.df(c)
+
+        if "location" not in df:
+            df["location"] = np.nan
+
+        ifind = pd.Series(df.index.str.find(" ", start=4), df.index)
+        for i in ifind.value_counts().index:
+            # these have already been assigned defaults
+            if i == -1:
+                continue
+            names = ifind.index[ifind == i]
+            df.loc[names, "location"] = names.str[:i]
+
+        bus_col = df.columns[df.columns.str.startswith("bus")][0]
+        df.location.fillna(df[bus_col], inplace=True)
+
+
+def assign_interconnection(n):
[…]

### 2023-02-07 workflow/scripts/constants.py

@@ -1,0 +1,57 @@
+KINDS = [
+    "electricity",
+    "gas",
+    "hydrogen",
+    "carbon",
+]
+ONEPORT_COMPS = ["Generator", "Store", "StorageUnit"]
+NODE_COMPS = ["Generator", "Link", "StorageUnit"]
+REGION_COMPS = ["Store"]
+BRANCH_COMPS = ["Line", "Link"]
+COMPS = set(NODE_COMPS + REGION_COMPS + BRANCH_COMPS)
+SITES = [
+    "residential urban decentral",
+    "services urban decentral",
+    "residential rural",
+    "services rural",
+    "services urban",
+    "urban central",
+]
+HEATING_TECHS = ["heat", "water tanks", "solar thermal"]
+CARRIER_TO_BUSES = {
+    "carbon": ["co2 stored"],
+    "hydrogen": ["H2"],
+    "electricity": ["AC", "DC"],
+    "gas": ["gas"],
+}
+
+PLOT_SPECS = {
+    "carbon": {
+        "region_cmap": "Reds",
+        "bus_scale": 1e-4,
+        "branch_scale": 5e-3,
+        "bus_sizes": [5_000, 10_000],
+        "branch_sizes": [1_000, 2_000],
+    },
+    "hydrogen": {
+        "region_cmap": "Blues",
+        "bus_scale": 4e-5,
+        "branch_scale": 3e-3,
+        "bus_sizes": [5_000, 10_000],
+        "branch_sizes": [1_000, 2_000],
+    },
+    "electricity": {
+        "region_cmap": "Greens",
+        "bus_scale": 5e-6,
+        "branch_scale": 2e-4,
+        "bus_sizes": [20_000, 50_000],
+        "branch_sizes": [5_000, 10_000],
+    },
+    "gas": {
+        "region_cmap": "Purples",
+        "bus_scale": 2e-6,
+        "branch_scale": 1e-4,
+        "bus_sizes": [100_000, 200_000],
+        "branch_sizes": [10_000, 20_000],
+    },
+}

### 2023-02-07 workflow/scripts/helpers.py

@@ -1,0 +1,284 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Wed Jan 18 13:55:01 2023
+
+@author: fabian
+"""
+
+# Sources and sinks
+# Focussing on CO2 network might be enough (turn on/off)
+# How to handle import of H2?
+
+
+from pathlib import Path
+import warnings
+import pypsa
+from pypsa.descriptors import Dict
+from pypsa.components import components, component_attrs
+import pandas as pd
+import yaml
+import numpy as np
+
+root = Path("/home/fabian/papers/co2-network")
+pypsa_eur = root / "workflow/subworkflows/pypsa-eur"
+pypsa_eur_sec = root / "workflow/subworkflows/pypsa-eur-sec"
+override_dir = pypsa_eur_sec / "data/override_component_attrs"
+
+SITES = [
+    "residential urban decentral",
+    "services urban decentral",
+    "residential rural",
+    "services rural",
+    "services urban",
+    "urban central",
+]
+HEATING_TECHS = ["heat", "water tanks", "solar thermal"]
+
+
+CARRIER_TO_BUSES = {
+    "carbon": ["co2 stored"],
+    "hydrogen": ["H2"],
+    "electricity": ["AC", "DC"],
+    "gas": ["gas"],
+}
+
+KIND_TO_CMAP = {
+    "carbon": "Reds",
+    "hydrogen": "Blues",
+    "electricity": "Greens",
+    "gas": "Purples",
+}
+
+
+ONEPORT_COMPS = ["Generator", "Store", "StorageUnit"]
+NODE_COMPS = ["Generator", "Link", "StorageUnit"]
+REGION_COMPS = ["Store"]
+BRANCH_COMPS = ["Line", "Link"]
+COMPS = set(NODE_COMPS + REGION_COMPS + BRANCH_COMPS)
+
+
+def create_carrier_network(n, kind, include_eu=False):
+    carriers = CARRIER_TO_BUSES[kind]
+    buses = n.buses.query("carrier in @carriers").index
+
+    o = pypsa.Network(snapshots=n.snapshots)
+    o.madd("Bus", n.buses.index, **n.buses, marginal_price=n.buses_t.marginal_price)
+    o.buses.loc[buses, "kind"] = True
+    o.buses.kind.fillna(False, inplace=True)
+
+    for c in ONEPORT_COMPS:
[…]

### 2023-02-07 workflow/scripts/leftover.py

@@ -1,0 +1,178 @@
+CARRIER_TO_PLOT = {
+    "carbon": {
+        "node_carrier": [
+            "process emissions CC",
+            "solid biomass for industry CC",
+            "SMR CC",
+            "gas for industry CC",
+            "solid biomass CHP CC",
+            "gas CHP CC",
+            "DAC",
+        ],
+        "region_carrier": ["co2 stored"],
+        "branch_carrier": ["CO2 pipeline"],
+    },
+    "hydrogen": {
+        "node_carrier": ["H2 Electrolysis"],
+        "region_carrier": ["H2"],
+        "branch_carrier": ["H2 pipeline"],
+    },
+    "electricity": {
+        "node_carrier": [
+            "solar",
+            "offwind",
+            "onwind",
+            "ror",
+            "hydro",
+            "PHS",
+            "OCGT",
+            "H2 Fuel Cell",
+            "CCGT",
+            "geothermal",
+            "biomass",
+        ],
+        "branch_carrier": ["AC", "DC"],
+        "region_carrier": ["battery", "home battery", "Li ion"],
+    },
+    "gas": {
+        "node_carrier": ["gas"],
+        "branch_carrier": ["gas pipeline"],
+        "region_carrier": ["gas", "biogas"],
+    },
+}
+
+
+CARRIER_GROUPS = {
+    "Battery Infrastructure": [
+        "battery",
+        "battery discharger",
+        "battery charger",
+        "Li ion",
+        "home battery",
+    ],
+    "Fossil Carriers": ["coal", "CCGT", "lignite", "OCGT"],
+    "Hydrogen Infrastructure": [
+        "H2",
+        "H2 fuel cell",
+        "H2 electrolysis",
+        "H2 Electrolysis",
+    ],
+    "Nuclear": ["nuclear"],
+    "Offshore Wind": ["offwind"],
+    "Onshore Wind": ["onwind"],
+    "Other Renewables": [
+        "geothermal",
+        "biomass",
+        "PHS",
+        "hydro",
+        "ror",
+        "biogas",
+        "solid biomass",
+    ],
+    "Solar": ["solar"],
+    "Gas": ["gas"],
+    "Transmission System": ["AC", "DC"],
+    "Heat": [f"{site} heat" for site in SITES]
+    + [
+        "heat",
+        "air heat pump",
+        "ground heat pump",
+        "biomass boiler",
+        "gas boiler",
+        "oil boiler",
+        "wood boiler",
+        "solar thermal",
[…]

### 2023-02-07 workflow/scripts/plot-networks-map.py

@@ -1,0 +1,79 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Nov 24 10:14:48 2022
+
+@author: fabian
+"""
+
+# Sources and sinks
+# Focussing on CO2 network might be enough (turn on/off)
+# How to handle import of H2?
+
+
+import seaborn as sns
+import pandas as pd
+import matplotlib.pyplot as plt
+import cartopy.crs as ccrs
+from common import import_network, pypsa_eur_sec, pypsa_eur
+from constants import PLOT_SPECS, KINDS
+
+from plotting import (
+    create_carrier_network,
+    plot_map,
+    add_legend,
+    get_carrier_network_plotting_data,
+)
+import geopandas as gpd
+
+
+sns.set_context("paper")
+kind = "carbon"
+column = "Optimal Capacity"
+alpha = 1
+region_alpha = 0.8
+
+
+n = import_network(
+    pypsa_eur_sec
+    / "results/your-run-name/postnetworks/elec_s_10_lv1.2__Co2L0-T-H-B-I-A-solar+p3-dist1_2050.nc"
+)
+regions = gpd.read_file(
+    pypsa_eur / "resources/regions_onshore_elec_s_10.geojson"
+).set_index("name")
+
+fig, axes = plt.subplots(
+    2, 2, figsize=(15, 13), subplot_kw={"projection": ccrs.EqualEarth()}
+)
+
+for kind, ax in zip(KINDS, axes.flatten()):
+
+    o = create_carrier_network(n, kind=kind)
+    data = get_carrier_network_plotting_data(o, "Optimal Capacity")
+    specs = PLOT_SPECS[kind]
+
+    plot_map(
+        ax,
+        o,
+        regions,
+        column=column,
+        bus_scale=specs["bus_scale"],
+        branch_scale=specs["branch_scale"],
+        alpha=alpha,
+        region_alpha=region_alpha,
+        region_cmap=specs["region_cmap"],
+    )
+    add_legend(
+        ax,
+        o,
+        bus_scale=specs["bus_scale"],
+        branch_scale=specs["branch_scale"],
+        bus_sizes=specs["bus_sizes"],
+        branch_sizes=specs["branch_sizes"],
+        alpha=alpha,
+    )
+    ax.set_extent(regions.total_bounds[[0, 2, 1, 3]])
+    ax.set_title(kind.title())
+
+fig.tight_layout()
[…]

### 2023-02-07 workflow/scripts/plotting.py

@@ -1,0 +1,200 @@
+import pandas as pd
+import pypsa
+from pypsa.descriptors import Dict
+from pypsa.plot import add_legend_circles, add_legend_lines, add_legend_patches
+from constants import CARRIER_TO_BUSES, ONEPORT_COMPS, BRANCH_COMPS, COMPS
+from common import get_transmission_links, get_generation_carriers
+from cartopy import crs as ccrs
+
+
+def plot_map(
+    ax,
+    n,
+    regions,
+    column="Optimal Capacity",
+    bus_scale=1.5e-5,
+    branch_scale=3e-3,
+    alpha=1,
+    region_alpha=0.8,
+    region_cmap="Blues",
+):
+    d = get_carrier_network_plotting_data(n, column=column)
+
+    n.bus_scale = bus_scale
+    n.branch_scale = branch_scale
+    n.alpha = alpha
+
+    # with plt.rc_context({"patch.linewidth": 0.0}):
+    n.plot(
+        bus_sizes=d.bus_sizes * bus_scale,
+        bus_alpha=alpha,
+        line_widths=d.line_widths * branch_scale,
+        link_widths=d.link_widths * branch_scale,
+        link_colors=d.branch_color,
+        line_colors=d.branch_color,
+        line_alpha=alpha,
+        link_alpha=alpha,
+        ax=ax,
+        margin=0.2,
+        color_geomap=None,
+        geomap="10m",
+    )
+    regions = regions.assign(color=d.region_color / 1e3)
+    regions.plot(
+        ax=ax,
+        column="color",
+        cmap=region_cmap,
+        vmin=0,
+        alpha=region_alpha,
+        facecolor="whitesmoke",
+        edgecolor="black",
+        aspect="equal",
+        transform=ccrs.PlateCarree(),
+        linewidth=0.3,
+        legend=True,
+        legend_kwds={
+            "label": n.kind.title() + " storage capacity [GWh]",
+            "orientation": "horizontal",
+            "shrink": 0.8,
+            "pad": 0.05,
+            "aspect": 30,
+            "alpha": region_alpha,
+        },
+    )
+
+
+def add_legend(
+    ax,
+    n,
+    bus_sizes=[10_000, 5_000],
+    branch_sizes=[2_000, 1_000],
+    bus_scale=None,
+    branch_scale=None,
+    alpha=None,
+):
+    legend_kwargs = {"loc": "upper left", "frameon": False}
+
+    add_legend_circles(
+        ax,
[…]

### 2023-02-23 report/sections/methods.tex
add plotting scripts
add cluster config

@@
-Energy system modelling in PyPSA-EUR-sec \cite{brownSynergiesSectorCoupling2018} with version
+Energy system modelling in PyPSA-EUR-sec~\cite{brownSynergiesSectorCoupling2018}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{sequestration_map.pdf}
+    \caption{Sequestration map}
+    \label{fig:sequestration_map}
+\end{figure}
+
+\begin{figure*}[h]
+    \centering
+    \includegraphics[width=\linewidth]{capacity_map_co2network_1000.pdf}
+    \caption{Optimal capacities per sector for a sequestration of 1000 Mt/a.}
+    \label{fig:capacity_map_noco2network_1000}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_1000.pdf}
+    \caption{Optimal operation per sector for a sequestration of 1000 Mt/a.}
+    \label{fig:operation_map_noco2network_1000}
+\end{figure*}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{operation_area_carbon_co2network.pdf}
+    \caption{Balance of captured CO$_2$ emissions for the optimal operation with CO$_2$ network.}
+    \label{fig:operation_area_carbon_co2network}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{operation_area_carbon_noco2network.pdf}
+    \caption{Balance of captured CO$_2$ emissions for the optimal operation without CO$_2$ network.}
+    \label{fig:operation_area_carbon_noco2network}
+\end{figure}

### 2023-02-24 workflow/scripts/common.py
adjust layouts to paper with

@@ -48,6 +48,27 @@ def import_network(path):
     return n
 
 
+def group_small_contributions(df, threshold=0.01):
+    if isinstance(df, pd.DataFrame):
+        is_small = df.sum() < threshold
+        if is_small.sum() <= 1:
+            return df
+        small = df.loc[:, is_small]
+        df = df.loc[:, ~is_small]
+        df = df.assign(Other=small.sum(1))
+        return df
+    elif isinstance(df, pd.Series):
+        is_small = df < threshold
+        if is_small.sum() <= 1:
+            return df
+        small = df[is_small]
+        df = df[~is_small]
+        df.assign(Other=small.sum())
+        return df
+    else:
+        raise TypeError("df must be a DataFrame or Series")
+
+
 def assert_carriers_existent(n, carriers, c):
     if not set(carriers).issubset(n.df(c).carrier.unique()):
         logger.warning(
@@ -387,6 +408,7 @@ def fill_missing_carriers(n):
 
 
 def modify_carrier_names(n):
+    n.add("Carrier", "other", nice_name="Other", color="lightgrey")
     n.add("Carrier", "offwind", nice_name="Offshore Wind", color="#6895dd")
     n.mremove("Carrier", ["offwind-ac", "offwind-dc"])
     n.generators.loc[

### 2023-02-24 workflow/scripts/plot-operation-area.py

@@ -5,6 +5,7 @@ from common import (
     mock_snakemake,
     get_carrier_consumption,
     get_carrier_production,
+    group_small_contributions,
 )
 import pandas as pd
 import seaborn as sns
@@ -17,7 +18,7 @@ if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
         "plot_operation_area",
         design="co2network",
-        kind="gas",
+        kind="electricity",
     )
 
 
@@ -62,10 +63,10 @@ for k, output in zip(keys, snakemake.output):
         df.loc["Consumption", k].rename(index=nice_name).div(norm).pipe(sort_by_color).T
     )
 
-    production = production.loc[:, production.sum() > 0.1]
-    consumption = consumption.loc[:, consumption.sum() > 0.1]
+    production = group_small_contributions(production, 1)
+    consumption = group_small_contributions(consumption, 1)
 
-    fig, ax = plt.subplots(1, 1, figsize=(8, 5))
+    fig, ax = plt.subplots(1, 1, figsize=(6, 5))
     production.plot(
         kind="area", stacked=True, ax=ax, color=colors.to_dict(), alpha=0.8, lw=0
     )
@@ -86,10 +87,10 @@ for k, output in zip(keys, snakemake.output):
     plabels, phandles = [], []
     clabels, chandles = [], []
     for label, handle in zip(labels[::-1], handles[::-1]):
-        if label in pcarriers:
+        if label in pcarriers and label not in plabels:
             plabels.append(label)
             phandles.append(handle)
-        if label in ccarriers:
+        if label in ccarriers and label not in clabels:
             clabels.append(label)
             chandles.append(handle)
 
@@ -97,11 +98,12 @@ for k, output in zip(keys, snakemake.output):
         phandles,
         plabels,
         loc="upper left",
-        bbox_to_anchor=(1, 1),
+        bbox_to_anchor=(1, 1.05),
         frameon=False,
-        ncol=2,
+        ncol=1,
         title="Production",
         labelcolor="k",
+        labelspacing=0.3,
     )
     fig.add_artist(legend)
 
@@ -109,11 +111,12 @@ for k, output in zip(keys, snakemake.output):
         chandles[::-1],
         clabels[::-1],
         loc="lower left",
-        bbox_to_anchor=(1, 0),
+        bbox_to_anchor=(1, -0.05),
         frameon=False,
-        ncol=2,
[…]

### 2023-03-07 report/sections/introduction.tex
update paper

@@
     \item CCUS or CCTUS seem to be covered mainly by IAM community, not energy system modelling.
-    \item \cite{CaptureMapGetSitespecific} Recent Swiss study on CCTS
-    \item \cite{weiProposedGlobalLayout2021} Chinese Nature study on global CCS potential: 92 GtCO$_2$ mitigated globally of which 59 Gt will be sequestered. All at cost of 0.12\% of global GDP   
-    \item \cite{ToolsGreenTransition} Endravas Capture Map estimates 1.4 Gt capturing potential from point sources in Europe.
-    \item SimCCS Optimization software for MIP problems with infrastructure decision \cite{middletonSimCCSOpensourceTool2020}
+    \item ~\cite{CaptureMapGetSitespecific} Recent Swiss study on CCTS
+    \item ~\cite{weiProposedGlobalLayout2021} Chinese Nature study on global CCS potential: 92 GtCO$_2$ mitigated globally of which 59 Gt will be sequestered. All at cost of 0.12\% of global GDP
+    \item ~\cite{ToolsGreenTransition} Endravas Capture Map estimates 1.4 Gt capturing potential from point sources in Europe.
+    \item SimCCS Optimization software for MIP problems with infrastructure decision ~\cite{middletonSimCCSOpensourceTool2020}
 \end{itemize}
@@
     \item Potential Carbon Management hubs are evolving Tree Energy, Carbfix, Equinor
-    \item Mission Innovation Hydrogen Valley Platform \cite{H2ValleysMissionInnovation} collects hydrogen flagship projects to facilitate 
-    \item European Hydrogen Backbone \cite{gasforclimateEuropeanHydrogenBackbone2022} 
+    \item Mission Innovation Hydrogen Valley Platform ~\cite{H2ValleysMissionInnovation} collects hydrogen flagship projects to facilitate
+    \item European Hydrogen Backbone ~\cite{gasforclimateEuropeanHydrogenBackbone2022}
 \end{itemize}
@@
 \begin{itemize}
-    \item German climate protection act \cite{KlimaschutzgesetzKlimaneutralitaetBis} targets climate neutrality by 2045, and 65\% CO$_2$ reduction by 2030.
+    \item German climate protection act ~\cite{KlimaschutzgesetzKlimaneutralitaetBis} targets climate neutrality by 2045, and 65\% CO$_2$ reduction by 2030.
 (762 Mt CO2-eq emitted in 2021; 438 Mt targeted for 2030)
     \item Fit-for-55 package by EU targets 55\% reduction by 2030 in the EU
-    \item REpowerEU \cite{REPowerEU} targets accelerated energy transition and security.
+    \item REpowerEU ~\cite{REPowerEU} targets accelerated energy transition and security.
 A fast ramp up of a green Hydrogen system should support transition of hard-to-abate sectors.
@@
 \end{itemize}
+
+In net-zero economies, carbon capture is considered a key technology for balancing emissions from hard-to-abate sectors and inevitable emissions from land use and agriculture.
+Capturing carbon at point sources and from the atmosphere by direct air capture (DAC) opens up new opportunities for producing drop-in fuels such as methane, methanol, or carbon-neutral liquid hydrocarbons, known as Carbon Capture and Utilization (CCU), and for storing carbon in geological formations, known as Carbon Capture and Storage (CCS).
+Together with carbon transport, these carbon management technologies are considered to be part of what is called the New Carbon Economy~\cite{arniehellerNewCarbonEconomy2019}.
+
+Individual potentials of carbon management technologies have been discussed in detail in the literature and included in previous modeling efforts~\cite{burandtDecarbonizingChinaEnergy2019,caprosEnergysystemModellingEU2019,damoreOptimalDesignEuropean2021,larsonNetZeroAmericaPotential2021,maModelingOptimizationCombined2021,mikulcicFlexibleCarbonCapture2019,williamsCarbonNeutralPathwaysUnited2021}.
+Recently, policymakers and industry have been committing to carbon management strategies, planning the first infrastructure components, and developing business models for emerging sectors of the economy~\cite{adomaitisEquinorRWEBuild2023,apnewswireGermanyDrawLegislation2023,KohlenstoffKannKlimaschutz2023,OGETESJoin2022,TESHydrogenLife2023}.
+
+However, there is still a lack of careful assessment of the benefits of carbon infrastructure in geographic terms.
+As we show in our study, the dynamics of carbon capture, transport, use, and storage in different regions are highly dependent on the availability of carbon pipelines and sequestration sinks.
+To our knowledge, there has never been a study that has examined carbon management with a detailed geographic representation for the future European energy system.
+
+In this paper, we therefore present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies.
+The study is conducted using the PyPSA-EUR-Sec energy system model and encompasses all relevant sectors.
+We enable the system to optimize the design of renewable energy sources and storage technologies, as well as the transmission of electricity, hydrogen, methane, and carbon.
+Our evaluation focuses on the transport dynamics of carbon and hydrogen through their respective networks on the European continent.
+We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available carbon sequestration potential.

### 2023-03-07 report/sections/methods.tex
adjust layouts to paper with
update paper

@@
-Energy system modelling in PyPSA-EUR-sec~\cite{brownSynergiesSectorCoupling2018}
+To model the detailed expansion of the carbon network, we use the open-source PyPSA-Eur-Sec~\cite{PyPSAEurSecSectorCoupledOpen2023} model.
+This is a fully sector-coupled model with high spatial and temporal resolution and detailed transmission infrastructure for electricity, hydrogen, methane, and carbon networks.
+The model enables joint optimization of investment and operational decisions for generation, storage, conversion, and transmission infrastructures.
+For our study, we use an approach with 181~geographic regions for Europe and a temporal resolution of 3~hours.
+These settings allow us to detail the future carbon network and capture the dynamics of material flow between industrial carbon sources and carbon sinks for both use and storage.
+
+PyPSA-Eur-Sec uses exogenous assumptions for the regional energy demand of the power sector, industry, transport, buildings and agriculture.
+This includes energy demand for shipping and aviation, and non-energy feedstocks for the chemical industry.
+To take account of climate neutrality targets, we set a binding cap on zero emissions in 2050.
+Biomass use in each region is limited by its biomass potential, but biomass transport between regions is allowed and the model accounts for transport costs.
+Furthermore, we limit the expansion of the electric transmission system to 50\% of today's total capacity to account for the difficulties in placing new transmission projects.
+The geographical layout of process emissions from industry is derived from~\cite{piamanzGeoreferencedIndustrialSites2018}.
+For carbon sequestration, only offshore sites are considered as potential sinks (see Figure~\ref{fig:carbonSequestrationPotentials}).
+This is partly because offshore storage potential tends to be greater than onshore potential, and partly because carbon storage infrastructure projects near population centers may be difficult to implement due to public safety concerns.
+We make conservative assumptions in estimating storage potential by truncating the total potential to 25~Mt per site and calculate annual storage availability when filled over 25 years.
+On the carbon utilization side, we include three drop-in fuel production technologies in our model: Methanation, Methanolization, and Fischer-Tropsch (FT) synthesis.
+For the fuels to be considered carbon-neutral substitutes, they must be produced using carbon from a technology with negative emissions.
+Synthetic methane can be used as a substitute for natural gas or biogas and is used in combined heat and power (CHP) plants, in gas boilers for home heating, or to meet the gas needs of industry.
+Synthetic methanol is used in the model to decarbonize fuel demand in the marine industry.
+Finally, FT fuels can replace fossil oil in the production of naphtha for industry and kerosene for aviation, or be used as machinery oil for agriculture.
+All technology cost assumptions for 2050 are listed taken from~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
+
+To explore the impact of carbon sequestration on the optimal layout of carbon management technologies, we vary the global carbon sequestration potential from 200~Mt to 1000~Mt per year in steps of 200~Mt.
 
@@
 
-\begin{figure*}[h]
+\begin{figure}[h]
     \centering
-    \includegraphics[width=\linewidth]{capacity_map_co2network_1000.pdf}
-    \caption{Optimal capacities per sector for a sequestration of 1000 Mt/a.}
+    \includegraphics[width=\linewidth]{capacity_map_electricity_co2network_1000.pdf}
+    \caption{Optimal capacities per of the electricity sector for a sequestration of 1000 Mt/a.}
     \label{fig:capacity_map_noco2network_1000}
-\end{figure*}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{capacity_map_carbon_co2network_1000.pdf}
+    \caption{Optimal capacities per of the carbon sector for a sequestration of 1000 Mt/a.}
+    \label{fig:capacity_map_noco2network_1000}
+\end{figure}
 
@@
 \end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{operation_bar_co2network_200.pdf}
+\end{figure}

### 2023-03-07 report/submissions/eem/paper.tex
update paper

\documentclass[conference]{IEEEtran}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{tabularx, multirow}
\usepackage{fancyhdr,lipsum}

% \IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
% \fancypagestyle{firstpage}{% Page style for first page
%   \fancyhf{}% Clear header/footer
%   \fancyfoot[l]{}% Footer
% }
% \thispagestyle{firstpage}

% \usepackage{cleveref}

% \crefname{align}{}{}
% \Crefname{align}{Equation}{Equations}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\usepackage[%
backend=biber,bibencoding=utf8, %instead of bibtex
language=auto,
style=ieee,
sorting=none, % nyt for name, year, title
maxbibnames=10, % default: 3, et al.
%backref=true,%
natbib=true % natbib compatibility mode (\citep and~\citet still work)
]{biblatex}
\bibliography{../../references.bib}
% Template from https://de.overleaf.com/latex/templates/ieee-conference-template/grfzhhncsfqn

%define approx proportional
\def\app#1#2{%
  \mathrel{%
    \setbox0=\hbox{$#1\sim$}%
    \setbox2=\hbox{%
      \rlap{\hbox{$#1\propto$}}%
      \lower1.1\ht0\box0%
    }%
    \raise0.25\ht2\box2%
  }%
}
\def\approxprop{\mathpalette\app\relax}

\graphicspath{
    {../../../results/eem-test/figures/},
}

\begin{document}

\title{Carbon Management Strategies for a Climate-Neutral European Economy}

\thanks{Fabian Hofmann is funded by the Breakthrough Energy Project "Hydrogen Integration and Carbon Management in Energy System Models".}

\author{
    \IEEEauthorblockN{Fabian Hofmann}
    \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
        \textit{Technical University of Berlin}\\
        Berlin, Germany \\
        m.hofmann@tu-berlin.de}
    \and
    \IEEEauthorblockN{Christoph Tries}
    \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
        \textit{Technical University of Berlin}\\
        Berlin, Germany \\
        christoph.tries@tu-berlin.de}
    \and
    \IEEEauthorblockN{Fabian Neumann}
    \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
        \textit{Technical University of Berlin}\\
        Berlin, Germany \\
        f.neumann@tu-berlin.de}
    \and
    \IEEEauthorblockN{Lisa Zeyen}
    \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
        \textit{Technical University of Berlin}\\
        Berlin, Germany \\
        e.zeyen@tu-berlin.de}
    \and
    \IEEEauthorblockN{Tom Brown}
    \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
        \textit{Technical University of Berlin}\\
        Berlin, Germany \\
        t.brown@tu-berlin.de}
}

\maketitle

\begin{abstract}
    Enabling a carbon-neutral European economy requires internationally coordinated investments into all sectors of the energy system. In the light of hard-to-abate emissions from process industry, energy-intensive transport and backup heat or power sources, it is important to determine the extent to which a system-wide carbon management is needed. Different technologies such as direct air capture, carbon capture at point sources and carbon transport, utilization in hydrocarbon synthesis and storage are considered as valuable contributors to formulate a comprehensive carbon strategy and take the final step to climate-neutrality.
    While the literature has already showed the synergies of these technologies, the question of how to design the infrastructure at high resolution together with a CO$_2$ transport system has not been addressed so far. Moreover, there is a growing number of individually planned carbon management projects whereas the formulation of a European carbon strategy is still missing from the political agenda.
    In this paper, we present the first study of a cost-optimal design of the European energy system with a full representation of carbon management technologies including a CO$_2$ transport network, considering all sectors with high spatial resolution. In addition, the system is allowed to optimize the layout of renewable energy sources as well as energy transmission and storage systems based on hydrogen and methane.
    Assuming net-zero emissions, we show that carbon management technologies provide critical cost-beneficial flexibilities to the European energy system. In particular, we show that building a CO$_2$ network complements the built out of a hydrogen network and is cost-effective across different rates of carbon sequestration in the offshore regions (100 -- 1000 Mt/a). Combined with a hydrogen grid, the CO$_2$ grid leads to cost savings of about 6\%, as it allows industrial and densely populated inland sites to source hydrogen from coastal regions with favorable renewable energy potential and, in return, supply CO$_2$ that can be sequestered or utilized for synthetic hydrocarbons for petrochemical feedstocks, aviation and shipping. At high sequestration rates, direct air capture facilities complement carbon capture technologies at point sources, leading however to higher use of fossil energy carriers in the power and heating sector. The optimization identifies three main carbon sinks that dominate and determine the CO$_2$ network design, namely a large sequestration cluster in the North Sea and two clusters on the coasts of Portugal and Greece.
    The results highlight the potential of a comprehensive carbon management strategy as a basis for designing future carbon networks and identifying the most promising sites for carbon capture and storage.
\end{abstract}

\begin{IEEEkeywords}
    power system analysis computing, power system management, power system planning, renewable energy source
\end{IEEEkeywords}

\section{Introduction}
\label{sec:introduction}

In net-zero economies, carbon capture is considered a key technology for balancing emissions from hard-to-abate sectors and inevitable emissions from land use and agriculture. Capturing carbon at point sources and from the atmosphere by direct air capture (DAC) opens up new opportunities for producing drop-in fuels such as methane, methanol, or carbon-neutral liquid hydrocarbons, known as Carbon Capture and Utilization (CCU), and for storing carbon in geological formations, known as Carbon Capture and Storage (CCS). Together with carbon transport, these technologies are considered to be part of what is called the New Carbon Economy~\cite{arniehellerNewCarbonEconomy2019}.

The literature has already extensively discussed the individual potentials of carbon management technologies~\cite{burandtDecarbonizingChinaEnergy2019,caprosEnergysystemModellingEU2019,damoreOptimalDesignEuropean2021,larsonNetZeroAmericaPotential2021,maModelingOptimizationCombined2021,mikulcicFlexibleCarbonCapture2019,williamsCarbonNeutralPathwaysUnited2021}.
Endrava's Capture Map estimates a potential of 1.4~Gt of carbon capture from point sources in Europe~\cite{ToolsGreenTransition}.
A recent study on carbon capture, transport and storage in~\cite{CaptureMapGetSitespecific} has shown that ...
A sequestration potential study in~\cite{weiProposedGlobalLayout2021} suggests that 92 GtCO$_2$ can be mitigated globally, of which 59 Gt can be sequestered, at a cost of 0.12\% of global GDP.
% TODO: what is that in Euro per tonne?
Finally, the SimCCS optimization software for mixed-integer programming (MIP) problems with infrastructure decisions~\cite{middletonSimCCSOpensourceTool2020} is a valuable tool for modeling and optimizing CCS infrastructure.

Recently, policymakers and industry in Europe have been committing to carbon management strategies, planning the first infrastructure components, and developing business models for emerging sectors of the economy~\cite{adomaitisEquinorRWEBuild2023,apnewswireGermanyDrawLegislation2023,KohlenstoffKannKlimaschutz2023,OGETESJoin2022,TESHydrogenLife2023}. Business models from companies like Tree Energy~\cite{TESHydrogenLife2023}, Carbfix [cite], and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offtake CO$_2$ on the other hand. ...
% TODO other businesses, sequestration business models
Despite the growing number of individually planned carbon management projects, an internationally coordinated carbon management strategy is still missing from the political agenda.

% Hydrogen
% Germany's climate protection act~\cite{KlimaschutzgesetzKlimaneutralitaetBis} targets a 65\% reduction in CO$_2$ emissions by 2030 and climate neutrality by 2045. Similarly, the EU's Fit-for-55 package aims for a 55\% reduction in emissions by 2030. To accelerate the energy transition and ensure energy security, REpowerEU~\cite{REPowerEU} proposes a fast ramp-up of a green Hydrogen system, with a domestic production target of 10 MT H$_2$ by 2030. Potential carbon management hubs such as Tree Energy, Carbfix, and Equinor are emerging as industry leaders.
% The Mission Innovation Hydrogen Valley Platform~\cite{H2ValleysMissionInnovation} is a platform that collects hydrogen flagship projects to facilitate collaboration and knowledge sharing. Finally, the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} is a project that aims to develop a cross-border hydrogen infrastructure network to enable the transport of renewable energy across Europe.

However, to formulate such a international strategy, it is important to carefully assess the benefits of carbon infrastructure in geographic terms and in terms of interaction with the energy system. As we show in our study, the dynamics of carbon capture, transport, use, and storage in different regions are highly dependent on the availability of renewable energy, carbon pipelines and sequestration sinks. To our knowledge, there has never been a study that has examined carbon management with a detailed geographic representation for the future European energy system.

In this paper, we therefore present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies. The study is conducted using the PyPSA-EUR-Sec energy system model and encompasses all relevant sectors. We enable the system to optimize the design of renewable energy sources and storage technologies, as well as the transmission of electricity, hydrogen, methane, and carbon. Our evaluation focuses on the transport dynamics of carbon and hydrogen through their respective networks on the European continent. We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available carbon sequestration potential.

\section{Methodology}
\label{sec:methodology}

To model the detailed expansion of the carbon network, we use the open-source PyPSA-Eur-Sec~\cite{PyPSAEurSecSectorCoupledOpen2023} model. This is a fully sector-coupled model with high spatial and temporal resolution and detailed transmission infrastructure for electricity, hydrogen, methane, and carbon networks. The model enables joint optimization of investment and operational decisions for generation, storage, conversion, and transmission infrastructures. For our study, we use an approach with 181~geographic regions for Europe and a temporal resolution of 3~hours. These settings allow us to detail the future carbon network and capture the dynamics of material flow between industrial carbon sources and carbon sinks for both use and storage.

PyPSA-Eur-Sec uses exogenous assumptions for the regional energy demand of the power sector, industry, transport, buildings and agriculture. This includes energy demand for shipping and aviation, and non-energy feedstocks for the chemical industry. To take account of climate neutrality targets, we set a binding cap on zero emissions in 2050. Biomass use in each region is limited by its biomass potential, but biomass transport between regions is allowed and the model accounts for transport costs. Furthermore, we limit the expansion of the electric transmission system to 50\% of today's total capacity to account for the difficulties in placing new transmission projects. The geographical layout of process emissions from industry is derived from~\cite{piamanzGeoreferencedIndustrialSites2018}. For carbon sequestration, only offshore sites are considered as potential sinks (see Figure~\ref{fig:carbonSequestrationPotentials}). This is partly because offshore storage potential tends to be greater than onshore potential, and partly because carbon storage infrastructure projects near population centers may be difficult to implement due to public safety concerns. We make conservative assumptions in estimating storage potential by truncating the total potential to 25~Mt per site and calculate annual storage availability when filled over 25 years.
On the carbon utilization side, we include three drop-in fuel production technologies in our model: Methanation, Methanolization, and Fischer-Tropsch (FT) synthesis. For the fuels to be considered carbon-neutral substitutes, they must be produced using carbon from a technology with negative emissions. Synthetic methane can be used as a substitute for natural gas or biogas and is used in combined heat and power (CHP) plants, in gas boilers for home heating, or to meet the gas needs of industry. Synthetic methanol is used in the model to decarbonize fuel demand in the marine industry. Finally, FT fuels can replace fossil oil in the production of naphtha for industry and kerosene for aviation, or be used as machinery oil for agriculture.
All technology cost assumptions for 2050 are listed taken from~\cite{lisazeyenPyPSATechnologydataTechnology2023}.

To explore the impact of carbon sequestration on the optimal layout of carbon management technologies, we vary the global carbon sequestration potential from 200~Mt to 1000~Mt per year in steps of 200~Mt.

\section{Results}
\label{sec:results}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{sequestration_map.pdf}
    \caption{Sequestration map}
    \label{fig:sequestration_map}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{capacity_map_electricity_co2network_1000.pdf}
    \caption{Optimal capacities per of the electricity sector for a sequestration of 1000 Mt/a.}
    \label{fig:capacity_map_noco2network_1000}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{capacity_map_carbon_co2network_1000.pdf}
    \caption{Optimal capacities per of the carbon sector for a sequestration of 1000 Mt/a.}
    \label{fig:capacity_map_noco2network_1000}
\end{figure}

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_1000.pdf}
    \caption{Optimal operation per sector for a sequestration of 1000 Mt/a.}
    \label{fig:operation_map_noco2network_1000}
\end{figure*}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{operation_area_carbon_co2network.pdf}
    \caption{Balance of captured CO$_2$ emissions for the optimal operation with CO$_2$ network.}
    \label{fig:operation_area_carbon_co2network}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{operation_area_carbon_noco2network.pdf}
    \caption{Balance of captured CO$_2$ emissions for the optimal operation without CO$_2$ network.}
    \label{fig:operation_area_carbon_noco2network}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{operation_bar_co2network_200.pdf}
\end{figure}

\section{Limitations}
\label{sec:limitations}
Economy of scale? Cost projections? Other uncertainties?

\section{Conclusion}
\label{sec:conclusion}

\section*{Acknowledgements}
\label{sec:acknowledgements}

\newpage

\printbibliography

\appendix

Here come the appendices

\end{document}

### 2023-03-07 report/submissions/iew/carbon-management-iew-long-abstract.tex
update paper

\documentclass[10pt,5p,reversenotenum,lefttitle]{elsarticle}

\journal{International Energy Workshop 2023}

\bibliographystyle{elsarticle-num}
\biboptions{numbers,sort&compress,super}

\usepackage{libertine}
\usepackage{libertinust1math}
\renewcommand{\ttdefault}{\sfdefault}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{eurosym}
\usepackage{url}
\usepackage{booktabs}
\usepackage{epstopdf}
\usepackage{lipsum}
\usepackage[colorlinks]{hyperref}
\usepackage[nameinlink,sort&compress,capitalise,noabbrev]{cleveref}

%% own packages
\usepackage{subfigure}
\usepackage{float}

%% correct bad hyphenation here
\hyphenation{PyPSA-Eur-Sec co-opti-mizes}

\graphicspath{
    {static/graphics/},
    {../results/},
    {../workflow/results/v8/},
    {../workflow/subworkflows/pypsa-eur-sec/results/v8/maps/}
}

\makeatletter
\long\def\MaketitleBox{%
  \resetTitleCounters
  \def\baselinestretch{1}%
   \def\baselinestretch{1}%
    \Large\@title\par\vskip18pt
    \normalsize\elsauthors\par\vskip10pt
    \footnotesize\itshape\elsaddress\par\vskip36pt
  }
\makeatother

\begin{document}

\begin{frontmatter}

	\title{Carbon Management Strategies for a Climate-Neutral European Economy}
    \author[tub]{Christoph Tries}
    \author[tub]{Fabian Hofmann}
    \author[tub]{Fabian Neumann}
    \author[tub]{Lisa Zeyen}
    \author[tub]{Tom Brown}
	\ead{t.brown@tu-berlin.de}
	\address[tub]{Department of Digital Transformation in Energy Systems, Institute of Energy Technology, Technische Universität Berlin, Fakultät III, Einsteinufer 25 (TA 8), 10587 Berlin, Germany}

\end{frontmatter}

\section*{Long Abstract -- Introduction}
\label{sec:long-abstract-introduction}
In net-zero economies, some degree of carbon capture and sequestration (CCS) is unavoidable to offset emissions from land-use and agriculture. The capturing of carbon at scale from negative emissions technologies such as biomass with carbon capture (BCC) or direct air capture (DAC) introduces new possibilities to not only sequester the carbon, but also to produce drop-in fuels like methane, methanol or climate-neutral liquid hydrocarbons from Fischer-Tropsch (FT) synthesis. These technologies are examples of carbon capture and utilization (CCU), and together with CCS they are the core of what is also called the New Carbon Economy~\cite{arniehellerNewCarbonEconomy2019}. Modelling of a net-zero economy with system-wide carbon management can provide insights into the economics as well as geographical links of these competing sources and sinks of carbon.

The elements and technologies of carbon management in a net-zero economy are well known: DAC, carbon capture (CC) at point sources as well as carbon transport, utilization and storage are considered in most comprehensive carbon management strategies. The literature has extensively discussed individual potentials of these technologies and included them in prior modelling
efforts.~\cite{burandtDecarbonizingChinaEnergy2019,caprosEnergysystemModellingEU2019,damoreOptimalDesignEuropean2021,larsonNetZeroAmericaPotential2021,maModelingOptimizationCombined2021,mikulcicFlexibleCarbonCapture2019,williamsCarbonNeutralPathwaysUnited2021} Recently, politics and industry are moving ahead with committing to carbon management strategies, planning the first infrastructure components, and developing business models around the emerging economic sectors.~\cite{adomaitisEquinorRWEBuild2023,apnewswireGermanyDrawLegislation2023,KohlenstoffKannKlimaschutz2023,OGETESJoin2022,TESHydrogenLife2023}

However, careful evaluations of the benefits of carbon infrastructure in geographical detail are still lacking. As we show in our study, the dynamics of carbon capture, transport, utilization and storage across regions depend highly on the availability of carbon pipelines and sequestration sinks. To our knowledge, there has never been a study that looked at carbon management with detailed geographical representation for the future European energy system.

In this paper, we thus present a study of the European energy system for 2050 with high geographical detail, a full representation of carbon management technologies, and with all sectors coupled in the energy system model PyPSA-Eur-Sec. We allow the system to optimize the layout of renewable energy sources and storage technologies as well as the transmission of electricity, hydrogen, methane and carbon. We evaluate in detail the transport dynamics of carbon and hydrogen through their respective transport networks across the European continent. For example, the local capture and utilization of carbon to produce synthetic fuels and replace fossil oil is competing with the carbon transport network connecting distributed carbon capture sources with centralized carbon utilization facilities and the remote offshore carbon sinks. We further investigate how an energy system with limited annual sequestration potentials prioritizes decarbonization and fuel switching across sectors, and how carbon network buildouts differ based on varying levels of available carbon sequestration potentials.

\section*{PyPSA-Eur-Sec: A sector-coupled energy system model with spatial resolution}
\label{sec:PyPSAEurSec}

To model the detailed carbon grid expansion, we deploy the open-source model PyPSA-Eur-Sec~\cite{PyPSAEurSecSectorCoupledOpen2023}. This is a fully sector-coupled model with high spatial and temporal resolution and detailed transmission infrastructure for electricity, hydrogen, methane and carbon grids. The model co-optimizes investment and operational dispatch decisions for generation, storage, conversion and transmission infrastructures. For our study, we apply an approach with 181 geographical regions for Europe and a 3-hourly temporal resolution. With these settings, we can present a detailed representation of the future carbon grid and capture the dynamics of material flow between industrial carbon sources and carbon sinks for both utilization and storage.

PyPSA-Eur-Sec uses exogeneous assumptions for the regional energy demands from the power sector, industry, transport, buildings and agriculture. This includes energy demand for both shipping and aviation, as well as non-energy feedstocks for the chemicals industry. To account for climate neutrality targets, we apply a binding cap for zero emissions in 2050. We further limit the buildout of additional transmission grids to 50\% of today's aggregated capacity to account for the difficulty in siting new transmission projects. Technology cost assumptions are taken for the year 2050~\cite{danishenergyagencyTechnologyDataGeneration2019}. We also introduce a limit on the total annual carbon sequestration, which we vary in a sensitivity analysis ranging from 200 Mt up to 1,000 Mt per year (in increments of 200 Mt). Biomass use in each region is limited by its biomass potential, but transport of biomass across regions is allowed and the model accounts for transport costs. Biomass transport is necessary because in some regions biomass demand is higher than biomass potential.

To model the carbon network, we include detailed data from industry on the type and geographical distribution of carbon sources (see Figure~\ref{fig:carbonSources}). As potential carbon sequestration sinks we consider only offshore sites (see Figure~\ref{fig:carbonSequestrationPotentials}). This is both because offshore storage potentials tend to be larger than onshore potentials, and because carbon storage infrastructure projects near population centers may be difficult to realize due to public safety concerns. To estimate storage potentials, we use conservative assumptions: We clip total potentials to 25 Gt per site and calculate annual storage availability with filling over 25 years.

On the carbon utilization side, we include in our model three technologies for drop-in fuel production: methanization, methanolization, and FT synthesis. For the fuels to be considered a climate-neutral replacement, they must be produced with carbon from a negative emissions technology. Synthetic methane can be used as drop-in for natural gas or biogas, and is used in combined heat and power (CHP) plants, in gas boilers for home heating or to meet gas demand from industry. Synthetic methanol is used by the model to decarbonize fuel demand in the shipping industry. Finally, FT fuels can replace fossil oil in the production of naphtha for industry, kerosene for aviation or be used as agriculture machinery oil.

\begin{figure}[!ht]
  \centering
  \includegraphics[width=.45\textwidth]{../hotmaps.pdf}\hfill
  \caption{Industrial carbon sources}
  \label{fig:carbonSources}
\end{figure}

\begin{figure}[!ht]
  \centering
  \includegraphics[width=.45\textwidth]{../20230125-carbon-small/sequestration_potential.pdf}\hfill
  \caption{Offshore carbon sequestration potentials}
  \label{fig:carbonSequestrationPotentials}
\end{figure}

\section*{Preliminary results indicate sequestration at three large offshore clusters across the continent}
\label{sec:results}

With the geographical detail of our model, we can analyze the buildout of the future carbon network for 181 nodes on the European continent. As our baseline, we present the network buildout with a sequestration limit of 200 Mt per year (see Figure~\ref{fig:map_w_co2net_200}). The model finds four main carbon sinks which are dominating and determining the carbon network design, namely a large sequestration cluster in the North Sea, and three locations in the Baltic Sea and off the coasts of Portugal and Greece.

With sequestration limits of 200 Mt per year, no fossil oil and almost no fossil gas are used (see Figure~\ref{fig:balancesEnergyCO2}). Instead, industry demand for liquid hydrocarbons is met with the production of FT fuels as was already shown in Victoria et al.~\cite{victoriaSpeedTechnologicalTransformations2022}. The remaining fossil gas is partially used in industry, gas boilers for home heating, and CHP plants.

\section*{The carbon utilization industry dictates material flows and switches locations to make use of "trapped carbon"}
\label{sec:carbonUtilization}
Next to carbon sequestration, three pathways for carbon utilization are considered in the model: methanation, methanolization, and the synthesis of FT fuels. Carbon for these uses is captured either from biomass, process emissions, CHP, direct use of gas in industry, SMR or directly from the air with DAC.

Methanation only has a role in our scenario with low sequestration volumes. Methanolization is constant throughout all scenarios, since decarbonization of shipping is lacking alternative fuel uses in a net-zero scenario. Naphtha for industry and kerosene for aviation is supplied by FT fuels in our scenarios with low sequestration potentials.

In our scenario without a carbon network, carbon capture activities change significantly across the continent (compare Figure~\ref{fig:map_w_co2net_200} and \ref{fig:map_wo_co2net_200}). With no carbon transport, overall capture volumes in regions away from offshore sequestration sites are much lower. Between the different capture technologies, BCC is preferred to all others. CC of process emissions, CC of gas used in industry and CC from CHP (CHPCC) are almost exclusively used in regions connected to sequestration sites. The carbon that is captured in each region serves the local demand for drop-in fuel production of either methane, methanol or FT fuels. Additional carbon that could be captured at low costs from additional BCC or from process emissions cannot be transported away from these inland regions. Thus, overall volumes of BCC are lower than in our base scenario. Instead, to meet net-zero conditions DAC in regions next to the sequestration sinks is increased (see Portugal and Greece in Figure~\ref{fig:map_wo_co2net_200}). DAC in Sweden powered by cheap electricity from hydro and wind power is no longer feasible in this scenario, not even in regions with access to sequestration sites.

\section*{Carbon sequestration limits of 1,000 Mt CO$_2$/a lead to significant cost reductions, but continued fossil fuel use}
\label{sec:sequestrationLimits}

Sequestering up to 1,000~Mt of carbon per year in offshore regions leads to total system cost savings of around 10\% compared to no sequestration allowed (see Figure~\ref{fig:systemCosts}). Building a carbon network is cost-effective across all levels of sequestration potentials and serves the system in several ways. The main driver of system cost reductions is the direct use of oil and gas combined with back-up DAC which is then sequestered, instead of FT fuels. These are much more expensive, because they require large amounts of hydrogen that is produced through electrolysis from wind and solar power.

Furthermore, large sequestration potentials lead to a buildout of the carbon network that focuses on three main carbon sequestration sinks: the North Sea, the Mediterranean Sea near Greece and on the Atlantic Coast off of Portugal (see Figure~\ref{fig:map_w_co2net_1000}). The sink in the Baltic Sea is used to a similar degree as with lower sequestration volumes, suggesting it is a cost-attractive location for sequestration, but with only limited potential.

In terms of end-use energy, with large sequestration volumes significant amounts of fossil oil and gas remain in use (see Figure~\ref{fig:balancesEnergyCO2}). The fossil oil replaces carbon-neutral FT fuels, and is used to produce naphtha for industry and kerosene for aviation. To reach the net-zero conditions, additional DAC is deployed (see explanation of carbon utilization below). The fossil gas is not used in industry as could be assumed, where the amount of gas used (with carbon capture) remains constant. But rather, additional fossil gas is used in CHPs, gas boilers and for the production of blue hydrogen, using steam methane reforming with carbon capture (SMR CC).

\section*{The New Carbon Economy under high sequestration: fossil fuels replace FT, unless cheap carbon is trapped again}
\label{sec:highSequestration}
With a carbon network and ample sequestration sinks, the system maximizes the capture of carbon from biomass, process emissions and CHP at clusters of industry and dense population, then transports it to carbon sequestration sinks (see Figure~\ref{fig:map_w_co2net_1000}). Secondly, DAC deployment is increased by about 200\% (see Figure~\ref{fig:balancesEnergyCO2}). This DAC happens in favorable regions, namely neighboring regions to sequestration sinks with favorable renewable resources. Thirdly, some SMR CC is deployed in regions next to sequestration sinks. If sequestration potentials are available, this is cheaper than electrolysis at large volumes. Some electrolysis remains highly competitive though, as it can take advantage of a limited amount of otherwise curtailed or low-cost electricity available in the system.

When also taking away the carbon network at high sequestration limits, we again trap the cheap carbon from BCC and process emissions CC (see Figure~\ref{fig:map_wo_co2net_1000}). Instead, this cheaply available carbon is used to produce FT fuels in place. Different than at 200 Mt per year of sequestration, hydrogen is procured from SMR CC located near the carbon sequestration sinks. The hydrogen is then transported to the distributed FT fuel demands, where it is combined with the locally captured carbon to produce naphtha for industry and kerosene for aviation. Here, it would also be interesting to explore the interactions with hydrogen imports as a potential competitor.

\section*{Impact and Outlook}
\label{sec:outlook}
As we show in this paper, the availability of carbon networks and carbon sequestration sinks is highly important in determining how “trapped cheap carbon” is utilized in the new carbon economy. Biomass transport and decisions of new carbon economy players about where to locate their activities can play the same role as a carbon network and the tapping of additional sequestration sinks in moving “trapped emissions” to places where drop-in fuels are produced, or where the carbon can be sequestered.

Regionally detailed modelling of the carbon management allows us to track the activities of the new carbon economy and interactions between local carbon capture from BCC or CHPCC and their transport to carbon sequestration sinks or carbon utilization sinks like synfuel production.

In 2050, either the technical potential or policy decisions may determine the amount of carbon sequestration available to the system. If technical potentials are not accessible or not deemed safe, significant increases in energy system costs must be anticipated when economies approach net-zero targets. We hope that our analysis can inject valuable and quantified inputs into the debate and the efforts to build out the carbon management infrastructure of the future.

\clearpage

\begin{figure*}[h!]
  \centering
  \subfigure[With carbon network]{\label{fig:map_w_co2net_200}\includegraphics[width=.49\textwidth]{../20230125-carbon-small/elec_s_181_lv1.0__Co2L0-25H-T-H-B-I-A-solar+p3-linemaxext15-seq200_2050/co2network.pdf}}
  \subfigure[Without carbon network]{\label{fig:map_wo_co2net_200}\includegraphics[width=.49\textwidth]{../20230125-carbon-small/elec_s_181_lv1.0__Co2L0-25H-T-H-B-I-A-solar+p3-linemaxext15-seq200-CF+sector+co2network+false_2050/co2network.pdf}}
  \caption{Captured carbon across Europe for 200 Mt per year sequestration limit}
  \label{fig:capacity_map_200}
\end{figure*}

\begin{figure*}[h!]
  \centering
  \subfigure[With carbon network]{\label{fig:map_w_co2net_1000}\includegraphics[width=.49\textwidth]{../20230125-carbon-small/elec_s_181_lv1.0__Co2L0-25H-T-H-B-I-A-solar+p3-linemaxext15-seq1000_2050/co2network.pdf}}
  \subfigure[Without carbon network]{\label{fig:map_wo_co2net_1000}\includegraphics[width=.49\textwidth]{../20230125-carbon-small/elec_s_181_lv1.0__Co2L0-25H-T-H-B-I-A-solar+p3-linemaxext15-seq1000-CF+sector+co2network+false_2050/co2network.pdf}}
  \caption{Captured carbon across Europe for 1,000 Mt per year sequestration limit}
  \label{fig:capacity_map_1000}
\end{figure*}

\begin{figure*}[h!]
  \centering
  \includegraphics[width=.45\textwidth]{../20230125-carbon-small/seq-sensitivity-co2-h2-1.pdf}\hfill
  \caption{System costs for sequestration potentials of 200 and 1,000 Mt per year, with and without CO$_2$ network}
  \label{fig:systemCosts}
\end{figure*}

\begin{figure*}[h!]
  \centering
  \includegraphics[width=.98\textwidth]{../20230125-carbon-small/balance.pdf}\hfill
  \caption{Energy and CO$_2$ balances for sequestration potentials of 200 and 1,000 Mt per year, with and without CO$_2$ network}
  \label{fig:balancesEnergyCO2}
\end{figure*}

\clearpage

\addcontentsline{toc}{section}{References}
\bibliography{bibliography-carbon-management.bib}

\end{document}

### 2023-03-09 workflow/scripts/plot-capacity-map.py
tidy up

@@ -31,135 +31,129 @@ region_alpha = 0.8
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
         "plot_capacity_map",
-        simpl="",
-        lv=1.2,
-        clusters=181,
-        opts="",
-        sector_opts="Co2L0-365H-T-H-B-I-A-solar+p3-linemaxext15-seq200",
-        planning_horizons=2050,
+        kind="carbon",
+        design="co2network",
+        sequestration=200,
     )
 
 
 n = import_network(snakemake.input.network)
 offshore_regions = gpd.read_file(snakemake.input.offshore_regions).set_index("name")
 onshore_regions = gpd.read_file(snakemake.input.onshore_regions).set_index("name")
-kinds = snakemake.config["constants"]["kinds"]
+kind = snakemake.wildcards.kind
 config = snakemake.config
 which = "capacity"
 
 
-fig, axes = plt.subplots(
-    2,
-    2,
-    figsize=(15, 13),
+fig, ax = plt.subplots(
+    figsize=(7.5, 6.5),
     subplot_kw={"projection": ccrs.EqualEarth()},
 )
 
-for kind, ax in zip(kinds, axes.flatten()):
 
-    regions = offshore_regions if kind == "carbon" else onshore_regions
+regions = offshore_regions if kind == "carbon" else onshore_regions
 
-    specs = config["plotting"]["capacity_map"][kind]
-    bus_scale = float(specs["bus_scale"])
-    branch_scale = float(specs["branch_scale"])
+specs = config["plotting"]["capacity_map"][kind]
+bus_scale = float(specs["bus_scale"])
+branch_scale = float(specs["branch_scale"])
 
-    bus_sizes = get_carrier_production(n, kind, config, which) * bus_scale
-    branch_widths = get_carrier_transport(n, kind, config, which)
-    branch_carriers = get_carrier_transport(n, kind, config, "carrier")
-    branch_colors = {
-        c: (n.carriers.color.get(vals[0], "lightgrey") if len(vals) else "lightgrey")
-        for c, vals in branch_carriers.items()
-    }
+bus_sizes = get_carrier_production(n, kind, config, which) * bus_scale
+branch_widths = get_carrier_transport(n, kind, config, which)
+branch_carriers = get_carrier_transport(n, kind, config, "carrier")
+branch_colors = {
[…]

### 2023-03-14 report/submissions/eem/paper.tex
tidy up
further work

@@
 \graphicspath{
-    {../../../results/eem-test/figures/},
+    {../../../results/eem/figures/},
 }
@@
 Different technologies such as direct air capture, carbon capture at point sources and carbon transport, utilization in hydrocarbon synthesis and storage are considered as valuable contributors to formulate a comprehensive carbon strategy and take the final step to climate-neutrality.
-    While the literature has already showed the synergies of these technologies, the question of how to design the infrastructure at high resolution together with a CO$_2$ transport system has not been addressed so far.
+    % While the literature has already showed the synergies of these technologies, the question of how to design the infrastructure at high resolution together with a CO$_2$ transport system has not been addressed so far.
 Moreover, there is a growing number of individually planned carbon management projects whereas the formulation of a European carbon strategy is still missing from the political agenda.
@@
     Assuming net-zero emissions, we show that carbon management technologies provide critical cost-beneficial flexibilities to the European energy system.
-In particular, we show that building a CO$_2$ network complements the built out of a hydrogen network and is cost-effective across different rates of carbon sequestration in the offshore regions (100 -- 1000 Mt/a).
+In particular, we show that building a CO$_2$ network complements the built out of a hydrogen network and is cost-effective across different rates of carbon sequestration in the offshore regions (200 -- 1000 Mt/a).
 Combined with a hydrogen grid, the CO$_2$ grid leads to cost savings of about 6\%, as it allows industrial and densely populated inland sites to source hydrogen from coastal regions with favorable renewable energy potential and, in return, supply CO$_2$ that can be sequestered or utilized for synthetic hydrocarbons for petrochemical feedstocks, aviation and shipping.
 At high sequestration rates, direct air capture facilities complement carbon capture technologies at point sources, leading however to higher use of fossil energy carriers in the power and heating sector.
-The optimization identifies three main carbon sinks that dominate and determine the CO$_2$ network design, namely a large sequestration cluster in the North Sea and two clusters on the coasts of Portugal and Greece.
+    % The optimization identifies three main carbon sinks that dominate and determine the CO$_2$ network design, namely a large sequestration cluster in the North Sea and two clusters on the coasts of Portugal and Greece.
     The results highlight the potential of a comprehensive carbon management strategy as a basis for designing future carbon networks and identifying the most promising sites for carbon capture and storage.
@@
 
-The literature has already extensively discussed the individual potentials of carbon management technologies~\cite{burandtDecarbonizingChinaEnergy2019,caprosEnergysystemModellingEU2019,damoreOptimalDesignEuropean2021,larsonNetZeroAmericaPotential2021,maModelingOptimizationCombined2021,mikulcicFlexibleCarbonCapture2019,williamsCarbonNeutralPathwaysUnited2021}.
-Endrava's Capture Map estimates a potential of 1.4~Gt of carbon capture from point sources in Europe~\cite{ToolsGreenTransition}.
-A recent study on carbon capture, transport and storage in~\cite{CaptureMapGetSitespecific} has shown that ...
-A sequestration potential study in~\cite{weiProposedGlobalLayout2021} suggests that 92 GtCO$_2$ can be mitigated globally, of which 59 Gt can be sequestered, at a cost of 0.12\% of global GDP.
-% TODO: what is that in Euro per tonne?
-Finally, the SimCCS optimization software for mixed-integer programming (MIP) problems with infrastructure decisions~\cite{middletonSimCCSOpensourceTool2020} is a valuable tool for modeling and optimizing CCS infrastructure.
+The literature has extensively discussed the individual potentials of carbon management technologies~\cite{burandtDecarbonizingChinaEnergy2019,caprosEnergysystemModellingEU2019,damoreOptimalDesignEuropean2021,larsonNetZeroAmericaPotential2021,maModelingOptimizationCombined2021,mikulcicFlexibleCarbonCapture2019,williamsCarbonNeutralPathwaysUnited2021}.
+% CC
+
+% CCU
+
+% CCS
+The Capture Map in~\cite{ToolsGreenTransition} estimates a potential of 1.4~Gt of carbon capture from point sources in Europe.
+A sequestration potential study in~\cite{weiProposedGlobalLayout2021} suggests that 92 GtCO$_2$ can be mitigated globally, of which 59 Gt can be sequestered.
+A recent study on carbon capture, transport and storage in~\cite{CaptureMapGetSitespecific} has shown that
+
+% combined analysis
+Finally, optimization tools have been developed in recent years to model synergies between different carbon management technologies and support infrastructure decisions~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
+In contrast to Integrated Assessment Models these consider the spatial distribution of carbon sources and sink.
+Among these, the recent work in~\cite{becattiniCarbonDioxideCapture2022} presents a comprehensive mixed-integer model to optimize the time-evolution of a CO$_2$ transport system in Switzerland with a connection to a remote sequestration site in Norway.
 
@@
 The model enables joint optimization of investment and operational decisions for generation, storage, conversion, and transmission infrastructures.
-For our study, we use an approach with 181~geographic regions for Europe and a temporal resolution of 3~hours.
-These settings allow us to detail the future carbon network and capture the dynamics of material flow between industrial carbon sources and carbon sinks for both use and storage.
+For our study, we use a model representation with 90~geographic regions for Europe and 4~hours time resolution to capture the dynamics of material flow between carbon sources and sinks while ensuring computational feasibility.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{sequestration_map.pdf}
+    \caption{Sequestration map}
+    \label{fig:sequestration_map}
+\end{figure}
 
@@
 This includes energy demand for shipping and aviation, and non-energy feedstocks for the chemical industry.
-To take account of climate neutrality targets, we set a binding cap on zero emissions in 2050.
+To take account of climate neutrality targets, we have set a binding upper limit for zero emissions in line with the EU's 2050 target.
 Biomass use in each region is limited by its biomass potential, but biomass transport between regions is allowed and the model accounts for transport costs.
-Furthermore, we limit the expansion of the electric transmission system to 50\% of today's total capacity to account for the difficulties in placing new transmission projects.
+Furthermore, we limit the expansion of the electric transmission system to 20\% of today's total capacity to account for the difficulties in placing new transmission projects.
 The geographical layout of process emissions from industry is derived from~\cite{piamanzGeoreferencedIndustrialSites2018}.
-For carbon sequestration, only offshore sites are considered as potential sinks (see Figure~\ref{fig:carbonSequestrationPotentials}).
+These include emissions from the cement, chemical, paper and printing, glass and steel industry as well as refineries and amount 158~Mt CO$_2$ per year in total.
+For carbon sequestration, only offshore sites are considered as potential sinks (see Figure~\ref{fig:sequestration_map}).
 This is partly because offshore storage potential tends to be greater than onshore potential, and partly because carbon storage infrastructure projects near population centers may be difficult to implement due to public safety concerns.
 We make conservative assumptions in estimating storage potential by truncating the total potential to 25~Mt per site and calculate annual storage availability when filled over 25 years.
-On the carbon utilization side, we include three drop-in fuel production technologies in our model: Methanation, Methanolization, and Fischer-Tropsch (FT) synthesis.
-For the fuels to be considered carbon-neutral substitutes, they must be produced using carbon from a technology with negative emissions.
+On the carbon utilization side, we include three drop-in fuel production technologies in our model: Steam Methane Reforming (SMR), Methanolization, and Fischer-Tropsch (FT) synthesis.
+After processing, the fuels are made available for free use and can transported accross regions without additional costs.
 Synthetic methane can be used as a substitute for natural gas or biogas and is used in combined heat and power (CHP) plants, in gas boilers for home heating, or to meet the gas needs of industry.
@@
 
-\begin{figure}
+Figure~\ref{fig:capacity_map_carbon_co2network_1000} shows the optimal capacities of carbon capture and transport technologies in the European sector-coupled system for a total sequestration of 1000~Mt/a.
+The colors of the offshore regions are reflecting the optimal sequestration capacity per region.
+As shown, the optimization identifies three main carbon sinks that dominate and determine the CO$_2$ network design, namely a large sequestration cluster in the North Sea and two clusters on the coasts of Portugal and Greece.
+The CO$_2$ network is primarily built around the North Sea, connecting onshore carbon capture facilities to the coast.
+At the same time, large SMR facilities with CC are placed close to sequestration areas.
+
+\begin{figure}[h]
     \centering
-    \includegraphics[width=\linewidth]{sequestration_map.pdf}
-    \caption{Sequestration map}
-    \label{fig:sequestration_map}
+    \includegraphics[width=\linewidth]{capacity_map_carbon_co2network_1000.pdf}
+    \caption{Optimal capacities per of the carbon sector for a sequestration of 1000 Mt/a. All technologies with an "*" are explicitely augmented technologies to capture carbon emissions.}
+    \label{fig:capacity_map_carbon_co2network_1000}
 \end{figure}
@@
     \caption{Optimal capacities per of the electricity sector for a sequestration of 1000 Mt/a.}
-    \label{fig:capacity_map_noco2network_1000}
+    \label{fig:capacity_map_electricity_co2network_1000}
 \end{figure}
@@
     \centering
-    \includegraphics[width=\linewidth]{capacity_map_carbon_co2network_1000.pdf}
-    \caption{Optimal capacities per of the carbon sector for a sequestration of 1000 Mt/a.}
-    \label{fig:capacity_map_noco2network_1000}
+    \includegraphics[width=\linewidth]{capacity_map_hydrogen_co2network_1000.pdf}
+    \caption{Optimal capacities per of the hydrogen sector for a sequestration of 1000 Mt/a.}
+    \label{fig:capacity_map_hydrogen_co2network_1000}
 \end{figure}
@@
     \centering
-    \includegraphics[width=\linewidth]{operation_bar_co2network_200.pdf}
+    \includegraphics[width=\linewidth]{operation_area_hydrogen_co2network.pdf}
+    \caption{Balance of H$_2$ production and consumption with CO$_2$ network.}
+    \label{fig:operation_area_carbon_co2network}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{operation_area_hydrogen_noco2network.pdf}
+    \caption{Balance of H$_2$ production and consumption without CO$_2$ network.}
+    \label{fig:operation_area_carbon_noco2network}
 \end{figure}
@@
 Economy of scale? Cost projections? Other uncertainties?
+Industry processes not endougenous.
 

### 2023-03-14 workflow/scripts/plot-capacity-map.py

@@ -14,7 +14,6 @@ from pypsa.plot import add_legend_circles, add_legend_patches, add_legend_lines
 from common import (
     import_network,
     mock_snakemake,
-    get_carrier_consumption,
     get_carrier_storage,
     get_carrier_transport,
     get_carrier_production,
@@ -22,7 +21,7 @@ from common import (
 
 import geopandas as gpd
 
-sns.set_theme(style="white", context="paper", rc={"patch.linewidth": 0.2})
+sns.set_theme(style="white", context="paper", rc={"patch.linewidth": 0.1}, font="serif")
 column = "Optimal Capacity"
 alpha = 1
 region_alpha = 0.8
@@ -46,7 +45,7 @@ which = "capacity"
 
 
 fig, ax = plt.subplots(
-    figsize=(7.5, 6.5),
+    figsize=(7, 6),
     subplot_kw={"projection": ccrs.EqualEarth()},
 )
 
@@ -107,15 +106,16 @@ regions.plot(
         "alpha": region_alpha,
     },
 )
+ax.set_extent(snakemake.config["plotting"]["extent"])
 
 legend_bus_sizes = specs["bus_sizes"]
-legend_kwargs = {"loc": "upper left", "frameon": False}
+legend_kwargs = {"framealpha": 0.7, "frameon": True, "loc": "upper left"}
 if legend_bus_sizes is not None:
     add_legend_circles(
         ax,
         [s * bus_scale for s in legend_bus_sizes],
         [f"{s // 1000} GW" for s in legend_bus_sizes],
-        legend_kw={"bbox_to_anchor": (1, 1), **legend_kwargs},
+        legend_kw={"bbox_to_anchor": (0, 1), **legend_kwargs},
     )
 
 legend_branch_sizes = specs["branch_sizes"]
@@ -124,7 +124,7 @@ if legend_branch_sizes is not None:
         ax,
         [s * branch_scale for s in legend_branch_sizes],
         [f"{s // 1000} GW" for s in legend_branch_sizes],
-        legend_kw={"bbox_to_anchor": (1, 0.9), **legend_kwargs},
+        legend_kw={"bbox_to_anchor": (0, 0.85), "framealpha": 0.7, **legend_kwargs},
     )
 
 gen_carriers = n.carriers.loc[bus_sizes.index.unique(1)]
@@ -133,29 +133,17 @@ add_legend_patches(
     gen_carriers.color,
     gen_carriers.nice_name,
     patch_kw={"alpha": alpha},
[…]

### 2023-03-14 workflow/scripts/plot-operation-map.py

@@ -17,14 +17,13 @@ from common import (
     import_network,
     mock_snakemake,
     get_carrier_consumption,
-    get_carrier_storage,
     get_carrier_transport,
     get_carrier_production,
 )
 
 import geopandas as gpd
 
-sns.set_theme(style="white", context="paper", rc={"patch.linewidth": 0.2})
+sns.set_theme(style="white", context="paper", rc={"patch.linewidth": 0.1}, font="serif")
 warnings.filterwarnings("ignore", category=UserWarning)
 alpha = 1
 region_alpha = 0.8
@@ -33,13 +32,9 @@ region_alpha = 0.8
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
         "plot_operation_map",
-        simpl="",
-        lv=1.2,
-        clusters=181,
-        opts="",
-        sector_opts="Co2L0-365H-T-H-B-I-A-solar+p3-linemaxext15-seq200",
-        planning_horizons=2050,
-        kind="electricity",
+        kind="hydrogen",
+        design="co2network",
+        sequestration=1000,
     )
 
 
@@ -53,7 +48,7 @@ kind = snakemake.wildcards.kind
 fig, axes = plt.subplots(
     1,
     2,
-    figsize=(15, 13),
+    figsize=(10, 8),
     squeeze=False,
     subplot_kw={"projection": ccrs.EqualEarth()},
 )
@@ -102,10 +97,10 @@ for tag, ax in zip(tags, axes.flatten()):
         gen_carriers.color,
         gen_carriers.nice_name,
         patch_kw={"alpha": alpha},
-        legend_kw={"bbox_to_anchor": (0, 0), "ncol": 3, **legend_kwargs},
+        legend_kw={"bbox_to_anchor": (0, 0), "ncol": 2, **legend_kwargs},
     )
 
-    ax.set_extent(regions.total_bounds[[0, 2, 1, 3]])
+    ax.set_extent(snakemake.config["plotting"]["extent"])
     ax.set_title(kind.title() + " " + tag.title())
 
     legend_bus_sizes = specs["bus_sizes"]

### 2023-03-14 workflow/scripts/plot-sequestration-map.py

@@ -10,7 +10,7 @@ import seaborn as sns
 import matplotlib.pyplot as plt
 import geopandas as gpd
 
-sns.set_theme(style="white", context="paper")
+sns.set_theme(style="white", context="paper", font="serif")
 
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():

### 2023-03-15 report/submissions/eem/paper.tex
eem: continue writing on paper

@@
 \graphicspath{
-    {../../../results/eem/figures/},
+    {../../../results/eem-test/figures/},
 }
@@
 
-PyPSA-Eur-Sec uses exogenous assumptions for the regional energy demand of the power sector, industry, transport, buildings and agriculture.
-This includes energy demand for shipping and aviation, and non-energy feedstocks for the chemical industry.
-To take account of climate neutrality targets, we have set a binding upper limit for zero emissions in line with the EU's 2050 target.
-Biomass use in each region is limited by its biomass potential, but biomass transport between regions is allowed and the model accounts for transport costs.
-Furthermore, we limit the expansion of the electric transmission system to 20\% of today's total capacity to account for the difficulties in placing new transmission projects.
-The geographical layout of process emissions from industry is derived from~\cite{piamanzGeoreferencedIndustrialSites2018}.
-These include emissions from the cement, chemical, paper and printing, glass and steel industry as well as refineries and amount 158~Mt CO$_2$ per year in total.
-For carbon sequestration, only offshore sites are considered as potential sinks (see Figure~\ref{fig:sequestration_map}).
-This is partly because offshore storage potential tends to be greater than onshore potential, and partly because carbon storage infrastructure projects near population centers may be difficult to implement due to public safety concerns.
-We make conservative assumptions in estimating storage potential by truncating the total potential to 25~Mt per site and calculate annual storage availability when filled over 25 years.
-On the carbon utilization side, we include three drop-in fuel production technologies in our model: Steam Methane Reforming (SMR), Methanolization, and Fischer-Tropsch (FT) synthesis.
-After processing, the fuels are made available for free use and can transported accross regions without additional costs.
-Synthetic methane can be used as a substitute for natural gas or biogas and is used in combined heat and power (CHP) plants, in gas boilers for home heating, or to meet the gas needs of industry.
-Synthetic methanol is used in the model to decarbonize fuel demand in the marine industry.
-Finally, FT fuels can replace fossil oil in the production of naphtha for industry and kerosene for aviation, or be used as machinery oil for agriculture.
-All technology cost assumptions for 2050 are listed taken from~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
+The PyPSA-EUR-Sec model incorporates exogenous assumptions for regional energy demand across various sectors, including power, industry, transport, buildings, and agriculture.
+This covers energy demand for shipping and aviation, as well as non-energy feedstocks for the chemical industry.
+In order to address climate neutrality targets, we establish a binding upper limit for zero emissions, consistent with the EU's 2050 target.
+The usage of biomass in each region is constrained by its biomass potential; however, biomass transport between regions is permitted, and the model takes transport costs into account.
+Additionally, we restrict the expansion of the electric transmission system to 20\% of its current capacity to account for challenges in establishing new transmission projects.
+The geographical distribution of process emissions from various industries, such as cement, chemical, paper and printing, glass, and steel, along with refineries, is obtained from~\cite{piamanzGeoreferencedIndustrialSites2018}, amounting to a total of 158~Mt CO$_2$ per year.
 
-To explore the impact of carbon sequestration on the optimal layout of carbon management technologies, we vary the global carbon sequestration potential from 200~Mt to 1000~Mt per year in steps of 200~Mt.
+For carbon sequestration, we only consider offshore sites as potential sinks (refer to Figure~\ref{fig:sequestration_map}).
+This is partly because offshore storage potential tends to be larger than onshore potential, and partly due to public safety concerns surrounding carbon storage infrastructure projects near populated areas.
+We adopt conservative assumptions when estimating storage potential, capping the total potential at 25~Mt per site and calculating annual storage availability over a 25-year period.
+
+Regarding carbon utilization, our model incorporates three drop-in fuel production technologies: Steam Methane Reforming (SMR), Methanolization, and Fischer-Tropsch (FT) synthesis.
+Once processed, the fuels are available for unrestricted use and can be transported across regions at no additional cost.
+Synthetic methane serves as a substitute for natural gas or biogas and is utilized in combined heat and power (CHP) plants, gas boilers for residential heating, or to fulfill the gas requirements of the industry.
+Synthetic methanol is employed to decarbonize fuel demand in the marine industry.
+Lastly, FT fuels can replace fossil oil for producing naphtha in the industry and kerosene for aviation or serve as machinery oil in agriculture.
+
+All technology cost assumptions for 2050 are derived from~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
+
+To investigate the influence of carbon sequestration on the optimal configuration of carbon management technologies, we incrementally vary the global carbon sequestration potential from 200~Mt to 1000~Mt per year in steps of 200~Mt.
 
@@
 
-Figure~\ref{fig:capacity_map_carbon_co2network_1000} shows the optimal capacities of carbon capture and transport technologies in the European sector-coupled system for a total sequestration of 1000~Mt/a.
-The colors of the offshore regions are reflecting the optimal sequestration capacity per region.
-As shown, the optimization identifies three main carbon sinks that dominate and determine the CO$_2$ network design, namely a large sequestration cluster in the North Sea and two clusters on the coasts of Portugal and Greece.
-The CO$_2$ network is primarily built around the North Sea, connecting onshore carbon capture facilities to the coast.
-At the same time, large SMR facilities with CC are placed close to sequestration areas.
+\begin{figure*}[h!]
+    \centering
+    \includegraphics*[width=\linewidth]{cost_area.pdf}
+    \caption[short]{Total annual system cost for the sector-coupled system with different levels of carbon sequestration potential, with (left) and without (right) CO$_2$ network.
+When increasing the annual sequestration of from 200~Mt to 1000~Mt, the system cost reduce up to 10\% in case of a network with CO$_2$ transport.
+Due to an increased flexibility from fossil carriers with subsequent sequestration, the need for FT synthesis and H$_2$ electrolysis and the corresponding renewable power supply is reduced.
+}
+    \label{fig:cost_area}
+\end{figure*}
+
+Firstly, we analyze the optimal deployment of technologies and their associated costs for different levels of carbon sequestration potential.
+Figure~\ref{fig:cost_area} shows the annual system cost per technology group for the sector-coupled model for different levels of sequestration allowed with and without CO$_2$-network.
+The costs are calculated as the sum of the annualized investment and operational costs for all technologies.
+Starting at the left, we see that most system costs are associated with the deployment of renewable energy sources, heating, hydrogen technologies and the electricity grid.
+The costs for fossil fuels and carbon capture and transport technologies are relatively low.
+In terms of capacities, xx~GW of offshore wind, xx~GW of solar PV, xx~GW of onshore wind, and xx~GW of electrolysis are installed.
+In both cases, with and without CO$_2$ transport, total costs decrease as sequestration potential increases.
+At high sequestration rates, the fossil fuels, gas and oil, remain in the system and reduce the need for synthetic FT fuel which requires significant renewable energy sources.
+At the same time, the system deploys more net negative technologies as DAC or Bioenergy with Carbon Capture (contained in "Carbon Capture Technologies") to balance out emissions from fossil carriers.
+Other Carbon Capture Technologies ramp up; in particular, SMR with Carbon Capture.
+The conversion of natural gas to hydrogen replaces the need for investments in electrolysis and a hydrogen network.
+
+The cost reduction for the scenario with CO$_2$ transport is about xx\%, which is slightly more than without CO$_2$-net.
+Across all sequestration levels, the system with CO$_2$ transport is on average xx\% cheaper.
+
+Moving on from the cost analysis, we now turn our attention to the spatial distribution of carbon capture and transport technologies, specifically focusing on the European sector-coupled system under a total sequestration of 1000~Mt/a.
+Figure~\ref{fig:capacity_map_carbon_co2network_1000} illustrates the optimal capacities of carbon capture and transport technologies in the European sector-coupled system, given a total sequestration of 1000~Mt/a.
+The colors within the offshore regions represent the optimal sequestration capacity for each region.
+As depicted, the optimization process identifies three primary carbon sinks that significantly influence the CO$_2$ network design.
+These include a substantial sequestration cluster in the North Sea, as well as two clusters located along the coasts of Portugal and Greece.
+The CO$_2$ network is predominantly developed around the North Sea, connecting onshore carbon capture facilities to the coast.
+Concurrently, large SMR facilities equipped with carbon capture are strategically well-positioned in proximity to the sequestration areas.
 
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{capacity_map_electricity_co2network_1000.pdf}
-    \caption{Optimal capacities per of the electricity sector for a sequestration of 1000 Mt/a.}
-    \label{fig:capacity_map_electricity_co2network_1000}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{capacity_map_hydrogen_co2network_1000.pdf}
-    \caption{Optimal capacities per of the hydrogen sector for a sequestration of 1000 Mt/a.}
-    \label{fig:capacity_map_hydrogen_co2network_1000}
-\end{figure}
-
-\begin{figure*}
+\begin{figure*}[h]
     \centering
@@
     \includegraphics[width=\linewidth]{operation_area_carbon_co2network.pdf}
-    \caption{Balance of captured CO$_2$ emissions for the optimal operation with CO$_2$ network.}
+    \caption{Balance of captured CO$_2$ for the optimal operation with CO$_2$ network.}
     \label{fig:operation_area_carbon_co2network}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{operation_area_carbon_noco2network.pdf}
-    \caption{Balance of captured CO$_2$ emissions for the optimal operation without CO$_2$ network.}
-    \label{fig:operation_area_carbon_noco2network}
 \end{figure}
@@
     \caption{Balance of H$_2$ production and consumption with CO$_2$ network.}
-    \label{fig:operation_area_carbon_co2network}
+    \label{fig:operation_area_hydrogen_co2network}
 \end{figure}
@@
     \centering
-    \includegraphics[width=\linewidth]{operation_area_hydrogen_noco2network.pdf}
-    \caption{Balance of H$_2$ production and consumption without CO$_2$ network.}
-    \label{fig:operation_area_carbon_noco2network}
+    \includegraphics[width=\linewidth]{operation_area_co2_co2network.pdf}
+    \caption{Balance of atmospheric CO$_2$ emissions for the optimal operation with CO$_2$ network.}
+    \label{fig:operation_area_co2_co2network}
 \end{figure}
@@
 
-Here come the appendices
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{capacity_map_electricity_co2network_1000.pdf}
+    \caption{Optimal capacities per of the electricity sector for a sequestration of 1000 Mt/a.}
+    \label{fig:capacity_map_electricity_co2network_1000}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{capacity_map_hydrogen_co2network_1000.pdf}
+    \caption{Optimal capacities per of the hydrogen sector for a sequestration of 1000 Mt/a.}
+    \label{fig:capacity_map_hydrogen_co2network_1000}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{operation_area_carbon_noco2network.pdf}
+    \caption{Balance of captured CO$_2$ emissions for the optimal operation without CO$_2$ network.}
+    \label{fig:operation_area_carbon_noco2network}
+\end{figure}
 

### 2023-03-20 report/submissions/eem/paper.tex
further work

@@
 
+% make abbreviation for co2
+\newcommand{\carbon}{CO$_2$ }
+
 \graphicspath{
-    {../../../results/eem-test/figures/},
+    {../../../results/eem/figures/},
 }
@@
 \author{
-    \IEEEauthorblockN{Fabian Hofmann}
+    \IEEEauthorblockN{Fabian Hofmann, Christoph Tries, Fabian Neumann, Lisa Zeyen, Tom Brown}
     \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
@@
         m.hofmann@tu-berlin.de}
-    \and
-    \IEEEauthorblockN{Christoph Tries}
-    \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
-        \textit{Technical University of Berlin}\\
-        Berlin, Germany \\
-        christoph.tries@tu-berlin.de}
-    \and
-    \IEEEauthorblockN{Fabian Neumann}
-    \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
-        \textit{Technical University of Berlin}\\
-        Berlin, Germany \\
-        f.neumann@tu-berlin.de}
-    \and
-    \IEEEauthorblockN{Lisa Zeyen}
-    \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
-        \textit{Technical University of Berlin}\\
-        Berlin, Germany \\
-        e.zeyen@tu-berlin.de}
-    \and
-    \IEEEauthorblockN{Tom Brown}
-    \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
-        \textit{Technical University of Berlin}\\
-        Berlin, Germany \\
-        t.brown@tu-berlin.de}
+    % \and
+    % \IEEEauthorblockN{Christoph Tries}
+    % \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
+    %     \textit{Technical University of Berlin}\\
+    %     Berlin, Germany \\
+    %     christoph.tries@tu-berlin.de}
+    % \and
+    % \IEEEauthorblockN{Fabian Neumann}
+    % \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
+    %     \textit{Technical University of Berlin}\\
+    %     Berlin, Germany \\
+    %     f.neumann@tu-berlin.de}
+    % \and
+    % \IEEEauthorblockN{Lisa Zeyen}
+    % \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
+    %     \textit{Technical University of Berlin}\\
+    %     Berlin, Germany \\
+    %     e.zeyen@tu-berlin.de}
+    % \and
+    % \IEEEauthorblockN{Tom Brown}
+    % \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
+    %     \textit{Technical University of Berlin}\\
+    %     Berlin, Germany \\
+    %     t.brown@tu-berlin.de}
 }
@@
 In the light of hard-to-abate emissions from process industry, energy-intensive transport and backup heat or power sources, it is important to determine the extent to which a system-wide carbon management is needed.
-Different technologies such as direct air capture, carbon capture at point sources and carbon transport, utilization in hydrocarbon synthesis and storage are considered as valuable contributors to formulate a comprehensive carbon strategy and take the final step to climate-neutrality.
-    % While the literature has already showed the synergies of these technologies, the question of how to design the infrastructure at high resolution together with a CO$_2$ transport system has not been addressed so far.
-Moreover, there is a growing number of individually planned carbon management projects whereas the formulation of a European carbon strategy is still missing from the political agenda.
-    In this paper, we present the first study of a cost-optimal design of the European energy system with a full representation of carbon management technologies including a CO$_2$ transport network, considering all sectors with high spatial resolution.
-In addition, the system is allowed to optimize the layout of renewable energy sources as well as energy transmission and storage systems based on hydrogen and methane.
-    Assuming net-zero emissions, we show that carbon management technologies provide critical cost-beneficial flexibilities to the European energy system.
-In particular, we show that building a CO$_2$ network complements the built out of a hydrogen network and is cost-effective across different rates of carbon sequestration in the offshore regions (200 -- 1000 Mt/a).
-Combined with a hydrogen grid, the CO$_2$ grid leads to cost savings of about 6\%, as it allows industrial and densely populated inland sites to source hydrogen from coastal regions with favorable renewable energy potential and, in return, supply CO$_2$ that can be sequestered or utilized for synthetic hydrocarbons for petrochemical feedstocks, aviation and shipping.
-At high sequestration rates, direct air capture facilities complement carbon capture technologies at point sources, leading however to higher use of fossil energy carriers in the power and heating sector.
-    % The optimization identifies three main carbon sinks that dominate and determine the CO$_2$ network design, namely a large sequestration cluster in the North Sea and two clusters on the coasts of Portugal and Greece.
-    The results highlight the potential of a comprehensive carbon management strategy as a basis for designing future carbon networks and identifying the most promising sites for carbon capture and storage.
+Different technologies such as carbon capture, transport, utilization and storage build the basis to formulate a carbon strategy to take the final step to climate-neutrality.
+    In this paper, we present the first study of a cost-optimal design of the European energy system with a full representation of carbon management technologies including a \carbon transport network, considering all sectors with high spatial resolution.
+Among other technologies, we are jointly optimizing carbon technologies, hydrogen transport and storage, and the design of renewable energy sources in the system.
+    Assuming net-zero emissions, we show that carbon management technologies provide systemic flexibilities to the European energy system, reducing system costs up to 14\%.
+In particular, we show that building a \carbon network is cost-effective across different rates of carbon sequestration in the offshore regions (200 -- 1000 Mt/a), leading to cost reduction of 13bn€/a in average.
+The network allows to capture carbon at point sources, instead of building large amounts of Direct Air Capture facilities.
+Moreover, we show that high sequestration rates, with or without \carbon network, reduce the need for hydrogen network and synthetic fuel production and allow fossil fuels to continue to be used in industry, aviation, agriculture, and the heating sector.
+In the light of potential "loopholes" for fossil fuels, the results should be interpreted in the context of moving towards a net-negative economy and should support creating carbon management strategy for that way.
 \end{abstract}
@@
 
-In net-zero economies, carbon capture is considered a key technology for balancing emissions from hard-to-abate sectors and inevitable emissions from land use and agriculture.
-Capturing carbon at point sources and from the atmosphere by direct air capture (DAC) opens up new opportunities for producing drop-in fuels such as methane, methanol, or carbon-neutral liquid hydrocarbons, known as Carbon Capture and Utilization (CCU), and for storing carbon in geological formations, known as Carbon Capture and Storage (CCS).
-Together with carbon transport, these technologies are considered to be part of what is called the New Carbon Economy~\cite{arniehellerNewCarbonEconomy2019}.
+In net-zero economies, carbon capture is regarded as a crucial technology for offsetting emissions from hard-to-abate sectors and unavoidable emissions from land use and agriculture.
+Capturing carbon at point sources and from the atmosphere via Direct Air Capture (DAC) presents new opportunities for producing drop-in fuels such as methane, methanol, or carbon-neutral liquid hydrocarbons.
+This process is known as Carbon Capture and Utilization (CCU).
+Additionally, storing carbon in geological formations is referred to as Carbon Capture and Storage (CCS).
+Along with carbon transport, these technologies are collectively considered part of what is called the New Carbon Economy~\cite{arniehellerNewCarbonEconomy2019}.
 
-The literature has extensively discussed the individual potentials of carbon management technologies~\cite{burandtDecarbonizingChinaEnergy2019,caprosEnergysystemModellingEU2019,damoreOptimalDesignEuropean2021,larsonNetZeroAmericaPotential2021,maModelingOptimizationCombined2021,mikulcicFlexibleCarbonCapture2019,williamsCarbonNeutralPathwaysUnited2021}.
-% CC
+For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.4Gt of carbon capture from point sources per year.
+In combination with large sequestration potentials as stated in~\cite{weiProposedGlobalLayout2021}, this highlights the vast potential for decarbonization.
+Recently, policymakers and industry in Europe have been committing to carbon management projects, planning the first infrastructure components, and developing business models for emerging sectors of the economy~\cite{adomaitisEquinorRWEBuild2023,apnewswireGermanyDrawLegislation2023,KohlenstoffKannKlimaschutz2023,OGETESJoin2022,TESHydrogenLife2023}.
+Business models from companies like Tree Energy~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offtake \carbon on the other hand.
+In this context, the need for an international strategy, analogous to the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022}, becomes apparent to coordinate and support deep decarbonization efforts.
+Important insights to formulate such a strategy can be gained from energy system models that provide a holistic view of the energy system and its technological interactions.
+To this end, a number of optimization tools have been developed in recent years to model synergies between various carbon management technologies~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
+In addition to Integrated Assessment Models, these tools account for the spatial distribution of carbon sources and sinks.
+% A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon transport system in Switzerland, connecting to a remote sequestration site in Norway.
+However, the models are often limited with regard to geographical scope and detail at the same time.
+While representing a single country with spatial resolution may neglect synergies of international cooperation, a coarse grained representation of multiple countries may neglect important geographical properties.
 
-% CCU
-
-% CCS
-The Capture Map in~\cite{ToolsGreenTransition} estimates a potential of 1.4~Gt of carbon capture from point sources in Europe.
-A sequestration potential study in~\cite{weiProposedGlobalLayout2021} suggests that 92 GtCO$_2$ can be mitigated globally, of which 59 Gt can be sequestered.
-A recent study on carbon capture, transport and storage in~\cite{CaptureMapGetSitespecific} has shown that
-
-% combined analysis
-Finally, optimization tools have been developed in recent years to model synergies between different carbon management technologies and support infrastructure decisions~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
-In contrast to Integrated Assessment Models these consider the spatial distribution of carbon sources and sink.
-Among these, the recent work in~\cite{becattiniCarbonDioxideCapture2022} presents a comprehensive mixed-integer model to optimize the time-evolution of a CO$_2$ transport system in Switzerland with a connection to a remote sequestration site in Norway.
-
-Recently, policymakers and industry in Europe have been committing to carbon management strategies, planning the first infrastructure components, and developing business models for emerging sectors of the economy~\cite{adomaitisEquinorRWEBuild2023,apnewswireGermanyDrawLegislation2023,KohlenstoffKannKlimaschutz2023,OGETESJoin2022,TESHydrogenLife2023}.
-Business models from companies like Tree Energy~\cite{TESHydrogenLife2023}, Carbfix [cite], and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offtake CO$_2$ on the other hand.
-...
-% TODO other businesses, sequestration business models
-Despite the growing number of individually planned carbon management projects, an internationally coordinated carbon management strategy is still missing from the political agenda.
-
-% Hydrogen
-% Germany's climate protection act~\cite{KlimaschutzgesetzKlimaneutralitaetBis} targets a 65\% reduction in CO$_2$ emissions by 2030 and climate neutrality by 2045.
-Similarly, the EU's Fit-for-55 package aims for a 55\% reduction in emissions by 2030.
-To accelerate the energy transition and ensure energy security, REpowerEU~\cite{REPowerEU} proposes a fast ramp-up of a green Hydrogen system, with a domestic production target of 10 MT H$_2$ by 2030.
-Potential carbon management hubs such as Tree Energy, Carbfix, and Equinor are emerging as industry leaders.
-% The Mission Innovation Hydrogen Valley Platform~\cite{H2ValleysMissionInnovation} is a platform that collects hydrogen flagship projects to facilitate collaboration and knowledge sharing.
-Finally, the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} is a project that aims to develop a cross-border hydrogen infrastructure network to enable the transport of renewable energy across Europe.
-
-However, to formulate such a international strategy, it is important to carefully assess the benefits of carbon infrastructure in geographic terms and in terms of interaction with the energy system.
-As we show in our study, the dynamics of carbon capture, transport, use, and storage in different regions are highly dependent on the availability of renewable energy, carbon pipelines and sequestration sinks.
-To our knowledge, there has never been a study that has examined carbon management with a detailed geographic representation for the future European energy system.
-
-In this paper, we therefore present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies.
-The study is conducted using the PyPSA-EUR-Sec energy system model and encompasses all relevant sectors.
+In this paper, we present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies.
+The study is conducted using the \textit{PyPSA-EUR-Sec} energy system model and encompasses all relevant energy sectors.
 We enable the system to optimize the design of renewable energy sources and storage technologies, as well as the transmission of electricity, hydrogen, methane, and carbon.
+We show that the dynamics of carbon capture, transport, use, and storage in different regions are highly dependent on the availability of renewable energy, carbon pipelines and sequestration sinks.
 Our evaluation focuses on the transport dynamics of carbon and hydrogen through their respective networks on the European continent.
-We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available carbon sequestration potential.
+We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available sequestration potential.
 
@@
 
-To model the detailed expansion of the carbon network, we use the open-source PyPSA-Eur-Sec~\cite{PyPSAEurSecSectorCoupledOpen2023} model.
+To model the detailed expansion of the carbon network, we use the open-source PyPSA-EUR-Sec~\cite{PyPSAEurSecSectorCoupledOpen2023} model.
 This is a fully sector-coupled model with high spatial and temporal resolution and detailed transmission infrastructure for electricity, hydrogen, methane, and carbon networks.
@@
     \centering
-    \includegraphics[width=\linewidth]{sequestration_map.pdf}
+    \includegraphics[width=\linewidth]{sequestration_map.png}
     \caption{Sequestration map}
@@
 Additionally, we restrict the expansion of the electric transmission system to 20\% of its current capacity to account for challenges in establishing new transmission projects.
-The geographical distribution of process emissions from various industries, such as cement, chemical, paper and printing, glass, and steel, along with refineries, is obtained from~\cite{piamanzGeoreferencedIndustrialSites2018}, amounting to a total of 158~Mt CO$_2$ per year.
+The geographical distribution of process emissions from various industries, such as cement, chemical, paper and printing, glass, and steel, along with refineries, is obtained from~\cite{piamanzGeoreferencedIndustrialSites2018}, amounting to a total of 158~Mt \carbon per year.
 
@@
 
-All technology cost assumptions for 2050 are derived from~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
+All technology cost assumptions for the year 2050 are derived from~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
 
@@
 
-\begin{figure*}[h!]
+\begin{figure*}[hbt]
     \centering
     \includegraphics*[width=\linewidth]{cost_area.pdf}
-    \caption[short]{Total annual system cost for the sector-coupled system with different levels of carbon sequestration potential, with (left) and without (right) CO$_2$ network.
-When increasing the annual sequestration of from 200~Mt to 1000~Mt, the system cost reduce up to 10\% in case of a network with CO$_2$ transport.
+    \caption[short]{Total annual system cost for the sector-coupled system with different levels of carbon sequestration potential, with (left) and without (right) \carbon network.
+When increasing the annual sequestration of from 200~Mt to 1000~Mt, the system cost reduce up to 10\% in case of a network with \carbon transport.
 Due to an increased flexibility from fossil carriers with subsequent sequestration, the need for FT synthesis and H$_2$ electrolysis and the corresponding renewable power supply is reduced.
@@
 
-Firstly, we analyze the optimal deployment of technologies and their associated costs for different levels of carbon sequestration potential.
-Figure~\ref{fig:cost_area} shows the annual system cost per technology group for the sector-coupled model for different levels of sequestration allowed with and without CO$_2$-network.
+First, we analyze the optimal deployment of technologies and their associated costs for different levels of carbon sequestration potential.
+Figure~\ref{fig:cost_area} displays the annual system cost per technology group for the sector-coupled model at various sequestration levels, both with and without a \carbon network.
 The costs are calculated as the sum of the annualized investment and operational costs for all technologies.
-Starting at the left, we see that most system costs are associated with the deployment of renewable energy sources, heating, hydrogen technologies and the electricity grid.
-The costs for fossil fuels and carbon capture and transport technologies are relatively low.
-In terms of capacities, xx~GW of offshore wind, xx~GW of solar PV, xx~GW of onshore wind, and xx~GW of electrolysis are installed.
-In both cases, with and without CO$_2$ transport, total costs decrease as sequestration potential increases.
-At high sequestration rates, the fossil fuels, gas and oil, remain in the system and reduce the need for synthetic FT fuel which requires significant renewable energy sources.
-At the same time, the system deploys more net negative technologies as DAC or Bioenergy with Carbon Capture (contained in "Carbon Capture Technologies") to balance out emissions from fossil carriers.
-Other Carbon Capture Technologies ramp up; in particular, SMR with Carbon Capture.
-The conversion of natural gas to hydrogen replaces the need for investments in electrolysis and a hydrogen network.
+Starting from the left, we observe that the majority of system costs are associated with the deployment of renewable energy sources, heating, hydrogen technologies, and the electricity grid.
+Costs for fossil fuels and carbon capture and transport technologies are relatively low.
+In terms of capacities, 200~GW of offshore wind, 3930~GW of solar PV, 1910~GW of onshore wind, and 1650~GW of electrolysis are installed in this setup.
+In both cases, with and without \carbon transport, total costs decrease as sequestration potential increases.
+At high sequestration rates, fossil fuels such as gas and oil are kept in the system and reduce the need for synthetic FT fuel, which requires significant renewable energy sources.
+Simultaneously, the system deploys more net negative technologies, such as DAC or Bioenergy with Carbon Capture (included in "Carbon Capture Technologies"), to offset emissions from fossil carriers.
 
-The cost reduction for the scenario with CO$_2$ transport is about xx\%, which is slightly more than without CO$_2$-net.
-Across all sequestration levels, the system with CO$_2$ transport is on average xx\% cheaper.
+The cost reduction for the scenario with \carbon transport is about 13\%, which is slightly more than without a \carbon network (11\%).
+Across all sequestration levels, the system with \carbon transport is, on average, 1.6\% less expensive, which is absolute terms is about €13 billion per year.
 
-Moving on from the cost analysis, we now turn our attention to the spatial distribution of carbon capture and transport technologies, specifically focusing on the European sector-coupled system under a total sequestration of 1000~Mt/a.
-Figure~\ref{fig:capacity_map_carbon_co2network_1000} illustrates the optimal capacities of carbon capture and transport technologies in the European sector-coupled system, given a total sequestration of 1000~Mt/a.
-The colors within the offshore regions represent the optimal sequestration capacity for each region.
-As depicted, the optimization process identifies three primary carbon sinks that significantly influence the CO$_2$ network design.
-These include a substantial sequestration cluster in the North Sea, as well as two clusters located along the coasts of Portugal and Greece.
-The CO$_2$ network is predominantly developed around the North Sea, connecting onshore carbon capture facilities to the coast.
-Concurrently, large SMR facilities equipped with carbon capture are strategically well-positioned in proximity to the sequestration areas.
+% Other Carbon Capture Technologies, particularly SMR with Carbon Capture, ramp up.
+The conversion of natural gas to hydrogen eliminates the need for investments in electrolysis and a hydrogen network.
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{capacity_map_carbon_co2network_1000.pdf}
-    \caption{Optimal capacities per of the carbon sector for a sequestration of 1000 Mt/a. All technologies with an "*" are explicitely augmented technologies to capture carbon emissions.}
-    \label{fig:capacity_map_carbon_co2network_1000}
-\end{figure}
-
-\begin{figure*}[h]
-    \centering
-    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_1000.pdf}
-    \caption{Optimal operation per sector for a sequestration of 1000 Mt/a.}
-    \label{fig:operation_map_noco2network_1000}
-\end{figure*}
+Figures~\ref{fig:operation_area_carbon_co2network} \ref{fig:operation_area_hydrogen_co2network} and \ref{fig:operation_area_co2_co2network} show the balances of capture carbon, hydrogen and atmospheric \carbon in the system as a function of the sequestration potential respectively.
+Note that technologies with a ``*`` are explicitly augmented technologies with integrated carbon capture.
+Underlying our previous findings, we see in Figure~\ref{fig:operation_area_carbon_co2network} that the deployment of FT synthesis decreases as the sequestration potential increases.
+At the same time, we perceive a strong increase in captured carbon from Gas Combined Heat and Power (CHP) plants, Allam Cycle plants, Steam Methane Reforming (SMR) and DAC.
+The remaining technologies are constant across varying level of sequestration: biomass is bound by its maximum available potential reported in~\cite{europeancommissionjointresearchcentreENSPRESOBIOMASS2019}; the industry processes with their associated emissions are fixed; the methanol production is bound by a lower limit representing the demand from the shipping sector.
 
@@
     \includegraphics[width=\linewidth]{operation_area_carbon_co2network.pdf}
-    \caption{Balance of captured CO$_2$ for the optimal operation with CO$_2$ network.}
+    \caption{Balance of captured \carbon for the optimal operation with \carbon network. Technologies with "*" are explicitly augmented technologies to capture carbon emissions.}
     \label{fig:operation_area_carbon_co2network}
@@
     \includegraphics[width=\linewidth]{operation_area_hydrogen_co2network.pdf}
-    \caption{Balance of H$_2$ production and consumption with CO$_2$ network.}
+    \caption{Balance of H$_2$ production and consumption with \carbon network.}
     \label{fig:operation_area_hydrogen_co2network}
@@
     \includegraphics[width=\linewidth]{operation_area_co2_co2network.pdf}
-    \caption{Balance of atmospheric CO$_2$ emissions for the optimal operation with CO$_2$ network.}
+    \caption{Balance of atmospheric \carbon emissions for the optimal operation with \carbon network.}
     \label{fig:operation_area_co2_co2network}
@@
 
+% TODO: kerosine emissions from fossil based kerosine are not taken into account by the system.
+In Figure~\ref{fig:operation_area_hydrogen_co2network} we see that hydrogen production decreases by about 50\% when going from 200~Mt/a sequestration to 1000~Mt/a.
+In the last segment, H$_2$ production from SMR starts to replace H$_2$ production from electrolysis.
+On the other hand, it can be observed that hydrogen demand for FT synthetic fuel production disappears, leaving only hydrogen demand for the marine sector.
+Finally, Figure~\ref{fig:operation_area_co2_co2network} shows that atmospheric \carbon emissions from gas and SMR plants increase, which is compensated by increased capture from DAC.
+We can conclude with that in combination with a \carbon transport network, a high sequestration reduces the feasibility of CCU and increases the feasibility of CCS, namely DAC and CC at point sources.
+To answer the question how the carbon capture technologies at point sources changes with the sequestration potential, we analyze the share of technologies with integrated carbon capture.
+Figure~\ref{fig:captureshare_line} shows the share of a given technology that is equipped with integrated carbon capture, i.e., carbon capture at point source, for both with and without \carbon transport network.
+For the case with \carbon network, all technologies except for SMR and Gas CHP are nearly fully equipped with CC across all sequestration potentials.
+For SMR the share rises with the sequestration from roughly 50\% to 100\%, for CHP from 0\% to 26\%.
+For the case without \carbon network the technologies behave differently.
+Besides the fact, that capturing at point source is less attractive in total, the CC share for the technologies rise slower with the sequestration.
+For process emissions we even perceive an decrease in CC share.
+This is mainly due to the fact that with increased sequestration, the system relies more on DAC and SMR with CC which can be placed flexibly at region close to the sequestration sites.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{captureshare_line.pdf}
+    \caption{Share of facilities with integrated Carbon Capture as a function of sequestration potential in a system with and without \carbon transport.}
+    \label{fig:captureshare_line}
+\end{figure}
+
+\begin{figure*}[h!]
+    \centering
+    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_1000.png}
+    \caption{Optimal operation per sector for a sequestration of 1000 Mt/a.}
+    \label{fig:operation_map_noco2network_1000}
+\end{figure*}
+
+Proceeding from the analysis of the optimal operation for different sequestration potentials, we now examine the spatial distribution of carbon capture and transport technologies for the most "optimistic" case of 1000 Mt/a sequestration potential and a \carbon transport network.
+Figure~\ref{fig:operation_area_carbon_co2network} illustrates the total production and consumption of captured carbon in the system.
+The left side depicts the combined carbon supply of Carbon Capture facilities, along with the carbon network flow, while the right side displays regional carbon usage and the flow.
+The optimization places large capacities of DAC and SMR with CC close to the sequestration areas.
+An outstanding DAC facility can be found in Scotland where large onshore wind farms are located with a total annual production of roughly 400~TWh per year.
+Besides being used for H$_2$ electrolysis and distributed through electricity grid, 13\% of it is directly used for the  local DAC facilities.
+The optimization identifies three primary carbon sinks that significantly influence the \carbon network design.
+These include a major sequestration cluster in the North Sea, as well as two clusters situated along the coasts of Portugal and Greece and roughly matching the distribution of sequestration sites displayed in Figure~\ref{fig:sequestration_map}.
+
+We observe that the carbon network primarily serves to transport captured carbon from process and biomass emissions in central Europe to the sequestration areas.
+Simultaneously, large DAC and SMR facilities with Carbon Capture (SMR*) along the coast generate significant amounts of carbon.
+The build out of the \carbon network leads to less investments in hydrogen network.
+
+...
+
 \section{Limitations}
 \label{sec:limitations}
-Economy of scale? Cost projections? Other uncertainties?
-Industry processes not endougenous.
+
+Despite the detailed model representation, there are some limitations to the validity of the results.
+The model itself is based on a linearized optimization with perfect foresight for the entire modeling year.
+In reality, long-term energy demand and renewable supply can only be roughly estimated, while short-term predictions still entail some uncertainty.
+The model's perfect foresight may lead to non-reproducible behavior, such as precisely aligning energy storage with future energy shortages at a specific point in time.
+
+The technology costs used in the model rely on cost projections that incorporate reductions based on learning rates.
+These learning rates are derived from historical data, which may not necessarily be indicative of future trends.
+The model does not account for uncertainties arising from disruptive market behavior, such as the gas price peak that occurred in 2021.
+
+Furthermore, the model assumes a fixed rate for industrial process emissions, which cannot be altered through investments.
+This simplification may not accurately represent real-world scenarios, where many industries are considering adopting low-carbon processes and technologies.
+
+Lastly, the model's spatial and temporal resolution is insufficient to capture all relevant dynamics.
+Variations in renewable energy supply and energy demand below the applied time resolution, as well as more detailed interregional energy transport constraints, are currently not considered in the optimization.
+Transitioning to an hourly representation with more representative nodes would increase the model's computational time and complexity, but would also enhance its robustness and validity.
 
@@
 
-\section*{Acknowledgements}
-\label{sec:acknowledgements}
+% \section*{Acknowledgements}
+% \label{sec:acknowledgements}
 
@@
     \centering
-    \includegraphics[width=\linewidth]{capacity_map_electricity_co2network_1000.pdf}
+    \includegraphics[width=\linewidth]{capacity_map_electricity_co2network_1000.png}
     \caption{Optimal capacities per of the electricity sector for a sequestration of 1000 Mt/a.}
@@
     \centering
-    \includegraphics[width=\linewidth]{capacity_map_hydrogen_co2network_1000.pdf}
+    \includegraphics[width=\linewidth]{capacity_map_hydrogen_co2network_1000.png}
     \caption{Optimal capacities per of the hydrogen sector for a sequestration of 1000 Mt/a.}
@@
     \includegraphics[width=\linewidth]{operation_area_carbon_noco2network.pdf}
-    \caption{Balance of captured CO$_2$ emissions for the optimal operation without CO$_2$ network.}
+    \caption{Balance of captured \carbon emissions for the optimal operation without \carbon network.}
     \label{fig:operation_area_carbon_noco2network}

### 2023-03-21 report/submissions/eem/paper.tex
partially incorp. review comment of TB

@@
 \newcommand{\carbon}{CO$_2$ }
+\newcommand{\hydrogen}{H$_2$ }
 
@@
 
-\title{Carbon Management Strategies for a Climate-Neutral European Economy}
+\title{Designing a \carbon Network for a Carbon-Neutral European Economy}
 
@@
         m.hofmann@tu-berlin.de}
-    % \and
-    % \IEEEauthorblockN{Christoph Tries}
-    % \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
-    %     \textit{Technical University of Berlin}\\
-    %     Berlin, Germany \\
-    %     christoph.tries@tu-berlin.de}
-    % \and
-    % \IEEEauthorblockN{Fabian Neumann}
-    % \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
-    %     \textit{Technical University of Berlin}\\
-    %     Berlin, Germany \\
-    %     f.neumann@tu-berlin.de}
-    % \and
-    % \IEEEauthorblockN{Lisa Zeyen}
-    % \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
-    %     \textit{Technical University of Berlin}\\
-    %     Berlin, Germany \\
-    %     e.zeyen@tu-berlin.de}
-    % \and
-    % \IEEEauthorblockN{Tom Brown}
-    % \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
-    %     \textit{Technical University of Berlin}\\
-    %     Berlin, Germany \\
-    %     t.brown@tu-berlin.de}
 }
@@
     Enabling a carbon-neutral European economy requires internationally coordinated investments into all sectors of the energy system.
-In the light of hard-to-abate emissions from process industry, energy-intensive transport and backup heat or power sources, it is important to determine the extent to which a system-wide carbon management is needed.
+In the light of hard-to-abate emissions from heavy industry including feedstocks, dense fuels needed for transport and backup heat or power sources, it is important to determine the extent to which a system-wide carbon management is needed.
 Different technologies such as carbon capture, transport, utilization and storage build the basis to formulate a carbon strategy to take the final step to climate-neutrality.
     In this paper, we present the first study of a cost-optimal design of the European energy system with a full representation of carbon management technologies including a \carbon transport network, considering all sectors with high spatial resolution.
-Among other technologies, we are jointly optimizing carbon technologies, hydrogen transport and storage, and the design of renewable energy sources in the system.
-    Assuming net-zero emissions, we show that carbon management technologies provide systemic flexibilities to the European energy system, reducing system costs up to 14\%.
-In particular, we show that building a \carbon network is cost-effective across different rates of carbon sequestration in the offshore regions (200 -- 1000 Mt/a), leading to cost reduction of 13bn€/a in average.
-The network allows to capture carbon at point sources, instead of building large amounts of Direct Air Capture facilities.
-Moreover, we show that high sequestration rates, with or without \carbon network, reduce the need for hydrogen network and synthetic fuel production and allow fossil fuels to continue to be used in industry, aviation, agriculture, and the heating sector.
-In the light of potential "loopholes" for fossil fuels, the results should be interpreted in the context of moving towards a net-negative economy and should support creating carbon management strategy for that way.
+Among other technologies, we jointly optimize carbon technologies, hydrogen transport and storage, and the design of renewable energy sources in the system.
+    Assuming net-zero emissions, we show that carbon management technologies provide systemic flexibilities to the European energy system.
+In particular, we show that building a \carbon network is cost-effective across different rates of carbon sequestration in the offshore regions (200 -- 1000 Mt/a), leading to cost reduction of 13bn€/a on average.
+The network drives the feasibility of carbon capture at point sources which is preferred over building large amounts of direct air capture facilities.
+Moreover, we show that high sequestration rates, with or without \carbon network, reduce the need for a hydrogen network and synthetic fuel production and allow fossil fuels to continue to be used in industry, aviation, agriculture, and the heating sector.
+    % However, such high rates of sequestration would require an unprecedented expansion from today's low sequestration capacities and, in case it is not feasible, may delay reaching net-zero compared to other abatement strategies.
+
+    % In the light of potential "loopholes" for fossil fuels, the results should be interpreted in the context of moving towards a net-negative economy and should support creating carbon management strategy along that path.
 \end{abstract}
@@
 Additionally, storing carbon in geological formations is referred to as Carbon Capture and Storage (CCS).
-Along with carbon transport, these technologies are collectively considered part of what is called the New Carbon Economy~\cite{arniehellerNewCarbonEconomy2019}.
+Together with carbon transport, these technologies are grouped under the term carbon management and are part of the so-called New Carbon Economy~\cite{arniehellerNewCarbonEconomy2019}.
 
-For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.4Gt of carbon capture from point sources per year.
+For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.4~Gt of carbon capture from point sources per year.
 In combination with large sequestration potentials as stated in~\cite{weiProposedGlobalLayout2021}, this highlights the vast potential for decarbonization.
 Recently, policymakers and industry in Europe have been committing to carbon management projects, planning the first infrastructure components, and developing business models for emerging sectors of the economy~\cite{adomaitisEquinorRWEBuild2023,apnewswireGermanyDrawLegislation2023,KohlenstoffKannKlimaschutz2023,OGETESJoin2022,TESHydrogenLife2023}.
-Business models from companies like Tree Energy~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offtake \carbon on the other hand.
+Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offtake \carbon on the other hand.
 In this context, the need for an international strategy, analogous to the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022}, becomes apparent to coordinate and support deep decarbonization efforts.
@@
 To this end, a number of optimization tools have been developed in recent years to model synergies between various carbon management technologies~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
-In addition to Integrated Assessment Models, these tools account for the spatial distribution of carbon sources and sinks.
+In contrast to Integrated Assessment Models, these tools account for the spatial distribution of carbon sources and sinks.
 % A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon transport system in Switzerland, connecting to a remote sequestration site in Norway.
@@
 In this paper, we present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies.
-The study is conducted using the \textit{PyPSA-EUR-Sec} energy system model and encompasses all relevant energy sectors.
-We enable the system to optimize the design of renewable energy sources and storage technologies, as well as the transmission of electricity, hydrogen, methane, and carbon.
-We show that the dynamics of carbon capture, transport, use, and storage in different regions are highly dependent on the availability of renewable energy, carbon pipelines and sequestration sinks.
-Our evaluation focuses on the transport dynamics of carbon and hydrogen through their respective networks on the European continent.
+The study is conducted using the PyPSA-Eur-Sec energy system model and encompasses all relevant energy sectors.
+We enable the system to optimize the design of renewable energy sources and storage technologies, as well as the transmission of electricity, hydrogen, methane, and \carbon.
+We show that the dynamics of carbon capture, transport, use, and storage in different regions are highly dependent on the availability of renewable energy, \carbon pipelines and sequestration sinks.
+Our evaluation focuses on the transport dynamics of \carbon and \hydrogen through their respective networks on the European continent.
 We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available sequestration potential.
@@
 
-To model the detailed expansion of the carbon network, we use the open-source PyPSA-EUR-Sec~\cite{PyPSAEurSecSectorCoupledOpen2023} model.
+To model the detailed expansion of the carbon network, we use the open-source PyPSA-Eur-Sec~\cite{PyPSAEurSecSectorCoupledOpen2023} model.
 This is a fully sector-coupled model with high spatial and temporal resolution and detailed transmission infrastructure for electricity, hydrogen, methane, and carbon networks.
 The model enables joint optimization of investment and operational decisions for generation, storage, conversion, and transmission infrastructures.
-For our study, we use a model representation with 90~geographic regions for Europe and 4~hours time resolution to capture the dynamics of material flow between carbon sources and sinks while ensuring computational feasibility.
+For our study, we use a model representation with 90~geographic regions for Europe and 4 hourly time resolution to capture the dynamics of material flow between carbon sources and sinks while ensuring computational feasibility.
 
@@
 
-The PyPSA-EUR-Sec model incorporates exogenous assumptions for regional energy demand across various sectors, including power, industry, transport, buildings, and agriculture.
+The PyPSA-Eur-Sec model incorporates exogenous assumptions for regional energy demand across various sectors, including power, industry, transport, buildings, and agriculture.
 This covers energy demand for shipping and aviation, as well as non-energy feedstocks for the chemical industry.
@@
 The geographical distribution of process emissions from various industries, such as cement, chemical, paper and printing, glass, and steel, along with refineries, is obtained from~\cite{piamanzGeoreferencedIndustrialSites2018}, amounting to a total of 158~Mt \carbon per year.
+% TODO: Mention exogenous decisions on carbon capture and 90% rates here for industry.
 
 For carbon sequestration, we only consider offshore sites as potential sinks (refer to Figure~\ref{fig:sequestration_map}).
-This is partly because offshore storage potential tends to be larger than onshore potential, and partly due to public safety concerns surrounding carbon storage infrastructure projects near populated areas.
+This is partly because offshore storage potentials tends to be larger than onshore potential, and partly due to public safety concerns surrounding carbon storage infrastructure projects near populated areas.
 We adopt conservative assumptions when estimating storage potential, capping the total potential at 25~Mt per site and calculating annual storage availability over a 25-year period.
+% TODO: what defines a site?
 
-Regarding carbon utilization, our model incorporates three drop-in fuel production technologies: Steam Methane Reforming (SMR), Methanolization, and Fischer-Tropsch (FT) synthesis.
-Once processed, the fuels are available for unrestricted use and can be transported across regions at no additional cost.
+Regarding carbon utilization, our model incorporates three drop-in fuel production technologies: methanation, methanolization, and Fischer-Tropsch (FT) synthesis.
+Once processed, the fuels are available for unrestricted use and can be transported across regions.
 Synthetic methane serves as a substitute for natural gas or biogas and is utilized in combined heat and power (CHP) plants, gas boilers for residential heating, or to fulfill the gas requirements of the industry.
@@
 
-All technology cost assumptions for the year 2050 are derived from~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
+All technology cost assumptions for the year 2050 are derived from a unified open-source technology database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
 
@@
 When increasing the annual sequestration of from 200~Mt to 1000~Mt, the system cost reduce up to 10\% in case of a network with \carbon transport.
-Due to an increased flexibility from fossil carriers with subsequent sequestration, the need for FT synthesis and H$_2$ electrolysis and the corresponding renewable power supply is reduced.
+Due to an increased flexibility from fossil carriers with subsequent sequestration, the need for FT synthesis and \hydrogen electrolysis and the corresponding renewable power supply is reduced.
+    % TODO: mention details on the grouped technologies
     }
@@
 At high sequestration rates, fossil fuels such as gas and oil are kept in the system and reduce the need for synthetic FT fuel, which requires significant renewable energy sources.
-Simultaneously, the system deploys more net negative technologies, such as DAC or Bioenergy with Carbon Capture (included in "Carbon Capture Technologies"), to offset emissions from fossil carriers.
+Simultaneously, the system deploys more \carbon removal technologies, such as DAC and Bioenergy with carbon capture (included in "Carbon Capture Technologies"), to offset emissions from fossil fuels.
 
-The cost reduction for the scenario with \carbon transport is about 13\%, which is slightly more than without a \carbon network (11\%).
-Across all sequestration levels, the system with \carbon transport is, on average, 1.6\% less expensive, which is absolute terms is about €13 billion per year.
+The cost reduction when going to high sequestration in a system with \carbon transport is about 13\%, which is slightly more than without a \carbon network (11\%).
+Across all sequestration levels, the system with \carbon transport is, on average, 1.6\% less expensive.
+The cost savings for a carbon transport system are lowest at 0.5\% (4 billion euros per year) for a sequestration of 200~Mt and highest at 3\% (21 billion euros per year) for a sequestration of 1000~Mt.
 
-% Other Carbon Capture Technologies, particularly SMR with Carbon Capture, ramp up.
-The conversion of natural gas to hydrogen eliminates the need for investments in electrolysis and a hydrogen network.
-
-Figures~\ref{fig:operation_area_carbon_co2network} \ref{fig:operation_area_hydrogen_co2network} and \ref{fig:operation_area_co2_co2network} show the balances of capture carbon, hydrogen and atmospheric \carbon in the system as a function of the sequestration potential respectively.
+To further investigate the switch in technologies, we analyze the annual operation of the system for different sequestration potentials.
+Figures~\ref{fig:operation_area_carbon_co2network}, \ref{fig:operation_area_hydrogen_co2network} and \ref{fig:operation_area_co2_co2network} show the balances of capture carbon, hydrogen and atmospheric \carbon in the system with \carbon transport as a function of the sequestration potential respectively.
+The areas in the positive segment indicated gross production of the respective commodity, while the areas in the negative segment indicate gross consumption.
 Note that technologies with a ``*`` are explicitly augmented technologies with integrated carbon capture.
-Underlying our previous findings, we see in Figure~\ref{fig:operation_area_carbon_co2network} that the deployment of FT synthesis decreases as the sequestration potential increases.
-At the same time, we perceive a strong increase in captured carbon from Gas Combined Heat and Power (CHP) plants, Allam Cycle plants, Steam Methane Reforming (SMR) and DAC.
-The remaining technologies are constant across varying level of sequestration: biomass is bound by its maximum available potential reported in~\cite{europeancommissionjointresearchcentreENSPRESOBIOMASS2019}; the industry processes with their associated emissions are fixed; the methanol production is bound by a lower limit representing the demand from the shipping sector.
-
+%
+Figure~\ref{fig:operation_area_carbon_co2network} confirms our previous results and shows that FT synthesis production decreases with sequestration availability.
+At the same time, integrated carbon capture from Gas Combined Heat and Power (CHP) plants, Allam Cycle plants, Steam Methane Reforming (SMR) and DAC increases.
+Other technologies are constant across varying level of sequestration due to different reasons: biomass is bound by its maximum available potential (as reported in~\cite{europeancommissionjointresearchcentreENSPRESOBIOMASS2019}); the industry processes with their associated emissions are set to fixed values; the methanol production is bound by a lower limit representing the demand from the shipping sector.
+%
 \begin{figure}
@@
 \end{figure}
-
+%
 \begin{figure}
@@
     \includegraphics[width=\linewidth]{operation_area_hydrogen_co2network.pdf}
-    \caption{Balance of H$_2$ production and consumption with \carbon network.}
+    \caption{Balance of \hydrogen production and consumption with \carbon network.}
+    % TODO: make colors of gas and gas* more distinct
     \label{fig:operation_area_hydrogen_co2network}
 \end{figure}
-
+%
 \begin{figure}
@@
 \end{figure}
+%
+% TODO: kerosine emissions from fossil based kerosine are not taken into account by the system.
+Figure~\ref{fig:operation_area_hydrogen_co2network} shows that hydrogen production decreases by about 50\% when going from 200~Mt/a sequestration to 1000~Mt/a.
+In the last segment, hydrogen production from SMR starts to replace hydrogen production from electrolysis.
+Correspondingly, hydrogen demand for FT synthetic fuel production disappears, leaving only hydrogen demand for the shipping, industry and transport sector.
+Finally, Figure~\ref{fig:operation_area_co2_co2network} shows that the balance of the atmospheric \carbon.
+Note that the absolute value in the atmospheric \carbon hardly changes with the sequestration potential.
+This is because the switch from FT to fossil fuels is mirrored by a switch from using captured \carbon from biomass and DAC for FT synthesis to sequestering it underground.
+The net increase of the DAC is solely due to an increase of unabated emissions from gas boilers.
 
-% TODO: kerosine emissions from fossil based kerosine are not taken into account by the system.
-In Figure~\ref{fig:operation_area_hydrogen_co2network} we see that hydrogen production decreases by about 50\% when going from 200~Mt/a sequestration to 1000~Mt/a.
-In the last segment, H$_2$ production from SMR starts to replace H$_2$ production from electrolysis.
-On the other hand, it can be observed that hydrogen demand for FT synthetic fuel production disappears, leaving only hydrogen demand for the marine sector.
-Finally, Figure~\ref{fig:operation_area_co2_co2network} shows that atmospheric \carbon emissions from gas and SMR plants increase, which is compensated by increased capture from DAC.
-We can conclude with that in combination with a \carbon transport network, a high sequestration reduces the feasibility of CCU and increases the feasibility of CCS, namely DAC and CC at point sources.
-To answer the question how the carbon capture technologies at point sources changes with the sequestration potential, we analyze the share of technologies with integrated carbon capture.
+We can conclude with that in combination with a \carbon transport network, high rates of sequestration reduce the feasibility of CCU and hydrogen infrastructure and increases the feasibility of CCS, namely DAC and CC at point sources.
+
+To answer the question how carbon capture technologies at point sources change with the sequestration potential, we analyze the shift to carbon capture integration for all relevant technologies.
 Figure~\ref{fig:captureshare_line} shows the share of a given technology that is equipped with integrated carbon capture, i.e., carbon capture at point source, for both with and without \carbon transport network.
-For the case with \carbon network, all technologies except for SMR and Gas CHP are nearly fully equipped with CC across all sequestration potentials.
+With \carbon transport, all technologies except for SMR and Gas CHP are nearly fully equipped with CC across all sequestration potentials.
 For SMR the share rises with the sequestration from roughly 50\% to 100\%, for CHP from 0\% to 26\%.
-For the case without \carbon network the technologies behave differently.
-Besides the fact, that capturing at point source is less attractive in total, the CC share for the technologies rise slower with the sequestration.
-For process emissions we even perceive an decrease in CC share.
-This is mainly due to the fact that with increased sequestration, the system relies more on DAC and SMR with CC which can be placed flexibly at region close to the sequestration sites.
+Without \carbon transport, the technologies behave differently.
+Besides the fact that capturing at point source is less attractive in total, the CC share for the technologies rise more slowly with the sequestration.
+The carbon capture share for process emissions even decreases.
+This is primarily due to the fact that as sequestration increases, the system relies more heavily on DAC and SMR with CC that can be flexibly placed near the sequestration sites.
 
-\begin{figure}
+\begin{figure}[h]
     \centering
     \includegraphics[width=\linewidth]{captureshare_line.pdf}
-    \caption{Share of facilities with integrated Carbon Capture as a function of sequestration potential in a system with and without \carbon transport.}
+    \caption{Share of facilities with integrated carbon capture as a function of sequestration potential in a system with and without \carbon transport.}
     \label{fig:captureshare_line}
 \end{figure}
+
+Building on the analysis of optimal operation for varying sequestration potentials, we explore the spatial distribution of carbon capture and transport technologies for the most "optimistic" case of 1000~Mt/a sequestration potential and a carbon transport network.
+
+Figure~\ref{fig:operation_map_noco2network_1000} shows a map of the total production and consumption of captured carbon in the system.
+The left side shows the combined carbon supply from carbon capture facilities and the carbon network flow.
+The right side illustrates regional carbon usage, again, together with the flow.
+Large capacity DACs and SMRs are placed close to shore with carbon capture in order to maximize efficiency.
+Significant facilities for DAC are built in the British Isles where carbon dioxide is used for methanol synthesis and also sequestration.
+Here substantial onshore wind farms generate a total annual production of approximately 400 TWh per year, of which 13\% is utilized by local DAC facilities.
+Other notable DAC installations are located on the Atlantic coast of Portugal, Spain, Ireland, and Denmark.
+Large facilities of SMR with CC are found in Portugal, South England and Greece.
+All these locations are in close proximity to sequestration sites, as indicated in Figure~\ref{fig:sequestration_map}.
+Note that the installation of SMR facilities close to the shore has the additional benefit of enabling easy transportation of hydrogen to ports for the shipping sector.
+
+Looking at the transport of \carbon across the system, we observe that the carbon network primarily transports captured carbon from process and biomass emissions from the inland to sequestration areas.
+Note that optimization installs a number of separate carbon transport networks instead of building one large connected grid.
+Therefore, a large network cluster transports carbon from Central Europe to the North Sea.
+In contrast, three smaller clusters in the south of Europe transport carbon to the Atlantic and the Mediterranean Sea.
+
+As shown on the right, methanol production using captured carbon is primarily located in the United Kingdom, Ireland, and South Spain.
+
+Finally, the deployment of carbon capture technologies in a system with high sequestration rates contrasts with the deployment of CCU facilities in a system with low carbon sequestration.
+On the one hand, with high sequestration, carbon capture management in the system is mainly driven by the location of the sequestration sites, leading to a spatially centralized deployment of carbon capture technologies.
+On the other hand, with low sequestration rates the use of CCU in a system is distributed more evenly across the system.
 
@@
     \includegraphics[width=\linewidth]{operation_map_carbon_co2network_1000.png}
-    \caption{Optimal operation per sector for a sequestration of 1000 Mt/a.}
+    \caption{Optimal operation per sector for a sequestration of 1000~Mt/a.
+The left side shows the combined carbon supply from carbon capture facilities and the carbon network flow, while the right side illustrates regional carbon usage together with the flow.}
+    % TODO: make colors more distinct
     \label{fig:operation_map_noco2network_1000}
 \end{figure*}
-
-Proceeding from the analysis of the optimal operation for different sequestration potentials, we now examine the spatial distribution of carbon capture and transport technologies for the most "optimistic" case of 1000 Mt/a sequestration potential and a \carbon transport network.
-Figure~\ref{fig:operation_area_carbon_co2network} illustrates the total production and consumption of captured carbon in the system.
-The left side depicts the combined carbon supply of Carbon Capture facilities, along with the carbon network flow, while the right side displays regional carbon usage and the flow.
-The optimization places large capacities of DAC and SMR with CC close to the sequestration areas.
-An outstanding DAC facility can be found in Scotland where large onshore wind farms are located with a total annual production of roughly 400~TWh per year.
-Besides being used for H$_2$ electrolysis and distributed through electricity grid, 13\% of it is directly used for the  local DAC facilities.
-The optimization identifies three primary carbon sinks that significantly influence the \carbon network design.
-These include a major sequestration cluster in the North Sea, as well as two clusters situated along the coasts of Portugal and Greece and roughly matching the distribution of sequestration sites displayed in Figure~\ref{fig:sequestration_map}.
-
-We observe that the carbon network primarily serves to transport captured carbon from process and biomass emissions in central Europe to the sequestration areas.
-Simultaneously, large DAC and SMR facilities with Carbon Capture (SMR*) along the coast generate significant amounts of carbon.
-The build out of the \carbon network leads to less investments in hydrogen network.
-
-...
 
@@
 These learning rates are derived from historical data, which may not necessarily be indicative of future trends.
-The model does not account for uncertainties arising from disruptive market behavior, such as the gas price peak that occurred in 2021.
+The model does not account for uncertainties arising from disruptive market behavior, such as the gas price peak that occurred in 2022.
 
+% TODO: which fixed rate? you mean that all emissions are captured by exogenous assumption or that we assume 90% of carbon can be captured?
 Furthermore, the model assumes a fixed rate for industrial process emissions, which cannot be altered through investments.
 This simplification may not accurately represent real-world scenarios, where many industries are considering adopting low-carbon processes and technologies.
+The transport of fuel (FT, gas, oil) between the regions is not unlimited which may overestimate the flexibility provided by these commodities.
 
@@
 
-% \section*{Acknowledgements}
-% \label{sec:acknowledgements}
+In this study, we investigated the impact of a \carbon transport system with different carbon sequestration potentials on the optimal configuration of technologies in a fully sector-coupled energy system in Europe, focusing on the deployment of carbon capture and Utilization (CCU) and Carbon Capture and Storage (CCS) technologies.
+Our results indicate that increasing carbon sequestration potential leads to a decrease in total system costs, with a more pronounced cost reduction observed when a carbon transport network is present.
+Higher sequestration rates result in a shift from CCU and hydrogen infrastructure towards CCS and unabated fossil fuel usage compensated by carbon dioxide removal elsewhere in the system, particularly based on direct air and biomass carbon capture with sequestration.
+
+Furthermore, we observed that with a \carbon transport network, the system tends to build out more carbon capture at point sources.
+With an increased sequestration potential, the share of technologies equipped with integrated carbon capture rises, particularly for Steam Methane Reforming (SMR) and Gas Combined Heat and Power (CHP) plants.
+Independent of a setup with or without \carbon transport, the spatial distribution of carbon capture technologies becomes more centralized, with installations concentrated near sequestration sites.
+
+Despite limitations of the study considered fixed model assumptions and cost uncertainties, our findings contribute valuable insights into the role of a carbon transport system and carbon sequestration potential in the transition towards a climate-neutral and net-negative energy system.
+As Europe strives to achieve its 2050 climate neutrality target, understanding the interplay between carbon sequestration, CCU, and CCS technologies will be crucial for the design of effective policies and strategies.
 
@@
 
-\appendix
+% \appendix
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{capacity_map_electricity_co2network_1000.png}
-    \caption{Optimal capacities per of the electricity sector for a sequestration of 1000 Mt/a.}
-    \label{fig:capacity_map_electricity_co2network_1000}
-\end{figure}
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=\linewidth]{capacity_map_electricity_co2network_1000.png}
+%     \caption{Optimal capacities per of the electricity sector for a sequestration of 1000 Mt/a.}
+%     \label{fig:capacity_map_electricity_co2network_1000}
+% \end{figure}
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{capacity_map_hydrogen_co2network_1000.png}
-    \caption{Optimal capacities per of the hydrogen sector for a sequestration of 1000 Mt/a.}
-    \label{fig:capacity_map_hydrogen_co2network_1000}
-\end{figure}
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=\linewidth]{capacity_map_hydrogen_co2network_1000.png}
+%     \caption{Optimal capacities per of the hydrogen sector for a sequestration of 1000 Mt/a.}
+%     \label{fig:capacity_map_hydrogen_co2network_1000}
+% \end{figure}
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{operation_area_carbon_noco2network.pdf}
-    \caption{Balance of captured \carbon emissions for the optimal operation without \carbon network.}
-    \label{fig:operation_area_carbon_noco2network}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=\linewidth]{operation_area_carbon_noco2network.pdf}
+%     \caption{Balance of captured \carbon emissions for the optimal operation without \carbon network.}
+%     \label{fig:operation_area_carbon_noco2network}
+% \end{figure}
 

### 2023-03-21 report/submissions/eem/paper.tex
continue review text

@@
 \begin{abstract}
-    Enabling a carbon-neutral European economy requires internationally coordinated investments into all sectors of the energy system.
+    Enabling a carbon-neutral European economy requires internationally coordinated investments in all sectors of the energy system.
 In the light of hard-to-abate emissions from heavy industry including feedstocks, dense fuels needed for transport and backup heat or power sources, it is important to determine the extent to which a system-wide carbon management is needed.
 Different technologies such as carbon capture, transport, utilization and storage build the basis to formulate a carbon strategy to take the final step to climate-neutrality.
+
     In this paper, we present the first study of a cost-optimal design of the European energy system with a full representation of carbon management technologies including a \carbon transport network, considering all sectors with high spatial resolution.
 Among other technologies, we jointly optimize carbon technologies, hydrogen transport and storage, and the design of renewable energy sources in the system.
-    Assuming net-zero emissions, we show that carbon management technologies provide systemic flexibilities to the European energy system.
-In particular, we show that building a \carbon network is cost-effective across different rates of carbon sequestration in the offshore regions (200 -- 1000 Mt/a), leading to cost reduction of 13bn€/a on average.
+
+    Assuming net-zero emissions, we argue that carbon management technologies provide systemic flexibility to the European energy system.
+In particular, we show that building a \carbon network is cost-effective across different rates of carbon sequestration in the offshore regions (200 -- 1000 Mt/a), leading to cost reductions of 13bn€/a on average.
 The network drives the feasibility of carbon capture at point sources which is preferred over building large amounts of direct air capture facilities.
-Moreover, we show that high sequestration rates, with or without \carbon network, reduce the need for a hydrogen network and synthetic fuel production and allow fossil fuels to continue to be used in industry, aviation, agriculture, and the heating sector.
+Moreover, we show that high sequestration rates, regardless of carbon network implementation, reduce the need for a hydrogen network and synthetic fuel production which allows fossil fuels to continue to be used in industry, aviation, agriculture, and the heating sector.
     % However, such high rates of sequestration would require an unprecedented expansion from today's low sequestration capacities and, in case it is not feasible, may delay reaching net-zero compared to other abatement strategies.
-
-    % In the light of potential "loopholes" for fossil fuels, the results should be interpreted in the context of moving towards a net-negative economy and should support creating carbon management strategy along that path.
 \end{abstract}
@@
 Recently, policymakers and industry in Europe have been committing to carbon management projects, planning the first infrastructure components, and developing business models for emerging sectors of the economy~\cite{adomaitisEquinorRWEBuild2023,apnewswireGermanyDrawLegislation2023,KohlenstoffKannKlimaschutz2023,OGETESJoin2022,TESHydrogenLife2023}.
-Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offtake \carbon on the other hand.
+Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offer \carbon offtake on the other hand.
 In this context, the need for an international strategy, analogous to the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022}, becomes apparent to coordinate and support deep decarbonization efforts.
@@
 % A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon transport system in Switzerland, connecting to a remote sequestration site in Norway.
-However, the models are often limited with regard to geographical scope and detail at the same time.
+However, the models are often limited with regard to both geographical scope and detail.
 While representing a single country with spatial resolution may neglect synergies of international cooperation, a coarse grained representation of multiple countries may neglect important geographical properties.
@@
 
-First, we analyze the optimal deployment of technologies and their associated costs for different levels of carbon sequestration potential.
+First, we analyze the optimal deployment of technologies and their associated costs at different levels of carbon sequestration potential.
 Figure~\ref{fig:cost_area} displays the annual system cost per technology group for the sector-coupled model at various sequestration levels, both with and without a \carbon network.
@@
 In terms of capacities, 200~GW of offshore wind, 3930~GW of solar PV, 1910~GW of onshore wind, and 1650~GW of electrolysis are installed in this setup.
-In both cases, with and without \carbon transport, total costs decrease as sequestration potential increases.
-At high sequestration rates, fossil fuels such as gas and oil are kept in the system and reduce the need for synthetic FT fuel, which requires significant renewable energy sources.
+In both cases, with and without \carbon transport, total costs decrease as the sequestration potential increases.
+At high sequestration rates, fossil fuels such as gas and oil are kept in the system.
+This reduces the need for synthetic FT fuel, which requires significant renewable energy sources.
 Simultaneously, the system deploys more \carbon removal technologies, such as DAC and Bioenergy with carbon capture (included in "Carbon Capture Technologies"), to offset emissions from fossil fuels.
@@
 The cost reduction when going to high sequestration in a system with \carbon transport is about 13\%, which is slightly more than without a \carbon network (11\%).
-Across all sequestration levels, the system with \carbon transport is, on average, 1.6\% less expensive.
+Across all sequestration levels, a system with \carbon transport is, on average, 1.6\% less expensive.
 The cost savings for a carbon transport system are lowest at 0.5\% (4 billion euros per year) for a sequestration of 200~Mt and highest at 3\% (21 billion euros per year) for a sequestration of 1000~Mt.
@@
 To further investigate the switch in technologies, we analyze the annual operation of the system for different sequestration potentials.
-Figures~\ref{fig:operation_area_carbon_co2network}, \ref{fig:operation_area_hydrogen_co2network} and \ref{fig:operation_area_co2_co2network} show the balances of capture carbon, hydrogen and atmospheric \carbon in the system with \carbon transport as a function of the sequestration potential respectively.
-The areas in the positive segment indicated gross production of the respective commodity, while the areas in the negative segment indicate gross consumption.
+Figures~\ref{fig:operation_area_carbon_co2network}, \ref{fig:operation_area_hydrogen_co2network} and \ref{fig:operation_area_co2_co2network} show the balances of captured carbon, hydrogen and atmospheric \carbon in the system with \carbon transport as a function of the sequestration potential respectively.
+The areas in the positive segment indicate gross production of the respective commodity, while the areas in the negative segment indicate gross consumption.
 Note that technologies with a ``*`` are explicitly augmented technologies with integrated carbon capture.
@@
 At the same time, integrated carbon capture from Gas Combined Heat and Power (CHP) plants, Allam Cycle plants, Steam Methane Reforming (SMR) and DAC increases.
-Other technologies are constant across varying level of sequestration due to different reasons: biomass is bound by its maximum available potential (as reported in~\cite{europeancommissionjointresearchcentreENSPRESOBIOMASS2019}); the industry processes with their associated emissions are set to fixed values; the methanol production is bound by a lower limit representing the demand from the shipping sector.
+Other technologies are constant across varying levels of sequestration due to different reasons: biomass is bound by its maximum available potential (as reported in~\cite{europeancommissionjointresearchcentreENSPRESOBIOMASS2019}); the industry processes with their associated emissions are set to fixed values; the methanol production is bound by a lower limit representing the demand from the shipping sector.
 %
@@
 This is because the switch from FT to fossil fuels is mirrored by a switch from using captured \carbon from biomass and DAC for FT synthesis to sequestering it underground.
-The net increase of the DAC is solely due to an increase of unabated emissions from gas boilers.
+The net increase in DAC is solely due to an increase in unabated emissions from gas boilers.
 
-We can conclude with that in combination with a \carbon transport network, high rates of sequestration reduce the feasibility of CCU and hydrogen infrastructure and increases the feasibility of CCS, namely DAC and CC at point sources.
+We can conclude that in combination with a \carbon transport network, high rates of sequestration reduce the feasibility of CCU and hydrogen infrastructure and increases the feasibility of CCS, namely DAC and CC at point sources.
 
-To answer the question how carbon capture technologies at point sources change with the sequestration potential, we analyze the shift to carbon capture integration for all relevant technologies.
+To answer the question of how carbon capture technologies at point sources change with the sequestration potential, we analyze the shift to carbon capture integration for all relevant technologies.
 Figure~\ref{fig:captureshare_line} shows the share of a given technology that is equipped with integrated carbon capture, i.e., carbon capture at point source, for both with and without \carbon transport network.
 With \carbon transport, all technologies except for SMR and Gas CHP are nearly fully equipped with CC across all sequestration potentials.
-For SMR the share rises with the sequestration from roughly 50\% to 100\%, for CHP from 0\% to 26\%.
-Without \carbon transport, the technologies behave differently.
-Besides the fact that capturing at point source is less attractive in total, the CC share for the technologies rise more slowly with the sequestration.
+As a result of sequestration, the share of SMR with carbon capture rises from 50\% to 100\%, and that of CHP from 0\% to 26\%.
+The technologies behave differently without carbon transport: besides the fact that capturing at point source is less attractive overall, their carbon capture shares rise more slowly as sequestration increases.
 The carbon capture share for process emissions even decreases.
@@
 % TODO: which fixed rate? you mean that all emissions are captured by exogenous assumption or that we assume 90% of carbon can be captured?
-Furthermore, the model assumes a fixed rate for industrial process emissions, which cannot be altered through investments.
+% Furthermore, the model assumes a fixed rate for industrial process emissions, which cannot be altered through investments.
 This simplification may not accurately represent real-world scenarios, where many industries are considering adopting low-carbon processes and technologies.
@@
 
-Despite limitations of the study considered fixed model assumptions and cost uncertainties, our findings contribute valuable insights into the role of a carbon transport system and carbon sequestration potential in the transition towards a climate-neutral and net-negative energy system.
+Despite limitations of the study considering fixed model assumptions and cost uncertainties, our findings contribute valuable insights into the role of a carbon transport system and carbon sequestration potential in the transition towards a climate-neutral and net-negative energy system.
 As Europe strives to achieve its 2050 climate neutrality target, understanding the interplay between carbon sequestration, CCU, and CCS technologies will be crucial for the design of effective policies and strategies.

### 2023-06-01 workflow/scripts/plot-captureshare-line.py
add config for presentation; make figure kwargs configurable

@@ -9,7 +9,7 @@ import seaborn as sns
 import matplotlib.pyplot as plt
 
 
-sns.set_theme(style="white", context="paper", font="serif")
+sns.set_theme(**snakemake.params["theme"])
 
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
@@ -49,7 +49,13 @@ df = pd.concat(df, axis=1)
 nice_name = n.carriers.nice_name
 colors = n.carriers.color.dropna().rename(nice_name)
 
-fig, axes = plt.subplots(2, 1, figsize=(5, 4), layout="constrained", sharex=True)
+fig, axes = plt.subplots(
+    2,
+    1,
+    figsize=snakemake.params.settings["figsize"],
+    layout="constrained",
+    sharex=True,
+)
 
 for ax, col in zip(axes, df.columns.unique(0)):
     df[col].mul(100).sort_values(by=200, ascending=False).T.plot(

### 2023-06-01 workflow/scripts/plot-operation-area.py

@@ -13,7 +13,7 @@ import seaborn as sns
 import matplotlib.pyplot as plt
 
 
-sns.set_theme(style="white", context="paper", font="serif")
+sns.set_theme(**snakemake.params["theme"])
 
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
@@ -60,7 +60,9 @@ consumption = (
 production = group_small_contributions(production, 1)
 consumption = group_small_contributions(consumption, 1)
 
-fig, ax = plt.subplots(1, 1, figsize=(3, 3.5), layout="constrained")
+fig, ax = plt.subplots(
+    1, 1, figsize=snakemake.params.settings["figsize"], layout="constrained"
+)
 
 production.plot(
     kind="area",

### 2023-06-01 workflow/scripts/plot-operation-bar.py

@@ -9,7 +9,7 @@ import pandas as pd
 import seaborn as sns
 import matplotlib.pyplot as plt
 
-sns.set_theme(style="white", context="paper")
+sns.set_theme(**snakemake.params["theme"])
 
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
@@ -38,7 +38,9 @@ for kind in kinds:
 df = pd.concat(df)
 
 
-fig, axes = plt.subplots(len(kinds), 1, figsize=(8, 10), sharex=True)
+fig, axes = plt.subplots(
+    len(kinds), 1, figsize=snakemake.params.settings["figsize"], sharex=True
+)
 colors = n.carriers.color.dropna()
 nice_name = n.carriers.nice_name
 

### 2023-06-01 workflow/scripts/plot-sequestration-map.py

@@ -10,7 +10,7 @@ import seaborn as sns
 import matplotlib.pyplot as plt
 import geopandas as gpd
 
-sns.set_theme(style="white", context="paper", font="serif")
+sns.set_theme(**snakemake.params["theme"])
 
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
@@ -30,7 +30,9 @@ offshore_regions["potential"] = (
 )  # Mt/a
 
 
-fig, ax = plt.subplots(figsize=(6, 6), subplot_kw={"projection": crs})
+fig, ax = plt.subplots(
+    figsize=snakemake.params.settings["figsize"], subplot_kw={"projection": crs}
+)
 offshore_regions.plot(
     ax=ax,
     column="potential",

### 2023-06-13 report/submissions/eem/paper/paper.tex
checkpoint: eem23 presentation plots

--- 2023-03-21 report/submissions/eem/paper.tex
@@
 \usepackage{amsmath,amssymb,amsfonts}
+\usepackage{cuted}  % Add the cuted package
+\usepackage{caption}
 \usepackage{algorithmic}
@@
 % \IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+% % The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
 % \fancypagestyle{firstpage}{% Page style for first page
@@
 % \Crefname{align}{Equation}{Equations}
+
+\IEEEoverridecommandlockouts
+\IEEEpubid{\makebox[\columnwidth]{979-8-3503-1258-4/23/\$31.00 ©2023 IEEE \hfill} \hspace{\columnsep}\makebox[\columnwidth]{ }}
 
@@
 % make abbreviation for co2
-\newcommand{\carbon}{CO$_2$ }
-\newcommand{\hydrogen}{H$_2$ }
+\newcommand{\carbon}{CO$_2$}
+\newcommand{\hydrogen}{H$_2$}
 
 \graphicspath{
-    {../../../results/eem/figures/},
+    {paper-figures},
 }
@@
 
-\title{Designing a \carbon Network for a Carbon-Neutral European Economy}
+\title{Designing a \carbon{} Network for a Carbon-Neutral European Economy}
 
-\thanks{Fabian Hofmann is funded by the Breakthrough Energy Project "Hydrogen Integration and Carbon Management in Energy System Models".}
+% \thanks{Fabian Hofmann is funded by the Breakthrough Energy Project "Hydrogen Integration and Carbon Management in Energy System Models".}
 
@@
         m.hofmann@tu-berlin.de}
+    % \IEEEauthorblockN{Anonym Author}
+
 }
@@
 \begin{abstract}
-    Enabling a carbon-neutral European economy requires internationally coordinated investments in all sectors of the energy system.
+    Achieving a carbon-neutral European economy requires internationally coordinated efforts across all energy sectors, especially those with hard-to-abate emissions like heavy industry.
+Against this backdrop, It is critical to assess the need for comprehensive carbon management involving technologies such as carbon capture, transport, use, and storage to develop an effective carbon strategy for climate neutrality.
+
+    This paper provides the first cost-optimal design for a European energy system that fully incorporates carbon management technologies including a carbon transport network, and considers all emissions-intensive sectors.
+It optimizes carbon technologies, hydrogen transport and storage, and the design of renewable energy sources.
+
+    We show that carbon management technologies aiming at net-zero emissions provide systemic flexibility to the European energy system.
+A carbon network is cost-effective at different carbon capture rates (200 - 1000 Mt/a), leading to average cost savings of €13 billion per year.
+The \carbon{} network enables viable point source carbon capture, reducing the need for direct air capture facilities.
+We show that high sequestration rates are cost-beneficial by allowing continued fossil fuel use in some sectors, but require massive capacity expansion which could jeopardize timely attainment of climate neutrality.
+
+\end{abstract}
+
+% Enabling a carbon-neutral European economy requires internationally coordinated investments in all sectors of the energy system.
 In the light of hard-to-abate emissions from heavy industry including feedstocks, dense fuels needed for transport and backup heat or power sources, it is important to determine the extent to which a system-wide carbon management is needed.
@@
 
-    In this paper, we present the first study of a cost-optimal design of the European energy system with a full representation of carbon management technologies including a \carbon transport network, considering all sectors with high spatial resolution.
+% In this paper, we present the first study of a cost-optimal design of the European energy system with a full representation of carbon management technologies including a \carbon{} transport network, considering all sectors with high spatial resolution.
 Among other technologies, we jointly optimize carbon technologies, hydrogen transport and storage, and the design of renewable energy sources in the system.
 
-    Assuming net-zero emissions, we argue that carbon management technologies provide systemic flexibility to the European energy system.
-In particular, we show that building a \carbon network is cost-effective across different rates of carbon sequestration in the offshore regions (200 -- 1000 Mt/a), leading to cost reductions of 13bn€/a on average.
+% Assuming net-zero emissions, we argue that carbon management technologies provide systemic flexibility to the European energy system.
+In particular, we show that building a \carbon{} network is cost-effective across different rates of carbon sequestration in the offshore regions (200 -- 1000 Mt/a), leading to cost reductions of 13bn€/a on average.
 The network drives the feasibility of carbon capture at point sources which is preferred over building large amounts of direct air capture facilities.
@@
 % However, such high rates of sequestration would require an unprecedented expansion from today's low sequestration capacities and, in case it is not feasible, may delay reaching net-zero compared to other abatement strategies.
-\end{abstract}
 
 \begin{IEEEkeywords}
-    power system analysis computing, power system management, power system planning, renewable energy source
+    carbon management, energy system modelling, CCU, CCS, sector-coupling
 \end{IEEEkeywords}
@@
 Recently, policymakers and industry in Europe have been committing to carbon management projects, planning the first infrastructure components, and developing business models for emerging sectors of the economy~\cite{adomaitisEquinorRWEBuild2023,apnewswireGermanyDrawLegislation2023,KohlenstoffKannKlimaschutz2023,OGETESJoin2022,TESHydrogenLife2023}.
-Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offer \carbon offtake on the other hand.
+Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offer \carbon{} offtake on the other hand.
 In this context, the need for an international strategy, analogous to the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022}, becomes apparent to coordinate and support deep decarbonization efforts.
@@
 In contrast to Integrated Assessment Models, these tools account for the spatial distribution of carbon sources and sinks.
-% A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon transport system in Switzerland, connecting to a remote sequestration site in Norway.
+% A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
 However, the models are often limited with regard to both geographical scope and detail.
@@
 The study is conducted using the PyPSA-Eur-Sec energy system model and encompasses all relevant energy sectors.
-We enable the system to optimize the design of renewable energy sources and storage technologies, as well as the transmission of electricity, hydrogen, methane, and \carbon.
-We show that the dynamics of carbon capture, transport, use, and storage in different regions are highly dependent on the availability of renewable energy, \carbon pipelines and sequestration sinks.
-Our evaluation focuses on the transport dynamics of \carbon and \hydrogen through their respective networks on the European continent.
+We enable the system to optimize the design of renewable energy sources and storage technologies, as well as the transmission of electricity, hydrogen, methane, and \carbon{}.
+We show that the dynamics of carbon capture, transport, use, and storage in different regions are highly dependent on the availability of renewable energy, \carbon{} pipelines and sequestration sinks.
+Our evaluation focuses on the transport dynamics of \carbon{} and \hydrogen{} through their respective networks on the European continent.
 We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available sequestration potential.
@@
 
-\begin{figure}
+\begin{figure}[h]
     \centering
@@
 Additionally, we restrict the expansion of the electric transmission system to 20\% of its current capacity to account for challenges in establishing new transmission projects.
-The geographical distribution of process emissions from various industries, such as cement, chemical, paper and printing, glass, and steel, along with refineries, is obtained from~\cite{piamanzGeoreferencedIndustrialSites2018}, amounting to a total of 158~Mt \carbon per year.
+The geographical distribution of process emissions from various industries, such as cement, chemical, paper and printing, glass, and steel, along with refineries, is obtained from~\cite{piamanzGeoreferencedIndustrialSites2018}, amounting to a total of 158~Mt \carbon{} per year.
 % TODO: Mention exogenous decisions on carbon capture and 90% rates here for industry.
@@
 
-\begin{figure*}[hbt]
+\begin{figure*}[ht!]
     \centering
     \includegraphics*[width=\linewidth]{cost_area.pdf}
-    \caption[short]{Total annual system cost for the sector-coupled system with different levels of carbon sequestration potential, with (left) and without (right) \carbon network.
-When increasing the annual sequestration of from 200~Mt to 1000~Mt, the system cost reduce up to 10\% in case of a network with \carbon transport.
-Due to an increased flexibility from fossil carriers with subsequent sequestration, the need for FT synthesis and \hydrogen electrolysis and the corresponding renewable power supply is reduced.
+    \caption[short]{Total annual system cost for the sector-coupled system with different levels of carbon sequestration potential, with (left) and without (right) \carbon{} network.
+"Gas Infrastructure" combines gas facilities for power and heat production, "H$_2$ Infrastructure" combines H$_2$ production, transport and re-electrification.
+Regardless of the implementation of a \carbon{} network, the system cost decrease as sequestration increases.
+Due to an increased flexibility from fossil carriers with subsequent sequestration, the need for FT synthesis and \hydrogen{} electrolysis and the corresponding renewable power supply is reduced.
         % TODO: mention details on the grouped technologies
@@
 First, we analyze the optimal deployment of technologies and their associated costs at different levels of carbon sequestration potential.
-Figure~\ref{fig:cost_area} displays the annual system cost per technology group for the sector-coupled model at various sequestration levels, both with and without a \carbon network.
+Figure~\ref{fig:cost_area} displays the annual system cost per technology group for the sector-coupled model at various sequestration levels, both with and without a \carbon{} network.
 The costs are calculated as the sum of the annualized investment and operational costs for all technologies.
@@
 In terms of capacities, 200~GW of offshore wind, 3930~GW of solar PV, 1910~GW of onshore wind, and 1650~GW of electrolysis are installed in this setup.
-In both cases, with and without \carbon transport, total costs decrease as the sequestration potential increases.
+In both cases, with and without \carbon{} transport, total costs decrease as the sequestration potential increases.
 At high sequestration rates, fossil fuels such as gas and oil are kept in the system.
 This reduces the need for synthetic FT fuel, which requires significant renewable energy sources.
-Simultaneously, the system deploys more \carbon removal technologies, such as DAC and Bioenergy with carbon capture (included in "Carbon Capture Technologies"), to offset emissions from fossil fuels.
+Simultaneously, the system deploys more \carbon{} removal technologies, such as DAC and Bioenergy with carbon capture (included in "Carbon Capture Technologies"), to offset emissions from fossil fuels.
 
-The cost reduction when going to high sequestration in a system with \carbon transport is about 13\%, which is slightly more than without a \carbon network (11\%).
-Across all sequestration levels, a system with \carbon transport is, on average, 1.6\% less expensive.
+The cost reduction when going to high sequestration in a system with \carbon{} transport is about 13\%, which is slightly more than without a \carbon{} network (11\%).
+Across all sequestration levels, a system with \carbon{} transport is, on average, 1.6\% less expensive.
 The cost savings for a carbon transport system are lowest at 0.5\% (4 billion euros per year) for a sequestration of 200~Mt and highest at 3\% (21 billion euros per year) for a sequestration of 1000~Mt.
@@
 To further investigate the switch in technologies, we analyze the annual operation of the system for different sequestration potentials.
-Figures~\ref{fig:operation_area_carbon_co2network}, \ref{fig:operation_area_hydrogen_co2network} and \ref{fig:operation_area_co2_co2network} show the balances of captured carbon, hydrogen and atmospheric \carbon in the system with \carbon transport as a function of the sequestration potential respectively.
+Figures~\ref{fig:operation_area_carbon_co2network}, \ref{fig:operation_area_hydrogen_co2network} and \ref{fig:operation_area_co2_co2network} show the balances of captured carbon, hydrogen and atmospheric \carbon{} in the system with \carbon{} transport as a function of the sequestration potential respectively.
 The areas in the positive segment indicate gross production of the respective commodity, while the areas in the negative segment indicate gross consumption.
@@
 %
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{operation_area_carbon_co2network.pdf}
+    \caption{Balance of captured \carbon{} for the optimal operation with \carbon{} network. Technologies with "*" are explicitly augmented technologies to capture carbon emissions.}
+    \label{fig:operation_area_carbon_co2network}
+\end{figure}
+
 Figure~\ref{fig:operation_area_carbon_co2network} confirms our previous results and shows that FT synthesis production decreases with sequestration availability.
@@
 Other technologies are constant across varying levels of sequestration due to different reasons: biomass is bound by its maximum available potential (as reported in~\cite{europeancommissionjointresearchcentreENSPRESOBIOMASS2019}); the industry processes with their associated emissions are set to fixed values; the methanol production is bound by a lower limit representing the demand from the shipping sector.
-%
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{operation_area_carbon_co2network.pdf}
-    \caption{Balance of captured \carbon for the optimal operation with \carbon network. Technologies with "*" are explicitly augmented technologies to capture carbon emissions.}
-    \label{fig:operation_area_carbon_co2network}
-\end{figure}
-%
-\begin{figure}
+
+\begin{figure}[h]
     \centering
     \includegraphics[width=\linewidth]{operation_area_hydrogen_co2network.pdf}
-    \caption{Balance of \hydrogen production and consumption with \carbon network.}
+    \caption{Balance of \hydrogen{} production and consumption with \carbon{} network.}
     % TODO: make colors of gas and gas* more distinct
@@
 %
-\begin{figure}
+\begin{figure}[h]
     \centering
     \includegraphics[width=\linewidth]{operation_area_co2_co2network.pdf}
-    \caption{Balance of atmospheric \carbon emissions for the optimal operation with \carbon network.}
+    \caption{Balance of atmospheric \carbon{} emissions for the optimal operation with \carbon{} network.}
     \label{fig:operation_area_co2_co2network}
@@
 Correspondingly, hydrogen demand for FT synthetic fuel production disappears, leaving only hydrogen demand for the shipping, industry and transport sector.
-Finally, Figure~\ref{fig:operation_area_co2_co2network} shows that the balance of the atmospheric \carbon.
-Note that the absolute value in the atmospheric \carbon hardly changes with the sequestration potential.
-This is because the switch from FT to fossil fuels is mirrored by a switch from using captured \carbon from biomass and DAC for FT synthesis to sequestering it underground.
+Finally, Figure~\ref{fig:operation_area_co2_co2network} shows that the balance of the atmospheric \carbon{}.
+Note that the absolute value in the atmospheric \carbon{} hardly changes with the sequestration potential.
+This is because the switch from FT to fossil fuels is mirrored by a switch from using captured \carbon{} from biomass and DAC for FT synthesis to sequestering it underground.
 The net increase in DAC is solely due to an increase in unabated emissions from gas boilers.
 
-We can conclude that in combination with a \carbon transport network, high rates of sequestration reduce the feasibility of CCU and hydrogen infrastructure and increases the feasibility of CCS, namely DAC and CC at point sources.
+We can conclude that in combination with a \carbon{} transport network, high rates of sequestration reduce the feasibility of CCU and hydrogen infrastructure and increases the feasibility of CCS, namely DAC and CC at point sources.
 
 To answer the question of how carbon capture technologies at point sources change with the sequestration potential, we analyze the shift to carbon capture integration for all relevant technologies.
-Figure~\ref{fig:captureshare_line} shows the share of a given technology that is equipped with integrated carbon capture, i.e., carbon capture at point source, for both with and without \carbon transport network.
-With \carbon transport, all technologies except for SMR and Gas CHP are nearly fully equipped with CC across all sequestration potentials.
+Figure~\ref{fig:captureshare_line} shows the share of a given technology that is equipped with integrated carbon capture, i.e., carbon capture at point source, for both with and without \carbon{} transport network.
+With \carbon{} transport, all technologies except for SMR and Gas CHP are nearly fully equipped with CC across all sequestration potentials.
 As a result of sequestration, the share of SMR with carbon capture rises from 50\% to 100\%, and that of CHP from 0\% to 26\%.
@@
     \includegraphics[width=\linewidth]{captureshare_line.pdf}
-    \caption{Share of facilities with integrated carbon capture as a function of sequestration potential in a system with and without \carbon transport.}
+    \caption{Share of facilities with integrated carbon capture as a function of sequestration potential in a system with and without \carbon{} transport.}
     \label{fig:captureshare_line}
@@
 
-Looking at the transport of \carbon across the system, we observe that the carbon network primarily transports captured carbon from process and biomass emissions from the inland to sequestration areas.
+Looking at the transport of \carbon{} across the system, we observe that the carbon network primarily transports captured carbon from process and biomass emissions from the inland to sequestration areas.
 Note that optimization installs a number of separate carbon transport networks instead of building one large connected grid.
@@
 
-\begin{figure*}[h!]
+\begin{figure*}[ht!]
     \centering
@@
 Variations in renewable energy supply and energy demand below the applied time resolution, as well as more detailed interregional energy transport constraints, are currently not considered in the optimization.
-Transitioning to an hourly representation with more representative nodes would increase the model's computational time and complexity, but would also enhance its robustness and validity.
+Transitioning to an hourly representation and a higher spatial resolution would increase the model's computational time and complexity, but would also enhance its robustness and validity.
 
@@
 
-In this study, we investigated the impact of a \carbon transport system with different carbon sequestration potentials on the optimal configuration of technologies in a fully sector-coupled energy system in Europe, focusing on the deployment of carbon capture and Utilization (CCU) and Carbon Capture and Storage (CCS) technologies.
+In this study, we investigated the impact of a \carbon{} transport system with different carbon sequestration potentials on the optimal configuration of technologies in a fully sector-coupled energy system in Europe, focusing on the deployment of carbon capture and Utilization (CCU) and Carbon Capture and Storage (CCS) technologies.
 Our results indicate that increasing carbon sequestration potential leads to a decrease in total system costs, with a more pronounced cost reduction observed when a carbon transport network is present.
@@
 
-Furthermore, we observed that with a \carbon transport network, the system tends to build out more carbon capture at point sources.
+Furthermore, we observed that with a \carbon{} transport network, the system tends to build out more carbon capture at point sources.
 With an increased sequestration potential, the share of technologies equipped with integrated carbon capture rises, particularly for Steam Methane Reforming (SMR) and Gas Combined Heat and Power (CHP) plants.
-Independent of a setup with or without \carbon transport, the spatial distribution of carbon capture technologies becomes more centralized, with installations concentrated near sequestration sites.
+Independent of a setup with or without \carbon{} transport, the spatial distribution of carbon capture technologies becomes more centralized, with installations concentrated near sequestration sites.
 
@@
 As Europe strives to achieve its 2050 climate neutrality target, understanding the interplay between carbon sequestration, CCU, and CCS technologies will be crucial for the design of effective policies and strategies.
-
-\newpage
 
@@
 %     \includegraphics[width=\linewidth]{operation_area_carbon_noco2network.pdf}
-%     \caption{Balance of captured \carbon emissions for the optimal operation without \carbon network.}
+%     \caption{Balance of captured \carbon{} emissions for the optimal operation without \carbon{} network.}
 %     \label{fig:operation_area_carbon_noco2network}

### 2023-06-13 report/submissions/eem/presentation/presentation.tex
checkpoint: eem23 presentation plots

\documentclass[12pt, aspectratio=169]{beamer}
\usetheme{metropolis}
\usepackage{graphicx}
\usepackage{xcolor}

\newcommand{\carbon}{\textcolor{teal}{CO$_2$}}
\newcommand{\hydrogen}{H$_2$}

\title{Designing a \carbon{} Network for a Carbon-Neutral European Economy}
\author{\textbf{Fabian Hofmann}, Christoph Tries, Fabian Neumann, Lisa Zeyen, Tom Brown}
\institute{Technical University Berlin}
\date{}
\urlstyle{blue}

\graphicspath{
    {figures},
}

\begin{document}

\begin{frame}[plain]
    \titlepage
\end{frame}

\begin{frame}
    \begin{columns}
        \column{0.4\textwidth}
        European \carbon{} emissions have to decrease \textbf{quickly}.
        \column{0.6\textwidth}
        \includegraphics[width=\linewidth]{emissions-timeline}
    \end{columns}
\end{frame}

\begin{frame}
    \begin{columns}
        \column{0.4\textwidth}
        Particular focus is needed on \textbf{hard-to-abate emissions} \\
        \vspace*{1em}
        \visible<2->{Formulate a strategy for \\ \textbf{carbon management}}
        \visible<3>{
            \begin{itemize}
                \item \carbon{} Capture
                \item \carbon{} Utilization
                \item \carbon{} Transport
                \item \carbon{} Sequestration
            \end{itemize}

        }
        \column{0.6\textwidth}
        \includegraphics[width=\linewidth]{emissions-timeline-alerted}
    \end{columns}
\end{frame}

\section*{Model Setup}

\begin{frame}
    \begin{columns}
        \column{0.55\textwidth}
        Investment optimization with \textbf{PyPSA-Eur}
        \visible<2>{
            \begin{itemize}
                \item European countries (EU+UK+NO+CH)
                \item 90 nodes
                \item 3 hour resolution
                \item electricity, heating, transport, industry \& waste sectors
                \item carbon management technologies (DAC, CC at point sources, Fischer-Trosch, Methanolization, )
            \end{itemize}
        }
        \column{0.5\textwidth}
        \includegraphics[width=\linewidth]{brownfield_capacities}
    \end{columns}
\end{frame}

\begin{frame}[plain]
    \begin{columns}
        \column{0.3\textwidth}
        Perform 2 $\times$ 5 model runs with
        \carbon{}~\textbf{transport network} on/off and 5~different rates of annual \textbf{sequestration}.\\
        \column{0.7\textwidth}
        \includegraphics[width=\linewidth]{sequestration_map.png}
    \end{columns}
\end{frame}

\section*{System Costs}
\begin{frame}[plain]
    % \qquad \textbf{System costs for different sequestration rates}
    \includegraphics[width=\linewidth]{cost_area.pdf}
\end{frame}

\section*{Carbon Capture Technologies}
\begin{frame}
    \qquad \textbf{Captured carbon balance with a CO$_2$ network}
    \begin{columns}
        \column{0.3\textwidth}
        \begin{itemize}
            \item  Fischer-Tropsch fuels are replaced by fossils fuels
            \item \hydrogen{} infrastructure is replaced by CCS
        \end{itemize}
        \column{0.7\textwidth}
        \begin{figure}
            \includegraphics[width=\linewidth]{operation_area_carbon_co2network.pdf}
            \par\footnotesize{*with Carbon Capture}
        \end{figure}
    \end{columns}
\end{frame}

\begin{frame}
    \textbf{Carbon Capture (left) and Utilization \& Storage (right) for 200~Mt$_{\text{seq}}$/a}
    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_200.png}
\end{frame}

\begin{frame}
    \textbf{Carbon Capture (left) and Utilization \& Storage (right) for 400~Mt$_{\text{seq}}$/a}
    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_400.png}
\end{frame}

\begin{frame}
    \textbf{Carbon Capture (left) and Utilization \& Storage (right) for 600~Mt$_{\text{seq}}$/a}
    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_600.png}
\end{frame}

\begin{frame}
    \textbf{Carbon Capture (left) and Utilization \& Storage (right) for 800~Mt$_{\text{seq}}$/a}
    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_800.png}
\end{frame}

\begin{frame}
    \textbf{Carbon Capture (left) and Utilization \& Storage (right) for 1000~Mt$_{\text{seq}}$/a}
    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_1000.png}
\end{frame}

\section*{Conclusion}

\begin{frame}
    \textbf{Conclusion}
    \begin{itemize}
        \pause
        \item \carbon{} transport is relatively cheap and leads to large system cost-benefits
              \pause
        \item Competition between CCU + \hydrogen{} network and Fossils + CCS
              \pause
        \item Optimal investments strongly dependent on sequestration rates
              \pause
        \item CCU is at point-sources in the inland while DAC is at the coasts
              \pause
        \item Sequestration technology is still to be elaborated $\rightarrow$ take results with care
    \end{itemize}
\end{frame}

\end{document}

### 2023-06-13 workflow/scripts/plot-operation-map.py

@@ -103,7 +103,8 @@ for tag, ax in zip(tags, axes.flatten()):
     )
 
     ax.set_extent(snakemake.config["plotting"]["extent"])
-    ax.set_title(labels[kind] + " " + tag.title())
+    if snakemake.params.settings.get("title", True):
+        ax.set_title(labels[kind] + " " + tag.title())
 
     legend_bus_sizes = specs["bus_sizes"]
     if legend_bus_sizes is not None:
@@ -118,7 +119,5 @@ for tag, ax in zip(tags, axes.flatten()):
 fig.tight_layout()
 fig.savefig(
     snakemake.output.map,
-    bbox_inches="tight",
     dpi=300,
-    facecolor="white",
 )

### 2023-09-19 report/sections/abstract.tex
init work for journal paper

--- 2022-11-30 report/sections/abstract.tex
@@
-The transformation of the global energy system to reach climate-neutrality requires large investments into new technologies. \dots
+Achieving a carbon-neutral European economy requires internationally coordinated efforts across all energy sectors, especially those with hard-to-abate emissions like heavy industry.
+Against this backdrop, It is critical to assess the need for comprehensive carbon management involving technologies such as carbon capture, transport, use, and storage to develop an effective carbon strategy for climate neutrality.
+
+This paper provides the first cost-optimal design for a European energy system that fully incorporates carbon management technologies including a carbon transport network, and considers all emissions-intensive sectors.
+It optimizes carbon technologies, hydrogen transport and storage, and the design of renewable energy sources.
+
+We show that carbon management technologies aiming at net-zero emissions provide systemic flexibility to the European energy system.
+A carbon network is cost-effective at different carbon capture rates (200 - 1000 Mt/a), leading to average cost savings of €13 billion per year.
+The \carbon{} network enables viable point source carbon capture, reducing the need for direct air capture facilities.
+We show that high sequestration rates are cost-beneficial by allowing continued fossil fuel use in some sectors, but require massive capacity expansion which could jeopardize timely attainment of climate neutrality.

### 2023-09-19 report/sections/methods.tex
init work for journal paper

@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{sequestration_map.pdf}
-    \caption{Sequestration map}
-    \label{fig:sequestration_map}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=\linewidth]{sequestration_map.pdf}
+%     \caption{Sequestration map}
+%     \label{fig:sequestration_map}
+% \end{figure}
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{capacity_map_electricity_co2network_1000.pdf}
-    \caption{Optimal capacities per of the electricity sector for a sequestration of 1000 Mt/a.}
-    \label{fig:capacity_map_noco2network_1000}
-\end{figure}
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=\linewidth]{capacity_map_electricity_co2network_1000.pdf}
+%     \caption{Optimal capacities per of the electricity sector for a sequestration of 1000 Mt/a.}
+%     \label{fig:capacity_map_noco2network_1000}
+% \end{figure}
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{capacity_map_carbon_co2network_1000.pdf}
-    \caption{Optimal capacities per of the carbon sector for a sequestration of 1000 Mt/a.}
-    \label{fig:capacity_map_noco2network_1000}
-\end{figure}
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=\linewidth]{capacity_map_carbon_co2network_1000.pdf}
+%     \caption{Optimal capacities per of the carbon sector for a sequestration of 1000 Mt/a.}
+%     \label{fig:capacity_map_noco2network_1000}
+% \end{figure}
 
-\begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_1000.pdf}
-    \caption{Optimal operation per sector for a sequestration of 1000 Mt/a.}
-    \label{fig:operation_map_noco2network_1000}
-\end{figure*}
+% \begin{figure*}
+%     \centering
+%     \includegraphics[width=\linewidth]{operation_map_carbon_co2network_1000.pdf}
+%     \caption{Optimal operation per sector for a sequestration of 1000 Mt/a.}
+%     \label{fig:operation_map_noco2network_1000}
+% \end{figure*}
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{operation_area_carbon_co2network.pdf}
-    \caption{Balance of captured CO$_2$ emissions for the optimal operation with CO$_2$ network.}
-    \label{fig:operation_area_carbon_co2network}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=\linewidth]{operation_area_carbon_co2network.pdf}
+%     \caption{Balance of captured CO$_2$ emissions for the optimal operation with CO$_2$ network.}
+%     \label{fig:operation_area_carbon_co2network}
+% \end{figure}
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{operation_area_carbon_noco2network.pdf}
-    \caption{Balance of captured CO$_2$ emissions for the optimal operation without CO$_2$ network.}
-    \label{fig:operation_area_carbon_noco2network}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=\linewidth]{operation_area_carbon_noco2network.pdf}
+%     \caption{Balance of captured CO$_2$ emissions for the optimal operation without CO$_2$ network.}
+%     \label{fig:operation_area_carbon_noco2network}
+% \end{figure}
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{operation_bar_co2network_200.pdf}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=\linewidth]{operation_bar_co2network_200.pdf}
+% \end{figure}

### 2023-09-19 report/submissions/journal/paper/paper.tex
init work for journal paper

--- 2023-06-13 report/submissions/eem/paper/paper.tex
@@
-\documentclass[conference]{IEEEtran}
+\documentclass[twocolumn]{article}
 \usepackage{amsmath,amssymb,amsfonts}
@@
 
-% \IEEEoverridecommandlockouts
-% % The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-% \fancypagestyle{firstpage}{% Page style for first page
-%   \fancyhf{}% Clear header/footer
-%   \fancyfoot[l]{}% Footer
-% }
-% \thispagestyle{firstpage}
-
-% \usepackage{cleveref}
-
-% \crefname{align}{}{}
-% \Crefname{align}{Equation}{Equations}
-
-\IEEEoverridecommandlockouts
-\IEEEpubid{\makebox[\columnwidth]{979-8-3503-1258-4/23/\$31.00 ©2023 IEEE \hfill} \hspace{\columnsep}\makebox[\columnwidth]{ }}
-
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
 \usepackage[%
@@
 \bibliography{../../references.bib}
-% Template from https://de.overleaf.com/latex/templates/ieee-conference-template/grfzhhncsfqn
 
@@
 
-% \thanks{Fabian Hofmann is funded by the Breakthrough Energy Project "Hydrogen Integration and Carbon Management in Energy System Models".}
-
 \author{
-    \IEEEauthorblockN{Fabian Hofmann, Christoph Tries, Fabian Neumann, Lisa Zeyen, Tom Brown}
-    \IEEEauthorblockA{\textit{Institute of Energy Technology} \\
+    Fabian Hofmann, Christoph Tries, Fabian Neumann, Lisa Zeyen, Tom Brown \\
+    \textit{Institute of Energy Technology} \\
     \textit{Technical University of Berlin}\\
     Berlin, Germany \\
-        m.hofmann@tu-berlin.de}
-    % \IEEEauthorblockN{Anonym Author}
-
+    m.hofmann@tu-berlin.de
 }
@@
 % However, such high rates of sequestration would require an unprecedented expansion from today's low sequestration capacities and, in case it is not feasible, may delay reaching net-zero compared to other abatement strategies.
-
-\begin{IEEEkeywords}
-    carbon management, energy system modelling, CCU, CCS, sector-coupling
-\end{IEEEkeywords}
 

### 2023-09-19 workflow/scripts/common.py

@@ -460,7 +460,7 @@ def add_carrier_nice_names(n):
     # replace abbreviations with capital letters
     replace = {
         "H2": "H$_2$",
-        "Co2 Stored": "CO$_2$ Sequestration",
+        "Hydrogen": "H$_2$",
         "Co2": "CO$_2$",
         "Chp": "CHP",
         "Dac": "DAC",
@@ -474,7 +474,8 @@ def add_carrier_nice_names(n):
         "Oil Emissions": "Aviation and Petrochemical\nEmissions",
         "Allam": "Allam Cycle",
     }
-    n.carriers.nice_name.replace(replace, regex=True, inplace=True)
+    nice_names = n.carriers.nice_name.str.title()
+    n.carriers.nice_name = nice_names.replace(replace, regex=True)
 
 
 def add_colors(n):

### 2023-09-19 workflow/scripts/plot-cost-bar.py

@@ -13,7 +13,7 @@ alpha = 1
 region_alpha = 0.8
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
-    snakemake = mock_snakemake("plot_cost_area", ext="png", clusters=90)
+    snakemake = mock_snakemake("plot_cost_bar", ext="pdf", clusters=90)
 
 sns.set_theme(**snakemake.params["theme"])
 
@@ -54,7 +54,7 @@ fig, ax = plt.subplots(
 )
 
 
-grouped.T.plot(kind="area", stacked=True, ax=ax, color=colors, alpha=0.8, lw=0)
+grouped.T.plot(kind="bar", stacked=True, ax=ax, color=colors, alpha=0.8, lw=0, rot=0)
 
 ax.axhline(0, color="k", lw=1)
 ax.set_ylabel(f"System cost [{unit}]")

### 2023-09-22 report/submissions/journal/paper/paper.tex
update plots

@@
 ]{biblatex}
-\bibliography{../../references.bib}
+\bibliography{../../../references.bib}
 
@@
 
-\title{Designing a \carbon{} Network for a Carbon-Neutral European Economy}
+\title{Designing a \carbon{} Network for a Net-Zero European Energy System}
 
@@
 \begin{abstract}
-    Achieving a carbon-neutral European economy requires internationally coordinated efforts across all energy sectors, especially those with hard-to-abate emissions like heavy industry.
+    The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors, particularly in emissions-intensive industries like heavy manufacturing.
+    A gap exists in understanding how comprehensive carbon management technologies, including carbon capture, transport, use, and storage, can be effectively integrated into a European energy system for climate neutrality.
+    To address this, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates carbon management technologies, hydrogen transport and storage, and renewable energy sources with a high spatial and temporal resolution.
+    Our findings reveal that implementing carbon management technologies can offer systemic flexibility to the European energy system.
+Specifically, a carbon transport network facilitates point source carbon capture, thereby reducing the necessity for direct air capture installation.
+The cost reduction of  average annual savings of €xx billion.
+    Overall, our work demonstrates the cost-effectiveness of a multi-grid system that includes both hydrogen and carbon transport networks and power grid expansion to achieve climate neutrality in Europe.
+
+    % Achieving a carbon-neutral European economy requires internationally coordinated efforts across all energy sectors, especially those with hard-to-abate emissions like heavy industry.
 Against this backdrop, It is critical to assess the need for comprehensive carbon management involving technologies such as carbon capture, transport, use, and storage to develop an effective carbon strategy for climate neutrality.
 
-    This paper provides the first cost-optimal design for a European energy system that fully incorporates carbon management technologies including a carbon transport network, and considers all emissions-intensive sectors.
+    % This paper provides the first cost-optimal design for a European energy system that fully incorporates carbon management technologies including a carbon transport network, and considers all emissions-intensive sectors.
 It optimizes carbon technologies, hydrogen transport and storage, and the design of renewable energy sources.
 
-    We show that carbon management technologies aiming at net-zero emissions provide systemic flexibility to the European energy system.
-A carbon network is cost-effective at different carbon capture rates (200 - 1000 Mt/a), leading to average cost savings of €13 billion per year.
+    % We show that carbon management technologies aiming at net-zero emissions provide systemic flexibility to the European energy system.
+A carbon network is cost-effective leading to average cost savings of €xx billion per year.
 The \carbon{} network enables viable point source carbon capture, reducing the need for direct air capture facilities.
-We show that high sequestration rates are cost-beneficial by allowing continued fossil fuel use in some sectors, but require massive capacity expansion which could jeopardize timely attainment of climate neutrality.
+The paper underpins the need for a complementing hydrogen network and \carbon{} network to achieve climate neutrality.
 
 \end{abstract}
-
-% Enabling a carbon-neutral European economy requires internationally coordinated investments in all sectors of the energy system.
-In the light of hard-to-abate emissions from heavy industry including feedstocks, dense fuels needed for transport and backup heat or power sources, it is important to determine the extent to which a system-wide carbon management is needed.
-Different technologies such as carbon capture, transport, utilization and storage build the basis to formulate a carbon strategy to take the final step to climate-neutrality.
-
-% In this paper, we present the first study of a cost-optimal design of the European energy system with a full representation of carbon management technologies including a \carbon{} transport network, considering all sectors with high spatial resolution.
-Among other technologies, we jointly optimize carbon technologies, hydrogen transport and storage, and the design of renewable energy sources in the system.
-
-% Assuming net-zero emissions, we argue that carbon management technologies provide systemic flexibility to the European energy system.
-In particular, we show that building a \carbon{} network is cost-effective across different rates of carbon sequestration in the offshore regions (200 -- 1000 Mt/a), leading to cost reductions of 13bn€/a on average.
-The network drives the feasibility of carbon capture at point sources which is preferred over building large amounts of direct air capture facilities.
-Moreover, we show that high sequestration rates, regardless of carbon network implementation, reduce the need for a hydrogen network and synthetic fuel production which allows fossil fuels to continue to be used in industry, aviation, agriculture, and the heating sector.
-% However, such high rates of sequestration would require an unprecedented expansion from today's low sequestration capacities and, in case it is not feasible, may delay reaching net-zero compared to other abatement strategies.
 

### 2023-09-22 workflow/scripts/plot-cost-bar.py

@@ -1,5 +1,4 @@
 import os
-from pathlib import Path
 import seaborn as sns
 import matplotlib.pyplot as plt
 import pandas as pd

### 2023-09-22 workflow/scripts/plot-operation-area.py

@@ -14,7 +14,7 @@ import matplotlib.pyplot as plt
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
-        "plot_operation_area", clusters=40, kind="carbon", ext="png"
+        "plot_operation_area", clusters=40, kind="hydrogen", ext="pdf"
     )
 
 sns.set_theme(**snakemake.params["theme"])

### 2023-09-22 workflow/scripts/plot-operation-bar.py

@@ -13,10 +13,10 @@ import matplotlib.pyplot as plt
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
         "plot_operation_bar",
-        kind="carbon",
+        kind="hydrogen",
         run="half-price",
         clusters=40,
-        ext="png",
+        ext="pdf",
     )
 
 sns.set_theme(**snakemake.params["theme"])

### 2023-09-27 report/submissions/ecemp/paper/paper.tex
udpate snakefile to auto create presentation in presentation mode

\documentclass[twocolumn]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{cuted}  % Add the cuted package
\usepackage{caption}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{tabularx, multirow}
\usepackage{fancyhdr,lipsum}

% \IEEEoverridecommandlockouts
% % The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
% \fancypagestyle{firstpage}{% Page style for first page
%   \fancyhf{}% Clear header/footer
%   \fancyfoot[l]{}% Footer
% }
% \thispagestyle{firstpage}

% \usepackage{cleveref}

% \crefname{align}{}{}
% \Crefname{align}{Equation}{Equations}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\usepackage[%
backend=biber,bibencoding=utf8, %instead of bibtex
language=auto,
style=ieee,
sorting=none, % nyt for name, year, title
maxbibnames=10, % default: 3, et al.
%backref=true,%
natbib=true % natbib compatibility mode (\citep and~\citet still work)
]{biblatex}
\bibliography{../../references.bib}
% Template from https://de.overleaf.com/latex/templates/ieee-conference-template/grfzhhncsfqn

%define approx proportional
\def\app#1#2{%
  \mathrel{%
    \setbox0=\hbox{$#1\sim$}%
    \setbox2=\hbox{%
      \rlap{\hbox{$#1\propto$}}%
      \lower1.1\ht0\box0%
    }%
    \raise0.25\ht2\box2%
  }%
}
\def\approxprop{\mathpalette\app\relax}

% make abbreviation for co2
\newcommand{\carbon}{CO$_2$}
\newcommand{\hydrogen}{H$_2$}

\graphicspath{
    {paper-figures},
}

\begin{document}

\title{\carbon{} Management Systems in a Carbon Neutral European Economy}

% \thanks{Fabian Hofmann is funded by the Breakthrough Energy Project "Hydrogen Integration and Carbon Management in Energy System Models".}
\author{
    Fabian Hofmann\textsuperscript{1,2},
    Christoph Tries\textsuperscript{1},
    Fabian Neumann\textsuperscript{1},
    Lisa Zeyen\textsuperscript{1},
    Tom Brown\textsuperscript{1}
    \vspace{.5em}
    \\
    \textsuperscript{1}Institute of Energy Technology, Technical University of Berlin, Berlin, Germany \\
    \textsuperscript{2}m.hofmann@tu-berlin.de
}

\maketitle

\begin{abstract}
    This paper presents a cost-optimal approach to a carbon-neutral European energy system, emphasizing the need for comprehensive carbon management across all emissions-intensive sectors. It is highlighted that carbon management technologies offer systemic flexibility and are cost-effective, which could save an average of EUR 13 billion annually. Incorporating a carbon network supports point source carbon capture and reduces the need for direct air capture systems. The paper points out that the assumption of high sequestration rates leads to continued use of fossil fuels in some sectors while reducing costs, but should be viewed with caution because significant capacity expansion would be required that could jeopardize timely climate neutrality.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

Carbon capture is recognized as a pivotal technology for reaching net-zero economies, particularly in offsetting emissions from hard-to-abate sectors, including land use and agriculture~\cite{arniehellerNewCarbonEconomy2019}. Carbon Capture and Utilization (CCU), Carbon Capture and Storage (CCS) and Carbon Transport, collectively referred to as carbon management, create new opportunities for producing carbon-neutral hydrocarbons and storing carbon in geological formations. The European region has an estimated potential for capturing roughly 1.4 Gt of carbon from point sources per annum, demonstrating a significant potential for decarbonization~\cite{ToolsGreenTransition,weiProposedGlobalLayout2021}. Recent industry and policymaker commitment in Europe to carbon management projects indicates emerging infrastructure components and business models~\cite{adomaitisEquinorRWEBuild2023,apnewswireGermanyDrawLegislation2023,KohlenstoffKannKlimaschutz2023,OGETESJoin2022,TESHydrogenLife2023}.

In this paper, we introduce a comprehensive analysis of the European energy system for 2050 using the PyPSA-Eur-Sec energy system model~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}. Our analysis highlights the dynamics of carbon capture, use, and storage for system with and without \carbon{} transport, and the prioritization of decarbonization and fuel switching in an energy system with limited annual sequestration potential.

\section{Methodology}
\label{sec:methodology}

\begin{figure*}[ht!]
    \centering
    \includegraphics*[width=\linewidth]{cost_area.pdf}
    \caption[short]{Total annual system cost for the sector-coupled system with different levels of carbon sequestration potential, with (left) and without (right) \carbon{} network. "Gas Infrastructure" combines gas facilities for power and heat production, "H$_2$ Infrastructure" combines H$_2$ production, transport and re-electrification. Regardless of the implementation of a \carbon{} network, the system cost decrease as sequestration increases. Due to an increased flexibility from fossil carriers with subsequent sequestration, the need for FT synthesis and \hydrogen{} electrolysis and the corresponding renewable power supply is reduced.
        % TODO: mention details on the grouped technologies
    }
    \label{fig:cost_area}
\end{figure*}

We base our analysis on the open-source model PyPSA-Eur~\cite{PyPSAEurSecSectorCoupledOpen2023}. It is a fully sector-coupled model with high spatial and temporal resolution that enables a joint optimization of investment and operational decisions in energy generation, storage, conversion, and transport infrastructures. In this study, a version with 90 European regions and a 3-hour time resolution is used. The model considers regional energy demand in sectors such as electricity, industry, transport, buildings, and agriculture, and sets a binding upper limit for zero emissions. Limits are also set for biomass use and expansion of the electricity transmission grid.

% TODO: Mention exogenous decisions on carbon capture and 90% rates here for industry.

The study only considers offshore sites for carbon sequestration due to their larger potential compared to onshore sites and public safety concerns and assumes a conservative approach to estimating storage potential, limiting it to 25~Mt per site and calculating annual storage availability over a 25-year period.

The model includes three drop-in fuel production technologies: methanation, methanolization, and Fischer-Tropsch (FT) synthesis. Synthetic methane substitutes natural gas or biogas, synthetic methanol helps decarbonize the marine industry, while Fischer-Tropsch fuels replace fossil oil for various industrial and aviation uses. All the technology costs for 2050 used in the model are derived from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.

In total, we run 10 simulations: To investigate the role of the \carbon{} network, we run different setups with the \carbon{} network on and off. To analyze the impact of sequestration, for each setup we vary the global sequestration potential from 200~Mt to 1000~Mt per year in steps of 200~Mt.

\section{Results}

As displayed in Figure~\ref{fig:cost_area}, the model results show that the majority of annual system costs are associated with renewable energy sources, heating, hydrogen technologies, and the electricity grid, while costs for fossil fuels and carbon capture and transport technologies are relatively low. As sequestration potential increases, costs decrease, with high sequestration rates allowing for continued use of fossil fuels and increased deployment of carbon removal technologies. Notably, a system with carbon transport is on average 1.6\% less expensive across all sequestration levels.

In particular, we perceive a clear shift in technology when increasing the sequestration rate: While FT synthesis decreases, integrated carbon capture from Gas Combined Heat and Power (CHP) plants, Allam Cycle plants, Steam Methane Reforming (SMR) and DAC increase. Other technologies are constant across varying levels of sequestration due to different reasons: biomass is bound by its maximum available potential (as reported in~\cite{europeancommissionjointresearchcentreENSPRESOBIOMASS2019}); the industry processes with their associated emissions are set to fixed values; the methanol production is bound by a lower limit representing the demand from the shipping sector.

As for hydrogen, the model with a \carbon{} transport system shows a 50\% decrease in hydrogen production as sequestration increases, with hydrogen production from steam methane reforming (SMR) beginning to replace electrolysis at high sequestration rates. With the decline of FT synfuel production, the demand for hydrogen declines, leaving only demand from the shipping, industry, and transport sectors.

\section{Limitations}
\label{sec:limitations}

The model's validity is limited due to its reliance on linear optimization with perfect foresight. Technology cost assumptions are based on historical learning rates, which might not reflect future trends and cannot account for market disruptions like the 2022 gas price peak. Lastly, the model's spatial and temporal resolution does not capture all relevant dynamics and variations in energy supply and demand, although increasing these would raise computational time and complexity.

\section{Conclusion}
\label{sec:conclusion}

This study explored the impact of a carbon transport system on the optimal configuration of technologies in a fully sector-coupled European energy system, focusing on carbon capture and Utilization (CCU) and Carbon Capture and Storage (CCS). Increasing carbon sequestration potential lowers total system costs, more significantly when a carbon transport network exists. Higher sequestration rates shift focus from CCU and hydrogen infrastructure to CCS and fossil fuels, offset by carbon removal elsewhere. A carbon transport network encourages more carbon capture at point sources. Despite limitations, our findings provide insights into the role of carbon sequestration and transport in achieving climate neutrality.
While carbon sequestration offers promising possibilities for climate mitigation, it is not yet fully matured and proven as a safe, reliable technology, thus caution is warranted against over-reliance on it as a sole solution.

\printbibliography

% \appendix

% \begin{figure}[h]
%     \centering
%     \includegraphics[width=\linewidth]{capacity_map_electricity_co2network_1000.png}
%     \caption{Optimal capacities per of the electricity sector for a sequestration of 1000 Mt/a.}
%     \label{fig:capacity_map_electricity_co2network_1000}
% \end{figure}

% \begin{figure}[h]
%     \centering
%     \includegraphics[width=\linewidth]{capacity_map_hydrogen_co2network_1000.png}
%     \caption{Optimal capacities per of the hydrogen sector for a sequestration of 1000 Mt/a.}
%     \label{fig:capacity_map_hydrogen_co2network_1000}
% \end{figure}

% \begin{figure}
%     \centering
%     \includegraphics[width=\linewidth]{operation_area_carbon_noco2network.pdf}
%     \caption{Balance of captured \carbon{} emissions for the optimal operation without \carbon{} network.}
%     \label{fig:operation_area_carbon_noco2network}
% \end{figure}

\end{document}

### 2023-09-27 workflow/scripts/plot-captureshare-line.py

@@ -37,7 +37,7 @@ for path in snakemake.input.networks:
     )
     caps = caps["cc"] / caps.sum(1)
 
-    key = snakemake.config["labels"][n.meta["wildcards"]["run"]]
+    key = n.meta["label"]
     df[key] = caps
 
 df = pd.concat(df, axis=1)

### 2023-09-27 workflow/scripts/plot-cost-diff-bar.py

@@ -30,7 +30,7 @@ for path in snakemake.input.networks:
     costs = costs.droplevel(0)[lambda x: x > 0]
     costs = costs.groupby(costs.index).sum()
 
-    key = snakemake.config["labels"][n.meta["wildcards"]["run"]]
+    key = n.meta["label"]
 
     df[key] = costs
     objectives[key] = n.objective

### 2023-09-27 workflow/scripts/plot-energy-balance-area.py

@@ -34,7 +34,7 @@ label = config["labels"].get(kind, kind.title())
 
 
 n = import_network(snakemake.input.network)
-key = snakemake.config["labels"][n.meta["wildcards"]["run"]]
+key = n.meta["label"]
 balance = n.statistics.energy_balance(aggregate_time=False)
 
 

### 2023-09-27 workflow/scripts/plot-operation-area.py

@@ -34,7 +34,7 @@ for path in snakemake.input.networks:
     if (prod.sum() / cons.sum()).round(3) != 1:
         print(f"Warning: {kind} production and consumption are not equal.")
 
-    key = snakemake.config["labels"][n.meta["wildcards"]["run"]]
+    key = n.meta["label"]
     df[key] = pd.concat([prod, cons], keys=["Production", "Consumption"])
 
 df = pd.concat(df, axis=1)

### 2023-09-27 workflow/scripts/leftover/leftover.py
follow up

@@ -176,3 +176,22 @@ def create_dummy_network(n):
     dummy.madd("Link", connections.index, **connections)
     dummy.madd("Carrier", n.carriers.index, **n.carriers)
     return dummy
+
+
+# rule create_gifs:
+#     input:
+#         # use input function here
+#         expand(
+#             "results/comparison-{comparison}/gifs/{clusters}_nodes/capacity_map_{kind}.gif",
+#             kind=config["constants"]["kinds"],
+#             clusters=config["scenario"]["clusters"],
+#         ),
+#         expand(
+#             "results/comparison-{comparison}/gifs/{clusters}_nodes/operation_map_{kind}.gif",
+#             kind=config["constants"]["kinds"],
+#             clusters=config["scenario"]["clusters"],
+#         ),
+#     output:
+#         "results/comparison-{comparison}/gifs/{clusters}_nodes/.gifs_created",
+#     shell:
+#         "touch {output}"

### 2023-09-27 workflow/scripts/plot-cost-diff-bar.py

@@ -40,7 +40,7 @@ df = pd.concat(df, axis=1).fillna(0)
 norm = 1e9
 unit = "bn€/a"
 
-diff = (df[labels["baseline"]] - df[labels["projected-price"]]).div(norm)
+diff = (df[df.columns[0]] - df[df.columns[1]]).div(norm)
 
 
 fig, ax = plt.subplots(

### 2023-09-27 workflow/scripts/plot-energy-balance-bar.py

@@ -17,18 +17,12 @@ from common import (
 )
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
-    snakemake = mock_snakemake(
-        "plot_energy_balance_bar", ext="pdf", kind="process emissions", clusters=40
-    )
+    snakemake = mock_snakemake("plot_energy_balance_bar", ext="pdf", clusters=40)
 
 sns.set_theme(**snakemake.params["theme"])
 
 config = snakemake.config
-kind = snakemake.wildcards.kind
-kind_to_carrier = {"hydrogen": "H2", "carbon": "co2 stored", "electricity": "AC"}
-carrier = kind_to_carrier.get(kind, kind)
-label = config["labels"].get(kind, kind.title())
-
+norm = 1e6
 
 df = {}
 objectives = {}
@@ -41,44 +35,49 @@ for path in snakemake.input.networks:
     df[key] = balance
     objectives[key] = n.objective
 
-norm = 1e6
-unit = "TWh" if kind not in ["carbon", "co2"] else "Mt"
 
 df = pd.concat(df, axis=1)
 
-ds = df.xs(carrier, level=2)
-ds = ds.div(norm)
-ds = ds[ds.round(2) != 0].droplevel(0).dropna(how="all").fillna(0)
-ds = ds.groupby(level=0).sum()
-ds = ds.sort_values(ds.columns[0], ascending=False)
-# ds = sort_rows_by_diff(ds)
+for kind, output in snakemake.output.items():
 
-fig, ax = plt.subplots(1, 1, figsize=snakemake.params.settings["figsize"])
+    kind_to_carrier = {"hydrogen": "H2", "carbon": "co2 stored", "electricity": "AC"}
+    carrier = kind_to_carrier.get(kind, kind)
+    label = config["labels"].get(kind, kind.title())
+    unit = "TWh" if kind not in ["carbon", "co2"] else "Mt"
 
-colors = n.carriers.set_index("nice_name").color.to_dict()
-ds.T.plot.bar(color=colors, ax=ax, stacked=True, alpha=0.8, lw=0, rot=0)
+    ds = df.xs(carrier, level=2)
+    ds = ds.div(norm)
+    ds = ds[ds.round(2) != 0].droplevel(0).dropna(how="all").fillna(0)
+    ds = ds.groupby(level=0).sum()
+    ds = ds.sort_values(ds.columns[0], ascending=False)
+    # ds = sort_rows_by_diff(ds)
 
[…]

### 2023-10-04 workflow/scripts/common.py
add transmission bar plot fix biogas CC upgrade

@@ -349,7 +349,7 @@ def get_carrier_transport(n, kind, config, which="capacity"):
         Whether to return the capacity or the operation of the carriers, by default 'capacity'
     """
     weights = n.snapshot_weightings.generators
-    specs = config["constants"]["kind_to_carrier"][kind].get("transport", {})
+    specs = config["constants"]["kind_to_carrier"].get(kind, {}).get("transport", {})
     res = {}
 
     carriers = specs.get("Link", [])
@@ -447,6 +447,7 @@ def modify_carrier_names(n):
         n.one_port_components | n.branch_components | {"Bus"}
     ):
         c.df.carrier.replace(replace, "", regex=True, inplace=True)
+    n.carriers = n.carriers.sort_values("nice_name")
     n.carriers.index = n.carriers.index.to_series().replace(replace, "", regex=True)
     n.carriers.nice_name = n.carriers.nice_name.replace(replace, "", regex=True)
     n.carriers = n.carriers[~n.carriers.index.duplicated()]

### 2023-10-04 workflow/scripts/plot-operation-map.py

@@ -15,6 +15,7 @@ from pypsa.plot import add_legend_circles, add_legend_patches, add_legend_lines
 from common import (
     import_network,
     mock_snakemake,
+    get_transmission_links,
     get_carrier_consumption,
     get_carrier_transport,
     get_carrier_production,
@@ -30,9 +31,8 @@ region_alpha = 0.8
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
         "plot_operation_map",
-        kind="carbon",
-        run="half-price",
-        clusters=40,
+        run="baseline",
+        clusters=90,
         ext="png",
     )
 
@@ -43,6 +43,16 @@ regions = gpd.read_file(snakemake.input.onshore_regions).set_index("name")
 which = "operation"
 config = snakemake.config
 
+is_transport = get_transmission_links(n)
+transport_carriers = [
+    *n.links.carrier[is_transport].unique(),
+    *n.lines.carrier.unique(),
+]
+transport_carriers = n.carriers.nice_name[transport_carriers]
+
+balance = n.statistics.energy_balance(aggregate_bus=False)
+balance = balance.rename(n.buses.location, level=3)
+dispatch = balance.drop(transport_carriers, level=1)
 
 for kind, output in snakemake.output.items():
 
@@ -53,6 +63,10 @@ for kind, output in snakemake.output.items():
         squeeze=False,
         subplot_kw={"projection": ccrs.EqualEarth()},
     )
+
+    carriers = config["constants"]["carrier_to_buses"].get(kind, [kind])
+    carriers = list(set(carriers) & set(dispatch.index.get_level_values(2)))
+    df = dispatch.loc[:, :, carriers].groupby(level=[3, 1]).sum()
 
     tags = ["production", "consumption"]
 
@@ -65,9 +79,9 @@ for kind, output in snakemake.output.items():
         unit = specs["region_unit"]
 
         if tag == "production":
-            bus_sizes = get_carrier_production(n, kind, config, which)
+            bus_sizes = df[df > 0]
         else:
-            bus_sizes = get_carrier_consumption(n, kind, config, which)
+            bus_sizes = df[df < 0].abs()
         branch_widths = get_carrier_transport(n, kind, config, which)
         flow = pd.concat(branch_widths)
         branch_carriers = get_carrier_transport(n, kind, config, "carrier")
@@ -80,6 +94,7 @@ for kind, output in snakemake.output.items():
 
         n.plot(
[…]

### 2023-10-04 workflow/scripts/plot-transmission-bar.py

@@ -3,6 +3,7 @@ import seaborn as sns
 import matplotlib.pyplot as plt
 import pandas as pd
 from common import (
+    get_transmission_links,
     import_network,
     mock_snakemake,
     sort_rows_by_diff,
@@ -12,7 +13,9 @@ alpha = 1
 region_alpha = 0.8
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
-    snakemake = mock_snakemake("plot_cost_bar", ext="pdf", clusters=90)
+    snakemake = mock_snakemake(
+        "plot_transmission_bar", ext="pdf", clusters=90, comparison="default"
+    )
 
 sns.set_theme(**snakemake.params["theme"])
 
@@ -20,33 +23,27 @@ df = {}
 objectives = {}
 for path in snakemake.input.networks:
     n = import_network(path)
-    capex = n.statistics.capex()
-    opex = n.statistics.opex(aggregate_time="sum")
 
-    costs = capex.add(opex, fill_value=0)
-    costs = costs.droplevel(0)[lambda x: x > 0]
-    costs = costs.groupby(costs.index).sum()
+    is_transport = get_transmission_links(n)
+    transport_carriers = [
+        *n.links.carrier[is_transport].unique(),
+        *n.lines.carrier.unique(),
+    ]
+    transport_carriers = n.carriers.nice_name[transport_carriers]
+
+    transmission = n.statistics.supply(comps=n.branch_components)
+    transmission = transmission.reindex(transport_carriers, level=1)
+    transmission = transmission.droplevel(0)
 
     key = snakemake.params.labels[n.meta["wildcards"]["run"]]
 
-    df[key] = costs
-    objectives[key] = n.objective
+    df[key] = transmission
 
-df = pd.concat(df, axis=1)
-
-groups = snakemake.config["plotting"]["technology_groups"]
-groups = {n.carriers.nice_name.get(k, k): v for k, v in groups.items()}
-grouped = df.groupby(groups).sum()
-grouped = grouped[grouped.max(axis=1) > 5e4]
-
-colors = snakemake.config["plotting"]["technology_group_colors"]
+data = pd.concat(df, axis=1)
+colors = n.carriers.set_index("nice_name").color.to_dict()
 
 norm = 1e9
-unit = "bn€/a"
-sort_by_color = (
[…]

### 2023-10-05 workflow/scripts/common.py
adjust for biogas upgrade CC use VOM for powertoX

@@ -109,8 +109,10 @@ def assert_carriers_existent(n, carriers, c):
         )
 
 
-def get_transmission_links(n):
+def get_transmission_links(n, with_eu=False):
     # only choose transmission links
+    if with_eu:
+        return n.links.bus0.map(n.buses.location) != n.links.bus1.map(n.buses.location)
     return (
         (n.links.bus0.map(n.buses.location) != n.links.bus1.map(n.buses.location))
         & ~n.links.bus0.map(n.buses.location).str.contains("EU")
@@ -165,7 +167,12 @@ def get_carrier_production(n, kind, config, which="capacity"):
 
         port = int(bus[-1])
         links = n.links.query("carrier in @carriers")
-        assert_carriers_existent(n, carriers, "Link")
+        check_carriers = set(carriers)
+        if not n.meta["sector"]["co2network"]:
+            check_carriers = check_carriers - {"co2 pipeline"}
+        if not n.meta["sector"]["gas_network"]:
+            check_carriers = check_carriers - {"gas pipeline", "gas pipeline new"}
+        assert_carriers_existent(n, check_carriers, "Link")
         groups = [links[bus].map(location), links.carrier]
         if which == "capacity":
             df = get_p_nom_opt_eff(links, bus).groupby(groups).sum()

### 2023-10-05 workflow/scripts/plot-operation-bar.py

@@ -13,7 +13,7 @@ import matplotlib.pyplot as plt
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
         "plot_operation_bar",
-        kind="hydrogen",
+        kind="carbon",
         run="half-price",
         clusters=40,
         ext="pdf",
@@ -108,5 +108,5 @@ for kind, ax in zip(kinds, axes):
     sns.despine()
     ax.set_ylabel(unit)
 
-fig.tight_layout()
+# fig.tight_layout()
 fig.savefig(snakemake.output[0], dpi=300, bbox_inches="tight")

### 2023-10-05 workflow/scripts/plot-transmission-bar.py

@@ -14,17 +14,17 @@ region_alpha = 0.8
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
-        "plot_transmission_bar", ext="pdf", clusters=90, comparison="default"
+        "plot_transmission_bar", ext="pdf", clusters=40, comparison="all"
     )
 
 sns.set_theme(**snakemake.params["theme"])
 
 df = {}
-objectives = {}
+carriers = []
 for path in snakemake.input.networks:
     n = import_network(path)
 
-    is_transport = get_transmission_links(n)
+    is_transport = get_transmission_links(n, with_eu=True)
     transport_carriers = [
         *n.links.carrier[is_transport].unique(),
         *n.lines.carrier.unique(),
@@ -38,9 +38,11 @@ for path in snakemake.input.networks:
     key = snakemake.params.labels[n.meta["wildcards"]["run"]]
 
     df[key] = transmission
+    carriers.append(n.carriers)
 
 data = pd.concat(df, axis=1)
-colors = n.carriers.set_index("nice_name").color.to_dict()
+carriers = pd.concat(carriers).drop_duplicates()
+colors = carriers.set_index("nice_name").color.to_dict()
 
 norm = 1e9
 data = sort_rows_by_diff(data).div(norm)
@@ -56,7 +58,7 @@ kwargs = {**defaults, **snakemake.params.settings.get("kwargs", {})}
 data[::-1].T.plot(ax=ax, color=colors, **kwargs)
 
 ax.axhline(0, color="k", lw=1)
-ax.set_ylabel("Transport Volume [pWh] / [Mt]")
+ax.set_ylabel("Transport Volume [pWh] / [Gt]")
 ax.grid(axis="y", alpha=0.5)
 handles, labels = ax.get_legend_handles_labels()
 ax.legend().remove()

### 2023-10-10 workflow/scripts/plot-operation-map.py
main: include gas-network by default

@@ -31,7 +31,7 @@ region_alpha = 0.8
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
         "plot_operation_map",
-        run="baseline",
+        run="with-gas/baseline",
         clusters=90,
         ext="png",
     )

### 2023-10-17 report/submissions/journal/paper/paper.tex
add transmission bar plot fix biogas CC upgrade
continued work on methodology

@@
 \graphicspath{
-    {paper-figures},
+    % {paper-figures}
+    {figures}
+    {../../../../results/}
+    {results/}
 }
@@
     The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors, particularly in emissions-intensive industries like heavy manufacturing.
-    A gap exists in understanding how comprehensive carbon management technologies, including carbon capture, transport, use, and storage, can be effectively integrated into a European energy system for climate neutrality.
+    A gap exists in understanding how comprehensive carbon management technologies, including carbon capture, transport, use, and storage, can be effectively integrated and interact with new infrastructures such as a hydrogen transportation network.
     To address this, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates carbon management technologies, hydrogen transport and storage, and renewable energy sources with a high spatial and temporal resolution.
@@
 
-To model the detailed expansion of the carbon network, we use the open-source PyPSA-Eur-Sec~\cite{PyPSAEurSecSectorCoupledOpen2023} model.
-This is a fully sector-coupled model with high spatial and temporal resolution and detailed transmission infrastructure for electricity, hydrogen, methane, and carbon networks.
-The model enables joint optimization of investment and operational decisions for generation, storage, conversion, and transmission infrastructures.
-For our study, we use a model representation with 90~geographic regions for Europe and 4 hourly time resolution to capture the dynamics of material flow between carbon sources and sinks while ensuring computational feasibility.
+The study is conducted on the basis of the open-source, capacity-expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
+The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, waste, and transport sectors.
+In our configuration, the model's time horizon spans one year with a temporal resolution of 3 hours and a spatial resolution of 90 regions.
+Each of the regions consists of a complex subsystem with technologies for supplying, converting, storing and transporting energy.
+This includes carbon neutral power supply from wind, solar, biomass and hydro.
+The regional renewable capacity expansion is limited by land-use restrictions, calculated from with atlite~\cite{hofmannAtliteLightweightPython2021}.
+
+%
+The PyPSA-Eur model integrates exogenous assumptions for energy demand on an hourly basis, the annual sums of which are shown in figure~\ref{fig:total-demand-bar}.
+Among those, the energy demand time-series for power, transport, heat, gas at rural, urban and industrial sites as well as for the agriculture sector are defined per region.
+Additionally, demands for kerosene for aviation, methanol for shipping as well of naphta for industry are aggregated and must be met by the sum of all regional supplies.
+Taking non-abatable emissions from fossil-based industrial processes into account, the system emits of 152~Mt \carbon{} per year, which can optionally be captured through corresponding investments.
+
+Carbon-neutral energy is provided by wind, solar, biomass and hydro. Solar and wind power can be expanded according to land-use constrained in alignment with
+
+, modeling the transmission and expansion of electricity (on extra-high voltage level), gas, and optionally \hydrogen{} and \carbon{}.
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/total-demand-bar.png}
+    \caption{Exogenous demand assumptions. The figure shows the total amount of yearly energy demands for each energy carrier and type of end-consumer considered in the model.}
+    \label{fig:total-demand-bar}
+\end{figure}
+
+Biomass usage in each region is bound by its biomass potential, but inter-regional biomass transport is permitted with transport costs considered.
+We restrict the electric transmission system's expansion to 20\% of its current capacity, acknowledging the challenges in inaugurating new transmission projects.
+The geographical spread of emissions from various industries is derived from~\cite{piamanzGeoreferencedIndustrialSites2018}, totaling 158~Mt \carbon{} annually.
 
@@
     \centering
-    \includegraphics[width=\linewidth]{sequestration_map.png}
+    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/sequestration_map.pdf}
     \caption{Sequestration map}
@@
 
-The PyPSA-Eur-Sec model incorporates exogenous assumptions for regional energy demand across various sectors, including power, industry, transport, buildings, and agriculture.
-This covers energy demand for shipping and aviation, as well as non-energy feedstocks for the chemical industry.
-In order to address climate neutrality targets, we establish a binding upper limit for zero emissions, consistent with the EU's 2050 target.
-The usage of biomass in each region is constrained by its biomass potential; however, biomass transport between regions is permitted, and the model takes transport costs into account.
-Additionally, we restrict the expansion of the electric transmission system to 20\% of its current capacity to account for challenges in establishing new transmission projects.
-The geographical distribution of process emissions from various industries, such as cement, chemical, paper and printing, glass, and steel, along with refineries, is obtained from~\cite{piamanzGeoreferencedIndustrialSites2018}, amounting to a total of 158~Mt \carbon{} per year.
-% TODO: Mention exogenous decisions on carbon capture and 90% rates here for industry.
-
-For carbon sequestration, we only consider offshore sites as potential sinks (refer to Figure~\ref{fig:sequestration_map}).
-This is partly because offshore storage potentials tends to be larger than onshore potential, and partly due to public safety concerns surrounding carbon storage infrastructure projects near populated areas.
-We adopt conservative assumptions when estimating storage potential, capping the total potential at 25~Mt per site and calculating annual storage availability over a 25-year period.
+For carbon sequestration, we only consider offshore sites as potential sinks, referencing Figure~\ref{fig:sequestration_map}.
+This choice favors offshore storage due to its typically larger capacity compared to onshore and concerns over public safety for infrastructure near populated areas.
+Our estimates for storage potential are conservative, limiting total potential to 25~Mt per site, calculating annual storage availability over 25 years, and capping the total sequestered \carbon{} to 200~Mt/a.
 % TODO: what defines a site?
 
-Regarding carbon utilization, our model incorporates three drop-in fuel production technologies: methanation, methanolization, and Fischer-Tropsch (FT) synthesis.
-Once processed, the fuels are available for unrestricted use and can be transported across regions.
-Synthetic methane serves as a substitute for natural gas or biogas and is utilized in combined heat and power (CHP) plants, gas boilers for residential heating, or to fulfill the gas requirements of the industry.
-Synthetic methanol is employed to decarbonize fuel demand in the marine industry.
-Lastly, FT fuels can replace fossil oil for producing naphtha in the industry and kerosene for aviation or serve as machinery oil in agriculture.
+Regarding carbon utilization, our model features three drop-in fuel production technologies: methanation, methanolization, and Fischer-Tropsch synthesis.
+Processed fuels have no usage restrictions.
+Methane is transported through the gas network, while methanol and FT-synfuel can be transferred inter-regionally without additional costs.
+Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial gas needs.
+Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production, aviation kerosene, or agricultural machinery oil.
 
-All technology cost assumptions for the year 2050 are derived from a unified open-source technology database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
+All technology cost assumptions for 2040 are sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
 
-To investigate the influence of carbon sequestration on the optimal configuration of carbon management technologies, we incrementally vary the global carbon sequestration potential from 200~Mt to 1000~Mt per year in steps of 200~Mt.
+Addressing climate targets, we define three \carbon{} target scenarios:
+\begin{itemize}
+    \item[] \textit{Net-zero:} Annual net-zero emissions, aligning with the EU's 2050 target.
+    \item[] \textit{5\% Net-negative:} Annual net-negative emissions of 5\% of 1990 emissions or 230~Mt \carbon{}.
+    \item[] \textit{10\% Net-negative:} 10\% of 1990 emissions net-negative annually, or 460~Mt \carbon{}.
+\end{itemize}
+Unless otherwise mentioned, our reference is the Net-zero scenario.
+For 5\% Net-negative and 10\% Net-negative scenarios, the carbon sequestration potential is adjusted to 400~Mt/a and 600~Mt/a, respectively.
+
+Beyond \carbon{} targets, we consider four expansion scenarios:
+\begin{itemize}
+    \item[] \textit{Baseline:} Neither \carbon{} nor \hydrogen{} networks are constructed.
+    \item[] \textit{\hydrogen{}-Net:} Only the \hydrogen{} network is developed.
+    \item[] \textit{\carbon{}-Net:} Only the \carbon{} network is developed.
+    \item[] \textit{Full:} Both \carbon{} and \hydrogen{} networks are developed.
+\end{itemize}
 
@@
     \centering
-    \includegraphics*[width=\linewidth]{cost_area.pdf}
+    % \includegraphics*[width=\linewidth]{cost_area.pdf}
     \caption[short]{Total annual system cost for the sector-coupled system with different levels of carbon sequestration potential, with (left) and without (right) \carbon{} network.
@@
     \centering
-    \includegraphics[width=\linewidth]{operation_area_carbon_co2network.pdf}
+    % \includegraphics[width=\linewidth]{operation_area_carbon_co2network.pdf}
     \caption{Balance of captured \carbon{} for the optimal operation with \carbon{} network. Technologies with "*" are explicitly augmented technologies to capture carbon emissions.}
@@
     \centering
-    \includegraphics[width=\linewidth]{operation_area_hydrogen_co2network.pdf}
+    % \includegraphics[width=\linewidth]{operation_area_hydrogen_co2network.pdf}
     \caption{Balance of \hydrogen{} production and consumption with \carbon{} network.}
@@
     \centering
-    \includegraphics[width=\linewidth]{operation_area_co2_co2network.pdf}
+    % \includegraphics[width=\linewidth]{operation_area_co2_co2network.pdf}
     \caption{Balance of atmospheric \carbon{} emissions for the optimal operation with \carbon{} network.}
@@
     \centering
-    \includegraphics[width=\linewidth]{captureshare_line.pdf}
+    % \includegraphics[width=\linewidth]{captureshare_line.pdf}
     \caption{Share of facilities with integrated carbon capture as a function of sequestration potential in a system with and without \carbon{} transport.}
@@
     \centering
-    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_1000.png}
+    % \includegraphics[width=\linewidth]{operation_map_carbon_co2network_1000.png}
     \caption{Optimal operation per sector for a sequestration of 1000~Mt/a.

### 2023-10-17 workflow/scripts/common.py

@@ -488,6 +488,7 @@ def add_carrier_nice_names(n):
         "Ac": "AC",
         "Dc": "DC",
         "Sabatier": "Methanation",
+        "Land Transport Ev": "Electric Vehicles",
         "Agriculture Machinery Oil Emissions": "Agriculture Emissions",
         "Oil Emissions": "Aviation and Petrochemical\nEmissions",
         "Allam": "Allam Cycle",

### 2023-10-17 workflow/scripts/plot-operation-bar.py

@@ -14,7 +14,7 @@ if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
         "plot_operation_bar",
         kind="co2",
-        run="net-negative-0.05/baseline",
+        run="co2-only",
         clusters=40,
         ext="pdf",
     )

### 2023-10-23 report/submissions/journal/paper/paper.tex
add more plots (diff, heatmap, capacity country)

@@
 
-\title{Designing a \carbon{} Network for a Net-Zero European Energy System}
+\title{From Net-Zero to Net-Negative: Assessing \hydrogen{} and \carbon{} Network Strategies in Europe}
 
@@
 Additionally, demands for kerosene for aviation, methanol for shipping as well of naphta for industry are aggregated and must be met by the sum of all regional supplies.
-Taking non-abatable emissions from fossil-based industrial processes into account, the system emits of 152~Mt \carbon{} per year, which can optionally be captured through corresponding investments.
+Taking non-abatable emissions from fossil-based industrial processes, aviation, shipping and agriculture into account, the system emits of 633~Mt \carbon{} per year, 153~Mt of which are industrial emissions which can optionally be captured through corresponding investments.
 

### 2023-10-23 workflow/scripts/plot-capacity-country-bar.py

@@ -1,0 +1,55 @@
+import os
+import seaborn as sns
+import pandas as pd
+import matplotlib.pyplot as plt
+import cartopy.crs as ccrs
+from pypsa.plot import add_legend_circles, add_legend_patches, add_legend_lines
+from common import (
+    import_network,
+    mock_snakemake,
+    get_carrier_storage,
+    get_carrier_transport,
+    get_carrier_production,
+)
+import geopandas as gpd
+
+
+column = "Optimal Capacity"
+alpha = 1
+region_alpha = 0.8
+
+
+if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
+    snakemake = mock_snakemake(
+        "plot_capacity_country_bar",
+        run="baseline",
+        clusters=90,
+        ext="pdf",
+    )
+
+sns.set_theme(**snakemake.params["theme"])
+
+n = import_network(snakemake.input.network)
+n.buses.country = n.buses.country.where(
+    n.buses.country.ne(""), n.buses.location.map(n.buses.country)
+)
+config = snakemake.config
+which = "capacity"
+colors = n.carriers.set_index("nice_name").color.to_dict()
+
+s = n.statistics
+
+ds = s.optimal_capacity(groupby=s.get_country_and_carrier)
+df = ds.Generator.unstack(fill_value=0).filter(regex="(?i)wind|solar|biomass")
+df = df / 1e3
+# df = ds.groupby(level=[1,2]).sum().unstack(fill_value=0).T
+
+# %%
+df.plot.bar(color=colors, subplots=True, legend=False, figsize=(5, 7), grid=True)
+plt.tight_layout()
+sns.despine()
+plt.xlabel("")
+plt.ylabel("GW")
+
+plt.savefig(snakemake.output[0], bbox_inches="tight")
+# %%

### 2023-10-23 workflow/scripts/plot-cost-diff-bar.py

@@ -13,7 +13,9 @@ alpha = 1
 region_alpha = 0.8
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
-    snakemake = mock_snakemake("plot_cost_diff_bar", ext="pdf", clusters=40)
+    snakemake = mock_snakemake(
+        "plot_cost_diff_bar", ext="pdf", clusters=90, difference="default"
+    )
 
 sns.set_theme(**snakemake.params["theme"])
 labels = snakemake.config["labels"]
@@ -21,6 +23,7 @@ labels = snakemake.config["labels"]
 
 df = {}
 objectives = {}
+carriers = []
 for path in snakemake.input.networks:
     n = import_network(path)
     capex = n.statistics.capex()
@@ -34,8 +37,10 @@ for path in snakemake.input.networks:
 
     df[key] = costs
     objectives[key] = n.objective
+    carriers.append(n.carriers)
 
 df = pd.concat(df, axis=1).fillna(0)
+carriers = pd.concat(carriers).drop_duplicates()
 
 norm = 1e9
 unit = "bn€/a"
@@ -48,13 +53,14 @@ fig, ax = plt.subplots(
 )
 
 diff = diff[diff.round(0) != 0].sort_values()
-colors = n.carriers.set_index("nice_name").color[diff.index]
+colors = carriers.set_index("nice_name").color[diff.index]
 diff.plot(kind="barh", ax=ax, color=colors, lw=0, legend=False)
 
 ax.vlines(0, -0.5, len(diff) - 0.5, color="k", lw=0.5)
 ax.set_xlabel(f"Cost benefit [{unit}]")
 ax.set_ylabel("")
 ax.grid(axis="both", alpha=0.5)
+ax.set_title(f"Cost difference ({df.columns[0]} - {df.columns[1]})")
 
 sns.despine(left=True)
 fig.savefig(snakemake.output[0], bbox_inches="tight", dpi=300)

### 2023-10-23 workflow/scripts/plot-energy-balance-area.py

@@ -22,8 +22,8 @@ if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
         "plot_energy_balance_area",
         ext="pdf",
-        clusters=40,
-        run="projected-price",
+        clusters=90,
+        run="co2-only",
     )
 
 sns.set_theme(**snakemake.params["theme"])

### 2023-10-23 workflow/scripts/plot-objective-heatmap.py

@@ -1,0 +1,63 @@
+# %%
+
+import os
+import pandas as pd
+import pypsa
+import matplotlib.pyplot as plt
+import seaborn as sns
+from common import (
+    import_network,
+    mock_snakemake,
+    sort_rows_by_diff,
+)
+
+if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
+    snakemake = mock_snakemake(
+        "plot_objectice_heatmap",
+        comparison="emission-reduction",
+        ext="pdf",
+        clusters=90,
+    )
+
+sns.set_theme(**snakemake.params["theme"])
+
+config = snakemake.config
+
+
+df = {}
+for path in snakemake.input.networks:
+    n = import_network(path)
+    balance = n.statistics.energy_balance()
+
+    # key = snakemake.params.labels[n.meta["wildcards"]["run"]]
+    key = n.meta["wildcards"]["run"]
+
+    df[key] = n.objective
+
+df = pd.Series(df)
+
+index = [l if len(l) == 2 else ["net-neutral"] + l for l in df.index.str.split("/")]
+df.index = pd.MultiIndex.from_tuples(index)
+df = df.unstack()
+
+df = df.loc[["net-neutral", "net-negative-0.05", "net-negative-0.1"]][
+    ["baseline", "co2-only", "h2-only", "full"]
+]
+df = df.div(1e9)  # bn €
+
+data = df.rename(config["labels"], axis=1).rename(config["labels"], axis=0).T
+
+fig, ax = plt.subplots(1, 1, figsize=snakemake.params.settings["figsize"])
+annot = data.applymap(lambda x: f"{x:.1f} bn €")
+sns.heatmap(
+    data,
+    ax=ax,
+    cmap="RdBu_r",
+    cbar=False,
+    annot=annot,
+    fmt="",
+    annot_kws={"fontsize": 11},
+)
+# fig.tight_layout()
+
+fig.savefig(snakemake.output[0])

### 2023-11-06 report/submissions/journal/paper/paper.tex
add sankey plot, further write methodology

@@
 
+The European Union's aim to become climate-neutral by 2050 require
+
 For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.4~Gt of carbon capture from point sources per year.
@@
 
-Carbon-neutral energy is provided by wind, solar, biomass and hydro. Solar and wind power can be expanded according to land-use constrained in alignment with
+Carbon-neutral energy is provided by wind, solar, biomass and hydro.
+Weather-dependent power potentials for solar, wind and hydro per region and time-stamp are calculated from reanalysis and satellite data sets Copernicus ERA5~\cite{hersbachERA5GlobalReanalysis2020} and Surface Radiation Data Set - Heliosat~\cite{pfeifrothSurfaceRadiationData2017}  using the open-source tool~\cite{hofmannAtliteLightweightPython2021}.
+Solar and wind power can be expanded in alignment with eligible land-use restrictions calculated on the basis of~\cite{eeaCorineLandCover2012,eeaNatura2000Data2016}.
+The biomass potential is calculated from the European Commission's Joint Research Centre's ENSPRESO database~\cite{enspreso_database} and the JRC-EU-TIMES model~\cite{instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
+Its usage in each region is bound by its biomass potential, but inter-regional biomass transport is permitted with transport costs considered.
 
-, modeling the transmission and expansion of electricity (on extra-high voltage level), gas, and optionally \hydrogen{} and \carbon{}.
+Salt caverns for \hydrogen{} storage are derived from~\cite{caglayanTechnicalPotentialSalt2020}.
+
+Industrial energy demand and excess heat potentials are calculated per node on the basis of~\cite{hotmaps_industrial_db}.
+
+% use https://pypsa-eur.readthedocs.io/en/latest/licenses.html
 
@@
 
-Biomass usage in each region is bound by its biomass potential, but inter-regional biomass transport is permitted with transport costs considered.
 We restrict the electric transmission system's expansion to 20\% of its current capacity, acknowledging the challenges in inaugurating new transmission projects.
@@
 \begin{itemize}
-    \item[] \textit{Net-zero:} Annual net-zero emissions, aligning with the EU's 2050 target.
-    \item[] \textit{5\% Net-negative:} Annual net-negative emissions of 5\% of 1990 emissions or 230~Mt \carbon{}.
-    \item[] \textit{10\% Net-negative:} 10\% of 1990 emissions net-negative annually, or 460~Mt \carbon{}.
+    \item[] Net-Zero: aligning with the EU's 2050 emission targets. Also denoted as NZ.
+    \item[] Net-Negative~5\%: relative to 1990 emissions, equaling 230~Mt \carbon{}. Also denoted as NN~5\%.
+    \item[] Net-Negative~10\%: relative to 1990 emissions, equaling 460~Mt \carbon. Also denoted as NN~10\%.
 \end{itemize}
@@
     \centering
-    % \includegraphics*[width=\linewidth]{cost_area.pdf}
-    \caption[short]{Total annual system cost for the sector-coupled system with different levels of carbon sequestration potential, with (left) and without (right) \carbon{} network.
+    \includegraphics*[width=\linewidth]{comparison/default/figures/90_nodes/cost_bar.png}
+    \caption[short]{Total annual system cost for the European energy system for the different scenarios subdivided into groups of technologies.
+While the `Baseline' scenario has neither a \hydrogen{} or \carbon{} network, `Full' is allowed to expand both.
 "Gas Infrastructure" combines gas facilities for power and heat production, "H$_2$ Infrastructure" combines H$_2$ production, transport and re-electrification.
-Regardless of the implementation of a \carbon{} network, the system cost decrease as sequestration increases.
-Due to an increased flexibility from fossil carriers with subsequent sequestration, the need for FT synthesis and \hydrogen{} electrolysis and the corresponding renewable power supply is reduced.
-        % TODO: mention details on the grouped technologies
+The
     }
@@
 At the same time, integrated carbon capture from Gas Combined Heat and Power (CHP) plants, Allam Cycle plants, Steam Methane Reforming (SMR) and DAC increases.
-Other technologies are constant across varying levels of sequestration due to different reasons: biomass is bound by its maximum available potential (as reported in~\cite{europeancommissionjointresearchcentreENSPRESOBIOMASS2019}); the industry processes with their associated emissions are set to fixed values; the methanol production is bound by a lower limit representing the demand from the shipping sector.
+Other technologies are constant across varying levels of sequestration due to different reasons: biomass is bound by its maximum available potential (as reported in~\cite{enspreso_database}); the industry processes with their associated emissions are set to fixed values; the methanol production is bound by a lower limit representing the demand from the shipping sector.
 

### 2023-11-06 workflow/scripts/plot-capacity-country-bar.py

@@ -8,11 +8,7 @@ from pypsa.plot import add_legend_circles, add_legend_patches, add_legend_lines
 from common import (
     import_network,
     mock_snakemake,
-    get_carrier_storage,
-    get_carrier_transport,
-    get_carrier_production,
 )
-import geopandas as gpd
 
 
 column = "Optimal Capacity"

### 2023-11-06 workflow/scripts/plot-energy-balance-bar.py

@@ -17,7 +17,12 @@ from common import (
 )
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
-    snakemake = mock_snakemake("plot_energy_balance_bar", ext="pdf", clusters=40)
+    snakemake = mock_snakemake(
+        "plot_energy_balance_bar",
+        ext="pdf",
+        clusters=40,
+        comparison="net-negative-0.05",
+    )
 
 sns.set_theme(**snakemake.params["theme"])
 
@@ -48,6 +53,8 @@ for kind, output in snakemake.output.items():
     ds = ds[ds.round(2) != 0].droplevel(0).dropna(how="all").fillna(0)
     ds = ds.groupby(level=0).sum()
     ds = ds.sort_values(ds.columns[0], ascending=False)
+    if kind == "co2":
+        ds.drop("CO$_2$", inplace=True, errors="ignore")
     # ds = sort_rows_by_diff(ds)
 
     fig, ax = plt.subplots(1, 1, figsize=snakemake.params.settings["figsize"])

### 2023-11-06 workflow/scripts/plot-objective-heatmap.py

@@ -43,12 +43,12 @@ df = df.unstack()
 df = df.loc[["net-neutral", "net-negative-0.05", "net-negative-0.1"]][
     ["baseline", "co2-only", "h2-only", "full"]
 ]
-df = df.div(1e9)  # bn €
+df = df.div(1e9).round(0).astype(int)  # bn €
 
 data = df.rename(config["labels"], axis=1).rename(config["labels"], axis=0).T
 
 fig, ax = plt.subplots(1, 1, figsize=snakemake.params.settings["figsize"])
-annot = data.applymap(lambda x: f"{x:.1f} bn €")
+annot = data.applymap(lambda x: f"{x} bn €")
 sns.heatmap(
     data,
     ax=ax,

### 2023-11-13 report/submissions/journal/paper/paper.tex
further adjustments for plots

@@
 \usepackage{fancyhdr,lipsum}
+\usepackage{subcaption}
 
@@
 
-The European Union's aim to become climate-neutral by 2050 require
+The European Union's goal to achieve climate neutrality by 2050 involves the implementation of carbon management strategies as a crucial component.
+This objective is integrated into a wide range of programs, funding schemes and initiatives~\cite{eu2023netzero,europeangreendeal,europeaninnovationfund}.
+
+%~\cite{lockwoodEuropeanStrategyCarbon} - Clean Air Task Force
+%~\cite{righettiSitingCarbonDioxide2017} - co2 pipeline are a mature technology, references to US and Canada co2 pipelines used for enhanced oil recovery
+%~\cite{friedmannNETZEROGEOSPHERICRETURN} - high learning rate for co2 pipeline costs
+%~\cite{NorthernLightsWhat} - Northern Lights project in Norway, 1.5 Mt/a co2 transport capacity, 1.5 Mt/a co2 storage capacity, 2024 start of operation, targeted scale of 5 Mtpa CO2 Seq. by 2030.
 
@@
 Important insights to formulate such a strategy can be gained from energy system models that provide a holistic view of the energy system and its technological interactions.
-To this end, a number of optimization tools have been developed in recent years to model synergies between various carbon management technologies~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
+To this end, a number of optimization tools have been developed in recent years to model synergies between various carbon management technologies~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
 In contrast to Integrated Assessment Models, these tools account for the spatial distribution of carbon sources and sinks.
+% point out that~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a co2 network, but only considers the power sector without co-optimizing renewable deployment.
+
 % A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
@@
 In this paper, we present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies.
-The study is conducted using the PyPSA-Eur-Sec energy system model and encompasses all relevant energy sectors.
-We enable the system to optimize the design of renewable energy sources and storage technologies, as well as the transmission of electricity, hydrogen, methane, and \carbon{}.
+The study is conducted using the PyPSA-Eur energy system model and encompasses all relevant energy sectors.
+We enable the system to optimize the design of renewable energy sources and storage technologies, as well as the transmission of electricity, \hydrogen{}, methane, and \carbon{}.
 We show that the dynamics of carbon capture, transport, use, and storage in different regions are highly dependent on the availability of renewable energy, \carbon{} pipelines and sequestration sinks.
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/sequestration_map.pdf}
-    \caption{Sequestration map}
-    \label{fig:sequestration_map}
-\end{figure}
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=\linewidth]{baseline/figures/90_nodes/sequestration_map.pdf}
+%     \caption{Sequestration map}
+%     \label{fig:sequestration_map}
+% \end{figure}
 
-For carbon sequestration, we only consider offshore sites as potential sinks, referencing Figure~\ref{fig:sequestration_map}.
+For carbon sequestration, we only consider offshore sites as potential sinks.
 This choice favors offshore storage due to its typically larger capacity compared to onshore and concerns over public safety for infrastructure near populated areas.
@@
 \begin{itemize}
-    \item[] Net-Zero: aligning with the EU's 2050 emission targets. Also denoted as NZ.
-    \item[] Net-Negative~5\%: relative to 1990 emissions, equaling 230~Mt \carbon{}. Also denoted as NN~5\%.
-    \item[] Net-Negative~10\%: relative to 1990 emissions, equaling 460~Mt \carbon. Also denoted as NN~10\%.
+    \item[] \textit{Net-Zero}: aligning with the EU's 2050 emission targets. Also denoted as NZ.
+    \item[] \textit{Net-Negative~5\%}: relative to 1990 emissions, equaling 230~Mt \carbon{}. Also denoted as NN~5\%.
+    \item[] \textit{Net-Negative~10\%}: relative to 1990 emissions, equaling 460~Mt \carbon. Also denoted as NN~10\%.
 \end{itemize}
@@
     \item[] \textit{\carbon{}-Net:} Only the \carbon{} network is developed.
-    \item[] \textit{Full:} Both \carbon{} and \hydrogen{} networks are developed.
+    \item[] \textit{Hybrid:} Both \carbon{} and \hydrogen{} networks are developed.
 \end{itemize}
@@
 
-\begin{figure*}[ht!]
+% hight level trends
+- cost reduction with carbon network is significant (2.5\%)
+
+\begin{figure}
     \centering
-    \includegraphics*[width=\linewidth]{comparison/default/figures/90_nodes/cost_bar.png}
+    \includegraphics*[width=0.8\linewidth]{comparison/emission-reduction/figures/90_nodes/objective_heatmap.png}
+    \caption{Total annual system cost for the different scenarios and \carbon{} targets.
+While the `Baseline' scenario has neither a \hydrogen{} or \carbon{} network, `Hybrid' is allowed to expand both.}
+    \label{fig:objective_heatmap}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/cost_bar.png}
     \caption[short]{Total annual system cost for the European energy system for the different scenarios subdivided into groups of technologies.
-While the `Baseline' scenario has neither a \hydrogen{} or \carbon{} network, `Full' is allowed to expand both.
+While the `Baseline' scenario has neither a \hydrogen{} or \carbon{} network, `Hybrid' is allowed to expand both.
 "Gas Infrastructure" combines gas facilities for power and heat production, "H$_2$ Infrastructure" combines H$_2$ production, transport and re-electrification.
@@
     \label{fig:cost_area}
+\end{figure}
+%
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/captureshare_line.png}
+    \caption{Share of facilities with integrated carbon capture as a function of emission targets for system with \hydrogen{} and \carbon{} network.}
+    \label{fig:captureshare_line}
+\end{figure}
+
+\begin{figure*}[h]
+    \centering
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{full/figures/90_nodes/capacity_map_carbon.png}
+        \caption{\carbon{} Sector.}
+        \label{fig:capacity_map_carbon_co2}
+    \end{subfigure}%
+    \hfill
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{full/figures/90_nodes/capacity_map_hydrogen.png}
+        \caption{\hydrogen{} Sector.}
+        \label{fig:capacity_map_hydrogen_co2}
+    \end{subfigure}
+    \caption{Optimal production and transport capacities of the carbon and hydrogen sector in a net-zero energy system in Europe with both \carbon{} and \hydrogen{} network expansion (Hybrid).
+        % Carbon network looks the same as in~\cite{morbeeOptimisedDeploymentEuropean2012}: two backbones, one in the nothern Europe other in south east.
+    }
+    \label{fig:capacity_maps}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \begin{subfigure}{\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_carbon.png}
+        \caption{\carbon{} Sector.}
+        \label{fig:operation_map_carbon}
+    \end{subfigure}
+    \begin{subfigure}{\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_hydrogen.png}
+        \caption{\hydrogen{} Sector.}
+        \label{fig:operation_map_hydrogen}
+    \end{subfigure}
+    \caption{Optimal operation per sector for a net-zero energy system in Europe with both \carbon{} and \hydrogen{} network expansion (Hybrid).}
+    \label{fig:operation_map_carbon_co2}
 \end{figure*}
@@
     \centering
-    % \includegraphics[width=\linewidth]{operation_area_carbon_co2network.pdf}
-    \caption{Balance of captured \carbon{} for the optimal operation with \carbon{} network. Technologies with "*" are explicitly augmented technologies to capture carbon emissions.}
-    \label{fig:operation_area_carbon_co2network}
+    \includegraphics[width=\linewidth]{comparison/emission-reduction-full/figures/90_nodes/transmission_cost_bar.png}
+    \caption{Annual transmission system cost as a function of the net carbon removal scenarios considered in the study.}
+    \label{fig:transmission_cost_bar}
 \end{figure}

### 2023-11-13 workflow/scripts/plot-capacity-map.py

@@ -7,8 +7,10 @@ Created on Thu Nov 24 10:14:48 2022
 """
 import os
 import seaborn as sns
+import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
+import matplotlib.colors as mcolors
 import cartopy.crs as ccrs
 from pypsa.plot import add_legend_circles, add_legend_patches, add_legend_lines
 from common import (
@@ -24,14 +26,15 @@ import geopandas as gpd
 column = "Optimal Capacity"
 alpha = 1
 region_alpha = 0.8
+aligned_carriers = ["carbon", "hydrogen"]
 
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
         "plot_capacity_map",
-        run="projected-price",
-        clusters=40,
-        ext="pdf",
+        run="full",
+        clusters=90,
+        ext="png",
     )
 
 sns.set_theme(**snakemake.params["theme"])
@@ -45,13 +48,18 @@ which = "capacity"
 for kind, output in snakemake.output.items():
 
     fig, ax = plt.subplots(
-        figsize=(7, 6),
+        figsize=snakemake.params["settings"]["figsize"],
         subplot_kw={"projection": ccrs.EqualEarth()},
+        layout="constrained" if kind in aligned_carriers else None,
     )
 
-    unit = "GW" if kind not in ["carbon", "co2"] else "kt/h"
-
     specs = config["plotting"]["capacity_map"][kind]
+    unit = specs["unit"]
+    unit_conversion = float(specs["unit_conversion"])
+    region_unit = specs["region_unit"]
+    region_unit_conversion = float(specs["region_unit_conversion"])
+    region_cmap = specs["region_cmap"]
+
     bus_scale = float(specs["bus_scale"])
     branch_scale = float(specs["branch_scale"])
 
@@ -64,12 +72,14 @@ for kind, output in snakemake.output.items():
     }
 
     n.plot(
-        bus_sizes=bus_sizes,
+        bus_sizes=bus_sizes * unit_conversion,
         bus_alpha=alpha,
         line_widths=branch_widths["Line"].reindex(n.lines.index, fill_value=0)
-        * branch_scale,
+        * branch_scale
+        * unit_conversion,
         link_widths=branch_widths["Link"].reindex(n.links.index, fill_value=0)
-        * branch_scale,
+        * branch_scale
+        * unit_conversion,
         line_colors=branch_colors["Line"],
         link_colors=branch_colors["Link"],
[…]

### 2023-11-13 workflow/scripts/plot-captureshare-line.py

@@ -12,7 +12,8 @@ import matplotlib.pyplot as plt
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
         "plot_captureshare_line",
-        clusters=40,
+        comparison="default",
+        clusters=90,
         ext="png",
     )
 
@@ -41,6 +42,8 @@ for path in snakemake.input.networks:
     df[key] = caps
 
 df = pd.concat(df, axis=1)
+df = df[df.sum().sort_values().index]
+df = df.reindex(df.iloc[:, 0].sort_values().index, axis=0)
 
 nice_name = n.carriers.nice_name
 colors = n.carriers.color.dropna().rename(nice_name)
@@ -48,7 +51,6 @@ colors = n.carriers.color.dropna().rename(nice_name)
 fig, ax = plt.subplots(
     figsize=snakemake.params.settings["figsize"],
     # layout="constrained",
-    sharex=True,
 )
 
 df.mul(100).T.plot(
@@ -57,10 +59,19 @@ df.mul(100).T.plot(
     color=colors.to_dict(),
     alpha=0.8,
     rot=90,
+    marker="o",
+    ls="-",
 )
 ax.set_ylabel(f"CC share [%]")
 ax.grid(axis="y", alpha=0.5)
-ax.legend(loc="center left", bbox_to_anchor=(1, 0.5), frameon=False)
+handles, labels = ax.get_legend_handles_labels()
+ax.legend(
+    handles[::-1],
+    labels[::-1],
+    loc="center left",
+    bbox_to_anchor=(1, 0.5),
+    frameon=False,
+)
 
 sns.despine()
 

### 2023-11-13 workflow/scripts/plot-cost-diff-map.py

@@ -1,9 +1,10 @@
 import os
-from pathlib import Path
 import cartopy.crs as ccrs
+import geopandas as gpd
 import seaborn as sns
 import matplotlib.pyplot as plt
 from pypsa.plot import add_legend_circles, add_legend_patches, add_legend_lines
+from pypsa.statistics import get_transmission_branches, get_transmission_carriers
 import pandas as pd
 from common import (
     import_network,
@@ -15,7 +16,10 @@ region_alpha = 0.8
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
-        "plot_cost_diff_map", ext="pdf", clusters=90, difference="default"
+        "plot_cost_diff_map",
+        ext="png",
+        clusters=90,
+        difference="emission-reduction-0.1-full",
     )
 
 sns.set_theme(**snakemake.params["theme"])
@@ -23,44 +27,56 @@ config = snakemake.config
 labels = config["labels"]
 specs = config["plotting"]["cost_diff_map"]
 
-df = {}
+networks = [import_network(path) for path in snakemake.input.networks]
+regions = gpd.read_file(snakemake.input.regions).set_index("name")
+df_bus = {}
+df_branch = {}
 carriers = []
-for path in snakemake.input.networks:
-    n = import_network(path)
+for n in networks:
     s = n.statistics
 
     # TODO: ensure this weird adjustment is not necessary anymore
     is_dac = n.links.carrier == "DAC"
-    n.links.loc[is_dac, "bus0"] = n.links.loc[is_dac, "bus1"]
+    n.links.loc[is_dac, ["bus0", "bus1"]] = n.links.loc[is_dac, ["bus1", "bus0"]].values
 
-    capex = n.statistics.capex(groupby=s.get_bus_and_carrier)
-    opex = n.statistics.opex(aggregate_time="sum", groupby=s.get_bus_and_carrier)
-
+    capex = s.capex(groupby=s.get_bus_and_carrier)
+    opex = s.opex(aggregate_time="sum", groupby=s.get_bus_and_carrier)
     costs = capex.add(opex, fill_value=0)
 
     key = snakemake.params.labels[n.meta["wildcards"]["run"]]
 
-    df[key] = costs
+    df_bus[key] = costs
+
+    branches = get_transmission_branches(n)
[…]

### 2023-11-13 workflow/scripts/plot-objective-heatmap.py

@@ -27,12 +27,11 @@ config = snakemake.config
 df = {}
 for path in snakemake.input.networks:
     n = import_network(path)
-    balance = n.statistics.energy_balance()
 
     # key = snakemake.params.labels[n.meta["wildcards"]["run"]]
     key = n.meta["wildcards"]["run"]
 
-    df[key] = n.objective
+    df[key] = n.statistics.capex().sum() + n.statistics.opex().sum()
 
 df = pd.Series(df)
 
@@ -56,8 +55,7 @@ sns.heatmap(
     cbar=False,
     annot=annot,
     fmt="",
-    annot_kws={"fontsize": 11},
+    annot_kws={"fontsize": 9},
 )
 # fig.tight_layout()
-
 fig.savefig(snakemake.output[0])

### 2023-11-13 workflow/scripts/plot-operation-map.py

@@ -31,7 +31,7 @@ region_alpha = 0.8
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
         "plot_operation_map",
-        run="with-gas/baseline",
+        run="full",
         clusters=90,
         ext="png",
     )
@@ -50,9 +50,14 @@ transport_carriers = [
 ]
 transport_carriers = n.carriers.nice_name[transport_carriers]
 
-balance = n.statistics.energy_balance(aggregate_bus=False)
-balance = balance.rename(n.buses.location, level=3)
-dispatch = balance.drop(transport_carriers, level=1)
+# TODO: ensure this weird adjustment is not necessary anymore
+dac = n.links.index[n.links.carrier == "DAC"]
+n.links.loc[dac, ["bus0", "bus2"]] = n.links.loc[dac, ["bus2", "bus0"]].values
+n.links.loc[dac, ["", "bus2"]] = n.links.loc[dac, ["bus2", "bus0"]].values
+n.links_t.p0[dac], n.links_t.p2[dac] = n.links_t.p2[dac], n.links_t.p0[dac].values
+
+s = n.statistics
+
 
 for kind, output in snakemake.output.items():
 
@@ -65,8 +70,12 @@ for kind, output in snakemake.output.items():
     )
 
     carriers = config["constants"]["carrier_to_buses"].get(kind, [kind])
-    carriers = list(set(carriers) & set(dispatch.index.get_level_values(2)))
-    df = dispatch.loc[:, :, carriers].groupby(level=[3, 1]).sum()
+
+    grouper = s.groupers.get_bus_and_carrier
+    df = s.dispatch(bus_carrier=carriers, groupby=grouper)
+    df = df.drop(transport_carriers, level=2, errors="ignore")
+    df = df.rename(lambda x: x.replace(" CC", ""), level=2)
+    df = df.groupby(level=[1, 2]).sum().rename(n.buses.location, level=0)
 
     tags = ["production", "consumption"]
 
@@ -107,6 +116,15 @@ for kind, output in snakemake.output.items():
             margin=0.2,
             color_geomap={"border": "darkgrey", "coastline": "darkgrey"},
             geomap="10m",
+        )
+        regions.plot(
+            ax=ax,
+            facecolor="whitesmoke",
+            edgecolor="darkgrey",
+            linewidth=0,
+            alpha=region_alpha,
+            transform=ccrs.PlateCarree(),
+            aspect="equal",
         )
 
         gen_carriers = (

### 2023-11-16 report/submissions/journal/paper/paper.tex
update methodology and operation map plots

@@
 The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, waste, and transport sectors.
+
 In our configuration, the model's time horizon spans one year with a temporal resolution of 3 hours and a spatial resolution of 90 regions.
 Each of the regions consists of a complex subsystem with technologies for supplying, converting, storing and transporting energy.
-This includes carbon neutral power supply from wind, solar, biomass and hydro.
-The regional renewable capacity expansion is limited by land-use restrictions, calculated from with atlite~\cite{hofmannAtliteLightweightPython2021}.
-
-%
-The PyPSA-Eur model integrates exogenous assumptions for energy demand on an hourly basis, the annual sums of which are shown in figure~\ref{fig:total-demand-bar}.
-Among those, the energy demand time-series for power, transport, heat, gas at rural, urban and industrial sites as well as for the agriculture sector are defined per region.
-Additionally, demands for kerosene for aviation, methanol for shipping as well of naphta for industry are aggregated and must be met by the sum of all regional supplies.
-Taking non-abatable emissions from fossil-based industrial processes, aviation, shipping and agriculture into account, the system emits of 633~Mt \carbon{} per year, 153~Mt of which are industrial emissions which can optionally be captured through corresponding investments.
-
-Carbon-neutral energy is provided by wind, solar, biomass and hydro.
-Weather-dependent power potentials for solar, wind and hydro per region and time-stamp are calculated from reanalysis and satellite data sets Copernicus ERA5~\cite{hersbachERA5GlobalReanalysis2020} and Surface Radiation Data Set - Heliosat~\cite{pfeifrothSurfaceRadiationData2017}  using the open-source tool~\cite{hofmannAtliteLightweightPython2021}.
-Solar and wind power can be expanded in alignment with eligible land-use restrictions calculated on the basis of~\cite{eeaCorineLandCover2012,eeaNatura2000Data2016}.
-The biomass potential is calculated from the European Commission's Joint Research Centre's ENSPRESO database~\cite{enspreso_database} and the JRC-EU-TIMES model~\cite{instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
-Its usage in each region is bound by its biomass potential, but inter-regional biomass transport is permitted with transport costs considered.
-
-Salt caverns for \hydrogen{} storage are derived from~\cite{caglayanTechnicalPotentialSalt2020}.
-
+Exogenous assumptions on energy demand and non-abatable emissions are taken from various sources~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023}.
+The energy demand for electricity, transport, biomass, heat and gas are defined per region and time-step with heat demand being regionally subdivided into shares of urban, rural and industrial sites.
+Demands for aviation kerosene, methanol for shipping and naphtha for industry are aggregated in the system scope and kept constant throughout all time-steps.
+We show the sum of all energy demands in \ref{fig:total-demand-bar}.
+The system emits of 633~Mt \carbon{} per year through non-abatable emissions from industry, aviation, shipping and agriculture into account, 153~Mt of which are fossil-based industrial emissions.
 Industrial energy demand and excess heat potentials are calculated per node on the basis of~\cite{hotmaps_industrial_db}.
-
 % use https://pypsa-eur.readthedocs.io/en/latest/licenses.html
@@
 
+Endogenous assumptions include the expansion of renewable energy sources, storage technologies, and transmission capacities.
+The model considers the various energy carriers like electricity, hydrogen, methan, methanol, liquid hydrocarbons and biomass, together with conversions technologies to convert these into each other.
+Carbon-neutral electricity is provided by wind, solar, biomass, hydro and nuclear power plants.
+Weather-dependent power potentials for solar, wind and hydro are calculated from reanalysis and satellite data sets~\cite{hersbachERA5GlobalReanalysis2020,pfeifrothSurfaceRadiationData2017}  per region and time-stamp, using the open-source tool Atlite~\cite{hofmannAtliteLightweightPython2021}.
+Solar and wind power can be expanded in alignment with eligible land-use restrictions calculated on the basis of~\cite{eeaCorineLandCover2012,eeaNatura2000Data2016}.
 We restrict the electric transmission system's expansion to 20\% of its current capacity, acknowledging the challenges in inaugurating new transmission projects.
-The geographical spread of emissions from various industries is derived from~\cite{piamanzGeoreferencedIndustrialSites2018}, totaling 158~Mt \carbon{} annually.
+The regional use of biomass is restricted by its feasible potential derived from~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
+Inter-regional biomass transport is permitted with transport costs considered.
+Hydrogen can be gained from electrolysis and steam methane reforming (SMR). Salt caverns for \hydrogen{} storage are derived from~\cite{caglayanTechnicalPotentialSalt2020}.
+
+\dots
+
+% carbon capture technologies
+Carbon capture technologies are represented by a set of augmented technologies with integrated carbon capture.
+These include Allam Cycle, Gas CHP, SMR, DAC, biomass, biogas and capture of process emissions from industry.
+
+Regarding carbon utilization, our model features three drop-in fuel production technologies: methanation, methanolization, and Fischer-Tropsch synthesis.
+Processed fuels have no usage restrictions.
+Methane is transported through the gas network, while methanol and FT-synfuel can be transferred inter-regionally without additional costs.
+Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial gas needs.
+Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production, aviation kerosene, or agricultural machinery oil.
+For carbon sequestration, we only consider offshore sites as potential sinks.
+This choice favors offshore storage due to its typically larger capacity compared to onshore and concerns over public safety for infrastructure near populated areas.
+Our estimates for storage potential are conservative, limiting total potential to 25~Mt per site, calculating annual storage availability over 25 years, and capping the total sequestered \carbon{} to 200~Mt/a.
+
+All technology cost assumptions for 2040 are sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
 
@@
 % \end{figure}
-
-For carbon sequestration, we only consider offshore sites as potential sinks.
-This choice favors offshore storage due to its typically larger capacity compared to onshore and concerns over public safety for infrastructure near populated areas.
-Our estimates for storage potential are conservative, limiting total potential to 25~Mt per site, calculating annual storage availability over 25 years, and capping the total sequestered \carbon{} to 200~Mt/a.
-% TODO: what defines a site?
-
-Regarding carbon utilization, our model features three drop-in fuel production technologies: methanation, methanolization, and Fischer-Tropsch synthesis.
-Processed fuels have no usage restrictions.
-Methane is transported through the gas network, while methanol and FT-synfuel can be transferred inter-regionally without additional costs.
-Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial gas needs.
-Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production, aviation kerosene, or agricultural machinery oil.
-
-All technology cost assumptions for 2040 are sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics*[width=0.8\linewidth]{comparison/emission-reduction/figures/90_nodes/objective_heatmap.png}
-    \caption{Total annual system cost for the different scenarios and \carbon{} targets.
-While the `Baseline' scenario has neither a \hydrogen{} or \carbon{} network, `Hybrid' is allowed to expand both.}
-    \label{fig:objective_heatmap}
-\end{figure}
-
 \begin{figure}[ht!]
@@
 \end{figure}
+
 %
@@
     \caption{Share of facilities with integrated carbon capture as a function of emission targets for system with \hydrogen{} and \carbon{} network.}
+    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
     \label{fig:captureshare_line}
 \end{figure}
+
+\begin{figure*}
+    \centering
+    \begin{subfigure}{\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_carbon.png}
+        \caption{\carbon{} Sector.}
+        \label{fig:operation_map_carbon}
+    \end{subfigure}
+    \begin{subfigure}{\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_hydrogen.png}
+        \caption{\hydrogen{} Sector.}
+        \label{fig:operation_map_hydrogen}
+    \end{subfigure}
+    \caption{Optimal operation per sector for a net-zero energy system in Europe with both \carbon{} and \hydrogen{} network expansion (Hybrid).}
+    \label{fig:operation_map_carbon_co2}
+\end{figure*}
 
@@
 
-\begin{figure*}
+\begin{figure}
     \centering
-    \begin{subfigure}{\linewidth}
+    \includegraphics*[width=0.8\linewidth]{comparison/emission-reduction/figures/90_nodes/objective_heatmap.png}
+    \caption{Total annual system cost for the different scenarios and \carbon{} targets.
+While the `Baseline' scenario has neither a \hydrogen{} or \carbon{} network, `Hybrid' is allowed to expand both.}
+    \label{fig:objective_heatmap}
+\end{figure}
+
+\begin{figure}[h]
     \centering
-        \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_carbon.png}
-        \caption{\carbon{} Sector.}
-        \label{fig:operation_map_carbon}
-    \end{subfigure}
-    \begin{subfigure}{\linewidth}
+    \includegraphics[width=\linewidth]{comparison/emission-reduction-full/figures/90_nodes/transmission_cost_bar.png}
+    \caption{Annual transmission system cost as a function of the net carbon removal scenarios considered in the study.}
+    \label{fig:transmission_cost_bar}
+\end{figure}
+
+\begin{figure}
     \centering
-        \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_hydrogen.png}
-        \caption{\hydrogen{} Sector.}
-        \label{fig:operation_map_hydrogen}
-    \end{subfigure}
-    \caption{Optimal operation per sector for a net-zero energy system in Europe with both \carbon{} and \hydrogen{} network expansion (Hybrid).}
-    \label{fig:operation_map_carbon_co2}
-\end{figure*}
+    \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
+    \caption{Sankey diagram of the optimal operation for a net-negative 10\% scenario.}
+    \label{fig:sankey_diagramm}
+\end{figure}
 
@@
 %
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{comparison/emission-reduction-full/figures/90_nodes/transmission_cost_bar.png}
-    \caption{Annual transmission system cost as a function of the net carbon removal scenarios considered in the study.}
-    \label{fig:transmission_cost_bar}
-\end{figure}
 
@@
 
-\begin{figure}[h]
-    \centering
-    % \includegraphics[width=\linewidth]{operation_area_hydrogen_co2network.pdf}
-    \caption{Balance of \hydrogen{} production and consumption with \carbon{} network.}
-    % TODO: make colors of gas and gas* more distinct
-    \label{fig:operation_area_hydrogen_co2network}
-\end{figure}
-%
-\begin{figure}[h]
-    \centering
-    % \includegraphics[width=\linewidth]{operation_area_co2_co2network.pdf}
-    \caption{Balance of atmospheric \carbon{} emissions for the optimal operation with \carbon{} network.}
-    \label{fig:operation_area_co2_co2network}
-\end{figure}
-%
 % TODO: kerosine emissions from fossil based kerosine are not taken into account by the system.
@@
 This is primarily due to the fact that as sequestration increases, the system relies more heavily on DAC and SMR with CC that can be flexibly placed near the sequestration sites.
-
-\begin{figure}[h]
-    \centering
-    % \includegraphics[width=\linewidth]{captureshare_line.pdf}
-    \caption{Share of facilities with integrated carbon capture as a function of sequestration potential in a system with and without \carbon{} transport.}
-    \label{fig:captureshare_line}
-\end{figure}
 
@@
 On the other hand, with low sequestration rates the use of CCU in a system is distributed more evenly across the system.
-
-\begin{figure*}[ht!]
-    \centering
-    % \includegraphics[width=\linewidth]{operation_map_carbon_co2network_1000.png}
-    \caption{Optimal operation per sector for a sequestration of 1000~Mt/a.
-The left side shows the combined carbon supply from carbon capture facilities and the carbon network flow, while the right side illustrates regional carbon usage together with the flow.}
-    % TODO: make colors more distinct
-    \label{fig:operation_map_noco2network_1000}
-\end{figure*}
 

### 2023-11-16 workflow/scripts/leftover/co2scenarios.py

@@ -219,7 +219,6 @@ def load_main(scenarios=None, rename=True):
         df = df.groupby(df.index.map(rename_techs_tyndp)).sum()
 
     to_drop = df.index[df.max(axis=1).fillna(0.0) < 1.2]
-    print(to_drop)
     df.drop(to_drop, inplace=True)
 
     order = preferred_order.intersection(df.index).append(

### 2023-11-20 report/submissions/journal/paper/paper.tex
further work

@@
     The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors, particularly in emissions-intensive industries like heavy manufacturing.
-    A gap exists in understanding how comprehensive carbon management technologies, including carbon capture, transport, use, and storage, can be effectively integrated and interact with new infrastructures such as a hydrogen transportation network.
-    To address this, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates carbon management technologies, hydrogen transport and storage, and renewable energy sources with a high spatial and temporal resolution.
-    Our findings reveal that implementing carbon management technologies can offer systemic flexibility to the European energy system.
-Specifically, a carbon transport network facilitates point source carbon capture, thereby reducing the necessity for direct air capture installation.
-The cost reduction of  average annual savings of €xx billion.
+While management of both \hydrogen{} and \carbon{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructures effectively interact with comprehensive carbon management technologies, including carbon capture, transport, use, and storage.
+In particular, the role of a carbon transport network as a complementing or substituting infrastructure to hydrogen transport remains unclear.
+To address this gap, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates carbon management technologies, hydrogen transport and storage, and renewable energy sources with a high spatial and temporal resolution considering all energy intensive sectors.
+    Our findings reveal that in a either-or scenario, the hydrogen network is superior to the carbon network as it enables the transportation cheap hydrogen inland to serve industrial processes and carbon capture and utilization at point sources.
+A carbon network facilitates point-source carbon capture and reduces the need for direct air capture, but leaves the scarcity of hydrogen in energy-intensive regions such as industrial clusters.
+    However, in a hybrid scenario, the carbon network is a cost-effective complement to the hydrogen network facilitating carbon capture at point sources and transport to sequestration sites.
+    We show that theses findings holds true against the backdrop of a net-zero energy system, as well as scenarios where the net-negative emissions are 5\% and 10\% lower than the 1990 baseline levels.
     Overall, our work demonstrates the cost-effectiveness of a multi-grid system that includes both hydrogen and carbon transport networks and power grid expansion to achieve climate neutrality in Europe.
-
-    % Achieving a carbon-neutral European economy requires internationally coordinated efforts across all energy sectors, especially those with hard-to-abate emissions like heavy industry.
-Against this backdrop, It is critical to assess the need for comprehensive carbon management involving technologies such as carbon capture, transport, use, and storage to develop an effective carbon strategy for climate neutrality.
-
-    % This paper provides the first cost-optimal design for a European energy system that fully incorporates carbon management technologies including a carbon transport network, and considers all emissions-intensive sectors.
-It optimizes carbon technologies, hydrogen transport and storage, and the design of renewable energy sources.
-
-    % We show that carbon management technologies aiming at net-zero emissions provide systemic flexibility to the European energy system.
-A carbon network is cost-effective leading to average cost savings of €xx billion per year.
-The \carbon{} network enables viable point source carbon capture, reducing the need for direct air capture facilities.
     The paper underpins the need for a complementing hydrogen network and \carbon{} network to achieve climate neutrality.
-
 \end{abstract}
@@
 
-In net-zero economies, carbon capture is regarded as a crucial technology for offsetting emissions from hard-to-abate sectors and unavoidable emissions from land use and agriculture.
-Capturing carbon at point sources and from the atmosphere via Direct Air Capture (DAC) presents new opportunities for producing drop-in fuels such as methane, methanol, or carbon-neutral liquid hydrocarbons.
-This process is known as Carbon Capture and Utilization (CCU).
-Additionally, storing carbon in geological formations is referred to as Carbon Capture and Storage (CCS).
-Together with carbon transport, these technologies are grouped under the term carbon management and are part of the so-called New Carbon Economy~\cite{arniehellerNewCarbonEconomy2019}.
+An integrated management of both hydrogen and carbon is considered a crucial component to achieve climate neutrality in energy systems including hard-to-abate sectors.
+Hydrogen offers an efficient way to transport and store energy from different sources.
+Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of both fossil and synthetic fuels through Carbon Capture (CC) techniques or harvested directly from the atmosphere using Direct-Air Capture (DAC).
+Additionally, it can be stored in geological formations, a process known as carbon sequestration or carbon capture and storage (CCS).
+In combination, both together build the basis for climate-neutral fuels like methane, methanol or Fischer-Tropsch fuels, central to Power-to-X (PtX) and Carbon Capture and Utilization (CCU) strategies.
 
-The European Union's goal to achieve climate neutrality by 2050 involves the implementation of carbon management strategies as a crucial component.
-This objective is integrated into a wide range of programs, funding schemes and initiatives~\cite{eu2023netzero,europeangreendeal,europeaninnovationfund}.
-
-%~\cite{lockwoodEuropeanStrategyCarbon} - Clean Air Task Force
-%~\cite{righettiSitingCarbonDioxide2017} - co2 pipeline are a mature technology, references to US and Canada co2 pipelines used for enhanced oil recovery
-%~\cite{friedmannNETZEROGEOSPHERICRETURN} - high learning rate for co2 pipeline costs
-%~\cite{NorthernLightsWhat} - Northern Lights project in Norway, 1.5 Mt/a co2 transport capacity, 1.5 Mt/a co2 storage capacity, 2024 start of operation, targeted scale of 5 Mtpa CO2 Seq. by 2030.
-
+Recently, policymakers and industry in Europe have been committing to both hydrogen and carbon management projects, planning infrastructure components, and developing business models for emerging sectors of the economy.
+With the European Union's goal of achieving climate neutrality by 2050, a wide range of programs, funding models and initiatives have been established in this regard~\cite{eu2023netzero,europeangreendeal,europeaninnovationfund}.
+Initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier.
+At the same time, business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offer \carbon{} offtake on the other hand.
 For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.4~Gt of carbon capture from point sources per year.
 In combination with large sequestration potentials as stated in~\cite{weiProposedGlobalLayout2021}, this highlights the vast potential for decarbonization.
-Recently, policymakers and industry in Europe have been committing to carbon management projects, planning the first infrastructure components, and developing business models for emerging sectors of the economy~\cite{adomaitisEquinorRWEBuild2023,apnewswireGermanyDrawLegislation2023,KohlenstoffKannKlimaschutz2023,OGETESJoin2022,TESHydrogenLife2023}.
-Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offer \carbon{} offtake on the other hand.
-In this context, the need for an international strategy, analogous to the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022}, becomes apparent to coordinate and support deep decarbonization efforts.
-Important insights to formulate such a strategy can be gained from energy system models that provide a holistic view of the energy system and its technological interactions.
-To this end, a number of optimization tools have been developed in recent years to model synergies between various carbon management technologies~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
-In contrast to Integrated Assessment Models, these tools account for the spatial distribution of carbon sources and sinks.
-% point out that~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a co2 network, but only considers the power sector without co-optimizing renewable deployment.
+The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a 1.5 Mt/a co2 transport and sequestration capacity to be operating in 2024, expanding to a targeted scale of 5 Mt/a sequestration by 2030.
+To this end, the Clean Air Task Force underlines the importance of a carbon transport system in Europe to facilitate the carbon economy~\cite{lockwoodEuropeanStrategyCarbon}.
 
-% A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
-However, the models are often limited with regard to both geographical scope and detail.
+However, up to this point, it remains unclear how the two management systems of hydrogen and carbon may complement or replace each other, particularly when it comes to their transport systems.
+A hydrogen network facilitates the transport of energy from renewable sources to regions with high energy demand, also enabling Carbon Capture and Utilization (CCU) from specific point sources.
+On the other hand, the carbon transport system allows for capturing and transporting carbon from areas with high emissions to regions with significant sequestration potential and abundant renewable sources, thereby enabling cost-effective CCU.
+In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
+In contrast to Integrated Assessment Models, the underlying optimization tools account for the spatial distribution of carbon sources and sinks which are crucial provide a holistic view of the energy system and its technological interactions..
+The work in~\cite{neumannBenefitsHydrogenNetwork2022} examines the effect of a hydrogen network in Europe...
+(extend on hydrogen literature)
+The publication by Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \carbon{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
+Another comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
+To this end, it is often argued that \carbon{} pipelines are a mature technology with an expected high learning rate, given the wide-spread installations in US and Canada for enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
+% However, the models are often limited with regard to both geographical scope and detail.
 While representing a single country with spatial resolution may neglect synergies of international cooperation, a coarse grained representation of multiple countries may neglect important geographical properties.
 
+To our knowledge, no study has yet considered the co-optimization and comprehensive assessment of both \hydrogen{} and \carbon{} networks in a fully sector-coupled energy system.
+However, we argue that such an assessment is strongly needed to avoid suboptimal investments and to identify synergies between hydrogen and carbon management technologies.
 In this paper, we present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies.
@@
 
+\begin{figure}
+    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/total-demand-bar.png}
+    \caption{Exogenous demand assumptions. The figure shows the total amount of yearly energy demands for each energy carrier and type of end-consumer considered in the model.}
+    \label{fig:total-demand-bar}
+\end{figure}
+%
 In our configuration, the model's time horizon spans one year with a temporal resolution of 3 hours and a spatial resolution of 90 regions.
@@
 % use https://pypsa-eur.readthedocs.io/en/latest/licenses.html
-
-\begin{figure}
-    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/total-demand-bar.png}
-    \caption{Exogenous demand assumptions. The figure shows the total amount of yearly energy demands for each energy carrier and type of end-consumer considered in the model.}
-    \label{fig:total-demand-bar}
-\end{figure}
-
+%
 Endogenous assumptions include the expansion of renewable energy sources, storage technologies, and transmission capacities.
@@
 Inter-regional biomass transport is permitted with transport costs considered.
-Hydrogen can be gained from electrolysis and steam methane reforming (SMR). Salt caverns for \hydrogen{} storage are derived from~\cite{caglayanTechnicalPotentialSalt2020}.
 
-\dots
+Hydrogen can be gained from electrolysis and steam methane reforming (SMR).
+Geological distribution and potentials for \hydrogen storage in salt caverns are derived from~\cite{caglayanTechnicalPotentialSalt2020}.
+Re-electrification of hydrogen is possible via fuel cells.
+If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limit.
+The topology of the \hydrogen network is mirroring the topology of the electricity network.
 
-% carbon capture technologies
-Carbon capture technologies are represented by a set of augmented technologies with integrated carbon capture.
-These include Allam Cycle, Gas CHP, SMR, DAC, biomass, biogas and capture of process emissions from industry.
+Carbon capture (CC) technologies consist of a range of augmented technologies with integrated carbon capture.
+These include Allam Cycle Power Plants, Gas Combined Heat and Power Plant (CHP), Steam Methan Reforming (SMR), Direc-Air Capture (DAC), biomass combustion, biogas combustion and capture of process emissions from industry.
 
-Regarding carbon utilization, our model features three drop-in fuel production technologies: methanation, methanolization, and Fischer-Tropsch synthesis.
-Processed fuels have no usage restrictions.
+Our model features three drop-in fuel production technologies for carbon utilization: methanation, methanolization, and Fischer-Tropsch synthesis.
+The processed fuels are not limited in their total quantity of use or production.
 Methane is transported through the gas network, while methanol and FT-synfuel can be transferred inter-regionally without additional costs.
@@
 Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production, aviation kerosene, or agricultural machinery oil.
-For carbon sequestration, we only consider offshore sites as potential sinks.
+We only consider offshore sites as potential sinks for carbon sequestration.
 This choice favors offshore storage due to its typically larger capacity compared to onshore and concerns over public safety for infrastructure near populated areas.
 Our estimates for storage potential are conservative, limiting total potential to 25~Mt per site, calculating annual storage availability over 25 years, and capping the total sequestered \carbon{} to 200~Mt/a.
-
 All technology cost assumptions for 2040 are sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
-
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=\linewidth]{baseline/figures/90_nodes/sequestration_map.pdf}
-%     \caption{Sequestration map}
-%     \label{fig:sequestration_map}
-% \end{figure}
 
@@
 
-% hight level trends
-- cost reduction with carbon network is significant (2.5\%)
+The comprehensive analysis of the European energy system scenarios in 2050 highlights significant variations in system costs and technology deployment across the different network strategies, as depicted in Figure \ref{fig:cost_bar}.
+The system cost decomposition for the Baseline, CO2-Net, H2-Net, and Hybrid scenarios are visualized as stacked bar plots, facilitating an insightful comparison of the economic implications of each scenario.
 
@@
 "Gas Infrastructure" combines gas facilities for power and heat production, "H$_2$ Infrastructure" combines H$_2$ production, transport and re-electrification.
-The
-    }
-    \label{fig:cost_area}
+"Carbon Capture at Point Sources" combines all technologies with integrated carbon capture.}
+    \label{fig:cost_bar}
+\end{figure}
+
+In the Baseline scenario, the absence of dedicated hydrogen and carbon networks necessitates a heavy reliance on local renewable generation and storage solutions, reflected in the substantial investments in solar and wind capacities.
+The cost of carbon removal technologies, CC and DAC, also forms a significant portion of the overall expenses.
+However, this scenario's limitations are apparent, as the lack of transport networks for hydrogen and carbon leads to higher system costs due to inefficiencies and lost opportunities for optimization.
+
+Transitioning to the CO2-Net scenario, we observe a significant cost reduction and a shift in cost allocation.
+The implementation of a \carbon{} transport network (\textit{\carbon{}-Net}) facilitates efficient point-source carbon capture, substantially reducing the dependency on direct air capture (DAC) technologies.
+The improved carbon management infrastructure also allows for a more strategic deployment of renewables, particularly in areas with high sequestration potential, leading to cost savings in renewable generation and storage.
+
+The H2-Net scenario underscores the pivotal role of hydrogen infrastructure.
+The creation of a hydrogen transport network enables the cost-effective delivery of hydrogen from regions with a surplus of cheap renewable energy to high-demand industrial centers.
+This scenario shows a pronounced decrease in investments in solar and gas  infrastructure.
+
+The Hybrid scenario represents the most integrated network strategy, combining both CO2 and H2 networks.
+While the cost-reduction compared to the \hydrogen-Net scenario is marginal, the allocation of costs reveals a propagation of cost savings in DAC facilities.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/captureshare_line.png}
+    \caption{Share of facilities with integrated carbon capture for different system setups for the net-zero emission scenario.}
+    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
+    \label{fig:captureshare_line}
+\end{figure}
+
+With the cost reduction across the different system setups, the utilization of carbon capture technologies varies significantly, as depicted in Figure \ref{fig:captureshare_line}.
+This variation highlights how the implementation of \hydrogen{} and \carbon{} networks unlocks specific carbon capture potentials.
+
+While the share of CC share for the combustion of upgraded biogas (100\%) and Gas in CHP (0\%) remains unchanged across all scenarios, the
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/energy_balance_bar_carbon.png}
+    \caption{Balance of captured carbon across all system setups. Positive values indicate carbon capture, negative values indicate carbon consumption.}
+    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
+    \label{fig:captureshare_line}
 \end{figure}
@@
 %
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/captureshare_line.png}
-    \caption{Share of facilities with integrated carbon capture as a function of emission targets for system with \hydrogen{} and \carbon{} network.}
-    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
-    \label{fig:captureshare_line}
-\end{figure}
 

### 2023-11-20 report/submissions/journal/paper/paper.tex
follow up

@@
 
-The Hybrid scenario represents the most integrated network strategy, combining both CO2 and H2 networks.
-While the cost-reduction compared to the \hydrogen-Net scenario is marginal, the allocation of costs reveals a propagation of cost savings in DAC facilities.
+While the cost-reduction in the Hybrid scenario compared to the \hydrogen-Net scenario is marginal, the allocation of costs reveals a propagation of reduced DAC facilities and increased CC at point sources.
+The Hybrid scenario relies much heavier on investments into the \hydrogen network which outweigh investments in the \carbon{} network by a factor 4.
 
@@
 
-While the share of CC share for the combustion of upgraded biogas (100\%) and Gas in CHP (0\%) remains unchanged across all scenarios, the
+While the share of CC share for the combustion of upgraded biogas (100\%) and Gas in CHP (0\%) remains unchanged across all scenarios, the introduction of a
 

### 2023-11-20 workflow/scripts/plot-demand.py

@@ -53,7 +53,7 @@ color = n.carriers.set_index("nice_name").color.to_dict()
 
 
 fig, ax = plt.subplots(figsize=(5, 4))
-df.T.plot.bar(stacked=True, color=color, xlabel="", ylabel="Energy demand [pWh]", ax=ax)
+df.T.plot.bar(stacked=True, color=color, xlabel="", ylabel="Energy demand [PWh]", ax=ax)
 ax.legend(title="", frameon=False, bbox_to_anchor=(0, 1.2), loc="upper left")
 
 sns.despine()

### 2023-11-20 report/submissions/journal/paper/paper.tex
add description of merit order of captured carbon in Methodology, add initial description of baseline and CO2 scenarios in Results

--- 2023-11-20 report/submissions/journal/paper/paper.tex
@@
 \usepackage{caption}
+\captionsetup{font=small}
 \usepackage{algorithmic}
@@
 \newcommand{\hydrogen}{H$_2$}
+\newcommand{\scbase}{Baseline scenario}
+\newcommand{\scCO}{CO$_2$~Grid scenario}
+\newcommand{\scH}{H$_2$~Grid scenario}
+\newcommand{\schybrid}{Hybrid scenario}
 
@@
 Hydrogen offers an efficient way to transport and store energy from different sources.
-Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of both fossil and synthetic fuels through Carbon Capture (CC) techniques or harvested directly from the atmosphere using Direct-Air Capture (DAC).
+Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of both fossil and synthetic fuels through Carbon Capture (CC) techniques or harvested directly from the atmosphere using Direct Air Capture (DAC).
 Additionally, it can be stored in geological formations, a process known as carbon sequestration or carbon capture and storage (CCS).
@@
 
-Carbon capture (CC) technologies consist of a range of augmented technologies with integrated carbon capture.
-These include Allam Cycle Power Plants, Gas Combined Heat and Power Plant (CHP), Steam Methan Reforming (SMR), Direc-Air Capture (DAC), biomass combustion, biogas combustion and capture of process emissions from industry.
-
 Our model features three drop-in fuel production technologies for carbon utilization: methanation, methanolization, and Fischer-Tropsch synthesis.
@@
 Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production, aviation kerosene, or agricultural machinery oil.
+
+To supply carbon needed for CS and CU, the system can choose to deploy carbon capture technology at various point sources, or centralized DAC facilities.
+The concept of a ``merit order for captured carbon'' plays a pivotal role in optimizing and prioritizing the deployment of carbon capture technologies based on their economic feasibility.
+This merit order ranks all CC technologies according to their relative costs to capture a ton of carbon (see Figure~X).
+At the lower end of the cost spectrum, CC technologies applied to process emissions, such as those from cement, offer a cost-effective starting point.
+Following this, biomass combined heat and power (CHP) systems provide a dual benefit of energy production and carbon capture.
+Moving up the scale, gas used in industrial applications and biomass employed for industrial processes represent more costly yet viable options to capture carbon.
+DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a higher cost level due to its current technological and operational expenses.
+Finally, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, typically incurs the highest costs in the merit order per ton of captured \carbon{} but offers a flexible (yet carbon emitting) energy source as a product.
+This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.
+
 We only consider offshore sites as potential sinks for carbon sequestration.
@@
     \item[] \textit{Net-Zero}: aligning with the EU's 2050 emission targets. Also denoted as NZ.
-    \item[] \textit{Net-Negative~5\%}: relative to 1990 emissions, equaling 230~Mt \carbon{}. Also denoted as NN~5\%.
-    \item[] \textit{Net-Negative~10\%}: relative to 1990 emissions, equaling 460~Mt \carbon. Also denoted as NN~10\%.
+    \item[] \textit{Net-Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \carbon{} annually. Also denoted as NN.
 \end{itemize}
@@
     \item[] \textit{Baseline:} Neither \carbon{} nor \hydrogen{} networks are constructed.
-    \item[] \textit{\hydrogen{}-Net:} Only the \hydrogen{} network is developed.
-    \item[] \textit{\carbon{}-Net:} Only the \carbon{} network is developed.
+    \item[] \textit{\hydrogen{}~Grid:} Only the \hydrogen{} network is developed.
+    \item[] \textit{\carbon{}~Grid:} Only the \carbon{} network is developed.
     \item[] \textit{Hybrid:} Both \carbon{} and \hydrogen{} networks are developed.
@@
 
-The comprehensive analysis of the European energy system scenarios in 2050 highlights significant variations in system costs and technology deployment across the different network strategies, as depicted in Figure \ref{fig:cost_bar}.
-The system cost decomposition for the Baseline, CO2-Net, H2-Net, and Hybrid scenarios are visualized as stacked bar plots, facilitating an insightful comparison of the economic implications of each scenario.
+Focusing first on the NZ scenarios, we can highlight significant variations in system costs and technology deployment depending on the network strategy employed (see Figure~\ref{fig:cost_bar}).
 
@@
 
-In the Baseline scenario, the absence of dedicated hydrogen and carbon networks necessitates a heavy reliance on local renewable generation and storage solutions, reflected in the substantial investments in solar and wind capacities.
-The cost of carbon removal technologies, CC and DAC, also forms a significant portion of the overall expenses.
-However, this scenario's limitations are apparent, as the lack of transport networks for hydrogen and carbon leads to higher system costs due to inefficiencies and lost opportunities for optimization.
+System cost differences between the scenarios are driven by renewable energy resources (solar, wind and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
+In the Baseline scenario, the absence of dedicated carbon and hydrogen transport networks requires the system to deploy carbon production technologies (DAC and CC at point sources) in the regions with carbon sequestration sites, and small amounts of hydrogen production technologies (mainly electrolysis) in each region to meet local hydrogen demand for industry and land transport fuel cells.
+For CU and its respective carbon and hydrogen demands, the system chooses the sites where on-site production of carbon, hydrogen and the synthesis process is cheapest.
+This leads to renewable resources and storage solutions being deployed not necessarily in the regions with the best resources or lowest costs for clean electricity generation, but where they are situated closer to the energy demand locations of electrolysers and DAC.
+Thus, total system costs in the baseline scenario are highest, while the scenarios with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs.
 
-Transitioning to the CO2-Net scenario, we observe a significant cost reduction and a shift in cost allocation.
-The implementation of a \carbon{} transport network (\textit{\carbon{}-Net}) facilitates efficient point-source carbon capture, substantially reducing the dependency on direct air capture (DAC) technologies.
+We can now closer examine the cost reductions and shifts in cost allocation when shifting from the baseline scenario to the other three network scenarios.
+The implementation of a \carbon{} transport network (\textit{\carbon{}~Grid}) facilitates efficient point-source carbon capture, substantially reducing the dependency on DAC technologies.
+While DAC supplies around 50\% of the carbon needed in the Baseline scenario (400 Mt \carbon{}, see Figure~\ref{fig:balance_captured_carbon}), in the \scCO scenario, DAC is only used to supply around 20\% of the carbon (150 Mt \carbon{}).
+The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Figure~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology is relatively cheaper than DAC.
+The \carbon{} transport network enables the transport of carbon from the locations of the cheapest captured carbon supply sources to the locations of the carbon demand sinks at CU amd CS sites, reducing the need for expensive DAC.
 The improved carbon management infrastructure also allows for a more strategic deployment of renewables, particularly in areas with high sequestration potential, leading to cost savings in renewable generation and storage.
-
-The H2-Net scenario underscores the pivotal role of hydrogen infrastructure.
-The creation of a hydrogen transport network enables the cost-effective delivery of hydrogen from regions with a surplus of cheap renewable energy to high-demand industrial centers.
-This scenario shows a pronounced decrease in investments in solar and gas  infrastructure.
-
-The Hybrid scenario represents the most integrated network strategy, combining both CO2 and H2 networks.
-While the cost-reduction compared to the \hydrogen-Net scenario is marginal, the allocation of costs reveals a propagation of cost savings in DAC facilities.
 
@@
 
-With the cost reduction across the different system setups, the utilization of carbon capture technologies varies significantly, as depicted in Figure \ref{fig:captureshare_line}.
-This variation highlights how the implementation of \hydrogen{} and \carbon{} networks unlocks specific carbon capture potentials.
-
-While the share of CC share for the combustion of upgraded biogas (100\%) and Gas in CHP (0\%) remains unchanged across all scenarios, the
-
 \begin{figure}
@@
     % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
-    \label{fig:captureshare_line}
+    \label{fig:balance_captured_carbon}
 \end{figure}
+
+The \hydrogen{}~Grid scenario underscores the pivotal role of hydrogen infrastructure.
+The creation of a hydrogen transport network enables the cost-effective delivery of hydrogen from regions with a surplus of cheap renewable energy to high-demand industrial centers.
+This scenario shows a pronounced decrease in investments in solar and gas  infrastructure.
+The \hydrogen{} network reduces DAC similarly to the \carbon{} network, but still produces 200 Mt of carbon from DAC.
+CC technology on process emissions increases from 90\% to 100\%, on biomass CHPs from 60\% to 100\%, and on facilities using gas in industry from 50\% to 75\%.
+CC on biomass use in industry remains at 50\%, same as the \scbase.
+This variation highlights how the implementation of \hydrogen{} and \carbon{} networks unlocks specific carbon capture potentials.
+
+While the cost-reduction in the Hybrid scenario compared to the \hydrogen~Grid scenario is marginal, the allocation of costs reveals a propagation of reduced DAC facilities and increased CC at point sources.
+The Hybrid scenario relies much heavier on investments into the \hydrogen network which outweigh investments in the \carbon{} network by a factor of 4.
 
@@
 
+[Start of old version]
 First, we analyze the optimal deployment of technologies and their associated costs at different levels of carbon sequestration potential.

### 2023-11-22 report/submissions/journal/paper/abstract.tex
further work

--- 2023-11-20 report/submissions/journal/paper/paper.tex
@@
-\documentclass[twocolumn]{article}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{cuted}  % Add the cuted package
-\usepackage{caption}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{tabularx, multirow}
-\usepackage{fancyhdr,lipsum}
-\usepackage{subcaption}
-
-\usepackage[%
-backend=biber,bibencoding=utf8, %instead of bibtex
-language=auto,
-style=ieee,
-sorting=none, % nyt for name, year, title
-maxbibnames=10, % default: 3, et al.
-%backref=true,%
-natbib=true % natbib compatibility mode (\citep and~\citet still work)
-]{biblatex}
-\bibliography{../../../references.bib}
-
-%define approx proportional
-\def\app#1#2{%
-  \mathrel{%
-    \setbox0=\hbox{$#1\sim$}%
-    \setbox2=\hbox{%
-      \rlap{\hbox{$#1\propto$}}%
-      \lower1.1\ht0\box0%
-    }%
-    \raise0.25\ht2\box2%
-  }%
-}
-\def\approxprop{\mathpalette\app\relax}
-
-% make abbreviation for co2
-\newcommand{\carbon}{CO$_2$}
-\newcommand{\hydrogen}{H$_2$}
-
-\graphicspath{
-    % {paper-figures}
-    {figures}
-    {../../../../results/}
-    {results/}
-}
-
-\begin{document}
-
-\title{From Net-Zero to Net-Negative: Assessing \hydrogen{} and \carbon{} Network Strategies in Europe}
-
-\author{
-    Fabian Hofmann, Christoph Tries, Fabian Neumann, Lisa Zeyen, Tom Brown \\
-    \textit{Institute of Energy Technology} \\
-    \textit{Technical University of Berlin}\\
-    Berlin, Germany \\
-    m.hofmann@tu-berlin.de
-}
-
-\maketitle
-
-\begin{abstract}
     The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors, particularly in emissions-intensive industries like heavy manufacturing.
@@
     The paper underpins the need for a complementing hydrogen network and \carbon{} network to achieve climate neutrality.
-\end{abstract}
-
-\section{Introduction}
-\label{sec:introduction}
-
-An integrated management of both hydrogen and carbon is considered a crucial component to achieve climate neutrality in energy systems including hard-to-abate sectors.
-Hydrogen offers an efficient way to transport and store energy from different sources.
-Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of both fossil and synthetic fuels through Carbon Capture (CC) techniques or harvested directly from the atmosphere using Direct-Air Capture (DAC).
-Additionally, it can be stored in geological formations, a process known as carbon sequestration or carbon capture and storage (CCS).
-In combination, both together build the basis for climate-neutral fuels like methane, methanol or Fischer-Tropsch fuels, central to Power-to-X (PtX) and Carbon Capture and Utilization (CCU) strategies.
-
-Recently, policymakers and industry in Europe have been committing to both hydrogen and carbon management projects, planning infrastructure components, and developing business models for emerging sectors of the economy.
-With the European Union's goal of achieving climate neutrality by 2050, a wide range of programs, funding models and initiatives have been established in this regard~\cite{eu2023netzero,europeangreendeal,europeaninnovationfund}.
-Initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier.
-At the same time, business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offer \carbon{} offtake on the other hand.
-For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.4~Gt of carbon capture from point sources per year.
-In combination with large sequestration potentials as stated in~\cite{weiProposedGlobalLayout2021}, this highlights the vast potential for decarbonization.
-The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a 1.5 Mt/a co2 transport and sequestration capacity to be operating in 2024, expanding to a targeted scale of 5 Mt/a sequestration by 2030.
-To this end, the Clean Air Task Force underlines the importance of a carbon transport system in Europe to facilitate the carbon economy~\cite{lockwoodEuropeanStrategyCarbon}.
-
-However, up to this point, it remains unclear how the two management systems of hydrogen and carbon may complement or replace each other, particularly when it comes to their transport systems.
-A hydrogen network facilitates the transport of energy from renewable sources to regions with high energy demand, also enabling Carbon Capture and Utilization (CCU) from specific point sources.
-On the other hand, the carbon transport system allows for capturing and transporting carbon from areas with high emissions to regions with significant sequestration potential and abundant renewable sources, thereby enabling cost-effective CCU.
-In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
-In contrast to Integrated Assessment Models, the underlying optimization tools account for the spatial distribution of carbon sources and sinks which are crucial provide a holistic view of the energy system and its technological interactions..
-The work in~\cite{neumannBenefitsHydrogenNetwork2022} examines the effect of a hydrogen network in Europe...
-(extend on hydrogen literature)
-The publication by Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \carbon{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
-Another comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
-To this end, it is often argued that \carbon{} pipelines are a mature technology with an expected high learning rate, given the wide-spread installations in US and Canada for enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
-% However, the models are often limited with regard to both geographical scope and detail.
-While representing a single country with spatial resolution may neglect synergies of international cooperation, a coarse grained representation of multiple countries may neglect important geographical properties.
-
-To our knowledge, no study has yet considered the co-optimization and comprehensive assessment of both \hydrogen{} and \carbon{} networks in a fully sector-coupled energy system.
-However, we argue that such an assessment is strongly needed to avoid suboptimal investments and to identify synergies between hydrogen and carbon management technologies.
-In this paper, we present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies.
-The study is conducted using the PyPSA-Eur energy system model and encompasses all relevant energy sectors.
-We enable the system to optimize the design of renewable energy sources and storage technologies, as well as the transmission of electricity, \hydrogen{}, methane, and \carbon{}.
-We show that the dynamics of carbon capture, transport, use, and storage in different regions are highly dependent on the availability of renewable energy, \carbon{} pipelines and sequestration sinks.
-Our evaluation focuses on the transport dynamics of \carbon{} and \hydrogen{} through their respective networks on the European continent.
-We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available sequestration potential.
-
-\section{Methodology}
-\label{sec:methodology}
-
-The study is conducted on the basis of the open-source, capacity-expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
-The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, waste, and transport sectors.
-
-\begin{figure}
-    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/total-demand-bar.png}
-    \caption{Exogenous demand assumptions. The figure shows the total amount of yearly energy demands for each energy carrier and type of end-consumer considered in the model.}
-    \label{fig:total-demand-bar}
-\end{figure}
-%
-In our configuration, the model's time horizon spans one year with a temporal resolution of 3 hours and a spatial resolution of 90 regions.
-Each of the regions consists of a complex subsystem with technologies for supplying, converting, storing and transporting energy.
-Exogenous assumptions on energy demand and non-abatable emissions are taken from various sources~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023}.
-The energy demand for electricity, transport, biomass, heat and gas are defined per region and time-step with heat demand being regionally subdivided into shares of urban, rural and industrial sites.
-Demands for aviation kerosene, methanol for shipping and naphtha for industry are aggregated in the system scope and kept constant throughout all time-steps.
-We show the sum of all energy demands in \ref{fig:total-demand-bar}.
-The system emits of 633~Mt \carbon{} per year through non-abatable emissions from industry, aviation, shipping and agriculture into account, 153~Mt of which are fossil-based industrial emissions.
-Industrial energy demand and excess heat potentials are calculated per node on the basis of~\cite{hotmaps_industrial_db}.
-% use https://pypsa-eur.readthedocs.io/en/latest/licenses.html
-%
-Endogenous assumptions include the expansion of renewable energy sources, storage technologies, and transmission capacities.
-The model considers the various energy carriers like electricity, hydrogen, methan, methanol, liquid hydrocarbons and biomass, together with conversions technologies to convert these into each other.
-Carbon-neutral electricity is provided by wind, solar, biomass, hydro and nuclear power plants.
-Weather-dependent power potentials for solar, wind and hydro are calculated from reanalysis and satellite data sets~\cite{hersbachERA5GlobalReanalysis2020,pfeifrothSurfaceRadiationData2017}  per region and time-stamp, using the open-source tool Atlite~\cite{hofmannAtliteLightweightPython2021}.
-Solar and wind power can be expanded in alignment with eligible land-use restrictions calculated on the basis of~\cite{eeaCorineLandCover2012,eeaNatura2000Data2016}.
-We restrict the electric transmission system's expansion to 20\% of its current capacity, acknowledging the challenges in inaugurating new transmission projects.
-The regional use of biomass is restricted by its feasible potential derived from~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
-Inter-regional biomass transport is permitted with transport costs considered.
-
-Hydrogen can be gained from electrolysis and steam methane reforming (SMR).
-Geological distribution and potentials for \hydrogen storage in salt caverns are derived from~\cite{caglayanTechnicalPotentialSalt2020}.
-Re-electrification of hydrogen is possible via fuel cells.
-If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limit.
-The topology of the \hydrogen network is mirroring the topology of the electricity network.
-
-Carbon capture (CC) technologies consist of a range of augmented technologies with integrated carbon capture.
-These include Allam Cycle Power Plants, Gas Combined Heat and Power Plant (CHP), Steam Methan Reforming (SMR), Direc-Air Capture (DAC), biomass combustion, biogas combustion and capture of process emissions from industry.
-
-Our model features three drop-in fuel production technologies for carbon utilization: methanation, methanolization, and Fischer-Tropsch synthesis.
-The processed fuels are not limited in their total quantity of use or production.
-Methane is transported through the gas network, while methanol and FT-synfuel can be transferred inter-regionally without additional costs.
-Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial gas needs.
-Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production, aviation kerosene, or agricultural machinery oil.
-We only consider offshore sites as potential sinks for carbon sequestration.
-This choice favors offshore storage due to its typically larger capacity compared to onshore and concerns over public safety for infrastructure near populated areas.
-Our estimates for storage potential are conservative, limiting total potential to 25~Mt per site, calculating annual storage availability over 25 years, and capping the total sequestered \carbon{} to 200~Mt/a.
-All technology cost assumptions for 2040 are sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
-
-Addressing climate targets, we define three \carbon{} target scenarios:
-\begin{itemize}
-    \item[] \textit{Net-Zero}: aligning with the EU's 2050 emission targets. Also denoted as NZ.
-    \item[] \textit{Net-Negative~5\%}: relative to 1990 emissions, equaling 230~Mt \carbon{}. Also denoted as NN~5\%.
-    \item[] \textit{Net-Negative~10\%}: relative to 1990 emissions, equaling 460~Mt \carbon. Also denoted as NN~10\%.
-\end{itemize}
-Unless otherwise mentioned, our reference is the Net-zero scenario.
-For 5\% Net-negative and 10\% Net-negative scenarios, the carbon sequestration potential is adjusted to 400~Mt/a and 600~Mt/a, respectively.
-
-Beyond \carbon{} targets, we consider four expansion scenarios:
-\begin{itemize}
-    \item[] \textit{Baseline:} Neither \carbon{} nor \hydrogen{} networks are constructed.
-    \item[] \textit{\hydrogen{}-Net:} Only the \hydrogen{} network is developed.
-    \item[] \textit{\carbon{}-Net:} Only the \carbon{} network is developed.
-    \item[] \textit{Hybrid:} Both \carbon{} and \hydrogen{} networks are developed.
-\end{itemize}
-
-\section{Results}
-\label{sec:results}
-
-The comprehensive analysis of the European energy system scenarios in 2050 highlights significant variations in system costs and technology deployment across the different network strategies, as depicted in Figure \ref{fig:cost_bar}.
-The system cost decomposition for the Baseline, CO2-Net, H2-Net, and Hybrid scenarios are visualized as stacked bar plots, facilitating an insightful comparison of the economic implications of each scenario.
-
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/cost_bar.png}
-    \caption[short]{Total annual system cost for the European energy system for the different scenarios subdivided into groups of technologies.
-While the `Baseline' scenario has neither a \hydrogen{} or \carbon{} network, `Hybrid' is allowed to expand both.
-"Gas Infrastructure" combines gas facilities for power and heat production, "H$_2$ Infrastructure" combines H$_2$ production, transport and re-electrification.
-"Carbon Capture at Point Sources" combines all technologies with integrated carbon capture.}
-    \label{fig:cost_bar}
-\end{figure}
-
-In the Baseline scenario, the absence of dedicated hydrogen and carbon networks necessitates a heavy reliance on local renewable generation and storage solutions, reflected in the substantial investments in solar and wind capacities.
-The cost of carbon removal technologies, CC and DAC, also forms a significant portion of the overall expenses.
-However, this scenario's limitations are apparent, as the lack of transport networks for hydrogen and carbon leads to higher system costs due to inefficiencies and lost opportunities for optimization.
-
-Transitioning to the CO2-Net scenario, we observe a significant cost reduction and a shift in cost allocation.
-The implementation of a \carbon{} transport network (\textit{\carbon{}-Net}) facilitates efficient point-source carbon capture, substantially reducing the dependency on direct air capture (DAC) technologies.
-The improved carbon management infrastructure also allows for a more strategic deployment of renewables, particularly in areas with high sequestration potential, leading to cost savings in renewable generation and storage.
-
-The H2-Net scenario underscores the pivotal role of hydrogen infrastructure.
-The creation of a hydrogen transport network enables the cost-effective delivery of hydrogen from regions with a surplus of cheap renewable energy to high-demand industrial centers.
-This scenario shows a pronounced decrease in investments in solar and gas  infrastructure.
-
-The Hybrid scenario represents the most integrated network strategy, combining both CO2 and H2 networks.
-While the cost-reduction compared to the \hydrogen-Net scenario is marginal, the allocation of costs reveals a propagation of cost savings in DAC facilities.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/captureshare_line.png}
-    \caption{Share of facilities with integrated carbon capture for different system setups for the net-zero emission scenario.}
-    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
-    \label{fig:captureshare_line}
-\end{figure}
-
-With the cost reduction across the different system setups, the utilization of carbon capture technologies varies significantly, as depicted in Figure \ref{fig:captureshare_line}.
-This variation highlights how the implementation of \hydrogen{} and \carbon{} networks unlocks specific carbon capture potentials.
-
-While the share of CC share for the combustion of upgraded biogas (100\%) and Gas in CHP (0\%) remains unchanged across all scenarios, the
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/energy_balance_bar_carbon.png}
-    \caption{Balance of captured carbon across all system setups. Positive values indicate carbon capture, negative values indicate carbon consumption.}
-    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
-    \label{fig:captureshare_line}
-\end{figure}
-
-%
-
-\begin{figure*}
-    \centering
-    \begin{subfigure}{\linewidth}
-        \centering
-        \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_carbon.png}
-        \caption{\carbon{} Sector.}
-        \label{fig:operation_map_carbon}
-    \end{subfigure}
-    \begin{subfigure}{\linewidth}
-        \centering
-        \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_hydrogen.png}
-        \caption{\hydrogen{} Sector.}
-        \label{fig:operation_map_hydrogen}
-    \end{subfigure}
-    \caption{Optimal operation per sector for a net-zero energy system in Europe with both \carbon{} and \hydrogen{} network expansion (Hybrid).}
-    \label{fig:operation_map_carbon_co2}
-\end{figure*}
-
-\begin{figure*}[h]
-    \centering
-    \begin{subfigure}{.5\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{full/figures/90_nodes/capacity_map_carbon.png}
-        \caption{\carbon{} Sector.}
-        \label{fig:capacity_map_carbon_co2}
-    \end{subfigure}%
-    \hfill
-    \begin{subfigure}{.5\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{full/figures/90_nodes/capacity_map_hydrogen.png}
-        \caption{\hydrogen{} Sector.}
-        \label{fig:capacity_map_hydrogen_co2}
-    \end{subfigure}
-    \caption{Optimal production and transport capacities of the carbon and hydrogen sector in a net-zero energy system in Europe with both \carbon{} and \hydrogen{} network expansion (Hybrid).
-    % Carbon network looks the same as in~\cite{morbeeOptimisedDeploymentEuropean2012}: two backbones, one in the nothern Europe other in south east.
-    }
-    \label{fig:capacity_maps}
-\end{figure*}
-
-\begin{figure}
-    \centering
-    \includegraphics*[width=0.8\linewidth]{comparison/emission-reduction/figures/90_nodes/objective_heatmap.png}
-    \caption{Total annual system cost for the different scenarios and \carbon{} targets.
-While the `Baseline' scenario has neither a \hydrogen{} or \carbon{} network, `Hybrid' is allowed to expand both.}
-    \label{fig:objective_heatmap}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{comparison/emission-reduction-full/figures/90_nodes/transmission_cost_bar.png}
-    \caption{Annual transmission system cost as a function of the net carbon removal scenarios considered in the study.}
-    \label{fig:transmission_cost_bar}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
-    \caption{Sankey diagram of the optimal operation for a net-negative 10\% scenario.}
-    \label{fig:sankey_diagramm}
-\end{figure}
-
-First, we analyze the optimal deployment of technologies and their associated costs at different levels of carbon sequestration potential.
-Figure~\ref{fig:cost_area} displays the annual system cost per technology group for the sector-coupled model at various sequestration levels, both with and without a \carbon{} network.
-The costs are calculated as the sum of the annualized investment and operational costs for all technologies.
-Starting from the left, we observe that the majority of system costs are associated with the deployment of renewable energy sources, heating, hydrogen technologies, and the electricity grid.
-Costs for fossil fuels and carbon capture and transport technologies are relatively low.
-In terms of capacities, 200~GW of offshore wind, 3930~GW of solar PV, 1910~GW of onshore wind, and 1650~GW of electrolysis are installed in this setup.
-In both cases, with and without \carbon{} transport, total costs decrease as the sequestration potential increases.
-At high sequestration rates, fossil fuels such as gas and oil are kept in the system.
-This reduces the need for synthetic FT fuel, which requires significant renewable energy sources.
-Simultaneously, the system deploys more \carbon{} removal technologies, such as DAC and Bioenergy with carbon capture (included in "Carbon Capture Technologies"), to offset emissions from fossil fuels.
-
-The cost reduction when going to high sequestration in a system with \carbon{} transport is about 13\%, which is slightly more than without a \carbon{} network (11\%).
-Across all sequestration levels, a system with \carbon{} transport is, on average, 1.6\% less expensive.
-The cost savings for a carbon transport system are lowest at 0.5\% (4 billion euros per year) for a sequestration of 200~Mt and highest at 3\% (21 billion euros per year) for a sequestration of 1000~Mt.
-
-To further investigate the switch in technologies, we analyze the annual operation of the system for different sequestration potentials.
-Figures~\ref{fig:operation_area_carbon_co2network}, \ref{fig:operation_area_hydrogen_co2network} and \ref{fig:operation_area_co2_co2network} show the balances of captured carbon, hydrogen and atmospheric \carbon{} in the system with \carbon{} transport as a function of the sequestration potential respectively.
-The areas in the positive segment indicate gross production of the respective commodity, while the areas in the negative segment indicate gross consumption.
-Note that technologies with a ``*`` are explicitly augmented technologies with integrated carbon capture.
-%
-
-Figure~\ref{fig:operation_area_carbon_co2network} confirms our previous results and shows that FT synthesis production decreases with sequestration availability.
-At the same time, integrated carbon capture from Gas Combined Heat and Power (CHP) plants, Allam Cycle plants, Steam Methane Reforming (SMR) and DAC increases.
-Other technologies are constant across varying levels of sequestration due to different reasons: biomass is bound by its maximum available potential (as reported in~\cite{enspreso_database}); the industry processes with their associated emissions are set to fixed values; the methanol production is bound by a lower limit representing the demand from the shipping sector.
-
-% TODO: kerosine emissions from fossil based kerosine are not taken into account by the system.
-Figure~\ref{fig:operation_area_hydrogen_co2network} shows that hydrogen production decreases by about 50\% when going from 200~Mt/a sequestration to 1000~Mt/a.
-In the last segment, hydrogen production from SMR starts to replace hydrogen production from electrolysis.
-Correspondingly, hydrogen demand for FT synthetic fuel production disappears, leaving only hydrogen demand for the shipping, industry and transport sector.
-Finally, Figure~\ref{fig:operation_area_co2_co2network} shows that the balance of the atmospheric \carbon{}.
-Note that the absolute value in the atmospheric \carbon{} hardly changes with the sequestration potential.
-This is because the switch from FT to fossil fuels is mirrored by a switch from using captured \carbon{} from biomass and DAC for FT synthesis to sequestering it underground.
-The net increase in DAC is solely due to an increase in unabated emissions from gas boilers.
-
-We can conclude that in combination with a \carbon{} transport network, high rates of sequestration reduce the feasibility of CCU and hydrogen infrastructure and increases the feasibility of CCS, namely DAC and CC at point sources.
-
-To answer the question of how carbon capture technologies at point sources change with the sequestration potential, we analyze the shift to carbon capture integration for all relevant technologies.
-Figure~\ref{fig:captureshare_line} shows the share of a given technology that is equipped with integrated carbon capture, i.e., carbon capture at point source, for both with and without \carbon{} transport network.
-With \carbon{} transport, all technologies except for SMR and Gas CHP are nearly fully equipped with CC across all sequestration potentials.
-As a result of sequestration, the share of SMR with carbon capture rises from 50\% to 100\%, and that of CHP from 0\% to 26\%.
-The technologies behave differently without carbon transport: besides the fact that capturing at point source is less attractive overall, their carbon capture shares rise more slowly as sequestration increases.
-The carbon capture share for process emissions even decreases.
-This is primarily due to the fact that as sequestration increases, the system relies more heavily on DAC and SMR with CC that can be flexibly placed near the sequestration sites.
-
-Building on the analysis of optimal operation for varying sequestration potentials, we explore the spatial distribution of carbon capture and transport technologies for the most "optimistic" case of 1000~Mt/a sequestration potential and a carbon transport network.
-
-Figure~\ref{fig:operation_map_noco2network_1000} shows a map of the total production and consumption of captured carbon in the system.
-The left side shows the combined carbon supply from carbon capture facilities and the carbon network flow.
-The right side illustrates regional carbon usage, again, together with the flow.
-Large capacity DACs and SMRs are placed close to shore with carbon capture in order to maximize efficiency.
-Significant facilities for DAC are built in the British Isles where carbon dioxide is used for methanol synthesis and also sequestration.
-Here substantial onshore wind farms generate a total annual production of approximately 400 TWh per year, of which 13\% is utilized by local DAC facilities.
-Other notable DAC installations are located on the Atlantic coast of Portugal, Spain, Ireland, and Denmark.
-Large facilities of SMR with CC are found in Portugal, South England and Greece.
-All these locations are in close proximity to sequestration sites, as indicated in Figure~\ref{fig:sequestration_map}.
-Note that the installation of SMR facilities close to the shore has the additional benefit of enabling easy transportation of hydrogen to ports for the shipping sector.
-
-Looking at the transport of \carbon{} across the system, we observe that the carbon network primarily transports captured carbon from process and biomass emissions from the inland to sequestration areas.
-Note that optimization installs a number of separate carbon transport networks instead of building one large connected grid.
-Therefore, a large network cluster transports carbon from Central Europe to the North Sea.
-In contrast, three smaller clusters in the south of Europe transport carbon to the Atlantic and the Mediterranean Sea.
-
-As shown on the right, methanol production using captured carbon is primarily located in the United Kingdom, Ireland, and South Spain.
-
-Finally, the deployment of carbon capture technologies in a system with high sequestration rates contrasts with the deployment of CCU facilities in a system with low carbon sequestration.
-On the one hand, with high sequestration, carbon capture management in the system is mainly driven by the location of the sequestration sites, leading to a spatially centralized deployment of carbon capture technologies.
-On the other hand, with low sequestration rates the use of CCU in a system is distributed more evenly across the system.
-
-\section{Limitations}
-\label{sec:limitations}
-
-Despite the detailed model representation, there are some limitations to the validity of the results.
-The model itself is based on a linearized optimization with perfect foresight for the entire modeling year.
-In reality, long-term energy demand and renewable supply can only be roughly estimated, while short-term predictions still entail some uncertainty.
-The model's perfect foresight may lead to non-reproducible behavior, such as precisely aligning energy storage with future energy shortages at a specific point in time.
-
-The technology costs used in the model rely on cost projections that incorporate reductions based on learning rates.
-These learning rates are derived from historical data, which may not necessarily be indicative of future trends.
-The model does not account for uncertainties arising from disruptive market behavior, such as the gas price peak that occurred in 2022.
-
-% TODO: which fixed rate? you mean that all emissions are captured by exogenous assumption or that we assume 90% of carbon can be captured?
-% Furthermore, the model assumes a fixed rate for industrial process emissions, which cannot be altered through investments.
-This simplification may not accurately represent real-world scenarios, where many industries are considering adopting low-carbon processes and technologies.
-The transport of fuel (FT, gas, oil) between the regions is not unlimited which may overestimate the flexibility provided by these commodities.
-
-Lastly, the model's spatial and temporal resolution is insufficient to capture all relevant dynamics.
-Variations in renewable energy supply and energy demand below the applied time resolution, as well as more detailed interregional energy transport constraints, are currently not considered in the optimization.
-Transitioning to an hourly representation and a higher spatial resolution would increase the model's computational time and complexity, but would also enhance its robustness and validity.
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-In this study, we investigated the impact of a \carbon{} transport system with different carbon sequestration potentials on the optimal configuration of technologies in a fully sector-coupled energy system in Europe, focusing on the deployment of carbon capture and Utilization (CCU) and Carbon Capture and Storage (CCS) technologies.
-Our results indicate that increasing carbon sequestration potential leads to a decrease in total system costs, with a more pronounced cost reduction observed when a carbon transport network is present.
-Higher sequestration rates result in a shift from CCU and hydrogen infrastructure towards CCS and unabated fossil fuel usage compensated by carbon dioxide removal elsewhere in the system, particularly based on direct air and biomass carbon capture with sequestration.
-
-Furthermore, we observed that with a \carbon{} transport network, the system tends to build out more carbon capture at point sources.
-With an increased sequestration potential, the share of technologies equipped with integrated carbon capture rises, particularly for Steam Methane Reforming (SMR) and Gas Combined Heat and Power (CHP) plants.
-Independent of a setup with or without \carbon{} transport, the spatial distribution of carbon capture technologies becomes more centralized, with installations concentrated near sequestration sites.
-
-Despite limitations of the study considering fixed model assumptions and cost uncertainties, our findings contribute valuable insights into the role of a carbon transport system and carbon sequestration potential in the transition towards a climate-neutral and net-negative energy system.
-As Europe strives to achieve its 2050 climate neutrality target, understanding the interplay between carbon sequestration, CCU, and CCS technologies will be crucial for the design of effective policies and strategies.
-
-\printbibliography
-
-% \appendix
-
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=\linewidth]{capacity_map_electricity_co2network_1000.png}
-%     \caption{Optimal capacities per of the electricity sector for a sequestration of 1000 Mt/a.}
-%     \label{fig:capacity_map_electricity_co2network_1000}
-% \end{figure}
-
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=\linewidth]{capacity_map_hydrogen_co2network_1000.png}
-%     \caption{Optimal capacities per of the hydrogen sector for a sequestration of 1000 Mt/a.}
-%     \label{fig:capacity_map_hydrogen_co2network_1000}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[width=\linewidth]{operation_area_carbon_noco2network.pdf}
-%     \caption{Balance of captured \carbon{} emissions for the optimal operation without \carbon{} network.}
-%     \label{fig:operation_area_carbon_noco2network}
-% \end{figure}
-
-\end{document}

### 2023-11-22 report/submissions/journal/paper/paper.tex
further work

@@
 \newcommand{\hydrogen}{H$_2$}
+\newcommand{\hydrogengrid}{\hydrogen{}--Grid}
+\newcommand{\carbongrid}{\carbon{}--Grid}
 \newcommand{\scbase}{Baseline scenario}
@@
 \begin{document}
+
+% FORMATTING: https://www.nature.com/nenergy/submission-guidelines/aip-and-formatting
 
@@
 \end{abstract}
-
-\section{Introduction}
-\label{sec:introduction}
 
@@
 To supply carbon needed for CS and CU, the system can choose to deploy carbon capture technology at various point sources, or centralized DAC facilities.
-The concept of a ``merit order for captured carbon'' plays a pivotal role in optimizing and prioritizing the deployment of carbon capture technologies based on their economic feasibility.
-This merit order ranks all CC technologies according to their relative costs to capture a ton of carbon (see Figure~X).
+The concept of a ``merit order for captured carbon`` plays a pivotal role in optimizing and prioritizing the deployment of carbon capture technologies based on their economic feasibility.
+This merit order ranks all CC technologies according to their relative costs to capture a ton of carbon (see Fig.~X).
 At the lower end of the cost spectrum, CC technologies applied to process emissions, such as those from cement, offer a cost-effective starting point.
@@
 
-Addressing climate targets, we define three \carbon{} target scenarios:
+Addressing climate targets, we define two \carbon{} target scenarios:
 \begin{itemize}
-    \item[] \textit{Net-Zero}: aligning with the EU's 2050 emission targets. Also denoted as NZ.
-    \item[] \textit{Net-Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \carbon{} annually. Also denoted as NN.
+    \item[] \textit{Net--Zero}: aligning with the EU's 2050 emission targets. Also denoted as NZ.
+    \item[] \textit{Net--Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \carbon{} annually. Also denoted as NN.
 \end{itemize}
@@
 
-Beyond \carbon{} targets, we consider four expansion scenarios:
+Beyond \carbon{} targets, we consider four expansion scenarios, referred to as models:
 \begin{itemize}
     \item[] \textit{Baseline:} Neither \carbon{} nor \hydrogen{} networks are constructed.
-    \item[] \textit{\hydrogen{}~Grid:} Only the \hydrogen{} network is developed.
-    \item[] \textit{\carbon{}~Grid:} Only the \carbon{} network is developed.
+    \item[] \textit{\hydrogengrid:} Only the \hydrogen{} network is developed.
+    \item[] \textit{\carbongrid:} Only the \carbon{} network is developed.
     \item[] \textit{Hybrid:} Both \carbon{} and \hydrogen{} networks are developed.
@@
 
-Focusing first on the NZ scenarios, we can highlight significant variations in system costs and technology deployment depending on the network strategy employed (see Figure~\ref{fig:cost_bar}).
+Focusing first on the net-zero scenario, we can highlight significant variations in system costs and technology deployment across the models (see Fig.~\ref{fig:cost_bar}).
 
@@
 The implementation of a \carbon{} transport network (\textit{\carbon{}~Grid}) facilitates efficient point-source carbon capture, substantially reducing the dependency on DAC technologies.
-While DAC supplies around 50\% of the carbon needed in the Baseline scenario (400 Mt \carbon{}, see Figure~\ref{fig:balance_captured_carbon}), in the \scCO scenario, DAC is only used to supply around 20\% of the carbon (150 Mt \carbon{}).
-The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Figure~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology is relatively cheaper than DAC.
+While DAC supplies around 50\% of the carbon needed in the Baseline scenario (400 Mt \carbon{}, see Fig.~\ref{fig:balance_captured_carbon}), in the \scCO scenario, DAC is only used to supply around 20\% of the carbon (150 Mt \carbon{}).
+The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Fig.~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology is relatively cheaper than DAC.
 The \carbon{} transport network enables the transport of carbon from the locations of the cheapest captured carbon supply sources to the locations of the carbon demand sinks at CU amd CS sites, reducing the need for expensive DAC.
@@
 While the cost-reduction in the Hybrid scenario compared to the \hydrogen~Grid scenario is marginal, the allocation of costs reveals a propagation of reduced DAC facilities and increased CC at point sources.
-The Hybrid scenario relies much heavier on investments into the \hydrogen network which outweigh investments in the \carbon{} network by a factor of 4.
+The Hybrid scenario relies much heavier on investments into the \hydrogen{} network which outweigh investments in the \carbon{} network by a factor of 4.
 
-%
-
-\begin{figure*}
+\begin{figure*}[h]
     \centering
@@
         \centering
-        \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_carbon.png}
+        \includegraphics[width=0.95\linewidth]{co2-only/figures/90_nodes/operation_map_carbon.png}
         \caption{\carbon{} Sector.}
@@
         \centering
-        \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_hydrogen.png}
+        \includegraphics[width=0.95\linewidth]{h2-only/figures/90_nodes/operation_map_hydrogen.png}
         \caption{\hydrogen{} Sector.}
@@
     \end{subfigure}
-    \caption{Optimal operation per sector for a net-zero energy system in Europe with both \carbon{} and \hydrogen{} network expansion (Hybrid).}
-    \label{fig:operation_map_carbon_co2}
+    \caption{Optimal operation per sector for a net-zero energy system in Europe with average production on the left and average consumption on the right for both, (a) the \carbon sector in the \carbon{}-Grid model and (b) the \hydrogen{} sector in the \hydrogen-Grid model.}
+    \label{fig:operation_map}
 \end{figure*}
+
+\subsection*{}
+
+Examining the cost-reductions across the different models raises an important question: why is the hydrogen network more cost-effective than its carbon counterpart, and how does it achieve similar cost reductions as the hybrid model?
+To answer this question, it is crucial to understand the underlying mechanisms and geographical implications of the two transportation systems.
+Fig.~\ref{fig:operation_map_carbon} maps the average production, consumption and flow in the carbon sector for the \carbongrid model.
+Fig.
+\ref{fig:operation_map_hydrogen} shows equivalent quantities in the hydrogen sector for the \hydrogengrid model.
+In each region, circles divide into contributions from technologies, and their sizes correspond proportionally to production on the left and consumption on the right.
+The region's colors represent the average price of the considered carrier.
+The two fundamental differences between the grid strategies are quickly evident: The \carbongrid model captures large amounts of carbon from bioenergy and industrial processes, especially in central and eastern Europe, at low cost.
+The carbon grid transports cheap carbon to regions where it is sequestered in offshore caverns or processed into synfuel, methane and methanol together with cheaply available green hydrogen.
+End-use primarily depends on the quality of renewable resources.
+In Italy, Greece, North Germany, and Denmark, the transported carbon is predominantly sequestered and reused to some extent.
+Conversely, in Spain and the United Kingdom, which both have high renewable potentials, the transported carbon is fully utilized, complementing air-captured carbon.
+Thus, the implementation of a \carbon network allows for connecting regions with abundant, low-price carbon to regions with carbon scarcity where in the absence of a \carbon network DAC facilities had to provide most carbon.
+% Sequestered carbon is nearly always transported before
+
+The \hydrogengrid on the other hand, transports cheap hydrogen from regions with abundant renewable resources to regions with high energy demand and abundant emissions.
+The main transportation routes go from Spain and the United Kingdom to Central Europe.
+In these regions, industries use the hydrogen to fuel processes and land transport.
+Additionally, the imported hydrogen allows for the direct reuse of local \carbon emissions in synfuel production, leading to decentralized synfuel production across the continent.
+Hydrogen produced in Italy contributes to local methane production, which then travels to Central Europe via gas pipelines to supply CHPs and industries."
+
+% Therefore, one can conclude that a \carbon network favors a decentralized CC and a CU centralized at regions with abundant renewable resources.
+A \hydrogen network favors a centralized electrolysis system at regions with abundant renewable resources and a decentralized CU system.
+These findings highlight the pivotal advantage of the \hydrogen network in contrast to the \carbon network.
+Central Europe with its industrial clusters and dense population is characterized by a scarcity in energy and a surplus and abundance of \carbon.
+While the \carbon network cost-efficiently resolves the surplus of carbon, the hydrogen resolves both the energy scarcity and the carbon surplus.
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics*[width=0.8\linewidth]{comparison/emission-reduction/figures/90_nodes/objective_heatmap.png}
-    \caption{Total annual system cost for the different scenarios and \carbon{} targets.
-While the `Baseline' scenario has neither a \hydrogen{} or \carbon{} network, `Hybrid' is allowed to expand both.}
-    \label{fig:objective_heatmap}
-\end{figure}
+As depicted in Fig.~\ref{fig:capacity_maps}, the hybrid model incorporates the advantages from the a large hydrogen network and complements it with a carbon network.
+Mirroring the behavior of \hydrogengrid model, it transports hydrogen from Western regions like Spain and the UK to Central Europe, with a centralized electrolysis and decentralized carbon utilization.
+The carbon network complements this by transporting cheap carbon to sequestration sites in the North Sea, Baltic Sea, Greece, and Italy, minimizing the need for Direct Air Capture (DAC) at these locations.
+A notable aspect of the hybrid model is the lack of overlap in the routes used by the hydrogen and carbon networks.
+The advantages of this model include a broader range of technologies contributing to its robustness, reduced reliance on DAC (which may be costlier than anticipated), and less land use for wind and solar due to the decreased necessity for DAC.
+
+% extend on flexibility gains
+
+\subsection*{Net-Negative Scenarios}
 
@@
     \centering
-    \includegraphics[width=\linewidth]{comparison/emission-reduction-full/figures/90_nodes/transmission_cost_bar.png}
-    \caption{Annual transmission system cost as a function of the net carbon removal scenarios considered in the study.}
-    \label{fig:transmission_cost_bar}
+    \includegraphics[width=0.9\linewidth]{difference/comparison/emission-reduction-0.1/figures/90_nodes/cost_bar.png}
+    \caption[short]{Net change in investments when tightening the \carbon{} emission target from net-zero to net-negative 10\% of 1990s emissions.}
+    \label{fig:net-negative_cost_bar}
 \end{figure}
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
-    \caption{Sankey diagram of the optimal operation for a net-negative 10\% scenario.}
-    \label{fig:sankey_diagramm}
-\end{figure}
+We now examine the case of moving from a net zero emissions target to a net emissions reduction target.
+Fig.~\ref{fig:net-negative_cost_bar} illustrates the shift in investments for the \hydrogengrid and the Hybrid model between the net-zero and the net-negative scenario.
 
-[Start of old version]
-First, we analyze the optimal deployment of technologies and their associated costs at different levels of carbon sequestration potential.
-Figure~\ref{fig:cost_area} displays the annual system cost per technology group for the sector-coupled model at various sequestration levels, both with and without a \carbon{} network.
-The costs are calculated as the sum of the annualized investment and operational costs for all technologies.
-Starting from the left, we observe that the majority of system costs are associated with the deployment of renewable energy sources, heating, hydrogen technologies, and the electricity grid.
-Costs for fossil fuels and carbon capture and transport technologies are relatively low.
-In terms of capacities, 200~GW of offshore wind, 3930~GW of solar PV, 1910~GW of onshore wind, and 1650~GW of electrolysis are installed in this setup.
-In both cases, with and without \carbon{} transport, total costs decrease as the sequestration potential increases.
-At high sequestration rates, fossil fuels such as gas and oil are kept in the system.
-This reduces the need for synthetic FT fuel, which requires significant renewable energy sources.
-Simultaneously, the system deploys more \carbon{} removal technologies, such as DAC and Bioenergy with carbon capture (included in "Carbon Capture Technologies"), to offset emissions from fossil fuels.
+The \hydrogengrid model emphasizes a substantial investment in hydrogen infrastructure, signifying a strategic bet on hydrogen's role in decarbonization.
+There is also a significant allocation for methanation and the expansion of electricity grids, underscoring the integration of hydrogen production with renewable energy sources.
 
-The cost reduction when going to high sequestration in a system with \carbon{} transport is about 13\%, which is slightly more than without a \carbon{} network (11\%).
-Across all sequestration levels, a system with \carbon{} transport is, on average, 1.6\% less expensive.
-The cost savings for a carbon transport system are lowest at 0.5\% (4 billion euros per year) for a sequestration of 200~Mt and highest at 3\% (21 billion euros per year) for a sequestration of 1000~Mt.
-
-To further investigate the switch in technologies, we analyze the annual operation of the system for different sequestration potentials.
-Figures~\ref{fig:operation_area_carbon_co2network}, \ref{fig:operation_area_hydrogen_co2network} and \ref{fig:operation_area_co2_co2network} show the balances of captured carbon, hydrogen and atmospheric \carbon{} in the system with \carbon{} transport as a function of the sequestration potential respectively.
-The areas in the positive segment indicate gross production of the respective commodity, while the areas in the negative segment indicate gross consumption.
-Note that technologies with a ``*`` are explicitly augmented technologies with integrated carbon capture.
-%
-
-Figure~\ref{fig:operation_area_carbon_co2network} confirms our previous results and shows that FT synthesis production decreases with sequestration availability.
-At the same time, integrated carbon capture from Gas Combined Heat and Power (CHP) plants, Allam Cycle plants, Steam Methane Reforming (SMR) and DAC increases.
-Other technologies are constant across varying levels of sequestration due to different reasons: biomass is bound by its maximum available potential (as reported in~\cite{enspreso_database}); the industry processes with their associated emissions are set to fixed values; the methanol production is bound by a lower limit representing the demand from the shipping sector.
-
-% TODO: kerosine emissions from fossil based kerosine are not taken into account by the system.
-Figure~\ref{fig:operation_area_hydrogen_co2network} shows that hydrogen production decreases by about 50\% when going from 200~Mt/a sequestration to 1000~Mt/a.
-In the last segment, hydrogen production from SMR starts to replace hydrogen production from electrolysis.
-Correspondingly, hydrogen demand for FT synthetic fuel production disappears, leaving only hydrogen demand for the shipping, industry and transport sector.
-Finally, Figure~\ref{fig:operation_area_co2_co2network} shows that the balance of the atmospheric \carbon{}.
-Note that the absolute value in the atmospheric \carbon{} hardly changes with the sequestration potential.
-This is because the switch from FT to fossil fuels is mirrored by a switch from using captured \carbon{} from biomass and DAC for FT synthesis to sequestering it underground.
-The net increase in DAC is solely due to an increase in unabated emissions from gas boilers.
-
-We can conclude that in combination with a \carbon{} transport network, high rates of sequestration reduce the feasibility of CCU and hydrogen infrastructure and increases the feasibility of CCS, namely DAC and CC at point sources.
-
-To answer the question of how carbon capture technologies at point sources change with the sequestration potential, we analyze the shift to carbon capture integration for all relevant technologies.
-Figure~\ref{fig:captureshare_line} shows the share of a given technology that is equipped with integrated carbon capture, i.e., carbon capture at point source, for both with and without \carbon{} transport network.
-With \carbon{} transport, all technologies except for SMR and Gas CHP are nearly fully equipped with CC across all sequestration potentials.
-As a result of sequestration, the share of SMR with carbon capture rises from 50\% to 100\%, and that of CHP from 0\% to 26\%.
-The technologies behave differently without carbon transport: besides the fact that capturing at point source is less attractive overall, their carbon capture shares rise more slowly as sequestration increases.
-The carbon capture share for process emissions even decreases.
-This is primarily due to the fact that as sequestration increases, the system relies more heavily on DAC and SMR with CC that can be flexibly placed near the sequestration sites.
-
-Building on the analysis of optimal operation for varying sequestration potentials, we explore the spatial distribution of carbon capture and transport technologies for the most "optimistic" case of 1000~Mt/a sequestration potential and a carbon transport network.
-
-Figure~\ref{fig:operation_map_noco2network_1000} shows a map of the total production and consumption of captured carbon in the system.
-The left side shows the combined carbon supply from carbon capture facilities and the carbon network flow.
-The right side illustrates regional carbon usage, again, together with the flow.
-Large capacity DACs and SMRs are placed close to shore with carbon capture in order to maximize efficiency.
-Significant facilities for DAC are built in the British Isles where carbon dioxide is used for methanol synthesis and also sequestration.
-Here substantial onshore wind farms generate a total annual production of approximately 400 TWh per year, of which 13\% is utilized by local DAC facilities.
-Other notable DAC installations are located on the Atlantic coast of Portugal, Spain, Ireland, and Denmark.
-Large facilities of SMR with CC are found in Portugal, South England and Greece.
-All these locations are in close proximity to sequestration sites, as indicated in Figure~\ref{fig:sequestration_map}.
-Note that the installation of SMR facilities close to the shore has the additional benefit of enabling easy transportation of hydrogen to ports for the shipping sector.
-
-Looking at the transport of \carbon{} across the system, we observe that the carbon network primarily transports captured carbon from process and biomass emissions from the inland to sequestration areas.
-Note that optimization installs a number of separate carbon transport networks instead of building one large connected grid.
-Therefore, a large network cluster transports carbon from Central Europe to the North Sea.
-In contrast, three smaller clusters in the south of Europe transport carbon to the Atlantic and the Mediterranean Sea.
-
-As shown on the right, methanol production using captured carbon is primarily located in the United Kingdom, Ireland, and South Spain.
-
-Finally, the deployment of carbon capture technologies in a system with high sequestration rates contrasts with the deployment of CCU facilities in a system with low carbon sequestration.
-On the one hand, with high sequestration, carbon capture management in the system is mainly driven by the location of the sequestration sites, leading to a spatially centralized deployment of carbon capture technologies.
-On the other hand, with low sequestration rates the use of CCU in a system is distributed more evenly across the system.
+% The Hybrid model, in contrast, adopts a more balanced investment approach, with notable allocations not just to hydrogen infrastructure but also to wind, solar, and CO2 infrastructure, as well as direct air capture (DAC) technology.
+This suggests a more comprehensive strategy that leverages a variety of technologies and energy sources to achieve emission reductions.
+The distributed investment in the Hybrid model indicates a systemic approach to decarbonization, integrating renewable generation, energy storage, carbon capture, and hydrogen production to address both energy and carbon management challenges.
 
@@
 
-% \appendix
+\appendix
 
-% \begin{figure}[h]
+% \begin{figure*}
 %     \centering
-%     \includegraphics[width=\linewidth]{capacity_map_electricity_co2network_1000.png}
-%     \caption{Optimal capacities per of the electricity sector for a sequestration of 1000 Mt/a.}
-%     \label{fig:capacity_map_electricity_co2network_1000}
+%     \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
+%     \caption{Sankey diagram of the optimal operation for a net-negative 10\% scenario.}
+%     \label{fig:sankey_diagramm}
+% \end{figure*}
+
+% \begin{figure}
+%     \centering
+%     \includegraphics*[width=0.8\linewidth]{comparison/emission-reduction/figures/90_nodes/objective_heatmap.png}
+%     \caption{Total annual system cost for the different scenarios and \carbon{} targets.
+While the `Baseline' scenario has neither a \hydrogen{} or \carbon{} network, `Hybrid' is allowed to expand both.}
+%     \label{fig:objective_heatmap}
 % \end{figure}
@@
 %     \centering
-%     \includegraphics[width=\linewidth]{capacity_map_hydrogen_co2network_1000.png}
-%     \caption{Optimal capacities per of the hydrogen sector for a sequestration of 1000 Mt/a.}
-%     \label{fig:capacity_map_hydrogen_co2network_1000}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[width=\linewidth]{operation_area_carbon_noco2network.pdf}
-%     \caption{Balance of captured \carbon{} emissions for the optimal operation without \carbon{} network.}
-%     \label{fig:operation_area_carbon_noco2network}
+%     \includegraphics[width=\linewidth]{comparison/emission-reduction-full/figures/90_nodes/transmission_cost_bar.png}
+%     \caption{Annual transmission system cost as a function of the net carbon removal scenarios considered in the study.}
+%     \label{fig:transmission_cost_bar}
 % \end{figure}

### 2023-11-22 workflow/scripts/plot-cost-compare-diff-bar.py

@@ -6,7 +6,8 @@ import seaborn as sns
 from common import (
     import_network,
     mock_snakemake,
-    sort_rows_by_relative_diff,
+    sort_rows_by_diff,
+    get_ordered_handles_labels,
 )
 
 alpha = 1
@@ -50,10 +51,10 @@ unit = "bn€/a"
 
 groups = snakemake.config["plotting"]["technology_groups"]
 groups = {v: groups[k] for k, v in n.carriers.nice_name.drop("").items()}
-grouped = diff.groupby(groups).sum().div(norm)
+grouped = diff.mul(-1).groupby(groups).sum().div(norm)
 # grouped = grouped[grouped.max(axis=1) > 5e4]
 colors = snakemake.config["plotting"]["technology_group_colors"]
-grouped = sort_rows_by_relative_diff(grouped)[::-1]
+grouped = sort_rows_by_diff(grouped)
 rename = {"Carbon Capt. at Point Sources": "Carbon Capture\nat Point Sources"}
 grouped = grouped.rename(rename, axis=0)
 colors = {rename.get(k, k): v for k, v in colors.items()}
@@ -62,9 +63,9 @@ fig, ax = plt.subplots(
     1, 1, figsize=snakemake.params.settings["figsize"], layout="constrained"
 )
 
-grouped = grouped[grouped.round(0).ne(0).any(axis=1)].sort_values(by=grouped.columns[0])
-grouped.T.mul(-1).plot(
-    kind="bar", stacked=True, ax=ax, color=colors, lw=0, legend=False, alpha=0.8
+grouped = grouped[grouped.round(0).ne(0).any(axis=1)]
+grouped.T.plot(
+    kind="bar", stacked=True, ax=ax, color=colors, lw=0, legend=True, alpha=0.8
 )
 
 # ax.vlines(0, -0.5, len(diff) - 0.5, color="k", lw=0.5)
@@ -72,11 +73,10 @@ grouped.T.mul(-1).plot(
 ax.set_ylabel(f"Net Investment Change [{unit}]")
 ax.set_xlabel("")
 ax.grid(axis="y", alpha=0.5)
-handles, labels = ax.get_legend_handles_labels()
+
 ax.legend().remove()
 ax.legend(
-    handles,
-    labels,
+    *get_ordered_handles_labels(ax, grouped),
     loc="center left",
     bbox_to_anchor=(1, 0.45),
     frameon=False,

### 2023-11-23 report/submissions/journal/paper/paper.tex
add subsections
review and add to subsections 1 and 2 in Results

@@
 
+\section{Introduction}
+
 An integrated management of both hydrogen and carbon is considered a crucial component to achieve climate neutrality in energy systems including hard-to-abate sectors.
@@
 Hydrogen can be gained from electrolysis and steam methane reforming (SMR).
-Geological distribution and potentials for \hydrogen storage in salt caverns are derived from~\cite{caglayanTechnicalPotentialSalt2020}.
+Geological distribution and potentials for \hydrogen{} storage in salt caverns are derived from~\cite{caglayanTechnicalPotentialSalt2020}.
 Re-electrification of hydrogen is possible via fuel cells.
 If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limit.
-The topology of the \hydrogen network is mirroring the topology of the electricity network.
+The topology of the \hydrogen{} network is mirroring the topology of the electricity network.
 
@@
     \end{subfigure}
-    \caption{Optimal operation per sector for a net-zero energy system in Europe with average production on the left and average consumption on the right for both, (a) the \carbon sector in the \carbon{}-Grid model and (b) the \hydrogen{} sector in the \hydrogen-Grid model.}
+    \caption{Optimal operation per sector for a net-zero energy system in Europe with average production on the left and average consumption on the right for both, (a) the \carbon{} sector in the \carbon{}-Grid model and (b) the \hydrogen{} sector in the \hydrogen-Grid model.}
     \label{fig:operation_map}
@@
 
-\subsection*{}
+\subsection*{Superiority of \hydrogen{} network}
 
@@
 To answer this question, it is crucial to understand the underlying mechanisms and geographical implications of the two transportation systems.
-Fig.~\ref{fig:operation_map_carbon} maps the average production, consumption and flow in the carbon sector for the \carbongrid model.
+Fig.~\ref{fig:operation_map_carbon} maps the average production, consumption and flow in the carbon sector for the \carbongrid{} model.
 Fig.
-\ref{fig:operation_map_hydrogen} shows equivalent quantities in the hydrogen sector for the \hydrogengrid model.
+\ref{fig:operation_map_hydrogen} shows equivalent quantities in the hydrogen sector for the \hydrogengrid{} model.
 In each region, circles divide into contributions from technologies, and their sizes correspond proportionally to production on the left and consumption on the right.
 The region's colors represent the average price of the considered carrier.
-The two fundamental differences between the grid strategies are quickly evident: The \carbongrid model captures large amounts of carbon from bioenergy and industrial processes, especially in central and eastern Europe, at low cost.
+
+The two fundamental differences between the grid strategies are quickly evident: The \carbongrid{} model captures large amounts of carbon from bioenergy and industrial processes, especially in central and eastern Europe, at low cost.
 The carbon grid transports cheap carbon to regions where it is sequestered in offshore caverns or processed into synfuel, methane and methanol together with cheaply available green hydrogen.
-End-use primarily depends on the quality of renewable resources.
-In Italy, Greece, North Germany, and Denmark, the transported carbon is predominantly sequestered and reused to some extent.
-Conversely, in Spain and the United Kingdom, which both have high renewable potentials, the transported carbon is fully utilized, complementing air-captured carbon.
-Thus, the implementation of a \carbon network allows for connecting regions with abundant, low-price carbon to regions with carbon scarcity where in the absence of a \carbon network DAC facilities had to provide most carbon.
+End-use primarily depends on the quality of renewable resources, and thus the availability of cheap hydrogen.
+In locations with relatively poor renewables and thus higher costs of hydrogen production, such as Italy, Greece, Northern Germany, and lcoations around the Baltic Sea, the transported carbon is predominantly sequestered.
+Conversely, in Spain and the United Kingdom, which both have high renewable potentials and cheap hydrogen production costs, the transported carbon is combined with hydrogen and fed into CU processes, with additional carbon being captured via DAC also running on the great renewable resources.
+
+In summary, the implementation of a \carbon{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce cheap hydrogen to make CU products (FT fuel, methanol, synthetic methane).
+In the absence of a \carbon{} network, sequestration sites first sequester all carbon from cheap point sources, and deploy additional DAC facilities at the locations with cheap renewables to capture the remaining necessary carbon for sequestration.
+Then, synfuel demand is met at the locations which offer the best mix of cheap carbon from point sources and cheap hydrogen from local renewables.
 % Sequestered carbon is nearly always transported before
 
-The \hydrogengrid on the other hand, transports cheap hydrogen from regions with abundant renewable resources to regions with high energy demand and abundant emissions.
+The \hydrogengrid{} on the other hand, transports cheap hydrogen from regions with abundant renewable resources to regions with high energy demand and abundant emissions.
 The main transportation routes go from Spain and the United Kingdom to Central Europe.
-In these regions, industries use the hydrogen to fuel processes and land transport.
-Additionally, the imported hydrogen allows for the direct reuse of local \carbon emissions in synfuel production, leading to decentralized synfuel production across the continent.
-Hydrogen produced in Italy contributes to local methane production, which then travels to Central Europe via gas pipelines to supply CHPs and industries."
+In these regions, industries use the hydrogen for demand in industry and fuel cell land transport, and to produce synfuels with locally captured carbon from point sources.
+This leads to a very decentralized synfuel production across the continent.
+Furthermore, the introduction of a hydrogen network enables to reduce both DAC production and FT fuels on the British Isles and Iberian Peninsula.
+In Great Britain an Ireland, the remaining local carbon is then sequestered, and all the hydrogen that previously went into FT fuel production is exported to Central Europe (see above).
+In Spain and Portugal, some DAC remains and the carbon sourced from DAC is used, along with locally captured carbon from point sources, to produce FT fuels (but less than without a hydrogen network).
+Surplus hydrogen from lower FT fuel production is exported to Central Europe.
 
-% Therefore, one can conclude that a \carbon network favors a decentralized CC and a CU centralized at regions with abundant renewable resources.
-A \hydrogen network favors a centralized electrolysis system at regions with abundant renewable resources and a decentralized CU system.
-These findings highlight the pivotal advantage of the \hydrogen network in contrast to the \carbon network.
-Central Europe with its industrial clusters and dense population is characterized by a scarcity in energy and a surplus and abundance of \carbon.
-While the \carbon network cost-efficiently resolves the surplus of carbon, the hydrogen resolves both the energy scarcity and the carbon surplus.
+% Therefore, one can conclude that a \carbon{} network favors a decentralized CC and a CU centralized at regions with abundant renewable resources.
+A \hydrogen{} network favors a centralized electrolysis system at regions with abundant renewable resources and a decentralized CU system.
+These findings explain the advantage of the \hydrogen{} network in contrast to the \carbon{} network.
+Central Europe with its industrial clusters and dense population is characterized by a scarcity in hydrogen and a surplus and abundance of \carbon.
+While the \carbon{} network cost-efficiently resolves the surplus of carbon in Central Europe by exporting it to regions with cheap renewable resources to produce synfuels, it needs to supply direct hydrogen demand in Central Europe (for industry and fuel cell land transport) with expensive local hydrogen production.
+The hydrogen network resolves both the energy scarcity by supplying hydrogen demand in Central Europe with cheap hydrogen imported from Spain and Great Britain, and the carbon surplus by transporting hydrogen to the sites with cheaper carbon available in Central Europe, and producing FT fuels highly decentralized throughout all of Central Europe.
 
@@
 
+\subsection*{Flexibility gains through hybrid configuration}
+
 As depicted in Fig.~\ref{fig:capacity_maps}, the hybrid model incorporates the advantages from the a large hydrogen network and complements it with a carbon network.
-Mirroring the behavior of \hydrogengrid model, it transports hydrogen from Western regions like Spain and the UK to Central Europe, with a centralized electrolysis and decentralized carbon utilization.
-The carbon network complements this by transporting cheap carbon to sequestration sites in the North Sea, Baltic Sea, Greece, and Italy, minimizing the need for Direct Air Capture (DAC) at these locations.
+The carbon network unlocks cheap carbon in landlocked locations like  Poland, Bulgaria, and Austria and transports it to sequestration sites in the North Sea, Baltic Sea, Greece, and Italy, minimizing the need for DAC directly at the sequestration sites.
+At the same time, there is reduced need to transport hydrogen from Western regions like Spain and the UK to Central Europe.
+
 A notable aspect of the hybrid model is the lack of overlap in the routes used by the hydrogen and carbon networks.
 The advantages of this model include a broader range of technologies contributing to its robustness, reduced reliance on DAC (which may be costlier than anticipated), and less land use for wind and solar due to the decreased necessity for DAC.
+
+In summary, sequestration moves to sites with or close to the cheapest carbon sources.
+Adding the carbon network unlocks cheap inland carbon point sources to be transported to sequestration.
+FT fuel production moves to sites with or close to cheap hydrogen.
+The model prefers sites with locally available carbon from point sources and transport hydrogen from nearby cheap hydrogen producing locations, rather than transporting the carbon to the hydrogen site.
+
+Furthermore, there are qualitative advantages to building out both a hydrogen and a carbon network that are not captured in the strict cost-benefit calculation employed in our optimization model.
+The resulting system may be more resilient towards unknown future developments in technologies, geopolitics, or policy goals.
 
@@
 We now examine the case of moving from a net zero emissions target to a net emissions reduction target.
-Fig.~\ref{fig:net-negative_cost_bar} illustrates the shift in investments for the \hydrogengrid and the Hybrid model between the net-zero and the net-negative scenario.
+Fig.~\ref{fig:net-negative_cost_bar} illustrates the shift in investments for the \hydrogengrid{} and the Hybrid model between the net-zero and the net-negative scenario.
 
-The \hydrogengrid model emphasizes a substantial investment in hydrogen infrastructure, signifying a strategic bet on hydrogen's role in decarbonization.
+The \hydrogengrid{} model emphasizes a substantial investment in hydrogen infrastructure, signifying a strategic bet on hydrogen's role in decarbonization.
 There is also a significant allocation for methanation and the expansion of electricity grids, underscoring the integration of hydrogen production with renewable energy sources.

### 2023-11-23 report/submissions/journal/paper/paper.tex
review initial Results section

@@
 \newcommand{\carbongrid}{\carbon{}--Grid}
-\newcommand{\scbase}{Baseline scenario}
-\newcommand{\scCO}{CO$_2$~Grid scenario}
-\newcommand{\scH}{H$_2$~Grid scenario}
-\newcommand{\schybrid}{Hybrid scenario}
+\newcommand{\modBase}{Baseline model}
+\newcommand{\modCO}{CO$_2$-Grid model}
+\newcommand{\modH}{H$_2$-Grid model}
+\newcommand{\modHybrid}{Hybrid model}
 
@@
     \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/cost_bar.png}
-    \caption[short]{Total annual system cost for the European energy system for the different scenarios subdivided into groups of technologies.
-While the `Baseline' scenario has neither a \hydrogen{} or \carbon{} network, `Hybrid' is allowed to expand both.
+    \caption[short]{Total annual system cost for the European energy system for the different models subdivided into groups of technologies.
+While the `Baseline' model has neither a \hydrogen{} or \carbon{} network, `Hybrid' is allowed to expand both.
 "Gas Infrastructure" combines gas facilities for power and heat production, "H$_2$ Infrastructure" combines H$_2$ production, transport and re-electrification.
@@
 
-System cost differences between the scenarios are driven by renewable energy resources (solar, wind and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
-In the Baseline scenario, the absence of dedicated carbon and hydrogen transport networks requires the system to deploy carbon production technologies (DAC and CC at point sources) in the regions with carbon sequestration sites, and small amounts of hydrogen production technologies (mainly electrolysis) in each region to meet local hydrogen demand for industry and land transport fuel cells.
+System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
+In the Baseline model, the absence of dedicated carbon and hydrogen transport networks requires the system to deploy carbon production technologies (DAC and CC at point sources) in the regions with carbon sequestration sites, and small amounts of hydrogen production technologies (mainly electrolysis) in each region to meet local hydrogen demand for industry and land transport fuel cells.
 For CU and its respective carbon and hydrogen demands, the system chooses the sites where on-site production of carbon, hydrogen and the synthesis process is cheapest.
-This leads to renewable resources and storage solutions being deployed not necessarily in the regions with the best resources or lowest costs for clean electricity generation, but where they are situated closer to the energy demand locations of electrolysers and DAC.
-Thus, total system costs in the baseline scenario are highest, while the scenarios with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs.
+This leads to renewable resources and storage solutions being deployed not necessarily in the regions with the best resources or lowest costs for clean electricity generation, but where they are situated closer to the locations with carbon supply from CC at point sources, and to the locations of energy demand for electrolysers and DAC.
+Thus, total system costs in the Baseline model are highest, while the models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs.
 
-We can now closer examine the cost reductions and shifts in cost allocation when shifting from the baseline scenario to the other three network scenarios.
+We can now closer examine the cost reductions and shifts in cost allocation when shifting from the Baseline model to the other three network models.
 The implementation of a \carbon{} transport network (\textit{\carbon{}~Grid}) facilitates efficient point-source carbon capture, substantially reducing the dependency on DAC technologies.
-While DAC supplies around 50\% of the carbon needed in the Baseline scenario (400 Mt \carbon{}, see Fig.~\ref{fig:balance_captured_carbon}), in the \scCO scenario, DAC is only used to supply around 20\% of the carbon (150 Mt \carbon{}).
+While DAC supplies around 50\% of the carbon needed in the Baseline model (400 Mt \carbon{}, see Fig.~\ref{fig:balance_captured_carbon}), in the \modCO{}, DAC is only used to supply around 20\% of the carbon (150 Mt \carbon{}).
 The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Fig.~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology is relatively cheaper than DAC.
 The \carbon{} transport network enables the transport of carbon from the locations of the cheapest captured carbon supply sources to the locations of the carbon demand sinks at CU amd CS sites, reducing the need for expensive DAC.
-The improved carbon management infrastructure also allows for a more strategic deployment of renewables, particularly in areas with high sequestration potential, leading to cost savings in renewable generation and storage.
+The improved carbon management infrastructure also allows for a more strategic deployment of renewables, particularly in areas that combine high quality renewables and high sequestration potentials, leading to system-wide cost savings in renewable generation and storage.
 
@@
 
-The \hydrogen{}~Grid scenario underscores the pivotal role of hydrogen infrastructure.
+The \hydrogen{}-Grid model underscores the pivotal role of hydrogen infrastructure.
 The creation of a hydrogen transport network enables the cost-effective delivery of hydrogen from regions with a surplus of cheap renewable energy to high-demand industrial centers.
-This scenario shows a pronounced decrease in investments in solar and gas  infrastructure.
+This model shows a pronounced decrease in investments in solar generation and gas infrastructure.
 The \hydrogen{} network reduces DAC similarly to the \carbon{} network, but still produces 200 Mt of carbon from DAC.
 CC technology on process emissions increases from 90\% to 100\%, on biomass CHPs from 60\% to 100\%, and on facilities using gas in industry from 50\% to 75\%.
-CC on biomass use in industry remains at 50\%, same as the \scbase.
+CC on biomass use in industry remains at 50\%, same as the \modCO{} (see Figure~\ref{fig:captureshare_line}).
 This variation highlights how the implementation of \hydrogen{} and \carbon{} networks unlocks specific carbon capture potentials.
 
-While the cost-reduction in the Hybrid scenario compared to the \hydrogen~Grid scenario is marginal, the allocation of costs reveals a propagation of reduced DAC facilities and increased CC at point sources.
-The Hybrid scenario relies much heavier on investments into the \hydrogen{} network which outweigh investments in the \carbon{} network by a factor of 4.
+While the cost-reduction in the Hybrid model compared to the \hydrogen-Grid model is marginal, the allocation of costs reveals a propagation of reduced DAC facilities and increased CC at point sources.
+The Hybrid model relies slightly less on investments into the \hydrogen{} network than the \modH, but adds some investments into the \carbon{} network.
+Still, investments into the \hydrogen{} network outweigh investments into the \carbon{} network by a factor of four.
 
@@
 
-\begin{figure*}[h]
+\subsection*{Flexibility gains through hybrid configuration}
+
+\begin{figure*}[h!]
     \centering
-    \begin{subfigure}{.5\textwidth}
+    \begin{subfigure}{.49\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{full/figures/90_nodes/capacity_map_hydrogen.png}
+        \caption{\hydrogen{} Sector.}
+        \label{fig:capacity_map_hydrogen_co2}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{.49\textwidth}
         \centering
@@
     \end{subfigure}%
-    \hfill
-    \begin{subfigure}{.5\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{full/figures/90_nodes/capacity_map_hydrogen.png}
-        \caption{\hydrogen{} Sector.}
-        \label{fig:capacity_map_hydrogen_co2}
-    \end{subfigure}
     \caption{Optimal production and transport capacities of the carbon and hydrogen sector in a net-zero energy system in Europe with both \carbon{} and \hydrogen{} network expansion (Hybrid).
@@
 
-\subsection*{Flexibility gains through hybrid configuration}
+% increased BECCS, decreased DACS
+% decreased methanation
 
-As depicted in Fig.~\ref{fig:capacity_maps}, the hybrid model incorporates the advantages from the a large hydrogen network and complements it with a carbon network.
-The carbon network unlocks cheap carbon in landlocked locations like  Poland, Bulgaria, and Austria and transports it to sequestration sites in the North Sea, Baltic Sea, Greece, and Italy, minimizing the need for DAC directly at the sequestration sites.
-At the same time, there is reduced need to transport hydrogen from Western regions like Spain and the UK to Central Europe.
+The Hybrid model combines the advantages of both hydrogen and carbon networks.
+While the topology of the \hydrogen{}-grid roughly corresponds to that of the \hydrogen grid model, the \carbon{}-grid plays more of a complementary role.
+Fig.~\ref{fig:capacity_maps} shows the capacities of the two grids in the hybrid model, with the production and transport capacities given in kt/h.
+Note that despite similar overall capacities, the investment in the hydrogen infrastructure is four times the investment in the carbon infrastructure.
 
-A notable aspect of the hybrid model is the lack of overlap in the routes used by the hydrogen and carbon networks.
-The advantages of this model include a broader range of technologies contributing to its robustness, reduced reliance on DAC (which may be costlier than anticipated), and less land use for wind and solar due to the decreased necessity for DAC.
+Similar to the \hydrogengrid model, the Hybrid model takes advantage of a large hydrogen network that enables hydrogen to be transported from centralized production sites in western regions such as Spain and the UK to decentralized FT production sites across the continent.
+The carbon network supplements this with smaller clusters of networks.
+In addition to a larger carbon grid cluster in Central Europe, which transports carbon to the North Sea, the system is building three large linear carbon routes.
+These transport carbon from Romania and Bulgaria to Greece, from northern Italy to central Italy and from Hungary, the Czech Republic and Poland to the Baltic Sea.
+These enable the system to transport cheaply available carbon from inland biomass sources and sequester it (see Fig.~\ref{fig:captureshare_line}).
+This reduces the need for DAC in the system and the sequestration sites are shifted to the end points of the carbon networks.
+Second-order effects can be seen in the shift in FT production compared to the \hydrogengrid model.
+Regions that have both good renewable resources and sequestration potential could now forego sequestration and use local process emissions for FT production, as can be observed in Ireland.
+A more detailed of differences between the \hydrogengrid and the Hybrid model can be found in Fig.~.
 
-In summary, sequestration moves to sites with or close to the cheapest carbon sources.
-Adding the carbon network unlocks cheap inland carbon point sources to be transported to sequestration.
-FT fuel production moves to sites with or close to cheap hydrogen.
-The model prefers sites with locally available carbon from point sources and transport hydrogen from nearby cheap hydrogen producing locations, rather than transporting the carbon to the hydrogen site.
+A notable aspect of the hybrid model is the lack of overlap in the hydrogen and carbon network topologies which is tightly connected to the nodal carbon strategy.
+Depending on the exogeneous need in hydrogen for example from industry, and the costs for locally produced carbon, a region can either import hydrogen to produce FT fuels or export carbon.
+\dots
 
-Furthermore, there are qualitative advantages to building out both a hydrogen and a carbon network that are not captured in the strict cost-benefit calculation employed in our optimization model.
-The resulting system may be more resilient towards unknown future developments in technologies, geopolitics, or policy goals.
-
-% extend on flexibility gains
+Despite only small decrease in cost in comparison to the \hydrogengrid model, hybrid approach brings several advantages.
+These include a broader range of technologies contributing to its robustness, reduced reliance on DAC, which may be costlier than anticipated in the cost projections, and less land use for wind and solar due to the decreased necessity for DAC.
+The increase used of biomass with carbon capture, transport and sequestration leads to slight cost advantages in comparison to the \hydrogengrid model.
 
@@
 
-The \hydrogengrid{} model emphasizes a substantial investment in hydrogen infrastructure, signifying a strategic bet on hydrogen's role in decarbonization.
-There is also a significant allocation for methanation and the expansion of electricity grids, underscoring the integration of hydrogen production with renewable energy sources.
+Both systems show a stron increase in DAC
 
-% The Hybrid model, in contrast, adopts a more balanced investment approach, with notable allocations not just to hydrogen infrastructure but also to wind, solar, and CO2 infrastructure, as well as direct air capture (DAC) technology.
-This suggests a more comprehensive strategy that leverages a variety of technologies and energy sources to achieve emission reductions.
-The distributed investment in the Hybrid model indicates a systemic approach to decarbonization, integrating renewable generation, energy storage, carbon capture, and hydrogen production to address both energy and carbon management challenges.
+\section{Conclusion}
+\label{sec:conclusion}
 
-\section{Limitations}
-\label{sec:limitations}
+In this study\dots
+
+\dots
 
@@
 These learning rates are derived from historical data, which may not necessarily be indicative of future trends.
-The model does not account for uncertainties arising from disruptive market behavior, such as the gas price peak that occurred in 2022.
+In particular, cost assumptions on DAC, electrolysis and \hydrogen{} pipeline costs, could have a strong impact in the model results.
+
+The modelling results are heavily driven by the demand and emissions from industrial clusters.
+Allowing the model to relocate these and/or incorporate flexibility measures, may lead to less dependency on both \hydrogen{} and \carbon{} networks.
+
+The technological flexibilities in the carbon sector might not be exploited to full extent.
+Therefore, the \carbon{} transport via truck or ship is not considered.
+Fischer-Tropsch facilities require to run on baseload with at least 90\% of their nominal capacity.
+
+\dots
 
@@
 
-Lastly, the model's spatial and temporal resolution is insufficient to capture all relevant dynamics.
-Variations in renewable energy supply and energy demand below the applied time resolution, as well as more detailed interregional energy transport constraints, are currently not considered in the optimization.
-Transitioning to an hourly representation and a higher spatial resolution would increase the model's computational time and complexity, but would also enhance its robustness and validity.
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-In this study, we investigated the impact of a \carbon{} transport system with different carbon sequestration potentials on the optimal configuration of technologies in a fully sector-coupled energy system in Europe, focusing on the deployment of carbon capture and Utilization (CCU) and Carbon Capture and Storage (CCS) technologies.
-Our results indicate that increasing carbon sequestration potential leads to a decrease in total system costs, with a more pronounced cost reduction observed when a carbon transport network is present.
-Higher sequestration rates result in a shift from CCU and hydrogen infrastructure towards CCS and unabated fossil fuel usage compensated by carbon dioxide removal elsewhere in the system, particularly based on direct air and biomass carbon capture with sequestration.
-
-Furthermore, we observed that with a \carbon{} transport network, the system tends to build out more carbon capture at point sources.
-With an increased sequestration potential, the share of technologies equipped with integrated carbon capture rises, particularly for Steam Methane Reforming (SMR) and Gas Combined Heat and Power (CHP) plants.
-Independent of a setup with or without \carbon{} transport, the spatial distribution of carbon capture technologies becomes more centralized, with installations concentrated near sequestration sites.
-
-Despite limitations of the study considering fixed model assumptions and cost uncertainties, our findings contribute valuable insights into the role of a carbon transport system and carbon sequestration potential in the transition towards a climate-neutral and net-negative energy system.
-As Europe strives to achieve its 2050 climate neutrality target, understanding the interplay between carbon sequestration, CCU, and CCS technologies will be crucial for the design of effective policies and strategies.
-
 \printbibliography
@@
 
-% \begin{figure*}
-%     \centering
-%     \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
-%     \caption{Sankey diagram of the optimal operation for a net-negative 10\% scenario.}
-%     \label{fig:sankey_diagramm}
-% \end{figure*}
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
+    \caption{Sankey diagram of the optimal operation for a net-negative 10\% scenario.}
+    \label{fig:sankey_diagramm}
+\end{figure*}
 
-% \begin{figure}
-%     \centering
-%     \includegraphics*[width=0.8\linewidth]{comparison/emission-reduction/figures/90_nodes/objective_heatmap.png}
-%     \caption{Total annual system cost for the different scenarios and \carbon{} targets.
+\begin{figure}
+    \centering
+    \includegraphics*[width=0.8\linewidth]{comparison/emission-reduction/figures/90_nodes/objective_heatmap.png}
+    \caption{Total annual system cost for the different scenarios and \carbon{} targets.
 While the `Baseline' scenario has neither a \hydrogen{} or \carbon{} network, `Hybrid' is allowed to expand both.}
-%     \label{fig:objective_heatmap}
-% \end{figure}
+    \label{fig:objective_heatmap}
+\end{figure}
 
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=\linewidth]{comparison/emission-reduction-full/figures/90_nodes/transmission_cost_bar.png}
-%     \caption{Annual transmission system cost as a function of the net carbon removal scenarios considered in the study.}
-%     \label{fig:transmission_cost_bar}
-% \end{figure}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{comparison/emission-reduction-full/figures/90_nodes/transmission_cost_bar.png}
+    \caption{Annual transmission system cost as a function of the net carbon removal scenarios considered in the study.}
+    \label{fig:transmission_cost_bar}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{difference/h2-only-full/figures/90_nodes/cost_map.png}
+    \caption{Difference in cost investments between \hydrogengrid and Hybrid model.
+The left subfigure shows higher spendings per technology and region and transport system for the \hydrogengrid model, the right shows higher spendings in the Hydrid model.}
+    \label{fig:cost_map_difference}
+\end{figure}
 

### 2023-11-23 report/submissions/journal/paper/paper.tex
continue writing on section 2.2 adjust capacity map colormap label

@@
 \newcommand{\carbongrid}{\carbon{}--Grid}
-\newcommand{\modBase}{Baseline model}
-\newcommand{\modCO}{CO$_2$-Grid model}
-\newcommand{\modH}{H$_2$-Grid model}
-\newcommand{\modHybrid}{Hybrid model}
+\newcommand{\scbase}{Baseline scenario}
+\newcommand{\scCO}{CO$_2$~Grid scenario}
+\newcommand{\scH}{H$_2$~Grid scenario}
+\newcommand{\schybrid}{Hybrid scenario}
 
@@
     \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/cost_bar.png}
-    \caption[short]{Total annual system cost for the European energy system for the different models subdivided into groups of technologies.
-While the `Baseline' model has neither a \hydrogen{} or \carbon{} network, `Hybrid' is allowed to expand both.
+    \caption[short]{Total annual system cost for the European energy system for the different scenarios subdivided into groups of technologies.
+While the `Baseline' scenario has neither a \hydrogen{} or \carbon{} network, `Hybrid' is allowed to expand both.
 "Gas Infrastructure" combines gas facilities for power and heat production, "H$_2$ Infrastructure" combines H$_2$ production, transport and re-electrification.
@@
 
-System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
-In the Baseline model, the absence of dedicated carbon and hydrogen transport networks requires the system to deploy carbon production technologies (DAC and CC at point sources) in the regions with carbon sequestration sites, and small amounts of hydrogen production technologies (mainly electrolysis) in each region to meet local hydrogen demand for industry and land transport fuel cells.
+System cost differences between the scenarios are driven by renewable energy resources (solar, wind and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
+In the Baseline scenario, the absence of dedicated carbon and hydrogen transport networks requires the system to deploy carbon production technologies (DAC and CC at point sources) in the regions with carbon sequestration sites, and small amounts of hydrogen production technologies (mainly electrolysis) in each region to meet local hydrogen demand for industry and land transport fuel cells.
 For CU and its respective carbon and hydrogen demands, the system chooses the sites where on-site production of carbon, hydrogen and the synthesis process is cheapest.
-This leads to renewable resources and storage solutions being deployed not necessarily in the regions with the best resources or lowest costs for clean electricity generation, but where they are situated closer to the locations with carbon supply from CC at point sources, and to the locations of energy demand for electrolysers and DAC.
-Thus, total system costs in the Baseline model are highest, while the models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs.
+This leads to renewable resources and storage solutions being deployed not necessarily in the regions with the best resources or lowest costs for clean electricity generation, but where they are situated closer to the energy demand locations of electrolysers and DAC.
+Thus, total system costs in the baseline scenario are highest, while the scenarios with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs.
 
-We can now closer examine the cost reductions and shifts in cost allocation when shifting from the Baseline model to the other three network models.
+We can now closer examine the cost reductions and shifts in cost allocation when shifting from the baseline scenario to the other three network scenarios.
 The implementation of a \carbon{} transport network (\textit{\carbon{}~Grid}) facilitates efficient point-source carbon capture, substantially reducing the dependency on DAC technologies.
-While DAC supplies around 50\% of the carbon needed in the Baseline model (400 Mt \carbon{}, see Fig.~\ref{fig:balance_captured_carbon}), in the \modCO{}, DAC is only used to supply around 20\% of the carbon (150 Mt \carbon{}).
+While DAC supplies around 50\% of the carbon needed in the Baseline scenario (400 Mt \carbon{}, see Fig.~\ref{fig:balance_captured_carbon}), in the \scCO scenario, DAC is only used to supply around 20\% of the carbon (150 Mt \carbon{}).
 The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Fig.~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology is relatively cheaper than DAC.
 The \carbon{} transport network enables the transport of carbon from the locations of the cheapest captured carbon supply sources to the locations of the carbon demand sinks at CU amd CS sites, reducing the need for expensive DAC.
-The improved carbon management infrastructure also allows for a more strategic deployment of renewables, particularly in areas that combine high quality renewables and high sequestration potentials, leading to system-wide cost savings in renewable generation and storage.
+The improved carbon management infrastructure also allows for a more strategic deployment of renewables, particularly in areas with high sequestration potential, leading to cost savings in renewable generation and storage.
 
@@
 
-The \hydrogen{}-Grid model underscores the pivotal role of hydrogen infrastructure.
+The \hydrogen{}~Grid scenario underscores the pivotal role of hydrogen infrastructure.
 The creation of a hydrogen transport network enables the cost-effective delivery of hydrogen from regions with a surplus of cheap renewable energy to high-demand industrial centers.
-This model shows a pronounced decrease in investments in solar generation and gas infrastructure.
+This scenario shows a pronounced decrease in investments in solar and gas  infrastructure.
 The \hydrogen{} network reduces DAC similarly to the \carbon{} network, but still produces 200 Mt of carbon from DAC.
 CC technology on process emissions increases from 90\% to 100\%, on biomass CHPs from 60\% to 100\%, and on facilities using gas in industry from 50\% to 75\%.
-CC on biomass use in industry remains at 50\%, same as the \modCO{} (see Figure~\ref{fig:captureshare_line}).
+CC on biomass use in industry remains at 50\%, same as the \scbase.
 This variation highlights how the implementation of \hydrogen{} and \carbon{} networks unlocks specific carbon capture potentials.
 
-While the cost-reduction in the Hybrid model compared to the \hydrogen-Grid model is marginal, the allocation of costs reveals a propagation of reduced DAC facilities and increased CC at point sources.
-The Hybrid model relies slightly less on investments into the \hydrogen{} network than the \modH, but adds some investments into the \carbon{} network.
-Still, investments into the \hydrogen{} network outweigh investments into the \carbon{} network by a factor of four.
+While the cost-reduction in the Hybrid scenario compared to the \hydrogen~Grid scenario is marginal, the allocation of costs reveals a propagation of reduced DAC facilities and increased CC at point sources.
+The Hybrid scenario relies much heavier on investments into the \hydrogen{} network which outweigh investments in the \carbon{} network by a factor of 4.
 

### 2023-11-23 workflow/scripts/plot-capacity-map.py

@@ -98,9 +98,11 @@ for kind, output in snakemake.output.items():
             offshore_regions.assign(color=offshore_data),
         ]
         regions = pd.concat(regions)
+        capacity = "sequestration capacity"
     else:
         region_data = region_data.groupby(level=0).sum()
         regions = onshore_regions.assign(color=region_data)
+        capacity = "storage capacity"
 
     regions.color = regions.color * region_unit_conversion
 
@@ -125,7 +127,7 @@ for kind, output in snakemake.output.items():
         linewidth=0.0,
         legend=True,
         legend_kwds={
-            "label": kind.title() + f" storage capacity [{region_unit}]",
+            "label": kind.title() + f" {capacity} [{region_unit}]",
             "orientation": "horizontal",
             "shrink": 0.8,
             "pad": 0.05,

### 2023-11-25 report/submissions/journal/paper/paper.tex
incorporate feedback from Tom

--- 2023-11-23 report/submissions/journal/paper/paper.tex
@@
 \newcommand{\hydrogen}{H$_2$}
+\newcommand{\carbongrid}{\carbon{}--Grid}
 \newcommand{\hydrogengrid}{\hydrogen{}--Grid}
-\newcommand{\carbongrid}{\carbon{}--Grid}
 \newcommand{\modBase}{Baseline model}
@@
 \title{From Net-Zero to Net-Negative: Assessing \hydrogen{} and \carbon{} Network Strategies in Europe}
+% \title{Synthetic fuels in Europe: Transport Hydrogen to Carbon, or Carbon to Hydrogen?}
 
@@
     The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors, particularly in emissions-intensive industries like heavy manufacturing.
-While management of both \hydrogen{} and \carbon{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructures effectively interact with comprehensive carbon management technologies, including carbon capture, transport, use, and storage.
+While management of both \carbon{} and \hydrogen{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructures effectively interact with comprehensive carbon management technologies, including carbon capture, transport, use, and storage.
 In particular, the role of a carbon transport network as a complementing or substituting infrastructure to hydrogen transport remains unclear.
@@
     Our findings reveal that in a either-or scenario, the hydrogen network is superior to the carbon network as it enables the transportation cheap hydrogen inland to serve industrial processes and carbon capture and utilization at point sources.
-A carbon network facilitates point-source carbon capture and reduces the need for direct air capture, but leaves the scarcity of hydrogen in energy-intensive regions such as industrial clusters.
+A carbon network facilitates point-source carbon capture and reduces the need for direct air capture, but leaves the scarcity of low-cost hydrogen in energy-intensive regions such as industrial clusters.
     However, in a hybrid scenario, the carbon network is a cost-effective complement to the hydrogen network facilitating carbon capture at point sources and transport to sequestration sites.
-    We show that theses findings holds true against the backdrop of a net-zero energy system, as well as scenarios where the net-negative emissions are 5\% and 10\% lower than the 1990 baseline levels.
+    We show that theses findings holds true against the backdrop of a net-zero energy system as well as a scenario with net-negative emissions.
     Overall, our work demonstrates the cost-effectiveness of a multi-grid system that includes both hydrogen and carbon transport networks and power grid expansion to achieve climate neutrality in Europe.
@@
 Hydrogen offers an efficient way to transport and store energy from different sources.
-Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of both fossil and synthetic fuels through Carbon Capture (CC) techniques or harvested directly from the atmosphere using Direct Air Capture (DAC).
-Additionally, it can be stored in geological formations, a process known as carbon sequestration or carbon capture and storage (CCS).
-In combination, both together build the basis for climate-neutral fuels like methane, methanol or Fischer-Tropsch fuels, central to Power-to-X (PtX) and Carbon Capture and Utilization (CCU) strategies.
+Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, and synthetic fuels through Carbon Capture (CC) techniques, or harvested directly from the atmosphere using Direct Air Capture (DAC).
+Additionally, it can be stored in geological formations, a process known as carbon capture and sequestration (CCS).
+In combination, \carbon{} and \hydrogen{} networks build the basis for climate-neutral fuels needed for aviation, shipping and industrial feedstocks, and are central to Power-to-X (PtX) and Carbon Capture and Utilization (CCU) strategies.
 
@@
 However, up to this point, it remains unclear how the two management systems of hydrogen and carbon may complement or replace each other, particularly when it comes to their transport systems.
-A hydrogen network facilitates the transport of energy from renewable sources to regions with high energy demand, also enabling Carbon Capture and Utilization (CCU) from specific point sources.
+A hydrogen network facilitates the transport of energy from renewable sources to regions with high energy demand, also enabling CCU at the site of point-source capture.
 On the other hand, the carbon transport system allows for capturing and transporting carbon from areas with high emissions to regions with significant sequestration potential and abundant renewable sources, thereby enabling cost-effective CCU.
-In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
+In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,bec attiniCarbonDioxideCapture2022}.
 In contrast to Integrated Assessment Models, the underlying optimization tools account for the spatial distribution of carbon sources and sinks which are crucial provide a holistic view of the energy system and its technological interactions..
@@
 
-To our knowledge, no study has yet considered the co-optimization and comprehensive assessment of both \hydrogen{} and \carbon{} networks in a fully sector-coupled energy system.
+To our knowledge, no study has yet considered the co-optimization and comprehensive assessment of both \carbon{} and \hydrogen{} networks in a fully sector-coupled energy system.
 However, we argue that such an assessment is strongly needed to avoid suboptimal investments and to identify synergies between hydrogen and carbon management technologies.
@@
 The study is conducted using the PyPSA-Eur energy system model and encompasses all relevant energy sectors.
-We enable the system to optimize the design of renewable energy sources and storage technologies, as well as the transmission of electricity, \hydrogen{}, methane, and \carbon{}.
-We show that the dynamics of carbon capture, transport, use, and storage in different regions are highly dependent on the availability of renewable energy, \carbon{} pipelines and sequestration sinks.
+We explore competition between \carbon{} and \hydrogen{} networks by adding them separately.
 Our evaluation focuses on the transport dynamics of \carbon{} and \hydrogen{} through their respective networks on the European continent.
@@
     \includegraphics[width=\linewidth]{baseline/figures/90_nodes/total-demand-bar.png}
-    \caption{Exogenous demand assumptions. The figure shows the total amount of yearly energy demands for each energy carrier and type of end-consumer considered in the model.}
+    \caption{Exogenous demand assumptions. The figure shows the total amount of yearly energy demands for each energy carrier and type of end-consumer considered in the model. }
     \label{fig:total-demand-bar}
@@
     \item[] \textit{Baseline:} Neither \carbon{} nor \hydrogen{} networks are constructed.
+    \item[] \textit{\carbongrid:} Only the \carbon{} network is developed.
     \item[] \textit{\hydrogengrid:} Only the \hydrogen{} network is developed.
-    \item[] \textit{\carbongrid:} Only the \carbon{} network is developed.
     \item[] \textit{Hybrid:} Both \carbon{} and \hydrogen{} networks are developed.
@@
     \caption[short]{Total annual system cost for the European energy system for the different models subdivided into groups of technologies.
-While the `Baseline' model has neither a \hydrogen{} or \carbon{} network, `Hybrid' is allowed to expand both.
+While the `Baseline' model has neither a \carbon{} nor a \hydrogen{} network, `Hybrid' is allowed to expand both.
 "Gas Infrastructure" combines gas facilities for power and heat production, "H$_2$ Infrastructure" combines H$_2$ production, transport and re-electrification.
@@
 CC on biomass use in industry remains at 50\%, same as the \modCO{} (see Figure~\ref{fig:captureshare_line}).
-This variation highlights how the implementation of \hydrogen{} and \carbon{} networks unlocks specific carbon capture potentials.
+This variation highlights how the implementation of \carbon{} and \hydrogen{} networks unlocks specific carbon capture potentials.
 
@@
 The modelling results are heavily driven by the demand and emissions from industrial clusters.
-Allowing the model to relocate these and/or incorporate flexibility measures, may lead to less dependency on both \hydrogen{} and \carbon{} networks.
+Allowing the model to relocate these and/or incorporate flexibility measures, may lead to less dependency on both \carbon{} and \hydrogen{} networks.
 
@@
     \caption{Total annual system cost for the different scenarios and \carbon{} targets.
-While the `Baseline' scenario has neither a \hydrogen{} or \carbon{} network, `Hybrid' is allowed to expand both.}
+While the `Baseline' scenario has neither a \carbon{} nor a \hydrogen{} network, `Hybrid' is allowed to expand both.}
     \label{fig:objective_heatmap}

### 2023-11-25 report/submissions/journal/paper/paper.tex
incorporate feedback from Tom 2/2

@@
 \begin{abstract}
-    The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors, particularly in emissions-intensive industries like heavy manufacturing.
-While management of both \carbon{} and \hydrogen{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructures effectively interact with comprehensive carbon management technologies, including carbon capture, transport, use, and storage.
-In particular, the role of a carbon transport network as a complementing or substituting infrastructure to hydrogen transport remains unclear.
+    In future climate-neutral scenarios, the transport of hydrogen and carbon dioxide can play competing roles to enable the production of carbonaceous fuels such as Fischer-Tropsch fuels, methanol or synthetic methane, made from hydrogen and carbon dioxide: If carbonaceous fuel refineries are located at or near hydrogen production or import site, a \carbon{} network is needed to move \carbon{} from the cheapest carbon capture sites.
+If carbonaceous fuel refineries are located at or near carbon capture sites, a \hydrogen{} network is needed to transport \hydrogen{} from the cheapest hydrogen production sites.
+Whether carbon and hydrogen transport networks are complementing or substituting each other remains unclear.
     To address this gap, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates carbon management technologies, hydrogen transport and storage, and renewable energy sources with a high spatial and temporal resolution considering all energy intensive sectors.
@@
 
-An integrated management of both hydrogen and carbon is considered a crucial component to achieve climate neutrality in energy systems including hard-to-abate sectors.
+The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors.
+While management of both \carbon{} and \hydrogen{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructures effectively interact with comprehensive carbon management technologies, including carbon capture, transport, use, and storage.
 Hydrogen offers an efficient way to transport and store energy from different sources.
@@
     \includegraphics[width=\linewidth]{baseline/figures/90_nodes/total-demand-bar.png}
-    \caption{Exogenous demand assumptions. The figure shows the total amount of yearly energy demands for each energy carrier and type of end-consumer considered in the model. }
+    \caption{Exogenous demand assumptions.
+The figure shows the total annual energy demands for each energy carrier which drive all model activities.
+Endogenous processes may lead to higher total production volumes of some carriers, e.g.: Demand in methanol requires more hydrogen and carbon as secondary (energy) inputs, which in turn need additional electricity to run electrolysis and carbon capture processes.}
     \label{fig:total-demand-bar}
@@
 Exogenous assumptions on energy demand and non-abatable emissions are taken from various sources~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023}.
-The energy demand for electricity, transport, biomass, heat and gas are defined per region and time-step with heat demand being regionally subdivided into shares of urban, rural and industrial sites.
-Demands for aviation kerosene, methanol for shipping and naphtha for industry are aggregated in the system scope and kept constant throughout all time-steps.
-We show the sum of all energy demands in \ref{fig:total-demand-bar}.
-The system emits of 633~Mt \carbon{} per year through non-abatable emissions from industry, aviation, shipping and agriculture into account, 153~Mt of which are fossil-based industrial emissions.
+The energy demand for electricity, transport, biomass, heat and gas are defined per region and time-step.
+Land transport demand is exogenously divided between electric vehicles (85\%) and fuel cell vehicles (15\%), the latter representing demand for heavy duty land transport.
+Demands for kerosene for aviation, methanol for shipping, and naphtha for industry are aggregated in the system scope and kept constant throughout all time-steps.
+Heat demand is regionally subdivided into shares of urban, rural and industrial sites.
+We show the sum of all energy demands in Fig.~\ref{fig:total-demand-bar}.
+The system emits 633~Mt \carbon{} per year from industry, aviation, shipping and agriculture, 153~Mt of which are fossil-based process emissions.
 Industrial energy demand and excess heat potentials are calculated per node on the basis of~\cite{hotmaps_industrial_db}.
@@
 %
-Endogenous assumptions include the expansion of renewable energy sources, storage technologies, and transmission capacities.
+Endogenous model results include the expansion of renewable energy sources, storage technologies, and transmission capacities.
 The model considers the various energy carriers like electricity, hydrogen, methan, methanol, liquid hydrocarbons and biomass, together with conversions technologies to convert these into each other.
@@
 Methane is transported through the gas network, while methanol and FT-synfuel can be transferred inter-regionally without additional costs.
-Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial gas needs.
+Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial process heat.
 Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production, aviation kerosene, or agricultural machinery oil.
 
-To supply carbon needed for CS and CU, the system can choose to deploy carbon capture technology at various point sources, or centralized DAC facilities.
+To supply carbon needed for Carbon Sequestration (CS) and Carbon Utilization (CU), the system can choose to deploy carbon capture technology at various point sources, or centralized DAC facilities.
 The concept of a ``merit order for captured carbon`` plays a pivotal role in optimizing and prioritizing the deployment of carbon capture technologies based on their economic feasibility.
@@
 Moving up the scale, gas used in industrial applications and biomass employed for industrial processes represent more costly yet viable options to capture carbon.
-DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a higher cost level due to its current technological and operational expenses.
-Finally, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, typically incurs the highest costs in the merit order per ton of captured \carbon{} but offers a flexible (yet carbon emitting) energy source as a product.
+DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand.
+Finally, based on our model results, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per ton of captured \carbon{}.
+Biogas input is an expensive fuel which based on the endogenous modeling decisions is not used in large quantities.
+Thus, the marginal price of capturing additional \carbon{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the cost of captured \carbon{}.
 This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.
@@
 While the `Baseline' model has neither a \carbon{} nor a \hydrogen{} network, `Hybrid' is allowed to expand both.
-"Gas Infrastructure" combines gas facilities for power and heat production, "H$_2$ Infrastructure" combines H$_2$ production, transport and re-electrification.
+"Gas Infrastructure" combines gas facilities for power and heat production,"\carbon{} Infrastrucure" combines transport, storage and sequestration, "H$_2$ Infrastructure" combines transport and storage.
 "Carbon Capture at Point Sources" combines all technologies with integrated carbon capture.}
@@
     \label{fig:captureshare_line}
-\end{figure}
+\end{figure}%
 
-\begin{figure}[h]
+\begin{figure}[ht!]
     \centering
@@
 CC technology on process emissions increases from 90\% to 100\%, on biomass CHPs from 60\% to 100\%, and on facilities using gas in industry from 50\% to 75\%.
-CC on biomass use in industry remains at 50\%, same as the \modCO{} (see Figure~\ref{fig:captureshare_line}).
+CC on biomass use in industry remains at 50\%, same as the \modCO{} (see Fig.~\ref{fig:captureshare_line}).
 This variation highlights how the implementation of \carbon{} and \hydrogen{} networks unlocks specific carbon capture potentials.
 
-While the cost-reduction in the Hybrid model compared to the \hydrogen-Grid model is marginal, the allocation of costs reveals a propagation of reduced DAC facilities and increased CC at point sources.
+While the cost reduction in the Hybrid model compared to the \hydrogen-Grid model is marginal, the allocation of costs reveals a propagation of reduced DAC facilities and increased CC at point sources.
 The Hybrid model relies slightly less on investments into the \hydrogen{} network than the \modH, but adds some investments into the \carbon{} network.
@@
 
-\begin{figure*}[h]
+\subsection*{Superiority of \hydrogen{} network}
+
+\begin{figure*}[ht!]
     \centering
-    \begin{subfigure}{\linewidth}
+    \begin{subfigure}{0.9\linewidth}
         \centering
-        \includegraphics[width=0.95\linewidth]{co2-only/figures/90_nodes/operation_map_carbon.png}
+        \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_carbon.png}
         \caption{\carbon{} Sector.}
@@
     \end{subfigure}
-    \begin{subfigure}{\linewidth}
+    \begin{subfigure}{0.9\linewidth}
         \centering
-        \includegraphics[width=0.95\linewidth]{h2-only/figures/90_nodes/operation_map_hydrogen.png}
+        \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_hydrogen.png}
         \caption{\hydrogen{} Sector.}
@@
 \end{figure*}
-
-\subsection*{Superiority of \hydrogen{} network}
 
@@
 The two fundamental differences between the grid strategies are quickly evident: The \carbongrid{} model captures large amounts of carbon from bioenergy and industrial processes, especially in central and eastern Europe, at low cost.
-The carbon grid transports cheap carbon to regions where it is sequestered in offshore caverns or processed into synfuel, methane and methanol together with cheaply available green hydrogen.
+The carbon grid transports cheap carbon to regions where it is sequestered in offshore caverns or processed into carbonaceous fuels together with cheaply available green hydrogen.
 End-use primarily depends on the quality of renewable resources, and thus the availability of cheap hydrogen.
@@
 
-Similar to the \hydrogengrid model, the Hybrid model takes advantage of a large hydrogen network that enables hydrogen to be transported from centralized production sites in western regions such as Spain and the UK to decentralized FT production sites across the continent.
+Similar to the \hydrogengrid model, the Hybrid model takes advantage of a large hydrogen network that enables hydrogen to be transported from centralized production sites in western regions such as Spain to decentralized FT production sites across the continent.
+However, in the Hybrid model hydrogen in the UK is used locally to produce synfuels, and less hydrogen is transported to Central Europe.
 The carbon network supplements this with smaller clusters of networks.
@@
 A notable aspect of the hybrid model is the lack of overlap in the hydrogen and carbon network topologies which is tightly connected to the nodal carbon strategy.
-Depending on the exogeneous need in hydrogen for example from industry, and the costs for locally produced carbon, a region can either import hydrogen to produce FT fuels or export carbon.
-\dots
+Depending on the exogeneous need in hydrogen for example from industry, and the costs for locally produced carbon, a region can either import hydrogen to produce FT fuels or export carbon, but it will neither import one and export the other nor import or export both.
 
@@
 
-\begin{figure}[h]
+\begin{figure}[htb!]
     \centering
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
 
-\begin{figure}[h]
+\begin{figure*}[ht]
     \centering
@@
     \label{fig:cost_map_difference}
-\end{figure}
+\end{figure*}
 

### 2023-11-25 report/submissions/journal/paper/paper.tex
rework abstract (still to be shortened significantly)

@@
 \begin{abstract}
-    In future climate-neutral scenarios, the transport of hydrogen and carbon dioxide can play competing roles to enable the production of carbonaceous fuels such as Fischer-Tropsch fuels, methanol or synthetic methane, made from hydrogen and carbon dioxide: If carbonaceous fuel refineries are located at or near hydrogen production or import site, a \carbon{} network is needed to move \carbon{} from the cheapest carbon capture sites.
-If carbonaceous fuel refineries are located at or near carbon capture sites, a \hydrogen{} network is needed to transport \hydrogen{} from the cheapest hydrogen production sites.
+    In future climate-neutral scenarios, the transport of hydrogen and carbon dioxide can play competing roles.
+    Hydrogen transport allows energy-intensive regions to import low-cost hydrogen.
+Carbon transport allows regions with carbon sinks such as carbon sequestration and synthetic fuel production to import low-cost carbon from distributed carbon capture sources.
     Whether carbon and hydrogen transport networks are complementing or substituting each other remains unclear.
+This is especially relevant for the production of carbonaceous fuels such as Fischer-Tropsch fuels, methanol or synthetic methane, which are made from both hydrogen and carbon dioxide, and in most locations require at least one of these feedstocks to be imported.
     To address this gap, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates carbon management technologies, hydrogen transport and storage, and renewable energy sources with a high spatial and temporal resolution considering all energy intensive sectors.
-    Our findings reveal that in a either-or scenario, the hydrogen network is superior to the carbon network as it enables the transportation cheap hydrogen inland to serve industrial processes and carbon capture and utilization at point sources.
-A carbon network facilitates point-source carbon capture and reduces the need for direct air capture, but leaves the scarcity of low-cost hydrogen in energy-intensive regions such as industrial clusters.
-    However, in a hybrid scenario, the carbon network is a cost-effective complement to the hydrogen network facilitating carbon capture at point sources and transport to sequestration sites.
+    Our findings reveal that when comparing either-or-models, a system with an hydrogen network provides lower system costs than a system with a carbon network.
+The former enables the transportation of hydrogen from centralized production sites with low-cost renewable energy in the Iberian Peninsula and the British Isles to serve distributed hydrogen demands across the European continent.
+The latter facilitates carbon capture at lower costs from distributed point sources and reduces the need for more expensive direct air capture.
+However, the system with only a carbon network needs to meet distributed hydrogen demand across the continent by locally deploying electrolysis regardless of renewable energy costs.
+    In a hybrid model with both types of networks, the carbon network is a cost-effective complement to the hydrogen network facilitating point-source carbon capture at low costs and transport to sequestration sites.
+In all models, additional demand for synthetic fuels is met at sites with favorable renewable energy resources that enable both electrolysis and direct air capture at low costs.
     We show that theses findings holds true against the backdrop of a net-zero energy system as well as a scenario with net-negative emissions.
@@
 While management of both \carbon{} and \hydrogen{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructures effectively interact with comprehensive carbon management technologies, including carbon capture, transport, use, and storage.
-Hydrogen offers an efficient way to transport and store energy from different sources.
-Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, and synthetic fuels through Carbon Capture (CC) techniques, or harvested directly from the atmosphere using Direct Air Capture (DAC).
-Additionally, it can be stored in geological formations, a process known as carbon capture and sequestration (CCS).
-In combination, \carbon{} and \hydrogen{} networks build the basis for climate-neutral fuels needed for aviation, shipping and industrial feedstocks, and are central to Power-to-X (PtX) and Carbon Capture and Utilization (CCU) strategies.
+Hydrogen offers an efficient way to transport and store energy.
+Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, and synthetic fuels through Carbon Capture (CC) techniques, or harvested from the atmosphere using Direct Air Capture (DAC).
+Additionally, it can be stored in geological formations, a process known as carbon sequestration (CS).
+In combination, \carbon{} and \hydrogen{} networks build the basis for climate-neutral fuels needed for aviation, shipping and industrial feedstocks, and are central to Power-to-X (PtX) and Carbon Utilization (CU) strategies.
 
@@
 However, up to this point, it remains unclear how the two management systems of hydrogen and carbon may complement or replace each other, particularly when it comes to their transport systems.
-A hydrogen network facilitates the transport of energy from renewable sources to regions with high energy demand, also enabling CCU at the site of point-source capture.
-On the other hand, the carbon transport system allows for capturing and transporting carbon from areas with high emissions to regions with significant sequestration potential and abundant renewable sources, thereby enabling cost-effective CCU.
-In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,bec attiniCarbonDioxideCapture2022}.
+A hydrogen network facilitates the transport of energy from renewable sources to regions with high energy demand, also enabling CU at the site of point-source capture.
+On the other hand, the carbon transport system allows for capturing and transporting carbon from areas with high emissions to regions with significant sequestration potential and abundant renewable sources, thereby enabling cost-effective CU.
+In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
 In contrast to Integrated Assessment Models, the underlying optimization tools account for the spatial distribution of carbon sources and sinks which are crucial provide a holistic view of the energy system and its technological interactions..
@@
 
-To supply carbon needed for Carbon Sequestration (CS) and Carbon Utilization (CU), the system can choose to deploy carbon capture technology at various point sources, or centralized DAC facilities.
+To supply carbon needed for CS and CU, the system can choose to deploy carbon capture technology at various point sources, or centralized DAC facilities.
 The concept of a ``merit order for captured carbon`` plays a pivotal role in optimizing and prioritizing the deployment of carbon capture technologies based on their economic feasibility.
@@
 In the Baseline model, the absence of dedicated carbon and hydrogen transport networks requires the system to deploy carbon production technologies (DAC and CC at point sources) in the regions with carbon sequestration sites, and small amounts of hydrogen production technologies (mainly electrolysis) in each region to meet local hydrogen demand for industry and land transport fuel cells.
-For CU and its respective carbon and hydrogen demands, the system chooses the sites where on-site production of carbon, hydrogen and the synthesis process is cheapest.
+For CU and its respective carbon and hydrogen demands, the system chooses the sites where on-site production of carbon, hydrogen and the synthesis process has the lowest cost.
 This leads to renewable resources and storage solutions being deployed not necessarily in the regions with the best resources or lowest costs for clean electricity generation, but where they are situated closer to the locations with carbon supply from CC at point sources, and to the locations of energy demand for electrolysers and DAC.
@@
 While DAC supplies around 50\% of the carbon needed in the Baseline model (400 Mt \carbon{}, see Fig.~\ref{fig:balance_captured_carbon}), in the \modCO{}, DAC is only used to supply around 20\% of the carbon (150 Mt \carbon{}).
-The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Fig.~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology is relatively cheaper than DAC.
-The \carbon{} transport network enables the transport of carbon from the locations of the cheapest captured carbon supply sources to the locations of the carbon demand sinks at CU amd CS sites, reducing the need for expensive DAC.
+The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Fig.~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology has relatively lower costs than DAC.
+The \carbon{} transport network enables the transport of carbon from the locations of the lowest-cost captured carbon sources to the locations of the carbon sinks at CU and CS sites, reducing the need for expensive DAC.
 The improved carbon management infrastructure also allows for a more strategic deployment of renewables, particularly in areas that combine high quality renewables and high sequestration potentials, leading to system-wide cost savings in renewable generation and storage.
@@
 The \hydrogen{}-Grid model underscores the pivotal role of hydrogen infrastructure.
-The creation of a hydrogen transport network enables the cost-effective delivery of hydrogen from regions with a surplus of cheap renewable energy to high-demand industrial centers.
+The creation of a hydrogen transport network enables the cost-effective delivery of hydrogen from regions with a surplus of low-cost renewable energy to high-demand industrial centers.
 This model shows a pronounced decrease in investments in solar generation and gas infrastructure.
@@
 The two fundamental differences between the grid strategies are quickly evident: The \carbongrid{} model captures large amounts of carbon from bioenergy and industrial processes, especially in central and eastern Europe, at low cost.
-The carbon grid transports cheap carbon to regions where it is sequestered in offshore caverns or processed into carbonaceous fuels together with cheaply available green hydrogen.
-End-use primarily depends on the quality of renewable resources, and thus the availability of cheap hydrogen.
+The carbon grid transports low-cost carbon to regions where it is sequestered in offshore caverns or processed into carbonaceous fuels together with green hydrogen available at low costs.
+End-use primarily depends on the quality of renewable resources, and thus the availability of low-cost hydrogen.
 In locations with relatively poor renewables and thus higher costs of hydrogen production, such as Italy, Greece, Northern Germany, and lcoations around the Baltic Sea, the transported carbon is predominantly sequestered.
-Conversely, in Spain and the United Kingdom, which both have high renewable potentials and cheap hydrogen production costs, the transported carbon is combined with hydrogen and fed into CU processes, with additional carbon being captured via DAC also running on the great renewable resources.
+Conversely, in Spain and the United Kingdom, which both have high renewable potentials and low-cost hydrogen production, the transported carbon is combined with hydrogen and fed into CU processes, with additional carbon being captured via DAC also running on the great renewable resources.
 
-In summary, the implementation of a \carbon{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce cheap hydrogen to make CU products (FT fuel, methanol, synthetic methane).
-In the absence of a \carbon{} network, sequestration sites first sequester all carbon from cheap point sources, and deploy additional DAC facilities at the locations with cheap renewables to capture the remaining necessary carbon for sequestration.
-Then, synfuel demand is met at the locations which offer the best mix of cheap carbon from point sources and cheap hydrogen from local renewables.
+In summary, the implementation of a \carbon{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce low-cost hydrogen to make CU products (FT fuel, methanol, synthetic methane).
+In the absence of a \carbon{} network, sequestration sites first sequester all carbon from low-cost point sources, and deploy additional DAC facilities at the locations with low-cost renewables to capture the remaining necessary carbon for sequestration.
+Then, synfuel demand is met at the locations which offer the best mix of low-cost carbon from point sources and low-cost hydrogen from local renewables.
 % Sequestered carbon is nearly always transported before
 
-The \hydrogengrid{} on the other hand, transports cheap hydrogen from regions with abundant renewable resources to regions with high energy demand and abundant emissions.
+The \hydrogengrid{} on the other hand, transports low-cost hydrogen from regions with abundant renewable resources to regions with high energy demand and abundant emissions.
 The main transportation routes go from Spain and the United Kingdom to Central Europe.
@@
 Central Europe with its industrial clusters and dense population is characterized by a scarcity in hydrogen and a surplus and abundance of \carbon.
-While the \carbon{} network cost-efficiently resolves the surplus of carbon in Central Europe by exporting it to regions with cheap renewable resources to produce synfuels, it needs to supply direct hydrogen demand in Central Europe (for industry and fuel cell land transport) with expensive local hydrogen production.
-The hydrogen network resolves both the energy scarcity by supplying hydrogen demand in Central Europe with cheap hydrogen imported from Spain and Great Britain, and the carbon surplus by transporting hydrogen to the sites with cheaper carbon available in Central Europe, and producing FT fuels highly decentralized throughout all of Central Europe.
+While the \carbon{} network cost-efficiently resolves the surplus of carbon in Central Europe by exporting it to regions with low-cost renewable resources to produce synfuels, it needs to supply direct hydrogen demand in Central Europe (for industry and fuel cell land transport) with expensive local hydrogen production.
+The hydrogen network resolves both the energy scarcity by supplying hydrogen demand in Central Europe with low-cost hydrogen imported from Spain and Great Britain, and the carbon surplus by transporting hydrogen to the sites with carbon available at lower costs in Central Europe, and producing FT fuels highly decentralized throughout all of Central Europe.
 
@@
 These transport carbon from Romania and Bulgaria to Greece, from northern Italy to central Italy and from Hungary, the Czech Republic and Poland to the Baltic Sea.
-These enable the system to transport cheaply available carbon from inland biomass sources and sequester it (see Fig.~\ref{fig:captureshare_line}).
+These enable the system to transport carbon available at low cost from inland biomass sources and sequester it (see Fig.~\ref{fig:captureshare_line}).
 This reduces the need for DAC in the system and the sequestration sites are shifted to the end points of the carbon networks.

### 2023-11-25 report/submissions/journal/paper/paper.tex
redact 'expensive' to 'high-cost'

@@
 The former enables the transportation of hydrogen from centralized production sites with low-cost renewable energy in the Iberian Peninsula and the British Isles to serve distributed hydrogen demands across the European continent.
-The latter facilitates carbon capture at lower costs from distributed point sources and reduces the need for more expensive direct air capture.
+The latter facilitates carbon capture at lower costs from distributed point sources and reduces the need for higher cost direct air capture.
 However, the system with only a carbon network needs to meet distributed hydrogen demand across the continent by locally deploying electrolysis regardless of renewable energy costs.
@@
 Finally, based on our model results, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per ton of captured \carbon{}.
-Biogas input is an expensive fuel which based on the endogenous modeling decisions is not used in large quantities.
+Biogas input is a high-cost fuel which based on the endogenous modeling decisions is not used in large quantities.
 Thus, the marginal price of capturing additional \carbon{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the cost of captured \carbon{}.
@@
 The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Fig.~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology has relatively lower costs than DAC.
-The \carbon{} transport network enables the transport of carbon from the locations of the lowest-cost captured carbon sources to the locations of the carbon sinks at CU and CS sites, reducing the need for expensive DAC.
+The \carbon{} transport network enables the transport of carbon from the locations of the lowest-cost captured carbon sources to the locations of the carbon sinks at CU and CS sites, reducing the need for higher cost DAC.
 The improved carbon management infrastructure also allows for a more strategic deployment of renewables, particularly in areas that combine high quality renewables and high sequestration potentials, leading to system-wide cost savings in renewable generation and storage.
@@
 Central Europe with its industrial clusters and dense population is characterized by a scarcity in hydrogen and a surplus and abundance of \carbon.
-While the \carbon{} network cost-efficiently resolves the surplus of carbon in Central Europe by exporting it to regions with low-cost renewable resources to produce synfuels, it needs to supply direct hydrogen demand in Central Europe (for industry and fuel cell land transport) with expensive local hydrogen production.
+While the \carbon{} network cost-efficiently resolves the surplus of carbon in Central Europe by exporting it to regions with low-cost renewable resources to produce synfuels, it needs to supply direct hydrogen demand in Central Europe (for industry and fuel cell land transport) with high-cost local hydrogen production.
 The hydrogen network resolves both the energy scarcity by supplying hydrogen demand in Central Europe with low-cost hydrogen imported from Spain and Great Britain, and the carbon surplus by transporting hydrogen to the sites with carbon available at lower costs in Central Europe, and producing FT fuels highly decentralized throughout all of Central Europe.

### 2023-11-26 report/submissions/journal/paper/paper.tex
incorporate comments from FN in Intro, Methodology

@@
 This is especially relevant for the production of carbonaceous fuels such as Fischer-Tropsch fuels, methanol or synthetic methane, which are made from both hydrogen and carbon dioxide, and in most locations require at least one of these feedstocks to be imported.
-    To address this gap, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates carbon management technologies, hydrogen transport and storage, and renewable energy sources with a high spatial and temporal resolution considering all energy intensive sectors.
+    To address this gap, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates transport, storage and sequestration for carbon dioxide, transport and storage for hydrogen, and renewable energy sources with a high spatial and temporal resolution considering all energy intensive sectors.
     Our findings reveal that when comparing either-or-models, a system with an hydrogen network provides lower system costs than a system with a carbon network.
@@
 However, the system with only a carbon network needs to meet distributed hydrogen demand across the continent by locally deploying electrolysis regardless of renewable energy costs.
-    In a hybrid model with both types of networks, the carbon network is a cost-effective complement to the hydrogen network facilitating point-source carbon capture at low costs and transport to sequestration sites.
+    In a hybrid model with both types of networks, the carbon network is a cost-effective complement to the hydrogen network facilitating point-source carbon capture at low costs and transport to sequestration or carbon utilization sites.
 In all models, additional demand for synthetic fuels is met at sites with favorable renewable energy resources that enable both electrolysis and direct air capture at low costs.
@@
     Overall, our work demonstrates the cost-effectiveness of a multi-grid system that includes both hydrogen and carbon transport networks and power grid expansion to achieve climate neutrality in Europe.
-    The paper underpins the need for a complementing hydrogen network and \carbon{} network to achieve climate neutrality.
+    The paper highlights the benefits of combining hydrogen and \carbon{} networks to achieve climate neutrality.
 \end{abstract}
@@
 Hydrogen offers an efficient way to transport and store energy.
-Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, and synthetic fuels through Carbon Capture (CC) techniques, or harvested from the atmosphere using Direct Air Capture (DAC).
+Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, and synthetic fuels through carbon capture (CC) techniques, or harvested from the atmosphere using direct air capture (DAC).
 Additionally, it can be stored in geological formations, a process known as carbon sequestration (CS).
-In combination, \carbon{} and \hydrogen{} networks build the basis for climate-neutral fuels needed for aviation, shipping and industrial feedstocks, and are central to Power-to-X (PtX) and Carbon Utilization (CU) strategies.
+In combination, \carbon{} and \hydrogen{} networks build the basis for climate-neutral fuels needed for aviation, shipping and industrial feedstocks, and are central to carbon utilization (CU) strategies.
 
-Recently, policymakers and industry in Europe have been committing to both hydrogen and carbon management projects, planning infrastructure components, and developing business models for emerging sectors of the economy.
+Recently, policymakers and industry in Europe have started developing carbon management strategies~\cite{GermanyDevelopingStrategy2023,CarbonManagementStrategie}, planning infrastructure components~\cite{CONetz}, and committing on the first carbon utilization projects~\cite{EFuelsPilotPlant2022,OrstedAssumesFull,GROUNDBREAKINGEFUELPRODUCTION,DLREfuelsDLR}.
 With the European Union's goal of achieving climate neutrality by 2050, a wide range of programs, funding models and initiatives have been established in this regard~\cite{eu2023netzero,europeangreendeal,europeaninnovationfund}.
-Initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier.
+Initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen Infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier, and some gas pipelines are already repurposed to transport hydrogen~\cite{RohrFreiFuer}.
 At the same time, business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offer \carbon{} offtake on the other hand.
-For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.4~Gt of carbon capture from point sources per year.
+For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.7~Gt of carbon capture from point sources per year, which represent roughly 50\% of all emissions.
 In combination with large sequestration potentials as stated in~\cite{weiProposedGlobalLayout2021}, this highlights the vast potential for decarbonization.
-The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a 1.5 Mt/a co2 transport and sequestration capacity to be operating in 2024, expanding to a targeted scale of 5 Mt/a sequestration by 2030.
+The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a transport and sequestration capacity of 1.5 Mt \carbon{} per year  to be operative in 2024, expanding to a targeted scale of 5 Mt per year of sequestration by 2030.
 To this end, the Clean Air Task Force underlines the importance of a carbon transport system in Europe to facilitate the carbon economy~\cite{lockwoodEuropeanStrategyCarbon}.
 
-However, up to this point, it remains unclear how the two management systems of hydrogen and carbon may complement or replace each other, particularly when it comes to their transport systems.
+However, up to this point, it remains unclear how the two transport systems of hydrogen and carbon may complement or replace each other, particularly when it comes to their transport systems.
 A hydrogen network facilitates the transport of energy from renewable sources to regions with high energy demand, also enabling CU at the site of point-source capture.
@@
 In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
-In contrast to Integrated Assessment Models, the underlying optimization tools account for the spatial distribution of carbon sources and sinks which are crucial provide a holistic view of the energy system and its technological interactions..
+Such techno-economic models, in contrast for example to Integrated Assessment Models, can account for the spatial distribution of carbon sources and sinks which are crucial provide a holistic view of the energy system and its technological interactions.
 The work in~\cite{neumannBenefitsHydrogenNetwork2022} examines the effect of a hydrogen network in Europe...
@@
 The publication by Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \carbon{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
+This limited sectoral scope cannot capture important dynamics of carbon management, since it neglects the sectors that will need to handle most \carbon{} in the future.
 Another comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
+Hoewever, this limited spatial scope fails to consider other sequestration sites and co-benefits from connecting the \carbon{} network to neighboring countries.
 To this end, it is often argued that \carbon{} pipelines are a mature technology with an expected high learning rate, given the wide-spread installations in US and Canada for enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
@@
 We explore competition between \carbon{} and \hydrogen{} networks by adding them separately.
-Our evaluation focuses on the transport dynamics of \carbon{} and \hydrogen{} through their respective networks on the European continent.
+Our evaluation focuses on the transport of \carbon{} and \hydrogen{} through their respective networks on the European continent.
 We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available sequestration potential.
@@
 The study is conducted on the basis of the open-source, capacity-expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
-The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, waste, and transport sectors.
+The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, waste, agriculture, and transport sectors (including international aviation and shipping).
 
@@
 The figure shows the total annual energy demands for each energy carrier which drive all model activities.
-Endogenous processes may lead to higher total production volumes of some carriers, e.g.: Demand in methanol requires more hydrogen and carbon as secondary (energy) inputs, which in turn need additional electricity to run electrolysis and carbon capture processes.}
+Endogenous processes may lead to higher total production volumes of some carriers, e.g., demand in methanol requires more hydrogen and carbon as secondary (energy) inputs, which in turn need additional electricity to run electrolysis and carbon capture processes.}
+    % TODO: adjust labels to show that these are exogenous assumptions
     \label{fig:total-demand-bar}
@@
 Carbon-neutral electricity is provided by wind, solar, biomass, hydro and nuclear power plants.
+Hydro and nuclear plant capacities cannot be extended.
 Weather-dependent power potentials for solar, wind and hydro are calculated from reanalysis and satellite data sets~\cite{hersbachERA5GlobalReanalysis2020,pfeifrothSurfaceRadiationData2017}  per region and time-stamp, using the open-source tool Atlite~\cite{hofmannAtliteLightweightPython2021}.
@@
 We restrict the electric transmission system's expansion to 20\% of its current capacity, acknowledging the challenges in inaugurating new transmission projects.
-The regional use of biomass is restricted by its feasible potential derived from~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
+For the use of biomass we consider only residual biomass products and no energy crops.
+We limit regional biomass use to the medium-level potentials from the ENPRESO database~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
 Inter-regional biomass transport is permitted with transport costs considered.
 
-Hydrogen can be gained from electrolysis and steam methane reforming (SMR).
+In our model, hydrogen can be produced from electrolysis and steam methane reforming (SMR).
 Geological distribution and potentials for \hydrogen{} storage in salt caverns are derived from~\cite{caglayanTechnicalPotentialSalt2020}.
 Re-electrification of hydrogen is possible via fuel cells.
-If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limit.
+If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limit, considering costs for pipeline segments and compressors.
+Pipeline flows are modelled using net transfer capacities and without flow dynamics, pressure valves, or energy demand for compression.
+No retrofitting of gas pipelines is considered, potentially overestimating hydrogen network cost.
+
 The topology of the \hydrogen{} network is mirroring the topology of the electricity network.
 
-Our model features three drop-in fuel production technologies for carbon utilization: methanation, methanolization, and Fischer-Tropsch synthesis.
+Our model features three drop-in fuel production technologies for CU: methanation, methanolization, and Fischer-Tropsch synthesis.
 The processed fuels are not limited in their total quantity of use or production.
-Methane is transported through the gas network, while methanol and FT-synfuel can be transferred inter-regionally without additional costs.
+Methane is transported through the gas network, while methanol and FT fuel can be transferred inter-regionally without additional costs, since dense fuels have negligible transport costs per unit of energy.
 Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial process heat.
@@
 
-To supply carbon needed for CS and CU, the system can choose to deploy carbon capture technology at various point sources, or centralized DAC facilities.
+To supply carbon needed for CS and CU, the system can choose to deploy carbon capture technology at various point sources (see below), or through DAC facilities.
 The concept of a ``merit order for captured carbon`` plays a pivotal role in optimizing and prioritizing the deployment of carbon capture technologies based on their economic feasibility.
@@
 This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.
+If we consider the spatial aspect of distributed carbon capture potentials, the \carbon{} network can exploit comparative cost advantages between ``CO2 bidding zones'' to utilize the lowest-cost CC facilities across the continent.
+We assume a capture rate of 90\% for CC on process emissions, SMR, biogas-to-gas, as well as gas and biomass used in industry, and 95\% for CC on biomass and gas CHPs.
 
-We only consider offshore sites as potential sinks for carbon sequestration.
-This choice favors offshore storage due to its typically larger capacity compared to onshore and concerns over public safety for infrastructure near populated areas.
-Our estimates for storage potential are conservative, limiting total potential to 25~Mt per site, calculating annual storage availability over 25 years, and capping the total sequestered \carbon{} to 200~Mt/a.
-All technology cost assumptions for 2040 are sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
+To store carbon, we differentiate between short-term storage in steel tanks and long-term, irreversible sequestration in underground sequestration sites such as porous rock formations or depleted gas reservoirs.
+Costs for both options are included in the model.
+For carbon sequestration, we only consider offshore sites as potential sinks.
+We make this choice because offshore sequestration sites typically have larger capacity compared to onshore sequestration sites in saline aquifers and due to concerns over public safety for infrastructure near populated areas.
+Our estimates for carbon sequestration potentials are conservative, limiting total sequestration to 25~Mt per site and calculating annual storage availability over 25 years.
+Furthermore, we decide to cap the total sequestered \carbon{} to 200~Mt per year for our Net-Zero scenario.
+This is enough to offset the hardest-to-abate fossil process emissions and some limited slack of fossil fuels to be used where they can be most valuable.
+This is in order to avoid offsetting fossil emissions with carbon removal if they can technologically be avoided in the first place.
+All technology cost assumptions are taken for the year 2040 and sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
 
@@
 \end{itemize}
-Unless otherwise mentioned, our reference is the Net-zero scenario.
-For 5\% Net-negative and 10\% Net-negative scenarios, the carbon sequestration potential is adjusted to 400~Mt/a and 600~Mt/a, respectively.
+Unless otherwise mentioned, our reference is the Net-Zero scenario. For the Net-Negative scenario, the cap of total carbon sequestration is adjusted from 200~Mt to 660~Mt per year accordingly.
 
@@
 
-Focusing first on the net-zero scenario, we can highlight significant variations in system costs and technology deployment across the models (see Fig.~\ref{fig:cost_bar}).
+Focusing first on the Net-Zero scenario, we can highlight significant variations in system costs and technology deployment across the models (see Fig.~\ref{fig:cost_bar}).
 
@@
     \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/captureshare_line.png}
-    \caption{Share of facilities with integrated carbon capture for different system setups for the net-zero emission scenario.}
+    \caption{Share of facilities with integrated carbon capture for different system setups for the Net-Zero scenario.}
     % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
@@
 
+% \begin{figure*}[ht!]
+%     \centering
+%     \begin{subfigure}{0.9\linewidth}
+%         \centering
+%         \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_carbon.png}
+%         \caption{\carbon{} Sector.}
+%         \label{fig:operation_map_carbon}
+%     \end{subfigure}
+%     \begin{subfigure}{0.9\linewidth}
+%         \centering
+%         \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_hydrogen.png}
+%         \caption{\hydrogen{} Sector.}
+%         \label{fig:operation_map_hydrogen}
+%     \end{subfigure}
+%     \caption{Optimal operation per sector for a net-zero energy system in Europe with average production on the left and average consumption on the right for both, (a) the \carbon{} sector in the \carbon{}-Grid model and (b) the \hydrogen{} sector in the \hydrogen-Grid model.}
+%     \label{fig:operation_map}
+% \end{figure*}
+
 \begin{figure*}[ht!]
     \centering
-    \begin{subfigure}{0.9\linewidth}
+    \begin{subfigure}{0.49\linewidth}
         \centering
-        \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_carbon.png}
-        \caption{\carbon{} Sector.}
-        \label{fig:operation_map_carbon}
+        \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_single_carbon.png}
+        \caption{\carbon{} Sector in \carbongrid{} model.}
+        \label{fig:co2-only-operation_map_carbon}
     \end{subfigure}
-    \begin{subfigure}{0.9\linewidth}
+    \begin{subfigure}{0.49\linewidth}
         \centering
-        \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_hydrogen.png}
-        \caption{\hydrogen{} Sector.}
-        \label{fig:operation_map_hydrogen}
+        \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_single_hydrogen.png}
+        \caption{\hydrogen{} Sector in \carbongrid{} model.}
+        \label{fig:co2-only-operation_map_hydrogen}
+    \end{subfigure}
+    \begin{subfigure}{0.49\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_single_carbon.png}
+        \caption{\hydrogen{} Sector in \hydrogengrid{} model.}
+        \label{fig:h2-only-operation_map_carbon}
+    \end{subfigure}
+    \begin{subfigure}{0.49\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_single_hydrogen.png}
+        \caption{\hydrogen{} Sector in \hydrogengrid{} model.}
+        \label{fig:h2-only-operation_map_hydrogen}
     \end{subfigure}
@@
 We now examine the case of moving from a net zero emissions target to a net emissions reduction target.
-Fig.~\ref{fig:net-negative_cost_bar} illustrates the shift in investments for the \hydrogengrid{} and the Hybrid model between the net-zero and the net-negative scenario.
+Fig.~\ref{fig:net-negative_cost_bar} illustrates the shift in investments for the \hydrogengrid{} and the Hybrid model between the Net-Zero and the Net-Negative scenario.
 

### 2023-11-26 report/submissions/journal/paper/paper.tex
incorporate feedback from FN in Results

@@
 DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand.
-Finally, based on our model results, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per ton of captured \carbon{}.
+Finally, based on our model results, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per marginal ton of captured \carbon{}.
 Biogas input is a high-cost fuel which based on the endogenous modeling decisions is not used in large quantities.
-Thus, the marginal price of capturing additional \carbon{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the cost of captured \carbon{}.
+To the extent that biogas-to-gas facilities are used by the model to supply additional (carbon-neutral) gas as fuel, adding CC infrastructure incurs low costs (and thus ranges on the left-hand side of the merit order curve).
+However, the price of capturing additional \carbon{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the marginal cost of captured \carbon{}.
 This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.
@@
     \caption[short]{Total annual system cost for the European energy system for the different models subdivided into groups of technologies.
-While the `Baseline' model has neither a \carbon{} nor a \hydrogen{} network, `Hybrid' is allowed to expand both.
-"Gas Infrastructure" combines gas facilities for power and heat production,"\carbon{} Infrastrucure" combines transport, storage and sequestration, "H$_2$ Infrastructure" combines transport and storage.
-"Carbon Capture at Point Sources" combines all technologies with integrated carbon capture.}
+While the Baseline model has neither a \carbon{} nor a \hydrogen{} network, the Hybrid model is allowed to expand both.
+``Gas Infrastructure'' combines gas facilities for power and heat production, ``\carbon{} Infrastrucure'' combines transport, storage and sequestration, and ``H$_2$ Infrastructure'' combines transport and storage.
+``Carbon Capture at Point Sources'' combines all technologies with integrated carbon capture, including the cost of the main facility (e.g., CHP unit) and the carbon capture application.}
     \label{fig:cost_bar}
@@
 System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
-In the Baseline model, the absence of dedicated carbon and hydrogen transport networks requires the system to deploy carbon production technologies (DAC and CC at point sources) in the regions with carbon sequestration sites, and small amounts of hydrogen production technologies (mainly electrolysis) in each region to meet local hydrogen demand for industry and land transport fuel cells.
-For CU and its respective carbon and hydrogen demands, the system chooses the sites where on-site production of carbon, hydrogen and the synthesis process has the lowest cost.
-This leads to renewable resources and storage solutions being deployed not necessarily in the regions with the best resources or lowest costs for clean electricity generation, but where they are situated closer to the locations with carbon supply from CC at point sources, and to the locations of energy demand for electrolysers and DAC.
-Thus, total system costs in the Baseline model are highest, while the models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs.
+In the Baseline model, the absence of dedicated carbon and hydrogen transport networks requires the system to deploy carbon production technologies (DAC and CC at point sources) in the regions with carbon sequestration sites, and distributed amounts of hydrogen production technologies (mainly electrolysis) in each region to meet local hydrogen demand for industry and land transport fuel cells.
+For CU and its respective carbon and hydrogen demands, the system chooses the sites where on-site production of carbon, hydrogen as well as the inputs and by-product outputs (e.g., waste heat) of the synthesis process have the lowest cost (or highest value).
+This leads to renewable resources and storage solutions being deployed not necessarily in the regions with the best resources or lowest costs for clean electricity generation, but where they are situated closer to the endogenously defined locations of energy demand for electrolysers and DAC.
+Thus, total annual system costs in the Baseline model are highest, totalling \label{}803 billion euros (see Fig.~\ref{fig:objective_heatmap}).
+The models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs, with around \label{}2\% cost reductions (\label{}781 billion euros) for the \carbon{} network, and around \label{}4\% cost reductions for the \hydrogen{} network and the hybrid scenario (\label{}765 and \label{}764 billion euros, respectively).
 
 We can now closer examine the cost reductions and shifts in cost allocation when shifting from the Baseline model to the other three network models.
-The implementation of a \carbon{} transport network (\textit{\carbon{}~Grid}) facilitates efficient point-source carbon capture, substantially reducing the dependency on DAC technologies.
-While DAC supplies around 50\% of the carbon needed in the Baseline model (400 Mt \carbon{}, see Fig.~\ref{fig:balance_captured_carbon}), in the \modCO{}, DAC is only used to supply around 20\% of the carbon (150 Mt \carbon{}).
+The implementation of a \carbon{} transport network (\textit{\carbon{}~Grid}) facilitates carbon capture from the point sources with the lowest costs across regions, substantially reducing the dependency on DAC.
+While DAC supplies around 50\% of the carbon needed in the Baseline model (400 Mt \carbon{}, see Fig.~\ref{fig:balance_captured_carbon}), in the \modCO{}, DAC is only used to supply around 20\% of the carbon handled by the system (150 Mt \carbon{}).
 The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Fig.~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology has relatively lower costs than DAC.
 The \carbon{} transport network enables the transport of carbon from the locations of the lowest-cost captured carbon sources to the locations of the carbon sinks at CU and CS sites, reducing the need for higher cost DAC.
-The improved carbon management infrastructure also allows for a more strategic deployment of renewables, particularly in areas that combine high quality renewables and high sequestration potentials, leading to system-wide cost savings in renewable generation and storage.
+The option to transport carbon also allows for the deployment of renewables in areas that combine high quality renewable resources and high sequestration potentials, leading to system-wide cost savings in renewable generation and storage.
 
@@
 
-The \hydrogen{}-Grid model underscores the pivotal role of hydrogen infrastructure.
+The \hydrogen{}-Grid model underscores the beneficial role of hydrogen infrastructure.
 The creation of a hydrogen transport network enables the cost-effective delivery of hydrogen from regions with a surplus of low-cost renewable energy to high-demand industrial centers.
-This model shows a pronounced decrease in investments in solar generation and gas infrastructure.
-The \hydrogen{} network reduces DAC similarly to the \carbon{} network, but still produces 200 Mt of carbon from DAC.
+This model shows a pronounced decrease in investments in DAC, and accordingly also solar generation which provides most of the electricity to run DAC facilities.
+The reduction in DAC is similar to the \carbon{} network, but the \hydrogen{} network still produces 200 Mt of carbon from DAC.
 CC technology on process emissions increases from 90\% to 100\%, on biomass CHPs from 60\% to 100\%, and on facilities using gas in industry from 50\% to 75\%.
 CC on biomass use in industry remains at 50\%, same as the \modCO{} (see Fig.~\ref{fig:captureshare_line}).
-This variation highlights how the implementation of \carbon{} and \hydrogen{} networks unlocks specific carbon capture potentials.
+This is because the \hydrogen{} network can supply low-cost \hydrogen{} to \carbon{} point sources, where without the \hydrogen{} network, fuel synthethis with locally captured \carbon{} and locally sourced electrolytic hydrogen would have been more expensive than synthesizing fuel with \carbon{} from DAC and hydrogen from electrolysis at the best renewable resource sites.
+This highlights how the implementation of \carbon{} and \hydrogen{} networks unlocks carbon capture potentials at distributed point sources across the continent.
 
 While the cost reduction in the Hybrid model compared to the \hydrogen-Grid model is marginal, the allocation of costs reveals a propagation of reduced DAC facilities and increased CC at point sources.
+In other words, the \carbon{} network allows to tap into far away biomass potentials to capture carbon indirectly from the air, which provides marginally lower costs than DAC, even considering carbon transport costs.
 The Hybrid model relies slightly less on investments into the \hydrogen{} network than the \modH, but adds some investments into the \carbon{} network.

### 2023-11-26 report/submissions/journal/paper/paper.tex
rewrite section 3.1, incorporate comments from FN in Results 3.1

@@
 However, up to this point, it remains unclear how the two transport systems of hydrogen and carbon may complement or replace each other, particularly when it comes to their transport systems.
-A hydrogen network facilitates the transport of energy from renewable sources to regions with high energy demand, also enabling CU at the site of point-source capture.
+A hydrogen network facilitates the transport of energy from renewable sources to regions with geographically fixed hydrogen demand, such as for steelmaking, and also enables CU at the site of point-source capture.
 On the other hand, the carbon transport system allows for capturing and transporting carbon from areas with high emissions to regions with significant sequestration potential and abundant renewable sources, thereby enabling cost-effective CU.
@@
 
-The topology of the \hydrogen{} network is mirroring the topology of the electricity network.
+The topology of the \carbon{}, \hydrogen{} and gas network is identical to the topology of the electricity network, connecting all neighboring regions.
+Transport of biomass is considered as a variable cost component, and transport of liquid fuels like oil, methanol and Fischer-Tropsch (FT) fuels is ``copper-plated'', since transport costs per unit of energy are negligible due to their high energy density.
+Throughout this study, we focus on \carbon{} and \hydrogen{} networks because of their high relevance for infrastructure investments and thus public policy decisions.
+The electricity network and gas network are both included in the model with full geographical detail, but not further analyzed.
+Electricity networks are already in place, and we restrict further extension due to concerns of public acceptance.
+Gas networks also already exist, and will most likely experience decreased use in the future, removing any bottlenecks or constraints on the optimal system buildout or operation.
 
@@
 System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
-In the Baseline model, the absence of dedicated carbon and hydrogen transport networks requires the system to deploy carbon production technologies (DAC and CC at point sources) in the regions with carbon sequestration sites, and distributed amounts of hydrogen production technologies (mainly electrolysis) in each region to meet local hydrogen demand for industry and land transport fuel cells.
+In the Baseline model, the absence of dedicated carbon and hydrogen transport networks requires the system to deploy carbon production technologies (DAC and CC at point sources) in the regions with carbon sequestration sites, and distributed amounts of hydrogen production technologies (mainly electrolysis) in each region to meet local hydrogen demand for industry and fuel cells in heavy-duty land transport vehicles.
 For CU and its respective carbon and hydrogen demands, the system chooses the sites where on-site production of carbon, hydrogen as well as the inputs and by-product outputs (e.g., waste heat) of the synthesis process have the lowest cost (or highest value).
@@
 Examining the cost-reductions across the different models raises an important question: why is the hydrogen network more cost-effective than its carbon counterpart, and how does it achieve similar cost reductions as the hybrid model?
-To answer this question, it is crucial to understand the underlying mechanisms and geographical implications of the two transportation systems.
-Fig.~\ref{fig:operation_map_carbon} maps the average production, consumption and flow in the carbon sector for the \carbongrid{} model.
+An initial hypothesis may be that the cost assumptions for \carbon{}
+ and \hydrogen{} networks play a role.
+However, the amount of system costs spent on either network is substantially lower than the cost difference between the models (20 billion euros less total system costs in the \hydrogengrid{} model, while it spends 10 billion euros on the \hydrogen{} grid and the \carbongrid{} model spends 8 billion euros on the \carbon{} grid).
+Furthermore, we have conducted a sensitivity analysis with cost assumptions for the \carbon{} network reduced to 50\% of the default assumptions, and the results show that the \hydrogen{} network still outperforms the \carbon{} network in terms of total system costs.
+To find the reason for why the \hydrogen{} network achieves lower system costs than the \carbon{} network, it is crucial to understand the underlying mechanisms and geographical implications of the two transportation systems.
+Fig.~\ref{fig:co2-only-operation_map_carbon} and \ref{fig:co2-only-operation_map_hydrogen} map the average production, consumption and flow in the carbon and hydrogen sectors for the \carbongrid{} model.
 Fig.
-\ref{fig:operation_map_hydrogen} shows equivalent quantities in the hydrogen sector for the \hydrogengrid{} model.
-In each region, circles divide into contributions from technologies, and their sizes correspond proportionally to production on the left and consumption on the right.
-The region's colors represent the average price of the considered carrier.
+\ref{fig:h2-only-operation_map_carbon} and \ref{fig:h2-only-operation_map_hydrogen} show equivalent quantities for the \hydrogengrid{} model.
+In each region, circles represent the production (top half-circle) and consumption (bottom half-circle) of the respective energy carrier, divided into shares of each technology and the area sizes correspond to production and consumption volumes.
+Differences between production and consumption in a given node correspond to import or export volumes through the connected networks, represented as lines with their size and arrows indicating volumes and direction of average flows.
+Each region's color represents the demand-weighted average price of the considered carrier in that region.
 
-The two fundamental differences between the grid strategies are quickly evident: The \carbongrid{} model captures large amounts of carbon from bioenergy and industrial processes, especially in central and eastern Europe, at low cost.
-The carbon grid transports low-cost carbon to regions where it is sequestered in offshore caverns or processed into carbonaceous fuels together with green hydrogen available at low costs.
-End-use primarily depends on the quality of renewable resources, and thus the availability of low-cost hydrogen.
-In locations with relatively poor renewables and thus higher costs of hydrogen production, such as Italy, Greece, Northern Germany, and lcoations around the Baltic Sea, the transported carbon is predominantly sequestered.
-Conversely, in Spain and the United Kingdom, which both have high renewable potentials and low-cost hydrogen production, the transported carbon is combined with hydrogen and fed into CU processes, with additional carbon being captured via DAC also running on the great renewable resources.
+There are four notable dynamics or differences between the two grid strategies: 
+First, the production of FT fuels always moves to where low-cost carbon and low-cost hydrogen can be brought together.
+In the \carbongrid{} model, carbon is moved from Western Europe to the Iberian Peninsula, the British Isles and to Denmark to produce FT fuels centrally in few regions.
+In the \hydrogengrid{} model, hydrogen is transported in the opposite direction to produce FT fuels decentrally across all regions in Central Europe with locally captured \carbon{} from point sources.
+Second, the \carbongrid{} model captures more low-cost carbon from biomass, especially in Central and Eastern Europe, and less high-cost carbon from DAC in Greece and Southern Italy.
+Instead, the carbon grid transports carbon from Central and Eastern Europe to Greece and Italy, where it is sequestered, or feeds into CU together with locally produced, low-cost electrolytic hydrogen.
+The \hydrogengrid{} model directly operates electrolyzers and DAC in Greece and Southern Italy, with electricity from low-cost renewables, and thus avoids significant network buildouts in this region.
+Third, sequestration in the \hydrogengrid{} model occurs more decentralized, spread out across several regions with access to offshore sequestration sites, making use of point-source CC potentials in each of these regions.
+In the \carbongrid{} model, sequestration is clustered at fewer sequestration sinks, each collecting low-cost \carbon{} from neighboring regions.
+However, there is no centralization at only one or two sites, since all the sequestration sites offer the same costs and the system minimizes system-wide carbon transport distances to about eight to ten sequestration sites.
+And fourth, how the three types of CU are spread across the continent is driven by the costs of carbon and hydrogen (the latter mostly driven by the costs of renewable electricity) as well as by renewables capacity factors: Clearly, all carbonaceous fuel production requires sites where both carbon and hydrogen are available, and they moreover prefer sites with low carbon and hydrogen prices.
+Based on our model assumptions, FT fuel synthesis has very limited flexibility and needs to be run alost at baseload (ramp-down only possible to 90\% of maximum load).
+Thus, for FT fuel synthesis sites in Northern Europe with more electricity from wind power are preferred to sites in Southern Europe, powered mostly by solar.
+Methanathion and Methanolisation both have higher flexibility, being able to ramp down to 50\% load during the night, for example.
+Thus, these fuel production sites prefer the best renewable resource sites in Southern Europe.
+And finally, as methanation produces synthetic methane that needs to be fed into the gas grid, methanation prefers those sites with good interconnection to the existing gas network.
+Since Italy possesses a large gas network with easy interconnection to Central Europe, methanation prefers sites in Southern Italy (and Greece, which would be connected to Southern Italy) over the Iberian Peninsula, where Methanolisation is more prevalent.
 
 In summary, the implementation of a \carbon{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce low-cost hydrogen to make CU products (FT fuel, methanol, synthetic methane).
-In the absence of a \carbon{} network, sequestration sites first sequester all carbon from low-cost point sources, and deploy additional DAC facilities at the locations with low-cost renewables to capture the remaining necessary carbon for sequestration.
-Then, synfuel demand is met at the locations which offer the best mix of low-cost carbon from point sources and low-cost hydrogen from local renewables.
+DAC and electrolysis are deployed in the regions with the best renewable resources to meet carbonaceous fuel demands.
+In the absence of a \carbon{} network and with a \hydrogengrid{} on the other hand, regions with access to offshore sequestration sites capture all carbon from low-cost point sources and then sequester it.
+The system deploys additional DAC facilities at the locations with sequestration access and low-cost renewables to capture the remaining necessary carbon for sequestration.
+Carbonaceous fuel demand is met at the locations which offer the best mix of low-cost carbon from point sources and low-cost hydrogen from renewables, with some additional electrolysis and DAC required to fully meet the demand.
 % Sequestered carbon is nearly always transported before
-
-The \hydrogengrid{} on the other hand, transports low-cost hydrogen from regions with abundant renewable resources to regions with high energy demand and abundant emissions.
-The main transportation routes go from Spain and the United Kingdom to Central Europe.
-In these regions, industries use the hydrogen for demand in industry and fuel cell land transport, and to produce synfuels with locally captured carbon from point sources.
-This leads to a very decentralized synfuel production across the continent.
-Furthermore, the introduction of a hydrogen network enables to reduce both DAC production and FT fuels on the British Isles and Iberian Peninsula.
-In Great Britain an Ireland, the remaining local carbon is then sequestered, and all the hydrogen that previously went into FT fuel production is exported to Central Europe (see above).
-In Spain and Portugal, some DAC remains and the carbon sourced from DAC is used, along with locally captured carbon from point sources, to produce FT fuels (but less than without a hydrogen network).
-Surplus hydrogen from lower FT fuel production is exported to Central Europe.
 
@@
 A \hydrogen{} network favors a centralized electrolysis system at regions with abundant renewable resources and a decentralized CU system.
-These findings explain the advantage of the \hydrogen{} network in contrast to the \carbon{} network.
-Central Europe with its industrial clusters and dense population is characterized by a scarcity in hydrogen and a surplus and abundance of \carbon.
-While the \carbon{} network cost-efficiently resolves the surplus of carbon in Central Europe by exporting it to regions with low-cost renewable resources to produce synfuels, it needs to supply direct hydrogen demand in Central Europe (for industry and fuel cell land transport) with high-cost local hydrogen production.
-The hydrogen network resolves both the energy scarcity by supplying hydrogen demand in Central Europe with low-cost hydrogen imported from Spain and Great Britain, and the carbon surplus by transporting hydrogen to the sites with carbon available at lower costs in Central Europe, and producing FT fuels highly decentralized throughout all of Central Europe.
+
+Together with understanding the dynamics described above, we can highlight three cost advantages of the \hydrogen{} network in contrast to the \carbon{} network: Capturing carbon at the lowest-cost regions reduces 4 billion euros in spending on DAC.
+However, the \hydrogengrid{} model produces electrolytic hydrogen in the lowest-cost regions and in total spends 7 billion euros less on renewable energy (14 billion euros less on solar, and 7 billion euros more on wind) as well as 1 billion euros less on SMR.
+Furthermore, without a \hydrogen{} network, additional \hydrogen{} storage is needed in Spain at a cost of 5 billion euros to store hydrogen produced during the summer for utilization in FT fuel production during the winter.
+And finally, the \carbongrid{} model spends 7 billion euros more on gas boilers, gas plants, gas infrastructure, and methanation.
+% TODO: why??
+The costs of the respective networks, on the other hand, are almost equal: the \carbongrid{} model spends 8 billion euros on the \carbon{} network, while the \hydrogengrid{} model even spends 10 billion euros on the \hydrogen{} network.
+These factors add up to account for 14 billion euros of the difference of 20 billion euros in system costs between the two scenarios (785 billion euros for the \carbongrid{} and 765 billion euros for the \hydrogengrid{} model, see Fig.~\ref{fig:objective_heatmap}).
 
@@
 The Hybrid model combines the advantages of both hydrogen and carbon networks.
-While the topology of the \hydrogen{}-grid roughly corresponds to that of the \hydrogen grid model, the \carbon{}-grid plays more of a complementary role.
+While the topology of the \hydrogen{}-grid roughly corresponds to that of the \hydrogen{} grid model, the \carbon{}-grid plays more of a complementary role.
 Fig.~\ref{fig:capacity_maps} shows the capacities of the two grids in the hybrid model, with the production and transport capacities given in kt/h.
@@
 
-Similar to the \hydrogengrid model, the Hybrid model takes advantage of a large hydrogen network that enables hydrogen to be transported from centralized production sites in western regions such as Spain to decentralized FT production sites across the continent.
+Similar to the \hydrogengrid{} model, the Hybrid model takes advantage of a large hydrogen network that enables hydrogen to be transported from centralized production sites in western regions such as Spain to decentralized FT production sites across the continent.
 However, in the Hybrid model hydrogen in the UK is used locally to produce synfuels, and less hydrogen is transported to Central Europe.
@@
 This reduces the need for DAC in the system and the sequestration sites are shifted to the end points of the carbon networks.
-Second-order effects can be seen in the shift in FT production compared to the \hydrogengrid model.
+Second-order effects can be seen in the shift in FT production compared to the \hydrogengrid{} model.
 Regions that have both good renewable resources and sequestration potential could now forego sequestration and use local process emissions for FT production, as can be observed in Ireland.
-A more detailed of differences between the \hydrogengrid and the Hybrid model can be found in Fig.~.
+A more detailed of differences between the \hydrogengrid{} and the Hybrid model can be found in Fig.~.
 
@@
 
-Despite only small decrease in cost in comparison to the \hydrogengrid model, hybrid approach brings several advantages.
+Despite only small decrease in cost in comparison to the \hydrogengrid{} model, hybrid approach brings several advantages.
 These include a broader range of technologies contributing to its robustness, reduced reliance on DAC, which may be costlier than anticipated in the cost projections, and less land use for wind and solar due to the decreased necessity for DAC.
-The increase used of biomass with carbon capture, transport and sequestration leads to slight cost advantages in comparison to the \hydrogengrid model.
+The increase used of biomass with carbon capture, transport and sequestration leads to slight cost advantages in comparison to the \hydrogengrid{} model.
 
@@
     \includegraphics[width=\linewidth]{difference/h2-only-full/figures/90_nodes/cost_map.png}
-    \caption{Difference in cost investments between \hydrogengrid and Hybrid model.
-The left subfigure shows higher spendings per technology and region and transport system for the \hydrogengrid model, the right shows higher spendings in the Hydrid model.}
+    \caption{Difference in cost investments between \hydrogengrid{} and Hybrid model.
+The left subfigure shows higher spendings per technology and region and transport system for the \hydrogengrid{} model, the right shows higher spendings in the Hydrid model.}
     \label{fig:cost_map_difference}

### 2023-11-27 report/submissions/journal/paper/paper.tex
add combined operation map per carrier

--- 2023-11-26 report/submissions/journal/paper/paper.tex
@@
 \begin{abstract}
-    In future climate-neutral scenarios, the transport of hydrogen and carbon dioxide can play competing roles.
-    Hydrogen transport allows energy-intensive regions to import low-cost hydrogen.
-Carbon transport allows regions with carbon sinks such as carbon sequestration and synthetic fuel production to import low-cost carbon from distributed carbon capture sources.
-    Whether carbon and hydrogen transport networks are complementing or substituting each other remains unclear.
-This is especially relevant for the production of carbonaceous fuels such as Fischer-Tropsch fuels, methanol or synthetic methane, which are made from both hydrogen and carbon dioxide, and in most locations require at least one of these feedstocks to be imported.
-    To address this gap, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates transport, storage and sequestration for carbon dioxide, transport and storage for hydrogen, and renewable energy sources with a high spatial and temporal resolution considering all energy intensive sectors.
-    Our findings reveal that when comparing either-or-models, a system with an hydrogen network provides lower system costs than a system with a carbon network.
-The former enables the transportation of hydrogen from centralized production sites with low-cost renewable energy in the Iberian Peninsula and the British Isles to serve distributed hydrogen demands across the European continent.
-The latter facilitates carbon capture at lower costs from distributed point sources and reduces the need for higher cost direct air capture.
-However, the system with only a carbon network needs to meet distributed hydrogen demand across the continent by locally deploying electrolysis regardless of renewable energy costs.
-    In a hybrid model with both types of networks, the carbon network is a cost-effective complement to the hydrogen network facilitating point-source carbon capture at low costs and transport to sequestration or carbon utilization sites.
-In all models, additional demand for synthetic fuels is met at sites with favorable renewable energy resources that enable both electrolysis and direct air capture at low costs.
+    The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors, particularly in emissions-intensive industries like heavy manufacturing.
+While management of both \carbon{} and \hydrogen{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructures effectively interact with comprehensive carbon management technologies, including carbon capture, transport, use, and storage.
+In particular, the role of a carbon transport network as a complementing or substituting infrastructure to hydrogen transport remains unclear.
+To address this gap, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates carbon management technologies, hydrogen transport and storage, and renewable energy sources with a high spatial and temporal resolution considering all energy intensive sectors.
+    Our findings reveal that in a either-or scenario, the hydrogen network is superior to the carbon network as it enables the transportation cheap hydrogen inland to serve industrial processes and carbon capture and utilization at point sources.
+A carbon network facilitates point-source carbon capture and reduces the need for direct air capture, but leaves the scarcity of low-cost hydrogen in energy-intensive regions such as industrial clusters.
+    However, in a hybrid scenario, the carbon network is a cost-effective complement to the hydrogen network facilitating carbon capture at point sources and transport to sequestration sites.
     We show that theses findings holds true against the backdrop of a net-zero energy system as well as a scenario with net-negative emissions.
     Overall, our work demonstrates the cost-effectiveness of a multi-grid system that includes both hydrogen and carbon transport networks and power grid expansion to achieve climate neutrality in Europe.
-    The paper highlights the benefits of combining hydrogen and \carbon{} networks to achieve climate neutrality.
+    The paper underpins the need for a complementing hydrogen network and \carbon{} network to achieve climate neutrality.
 \end{abstract}
@@
 
-The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors.
-While management of both \carbon{} and \hydrogen{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructures effectively interact with comprehensive carbon management technologies, including carbon capture, transport, use, and storage.
-Hydrogen offers an efficient way to transport and store energy.
-Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, and synthetic fuels through carbon capture (CC) techniques, or harvested from the atmosphere using direct air capture (DAC).
-Additionally, it can be stored in geological formations, a process known as carbon sequestration (CS).
-In combination, \carbon{} and \hydrogen{} networks build the basis for climate-neutral fuels needed for aviation, shipping and industrial feedstocks, and are central to carbon utilization (CU) strategies.
+An integrated management of both hydrogen and carbon is considered a crucial component to achieve climate neutrality in energy systems including hard-to-abate sectors.
+Hydrogen offers an efficient way to transport and store energy from different sources.
+Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, and synthetic fuels through Carbon Capture (CC) techniques, or harvested directly from the atmosphere using Direct Air Capture (DAC).
+Additionally, it can be stored in geological formations, a process known as carbon capture and sequestration (CCS).
+In combination, \carbon{} and \hydrogen{} networks build the basis for climate-neutral fuels needed for aviation, shipping and industrial feedstocks, and are central to Power-to-X (PtX) and Carbon Capture and Utilization (CCU) strategies.
 
-Recently, policymakers and industry in Europe have started developing carbon management strategies~\cite{GermanyDevelopingStrategy2023,CarbonManagementStrategie}, planning infrastructure components~\cite{CONetz}, and committing on the first carbon utilization projects~\cite{EFuelsPilotPlant2022,OrstedAssumesFull,GROUNDBREAKINGEFUELPRODUCTION,DLREfuelsDLR}.
+Recently, policymakers and industry in Europe have been committing to both hydrogen and carbon management projects, planning infrastructure components, and developing business models for emerging sectors of the economy.
 With the European Union's goal of achieving climate neutrality by 2050, a wide range of programs, funding models and initiatives have been established in this regard~\cite{eu2023netzero,europeangreendeal,europeaninnovationfund}.
-Initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen Infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier, and some gas pipelines are already repurposed to transport hydrogen~\cite{RohrFreiFuer}.
+Initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier.
 At the same time, business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offer \carbon{} offtake on the other hand.
-For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.7~Gt of carbon capture from point sources per year, which represent roughly 50\% of all emissions.
+For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.4~Gt of carbon capture from point sources per year.
 In combination with large sequestration potentials as stated in~\cite{weiProposedGlobalLayout2021}, this highlights the vast potential for decarbonization.
-The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a transport and sequestration capacity of 1.5 Mt \carbon{} per year  to be operative in 2024, expanding to a targeted scale of 5 Mt per year of sequestration by 2030.
+The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a 1.5 Mt/a co2 transport and sequestration capacity to be operating in 2024, expanding to a targeted scale of 5 Mt/a sequestration by 2030.
 To this end, the Clean Air Task Force underlines the importance of a carbon transport system in Europe to facilitate the carbon economy~\cite{lockwoodEuropeanStrategyCarbon}.
 
-However, up to this point, it remains unclear how the two transport systems of hydrogen and carbon may complement or replace each other, particularly when it comes to their transport systems.
-A hydrogen network facilitates the transport of energy from renewable sources to regions with high energy demand, also enabling CU at the site of point-source capture.
-On the other hand, the carbon transport system allows for capturing and transporting carbon from areas with high emissions to regions with significant sequestration potential and abundant renewable sources, thereby enabling cost-effective CU.
-In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
-Such techno-economic models, in contrast for example to Integrated Assessment Models, can account for the spatial distribution of carbon sources and sinks which are crucial provide a holistic view of the energy system and its technological interactions.
+However, up to this point, it remains unclear how the two management systems of hydrogen and carbon may complement or replace each other, particularly when it comes to their transport systems.
+A hydrogen network facilitates the transport of energy from renewable sources to regions with high energy demand, also enabling CCU at the site of point-source capture.
+On the other hand, the carbon transport system allows for capturing and transporting carbon from areas with high emissions to regions with significant sequestration potential and abundant renewable sources, thereby enabling cost-effective CCU.
+In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,bec attiniCarbonDioxideCapture2022}.
+In contrast to Integrated Assessment Models, the underlying optimization tools account for the spatial distribution of carbon sources and sinks which are crucial provide a holistic view of the energy system and its technological interactions..
 The work in~\cite{neumannBenefitsHydrogenNetwork2022} examines the effect of a hydrogen network in Europe...
@@
 The publication by Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \carbon{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
-This limited sectoral scope cannot capture important dynamics of carbon management, since it neglects the sectors that will need to handle most \carbon{} in the future.
 Another comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
-Hoewever, this limited spatial scope fails to consider other sequestration sites and co-benefits from connecting the \carbon{} network to neighboring countries.
 To this end, it is often argued that \carbon{} pipelines are a mature technology with an expected high learning rate, given the wide-spread installations in US and Canada for enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
@@
 We explore competition between \carbon{} and \hydrogen{} networks by adding them separately.
-Our evaluation focuses on the transport of \carbon{} and \hydrogen{} through their respective networks on the European continent.
+Our evaluation focuses on the transport dynamics of \carbon{} and \hydrogen{} through their respective networks on the European continent.
 We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available sequestration potential.
@@
 The study is conducted on the basis of the open-source, capacity-expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
-The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, waste, agriculture, and transport sectors (including international aviation and shipping).
+The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, waste, and transport sectors.
 
@@
     \includegraphics[width=\linewidth]{baseline/figures/90_nodes/total-demand-bar.png}
-    \caption{Exogenous demand assumptions.
-The figure shows the total annual energy demands for each energy carrier which drive all model activities.
-Endogenous processes may lead to higher total production volumes of some carriers, e.g., demand in methanol requires more hydrogen and carbon as secondary (energy) inputs, which in turn need additional electricity to run electrolysis and carbon capture processes.}
-    % TODO: adjust labels to show that these are exogenous assumptions
+    \caption{Exogenous demand assumptions. The figure shows the total amount of yearly energy demands for each energy carrier and type of end-consumer considered in the model. }
     \label{fig:total-demand-bar}
@@
 Exogenous assumptions on energy demand and non-abatable emissions are taken from various sources~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023}.
-The energy demand for electricity, transport, biomass, heat and gas are defined per region and time-step.
-Land transport demand is exogenously divided between electric vehicles (85\%) and fuel cell vehicles (15\%), the latter representing demand for heavy duty land transport.
-Demands for kerosene for aviation, methanol for shipping, and naphtha for industry are aggregated in the system scope and kept constant throughout all time-steps.
-Heat demand is regionally subdivided into shares of urban, rural and industrial sites.
-We show the sum of all energy demands in Fig.~\ref{fig:total-demand-bar}.
-The system emits 633~Mt \carbon{} per year from industry, aviation, shipping and agriculture, 153~Mt of which are fossil-based process emissions.
+The energy demand for electricity, transport, biomass, heat and gas are defined per region and time-step with heat demand being regionally subdivided into shares of urban, rural and industrial sites.
+Demands for aviation kerosene, methanol for shipping and naphtha for industry are aggregated in the system scope and kept constant throughout all time-steps.
+We show the sum of all energy demands in \ref{fig:total-demand-bar}.
+The system emits of 633~Mt \carbon{} per year through non-abatable emissions from industry, aviation, shipping and agriculture into account, 153~Mt of which are fossil-based industrial emissions.
 Industrial energy demand and excess heat potentials are calculated per node on the basis of~\cite{hotmaps_industrial_db}.
@@
 %
-Endogenous model results include the expansion of renewable energy sources, storage technologies, and transmission capacities.
+Endogenous assumptions include the expansion of renewable energy sources, storage technologies, and transmission capacities.
 The model considers the various energy carriers like electricity, hydrogen, methan, methanol, liquid hydrocarbons and biomass, together with conversions technologies to convert these into each other.
 Carbon-neutral electricity is provided by wind, solar, biomass, hydro and nuclear power plants.
-Hydro and nuclear plant capacities cannot be extended.
 Weather-dependent power potentials for solar, wind and hydro are calculated from reanalysis and satellite data sets~\cite{hersbachERA5GlobalReanalysis2020,pfeifrothSurfaceRadiationData2017}  per region and time-stamp, using the open-source tool Atlite~\cite{hofmannAtliteLightweightPython2021}.
@@
 We restrict the electric transmission system's expansion to 20\% of its current capacity, acknowledging the challenges in inaugurating new transmission projects.
-For the use of biomass we consider only residual biomass products and no energy crops.
-We limit regional biomass use to the medium-level potentials from the ENPRESO database~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
+The regional use of biomass is restricted by its feasible potential derived from~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
 Inter-regional biomass transport is permitted with transport costs considered.
 
-In our model, hydrogen can be produced from electrolysis and steam methane reforming (SMR).
+Hydrogen can be gained from electrolysis and steam methane reforming (SMR).
 Geological distribution and potentials for \hydrogen{} storage in salt caverns are derived from~\cite{caglayanTechnicalPotentialSalt2020}.
 Re-electrification of hydrogen is possible via fuel cells.
-If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limit, considering costs for pipeline segments and compressors.
-Pipeline flows are modelled using net transfer capacities and without flow dynamics, pressure valves, or energy demand for compression.
-No retrofitting of gas pipelines is considered, potentially overestimating hydrogen network cost.
-
+If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limit.
 The topology of the \hydrogen{} network is mirroring the topology of the electricity network.
 
-Our model features three drop-in fuel production technologies for CU: methanation, methanolization, and Fischer-Tropsch synthesis.
+Our model features three drop-in fuel production technologies for carbon utilization: methanation, methanolization, and Fischer-Tropsch synthesis.
 The processed fuels are not limited in their total quantity of use or production.
-Methane is transported through the gas network, while methanol and FT fuel can be transferred inter-regionally without additional costs, since dense fuels have negligible transport costs per unit of energy.
-Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial process heat.
+Methane is transported through the gas network, while methanol and FT-synfuel can be transferred inter-regionally without additional costs.
+Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial gas needs.
 Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production, aviation kerosene, or agricultural machinery oil.
 
-To supply carbon needed for CS and CU, the system can choose to deploy carbon capture technology at various point sources (see below), or through DAC facilities.
+To supply carbon needed for CS and CU, the system can choose to deploy carbon capture technology at various point sources, or centralized DAC facilities.
 The concept of a ``merit order for captured carbon`` plays a pivotal role in optimizing and prioritizing the deployment of carbon capture technologies based on their economic feasibility.
@@
 Moving up the scale, gas used in industrial applications and biomass employed for industrial processes represent more costly yet viable options to capture carbon.
-DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand.
-Finally, based on our model results, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per ton of captured \carbon{}.
-Biogas input is a high-cost fuel which based on the endogenous modeling decisions is not used in large quantities.
-Thus, the marginal price of capturing additional \carbon{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the cost of captured \carbon{}.
+DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a higher cost level due to its current technological and operational expenses.
+Finally, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, typically incurs the highest costs in the merit order per ton of captured \carbon{} but offers a flexible (yet carbon emitting) energy source as a product.
 This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.
-If we consider the spatial aspect of distributed carbon capture potentials, the \carbon{} network can exploit comparative cost advantages between ``CO2 bidding zones'' to utilize the lowest-cost CC facilities across the continent.
-We assume a capture rate of 90\% for CC on process emissions, SMR, biogas-to-gas, as well as gas and biomass used in industry, and 95\% for CC on biomass and gas CHPs.
 
-To store carbon, we differentiate between short-term storage in steel tanks and long-term, irreversible sequestration in underground sequestration sites such as porous rock formations or depleted gas reservoirs.
-Costs for both options are included in the model.
-For carbon sequestration, we only consider offshore sites as potential sinks.
-We make this choice because offshore sequestration sites typically have larger capacity compared to onshore sequestration sites in saline aquifers and due to concerns over public safety for infrastructure near populated areas.
-Our estimates for carbon sequestration potentials are conservative, limiting total sequestration to 25~Mt per site and calculating annual storage availability over 25 years.
-Furthermore, we decide to cap the total sequestered \carbon{} to 200~Mt per year for our Net-Zero scenario.
-This is enough to offset the hardest-to-abate fossil process emissions and some limited slack of fossil fuels to be used where they can be most valuable.
-This is in order to avoid offsetting fossil emissions with carbon removal if they can technologically be avoided in the first place.
-All technology cost assumptions are taken for the year 2040 and sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
+We only consider offshore sites as potential sinks for carbon sequestration.
+This choice favors offshore storage due to its typically larger capacity compared to onshore and concerns over public safety for infrastructure near populated areas.
+Our estimates for storage potential are conservative, limiting total potential to 25~Mt per site, calculating annual storage availability over 25 years, and capping the total sequestered \carbon{} to 200~Mt/a.
+All technology cost assumptions for 2040 are sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
 
@@
 \end{itemize}
-Unless otherwise mentioned, our reference is the Net-Zero scenario. For the Net-Negative scenario, the cap of total carbon sequestration is adjusted from 200~Mt to 660~Mt per year accordingly.
+Unless otherwise mentioned, our reference is the Net-zero scenario.
+For 5\% Net-negative and 10\% Net-negative scenarios, the carbon sequestration potential is adjusted to 400~Mt/a and 600~Mt/a, respectively.
 
@@
 
-Focusing first on the Net-Zero scenario, we can highlight significant variations in system costs and technology deployment across the models (see Fig.~\ref{fig:cost_bar}).
+Focusing first on the net-zero scenario, we can highlight significant variations in system costs and technology deployment across the models (see Fig.~\ref{fig:cost_bar}).
 
@@
 While the `Baseline' model has neither a \carbon{} nor a \hydrogen{} network, `Hybrid' is allowed to expand both.
-"Gas Infrastructure" combines gas facilities for power and heat production,"\carbon{} Infrastrucure" combines transport, storage and sequestration, "H$_2$ Infrastructure" combines transport and storage.
+"Gas Infrastructure" combines gas facilities for power and heat production, "H$_2$ Infrastructure" combines H$_2$ production, transport and re-electrification.
 "Carbon Capture at Point Sources" combines all technologies with integrated carbon capture.}
@@
 In the Baseline model, the absence of dedicated carbon and hydrogen transport networks requires the system to deploy carbon production technologies (DAC and CC at point sources) in the regions with carbon sequestration sites, and small amounts of hydrogen production technologies (mainly electrolysis) in each region to meet local hydrogen demand for industry and land transport fuel cells.
-For CU and its respective carbon and hydrogen demands, the system chooses the sites where on-site production of carbon, hydrogen and the synthesis process has the lowest cost.
+For CU and its respective carbon and hydrogen demands, the system chooses the sites where on-site production of carbon, hydrogen and the synthesis process is cheapest.
 This leads to renewable resources and storage solutions being deployed not necessarily in the regions with the best resources or lowest costs for clean electricity generation, but where they are situated closer to the locations with carbon supply from CC at point sources, and to the locations of energy demand for electrolysers and DAC.
@@
 While DAC supplies around 50\% of the carbon needed in the Baseline model (400 Mt \carbon{}, see Fig.~\ref{fig:balance_captured_carbon}), in the \modCO{}, DAC is only used to supply around 20\% of the carbon (150 Mt \carbon{}).
-The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Fig.~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology has relatively lower costs than DAC.
-The \carbon{} transport network enables the transport of carbon from the locations of the lowest-cost captured carbon sources to the locations of the carbon sinks at CU and CS sites, reducing the need for higher cost DAC.
+The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Fig.~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology is relatively cheaper than DAC.
+The \carbon{} transport network enables the transport of carbon from the locations of the cheapest captured carbon supply sources to the locations of the carbon demand sinks at CU amd CS sites, reducing the need for expensive DAC.
 The improved carbon management infrastructure also allows for a more strategic deployment of renewables, particularly in areas that combine high quality renewables and high sequestration potentials, leading to system-wide cost savings in renewable generation and storage.
@@
     \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/captureshare_line.png}
-    \caption{Share of facilities with integrated carbon capture for different system setups for the Net-Zero scenario.}
+    \caption{Share of facilities with integrated carbon capture for different system setups for the net-zero emission scenario.}
     % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
@@
 The \hydrogen{}-Grid model underscores the pivotal role of hydrogen infrastructure.
-The creation of a hydrogen transport network enables the cost-effective delivery of hydrogen from regions with a surplus of low-cost renewable energy to high-demand industrial centers.
+The creation of a hydrogen transport network enables the cost-effective delivery of hydrogen from regions with a surplus of cheap renewable energy to high-demand industrial centers.
 This model shows a pronounced decrease in investments in solar generation and gas infrastructure.
@@
 CC technology on process emissions increases from 90\% to 100\%, on biomass CHPs from 60\% to 100\%, and on facilities using gas in industry from 50\% to 75\%.
-CC on biomass use in industry remains at 50\%, same as the \modCO{} (see Fig.~\ref{fig:captureshare_line}).
+CC on biomass use in industry remains at 50\%, same as the \modCO{} (see Figure~\ref{fig:captureshare_line}).
 This variation highlights how the implementation of \carbon{} and \hydrogen{} networks unlocks specific carbon capture potentials.
 
-While the cost reduction in the Hybrid model compared to the \hydrogen-Grid model is marginal, the allocation of costs reveals a propagation of reduced DAC facilities and increased CC at point sources.
+While the cost-reduction in the Hybrid model compared to the \hydrogen-Grid model is marginal, the allocation of costs reveals a propagation of reduced DAC facilities and increased CC at point sources.
 The Hybrid model relies slightly less on investments into the \hydrogen{} network than the \modH, but adds some investments into the \carbon{} network.
@@
 
-% \begin{figure*}[ht!]
-%     \centering
-%     \begin{subfigure}{0.9\linewidth}
-%         \centering
-%         \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_carbon.png}
-%         \caption{\carbon{} Sector.}
-%         \label{fig:operation_map_carbon}
-%     \end{subfigure}
-%     \begin{subfigure}{0.9\linewidth}
-%         \centering
-%         \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_hydrogen.png}
-%         \caption{\hydrogen{} Sector.}
-%         \label{fig:operation_map_hydrogen}
-%     \end{subfigure}
-%     \caption{Optimal operation per sector for a net-zero energy system in Europe with average production on the left and average consumption on the right for both, (a) the \carbon{} sector in the \carbon{}-Grid model and (b) the \hydrogen{} sector in the \hydrogen-Grid model.}
-%     \label{fig:operation_map}
-% \end{figure*}
-
 \begin{figure*}[ht!]
     \centering
-    \begin{subfigure}{0.49\linewidth}
+    \begin{subfigure}{0.9\linewidth}
         \centering
-        \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_single_carbon.png}
-        \caption{\carbon{} Sector in \carbongrid{} model.}
-        \label{fig:co2-only-operation_map_carbon}
+        \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_carbon.png}
+        \caption{\carbon{} Sector.}
+        \label{fig:operation_map_carbon}
     \end{subfigure}
-    \begin{subfigure}{0.49\linewidth}
+    \begin{subfigure}{0.9\linewidth}
         \centering
-        \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_single_hydrogen.png}
-        \caption{\hydrogen{} Sector in \carbongrid{} model.}
-        \label{fig:co2-only-operation_map_hydrogen}
-    \end{subfigure}
-    \begin{subfigure}{0.49\linewidth}
-        \centering
-        \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_single_carbon.png}
-        \caption{\hydrogen{} Sector in \hydrogengrid{} model.}
-        \label{fig:h2-only-operation_map_carbon}
-    \end{subfigure}
-    \begin{subfigure}{0.49\linewidth}
-        \centering
-        \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_single_hydrogen.png}
-        \caption{\hydrogen{} Sector in \hydrogengrid{} model.}
-        \label{fig:h2-only-operation_map_hydrogen}
+        \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_hydrogen.png}
+        \caption{\hydrogen{} Sector.}
+        \label{fig:operation_map_hydrogen}
     \end{subfigure}
@@
 The two fundamental differences between the grid strategies are quickly evident: The \carbongrid{} model captures large amounts of carbon from bioenergy and industrial processes, especially in central and eastern Europe, at low cost.
-The carbon grid transports low-cost carbon to regions where it is sequestered in offshore caverns or processed into carbonaceous fuels together with green hydrogen available at low costs.
-End-use primarily depends on the quality of renewable resources, and thus the availability of low-cost hydrogen.
+The carbon grid transports cheap carbon to regions where it is sequestered in offshore caverns or processed into synfuel, methane and methanol together with cheaply available green hydrogen.
+End-use primarily depends on the quality of renewable resources, and thus the availability of cheap hydrogen.
 In locations with relatively poor renewables and thus higher costs of hydrogen production, such as Italy, Greece, Northern Germany, and lcoations around the Baltic Sea, the transported carbon is predominantly sequestered.
-Conversely, in Spain and the United Kingdom, which both have high renewable potentials and low-cost hydrogen production, the transported carbon is combined with hydrogen and fed into CU processes, with additional carbon being captured via DAC also running on the great renewable resources.
+Conversely, in Spain and the United Kingdom, which both have high renewable potentials and cheap hydrogen production costs, the transported carbon is combined with hydrogen and fed into CU processes, with additional carbon being captured via DAC also running on the great renewable resources.
 
-In summary, the implementation of a \carbon{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce low-cost hydrogen to make CU products (FT fuel, methanol, synthetic methane).
-In the absence of a \carbon{} network, sequestration sites first sequester all carbon from low-cost point sources, and deploy additional DAC facilities at the locations with low-cost renewables to capture the remaining necessary carbon for sequestration.
-Then, synfuel demand is met at the locations which offer the best mix of low-cost carbon from point sources and low-cost hydrogen from local renewables.
+In summary, the implementation of a \carbon{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce cheap hydrogen to make CU products (FT fuel, methanol, synthetic methane).
+In the absence of a \carbon{} network, sequestration sites first sequester all carbon from cheap point sources, and deploy additional DAC facilities at the locations with cheap renewables to capture the remaining necessary carbon for sequestration.
+Then, synfuel demand is met at the locations which offer the best mix of cheap carbon from point sources and cheap hydrogen from local renewables.
 % Sequestered carbon is nearly always transported before
 
-The \hydrogengrid{} on the other hand, transports low-cost hydrogen from regions with abundant renewable resources to regions with high energy demand and abundant emissions.
+The \hydrogengrid{} on the other hand, transports cheap hydrogen from regions with abundant renewable resources to regions with high energy demand and abundant emissions.
 The main transportation routes go from Spain and the United Kingdom to Central Europe.
@@
 Central Europe with its industrial clusters and dense population is characterized by a scarcity in hydrogen and a surplus and abundance of \carbon.
-While the \carbon{} network cost-efficiently resolves the surplus of carbon in Central Europe by exporting it to regions with low-cost renewable resources to produce synfuels, it needs to supply direct hydrogen demand in Central Europe (for industry and fuel cell land transport) with high-cost local hydrogen production.
-The hydrogen network resolves both the energy scarcity by supplying hydrogen demand in Central Europe with low-cost hydrogen imported from Spain and Great Britain, and the carbon surplus by transporting hydrogen to the sites with carbon available at lower costs in Central Europe, and producing FT fuels highly decentralized throughout all of Central Europe.
+While the \carbon{} network cost-efficiently resolves the surplus of carbon in Central Europe by exporting it to regions with cheap renewable resources to produce synfuels, it needs to supply direct hydrogen demand in Central Europe (for industry and fuel cell land transport) with expensive local hydrogen production.
+The hydrogen network resolves both the energy scarcity by supplying hydrogen demand in Central Europe with cheap hydrogen imported from Spain and Great Britain, and the carbon surplus by transporting hydrogen to the sites with cheaper carbon available in Central Europe, and producing FT fuels highly decentralized throughout all of Central Europe.
 
@@
 These transport carbon from Romania and Bulgaria to Greece, from northern Italy to central Italy and from Hungary, the Czech Republic and Poland to the Baltic Sea.
-These enable the system to transport carbon available at low cost from inland biomass sources and sequester it (see Fig.~\ref{fig:captureshare_line}).
+These enable the system to transport cheaply available carbon from inland biomass sources and sequester it (see Fig.~\ref{fig:captureshare_line}).
 This reduces the need for DAC in the system and the sequestration sites are shifted to the end points of the carbon networks.
@@
 We now examine the case of moving from a net zero emissions target to a net emissions reduction target.
-Fig.~\ref{fig:net-negative_cost_bar} illustrates the shift in investments for the \hydrogengrid{} and the Hybrid model between the Net-Zero and the Net-Negative scenario.
+Fig.~\ref{fig:net-negative_cost_bar} illustrates the shift in investments for the \hydrogengrid{} and the Hybrid model between the net-zero and the net-negative scenario.
 

### 2023-11-27 workflow/scripts/plot-operation-map-single.py

@@ -1,0 +1,187 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Thu Nov 24 10:14:48 2022
+
+@author: fabian
+"""
+import os
+import warnings
+import seaborn as sns
+import pandas as pd
+import matplotlib.pyplot as plt
+import cartopy.crs as ccrs
+from pypsa.plot import add_legend_circles, add_legend_patches
+from matplotlib.gridspec import GridSpec
+from common import (
+    import_network,
+    mock_snakemake,
+    get_transmission_links,
+    get_carrier_transport,
+)
+import geopandas as gpd
+
+
+warnings.filterwarnings("ignore", category=UserWarning)
+
+if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
+    snakemake = mock_snakemake(
+        "plot_operation_map_single",
+        run="co2-only",
+        clusters=90,
+        ext="png",
+    )
+
+sns.set_theme(**snakemake.params["theme"])
+
+n = import_network(snakemake.input.network)
+regions = gpd.read_file(snakemake.input.onshore_regions).set_index("name")
+which = "operation"
+config = snakemake.config
+
+alpha = 1
+region_alpha = 0.6
+
+is_transport = get_transmission_links(n)
+transport_carriers = [
+    *n.links.carrier[is_transport].unique(),
+    *n.lines.carrier.unique(),
+]
+transport_carriers = n.carriers.nice_name[transport_carriers]
+
+# TODO: ensure this weird adjustment is not necessary anymore
+dac = n.links.index[n.links.carrier == "DAC"]
+n.links.loc[dac, ["bus0", "bus2"]] = n.links.loc[dac, ["bus2", "bus0"]].values
+n.links.loc[dac, ["", "bus2"]] = n.links.loc[dac, ["bus2", "bus0"]].values
+n.links_t.p0[dac], n.links_t.p2[dac] = n.links_t.p2[dac], n.links_t.p0[dac].values
+
+s = n.statistics
+
+run = snakemake.wildcards.run
+draw_legend = snakemake.params.settings["legend"].get(run, True)
+
+for kind, output in snakemake.output.items():
+
+    # Create a GridSpec object
+    # gs = GridSpec(1, 2, width_ratios=[20, 20])
+
+    figsize = snakemake.params.settings["figsize"]
+    figsize = figsize["single"] if draw_legend else figsize["single_wo_legend"]
+    fig, ax = plt.subplots(
[…]

### 2023-11-27 workflow/scripts/plot-operation-map.py

@@ -62,7 +62,7 @@ for kind, output in snakemake.output.items():
     # Create a GridSpec object
     gs = GridSpec(1, 2, width_ratios=[20, 20])
 
-    fig = plt.figure(figsize=snakemake.params.settings["figsize"])
+    fig = plt.figure(figsize=snakemake.params.settings["figsize"]["double"])
 
     # Create the first axes in the first column
     ax1 = fig.add_subplot(gs[0, 0], projection=ccrs.EqualEarth())

### 2023-11-27 report/submissions/journal/paper/paper.tex
adjust capture share plot

@@
 \begin{abstract}
-    The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors, particularly in emissions-intensive industries like heavy manufacturing.
-While management of both \carbon{} and \hydrogen{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructures effectively interact with comprehensive carbon management technologies, including carbon capture, transport, use, and storage.
-In particular, the role of a carbon transport network as a complementing or substituting infrastructure to hydrogen transport remains unclear.
+    In future climate-neutral scenarios, the transport of hydrogen and carbon dioxide can play competing roles.
+    Hydrogen transport allows energy-intensive regions to import low-cost hydrogen.
+Carbon transport allows regions with carbon sinks such as carbon sequestration and synthetic fuel production to import low-cost carbon from distributed carbon capture sources.
+    Whether carbon and hydrogen transport networks are complementing or substituting each other remains unclear.
+This is especially relevant for the production of carbonaceous fuels such as Fischer-Tropsch fuels, methanol or synthetic methane, which are made from both hydrogen and carbon dioxide, and in most locations require at least one of these feedstocks to be imported.
     To address this gap, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates carbon management technologies, hydrogen transport and storage, and renewable energy sources with a high spatial and temporal resolution considering all energy intensive sectors.
-    Our findings reveal that in a either-or scenario, the hydrogen network is superior to the carbon network as it enables the transportation cheap hydrogen inland to serve industrial processes and carbon capture and utilization at point sources.
-A carbon network facilitates point-source carbon capture and reduces the need for direct air capture, but leaves the scarcity of low-cost hydrogen in energy-intensive regions such as industrial clusters.
-    However, in a hybrid scenario, the carbon network is a cost-effective complement to the hydrogen network facilitating carbon capture at point sources and transport to sequestration sites.
+    Our findings reveal that when comparing either-or-models, a system with an hydrogen network provides lower system costs than a system with a carbon network.
+The former enables the transportation of hydrogen from centralized production sites with low-cost renewable energy in the Iberian Peninsula and the British Isles to serve distributed hydrogen demands across the European continent.
+The latter facilitates carbon capture at lower costs from distributed point sources and reduces the need for higher cost direct air capture.
+However, the system with only a carbon network needs to meet distributed hydrogen demand across the continent by locally deploying electrolysis regardless of renewable energy costs.
+    In a hybrid model with both types of networks, the carbon network is a cost-effective complement to the hydrogen network facilitating point-source carbon capture at low costs and transport to sequestration sites.
+In all models, additional demand for synthetic fuels is met at sites with favorable renewable energy resources that enable both electrolysis and direct air capture at low costs.
     We show that theses findings holds true against the backdrop of a net-zero energy system as well as a scenario with net-negative emissions.
@@
 
-An integrated management of both hydrogen and carbon is considered a crucial component to achieve climate neutrality in energy systems including hard-to-abate sectors.
-Hydrogen offers an efficient way to transport and store energy from different sources.
-Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, and synthetic fuels through Carbon Capture (CC) techniques, or harvested directly from the atmosphere using Direct Air Capture (DAC).
-Additionally, it can be stored in geological formations, a process known as carbon capture and sequestration (CCS).
-In combination, \carbon{} and \hydrogen{} networks build the basis for climate-neutral fuels needed for aviation, shipping and industrial feedstocks, and are central to Power-to-X (PtX) and Carbon Capture and Utilization (CCU) strategies.
+The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors.
+While management of both \carbon{} and \hydrogen{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructures effectively interact with comprehensive carbon management technologies, including carbon capture, transport, use, and storage.
+Hydrogen offers an efficient way to transport and store energy.
+Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, and synthetic fuels through Carbon Capture (CC) techniques, or harvested from the atmosphere using Direct Air Capture (DAC).
+Additionally, it can be stored in geological formations, a process known as carbon sequestration (CS).
+In combination, \carbon{} and \hydrogen{} networks build the basis for climate-neutral fuels needed for aviation, shipping and industrial feedstocks, and are central to Power-to-X (PtX) and Carbon Utilization (CU) strategies.
 
@@
     \includegraphics[width=\linewidth]{baseline/figures/90_nodes/total-demand-bar.png}
-    \caption{Exogenous demand assumptions. The figure shows the total amount of yearly energy demands for each energy carrier and type of end-consumer considered in the model. }
+    \caption{Exogenous demand assumptions.
+The figure shows the total annual energy demands for each energy carrier which drive all model activities.
+Endogenous processes may lead to higher total production volumes of some carriers, e.g.: Demand in methanol requires more hydrogen and carbon as secondary (energy) inputs, which in turn need additional electricity to run electrolysis and carbon capture processes.}
+    % TODO: adjust labels to show that these are exogenous assumptions
     \label{fig:total-demand-bar}
@@
 Methane is transported through the gas network, while methanol and FT-synfuel can be transferred inter-regionally without additional costs.
-Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial gas needs.
+Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial process heat.
 Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production, aviation kerosene, or agricultural machinery oil.
@@
 Moving up the scale, gas used in industrial applications and biomass employed for industrial processes represent more costly yet viable options to capture carbon.
-DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a higher cost level due to its current technological and operational expenses.
-Finally, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, typically incurs the highest costs in the merit order per ton of captured \carbon{} but offers a flexible (yet carbon emitting) energy source as a product.
+DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand.
+Finally, based on our model results, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per ton of captured \carbon{}.
+Biogas input is a high-cost fuel which based on the endogenous modeling decisions is not used in large quantities.
+Thus, the marginal price of capturing additional \carbon{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the cost of captured \carbon{}.
 This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.
@@
 While DAC supplies around 50\% of the carbon needed in the Baseline model (400 Mt \carbon{}, see Fig.~\ref{fig:balance_captured_carbon}), in the \modCO{}, DAC is only used to supply around 20\% of the carbon (150 Mt \carbon{}).
-The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Fig.~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology is relatively cheaper than DAC.
-The \carbon{} transport network enables the transport of carbon from the locations of the cheapest captured carbon supply sources to the locations of the carbon demand sinks at CU amd CS sites, reducing the need for expensive DAC.
+The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Fig.~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology has relatively lower costs than DAC.
+The \carbon{} transport network enables the transport of carbon from the locations of the lowest-cost captured carbon sources to the locations of the carbon sinks at CU and CS sites, reducing the need for higher cost DAC.
 The improved carbon management infrastructure also allows for a more strategic deployment of renewables, particularly in areas that combine high quality renewables and high sequestration potentials, leading to system-wide cost savings in renewable generation and storage.
@@
     \caption{Share of facilities with integrated carbon capture for different system setups for the net-zero emission scenario.}
+    % TODO: Add capacity factors into the figure
     % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
@@
 The two fundamental differences between the grid strategies are quickly evident: The \carbongrid{} model captures large amounts of carbon from bioenergy and industrial processes, especially in central and eastern Europe, at low cost.
-The carbon grid transports cheap carbon to regions where it is sequestered in offshore caverns or processed into synfuel, methane and methanol together with cheaply available green hydrogen.
-End-use primarily depends on the quality of renewable resources, and thus the availability of cheap hydrogen.
+The carbon grid transports low-cost carbon to regions where it is sequestered in offshore caverns or processed into carbonaceous fuels together with green hydrogen available at low costs.
+End-use primarily depends on the quality of renewable resources, and thus the availability of low-cost hydrogen.
 In locations with relatively poor renewables and thus higher costs of hydrogen production, such as Italy, Greece, Northern Germany, and lcoations around the Baltic Sea, the transported carbon is predominantly sequestered.
-Conversely, in Spain and the United Kingdom, which both have high renewable potentials and cheap hydrogen production costs, the transported carbon is combined with hydrogen and fed into CU processes, with additional carbon being captured via DAC also running on the great renewable resources.
+Conversely, in Spain and the United Kingdom, which both have high renewable potentials and low-cost hydrogen production, the transported carbon is combined with hydrogen and fed into CU processes, with additional carbon being captured via DAC also running on the great renewable resources.
 
-In summary, the implementation of a \carbon{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce cheap hydrogen to make CU products (FT fuel, methanol, synthetic methane).
-In the absence of a \carbon{} network, sequestration sites first sequester all carbon from cheap point sources, and deploy additional DAC facilities at the locations with cheap renewables to capture the remaining necessary carbon for sequestration.
-Then, synfuel demand is met at the locations which offer the best mix of cheap carbon from point sources and cheap hydrogen from local renewables.
+In summary, the implementation of a \carbon{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce low-cost hydrogen to make CU products (FT fuel, methanol, synthetic methane).
+In the absence of a \carbon{} network, sequestration sites first sequester all carbon from low-cost point sources, and deploy additional DAC facilities at the locations with low-cost renewables to capture the remaining necessary carbon for sequestration.
+Then, synfuel demand is met at the locations which offer the best mix of low-cost carbon from point sources and low-cost hydrogen from local renewables.
 % Sequestered carbon is nearly always transported before
 
-The \hydrogengrid{} on the other hand, transports cheap hydrogen from regions with abundant renewable resources to regions with high energy demand and abundant emissions.
+The \hydrogengrid{} on the other hand, transports low-cost hydrogen from regions with abundant renewable resources to regions with high energy demand and abundant emissions.
 The main transportation routes go from Spain and the United Kingdom to Central Europe.
@@
 Central Europe with its industrial clusters and dense population is characterized by a scarcity in hydrogen and a surplus and abundance of \carbon.
-While the \carbon{} network cost-efficiently resolves the surplus of carbon in Central Europe by exporting it to regions with cheap renewable resources to produce synfuels, it needs to supply direct hydrogen demand in Central Europe (for industry and fuel cell land transport) with expensive local hydrogen production.
-The hydrogen network resolves both the energy scarcity by supplying hydrogen demand in Central Europe with cheap hydrogen imported from Spain and Great Britain, and the carbon surplus by transporting hydrogen to the sites with cheaper carbon available in Central Europe, and producing FT fuels highly decentralized throughout all of Central Europe.
+While the \carbon{} network cost-efficiently resolves the surplus of carbon in Central Europe by exporting it to regions with low-cost renewable resources to produce synfuels, it needs to supply direct hydrogen demand in Central Europe (for industry and fuel cell land transport) with high-cost local hydrogen production.
+The hydrogen network resolves both the energy scarcity by supplying hydrogen demand in Central Europe with low-cost hydrogen imported from Spain and Great Britain, and the carbon surplus by transporting hydrogen to the sites with carbon available at lower costs in Central Europe, and producing FT fuels highly decentralized throughout all of Central Europe.
 

### 2023-11-28 report/submissions/journal/paper/paper.tex
add dedicated carbon and hydrogen map

@@
 The figure shows the total annual energy demands for each energy carrier which drive all model activities.
-Endogenous processes may lead to higher total production volumes of some carriers, e.g.: Demand in methanol requires more hydrogen and carbon as secondary (energy) inputs, which in turn need additional electricity to run electrolysis and carbon capture processes.}
+Endogenous processes may lead to higher total production volumes of some carriers, e.g., demand in methanol requires more hydrogen and carbon as secondary (energy) inputs, which in turn need additional electricity to run electrolysis and carbon capture processes.}
     % TODO: adjust labels to show that these are exogenous assumptions
@@
     \caption{Share of facilities with integrated carbon capture for different system setups for the net-zero emission scenario.}
-    % TODO: Add capacity factors into the figure
     % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}

### 2023-11-28 workflow/scripts/plot-captureshare-line.py

@@ -59,7 +59,7 @@ data = pd.concat(
     axis=1,
     keys=["Capacity Factor", "CC Share [%]"],
 ).reset_index()
-# %%
+
 fig, ax = plt.subplots(
     figsize=snakemake.params.settings["figsize"],
     layout="constrained",
@@ -99,3 +99,33 @@ plt.xticks(rotation=90)
 sns.despine()
 
 fig.savefig(snakemake.output[0], dpi=300)
+
+# ALTERATIVE PLOT
+# fig, ax = plt.subplots(
+#     figsize=snakemake.params.settings["figsize"],
+#     layout="constrained",
+# )
+# plot = sns.scatterplot(
+#     ax=ax,
+#     data=data,
+#     style="Model",
+#     y="CC Share [%]",
+#     x="Capacity Factor",
+#     hue="Carrier",
+#     palette=colors.to_dict(),
+#     legend="auto",
+# )
+# ax.legend(loc="center left", bbox_to_anchor=(1.0, 0.5), frameon=False, ncol=1)
+# x = data["Capacity Factor"]
+# y = data["CC Share [%]"]
+# ax.plot(
+#     [x.min(), x.max()],
+#     [y.min(), y.max()],
+#     ls="--",
+#     color="grey",
+#     alpha=0.5,
+#     zorder=-1,
+# )
+# ax.set_xlabel("Capacity Factor [%]")
+# sns.despine()
+# fig.savefig(snakemake.output[0], dpi=300)

### 2023-11-28 report/submissions/journal/paper/paper.tex
cosmetic changes to operation map plot

@@
 Hydrogen offers an efficient way to transport and store energy.
-Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, and synthetic fuels through Carbon Capture (CC) techniques, or harvested from the atmosphere using Direct Air Capture (DAC).
+Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, and synthetic fuels through carbon capture (CC) techniques, or harvested from the atmosphere using direct air capture (DAC).
 Additionally, it can be stored in geological formations, a process known as carbon sequestration (CS).
-In combination, \carbon{} and \hydrogen{} networks build the basis for climate-neutral fuels needed for aviation, shipping and industrial feedstocks, and are central to Power-to-X (PtX) and Carbon Utilization (CU) strategies.
+In combination, \carbon{} and \hydrogen{} networks build the basis for climate-neutral fuels needed for aviation, shipping and industrial feedstocks, and are central to carbon utilization (CU) strategies.
 
-Recently, policymakers and industry in Europe have been committing to both hydrogen and carbon management projects, planning infrastructure components, and developing business models for emerging sectors of the economy.
+Recently, policymakers and industry in Europe have started developing carbon management strategies~\cite{GermanyDevelopingStrategy2023,CarbonManagementStrategie}, planning infrastructure components~\cite{CONetz}, and committing on the first carbon utilization projects~\cite{EFuelsPilotPlant2022,OrstedAssumesFull,GROUNDBREAKINGEFUELPRODUCTION,DLREfuelsDLR}.
 With the European Union's goal of achieving climate neutrality by 2050, a wide range of programs, funding models and initiatives have been established in this regard~\cite{eu2023netzero,europeangreendeal,europeaninnovationfund}.
-Initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier.
+Initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen Infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier, and some gas pipelines are already repurposed to transport hydrogen~\cite{RohrFreiFuer}.
 At the same time, business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offer \carbon{} offtake on the other hand.
-For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.4~Gt of carbon capture from point sources per year.
+For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.7~Gt of carbon capture from point sources per year, which represent roughly 50\% of all emissions.
 In combination with large sequestration potentials as stated in~\cite{weiProposedGlobalLayout2021}, this highlights the vast potential for decarbonization.
-The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a 1.5 Mt/a co2 transport and sequestration capacity to be operating in 2024, expanding to a targeted scale of 5 Mt/a sequestration by 2030.
+The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a transport and sequestration capacity of 1.5 Mt \carbon{} per year  to be operative in 2024, expanding to a targeted scale of 5 Mt per year of sequestration by 2030.
 To this end, the Clean Air Task Force underlines the importance of a carbon transport system in Europe to facilitate the carbon economy~\cite{lockwoodEuropeanStrategyCarbon}.
 
-However, up to this point, it remains unclear how the two management systems of hydrogen and carbon may complement or replace each other, particularly when it comes to their transport systems.
-A hydrogen network facilitates the transport of energy from renewable sources to regions with high energy demand, also enabling CU at the site of point-source capture.
+However, up to this point, it remains unclear how the two transport systems of hydrogen and carbon may complement or replace each other, particularly when it comes to their transport systems.
+A hydrogen network facilitates the transport of energy from renewable sources to regions with geographically fixed hydrogen demand, such as for steelmaking, and also enables CU at the site of point-source capture.
 On the other hand, the carbon transport system allows for capturing and transporting carbon from areas with high emissions to regions with significant sequestration potential and abundant renewable sources, thereby enabling cost-effective CU.
 In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
-In contrast to Integrated Assessment Models, the underlying optimization tools account for the spatial distribution of carbon sources and sinks which are crucial provide a holistic view of the energy system and its technological interactions..
+Such techno-economic models, in contrast for example to Integrated Assessment Models, can account for the spatial distribution of carbon sources and sinks which are crucial provide a holistic view of the energy system and its technological interactions.
 The work in~\cite{neumannBenefitsHydrogenNetwork2022} examines the effect of a hydrogen network in Europe...
@@
 We restrict the electric transmission system's expansion to 20\% of its current capacity, acknowledging the challenges in inaugurating new transmission projects.
-The regional use of biomass is restricted by its feasible potential derived from~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
+For the use of biomass we consider only residual biomass products and no energy crops.
+We limit regional biomass use to the medium-level potentials from the ENPRESO database~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
 Inter-regional biomass transport is permitted with transport costs considered.
 
-Hydrogen can be gained from electrolysis and steam methane reforming (SMR).
+In our model, hydrogen can be produced from electrolysis and steam methane reforming (SMR).
 Geological distribution and potentials for \hydrogen{} storage in salt caverns are derived from~\cite{caglayanTechnicalPotentialSalt2020}.
 Re-electrification of hydrogen is possible via fuel cells.
-If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limit.
-The topology of the \hydrogen{} network is mirroring the topology of the electricity network.
+If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limit, considering costs for pipeline segments and compressors.
+Pipeline flows are modelled using net transfer capacities and without flow dynamics, pressure valves, or energy demand for compression.
+No retrofitting of gas pipelines is considered, potentially overestimating hydrogen network cost.
 
-Our model features three drop-in fuel production technologies for carbon utilization: methanation, methanolization, and Fischer-Tropsch synthesis.
+The topology of the \carbon{}, \hydrogen{} and gas network is identical to the topology of the electricity network, connecting all neighboring regions.
+Transport of biomass is considered as a variable cost component, and transport of liquid fuels like oil, methanol and Fischer-Tropsch (FT) fuels is ``copper-plated'', since transport costs per unit of energy are negligible due to their high energy density.
+Throughout this study, we focus on \carbon{} and \hydrogen{} networks because of their high relevance for infrastructure investments and thus public policy decisions.
+The electricity network and gas network are both included in the model with full geographical detail, but not further analyzed.
+Electricity networks are already in place, and we restrict further extension due to concerns of public acceptance.
+Gas networks also already exist, and will most likely experience decreased use in the future, removing any bottlenecks or constraints on the optimal system buildout or operation.
+
+Our model features three drop-in fuel production technologies for CU: methanation, methanolization, and Fischer-Tropsch synthesis.
 The processed fuels are not limited in their total quantity of use or production.
-Methane is transported through the gas network, while methanol and FT-synfuel can be transferred inter-regionally without additional costs.
+Methane is transported through the gas network, while methanol and FT fuel can be transferred inter-regionally without additional costs, since dense fuels have negligible transport costs per unit of energy.
 Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial process heat.
@@
 DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand.
-Finally, based on our model results, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per ton of captured \carbon{}.
+Finally, based on our model results, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per marginal ton of captured \carbon{}.
 Biogas input is a high-cost fuel which based on the endogenous modeling decisions is not used in large quantities.
-Thus, the marginal price of capturing additional \carbon{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the cost of captured \carbon{}.
+To the extent that biogas-to-gas facilities are used by the model to supply additional (carbon-neutral) gas as fuel, adding CC infrastructure incurs low costs (and thus ranges on the left-hand side of the merit order curve).
+However, the price of capturing additional \carbon{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the marginal cost of captured \carbon{}.
 This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.
+If we consider the spatial aspect of distributed carbon capture potentials, the \carbon{} network can exploit comparative cost advantages between ``CO2 bidding zones'' to utilize the lowest-cost CC facilities across the continent.
+We assume a capture rate of 90\% for CC on process emissions, SMR, biogas-to-gas, as well as gas and biomass used in industry, and 95\% for CC on biomass and gas CHPs.
 
-We only consider offshore sites as potential sinks for carbon sequestration.
-This choice favors offshore storage due to its typically larger capacity compared to onshore and concerns over public safety for infrastructure near populated areas.
-Our estimates for storage potential are conservative, limiting total potential to 25~Mt per site, calculating annual storage availability over 25 years, and capping the total sequestered \carbon{} to 200~Mt/a.
-All technology cost assumptions for 2040 are sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
+To store carbon, we differentiate between short-term storage in steel tanks and long-term, irreversible sequestration in underground sequestration sites such as porous rock formations or depleted gas reservoirs.
+Costs for both options are included in the model.
+For carbon sequestration, we only consider offshore sites as potential sinks.
+We make this choice because offshore sequestration sites typically have larger capacity compared to onshore sequestration sites in saline aquifers and due to concerns over public safety for infrastructure near populated areas.
+Our estimates for carbon sequestration potentials are conservative, limiting total sequestration to 25~Mt per site and calculating annual storage availability over 25 years.
+Furthermore, we decide to cap the total sequestered \carbon{} to 200~Mt per year for our Net-Zero scenario.
+This is enough to offset the hardest-to-abate fossil process emissions and some limited slack of fossil fuels to be used where they can be most valuable.
+This is in order to avoid offsetting fossil emissions with carbon removal if they can technologically be avoided in the first place.
+All technology cost assumptions are taken for the year 2040 and sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
 
@@
 System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
-In the Baseline model, the absence of dedicated carbon and hydrogen transport networks requires the system to deploy carbon production technologies (DAC and CC at point sources) in the regions with carbon sequestration sites, and small amounts of hydrogen production technologies (mainly electrolysis) in each region to meet local hydrogen demand for industry and land transport fuel cells.
-For CU and its respective carbon and hydrogen demands, the system chooses the sites where on-site production of carbon, hydrogen and the synthesis process has the lowest cost.
-This leads to renewable resources and storage solutions being deployed not necessarily in the regions with the best resources or lowest costs for clean electricity generation, but where they are situated closer to the locations with carbon supply from CC at point sources, and to the locations of energy demand for electrolysers and DAC.
-Thus, total system costs in the Baseline model are highest, while the models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs.
+In the Baseline model, the absence of dedicated carbon and hydrogen transport networks requires the system to deploy carbon production technologies (DAC and CC at point sources) in the regions with carbon sequestration sites, and distributed amounts of hydrogen production technologies (mainly electrolysis) in each region to meet local hydrogen demand for industry and fuel cells in heavy-duty land transport vehicles.
+For CU and its respective carbon and hydrogen demands, the system chooses the sites where on-site production of carbon, hydrogen as well as the inputs and by-product outputs (e.g., waste heat) of the synthesis process have the lowest cost (or highest value).
+This leads to renewable resources and storage solutions being deployed not necessarily in the regions with the best resources or lowest costs for clean electricity generation, but where they are situated closer to the endogenously defined locations of energy demand for electrolysers and DAC.
+Thus, total annual system costs in the Baseline model are highest, totalling \label{}803 billion euros (see Fig.~\ref{fig:objective_heatmap}).
+The models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs, with around \label{}2\% cost reductions (\label{}781 billion euros) for the \carbon{} network, and around \label{}4\% cost reductions for the \hydrogen{} network and the hybrid scenario (\label{}765 and \label{}764 billion euros, respectively).
 
 We can now closer examine the cost reductions and shifts in cost allocation when shifting from the Baseline model to the other three network models.
-The implementation of a \carbon{} transport network (\textit{\carbon{}~Grid}) facilitates efficient point-source carbon capture, substantially reducing the dependency on DAC technologies.
-While DAC supplies around 50\% of the carbon needed in the Baseline model (400 Mt \carbon{}, see Fig.~\ref{fig:balance_captured_carbon}), in the \modCO{}, DAC is only used to supply around 20\% of the carbon (150 Mt \carbon{}).
+The implementation of a \carbon{} transport network (\textit{\carbon{}~Grid}) facilitates carbon capture from the point sources with the lowest costs across regions, substantially reducing the dependency on DAC.
+While DAC supplies around 50\% of the carbon needed in the Baseline model (400 Mt \carbon{}, see Fig.~\ref{fig:balance_captured_carbon}), in the \modCO{}, DAC is only used to supply around 20\% of the carbon handled by the system (150 Mt \carbon{}).
 The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Fig.~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology has relatively lower costs than DAC.
 The \carbon{} transport network enables the transport of carbon from the locations of the lowest-cost captured carbon sources to the locations of the carbon sinks at CU and CS sites, reducing the need for higher cost DAC.
-The improved carbon management infrastructure also allows for a more strategic deployment of renewables, particularly in areas that combine high quality renewables and high sequestration potentials, leading to system-wide cost savings in renewable generation and storage.
+The option to transport carbon also allows for the deployment of renewables in areas that combine high quality renewable resources and high sequestration potentials, leading to system-wide cost savings in renewable generation and storage.
 
@@
     \centering
-    \begin{subfigure}{0.49\linewidth}
-        \centering
-        \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_single_carbon.png}
-        \caption{\carbon{} Sector in \carbongrid{} model.}
-        \label{fig:co2-only-operation_map_carbon}
-    \end{subfigure}
-    \begin{subfigure}{0.49\linewidth}
-        \centering
-        \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_single_hydrogen.png}
-        \caption{\hydrogen{} Sector in \carbongrid{} model.}
-        \label{fig:co2-only-operation_map_hydrogen}
-    \end{subfigure}
-    \begin{subfigure}{0.49\linewidth}
-        \centering
-        \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_single_carbon.png}
-        \caption{\hydrogen{} Sector in \hydrogengrid{} model.}
-        \label{fig:h2-only-operation_map_carbon}
-    \end{subfigure}
-    \begin{subfigure}{0.49\linewidth}
-        \centering
-        \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_single_hydrogen.png}
-        \caption{\hydrogen{} Sector in \hydrogengrid{} model.}
-        \label{fig:h2-only-operation_map_hydrogen}
-    \end{subfigure}
+    \includegraphics[width=\linewidth]{comparison/single-technologies/figures/90_nodes/operation_map_dedicated.png}
     \caption{Optimal operation per sector for a net-zero energy system in Europe with average production on the left and average consumption on the right for both, (a) the \carbon{} sector in the \carbon{}-Grid model and (b) the \hydrogen{} sector in the \hydrogen-Grid model.}

### 2023-11-28 report/submissions/journal/paper/word-count.tex
cosmetic changes to operation map plot

--- 2023-11-27 report/submissions/journal/paper/paper.tex
@@
-\documentclass[twocolumn]{article}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{cuted}  % Add the cuted package
-\usepackage{caption}
-\captionsetup{font=small}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{tabularx, multirow}
-\usepackage{fancyhdr,lipsum}
-\usepackage{subcaption}
-
-\usepackage[%
-backend=biber,bibencoding=utf8, %instead of bibtex
-language=auto,
-style=ieee,
-sorting=none, % nyt for name, year, title
-maxbibnames=10, % default: 3, et al.
-%backref=true,%
-natbib=true % natbib compatibility mode (\citep and~\citet still work)
-]{biblatex}
-\bibliography{../../../references.bib}
-
-%define approx proportional
-\def\app#1#2{%
-  \mathrel{%
-    \setbox0=\hbox{$#1\sim$}%
-    \setbox2=\hbox{%
-      \rlap{\hbox{$#1\propto$}}%
-      \lower1.1\ht0\box0%
-    }%
-    \raise0.25\ht2\box2%
-  }%
-}
-\def\approxprop{\mathpalette\app\relax}
-
-% make abbreviation for co2
-\newcommand{\carbon}{CO$_2$}
-\newcommand{\hydrogen}{H$_2$}
-\newcommand{\carbongrid}{\carbon{}--Grid}
-\newcommand{\hydrogengrid}{\hydrogen{}--Grid}
-\newcommand{\modBase}{Baseline model}
-\newcommand{\modCO}{CO$_2$-Grid model}
-\newcommand{\modH}{H$_2$-Grid model}
-\newcommand{\modHybrid}{Hybrid model}
-
-\graphicspath{
-    % {paper-figures}
-    {figures}
-    {../../../../results/}
-    {results/}
-}
-
-\begin{document}
-
-% FORMATTING: https://www.nature.com/nenergy/submission-guidelines/aip-and-formatting
-
-\title{From Net-Zero to Net-Negative: Assessing \hydrogen{} and \carbon{} Network Strategies in Europe}
-% \title{Synthetic fuels in Europe: Transport Hydrogen to Carbon, or Carbon to Hydrogen?}
-
-\author{
-    Fabian Hofmann, Christoph Tries, Fabian Neumann, Lisa Zeyen, Tom Brown \\
-    \textit{Institute of Energy Technology} \\
-    \textit{Technical University of Berlin}\\
-    Berlin, Germany \\
-    m.hofmann@tu-berlin.de
-}
-
-\maketitle
-
-\begin{abstract}
     In future climate-neutral scenarios, the transport of hydrogen and carbon dioxide can play competing roles.
@@
     The paper underpins the need for a complementing hydrogen network and \carbon{} network to achieve climate neutrality.
-\end{abstract}
-
-\section{Introduction}
-
-The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors.
-While management of both \carbon{} and \hydrogen{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructures effectively interact with comprehensive carbon management technologies, including carbon capture, transport, use, and storage.
-Hydrogen offers an efficient way to transport and store energy.
-Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, and synthetic fuels through Carbon Capture (CC) techniques, or harvested from the atmosphere using Direct Air Capture (DAC).
-Additionally, it can be stored in geological formations, a process known as carbon sequestration (CS).
-In combination, \carbon{} and \hydrogen{} networks build the basis for climate-neutral fuels needed for aviation, shipping and industrial feedstocks, and are central to Power-to-X (PtX) and Carbon Utilization (CU) strategies.
-
-Recently, policymakers and industry in Europe have been committing to both hydrogen and carbon management projects, planning infrastructure components, and developing business models for emerging sectors of the economy.
-With the European Union's goal of achieving climate neutrality by 2050, a wide range of programs, funding models and initiatives have been established in this regard~\cite{eu2023netzero,europeangreendeal,europeaninnovationfund}.
-Initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier.
-At the same time, business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offer \carbon{} offtake on the other hand.
-For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.4~Gt of carbon capture from point sources per year.
-In combination with large sequestration potentials as stated in~\cite{weiProposedGlobalLayout2021}, this highlights the vast potential for decarbonization.
-The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a 1.5 Mt/a co2 transport and sequestration capacity to be operating in 2024, expanding to a targeted scale of 5 Mt/a sequestration by 2030.
-To this end, the Clean Air Task Force underlines the importance of a carbon transport system in Europe to facilitate the carbon economy~\cite{lockwoodEuropeanStrategyCarbon}.
-
-However, up to this point, it remains unclear how the two management systems of hydrogen and carbon may complement or replace each other, particularly when it comes to their transport systems.
-A hydrogen network facilitates the transport of energy from renewable sources to regions with high energy demand, also enabling CU at the site of point-source capture.
-On the other hand, the carbon transport system allows for capturing and transporting carbon from areas with high emissions to regions with significant sequestration potential and abundant renewable sources, thereby enabling cost-effective CU.
-In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
-In contrast to Integrated Assessment Models, the underlying optimization tools account for the spatial distribution of carbon sources and sinks which are crucial provide a holistic view of the energy system and its technological interactions..
-The work in~\cite{neumannBenefitsHydrogenNetwork2022} examines the effect of a hydrogen network in Europe...
-(extend on hydrogen literature)
-The publication by Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \carbon{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
-Another comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
-To this end, it is often argued that \carbon{} pipelines are a mature technology with an expected high learning rate, given the wide-spread installations in US and Canada for enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
-% However, the models are often limited with regard to both geographical scope and detail.
-While representing a single country with spatial resolution may neglect synergies of international cooperation, a coarse grained representation of multiple countries may neglect important geographical properties.
-
-To our knowledge, no study has yet considered the co-optimization and comprehensive assessment of both \carbon{} and \hydrogen{} networks in a fully sector-coupled energy system.
-However, we argue that such an assessment is strongly needed to avoid suboptimal investments and to identify synergies between hydrogen and carbon management technologies.
-In this paper, we present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies.
-The study is conducted using the PyPSA-Eur energy system model and encompasses all relevant energy sectors.
-We explore competition between \carbon{} and \hydrogen{} networks by adding them separately.
-Our evaluation focuses on the transport dynamics of \carbon{} and \hydrogen{} through their respective networks on the European continent.
-We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available sequestration potential.
-
-\section{Methodology}
-\label{sec:methodology}
-
-The study is conducted on the basis of the open-source, capacity-expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
-The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, waste, and transport sectors.
-
-\begin{figure}
-    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/total-demand-bar.png}
-    \caption{Exogenous demand assumptions.
-The figure shows the total annual energy demands for each energy carrier which drive all model activities.
-Endogenous processes may lead to higher total production volumes of some carriers, e.g.: Demand in methanol requires more hydrogen and carbon as secondary (energy) inputs, which in turn need additional electricity to run electrolysis and carbon capture processes.}
-    % TODO: adjust labels to show that these are exogenous assumptions
-    \label{fig:total-demand-bar}
-\end{figure}
-%
-In our configuration, the model's time horizon spans one year with a temporal resolution of 3 hours and a spatial resolution of 90 regions.
-Each of the regions consists of a complex subsystem with technologies for supplying, converting, storing and transporting energy.
-Exogenous assumptions on energy demand and non-abatable emissions are taken from various sources~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023}.
-The energy demand for electricity, transport, biomass, heat and gas are defined per region and time-step.
-Land transport demand is exogenously divided between electric vehicles (85\%) and fuel cell vehicles (15\%), the latter representing demand for heavy duty land transport.
-Demands for kerosene for aviation, methanol for shipping, and naphtha for industry are aggregated in the system scope and kept constant throughout all time-steps.
-Heat demand is regionally subdivided into shares of urban, rural and industrial sites.
-We show the sum of all energy demands in Fig.~\ref{fig:total-demand-bar}.
-The system emits 633~Mt \carbon{} per year from industry, aviation, shipping and agriculture, 153~Mt of which are fossil-based process emissions.
-Industrial energy demand and excess heat potentials are calculated per node on the basis of~\cite{hotmaps_industrial_db}.
-% use https://pypsa-eur.readthedocs.io/en/latest/licenses.html
-%
-Endogenous model results include the expansion of renewable energy sources, storage technologies, and transmission capacities.
-The model considers the various energy carriers like electricity, hydrogen, methan, methanol, liquid hydrocarbons and biomass, together with conversions technologies to convert these into each other.
-Carbon-neutral electricity is provided by wind, solar, biomass, hydro and nuclear power plants.
-Weather-dependent power potentials for solar, wind and hydro are calculated from reanalysis and satellite data sets~\cite{hersbachERA5GlobalReanalysis2020,pfeifrothSurfaceRadiationData2017}  per region and time-stamp, using the open-source tool Atlite~\cite{hofmannAtliteLightweightPython2021}.
-Solar and wind power can be expanded in alignment with eligible land-use restrictions calculated on the basis of~\cite{eeaCorineLandCover2012,eeaNatura2000Data2016}.
-We restrict the electric transmission system's expansion to 20\% of its current capacity, acknowledging the challenges in inaugurating new transmission projects.
-The regional use of biomass is restricted by its feasible potential derived from~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
-Inter-regional biomass transport is permitted with transport costs considered.
-
-Hydrogen can be gained from electrolysis and steam methane reforming (SMR).
-Geological distribution and potentials for \hydrogen{} storage in salt caverns are derived from~\cite{caglayanTechnicalPotentialSalt2020}.
-Re-electrification of hydrogen is possible via fuel cells.
-If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limit.
-The topology of the \hydrogen{} network is mirroring the topology of the electricity network.
-
-Our model features three drop-in fuel production technologies for carbon utilization: methanation, methanolization, and Fischer-Tropsch synthesis.
-The processed fuels are not limited in their total quantity of use or production.
-Methane is transported through the gas network, while methanol and FT-synfuel can be transferred inter-regionally without additional costs.
-Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial process heat.
-Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production, aviation kerosene, or agricultural machinery oil.
-
-To supply carbon needed for CS and CU, the system can choose to deploy carbon capture technology at various point sources, or centralized DAC facilities.
-The concept of a ``merit order for captured carbon`` plays a pivotal role in optimizing and prioritizing the deployment of carbon capture technologies based on their economic feasibility.
-This merit order ranks all CC technologies according to their relative costs to capture a ton of carbon (see Fig.~X).
-At the lower end of the cost spectrum, CC technologies applied to process emissions, such as those from cement, offer a cost-effective starting point.
-Following this, biomass combined heat and power (CHP) systems provide a dual benefit of energy production and carbon capture.
-Moving up the scale, gas used in industrial applications and biomass employed for industrial processes represent more costly yet viable options to capture carbon.
-DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand.
-Finally, based on our model results, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per ton of captured \carbon{}.
-Biogas input is a high-cost fuel which based on the endogenous modeling decisions is not used in large quantities.
-Thus, the marginal price of capturing additional \carbon{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the cost of captured \carbon{}.
-This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.
-
-We only consider offshore sites as potential sinks for carbon sequestration.
-This choice favors offshore storage due to its typically larger capacity compared to onshore and concerns over public safety for infrastructure near populated areas.
-Our estimates for storage potential are conservative, limiting total potential to 25~Mt per site, calculating annual storage availability over 25 years, and capping the total sequestered \carbon{} to 200~Mt/a.
-All technology cost assumptions for 2040 are sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
-
-Addressing climate targets, we define two \carbon{} target scenarios:
-\begin{itemize}
-    \item[] \textit{Net--Zero}: aligning with the EU's 2050 emission targets. Also denoted as NZ.
-    \item[] \textit{Net--Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \carbon{} annually. Also denoted as NN.
-\end{itemize}
-Unless otherwise mentioned, our reference is the Net-zero scenario.
-For 5\% Net-negative and 10\% Net-negative scenarios, the carbon sequestration potential is adjusted to 400~Mt/a and 600~Mt/a, respectively.
-
-Beyond \carbon{} targets, we consider four expansion scenarios, referred to as models:
-\begin{itemize}
-    \item[] \textit{Baseline:} Neither \carbon{} nor \hydrogen{} networks are constructed.
-    \item[] \textit{\carbongrid:} Only the \carbon{} network is developed.
-    \item[] \textit{\hydrogengrid:} Only the \hydrogen{} network is developed.
-    \item[] \textit{Hybrid:} Both \carbon{} and \hydrogen{} networks are developed.
-\end{itemize}
-
-\section{Results}
-\label{sec:results}
-
-Focusing first on the net-zero scenario, we can highlight significant variations in system costs and technology deployment across the models (see Fig.~\ref{fig:cost_bar}).
-
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/cost_bar.png}
-    \caption[short]{Total annual system cost for the European energy system for the different models subdivided into groups of technologies.
-While the `Baseline' model has neither a \carbon{} nor a \hydrogen{} network, `Hybrid' is allowed to expand both.
-"Gas Infrastructure" combines gas facilities for power and heat production,"\carbon{} Infrastrucure" combines transport, storage and sequestration, "H$_2$ Infrastructure" combines transport and storage.
-"Carbon Capture at Point Sources" combines all technologies with integrated carbon capture.}
-    \label{fig:cost_bar}
-\end{figure}
-
-System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
-In the Baseline model, the absence of dedicated carbon and hydrogen transport networks requires the system to deploy carbon production technologies (DAC and CC at point sources) in the regions with carbon sequestration sites, and small amounts of hydrogen production technologies (mainly electrolysis) in each region to meet local hydrogen demand for industry and land transport fuel cells.
-For CU and its respective carbon and hydrogen demands, the system chooses the sites where on-site production of carbon, hydrogen and the synthesis process has the lowest cost.
-This leads to renewable resources and storage solutions being deployed not necessarily in the regions with the best resources or lowest costs for clean electricity generation, but where they are situated closer to the locations with carbon supply from CC at point sources, and to the locations of energy demand for electrolysers and DAC.
-Thus, total system costs in the Baseline model are highest, while the models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs.
-
-We can now closer examine the cost reductions and shifts in cost allocation when shifting from the Baseline model to the other three network models.
-The implementation of a \carbon{} transport network (\textit{\carbon{}~Grid}) facilitates efficient point-source carbon capture, substantially reducing the dependency on DAC technologies.
-While DAC supplies around 50\% of the carbon needed in the Baseline model (400 Mt \carbon{}, see Fig.~\ref{fig:balance_captured_carbon}), in the \modCO{}, DAC is only used to supply around 20\% of the carbon (150 Mt \carbon{}).
-The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Fig.~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology has relatively lower costs than DAC.
-The \carbon{} transport network enables the transport of carbon from the locations of the lowest-cost captured carbon sources to the locations of the carbon sinks at CU and CS sites, reducing the need for higher cost DAC.
-The improved carbon management infrastructure also allows for a more strategic deployment of renewables, particularly in areas that combine high quality renewables and high sequestration potentials, leading to system-wide cost savings in renewable generation and storage.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/captureshare_line.png}
-    \caption{Share of facilities with integrated carbon capture for different system setups for the net-zero emission scenario.}
-    % TODO: Add capacity factors into the figure
-    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
-    \label{fig:captureshare_line}
-\end{figure}%
-
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/energy_balance_bar_carbon.png}
-    \caption{Balance of captured carbon across all system setups. Positive values indicate carbon capture, negative values indicate carbon consumption.}
-    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
-    \label{fig:balance_captured_carbon}
-\end{figure}
-
-The \hydrogen{}-Grid model underscores the pivotal role of hydrogen infrastructure.
-The creation of a hydrogen transport network enables the cost-effective delivery of hydrogen from regions with a surplus of low-cost renewable energy to high-demand industrial centers.
-This model shows a pronounced decrease in investments in solar generation and gas infrastructure.
-The \hydrogen{} network reduces DAC similarly to the \carbon{} network, but still produces 200 Mt of carbon from DAC.
-CC technology on process emissions increases from 90\% to 100\%, on biomass CHPs from 60\% to 100\%, and on facilities using gas in industry from 50\% to 75\%.
-CC on biomass use in industry remains at 50\%, same as the \modCO{} (see Fig.~\ref{fig:captureshare_line}).
-This variation highlights how the implementation of \carbon{} and \hydrogen{} networks unlocks specific carbon capture potentials.
-
-While the cost reduction in the Hybrid model compared to the \hydrogen-Grid model is marginal, the allocation of costs reveals a propagation of reduced DAC facilities and increased CC at point sources.
-The Hybrid model relies slightly less on investments into the \hydrogen{} network than the \modH, but adds some investments into the \carbon{} network.
-Still, investments into the \hydrogen{} network outweigh investments into the \carbon{} network by a factor of four.
-
-\subsection*{Superiority of \hydrogen{} network}
-
-\begin{figure*}[ht!]
-    \centering
-    \begin{subfigure}{0.9\linewidth}
-        \centering
-        \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_carbon.png}
-        \caption{\carbon{} Sector.}
-        \label{fig:operation_map_carbon}
-    \end{subfigure}
-    \begin{subfigure}{0.9\linewidth}
-        \centering
-        \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_hydrogen.png}
-        \caption{\hydrogen{} Sector.}
-        \label{fig:operation_map_hydrogen}
-    \end{subfigure}
-    \caption{Optimal operation per sector for a net-zero energy system in Europe with average production on the left and average consumption on the right for both, (a) the \carbon{} sector in the \carbon{}-Grid model and (b) the \hydrogen{} sector in the \hydrogen-Grid model.}
-    \label{fig:operation_map}
-\end{figure*}
-
-Examining the cost-reductions across the different models raises an important question: why is the hydrogen network more cost-effective than its carbon counterpart, and how does it achieve similar cost reductions as the hybrid model?
-To answer this question, it is crucial to understand the underlying mechanisms and geographical implications of the two transportation systems.
-Fig.~\ref{fig:operation_map_carbon} maps the average production, consumption and flow in the carbon sector for the \carbongrid{} model.
-Fig.
-\ref{fig:operation_map_hydrogen} shows equivalent quantities in the hydrogen sector for the \hydrogengrid{} model.
-In each region, circles divide into contributions from technologies, and their sizes correspond proportionally to production on the left and consumption on the right.
-The region's colors represent the average price of the considered carrier.
-
-The two fundamental differences between the grid strategies are quickly evident: The \carbongrid{} model captures large amounts of carbon from bioenergy and industrial processes, especially in central and eastern Europe, at low cost.
-The carbon grid transports low-cost carbon to regions where it is sequestered in offshore caverns or processed into carbonaceous fuels together with green hydrogen available at low costs.
-End-use primarily depends on the quality of renewable resources, and thus the availability of low-cost hydrogen.
-In locations with relatively poor renewables and thus higher costs of hydrogen production, such as Italy, Greece, Northern Germany, and lcoations around the Baltic Sea, the transported carbon is predominantly sequestered.
-Conversely, in Spain and the United Kingdom, which both have high renewable potentials and low-cost hydrogen production, the transported carbon is combined with hydrogen and fed into CU processes, with additional carbon being captured via DAC also running on the great renewable resources.
-
-In summary, the implementation of a \carbon{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce low-cost hydrogen to make CU products (FT fuel, methanol, synthetic methane).
-In the absence of a \carbon{} network, sequestration sites first sequester all carbon from low-cost point sources, and deploy additional DAC facilities at the locations with low-cost renewables to capture the remaining necessary carbon for sequestration.
-Then, synfuel demand is met at the locations which offer the best mix of low-cost carbon from point sources and low-cost hydrogen from local renewables.
-% Sequestered carbon is nearly always transported before
-
-The \hydrogengrid{} on the other hand, transports low-cost hydrogen from regions with abundant renewable resources to regions with high energy demand and abundant emissions.
-The main transportation routes go from Spain and the United Kingdom to Central Europe.
-In these regions, industries use the hydrogen for demand in industry and fuel cell land transport, and to produce synfuels with locally captured carbon from point sources.
-This leads to a very decentralized synfuel production across the continent.
-Furthermore, the introduction of a hydrogen network enables to reduce both DAC production and FT fuels on the British Isles and Iberian Peninsula.
-In Great Britain an Ireland, the remaining local carbon is then sequestered, and all the hydrogen that previously went into FT fuel production is exported to Central Europe (see above).
-In Spain and Portugal, some DAC remains and the carbon sourced from DAC is used, along with locally captured carbon from point sources, to produce FT fuels (but less than without a hydrogen network).
-Surplus hydrogen from lower FT fuel production is exported to Central Europe.
-
-% Therefore, one can conclude that a \carbon{} network favors a decentralized CC and a CU centralized at regions with abundant renewable resources.
-A \hydrogen{} network favors a centralized electrolysis system at regions with abundant renewable resources and a decentralized CU system.
-These findings explain the advantage of the \hydrogen{} network in contrast to the \carbon{} network.
-Central Europe with its industrial clusters and dense population is characterized by a scarcity in hydrogen and a surplus and abundance of \carbon.
-While the \carbon{} network cost-efficiently resolves the surplus of carbon in Central Europe by exporting it to regions with low-cost renewable resources to produce synfuels, it needs to supply direct hydrogen demand in Central Europe (for industry and fuel cell land transport) with high-cost local hydrogen production.
-The hydrogen network resolves both the energy scarcity by supplying hydrogen demand in Central Europe with low-cost hydrogen imported from Spain and Great Britain, and the carbon surplus by transporting hydrogen to the sites with carbon available at lower costs in Central Europe, and producing FT fuels highly decentralized throughout all of Central Europe.
-
-\subsection*{Flexibility gains through hybrid configuration}
-
-\begin{figure*}[ht!]
-    \centering
-    \begin{subfigure}{.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{full/figures/90_nodes/capacity_map_hydrogen.png}
-        \caption{\hydrogen{} Sector.}
-        \label{fig:capacity_map_hydrogen_co2}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{full/figures/90_nodes/capacity_map_carbon.png}
-        \caption{\carbon{} Sector.}
-        \label{fig:capacity_map_carbon_co2}
-    \end{subfigure}%
-    \caption{Optimal production and transport capacities of the carbon and hydrogen sector in a net-zero energy system in Europe with both \carbon{} and \hydrogen{} network expansion (Hybrid).
-    % Carbon network looks the same as in~\cite{morbeeOptimisedDeploymentEuropean2012}: two backbones, one in the nothern Europe other in south east.
-    }
-    \label{fig:capacity_maps}
-\end{figure*}
-
-% increased BECCS, decreased DACS
-% decreased methanation
-
-The Hybrid model combines the advantages of both hydrogen and carbon networks.
-While the topology of the \hydrogen{}-grid roughly corresponds to that of the \hydrogen grid model, the \carbon{}-grid plays more of a complementary role.
-Fig.~\ref{fig:capacity_maps} shows the capacities of the two grids in the hybrid model, with the production and transport capacities given in kt/h.
-Note that despite similar overall capacities, the investment in the hydrogen infrastructure is four times the investment in the carbon infrastructure.
-
-Similar to the \hydrogengrid model, the Hybrid model takes advantage of a large hydrogen network that enables hydrogen to be transported from centralized production sites in western regions such as Spain to decentralized FT production sites across the continent.
-However, in the Hybrid model hydrogen in the UK is used locally to produce synfuels, and less hydrogen is transported to Central Europe.
-The carbon network supplements this with smaller clusters of networks.
-In addition to a larger carbon grid cluster in Central Europe, which transports carbon to the North Sea, the system is building three large linear carbon routes.
-These transport carbon from Romania and Bulgaria to Greece, from northern Italy to central Italy and from Hungary, the Czech Republic and Poland to the Baltic Sea.
-These enable the system to transport carbon available at low cost from inland biomass sources and sequester it (see Fig.~\ref{fig:captureshare_line}).
-This reduces the need for DAC in the system and the sequestration sites are shifted to the end points of the carbon networks.
-Second-order effects can be seen in the shift in FT production compared to the \hydrogengrid model.
-Regions that have both good renewable resources and sequestration potential could now forego sequestration and use local process emissions for FT production, as can be observed in Ireland.
-A more detailed of differences between the \hydrogengrid and the Hybrid model can be found in Fig.~.
-
-A notable aspect of the hybrid model is the lack of overlap in the hydrogen and carbon network topologies which is tightly connected to the nodal carbon strategy.
-Depending on the exogeneous need in hydrogen for example from industry, and the costs for locally produced carbon, a region can either import hydrogen to produce FT fuels or export carbon, but it will neither import one and export the other nor import or export both.
-
-Despite only small decrease in cost in comparison to the \hydrogengrid model, hybrid approach brings several advantages.
-These include a broader range of technologies contributing to its robustness, reduced reliance on DAC, which may be costlier than anticipated in the cost projections, and less land use for wind and solar due to the decreased necessity for DAC.
-The increase used of biomass with carbon capture, transport and sequestration leads to slight cost advantages in comparison to the \hydrogengrid model.
-
-\subsection*{Net-Negative Scenarios}
-
-\begin{figure}[htb!]
-    \centering
-    \includegraphics[width=0.9\linewidth]{difference/comparison/emission-reduction-0.1/figures/90_nodes/cost_bar.png}
-    \caption[short]{Net change in investments when tightening the \carbon{} emission target from net-zero to net-negative 10\% of 1990s emissions.}
-    \label{fig:net-negative_cost_bar}
-\end{figure}
-
-We now examine the case of moving from a net zero emissions target to a net emissions reduction target.
-Fig.~\ref{fig:net-negative_cost_bar} illustrates the shift in investments for the \hydrogengrid{} and the Hybrid model between the net-zero and the net-negative scenario.
-
-Both systems show a stron increase in DAC
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-In this study\dots
-
-\dots
-
-Despite the detailed model representation, there are some limitations to the validity of the results.
-The model itself is based on a linearized optimization with perfect foresight for the entire modeling year.
-In reality, long-term energy demand and renewable supply can only be roughly estimated, while short-term predictions still entail some uncertainty.
-The model's perfect foresight may lead to non-reproducible behavior, such as precisely aligning energy storage with future energy shortages at a specific point in time.
-
-The technology costs used in the model rely on cost projections that incorporate reductions based on learning rates.
-These learning rates are derived from historical data, which may not necessarily be indicative of future trends.
-In particular, cost assumptions on DAC, electrolysis and \hydrogen{} pipeline costs, could have a strong impact in the model results.
-
-The modelling results are heavily driven by the demand and emissions from industrial clusters.
-Allowing the model to relocate these and/or incorporate flexibility measures, may lead to less dependency on both \carbon{} and \hydrogen{} networks.
-
-The technological flexibilities in the carbon sector might not be exploited to full extent.
-Therefore, the \carbon{} transport via truck or ship is not considered.
-Fischer-Tropsch facilities require to run on baseload with at least 90\% of their nominal capacity.
-
-\dots
-
-% TODO: which fixed rate? you mean that all emissions are captured by exogenous assumption or that we assume 90% of carbon can be captured?
-% Furthermore, the model assumes a fixed rate for industrial process emissions, which cannot be altered through investments.
-This simplification may not accurately represent real-world scenarios, where many industries are considering adopting low-carbon processes and technologies.
-The transport of fuel (FT, gas, oil) between the regions is not unlimited which may overestimate the flexibility provided by these commodities.
-
-\printbibliography
-
-\appendix
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
-    \caption{Sankey diagram of the optimal operation for a net-negative 10\% scenario.}
-    \label{fig:sankey_diagramm}
-\end{figure*}
-
-\begin{figure}
-    \centering
-    \includegraphics*[width=0.8\linewidth]{comparison/emission-reduction/figures/90_nodes/objective_heatmap.png}
-    \caption{Total annual system cost for the different scenarios and \carbon{} targets.
-While the `Baseline' scenario has neither a \carbon{} nor a \hydrogen{} network, `Hybrid' is allowed to expand both.}
-    \label{fig:objective_heatmap}
-\end{figure}
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=\linewidth]{comparison/emission-reduction-full/figures/90_nodes/transmission_cost_bar.png}
-    \caption{Annual transmission system cost as a function of the net carbon removal scenarios considered in the study.}
-    \label{fig:transmission_cost_bar}
-\end{figure}
-
-\begin{figure*}[ht]
-    \centering
-    \includegraphics[width=\linewidth]{difference/h2-only-full/figures/90_nodes/cost_map.png}
-    \caption{Difference in cost investments between \hydrogengrid and Hybrid model.
-The left subfigure shows higher spendings per technology and region and transport system for the \hydrogengrid model, the right shows higher spendings in the Hydrid model.}
-    \label{fig:cost_map_difference}
-\end{figure*}
-
-\end{document}

### 2023-11-28 workflow/scripts/plot-operation-map.py

@@ -67,7 +67,7 @@ for kind, output in snakemake.output.items():
 
     grouper = s.groupers.get_bus_and_carrier
     df = s.dispatch(bus_carrier=carriers, groupby=grouper)
-    df = df.drop(transport_carriers, level=2, errors="ignore")
+    df.Link = df.Link.drop(transport_carriers, level=1, errors="ignore")
     df = df.rename(lambda x: x.replace(" CC", ""), level=2)
     df = df.groupby(level=[1, 2]).sum().rename(n.buses.location, level=0)
 

### 2023-11-28 report/submissions/journal/paper/paper.tex
add basic 3.0 Results subsection, move 'unlocking carbon' storyline to new subsection 3.1

@@
 \newcommand{\carbon}{CO$_2$}
+\newcommand{\CO}{CO$_2$}
 \newcommand{\hydrogen}{H$_2$}
+\newcommand{\Htwo}{H$_2$}
 \newcommand{\carbongrid}{\carbon{}--Grid}
@@
 
-% FORMATTING: https://www.nature.com/nenergy/submission-guidelines/aip-and-formatting
+% FORMATTING:
+% - https://www.nature.com/nenergy/submission-guidelines/aip-and-formatting
+% - https://www.nature.com/nenergy/content abstract: 150 words, main text: 3000 words, figures: 8
 
-\title{From Net-Zero to Net-Negative: Assessing \hydrogen{} and \carbon{} Network Strategies in Europe}
+\title{Assessing \hydrogen{} and \carbon{} Network Strategies in Europe}
 % \title{Synthetic fuels in Europe: Transport Hydrogen to Carbon, or Carbon to Hydrogen?}
@@
 \begin{abstract}
-    In future climate-neutral scenarios, the transport of hydrogen and carbon dioxide can play competing roles.
-    Hydrogen transport allows energy-intensive regions to import low-cost hydrogen.
-Carbon transport allows regions with carbon sinks such as carbon sequestration and synthetic fuel production to import low-cost carbon from distributed carbon capture sources.
-    Whether carbon and hydrogen transport networks are complementing or substituting each other remains unclear.
-This is especially relevant for the production of carbonaceous fuels such as Fischer-Tropsch fuels, methanol or synthetic methane, which are made from both hydrogen and carbon dioxide, and in most locations require at least one of these feedstocks to be imported.
-    To address this gap, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates transport, storage and sequestration for carbon dioxide, transport and storage for hydrogen, and renewable energy sources with a high spatial and temporal resolution considering all energy intensive sectors.
-    Our findings reveal that when comparing either-or-models, a system with an hydrogen network provides lower system costs than a system with a carbon network.
-The former enables the transportation of hydrogen from centralized production sites with low-cost renewable energy in the Iberian Peninsula and the British Isles to serve distributed hydrogen demands across the European continent.
-The latter facilitates carbon capture at lower costs from distributed point sources and reduces the need for higher cost direct air capture.
-However, the system with only a carbon network needs to meet distributed hydrogen demand across the continent by locally deploying electrolysis regardless of renewable energy costs.
-    In a hybrid model with both types of networks, the carbon network is a cost-effective complement to the hydrogen network facilitating point-source carbon capture at low costs and transport to sequestration or carbon utilization sites.
-In all models, additional demand for synthetic fuels is met at sites with favorable renewable energy resources that enable both electrolysis and direct air capture at low costs.
-    We show that theses findings holds true against the backdrop of a net-zero energy system as well as a scenario with net-negative emissions.
-    Overall, our work demonstrates the cost-effectiveness of a multi-grid system that includes both hydrogen and carbon transport networks and power grid expansion to achieve climate neutrality in Europe.
-    The paper highlights the benefits of combining hydrogen and \carbon{} networks to achieve climate neutrality.
+    % Hydrogen and carbon dioxide transportation in climate-neutral energy systems fulfills distinct yet connected roles: Hydrogen transport allows energy-intensive areas to import low-cost hydrogen, while carbon transportation allows high-emitting regions to capture and export carbon for sequestration or re-use.
+The synergy or competition of hydrogen and carbon transport is yet to be determined.
+    % To address this gap, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates transport, storage and sequestration for carbon dioxide, transport and storage for hydrogen, and renewable energy sources with a high spatial and temporal resolution considering all energy intensive sectors.
+    % Our research shows that in direct comparison a hydrogen network is more cost-effective than a carbon network.
+It allows regions across Europe to import low-cost hydrogen from centralized production sites, supplying local hydrogen demand and enabling local carbon capture and utilization.
+A carbon network facilitates point-source carbon capture and reduces the need for direct air capture, but leaves the scarcity of hydrogen in energy-intensive regions such as industrial clusters.
+    % However, in a hybrid scenario, the carbon network is a cost-effective complement to the hydrogen network facilitating carbon capture at point sources and transport to sequestration sites.
+    % In a hybrid model, the carbon network is a cost-effective complement to the hydrogen network, unlocking carbon capture potentials from biomass with subsequent transport to sequestration and leading to a reduced deployment of direct air capture.
+    % When strengthening the emission reduction target to net-negative realm, additional carbon removal is achieved by expanding direct air capture at coastal regions with favorable renewable resources and sequestration potentials.
+    % Our research underscores the efficiency of integrating hydrogen, carbon transport networks, and power grid expansion in infrastructure planning for achieving climate neutrality in Europe.
+
+    % Condensed version with 150 words
+    Hydrogen and carbon dioxide transportation are considered crucial in climate-neutral energy systems, with hydrogen enabling energy imports to high-demand areas and carbon transport aiding in emissions export from high-emission areas.
+Yet, possible synergies and competitions between the two systems are not fully understood.
+Our study employs optimization techniques to develop a cost-optimal European energy system, integrating transport, storage, and sequestration of both carbon dioxide and hydrogen, along with renewable energy sources.
+Results indicate that a hydrogen network is more cost-effective than a carbon network, facilitating low-cost hydrogen imports and local carbon capture and utilization.
+However, in a hybrid scenario, the carbon network effectively complements the hydrogen network, promoting carbon capture from biomass and reducing reliance on direct air capture.
+This research demonstrates the effectiveness of combining hydrogen and carbon transport networks with power grid expansion in achieving climate neutrality in Europe.
 \end{abstract}
@@
 Initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen Infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier, and some gas pipelines are already repurposed to transport hydrogen~\cite{RohrFreiFuer}.
-At the same time, business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offer \carbon{} offtake on the other hand.
+At the same time, business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synfuels on the one hand and offer \carbon{} offtake on the other hand.
+The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a transport and sequestration capacity of 1.5 Mt \carbon{} per year  to be operative in 2024, expanding to a targeted scale of 5 Mt per year of sequestration by 2030.
+% TODO: frame next two sentences to be more market oriented
 For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.7~Gt of carbon capture from point sources per year, which represent roughly 50\% of all emissions.
 In combination with large sequestration potentials as stated in~\cite{weiProposedGlobalLayout2021}, this highlights the vast potential for decarbonization.
-The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a transport and sequestration capacity of 1.5 Mt \carbon{} per year  to be operative in 2024, expanding to a targeted scale of 5 Mt per year of sequestration by 2030.
 To this end, the Clean Air Task Force underlines the importance of a carbon transport system in Europe to facilitate the carbon economy~\cite{lockwoodEuropeanStrategyCarbon}.
@@
 The study is conducted on the basis of the open-source, capacity-expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
-The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, waste, agriculture, and transport sectors (including international aviation and shipping).
+The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, waste, agriculture, and transport sectors, including international aviation and shipping.
 
@@
 To supply carbon needed for CS and CU, the system can choose to deploy carbon capture technology at various point sources (see below), or through DAC facilities.
-The concept of a ``merit order for captured carbon`` plays a pivotal role in optimizing and prioritizing the deployment of carbon capture technologies based on their economic feasibility.
+The concept of a merit order for captured carbon plays a pivotal role in optimizing and prioritizing the deployment of carbon capture technologies based on their economic feasibility.
 This merit order ranks all CC technologies according to their relative costs to capture a ton of carbon (see Fig.~X).
@@
 This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.
-If we consider the spatial aspect of distributed carbon capture potentials, the \carbon{} network can exploit comparative cost advantages between ``CO2 bidding zones'' to utilize the lowest-cost CC facilities across the continent.
+If we consider the spatial aspect of distributed carbon capture potentials, the \carbon{} network can exploit comparative cost advantages between CO2 bidding zones to utilize the lowest-cost CC facilities across the continent.
 We assume a capture rate of 90\% for CC on process emissions, SMR, biogas-to-gas, as well as gas and biomass used in industry, and 95\% for CC on biomass and gas CHPs.
@@
 
-Focusing first on the Net-Zero scenario, we can highlight significant variations in system costs and technology deployment across the models (see Fig.~\ref{fig:cost_bar}).
+Focusing first on the Net-Zero scenario, we can highlight notable characteristics of the energy system buildouts shared by all models, as well as significant variations in system costs and technology deployment between the models (see Fig.~\ref{fig:cost_bar}).
 
@@
 
-System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
-In the Baseline model, the absence of dedicated carbon and hydrogen transport networks requires the system to deploy carbon production technologies (DAC and CC at point sources) in the regions with carbon sequestration sites, and distributed amounts of hydrogen production technologies (mainly electrolysis) in each region to meet local hydrogen demand for industry and fuel cells in heavy-duty land transport vehicles.
-For CU and its respective carbon and hydrogen demands, the system chooses the sites where on-site production of carbon, hydrogen as well as the inputs and by-product outputs (e.g., waste heat) of the synthesis process have the lowest cost (or highest value).
-This leads to renewable resources and storage solutions being deployed not necessarily in the regions with the best resources or lowest costs for clean electricity generation, but where they are situated closer to the endogenously defined locations of energy demand for electrolysers and DAC.
-Thus, total annual system costs in the Baseline model are highest, totalling \label{}803 billion euros (see Fig.~\ref{fig:objective_heatmap}).
+Total annual system costs in the Baseline model are highest, totalling \label{}803 billion euros.
 The models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs, with around \label{}2\% cost reductions (\label{}781 billion euros) for the \carbon{} network, and around \label{}4\% cost reductions for the \hydrogen{} network and the hybrid scenario (\label{}765 and \label{}764 billion euros, respectively).
 
-We can now closer examine the cost reductions and shifts in cost allocation when shifting from the Baseline model to the other three network models.
-The implementation of a \carbon{} transport network (\textit{\carbon{}~Grid}) facilitates carbon capture from the point sources with the lowest costs across regions, substantially reducing the dependency on DAC.
-While DAC supplies around 50\% of the carbon needed in the Baseline model (400 Mt \carbon{}, see Fig.~\ref{fig:balance_captured_carbon}), in the \modCO{}, DAC is only used to supply around 20\% of the carbon handled by the system (150 Mt \carbon{}).
-The remaining carbon is supplied by additional carbon capture technology on all plants burning biomass in industry facilities and in CHPs as well as all industry facilities using gas (see Fig.~\ref{fig:captureshare_line}), because the cost to deploy on-site carbon capture technology has relatively lower costs than DAC.
-The \carbon{} transport network enables the transport of carbon from the locations of the lowest-cost captured carbon sources to the locations of the carbon sinks at CU and CS sites, reducing the need for higher cost DAC.
-The option to transport carbon also allows for the deployment of renewables in areas that combine high quality renewable resources and high sequestration potentials, leading to system-wide cost savings in renewable generation and storage.
+In all models, almost half of the system costs are spent on primary electricity production from wind and solar power (\label{}45\% or \label{}360 of \label{}799 billion euros total costs in the Baseline model), with another \label{}12\% (\label{}100 billion euros) spent on hydro, nuclear and biomass.
+Costs for energy transport networks for electricity, gas, hydrogen and carbon are below \label{}12\% (\label{}100 billion euros for the Baseline model), with the electricity grid making up more of the system costs than the other three networks combined in all models.
+System costs spent on carbonaceous fuel production (only considering investment costs, not the carbon, hydrogen and electricity inputs) are around \label{}3\% (\label{}25 billion euros) in each model, while electrolysis makes up around \label{}5\% (\label{}40 billion euros).
+Heat pump installations make up around \label{}8\% (60 billion euros).
+
+System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
+In the Baseline model, the absence of dedicated carbon and hydrogen transport networks means that the system cannot use all the CC potentials of geographically fixed point sources, but instead needs to deploy more DAC in the locations where carbon is needed for CS and CU.
+Naturally, costs for the \carbon{} and \hydrogen{} networks differ between models as they are being deployed.
+Finally, we see that the Baseline model deploys more wind and solar than the network models.
+This is driven by two aspects of the electrolytic hydrogen production: First, the Baseline model deploys more methanation, leading to a larger endogenous hydrogen demand and thus more electricity production.
+And second, in the network models, the electrolysis is deployed at sites with better renewable resources, thus decreasing the total investments in solar and wind power necessary to run the electrolysis.
+
+\subsection*{Unlocking Point-Source Carbon Capture}
+
+We can now closer examine the carbon capture activity when shifting from the \modBase{} to the other three network models.
+CU is always deployed in the regions where carbon, hydrogen as well as the inputs and by-product outputs (e.g., waste heat) of the synthesis processes have the lowest cost (or highest value).
+In the \modBase{}, this means combining the lowest-cost CC source (process emissions) with electrolysis in all regions, even if the renewable resources are poor and electrolysis costs are thus relatively high (e.g., non-coastal regions in Germany).
+In regions with medium-quality renewables (e.g., France and Scandinavia), CC biomass is also deployed -- and combined with medium-price electrolytic hydrogen -- to run additional CU.
+But the largest volumes of CU are deployed in the regions with the best renewables and lowest-cost electrolysis (e.g., Iberian Peninsula and British Isles), where all point-source CC potentials are exhausted and additionally necessary carbon is sourced from high-cost DAC.
+
+In the \modH{}, the hydrogen network enables the system to transport low-cost hydrogen to make use of low-cost biomass CC potentials that were unused in the \modBase{}.
+Thus, less additional DAC is needed to run CU processes, contributing to the reduction of DAC to \label{}200 Mt (see Fig.~\ref{fig:balance_captured_carbon}).
+CC technology on process emissions increases from \label{}90\% to \label{}100\%, on biomass in industry from \label{}65\% to \label{}100\%, and on facilities using gas in industry from \label{}40\% to \label{}75\%.
+CC on biomass CHPs increases slightly from \label{}45\% to \label{}50\% (see Fig.~\ref{fig:captureshare_line}).
+
+In the \modCO{}, a similar effect is achieved by transporting the carbon from unused low-cost biomass CC potentials to the sites where the lowest-cost hydrogen is produced.
 
@@
 
-The \hydrogen{}-Grid model underscores the beneficial role of hydrogen infrastructure.
-The creation of a hydrogen transport network enables the cost-effective delivery of hydrogen from regions with a surplus of low-cost renewable energy to high-demand industrial centers.
-This model shows a pronounced decrease in investments in DAC, and accordingly also solar generation which provides most of the electricity to run DAC facilities.
-The reduction in DAC is similar to the \carbon{} network, but the \hydrogen{} network still produces 200 Mt of carbon from DAC.
-CC technology on process emissions increases from 90\% to 100\%, on biomass CHPs from 60\% to 100\%, and on facilities using gas in industry from 50\% to 75\%.
-CC on biomass use in industry remains at 50\%, same as the \modCO{} (see Fig.~\ref{fig:captureshare_line}).
-This is because the \hydrogen{} network can supply low-cost \hydrogen{} to \carbon{} point sources, where without the \hydrogen{} network, fuel synthethis with locally captured \carbon{} and locally sourced electrolytic hydrogen would have been more expensive than synthesizing fuel with \carbon{} from DAC and hydrogen from electrolysis at the best renewable resource sites.
-This highlights how the implementation of \carbon{} and \hydrogen{} networks unlocks carbon capture potentials at distributed point sources across the continent.
+Now we turn to how carbon for CS is sourced by the different models.
+In the \modBase{}, because of the absence of a carbon network, all carbon that is sequestered (200 Mt, the same in each model) needs to be captured in the regions with access to an offshore CS site.
+The \modBase{} deploys all available low-cost point-source CC in these regions, and adds DAC in the sequestration locations with the lowest-cost electricity.
+In other words, the \modBase{} uses the atmosphere as a substitute carbon transport pathway, allowing \CO{} from inland point sources to escape into the atmosphere, and extracting it again from the atmosphere with DAC at the offshore locations.
 
-While the cost reduction in the Hybrid model compared to the \hydrogen-Grid model is marginal, the allocation of costs reveals a propagation of reduced DAC facilities and increased CC at point sources.
-In other words, the \carbon{} network allows to tap into far away biomass potentials to capture carbon indirectly from the air, which provides marginally lower costs than DAC, even considering carbon transport costs.
-The Hybrid model relies slightly less on investments into the \hydrogen{} network than the \modH, but adds some investments into the \carbon{} network.
-Still, investments into the \hydrogen{} network outweigh investments into the \carbon{} network by a factor of four.
+In the \modCO{}, the carbon network enables the system to capture carbon from low-cost point sources and transport it to the sequestration sites, trading off CC costs with transport costs.
+Together with the reduction of DAC for CU explained above, this contributes to the reduction of DAC from almost \label{}400 Mt in the Baseline model to \label{}150 Mt in the \modCO{} (see Fig.~\ref{fig:balance_captured_carbon}).
+The CC share on process emissions, gas and biomass use in industry, as well as biomass CHPs increase to 100\% (see Fig.~\ref{fig:captureshare_line}).
+The increased CC from point sources compared to the \modH{} thus derives from more CC on gas for industry and biomass CHPs.
+
+In the \modHybrid{}, both networks contribute together to an even larger reduction of DAC to only \label{}100 Mt, with a larger deployment of biomass CHPs than in the \modCO{}, and of which also 100\% are equipped with CC installations.
+
+In summary, the \carbon{} network enables the transport of carbon from the lowest-cost CC sources to CU and CS carbon sinks, reducing the need for high-cost DAC.
+The creation of a \hydrogen{} network enables the cost-effective delivery of hydrogen from regions with a surplus of low-cost renewable energy to high-demand industrial centers with exogenous hydrogen demand and a surplus of low-cost CC potentials.
+This highlights how both the implementation of a \carbon{} or a \hydrogen{} network unlocks low-cost carbon capture potentials at geographically fixed point sources across the continent.
 
@@
 
+% TODO: distribution of current refineries might shape the CU network. These are quite decentralized and might fit to the cost-optimal layout.
+
 In this study\dots

### 2023-11-30 report/submissions/journal/paper/word-count.tex
general figure update

@@
-    In future climate-neutral scenarios, the transport of hydrogen and carbon dioxide can play competing roles.
-    Hydrogen transport allows energy-intensive regions to import low-cost hydrogen.
-Carbon transport allows regions with carbon sinks such as carbon sequestration and synthetic fuel production to import low-cost carbon from distributed carbon capture sources.
-    Whether carbon and hydrogen transport networks are complementing or substituting each other remains unclear.
-This is especially relevant for the production of carbonaceous fuels such as Fischer-Tropsch fuels, methanol or synthetic methane, which are made from both hydrogen and carbon dioxide, and in most locations require at least one of these feedstocks to be imported.
-    To address this gap, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates carbon management technologies, hydrogen transport and storage, and renewable energy sources with a high spatial and temporal resolution considering all energy intensive sectors.
-    Our findings reveal that when comparing either-or-models, a system with an hydrogen network provides lower system costs than a system with a carbon network.
-The former enables the transportation of hydrogen from centralized production sites with low-cost renewable energy in the Iberian Peninsula and the British Isles to serve distributed hydrogen demands across the European continent.
-The latter facilitates carbon capture at lower costs from distributed point sources and reduces the need for higher cost direct air capture.
-However, the system with only a carbon network needs to meet distributed hydrogen demand across the continent by locally deploying electrolysis regardless of renewable energy costs.
-    In a hybrid model with both types of networks, the carbon network is a cost-effective complement to the hydrogen network facilitating point-source carbon capture at low costs and transport to sequestration sites.
-In all models, additional demand for synthetic fuels is met at sites with favorable renewable energy resources that enable both electrolysis and direct air capture at low costs.
-    We show that theses findings holds true against the backdrop of a net-zero energy system as well as a scenario with net-negative emissions.
-    Overall, our work demonstrates the cost-effectiveness of a multi-grid system that includes both hydrogen and carbon transport networks and power grid expansion to achieve climate neutrality in Europe.
-    The paper underpins the need for a complementing hydrogen network and \carbon{} network to achieve climate neutrality.
+Hydrogen and carbon dioxide transportation are considered crucial in climate-neutral energy systems, with hydrogen enabling energy imports to high-demand areas and carbon transport aiding in emissions export from high-emission zones.
+Yet, the synergy or competition between these systems is not fully understood.
+Our study employs optimization techniques to develop a cost-optimal European energy system, integrating transport, storage, and sequestration of both carbon dioxide and hydrogen, along with renewable energy sources.
+Results indicate that a hydrogen network is more cost-effective than a carbon network, facilitating low-cost hydrogen imports and local carbon capture.
+However, in a hybrid scenario, the carbon network effectively complements the hydrogen network, promoting carbon capture from biomass and reducing reliance on direct air capture.
+This research demonstrates the effectiveness of combining hydrogen and carbon transport networks with power grid expansion in achieving climate neutrality in Europe.

### 2023-11-30 workflow/scripts/common.py

@@ -126,6 +126,10 @@ def sort_rows_by_relative_diff(df: pd.DataFrame, consider_sign=False):
     )
 
 
+def groupby_carrier_across_cc(n, c, nice_names) -> pd.Series:
+    return n.df(c).carrier.replace(" CC", "", regex=True).replace(n.carriers.nice_name)
+
+
 def get_ordered_handles_labels(ax, data, wrap=20):
     """
     Use this for stacked bar plots with negative values.

### 2023-11-30 workflow/scripts/plot-cost-bar.py

@@ -1,4 +1,5 @@
 import os
+import numpy as np
 import seaborn as sns
 import matplotlib.pyplot as plt
 import pandas as pd
@@ -49,10 +50,11 @@ unit = "bn€/a"
 sort_by_color = (
     lambda df: df.assign(color=colors).sort_values(by="color").drop("color", axis=1)
 )
-grouped = sort_rows_by_diff(grouped).div(norm)[::-1]
+grouped = sort_rows_by_diff(grouped).div(norm)
 rename = {"Carbon Capt. at Point Sources": "Carbon Capture\nat Point Sources"}
 grouped = grouped.rename(rename, axis=0)
 colors = {rename.get(k, k): v for k, v in colors.items()}
+
 # %%
 fig, ax = plt.subplots(
     1, 1, figsize=snakemake.params.settings["figsize"], layout="constrained"
@@ -61,11 +63,43 @@ fig, ax = plt.subplots(
 defaults = dict(kind="bar", stacked=True, rot=90, lw=0.2, alpha=0.8)
 
 kwargs = {**defaults, **snakemake.params.settings.get("kwargs", {})}
+
 grouped.T.plot(ax=ax, color=colors, **kwargs)
+for container in ax.containers:
+    if container.datavalues.sum() > grouped.sum().sum() / 20:
+        ax.bar_label(
+            container,
+            label_type="center",
+            fmt=lambda x: int(round(x, 0)),
+            fontsize=7,
+            color="grey",
+        )
+
+if snakemake.wildcards.comparison == "default":
+    pad = 5
+    total = grouped.sum()
+    baseline = total["Baseline"]
+    decrease = 1 - total / baseline
+    for i, (val, y) in enumerate(zip(decrease, total)):
+        if val > 0:
+            ax.text(
+                i,
+                y + pad,
+                f"{-val:.1%}",
+                ha="center",
+                va="bottom",
+                fontsize=7,
+                color="gray",
+            )
+
 
 ax.axhline(0, color="k", lw=1)
 ax.set_ylabel(f"System cost [{unit}]")
-ax.grid(axis="y", alpha=0.5)
+
+# ax.grid(axis="y", alpha=0.5)
+
+plt.xticks(rotation=90)
+
 handles, labels = ax.get_legend_handles_labels()
 ax.legend().remove()
 ax.legend(

### 2023-11-30 workflow/scripts/plot-cost-compare-diff-bar.py

@@ -3,6 +3,7 @@ import os
 import matplotlib.pyplot as plt
 import pandas as pd
 import seaborn as sns
+import math
 from common import (
     import_network,
     mock_snakemake,
@@ -58,6 +59,7 @@ grouped = sort_rows_by_diff(grouped)
 rename = {"Carbon Capt. at Point Sources": "Carbon Capture\nat Point Sources"}
 grouped = grouped.rename(rename, axis=0)
 colors = {rename.get(k, k): v for k, v in colors.items()}
+# %%
 
 fig, ax = plt.subplots(
     1, 1, figsize=snakemake.params.settings["figsize"], layout="constrained"
@@ -67,6 +69,15 @@ grouped = grouped[grouped.round(0).ne(0).any(axis=1)]
 grouped.T.plot(
     kind="bar", stacked=True, ax=ax, color=colors, lw=0, legend=True, alpha=0.8
 )
+for container in ax.containers:
+    if abs(container.datavalues).sum() > grouped.abs().sum().sum() / 20:
+        ax.bar_label(
+            container,
+            fmt=lambda x: int(round(x, 0)),
+            label_type="center",
+            fontsize=7,
+            color="grey",
+        )
 
 # ax.vlines(0, -0.5, len(diff) - 0.5, color="k", lw=0.5)
 

### 2023-11-30 workflow/scripts/plot-cost-diff-bar.py

@@ -44,6 +44,7 @@ unit = "bn€/a"
 
 diff = (df[df.columns[0]] - df[df.columns[1]]).div(norm)
 
+# %%
 
 fig, ax = plt.subplots(
     1, 1, figsize=snakemake.params.settings["figsize"], layout="constrained"
@@ -52,6 +53,15 @@ fig, ax = plt.subplots(
 diff = diff[diff.round(0) != 0].sort_values()
 colors = carriers.set_index("nice_name").color[diff.index]
 diff.plot(kind="barh", ax=ax, color=colors, lw=0, legend=False)
+for container in ax.containers:
+    if abs(container.datavalues).sum() > diff.abs().sum().sum() / 20:
+        ax.bar_label(
+            container,
+            label_type="center",
+            fontsize=7,
+            color="grey",
+            labels=diff.round(0).astype(int),
+        )
 
 ax.vlines(0, -0.5, len(diff) - 0.5, color="k", lw=0.5)
 ax.set_xlabel(f"Cost benefit [{unit}]")

### 2023-11-30 workflow/scripts/plot-demand.py

@@ -18,7 +18,7 @@ n = import_network(
 )
 s = n.statistics
 
-
+# %%
 demand = s.withdrawal(groupby=s.groupers.get_carrier_and_bus_carrier).Load
 
 
@@ -26,17 +26,18 @@ rename_bus_carrier = {
     "solid biomass for industry": "biomass",
     "gas for industry": "gas",
     "low voltage": "electricity",
-    "services rural heat": "Low-Temperature\nHeat",
-    "urban central heat": "Low-Temperature\nHeat",
-    "residential rural heat": "Low-Temperature\nHeat",
-    "residential urban decentral heat": "Low-Temperature\nHeat",
-    "services urban decentral heat": "Low-Temperature\nHeat",
+    "heat": "low-t heat",
     "Li ion": "electricity",
+    "oil": "carbonaceous\nfuel",
 }
-
+rename_carrier = {
+    "Heat": "Residential and Services Heat",
+    "Electricity": "Residential and Services Elec.",
+}
 
 demand = (
     demand.rename(rename_bus_carrier, level=1)
+    .rename(rename_carrier, level=0)
     .rename(str.title, level=1)
     .rename(n.carriers.nice_name, level=0)
 )
@@ -47,17 +48,18 @@ carrier_order = demand.loc[
 
 df = demand.unstack().div(1e9)
 df = df[df.sum().sort_values().index].loc[carrier_order]
-df = df.rename({"Heat": "Low-T Heat", "Oil": "Carbonaceous\nFuel"}, axis=1)
 
-color = n.carriers.set_index("nice_name").color.to_dict()
+color = n.carriers.set_index("nice_name").rename(rename_carrier).color.to_dict()
 
 
 fig, ax = plt.subplots(figsize=(5, 4))
 df.T.plot.bar(stacked=True, color=color, xlabel="", ylabel="Energy demand [PWh]", ax=ax)
+pad = 0.01
+for i, val in enumerate(df.sum()):
+    ax.text(i, val + pad, f"{val:.2f}", ha="center", va="bottom", fontsize=8, color="k")
+
 ax.legend(title="", frameon=False, bbox_to_anchor=(0, 1.2), loc="upper left")
-
 sns.despine()
-
 
 fig.savefig(
     "/home/fabian/papers/co2-network/results/baseline/figures/90_nodes/total-demand-bar.png",

### 2023-11-30 report/submissions/journal/paper/paper.tex
rename file and rules I
rename III

--- 2023-11-28 report/submissions/journal/paper/paper.tex
@@
     \centering
-    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/energy_balance_bar_carbon.png}
+    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/balance_bar_carbon.png}
     \caption{Balance of captured carbon across all system setups. Positive values indicate carbon capture, negative values indicate carbon consumption.}
@@
     \centering
-    \includegraphics[width=\linewidth]{comparison/emission-reduction-full/figures/90_nodes/transmission_cost_bar.png}
+    \includegraphics[width=\linewidth]{comparison/emission-reduction-full/figures/90_nodes/cost_bar_transmission.png}
     \caption{Annual transmission system cost as a function of the net carbon removal scenarios considered in the study.}
-    \label{fig:transmission_cost_bar}
+    \label{fig:cost_bar_transmission}
 \end{figure}

### 2023-11-30 workflow/scripts/plot_cost_bar_compare_diff.py

@@ -16,7 +16,7 @@ region_alpha = 0.8
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
-        "plot_cost_compare_diff_bar", ext="png", clusters=90, comparison="baseline"
+        "plot_cost_bar_compare_diff", ext="png", clusters=90, comparison="baseline"
     )
 
 sns.set_theme(**snakemake.params["theme"])
@@ -61,7 +61,7 @@ grouped = grouped.rename(rename, axis=0)
 colors = {rename.get(k, k): v for k, v in colors.items()}
 
 snakemake = mock_snakemake(
-    "plot_cost_compare_diff_bar", ext="png", clusters=90, comparison="baseline"
+    "plot_cost_bar_compare_diff", ext="png", clusters=90, comparison="baseline"
 )
 
 sns.set_theme(**snakemake.params["theme"])

### 2023-12-04 report/report.tex
add plotting scripts
add cluster config
update paper
follow up renaming; separate configs; run scenarios wo fuel cell land transport demand

@@
 
+\newcommand{\baseline}{../results/baseline/figures/40_nodes/}
+\newcommand{\halfprice}{../results/half-price/figures/40_nodes/}
+\newcommand{\projectedprice}{../results/projected-price/figures/40_nodes/}
+\newcommand{\comparison}{../results/comparison/figures/40_nodes/}
+
 \graphicspath{
-    {../results/graphics/},
+    {../results/},
 }
@@
 \newcommand{\co}{\ce{CO2}~}
+\newcommand{\carbon}{CO$_2$}
+\newcommand{\hydrogen}{H$_2$}
 \def\el{${}_{\textrm{el}}$}
@@
 
-	\title{Competing networks: Co-optimizing Hydrogen and Carbon Management for a Climate-neutral German Energy System}
+    \title{Carbon Management Strategies for a Climate-Neutral European Economy}
 
-	\author[affil]{ Fabian Hofmann, Christoph Tries, Tom Brown } % \corref{correspondingauthor}
+    \author[tub]{Fabian Hofmann}
     \ead{ m.hofmann@tu-berlin.de }
+    \author[tub]{Tom Brown}
+    \ead{ t.brown@tu-berlin.de }
 
@@
     \begin{abstract}
-		\input{sections/abstract.tex}
+        Achieving a carbon-neutral European economy requires internationally coordinated efforts across all energy sectors, especially those with hard-to-abate emissions like heavy industry.
+Against this backdrop, It is critical to assess the need for comprehensive carbon management involving technologies such as carbon capture, transport, use, and storage to develop an effective carbon strategy for climate neutrality.
+
+        This paper provides the first cost-optimal design for a European energy system that fully incorporates carbon management technologies including a carbon transport network, and considers all emissions-intensive sectors.
+It optimizes carbon technologies, hydrogen transport and storage, and the design of renewable energy sources.
+
+        We show that carbon management technologies aiming at net-zero emissions provide systemic flexibility to the European energy system.
+A carbon network is cost-effective at different carbon capture rates (200 - 1000 Mt/a), leading to average cost savings of €13 billion per year.
+The \carbon{} network enables viable point source carbon capture, reducing the need for direct air capture facilities.
+We show that high sequestration rates are cost-beneficial by allowing continued fossil fuel use in some sectors, but require massive capacity expansion which could jeopardize timely attainment of climate neutrality.
+
     \end{abstract}
@@
 
-% \listoftodos[TODOs]
+\section{Introduction}
+\label{sec:introduction}
 
-% \tableofcontents
+In net-zero economies, carbon capture is regarded as a crucial technology for offsetting emissions from hard-to-abate sectors and unavoidable emissions from land use and agriculture.
+Capturing carbon at point sources and from the atmosphere via Direct Air Capture (DAC) presents new opportunities for producing drop-in fuels such as methane, methanol, or carbon-neutral liquid hydrocarbons.
+This process is known as Carbon Capture and Utilization (CCU).
+Additionally, storing carbon in geological formations is referred to as Carbon Capture and Storage (CCS).
+Together with carbon transport, these technologies are grouped under the term carbon management and are part of the so-called New Carbon Economy~\cite{arniehellerNewCarbonEconomy2019}.
 
-\section{Introduction}
-\label{sec:intro}
+For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.4~Gt of carbon capture from point sources per year.
+In combination with large sequestration potentials as stated in~\cite{weiProposedGlobalLayout2021}, this highlights the vast potential for decarbonization.
+Recently, policymakers and industry in Europe have been committing to carbon management projects, planning the first infrastructure components, and developing business models for emerging sectors of the economy~\cite{adomaitisEquinorRWEBuild2023,apnewswireGermanyDrawLegislation2023,KohlenstoffKannKlimaschutz2023,OGETESJoin2022,TESHydrogenLife2023}.
+Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offer \carbon{} offtake on the other hand.
+In this context, the need for an international strategy, analogous to the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022}, becomes apparent to coordinate and support deep decarbonization efforts.
+Important insights to formulate such a strategy can be gained from energy system models that provide a holistic view of the energy system and its technological interactions.
+To this end, a number of optimization tools have been developed in recent years to model synergies between various carbon management technologies~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
+In contrast to Integrated Assessment Models, these tools account for the spatial distribution of carbon sources and sinks.
+% A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
+However, the models are often limited with regard to both geographical scope and detail.
+While representing a single country with spatial resolution may neglect synergies of international cooperation, a coarse grained representation of multiple countries may neglect important geographical properties.
 
-\input{sections/introduction.tex}
+In this paper, we present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies.
+The study is conducted using the PyPSA-Eur-Sec energy system model and encompasses all relevant energy sectors.
+We enable the system to optimize the design of renewable energy sources and storage technologies, as well as the transmission of electricity, hydrogen, methane, and \carbon{}.
+We show that the dynamics of carbon capture, transport, use, and storage in different regions are highly dependent on the availability of renewable energy, \carbon{} pipelines and sequestration sinks.
+Our evaluation focuses on the transport dynamics of \carbon{} and \hydrogen{} through their respective networks on the European continent.
+We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available sequestration potential.
 
-\section{Methods}
-\label{sec:methods}
+\section{Methodology}
+\label{sec:methodology}
 
-\input{sections/methods.tex}
+To model the detailed expansion of the carbon network, we use the open-source PyPSA-Eur-Sec~\cite{PyPSAEurSecSectorCoupledOpen2023} model.
+This is a fully sector-coupled model with high spatial and temporal resolution and detailed transmission infrastructure for electricity, hydrogen, methane, and carbon networks.
+The model enables joint optimization of investment and operational decisions for generation, storage, conversion, and transmission infrastructures.
+For our study, we use a model representation with 90~geographic regions for Europe and 4 hourly time resolution to capture the dynamics of material flow between carbon sources and sinks while ensuring computational feasibility.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{\baseline sequestration_map.png}
+    \caption{Sequestration map}
+    \label{fig:sequestration_map}
+\end{figure}
+
+The PyPSA-Eur-Sec model incorporates exogenous assumptions for regional energy demand across various sectors, including power, industry, transport, buildings, and agriculture.
+This covers energy demand for shipping and aviation, as well as non-energy feedstocks for the chemical industry.
+In order to address climate neutrality targets, we establish a binding upper limit for zero emissions, consistent with the EU's 2050 target.
+The usage of biomass in each region is constrained by its biomass potential; however, biomass transport between regions is permitted, and the model takes transport costs into account.
+Additionally, we restrict the expansion of the electric transmission system to 20\% of its current capacity to account for challenges in establishing new transmission projects.
+The geographical distribution of process emissions from various industries, such as cement, chemical, paper and printing, glass, and steel, along with refineries, is obtained from~\cite{piamanzGeoreferencedIndustrialSites2018}, amounting to a total of 158~Mt \carbon{} per year.
+% TODO: Mention exogenous decisions on carbon capture and 90% rates here for industry.
+
+For carbon sequestration, we only consider offshore sites as potential sinks (refer to Figure~\ref{fig:sequestration_map}).
+This is partly because offshore storage potentials tends to be larger than onshore potential, and partly due to public safety concerns surrounding carbon storage infrastructure projects near populated areas.
+We adopt conservative assumptions when estimating storage potential, capping the total potential at 25~Mt per site and calculating annual storage availability over a 25-year period.
+% TODO: what defines a site?
+
+Regarding carbon utilization, our model incorporates three drop-in fuel production technologies: methanation, methanolization, and Fischer-Tropsch (FT) synthesis.
+Once processed, the fuels are available for unrestricted use and can be transported across regions.
+Synthetic methane serves as a substitute for natural gas or biogas and is utilized in combined heat and power (CHP) plants, gas boilers for residential heating, or to fulfill the gas requirements of the industry.
+Synthetic methanol is employed to decarbonize fuel demand in the marine industry.
+Lastly, FT fuels can replace fossil oil for producing naphtha in the industry and kerosene for aviation or serve as machinery oil in agriculture.
+
+All technology cost assumptions for the year 2050 are derived from a unified open-source technology database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
+
+To investigate the influence of carbon sequestration on the optimal configuration of carbon management technologies, we incrementally vary the global carbon sequestration potential from 200~Mt to 1000~Mt per year in steps of 200~Mt.
 
@@
 
-\input{sections/results.tex}
+\begin{figure*}[ht!]
+    \centering
+    \includegraphics*[width=\linewidth]{\comparison cost_area.pdf}
+    \caption[short]{Total annual system cost for the sector-coupled system with different levels of carbon sequestration potential, with (left) and without (right) \carbon{} network.
+"Gas Infrastructure" combines gas facilities for power and heat production, "H$_2$ Infrastructure" combines H$_2$ production, transport and re-electrification.
+Regardless of the implementation of a \carbon{} network, the system cost decrease as sequestration increases.
+Due to an increased flexibility from fossil carriers with subsequent sequestration, the need for FT synthesis and \hydrogen{} electrolysis and the corresponding renewable power supply is reduced.
+        % TODO: mention details on the grouped technologies
+    }
+    \label{fig:cost_area}
+\end{figure*}
 
-\section{Discussion}
-\label{sec:discussion}
+First, we analyze the optimal deployment of technologies and their associated costs at different levels of carbon sequestration potential.
+Figure~\ref{fig:cost_area} displays the annual system cost per technology group for the sector-coupled model at various sequestration levels, both with and without a \carbon{} network.
+The costs are calculated as the sum of the annualized investment and operational costs for all technologies.
+Starting from the left, we observe that the majority of system costs are associated with the deployment of renewable energy sources, heating, hydrogen technologies, and the electricity grid.
+Costs for fossil fuels and carbon capture and transport technologies are relatively low.
+In terms of capacities, 200~GW of offshore wind, 3930~GW of solar PV, 1910~GW of onshore wind, and 1650~GW of electrolysis are installed in this setup.
+In both cases, with and without \carbon{} transport, total costs decrease as the sequestration potential increases.
+At high sequestration rates, fossil fuels such as gas and oil are kept in the system.
+This reduces the need for synthetic FT fuel, which requires significant renewable energy sources.
+Simultaneously, the system deploys more \carbon{} removal technologies, such as DAC and Bioenergy with carbon capture (included in "Carbon Capture Technologies"), to offset emissions from fossil fuels.
 
-\input{sections/discussion.tex}
+The cost reduction when going to high sequestration in a system with \carbon{} transport is about 13\%, which is slightly more than without a \carbon{} network (11\%).
+Across all sequestration levels, a system with \carbon{} transport is, on average, 1.6\% less expensive.
+The cost savings for a carbon transport system are lowest at 0.5\% (4 billion euros per year) for a sequestration of 200~Mt and highest at 3\% (21 billion euros per year) for a sequestration of 1000~Mt.
+
+To further investigate the switch in technologies, we analyze the annual operation of the system for different sequestration potentials.
+Figures~\ref{fig:operation_area_carbon_co2network}, \ref{fig:operation_area_hydrogen_co2network} and \ref{fig:operation_area_co2_co2network} show the balances of captured carbon, hydrogen and atmospheric \carbon{} in the system with \carbon{} transport as a function of the sequestration potential respectively.
+The areas in the positive segment indicate gross production of the respective commodity, while the areas in the negative segment indicate gross consumption.
+Note that technologies with a ``*`` are explicitly augmented technologies with integrated carbon capture.
+%
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{\comparison operation_area_carbon.pdf}
+    \caption{Balance of captured \carbon{} for the optimal operation with \carbon{} network. Technologies with "*" are explicitly augmented technologies to capture carbon emissions.}
+    \label{fig:operation_area_carbon_co2network}
+\end{figure}
+
+Figure~\ref{fig:operation_area_carbon_co2network} confirms our previous results and shows that FT synthesis production decreases with sequestration availability.
+At the same time, integrated carbon capture from Gas Combined Heat and Power (CHP) plants, Allam Cycle plants, Steam Methane Reforming (SMR) and DAC increases.
+Other technologies are constant across varying levels of sequestration due to different reasons: biomass is bound by its maximum available potential (as reported in~\cite{europeancommissionjointresearchcentreENSPRESOBIOMASS2019}); the industry processes with their associated emissions are set to fixed values; the methanol production is bound by a lower limit representing the demand from the shipping sector.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{\comparison operation_area_hydrogen.pdf}
+    \caption{Balance of \hydrogen{} production and consumption with \carbon{} network.}
+    % TODO: make colors of gas and gas* more distinct
+    \label{fig:operation_area_hydrogen_co2network}
+\end{figure}
+%
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{\comparison operation_area_co2.pdf}
+    \caption{Balance of atmospheric \carbon{} emissions for the optimal operation with \carbon{} network.}
+    \label{fig:operation_area_co2_co2network}
+\end{figure}
+%
+% TODO: kerosine emissions from fossil based kerosine are not taken into account by the system.
+Figure~\ref{fig:operation_area_hydrogen_co2network} shows that hydrogen production decreases by about 50\% when going from 200~Mt/a sequestration to 1000~Mt/a.
+In the last segment, hydrogen production from SMR starts to replace hydrogen production from electrolysis.
+Correspondingly, hydrogen demand for FT synthetic fuel production disappears, leaving only hydrogen demand for the shipping, industry and transport sector.
+Finally, Figure~\ref{fig:operation_area_co2_co2network} shows that the balance of the atmospheric \carbon{}.
+Note that the absolute value in the atmospheric \carbon{} hardly changes with the sequestration potential.
+This is because the switch from FT to fossil fuels is mirrored by a switch from using captured \carbon{} from biomass and DAC for FT synthesis to sequestering it underground.
+The net increase in DAC is solely due to an increase in unabated emissions from gas boilers.
+
+We can conclude that in combination with a \carbon{} transport network, high rates of sequestration reduce the feasibility of CCU and hydrogen infrastructure and increases the feasibility of CCS, namely DAC and CC at point sources.
+
+To answer the question of how carbon capture technologies at point sources change with the sequestration potential, we analyze the shift to carbon capture integration for all relevant technologies.
+Figure~\ref{fig:captureshare_line} shows the share of a given technology that is equipped with integrated carbon capture, i.e., carbon capture at point source, for both with and without \carbon{} transport network.
+With \carbon{} transport, all technologies except for SMR and Gas CHP are nearly fully equipped with CC across all sequestration potentials.
+As a result of sequestration, the share of SMR with carbon capture rises from 50\% to 100\%, and that of CHP from 0\% to 26\%.
+The technologies behave differently without carbon transport: besides the fact that capturing at point source is less attractive overall, their carbon capture shares rise more slowly as sequestration increases.
+The carbon capture share for process emissions even decreases.
+This is primarily due to the fact that as sequestration increases, the system relies more heavily on DAC and SMR with CC that can be flexibly placed near the sequestration sites.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{\comparison captureshare_line.pdf}
+    \caption{Share of facilities with integrated carbon capture as a function of sequestration potential in a system with and without \carbon{} transport.}
+    \label{fig:captureshare_line}
+\end{figure}
+
+Building on the analysis of optimal operation for varying sequestration potentials, we explore the spatial distribution of carbon capture and transport technologies for the most "optimistic" case of 1000~Mt/a sequestration potential and a carbon transport network.
+
+Figure~\ref{fig:balance_map_noco2network_1000} shows a map of the total production and consumption of captured carbon in the system.
+The left side shows the combined carbon supply from carbon capture facilities and the carbon network flow.
+The right side illustrates regional carbon usage, again, together with the flow.
+Large capacity DACs and SMRs are placed close to shore with carbon capture in order to maximize efficiency.
+Significant facilities for DAC are built in the British Isles where carbon dioxide is used for methanol synthesis and also sequestration.
+Here substantial onshore wind farms generate a total annual production of approximately 400 TWh per year, of which 13\% is utilized by local DAC facilities.
+Other notable DAC installations are located on the Atlantic coast of Portugal, Spain, Ireland, and Denmark.
+Large facilities of SMR with CC are found in Portugal, South England and Greece.
+All these locations are in close proximity to sequestration sites, as indicated in Figure~\ref{fig:sequestration_map}.
+Note that the installation of SMR facilities close to the shore has the additional benefit of enabling easy transportation of hydrogen to ports for the shipping sector.
+
+Looking at the transport of \carbon{} across the system, we observe that the carbon network primarily transports captured carbon from process and biomass emissions from the inland to sequestration areas.
+Note that optimization installs a number of separate carbon transport networks instead of building one large connected grid.
+Therefore, a large network cluster transports carbon from Central Europe to the North Sea.
+In contrast, three smaller clusters in the south of Europe transport carbon to the Atlantic and the Mediterranean Sea.
+
+As shown on the right, methanol production using captured carbon is primarily located in the United Kingdom, Ireland, and South Spain.
+
+Finally, the deployment of carbon capture technologies in a system with high sequestration rates contrasts with the deployment of CCU facilities in a system with low carbon sequestration.
+On the one hand, with high sequestration, carbon capture management in the system is mainly driven by the location of the sequestration sites, leading to a spatially centralized deployment of carbon capture technologies.
+On the other hand, with low sequestration rates the use of CCU in a system is distributed more evenly across the system.
+
+\begin{figure*}[ht!]
+    \centering
+    \includegraphics[width=\linewidth]{\projectedprice balance_map_carbon.png}
+    \caption{Optimal operation per sector for a sequestration of 1000~Mt/a.
+The left side shows the combined carbon supply from carbon capture facilities and the carbon network flow, while the right side illustrates regional carbon usage together with the flow.}
+    % TODO: make colors more distinct
+    \label{fig:balance_map_noco2network_1000}
+\end{figure*}
+
+\section{Limitations}
+\label{sec:limitations}
+
+Despite the detailed model representation, there are some limitations to the validity of the results.
+The model itself is based on a linearized optimization with perfect foresight for the entire modeling year.
+In reality, long-term energy demand and renewable supply can only be roughly estimated, while short-term predictions still entail some uncertainty.
+The model's perfect foresight may lead to non-reproducible behavior, such as precisely aligning energy storage with future energy shortages at a specific point in time.
+
+The technology costs used in the model rely on cost projections that incorporate reductions based on learning rates.
+These learning rates are derived from historical data, which may not necessarily be indicative of future trends.
+The model does not account for uncertainties arising from disruptive market behavior, such as the gas price peak that occurred in 2022.
+
+% TODO: which fixed rate? you mean that all emissions are captured by exogenous assumption or that we assume 90% of carbon can be captured?
+% Furthermore, the model assumes a fixed rate for industrial process emissions, which cannot be altered through investments.
+This simplification may not accurately represent real-world scenarios, where many industries are considering adopting low-carbon processes and technologies.
+The transport of fuel (FT, gas, oil) between the regions is not unlimited which may overestimate the flexibility provided by these commodities.
+
+Lastly, the model's spatial and temporal resolution is insufficient to capture all relevant dynamics.
+Variations in renewable energy supply and energy demand below the applied time resolution, as well as more detailed interregional energy transport constraints, are currently not considered in the optimization.
+Transitioning to an hourly representation and a higher spatial resolution would increase the model's computational time and complexity, but would also enhance its robustness and validity.
 
@@
 
-\input{sections/conclusion.tex}
+In this study, we investigated the impact of a \carbon{} transport system with different carbon sequestration potentials on the optimal configuration of technologies in a fully sector-coupled energy system in Europe, focusing on the deployment of carbon capture and Utilization (CCU) and Carbon Capture and Storage (CCS) technologies.
+Our results indicate that increasing carbon sequestration potential leads to a decrease in total system costs, with a more pronounced cost reduction observed when a carbon transport network is present.
+Higher sequestration rates result in a shift from CCU and hydrogen infrastructure towards CCS and unabated fossil fuel usage compensated by carbon dioxide removal elsewhere in the system, particularly based on direct air and biomass carbon capture with sequestration.
+
+Furthermore, we observed that with a \carbon{} transport network, the system tends to build out more carbon capture at point sources.
+With an increased sequestration potential, the share of technologies equipped with integrated carbon capture rises, particularly for Steam Methane Reforming (SMR) and Gas Combined Heat and Power (CHP) plants.
+Independent of a setup with or without \carbon{} transport, the spatial distribution of carbon capture technologies becomes more centralized, with installations concentrated near sequestration sites.
+
+Despite limitations of the study considering fixed model assumptions and cost uncertainties, our findings contribute valuable insights into the role of a carbon transport system and carbon sequestration potential in the transition towards a climate-neutral and net-negative energy system.
+As Europe strives to achieve its 2050 climate neutrality target, understanding the interplay between carbon sequestration, CCU, and CCS technologies will be crucial for the design of effective policies and strategies.
 
 \section*{Acknowledgements}
+\label{sec:acknowledgements}
 
@@
 \section*{License}
+\label{sec:license}
 % \doclicenseLongText
@@
 \section*{Author Contributions}
-
+\label{sec:author-contributions}
 % following https://casrai.org/credit/
@@
 \section*{Data and Code Availability}
+\label{sec:data-and-code-availability}
 
-A dataset of the model results is available at \href{zenodoTBA}{doi:zenodoTBA}.
-The code to reproduce the experiments is available at 
+A dataset of the model results is available at \href{zenodo}{doi:zenodo}.
+The code to reproduce the experiments is available at \newline
 \href{https://github.com/https://github.com/FabianHofmann/co2-network.git}{github.com/https://github.com/FabianHofmann/co2-network.git}.
@@
 \section{Supplementary Information}
-\label{sec:si}
+\label{sec:supplementary-information}
 

### 2023-12-04 report/submissions/eem/paper/paper.tex
follow up renaming; separate configs; run scenarios wo fuel cell land transport demand

@@
 
-Figure~\ref{fig:operation_map_noco2network_1000} shows a map of the total production and consumption of captured carbon in the system.
+Figure~\ref{fig:balance_map_noco2network_1000} shows a map of the total production and consumption of captured carbon in the system.
 The left side shows the combined carbon supply from carbon capture facilities and the carbon network flow.
@@
     \centering
-    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_1000.png}
+    \includegraphics[width=\linewidth]{balance_map_carbon_co2network_1000.png}
     \caption{Optimal operation per sector for a sequestration of 1000~Mt/a.
@@
     % TODO: make colors more distinct
-    \label{fig:operation_map_noco2network_1000}
+    \label{fig:balance_map_noco2network_1000}
 \end{figure*}

### 2023-12-04 report/submissions/eem/presentation/presentation.tex
follow up renaming; separate configs; run scenarios wo fuel cell land transport demand

@@
     \textbf{Carbon Capture (left) and Utilization \& Storage (right) for 200~Mt$_{\text{seq}}$/a}
-    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_200.png}
+    \includegraphics[width=\linewidth]{balance_map_carbon_co2network_200.png}
 \end{frame}
@@
     \textbf{Carbon Capture (left) and Utilization \& Storage (right) for 400~Mt$_{\text{seq}}$/a}
-    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_400.png}
+    \includegraphics[width=\linewidth]{balance_map_carbon_co2network_400.png}
 \end{frame}
@@
     \textbf{Carbon Capture (left) and Utilization \& Storage (right) for 600~Mt$_{\text{seq}}$/a}
-    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_600.png}
+    \includegraphics[width=\linewidth]{balance_map_carbon_co2network_600.png}
 \end{frame}
@@
     \textbf{Carbon Capture (left) and Utilization \& Storage (right) for 800~Mt$_{\text{seq}}$/a}
-    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_800.png}
+    \includegraphics[width=\linewidth]{balance_map_carbon_co2network_800.png}
 \end{frame}
@@
     \textbf{Carbon Capture (left) and Utilization \& Storage (right) for 1000~Mt$_{\text{seq}}$/a}
-    \includegraphics[width=\linewidth]{operation_map_carbon_co2network_1000.png}
+    \includegraphics[width=\linewidth]{balance_map_carbon_co2network_1000.png}
 \end{frame}

### 2023-12-04 workflow/scripts/leftover/leftover.py

@@ -187,7 +187,7 @@ def create_dummy_network(n):
 #             clusters=config["scenario"]["clusters"],
 #         ),
 #         expand(
-#             "results/comparison-{comparison}/gifs/{clusters}_nodes/operation_map_{kind}.gif",
+#             "results/comparison-{comparison}/gifs/{clusters}_nodes/balance_map_{kind}.gif",
 #             kind=config["constants"]["kinds"],
 #             clusters=config["scenario"]["clusters"],
 #         ),

### 2023-12-04 workflow/scripts/plot_balance_area.py

@@ -35,7 +35,7 @@ kind_to_carrier = {"hydrogen": "H2", "carbon": "co2 stored", "electricity": "AC"
 
 n = import_network(snakemake.input.network)
 key = snakemake.params.label
-balance = n.statistics.balance(aggregate_time=False)
+balance = n.statistics.energy_balance(aggregate_time=False)
 
 
 for kind, output in snakemake.output.items():

### 2023-12-04 workflow/scripts/plot_balance_bar.py

@@ -12,7 +12,7 @@ import seaborn as sns
 from common import (
     import_network,
     mock_snakemake,
-    sort_rows_by_diff,
+    sort_rows_by_relative_diff,
     get_ordered_handles_labels,
 )
 
@@ -20,8 +20,8 @@ if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
         "plot_balance_bar",
         ext="png",
-        clusters=90,
-        comparison="default",
+        clusters=40,
+        comparison="emission-reduction-full",
     )
 
 sns.set_theme(**snakemake.params["theme"])
@@ -31,14 +31,14 @@ config = snakemake.config
 df = {}
 for path in snakemake.input.networks:
     n = import_network(path)
-    balance = n.statistics.balance()
+    balance = n.statistics.energy_balance()
 
     key = snakemake.params.labels[n.meta["wildcards"]["run"]]
 
     df[key] = balance
 
 
-df = pd.concat(df, axis=1)
+df = pd.concat(df, axis=1).fillna(0)
 df = df.rename(lambda k: k.replace(" CC", ""), level=1).groupby(level=[0, 1, 2]).sum()
 
 for kind, output in snakemake.output.items():
@@ -60,13 +60,15 @@ for kind, output in snakemake.output.items():
         fmt = ".0f"
 
     ds = df.xs(carrier, level=2)
-    ds = ds.div(norm)
-    ds = ds[ds.round(2) != 0].droplevel(0).dropna(how="all").fillna(0)
+    ds = ds.div(norm).droplevel(0)
+    reduced = ds.where(ds.round(2) != 0).dropna(how="all").fillna(0)
+    if not reduced.empty:
+        ds = reduced
     ds = ds.groupby(level=0).sum()
     ds = ds.sort_values(ds.columns[0], ascending=False)
     if kind == "co2":
         ds.drop("CO$_2$", inplace=True, errors="ignore")
-    ds = sort_rows_by_diff(ds)
+    ds = sort_rows_by_relative_diff(ds)
 
     fig, ax = plt.subplots(
         1, 1, figsize=snakemake.params.settings["figsize"], layout="constrained"

### 2023-12-04 workflow/scripts/plot_balance_map.py

@@ -16,9 +16,8 @@ from matplotlib.gridspec import GridSpec
 from common import (
     import_network,
     mock_snakemake,
-    get_transmission_links,
-    get_carrier_transport,
 )
+from pypsa.statistics import get_transmission_carriers
 import geopandas as gpd
 
 
@@ -26,13 +25,14 @@ warnings.filterwarnings("ignore", category=UserWarning)
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
-        "plot_operation_map",
-        run="h2-only",
-        clusters=90,
+        "plot_balance_map",
+        run="co2-only",
+        clusters=40,
         ext="png",
     )
 
 sns.set_theme(**snakemake.params["theme"])
+plt.rc("patch", linewidth=0.05)
 
 n = import_network(snakemake.input.network, revert_dac=True)
 regions = gpd.read_file(snakemake.input.onshore_regions).set_index("name")
@@ -42,14 +42,9 @@ config = snakemake.config
 alpha = 1
 region_alpha = 0.6
 
-is_transport = get_transmission_links(n)
-transport_carriers = [
-    *n.links.carrier[is_transport].unique(),
-    *n.lines.carrier.unique(),
-]
-transport_carriers = n.carriers.nice_name[transport_carriers]
 
 s = n.statistics
+colors = n.carriers.set_index("nice_name").color
 
 for kind, output in snakemake.output.items():
 
@@ -67,14 +62,21 @@ for kind, output in snakemake.output.items():
 
     grouper = s.groupers.get_bus_and_carrier
     df = s.dispatch(bus_carrier=carriers, groupby=grouper)
-    df.Link = df.Link.drop(transport_carriers, level=1, errors="ignore")
+    _ = get_transmission_carriers(n, bus_carrier=carriers)
+    transmission_carriers = _.set_levels(
+        n.carriers.nice_name[_.get_level_values(1)], level=1
+    )
+    sub = df.loc[["Link"]].drop(
+        transmission_carriers.unique(1), level=2, errors="ignore"
+    )
+    df = pd.concat([df.drop("Link"), sub])
     df = df.rename(lambda x: x.replace(" CC", ""), level=2)
     df = df.groupby(level=[1, 2]).sum().rename(n.buses.location, level=0)
 
     tags = ["production", "consumption"]
 
     for tag, ax in zip(tags, axes):
[…]

### 2023-12-04 workflow/scripts/plot_cost_bar.py

@@ -1,12 +1,12 @@
 import os
-import numpy as np
 import seaborn as sns
 import matplotlib.pyplot as plt
 import pandas as pd
 from common import (
+    get_ordered_handles_labels,
     import_network,
     mock_snakemake,
-    sort_rows_by_diff,
+    sort_rows_by_relative_diff,
 )
 
 alpha = 1
@@ -20,7 +20,6 @@ if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
 sns.set_theme(**snakemake.params["theme"])
 
 df = {}
-objectives = {}
 for path in snakemake.input.networks:
     n = import_network(path)
     capex = n.statistics.capex()
@@ -33,7 +32,6 @@ for path in snakemake.input.networks:
     key = snakemake.params.labels[n.meta["wildcards"]["run"]]
 
     df[key] = costs
-    objectives[key] = n.objective
 
 df = pd.concat(df, axis=1)
 
@@ -47,23 +45,17 @@ colors = snakemake.config["plotting"]["technology_group_colors"]
 
 norm = 1e9
 unit = "bn€/a"
-sort_by_color = (
-    lambda df: df.assign(color=colors).sort_values(by="color").drop("color", axis=1)
-)
-grouped = sort_rows_by_diff(grouped).div(norm)
+grouped = sort_rows_by_relative_diff(grouped).div(norm)
 rename = {"Carbon Capt. at Point Sources": "Carbon Capture\nat Point Sources"}
 grouped = grouped.rename(rename, axis=0)
 colors = {rename.get(k, k): v for k, v in colors.items()}
 
-# %%
 fig, ax = plt.subplots(
     1, 1, figsize=snakemake.params.settings["figsize"], layout="constrained"
 )
 
 defaults = dict(kind="bar", stacked=True, rot=90, lw=0.2, alpha=0.8)
-
 kwargs = {**defaults, **snakemake.params.settings.get("kwargs", {})}
-
 grouped.T.plot(ax=ax, color=colors, **kwargs)
 for container in ax.containers:
     if container.datavalues.sum() > grouped.sum().sum() / 20:
@@ -95,16 +87,12 @@ if snakemake.wildcards.comparison == "default":
 
 ax.axhline(0, color="k", lw=1)
 ax.set_ylabel(f"System cost [{unit}]")
-
-# ax.grid(axis="y", alpha=0.5)
-
 plt.xticks(rotation=90)
 
[…]

### 2023-12-04 workflow/scripts/plot_cost_bar_compare_diff.py

@@ -7,16 +7,17 @@ import math
 from common import (
     import_network,
     mock_snakemake,
-    sort_rows_by_diff,
+    sort_rows_by_relative_diff,
     get_ordered_handles_labels,
 )
 
-alpha = 1
-region_alpha = 0.8
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
-        "plot_cost_bar_compare_diff", ext="png", clusters=90, comparison="baseline"
+        "plot_cost_bar_compare_diff",
+        ext="png",
+        clusters=90,
+        comparison="emission-reduction-0.1",
     )
 
 sns.set_theme(**snakemake.params["theme"])
@@ -55,24 +56,18 @@ groups = {v: groups[k] for k, v in n.carriers.nice_name.drop("").items()}
 grouped = diff.mul(-1).groupby(groups).sum().div(norm)
 # grouped = grouped[grouped.max(axis=1) > 5e4]
 colors = snakemake.config["plotting"]["technology_group_colors"]
-grouped = sort_rows_by_diff(grouped)
+grouped = sort_rows_by_relative_diff(grouped)
 rename = {"Carbon Capt. at Point Sources": "Carbon Capture\nat Point Sources"}
 grouped = grouped.rename(rename, axis=0)
 colors = {rename.get(k, k): v for k, v in colors.items()}
-
-snakemake = mock_snakemake(
-    "plot_cost_bar_compare_diff", ext="png", clusters=90, comparison="baseline"
-)
-
-sns.set_theme(**snakemake.params["theme"])
-
+# %%
 
 fig, ax = plt.subplots(
     1, 1, figsize=snakemake.params.settings["figsize"], layout="constrained"
 )
 
 grouped = grouped[grouped.round(0).ne(0).any(axis=1)]
-grouped.T.plot(kind="bar", stacked=True, ax=ax, color=colors, legend=True, alpha=0.8)
+grouped.T.plot(kind="bar", stacked=True, ax=ax, color=colors, legend=True, alpha=0.9)
 for container in ax.containers:
     if abs(container.datavalues).sum() > grouped.abs().sum().sum() / 20:
         ax.bar_label(
@@ -82,6 +77,13 @@ for container in ax.containers:
             fontsize=7,
             color="grey",
         )
+pad = ax.get_ylim()[1] * 0.02
[…]

### 2023-12-04 workflow/scripts/plot_cost_bar_diff.py

@@ -18,7 +18,7 @@ if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
 
 sns.set_theme(**snakemake.params["theme"])
 labels = snakemake.config["labels"]
-
+cutoff = snakemake.config["plotting"]["cost_bar_diff"]["cutoff"]
 
 df = {}
 carriers = []
@@ -42,15 +42,14 @@ carriers = pd.concat(carriers).drop_duplicates()
 norm = 1e9
 unit = "bn€/a"
 
-diff = (df[df.columns[0]] - df[df.columns[1]]).div(norm)
-
-# %%
+diff = df[df.columns[0]] - df[df.columns[1]]
+diff = diff[diff.abs() > cutoff].div(norm)
 
 fig, ax = plt.subplots(
     1, 1, figsize=snakemake.params.settings["figsize"], layout="constrained"
 )
 
-diff = diff[diff.round(0) != 0].sort_values()
+diff = diff[diff.round(1) != 0].sort_values()
 colors = carriers.set_index("nice_name").color[diff.index]
 diff.plot(kind="barh", ax=ax, color=colors, lw=0, legend=False)
 for container in ax.containers:

### 2023-12-04 workflow/scripts/plot_objective_heatmap.py

@@ -8,7 +8,6 @@ import seaborn as sns
 from common import (
     import_network,
     mock_snakemake,
-    sort_rows_by_diff,
 )
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
@@ -39,7 +38,7 @@ index = [l if len(l) == 2 else ["net-neutral"] + l for l in df.index.str.split("
 df.index = pd.MultiIndex.from_tuples(index)
 df = df.unstack()
 
-df = df.loc[["net-neutral", "net-negative-0.05", "net-negative-0.1"]][
+df = df.loc[["net-neutral", "net-negative-0.1"]][
     ["baseline", "co2-only", "h2-only", "full"]
 ]
 df = df.div(1e9).round(0).astype(int)  # bn €

### 2023-12-04 report/submissions/journal/paper/paper.tex
add section 3.4

@@
 % use https://pypsa-eur.readthedocs.io/en/latest/licenses.html
-%
+
 Endogenous model results include the expansion of renewable energy sources, storage technologies, and transmission capacities.
-The model considers the various energy carriers like electricity, hydrogen, methan, methanol, liquid hydrocarbons and biomass, together with conversions technologies to convert these into each other.
+The model considers various energy carriers like electricity, hydrogen, methan, methanol, liquid hydrocarbons and biomass, together with conversions technologies.
 Carbon-neutral electricity is provided by wind, solar, biomass, hydro and nuclear power plants.
@@
 Pipeline flows are modelled using net transfer capacities and without flow dynamics, pressure valves, or energy demand for compression.
-No retrofitting of gas pipelines is considered, potentially overestimating hydrogen network cost.
+% No retrofitting of gas pipelines is considered, potentially overestimating hydrogen network cost.
 
 The topology of the \carbon{}, \hydrogen{} and gas network is identical to the topology of the electricity network, connecting all neighboring regions.
-Transport of biomass is considered as a variable cost component, and transport of liquid fuels like oil, methanol and Fischer-Tropsch (FT) fuels is ``copper-plated'', since transport costs per unit of energy are negligible due to their high energy density.
+Transport of liquid fuels like oil, methanol and Fischer-Tropsch (FT) fuels is ``copper-plated'', since transport costs per unit of energy are negligible due to their high energy density.
 Throughout this study, we focus on \carbon{} and \hydrogen{} networks because of their high relevance for infrastructure investments and thus public policy decisions.
@@
 Costs for energy transport networks for electricity, gas, hydrogen and carbon are below \label{}12\% (\label{}100 billion euros for the Baseline model), with the electricity grid making up more of the system costs than the other three networks combined in all models.
-System costs spent on carbonaceous fuel production (only considering investment costs, not the carbon, hydrogen and electricity inputs) are around \label{}3\% (\label{}25 billion euros) in each model, while electrolysis makes up around \label{}5\% (\label{}40 billion euros).
+Capital expenditures on carbonaceous fuel production are around \label{}3\% (\label{}25 billion euros) in each model, while electrolysis makes up around \label{}5\% (\label{}40 billion euros).
 Heat pump installations make up around \label{}8\% (60 billion euros).
@@
 
-\subsection*{Unlocking Point-Source Carbon Capture}
+\subsection*{Transport systems unlock point-source carbon capture}
 
@@
 
-\subsection*{Superiority of \hydrogen{} network}
+\subsection*{\hydrogen{}--Network enables decentral carbon utilization}
 
@@
 %         \centering
-%         \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/operation_map_carbon.png}
+%         \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/balance_map_carbon.png}
 %         \caption{\carbon{} Sector.}
-%         \label{fig:operation_map_carbon}
+%         \label{fig:balance_map_carbon}
 %     \end{subfigure}
@@
 %         \centering
-%         \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/operation_map_hydrogen.png}
+%         \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/balance_map_hydrogen.png}
 %         \caption{\hydrogen{} Sector.}
-%         \label{fig:operation_map_hydrogen}
+%         \label{fig:balance_map_hydrogen}
 %     \end{subfigure}
 %     \caption{Optimal operation per sector for a net-zero energy system in Europe with average production on the left and average consumption on the right for both, (a) the \carbon{} sector in the \carbon{}-Grid model and (b) the \hydrogen{} sector in the \hydrogen-Grid model.}
-%     \label{fig:operation_map}
+%     \label{fig:balance_map}
 % \end{figure*}
@@
     \centering
-    \includegraphics[width=\linewidth]{comparison/single-technologies/figures/90_nodes/operation_map_dedicated.png}
+    \includegraphics[width=\linewidth]{comparison/single-technologies/figures/90_nodes/balance_map_dedicated.png}
     \caption{Optimal operation per sector for a net-zero energy system in Europe with average production on the left and average consumption on the right for both, (a) the \carbon{} sector in the \carbon{}-Grid model and (b) the \hydrogen{} sector in the \hydrogen-Grid model.}
-    \label{fig:operation_map}
+    \label{fig:balance_map}
 \end{figure*}
 
-Examining the cost-reductions across the different models raises an important question: why is the hydrogen network more cost-effective than its carbon counterpart, and how does it achieve similar cost reductions as the hybrid model?
+% Examining the cost-reductions across the different models raises an important question: why is the hydrogen network more cost-effective than its carbon counterpart, and how does it achieve similar cost reductions as the hybrid model?
 An initial hypothesis may be that the cost assumptions for \carbon{}
- and \hydrogen{} networks play a role.
+%  and \hydrogen{} networks play a role.
 However, the amount of system costs spent on either network is substantially lower than the cost difference between the models (20 billion euros less total system costs in the \hydrogengrid{} model, while it spends 10 billion euros on the \hydrogen{} grid and the \carbongrid{} model spends 8 billion euros on the \carbon{} grid).
@@
 To find the reason for why the \hydrogen{} network achieves lower system costs than the \carbon{} network, it is crucial to understand the underlying mechanisms and geographical implications of the two transportation systems.
-Fig.~\ref{fig:co2-only-operation_map_carbon} and \ref{fig:co2-only-operation_map_hydrogen} map the average production, consumption and flow in the carbon and hydrogen sectors for the \carbongrid{} model.
-Fig.
-\ref{fig:h2-only-operation_map_carbon} and \ref{fig:h2-only-operation_map_hydrogen} show equivalent quantities for the \hydrogengrid{} model.
+Fig.~\ref{fig:balance_map} maps the average production, consumption and flow in the carbon (upper row) and hydrogen (lower row) sectors for the \carbongrid{} (left) and \hydrogen{} (right) model.
 In each region, circles represent the production (top half-circle) and consumption (bottom half-circle) of the respective energy carrier, divided into shares of each technology and the area sizes correspond to production and consumption volumes.
@@
 
-\subsection*{Flexibility gains through hybrid configuration}
+\subsection*{Hybrid configuration provides flexibilities}
 
@@
         \centering
-        \includegraphics[width=\linewidth]{full/figures/90_nodes/capacity_map_hydrogen.png}
+        \includegraphics[width=\linewidth]{full/figures/90_nodes/balance_map_single_hydrogen.png}
         \caption{\hydrogen{} Sector.}
@@
         \centering
-        \includegraphics[width=\linewidth]{full/figures/90_nodes/capacity_map_carbon.png}
+        \includegraphics[width=\linewidth]{full/figures/90_nodes/balance_map_single_carbon.png}
         \caption{\carbon{} Sector.}
@@
 Regions that have both good renewable resources and sequestration potential could now forego sequestration and use local process emissions for FT production, as can be observed in Ireland.
+A notable aspect of the hybrid model is the lack of overlap in the hydrogen and carbon network topologies.
+This finding is tightly linked to the fact of competing functionalities to offtake carbon emissions.
 A more detailed of differences between the \hydrogengrid{} and the Hybrid model can be found in Fig.~.
-
-A notable aspect of the hybrid model is the lack of overlap in the hydrogen and carbon network topologies which is tightly connected to the nodal carbon strategy.
-Depending on the exogeneous need in hydrogen for example from industry, and the costs for locally produced carbon, a region can either import hydrogen to produce FT fuels or export carbon, but it will neither import one and export the other nor import or export both.
 
@@
 
-\subsection*{Net-Negative Scenarios}
+\subsection*{Net-negativity amplifies DAC and Bioenergy with CU and CS}
 
@@
 
-We now examine the case of moving from a net zero emissions target to a net emissions reduction target.
-Fig.~\ref{fig:net-negative_cost_bar} illustrates the shift in investments for the \hydrogengrid{} and the Hybrid model between the Net-Zero and the Net-Negative scenario.
+In transitioning from a net zero to a net emissions reduction target, a continuation of investment trends are perceived.
+Fig.~\ref{fig:net-negative_cost_bar} shows the net investment changes per technology group for the \hydrogengrid{} and Hybrid.
+The scenario imposes a net carbon removal of 460 Mt/a, with both models incurring a similar net cost increase - 95 billion €/a for \hydrogengrid{} and 96 billion €/a for Hybrid, balancing the earlier cost advantage of the Hybrid model.
+Both models invest relatively evenly in additional DAC, wind and heat pump systems, with DAC accounting for more than a third of the additional system costs.
+These remove carbon directly at sequestration sites at the coast, in particular in Portugal and UK, together with heat and power from new heat pumps at site and new solar and wind installation in the vicinity.
+More heat pumps, solar and wind installations across all Central and Eastern European countries as well as biogas-to-gas installations compensate for the reduction of fossil gas based heating and powering.
+Reduced usage of gas in general lower the investments in gas infrastructure in both models.
+Fundamental difference in investment strategies occur in the carbon capture sector: In contrast to the \hydrogengrid{} model, the Hybrid model continues the strategy of sequestering carbon captured and transported from bio-energetic inputs, expanding the carbon transport routes from Central and Eastern Europe to the near shores.
+At the same time, carbonaceous fuel production in Central Europe partially moves to Spain and UK, where it uses carbon from new DAC facilities without the installation of a \carbon{} network.
+At the same time, solar based electrolysis in France, Nothern Italy and Switzerland expands requiring more solar capacities but less hydrogen imports from the Iberian Peninsula.
+The \hydrogengrid{} model on the other hand, expands decentralized CU systems from the Net-Zero scenario, which are now increasingly supplied with carbon from biogas-to-gas facilities, showing up in both "Bioenergy" and "Carbon Capture at Point Sources" cost-contributions.
+The centralized DAC facilities at sequestration sites are more expanded than in the Hybrid model and are requiring additional heat from biomass CHPs.
 
-Both systems show a stron increase in DAC
+These findings highlight the continuation of trends of the two systemic approaches, which are interchangable at marginal cost-difference: While a stand-alone \hydrogen{} grid combines well with highly centralized DAC and decentralized CU with bioenergy as primary carbon source, the complementation with a \carbon{} grid promotes combining DAC and CU where good renewable resources are and collecting, transporting and sequestering carbon from decentral bioenergetic processes.
+
+% The difference in absolute investment changes between the two models, regionally displayed in Fig.~\ref{fig:net-negative_cost_map}.
+% \begin{figure*}[htb!]
+%     \centering
+%     \includegraphics[width=\linewidth]{difference/net-negative-0.1-h2-only-full/figures/90_nodes/cost_map.png}
+%     \caption[short]{Net difference in investments between the \hydrogengrid{} model and the Hybrid model for the Net-Negative scenario.}
+%     \label{fig:net-negative_cost_map}
+% \end{figure*}
 
@@
 
+% \begin{figure*}
+%     \centering
+%     \includegraphics[width=\linewidth]{full/figures/90_nodes/sankey_diagramm.png}
+%     \caption{Sankey diagram of the optimal operation for a net-zero scenario.}
+%     \label{fig:sankey_diagramm}
+% \end{figure*}
+
 \begin{figure*}

### 2023-12-04 report/submissions/journal/paper/paper.tex
start reviewing results section

@@
 System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
-In the Baseline model, the absence of dedicated carbon and hydrogen transport networks means that the system cannot use all the CC potentials of geographically fixed point sources, but instead needs to deploy more DAC in the locations where carbon is needed for CS and CU.
-Naturally, costs for the \carbon{} and \hydrogen{} networks differ between models as they are being deployed.
-Finally, we see that the Baseline model deploys more wind and solar than the network models.
+% In the Baseline model, the absence of dedicated carbon and hydrogen transport networks means that the system cannot use all the CC potentials of geographically fixed point sources, but instead needs to deploy more DAC in the locations where carbon is needed for CS and CU.
+% Naturally, costs for the \carbon{} and \hydrogen{} networks differ between models as they are being deployed.
+% Finally, we see that the Baseline model deploys more wind and solar than the network models.
 This is driven by two aspects of the electrolytic hydrogen production: First, the Baseline model deploys more methanation, leading to a larger endogenous hydrogen demand and thus more electricity production.
 And second, in the network models, the electrolysis is deployed at sites with better renewable resources, thus decreasing the total investments in solar and wind power necessary to run the electrolysis.
+As we show in the following, these are due to essential though fundamentally different flexibility gains and a shift in carbon utilization and carbon sequestration strategies through the \hydrogen{} and \carbon{} networks.
 
@@
 
+% This is a bit low-level on regional scale, start with the big picture and dynamics, then zoom in on the regional scale in 3.2 (next section) together with end-use of hydrogen and carbon
+% alternative outline:
+% 1.
+baseline model is sub-optimal -> un-used landlocked carbon capture potentials; scarcity in hydrogen and power in Central Europe due to industrial clusters and high population density; costly DAC + additional renewables at sequestration sites
+% 2.
+introduce carbon balance figure -> transport systems unlock point-source CC potentials, in particular biomass and gas for industry, less methanation, less DAC and therefore less DAC specific renewables
+% 3. introduce capture share figure -> correlation between average cf and cc share, since gas CHPs and PPs are peak load techs, no cc infrastructure;
+% 4. carbon-grid has higher absolute CC and CC shares than hydrogen-grid and less DAC however hydrogen-grid is more cost-effective
+
 We can now closer examine the carbon capture activity when shifting from the \modBase{} to the other three network models.
-CU is always deployed in the regions where carbon, hydrogen as well as the inputs and by-product outputs (e.g., waste heat) of the synthesis processes have the lowest cost (or highest value).
+CU is preferably deployed in the regions where carbon, hydrogen as well as the inputs and by-product outputs (e.g., waste heat) of the synthesis processes have the lowest cost (or highest value).
 In the \modBase{}, this means combining the lowest-cost CC source (process emissions) with electrolysis in all regions, even if the renewable resources are poor and electrolysis costs are thus relatively high (e.g., non-coastal regions in Germany).
@@
 
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/balance_bar_carbon.png}
+    \caption{Balance of captured carbon across all system setups. Positive values indicate carbon capture, negative values indicate carbon consumption.}
+    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
+    \label{fig:balance_captured_carbon}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/balance_bar_hydrogen.png}
+    \caption{Balance of captured hydrogen across all system setups. Positive values indicate carbon capture, negative values indicate carbon consumption.}
+    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
+    \label{fig:balance_captured_carbon}
+\end{figure}
+
 \begin{figure}[h]
@@
 \end{figure}%
-
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/balance_bar_carbon.png}
-    \caption{Balance of captured carbon across all system setups. Positive values indicate carbon capture, negative values indicate carbon consumption.}
-    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
-    \label{fig:balance_captured_carbon}
-\end{figure}
 
@@
 
-\subsection*{Hybrid configuration provides flexibilities}
+\subsection*{Hybrid configuration provides further flexibilities}
 
@@
 The Hybrid model combines the advantages of both hydrogen and carbon networks.
-While the topology of the \hydrogen{}-grid roughly corresponds to that of the \hydrogen{} grid model, the \carbon{}-grid plays more of a complementary role.
-Fig.~\ref{fig:capacity_maps} shows the capacities of the two grids in the hybrid model, with the production and transport capacities given in kt/h.
+While the topology of the \hydrogen{}-grid roughly corresponds to that of the \hydrogengrid{} model, the \carbon{}-grid plays more of a complementary role.
+Fig.~\ref{fig:capacity_maps} shows the average production, consumption and transport of the two grids in the hybrid model.
 Note that despite similar overall capacities, the investment in the hydrogen infrastructure is four times the investment in the carbon infrastructure.

### 2023-12-05 report/submissions/journal/paper/paper.tex
rewrite 3.1 without regional context

@@
 \usepackage{subcaption}
+\usepackage[shortcuts]{extdash}
 
@@
 
-Total annual system costs in the Baseline model are highest, totalling \label{}803 billion euros.
-The models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs, with around \label{}2\% cost reductions (\label{}781 billion euros) for the \carbon{} network, and around \label{}4\% cost reductions for the \hydrogen{} network and the hybrid scenario (\label{}765 and \label{}764 billion euros, respectively).
+Total annual system costs in the \modBase{} are highest, totalling \label{}803 billion euros.
+The models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs, with around \label{}2\% cost reduction (\label{}781 billion euros) in the \modCO{}, and around \label{}4\% cost reduction in the \modH{} and the \modHybrid{} (\label{}765 and \label{}764 billion euros, respectively).
 
 In all models, almost half of the system costs are spent on primary electricity production from wind and solar power (\label{}45\% or \label{}360 of \label{}799 billion euros total costs in the Baseline model), with another \label{}12\% (\label{}100 billion euros) spent on hydro, nuclear and biomass.
-Costs for energy transport networks for electricity, gas, hydrogen and carbon are below \label{}12\% (\label{}100 billion euros for the Baseline model), with the electricity grid making up more of the system costs than the other three networks combined in all models.
+Costs for energy transport networks for electricity, gas, hydrogen and carbon are below \label{}12\% (\label{}100 billion euros), with the electricity grid making up more of the system costs than the other three networks combined.
+This is the case for all four models.
 Capital expenditures on carbonaceous fuel production are around \label{}3\% (\label{}25 billion euros) in each model, while electrolysis makes up around \label{}5\% (\label{}40 billion euros).
-Heat pump installations make up around \label{}8\% (60 billion euros).
+Heat pump installations account for around \label{}8\% (60 billion euros) of total system costs.
 
@@
 
-\subsection*{Transport systems unlock point-source carbon capture}
+\subsection*{Transport Systems Unlock Point\=/Source Carbon Capture}
 
@@
 
+% 1.
+baseline model is sub-optimal -> un-used landlocked carbon capture potentials; scarcity in hydrogen and power in Central Europe due to industrial clusters and high population density; costly DAC + additional renewables at sequestration sites
+We can now closer examine the carbon capture activity when shifting from the \modBase{} to the other three network models.
+In comparison, the \modBase{} reveals significant inefficiencies: On the one hand, regions with no access to sequestration sites (i.e., ``landlocked'' regions) underutilize CC potentials from process emissions, fossil gas and biomass.
+On the other hand, sequestration sites rely on costly DAC to capture the carbon for CS from the air.
+Furthermore, hydrogen and power prices show high regional differences, indicating system inefficiencies due to transport constraints caused by non-existent (in case of hydrogen) or insufficient (in case of power) network connections.
+Prices are significantly higher in Central Europe, a region with high population density and high concentration of energy-intensive industries.
+
+% 2.
+introduce carbon balance figure -> transport systems unlock point-source CC potentials, in particular biomass and gas for industry, less methanation, less DAC and therefore less DAC specific renewables
+Transport systems emerge as a crucial component in optimizing CC.
+By enabling the efficient transportation of carbon, hydrogen, or both, these models reduce the reliance on high-cost DAC, and source more carbon from low-cost point-source CC on biomass and gas in industrial applications (see Fig.~\ref{fig:balance_captured_carbon}).
+The \modH{} only captures \label{}200 Mt of \carbon{} from DAC compared to \label{}400 Mt in the \modBase{}.
+Instead, the \modH{} captures more carbon from process emissions as well as gas and biomass for industry.
+The \modCO{} uses only \label{}150 Mt from DAC, and instead also captures more carbon than the \modH{} from gas for industry and biomass CHPs.
+And finally, the \modHybrid{} captures only \label{}100 Mt from DAC, but slightly more carbon from all other CC technologies.
+Also, the networks reduce the reliance on methanation in the model, reducing the overall demand for CC volumes.
+Another knock-on effect of the reduced reliance on DAC is reduced endogenous demand for renewables to power DAC facilities.
+
+% 3. introduce capture share figure -> correlation between average cf and cc share, since gas CHPs and PPs are peak load techs, no cc infrastructure;
+The increase in captured carbon from point sources in the network models is mostly due to CC infrastructure being deployed on industrial or power plant facilities that are present in all models (as opposed to the deployment of additional electricity generating capacity for the purpose of carbon capture).
+This is evident when analyzing the capture share of the different CC technologies (see Fig.~\ref{fig:captureshare_line}).
+
+In the \modBase{}, the capture share is highest for biogas-to-gas facilities, followed by process emissions, biomass for industry, biomass CHPs, gas for industry, SMR and gas CHPs.
+For the network models, we can observe a correlation between CC share and capacity factors of the underlying technologies: Process emissions as well as gas and biomass for industry have high capacity factors (above 80\%), and CC shares close to 100\% (the only exception being gas for industry in the \modH{}).
+Biomass CHPs have a capacity factor of around 40\%, and CC shares of 50\% in the \modH{} and 95\% in the \modCO{}.
+SMR and gas CHPs have capacity factors below 20\%, and CC shares around 50\% and 0\%, respectively, across all network models.
+Gas CHPs serve as peak load electricity production technology, and with very few operating hours, the high investment costs into CC applications are not efficient.\footnote[1]{Similarly, SMR serves as ``peak-load'' hydrogen production technology in regions with poor renewable resources for electrolysis.}
+
+% 4. carbon-grid has higher absolute CC and CC shares than hydrogen-grid and less DAC however hydrogen-grid is more cost-effective
+
+[old text]
 We can now closer examine the carbon capture activity when shifting from the \modBase{} to the other three network models.
@@
     % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
-    \label{fig:balance_captured_carbon}
+    \label{fig:balance_hydrogen}
 \end{figure}

### 2023-12-06 report/submissions/journal/paper/paper.tex
finish rewrite of 3.1

@@
 \subsection*{Transport Systems Unlock Point\=/Source Carbon Capture}
-
-% This is a bit low-level on regional scale, start with the big picture and dynamics, then zoom in on the regional scale in 3.2 (next section) together with end-use of hydrogen and carbon
-% alternative outline:
-% 1.
-baseline model is sub-optimal -> un-used landlocked carbon capture potentials; scarcity in hydrogen and power in Central Europe due to industrial clusters and high population density; costly DAC + additional renewables at sequestration sites
-% 2.
-introduce carbon balance figure -> transport systems unlock point-source CC potentials, in particular biomass and gas for industry, less methanation, less DAC and therefore less DAC specific renewables
-% 3. introduce capture share figure -> correlation between average cf and cc share, since gas CHPs and PPs are peak load techs, no cc infrastructure;
-% 4. carbon-grid has higher absolute CC and CC shares than hydrogen-grid and less DAC however hydrogen-grid is more cost-effective
-
 % 1.
@@
 
-% 3. introduce capture share figure -> correlation between average cf and cc share, since gas CHPs and PPs are peak load techs, no cc infrastructure;
-The increase in captured carbon from point sources in the network models is mostly due to CC infrastructure being deployed on industrial or power plant facilities that are present in all models (as opposed to the deployment of additional electricity generating capacity for the purpose of carbon capture).
-This is evident when analyzing the capture share of the different CC technologies (see Fig.~\ref{fig:captureshare_line}).
-
-In the \modBase{}, the capture share is highest for biogas-to-gas facilities, followed by process emissions, biomass for industry, biomass CHPs, gas for industry, SMR and gas CHPs.
-For the network models, we can observe a correlation between CC share and capacity factors of the underlying technologies: Process emissions as well as gas and biomass for industry have high capacity factors (above 80\%), and CC shares close to 100\% (the only exception being gas for industry in the \modH{}).
-Biomass CHPs have a capacity factor of around 40\%, and CC shares of 50\% in the \modH{} and 95\% in the \modCO{}.
-SMR and gas CHPs have capacity factors below 20\%, and CC shares around 50\% and 0\%, respectively, across all network models.
-Gas CHPs serve as peak load electricity production technology, and with very few operating hours, the high investment costs into CC applications are not efficient.\footnote[1]{Similarly, SMR serves as ``peak-load'' hydrogen production technology in regions with poor renewable resources for electrolysis.}
-
-% 4. carbon-grid has higher absolute CC and CC shares than hydrogen-grid and less DAC however hydrogen-grid is more cost-effective
-
-[old text]
-We can now closer examine the carbon capture activity when shifting from the \modBase{} to the other three network models.
-CU is preferably deployed in the regions where carbon, hydrogen as well as the inputs and by-product outputs (e.g., waste heat) of the synthesis processes have the lowest cost (or highest value).
-In the \modBase{}, this means combining the lowest-cost CC source (process emissions) with electrolysis in all regions, even if the renewable resources are poor and electrolysis costs are thus relatively high (e.g., non-coastal regions in Germany).
-In regions with medium-quality renewables (e.g., France and Scandinavia), CC biomass is also deployed -- and combined with medium-price electrolytic hydrogen -- to run additional CU.
-But the largest volumes of CU are deployed in the regions with the best renewables and lowest-cost electrolysis (e.g., Iberian Peninsula and British Isles), where all point-source CC potentials are exhausted and additionally necessary carbon is sourced from high-cost DAC.
-
-In the \modH{}, the hydrogen network enables the system to transport low-cost hydrogen to make use of low-cost biomass CC potentials that were unused in the \modBase{}.
-Thus, less additional DAC is needed to run CU processes, contributing to the reduction of DAC to \label{}200 Mt (see Fig.~\ref{fig:balance_captured_carbon}).
-CC technology on process emissions increases from \label{}90\% to \label{}100\%, on biomass in industry from \label{}65\% to \label{}100\%, and on facilities using gas in industry from \label{}40\% to \label{}75\%.
-CC on biomass CHPs increases slightly from \label{}45\% to \label{}50\% (see Fig.~\ref{fig:captureshare_line}).
-
-In the \modCO{}, a similar effect is achieved by transporting the carbon from unused low-cost biomass CC potentials to the sites where the lowest-cost hydrogen is produced.
-
 \begin{figure}[ht!]
@@
 
+% 3. introduce capture share figure -> correlation between average cf and cc share, since gas CHPs and PPs are peak load techs, no cc infrastructure;
+The increase in captured carbon from point sources in the network models is mostly due to CC infrastructure being deployed on industrial or power plant facilities that are present in all models (as opposed to the deployment of additional electricity generating capacity for the purpose of carbon capture).
+This is evident when analyzing the capture share of the different CC technologies (see Fig.~\ref{fig:captureshare_line}).
+
 \begin{figure}[h]
@@
 
-Now we turn to how carbon for CS is sourced by the different models.
+In the \modBase{}, the capture share is highest for biogas-to-gas facilities, followed by process emissions, biomass for industry, biomass CHPs, gas for industry, SMR and gas CHPs.
+For the network models, we can observe a correlation between CC share and capacity factors of the underlying technologies: Process emissions as well as gas and biomass for industry have high capacity factors (above 80\%), and CC shares close to 100\% (the only exception being gas for industry in the \modH{}).
+Biomass CHPs have a capacity factor of around 40\%, and CC shares of 50\% in the \modH{} and 95\% in the \modCO{}.
+SMR and gas CHPs have capacity factors below 20\%, and CC shares around 50\% and 0\%, respectively, across all network models.
+Gas CHPs serve as peak load electricity production technology, and with very few operating hours, the high investment costs into CC applications are not efficient.\footnote[1]{Similarly, SMR serves as ``peak-load'' hydrogen production technology in regions with poor renewable resources for electrolysis.}
+
+% 4. carbon-grid has higher absolute CC and CC shares than hydrogen-grid and less DAC however hydrogen-grid is more cost-effective
+In summary, the implementation of \carbon{} and \hydrogen{} networks is instrumental in unlocking low-cost CC potentials at geographically fixed point sources across the continent.
+While the \carbon{} network facilitates the movement of carbon from low-cost sources to CU and CS sinks, reducing the reliance on high-cost DAC, the \hydrogen{} network enables the cost-effective delivery of hydrogen from regions rich in renewable energy to regions with low-cost CC potentials to produce CU.
+However, despite the \carbongrid{} model demonstrating higher absolute amounts of CC from point sources and higher CC shares across all point-source CC technologies, the \hydrogengrid{} model achieves lower total system costs (see Fig.~\ref{fig:cost_bar}).
+In the following section, we will explore the reasons for this.
+
+% [potentially relevant text blobs from old 3.1]
+% CU is preferably deployed in the regions where carbon, hydrogen as well as the inputs and by-product outputs (e.g., waste heat) of the synthesis processes have the lowest cost (or highest value).
+
+% Now we turn to how carbon for CS is sourced by the different models.
 In the \modBase{}, because of the absence of a carbon network, all carbon that is sequestered (200 Mt, the same in each model) needs to be captured in the regions with access to an offshore CS site.
@@
 
-In the \modCO{}, the carbon network enables the system to capture carbon from low-cost point sources and transport it to the sequestration sites, trading off CC costs with transport costs.
-Together with the reduction of DAC for CU explained above, this contributes to the reduction of DAC from almost \label{}400 Mt in the Baseline model to \label{}150 Mt in the \modCO{} (see Fig.~\ref{fig:balance_captured_carbon}).
-The CC share on process emissions, gas and biomass use in industry, as well as biomass CHPs increase to 100\% (see Fig.~\ref{fig:captureshare_line}).
-The increased CC from point sources compared to the \modH{} thus derives from more CC on gas for industry and biomass CHPs.
+% In the \modCO{}, the carbon network enables the system to capture carbon from low-cost point sources and transport it to the sequestration sites, trading off CC costs with transport costs.
 
-In the \modHybrid{}, both networks contribute together to an even larger reduction of DAC to only \label{}100 Mt, with a larger deployment of biomass CHPs than in the \modCO{}, and of which also 100\% are equipped with CC installations.
-
-In summary, the \carbon{} network enables the transport of carbon from the lowest-cost CC sources to CU and CS carbon sinks, reducing the need for high-cost DAC.
-The creation of a \hydrogen{} network enables the cost-effective delivery of hydrogen from regions with a surplus of low-cost renewable energy to high-demand industrial centers with exogenous hydrogen demand and a surplus of low-cost CC potentials.
-This highlights how both the implementation of a \carbon{} or a \hydrogen{} network unlocks low-cost carbon capture potentials at geographically fixed point sources across the continent.
+% In summary, the \carbon{} network enables the transport of carbon from the lowest-cost CC sources to CU and CS carbon sinks, reducing the need for high-cost DAC.
+% The creation of a \hydrogen{} network enables the cost-effective delivery of hydrogen from regions with a surplus of low-cost renewable energy to high-demand industrial centers with exogenous hydrogen demand and a surplus of low-cost CC potentials, reducing the need for DAC and for high-cost electrolysis at the industrial centers.
+% This highlights how both the implementation of a \carbon{} or a \hydrogen{} network unlocks low-cost carbon capture potentials at geographically fixed point sources across the continent.
 

### 2023-12-06 report/submissions/journal/paper/paper.tex
shorten/find comprimize for 3.2

@@
 Fig.~\ref{fig:balance_map} maps the average production, consumption and flow in the carbon (upper row) and hydrogen (lower row) sectors for the \carbongrid{} (left) and \hydrogen{} (right) model.
-In each region, circles represent the production (top half-circle) and consumption (bottom half-circle) of the respective energy carrier, divided into shares of each technology and the area sizes correspond to production and consumption volumes.
-Differences between production and consumption in a given node correspond to import or export volumes through the connected networks, represented as lines with their size and arrows indicating volumes and direction of average flows.
-Each region's color represents the demand-weighted average price of the considered carrier in that region.
+Each map includes circles indicating the production (top half-circle) and consumption (bottom half-circle) for each region, with lines representing the import/export volumes and flow directions.
+The color of each region corresponds to the average price of the carrier.
 
-There are four notable dynamics or differences between the two grid strategies:
-First, the production of FT fuels always moves to where low-cost carbon and low-cost hydrogen can be brought together.
+The \carbongrid{} model captures large amounts of low-cost carbon from bioenergy and industrial activities, especially in Central and Eastern Europe (top left), almost none of which is used locally.
+A large part is transported from Western Europe to the Iberian Peninsula, the British Isles and Denmark, where it is used to supplement low-cost hydrogen in central plants for electrolysis, Fischer-Tropsch and methanolization (bottom left).
+A second part is transported to methanation plants such as those in Italy and Greece, from where the gas is fed into the gas grid, which also is set up for distribution from large gas import hubs.
+Finally, the third part is transported directly to sites in the North Sea, Baltic Sea and Mediterranean Sea where it is eventually enriched with \carbon{} from DAC to be finally sequestered.
+In absence of a \hydrogen{} network, the system distributes electrolysis where it is needed for industrial processes and fuel cell land transport and emphasizes gas imports into Central Europe.
+
+The \hydrogengrid{} on the other hand, transports low-cost hydrogen from regions with abundant renewable resources to regions with high energy demand and abundant emissions (lower right).
+The main transportation routes go from Spain and the United Kingdom to Central Europe.
+In these regions, industries use the hydrogen for demand in industry and fuel cell land transport, and to produce synfuels with locally captured carbon from point sources, leading to a decentralized CU system across the continent.
+DAC and local emissions nearby the coast enable sequestration in the North Sea, Baltic Sea, Mediterranean Sea and, in particular, in Portugal (upper right).
+% Furthermore, the introduction of a hydrogen network enables to reduce both DAC production and FT fuels on the British Isles and Iberian Peninsula.
+In Great Britain an Ireland, the remaining local carbon is then sequestered, and all the hydrogen that previously went into FT fuel production is exported to Central Europe (see above).
+In Spain and Portugal, some DAC remains and the carbon sourced from DAC is used, along with locally captured carbon from point sources, to produce FT fuels (but less than without a hydrogen network).
+Surplus hydrogen from lower FT fuel production is exported to Central Europe.
+
+The results illustrate the fundamentally different approaches of the grids: a \carbon{} grid prefers a decentralized CC system and a co-location of electrolysis and CU centrally in regions with abundant renewable resources; a \hydrogen{} grid prefers a centralized electrolysis system and a decentralized CC and CU system distributed jointly across the continent.
+A \hydrogen{} network not only unlocks the potential of carbon capture from point sources, but also reduces the energy shortages in energy-intensive areas.
+
+% There are four notable dynamics or differences between the two grid strategies:
+% First, the production of FT fuels always moves to where low-cost carbon and low-cost hydrogen can be brought together.
 In the \carbongrid{} model, carbon is moved from Western Europe to the Iberian Peninsula, the British Isles and to Denmark to produce FT fuels centrally in few regions.
 In the \hydrogengrid{} model, hydrogen is transported in the opposite direction to produce FT fuels decentrally across all regions in Central Europe with locally captured \carbon{} from point sources.
-Second, the \carbongrid{} model captures more low-cost carbon from biomass, especially in Central and Eastern Europe, and less high-cost carbon from DAC in Greece and Southern Italy.
+% Second, the \carbongrid{} model captures more low-cost carbon from biomass, especially in Central and Eastern Europe, and less high-cost carbon from DAC in Greece and Southern Italy.
 Instead, the carbon grid transports carbon from Central and Eastern Europe to Greece and Italy, where it is sequestered, or feeds into CU together with locally produced, low-cost electrolytic hydrogen.
 The \hydrogengrid{} model directly operates electrolyzers and DAC in Greece and Southern Italy, with electricity from low-cost renewables, and thus avoids significant network buildouts in this region.
-Third, sequestration in the \hydrogengrid{} model occurs more decentralized, spread out across several regions with access to offshore sequestration sites, making use of point-source CC potentials in each of these regions.
+% Third, sequestration in the \hydrogengrid{} model occurs more decentralized, spread out across several regions with access to offshore sequestration sites, making use of point-source CC potentials in each of these regions.
 In the \carbongrid{} model, sequestration is clustered at fewer sequestration sinks, each collecting low-cost \carbon{} from neighboring regions.
 However, there is no centralization at only one or two sites, since all the sequestration sites offer the same costs and the system minimizes system-wide carbon transport distances to about eight to ten sequestration sites.
-And fourth, how the three types of CU are spread across the continent is driven by the costs of carbon and hydrogen (the latter mostly driven by the costs of renewable electricity) as well as by renewables capacity factors: Clearly, all carbonaceous fuel production requires sites where both carbon and hydrogen are available, and they moreover prefer sites with low carbon and hydrogen prices.
+% And fourth, how the three types of CU are spread across the continent is driven by the costs of carbon and hydrogen (the latter mostly driven by the costs of renewable electricity) as well as by renewables capacity factors: Clearly, all carbonaceous fuel production requires sites where both carbon and hydrogen are available, and they moreover prefer sites with low carbon and hydrogen prices.
 Based on our model assumptions, FT fuel synthesis has very limited flexibility and needs to be run alost at baseload (ramp-down only possible to 90\% of maximum load).
@@
 
-In summary, the implementation of a \carbon{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce low-cost hydrogen to make CU products (FT fuel, methanol, synthetic methane).
+% In summary, the implementation of a \carbon{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce low-cost hydrogen to make CU products (FT fuel, methanol, synthetic methane).
 DAC and electrolysis are deployed in the regions with the best renewable resources to meet carbonaceous fuel demands.
-In the absence of a \carbon{} network and with a \hydrogengrid{} on the other hand, regions with access to offshore sequestration sites capture all carbon from low-cost point sources and then sequester it.
+% In the absence of a \carbon{} network and with a \hydrogengrid{} on the other hand, regions with access to offshore sequestration sites capture all carbon from low-cost point sources and then sequester it.
 The system deploys additional DAC facilities at the locations with sequestration access and low-cost renewables to capture the remaining necessary carbon for sequestration.
@@
 
-Together with understanding the dynamics described above, we can highlight three cost advantages of the \hydrogen{} network in contrast to the \carbon{} network: Capturing carbon at the lowest-cost regions reduces 4 billion euros in spending on DAC.
+% Together with understanding the dynamics described above, we can highlight three cost advantages of the \hydrogen{} network in contrast to the \carbon{} network: Capturing carbon at the lowest-cost regions reduces 4 billion euros in spending on DAC.
 However, the \hydrogengrid{} model produces electrolytic hydrogen in the lowest-cost regions and in total spends 7 billion euros less on renewable energy (14 billion euros less on solar, and 7 billion euros more on wind) as well as 1 billion euros less on SMR.
 Furthermore, without a \hydrogen{} network, additional \hydrogen{} storage is needed in Spain at a cost of 5 billion euros to store hydrogen produced during the summer for utilization in FT fuel production during the winter.
-And finally, the \carbongrid{} model spends 7 billion euros more on gas boilers, gas plants, gas infrastructure, and methanation.
+% And finally, the \carbongrid{} model spends 7 billion euros more on gas boilers, gas plants, gas infrastructure, and methanation.
 % TODO: why??
-The costs of the respective networks, on the other hand, are almost equal: the \carbongrid{} model spends 8 billion euros on the \carbon{} network, while the \hydrogengrid{} model even spends 10 billion euros on the \hydrogen{} network.
+% The costs of the respective networks, on the other hand, are almost equal: the \carbongrid{} model spends 8 billion euros on the \carbon{} network, while the \hydrogengrid{} model even spends 10 billion euros on the \hydrogen{} network.
 These factors add up to account for 14 billion euros of the difference of 20 billion euros in system costs between the two scenarios (785 billion euros for the \carbongrid{} and 765 billion euros for the \hydrogengrid{} model, see Fig.~\ref{fig:objective_heatmap}).

### 2023-12-06 report/submissions/journal/paper/paper.tex
paper: tidy up a little, marginally review 3.1

@@
 \begin{abstract}
-    % Hydrogen and carbon dioxide transportation in climate-neutral energy systems fulfills distinct yet connected roles: Hydrogen transport allows energy-intensive areas to import low-cost hydrogen, while carbon transportation allows high-emitting regions to capture and export carbon for sequestration or re-use.
-The synergy or competition of hydrogen and carbon transport is yet to be determined.
-    % To address this gap, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates transport, storage and sequestration for carbon dioxide, transport and storage for hydrogen, and renewable energy sources with a high spatial and temporal resolution considering all energy intensive sectors.
-    % Our research shows that in direct comparison a hydrogen network is more cost-effective than a carbon network.
-It allows regions across Europe to import low-cost hydrogen from centralized production sites, supplying local hydrogen demand and enabling local carbon capture and utilization.
-A carbon network facilitates point-source carbon capture and reduces the need for direct air capture, but leaves the scarcity of hydrogen in energy-intensive regions such as industrial clusters.
-    % However, in a hybrid scenario, the carbon network is a cost-effective complement to the hydrogen network facilitating carbon capture at point sources and transport to sequestration sites.
-    % In a hybrid model, the carbon network is a cost-effective complement to the hydrogen network, unlocking carbon capture potentials from biomass with subsequent transport to sequestration and leading to a reduced deployment of direct air capture.
-    % When strengthening the emission reduction target to net-negative realm, additional carbon removal is achieved by expanding direct air capture at coastal regions with favorable renewable resources and sequestration potentials.
-    % Our research underscores the efficiency of integrating hydrogen, carbon transport networks, and power grid expansion in infrastructure planning for achieving climate neutrality in Europe.
-
-    % Condensed version with 150 words
     Hydrogen and carbon dioxide transportation are considered crucial in climate-neutral energy systems, with hydrogen enabling energy imports to high-demand areas and carbon transport aiding in emissions export from high-emission areas.
@@
 Our study employs optimization techniques to develop a cost-optimal European energy system, integrating transport, storage, and sequestration of both carbon dioxide and hydrogen, along with renewable energy sources.
-Results indicate that a hydrogen network is more cost-effective than a carbon network, facilitating low-cost hydrogen imports and local carbon capture and utilization.
+Results indicate that a hydrogen network is more cost-effective than a carbon network, facilitating low-cost hydrogen imports to Central Europe and local carbon capture and utilization.
 However, in a hybrid scenario, the carbon network effectively complements the hydrogen network, promoting carbon capture from biomass and reducing reliance on direct air capture.
@@
 And second, in the network models, the electrolysis is deployed at sites with better renewable resources, thus decreasing the total investments in solar and wind power necessary to run the electrolysis.
-As we show in the following, these are due to essential though fundamentally different flexibility gains and a shift in carbon utilization and carbon sequestration strategies through the \hydrogen{} and \carbon{} networks.
+As we show below, these are due to substantial, albeit fundamentally different, gains in flexibility and a shift in carbon utilization and storage strategies by the \hydrogen{} and \carbon{} networks.
 
 \subsection*{Transport Systems Unlock Point\=/Source Carbon Capture}
+
 % 1.
 baseline model is sub-optimal -> un-used landlocked carbon capture potentials; scarcity in hydrogen and power in Central Europe due to industrial clusters and high population density; costly DAC + additional renewables at sequestration sites
-We can now closer examine the carbon capture activity when shifting from the \modBase{} to the other three network models.
-In comparison, the \modBase{} reveals significant inefficiencies: On the one hand, regions with no access to sequestration sites (i.e., ``landlocked'' regions) underutilize CC potentials from process emissions, fossil gas and biomass.
+In comparison to the other three models, the \modBase{} reveals significant inefficiencies: On the one hand, regions with no access to sequestration sites (i.e., ``landlocked'' regions) underutilize CC potentials from process emissions, fossil gas and biomass.
 On the other hand, sequestration sites rely on costly DAC to capture the carbon for CS from the air.
-Furthermore, hydrogen and power prices show high regional differences, indicating system inefficiencies due to transport constraints caused by non-existent (in case of hydrogen) or insufficient (in case of power) network connections.
+Furthermore, hydrogen and power prices show high regional differences, indicating system inefficiencies due to transport constraints caused by non-existent hydrogen or insufficient power transmission.
 Prices are significantly higher in Central Europe, a region with high population density and high concentration of energy-intensive industries.
@@
 introduce carbon balance figure -> transport systems unlock point-source CC potentials, in particular biomass and gas for industry, less methanation, less DAC and therefore less DAC specific renewables
-Transport systems emerge as a crucial component in optimizing CC.
 By enabling the efficient transportation of carbon, hydrogen, or both, these models reduce the reliance on high-cost DAC, and source more carbon from low-cost point-source CC on biomass and gas in industrial applications (see Fig.~\ref{fig:balance_captured_carbon}).
-The \modH{} only captures \label{}200 Mt of \carbon{} from DAC compared to \label{}400 Mt in the \modBase{}.
+The \modH{} only captures \label{}200~Mt of \carbon{} from DAC compared to \label{}400~Mt in the \modBase{}.
 Instead, the \modH{} captures more carbon from process emissions as well as gas and biomass for industry.
-The \modCO{} uses only \label{}150 Mt from DAC, and instead also captures more carbon than the \modH{} from gas for industry and biomass CHPs.
-And finally, the \modHybrid{} captures only \label{}100 Mt from DAC, but slightly more carbon from all other CC technologies.
+The \modCO{} uses only \label{}150~Mt from DAC, and instead also captures more carbon than the \modH{} from gas for industry and biomass CHPs.
+And finally, the \modHybrid{} captures only \label{}100~Mt from DAC, but slightly more carbon from all other CC technologies.
 Also, the networks reduce the reliance on methanation in the model, reducing the overall demand for CC volumes.
@@
 
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/balance_bar_hydrogen.png}
-    \caption{Balance of captured hydrogen across all system setups. Positive values indicate carbon capture, negative values indicate carbon consumption.}
-    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
-    \label{fig:balance_hydrogen}
-\end{figure}
+% \begin{figure}[ht!]
+%     \centering
+%     \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/balance_bar_hydrogen.png}
+%     \caption{Balance of captured hydrogen across all system setups. Positive values indicate carbon capture, negative values indicate carbon consumption.}
+%     % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
+%     \label{fig:balance_hydrogen}
+% \end{figure}
 
@@
 
-% increased BECCS, decreased DACS
-% decreased methanation
-
-The Hybrid model combines the advantages of both hydrogen and carbon networks.
+The Hybrid model combines the advantages of hydrogen and carbon networks.
 While the topology of the \hydrogen{}-grid roughly corresponds to that of the \hydrogengrid{} model, the \carbon{}-grid plays more of a complementary role.
@@
 
+% \section{Limitations}
+% \label{sec:limitations}
+% Despite the detailed model representation, there are some limitations to the validity of the results.
+The model itself is based on a linearized optimization with perfect foresight for the entire modeling year.
+In reality, long-term energy demand and renewable supply can only be roughly estimated, while short-term predictions still entail some uncertainty.
+The model's perfect foresight may lead to non-reproducible behavior, such as precisely aligning energy storage with future energy shortages at a specific point in time.
+% The technology costs used in the model rely on cost projections that incorporate reductions based on learning rates.
+These learning rates are derived from historical data, which may not necessarily be indicative of future trends.
+In particular, cost assumptions on DAC, electrolysis and \hydrogen{} pipeline costs, could have a strong impact in the model results.
+% The modelling results are heavily driven by the demand and emissions from industrial clusters.
+Allowing the model to relocate these and/or incorporate flexibility measures, may lead to less dependency on both \carbon{} and \hydrogen{} networks.
+% The technological flexibilities in the carbon sector might not be exploited to full extent.
+Therefore, the \carbon{} transport via truck or ship is not considered.
+Fischer-Tropsch facilities require to run on baseload with at least 90\% of their nominal capacity.
+% Furthermore, the model assumes a fixed rate for industrial process emissions, which cannot be altered through investments.
+This simplification may not accurately represent real-world scenarios, where many industries are considering adopting low-carbon processes and technologies.
+The transport of fuel (FT, gas, oil) between the regions is not unlimited which may overestimate the flexibility provided by these commodities.
+
 \section{Conclusion}
@@
 
+This study offers a detailed exploration of Europe's energy system in 2050, focusing on the integration and optimization of hydrogen and carbon networks to achieve climate neutrality.
+The analysis reveals significant efficiencies gained through these networks, particularly in terms of leveraging low-cost carbon sources and facilitating a decentralized approach to carbon utilization.
+
+The hydrogen network demonstrates superior cost-effectiveness, primarily due to its capacity to transport low-cost hydrogen to high-demand regions and promote decentralized carbon capture and utilization.
+This network reduces reliance on direct air capture (DAC), offering a shift towards more economically viable and sustainable solutions.
+The combination of hydrogen and carbon networks in the hybrid model introduces additional flexibilities.
+This model not only achieves further cost reductions but also shows resilience against potential increases in DAC costs and reduces land use for renewable energy installations.
+Under net-negative emission scenarios, the importance of DAC and bioenergy with carbon capture and storage (BECCS) is amplified.
+Both the hydrogen and hybrid models demonstrate adaptability to these more stringent climate targets, highlighting the robustness of the networked approaches.
+
 % TODO: distribution of current refineries might shape the CU network. These are quite decentralized and might fit to the cost-optimal layout.
 
-In this study\dots
-
-\dots
-
-Despite the detailed model representation, there are some limitations to the validity of the results.
-The model itself is based on a linearized optimization with perfect foresight for the entire modeling year.
-In reality, long-term energy demand and renewable supply can only be roughly estimated, while short-term predictions still entail some uncertainty.
-The model's perfect foresight may lead to non-reproducible behavior, such as precisely aligning energy storage with future energy shortages at a specific point in time.
-
-The technology costs used in the model rely on cost projections that incorporate reductions based on learning rates.
-These learning rates are derived from historical data, which may not necessarily be indicative of future trends.
-In particular, cost assumptions on DAC, electrolysis and \hydrogen{} pipeline costs, could have a strong impact in the model results.
-
-The modelling results are heavily driven by the demand and emissions from industrial clusters.
-Allowing the model to relocate these and/or incorporate flexibility measures, may lead to less dependency on both \carbon{} and \hydrogen{} networks.
-
-The technological flexibilities in the carbon sector might not be exploited to full extent.
-Therefore, the \carbon{} transport via truck or ship is not considered.
-Fischer-Tropsch facilities require to run on baseload with at least 90\% of their nominal capacity.
-
-\dots
-
-% TODO: which fixed rate? you mean that all emissions are captured by exogenous assumption or that we assume 90% of carbon can be captured?
-% Furthermore, the model assumes a fixed rate for industrial process emissions, which cannot be altered through investments.
-This simplification may not accurately represent real-world scenarios, where many industries are considering adopting low-carbon processes and technologies.
-The transport of fuel (FT, gas, oil) between the regions is not unlimited which may overestimate the flexibility provided by these commodities.
+The study conclusively demonstrates the pivotal role of strategically integrated hydrogen and carbon networks in steering Europe towards a sustainable and climate-neutral energy future.
+It emphasizes the necessity of comprehensive infrastructure planning, integrating diverse energy carriers and technologies to achieve these objectives effectively.
 

### 2023-12-07 report/submissions/journal/paper/paper.tex
paper: strip some comments and revise figure captions

@@
     \includegraphics[width=\linewidth]{baseline/figures/90_nodes/total-demand-bar.png}
-    \caption{Exogenous demand assumptions.
-The figure shows the total annual energy demands for each energy carrier which drive all model activities.
-Endogenous processes may lead to higher total production volumes of some carriers, e.g., demand in methanol requires more hydrogen and carbon as secondary (energy) inputs, which in turn need additional electricity to run electrolysis and carbon capture processes.}
+    \caption{Assumptions on exogenous demand, derived from~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023}.
+The figure shows the total annual energy demand for each energy source which determine the model's endogenous investments and operation.
+Endogenous processes can lead to higher total production volumes of some energy carriers, e.g., the demand for methanol requires more hydrogen and carbon as secondary (energy) inputs, which are not considered here.
+The model are defined per region and time stamp.}
     % TODO: adjust labels to show that these are exogenous assumptions
@@
     \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/cost_bar.png}
-    \caption[short]{Total annual system cost for the European energy system for the different models subdivided into groups of technologies.
+    \caption[short]{Total annual system cost subdivided into groups of technologies for the different models of the European energy system with a net-zero emission target.
 While the Baseline model has neither a \carbon{} nor a \hydrogen{} network, the Hybrid model is allowed to expand both.
@@
 System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
-% In the Baseline model, the absence of dedicated carbon and hydrogen transport networks means that the system cannot use all the CC potentials of geographically fixed point sources, but instead needs to deploy more DAC in the locations where carbon is needed for CS and CU.
-% Naturally, costs for the \carbon{} and \hydrogen{} networks differ between models as they are being deployed.
-% Finally, we see that the Baseline model deploys more wind and solar than the network models.
-This is driven by two aspects of the electrolytic hydrogen production: First, the Baseline model deploys more methanation, leading to a larger endogenous hydrogen demand and thus more electricity production.
-And second, in the network models, the electrolysis is deployed at sites with better renewable resources, thus decreasing the total investments in solar and wind power necessary to run the electrolysis.
 As we show below, these are due to substantial, albeit fundamentally different, gains in flexibility and a shift in carbon utilization and storage strategies by the \hydrogen{} and \carbon{} networks.
@@
 
-% 1.
-baseline model is sub-optimal -> un-used landlocked carbon capture potentials; scarcity in hydrogen and power in Central Europe due to industrial clusters and high population density; costly DAC + additional renewables at sequestration sites
 In comparison to the other three models, the \modBase{} reveals significant inefficiencies: On the one hand, regions with no access to sequestration sites (i.e., ``landlocked'' regions) underutilize CC potentials from process emissions, fossil gas and biomass.
@@
 
-% 2.
-introduce carbon balance figure -> transport systems unlock point-source CC potentials, in particular biomass and gas for industry, less methanation, less DAC and therefore less DAC specific renewables
 By enabling the efficient transportation of carbon, hydrogen, or both, these models reduce the reliance on high-cost DAC, and source more carbon from low-cost point-source CC on biomass and gas in industrial applications (see Fig.~\ref{fig:balance_captured_carbon}).
@@
     \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/balance_bar_carbon.png}
-    \caption{Balance of captured carbon across all system setups. Positive values indicate carbon capture, negative values indicate carbon consumption.}
-    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
+    \caption{Balance of captured carbon for all models in the Net-Zero scenario.
+Positive values indicate carbon capture, negative values indicate carbon consumption.
+By integrating \hydrogen{} and \carbon{} networks, the predominant method for carbon removal shifts from Direct Air Capture (DAC) to a bioenergetic process that incorporates carbon capture.
+At the same time the reliance on methanation decreases.}
     \label{fig:balance_captured_carbon}
@@
 
-% \begin{figure}[ht!]
-%     \centering
-%     \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/balance_bar_hydrogen.png}
-%     \caption{Balance of captured hydrogen across all system setups. Positive values indicate carbon capture, negative values indicate carbon consumption.}
-%     % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
-%     \label{fig:balance_hydrogen}
-% \end{figure}
-
-% 3. introduce capture share figure -> correlation between average cf and cc share, since gas CHPs and PPs are peak load techs, no cc infrastructure;
 The increase in captured carbon from point sources in the network models is mostly due to CC infrastructure being deployed on industrial or power plant facilities that are present in all models (as opposed to the deployment of additional electricity generating capacity for the purpose of carbon capture).
@@
     \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/captureshare_line.png}
-    \caption{Share of facilities with integrated carbon capture for different system setups for the Net-Zero scenario.}
+    \caption{Share of facilities with integrated carbon capture for all models .}
     % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
@@
 
-% 4. carbon-grid has higher absolute CC and CC shares than hydrogen-grid and less DAC however hydrogen-grid is more cost-effective
 In summary, the implementation of \carbon{} and \hydrogen{} networks is instrumental in unlocking low-cost CC potentials at geographically fixed point sources across the continent.
@@
 
-% In the \modCO{}, the carbon network enables the system to capture carbon from low-cost point sources and transport it to the sequestration sites, trading off CC costs with transport costs.
-
-% In summary, the \carbon{} network enables the transport of carbon from the lowest-cost CC sources to CU and CS carbon sinks, reducing the need for high-cost DAC.
-% The creation of a \hydrogen{} network enables the cost-effective delivery of hydrogen from regions with a surplus of low-cost renewable energy to high-demand industrial centers with exogenous hydrogen demand and a surplus of low-cost CC potentials, reducing the need for DAC and for high-cost electrolysis at the industrial centers.
-% This highlights how both the implementation of a \carbon{} or a \hydrogen{} network unlocks low-cost carbon capture potentials at geographically fixed point sources across the continent.
-
 \subsection*{\hydrogen{}--Network enables decentral carbon utilization}
-
-% \begin{figure*}[ht!]
-%     \centering
-%     \begin{subfigure}{0.9\linewidth}
-%         \centering
-%         \includegraphics[width=\linewidth]{co2-only/figures/90_nodes/balance_map_carbon.png}
-%         \caption{\carbon{} Sector.}
-%         \label{fig:balance_map_carbon}
-%     \end{subfigure}
-%     \begin{subfigure}{0.9\linewidth}
-%         \centering
-%         \includegraphics[width=\linewidth]{h2-only/figures/90_nodes/balance_map_hydrogen.png}
-%         \caption{\hydrogen{} Sector.}
-%         \label{fig:balance_map_hydrogen}
-%     \end{subfigure}
-%     \caption{Optimal operation per sector for a net-zero energy system in Europe with average production on the left and average consumption on the right for both, (a) the \carbon{} sector in the \carbon{}-Grid model and (b) the \hydrogen{} sector in the \hydrogen-Grid model.}
-%     \label{fig:balance_map}
-% \end{figure*}
 
@@
 
-The results illustrate the fundamentally different approaches of the grids: a \carbon{} grid prefers a decentralized CC system and a co-location of electrolysis and CU centrally in regions with abundant renewable resources; a \hydrogen{} grid prefers a centralized electrolysis system and a decentralized CC and CU system distributed jointly across the continent.
+The results illustrate the fundamentally different approaches of the grids: a \carbon{} grid prefers a decentralized CC system and a co-location of electrolysis and CU centrally in regions with abundant renewable resources; a \hydrogen{} grid prefers a centralized electrolysis system and a co-location of decentralized CC and CU systems across the continent.
 A \hydrogen{} network not only unlocks the potential of carbon capture from point sources, but also reduces the energy shortages in energy-intensive areas.
-
-% There are four notable dynamics or differences between the two grid strategies:
-% First, the production of FT fuels always moves to where low-cost carbon and low-cost hydrogen can be brought together.
-In the \carbongrid{} model, carbon is moved from Western Europe to the Iberian Peninsula, the British Isles and to Denmark to produce FT fuels centrally in few regions.
-In the \hydrogengrid{} model, hydrogen is transported in the opposite direction to produce FT fuels decentrally across all regions in Central Europe with locally captured \carbon{} from point sources.
-% Second, the \carbongrid{} model captures more low-cost carbon from biomass, especially in Central and Eastern Europe, and less high-cost carbon from DAC in Greece and Southern Italy.
-Instead, the carbon grid transports carbon from Central and Eastern Europe to Greece and Italy, where it is sequestered, or feeds into CU together with locally produced, low-cost electrolytic hydrogen.
-The \hydrogengrid{} model directly operates electrolyzers and DAC in Greece and Southern Italy, with electricity from low-cost renewables, and thus avoids significant network buildouts in this region.
-% Third, sequestration in the \hydrogengrid{} model occurs more decentralized, spread out across several regions with access to offshore sequestration sites, making use of point-source CC potentials in each of these regions.
-In the \carbongrid{} model, sequestration is clustered at fewer sequestration sinks, each collecting low-cost \carbon{} from neighboring regions.
-However, there is no centralization at only one or two sites, since all the sequestration sites offer the same costs and the system minimizes system-wide carbon transport distances to about eight to ten sequestration sites.
-% And fourth, how the three types of CU are spread across the continent is driven by the costs of carbon and hydrogen (the latter mostly driven by the costs of renewable electricity) as well as by renewables capacity factors: Clearly, all carbonaceous fuel production requires sites where both carbon and hydrogen are available, and they moreover prefer sites with low carbon and hydrogen prices.
-Based on our model assumptions, FT fuel synthesis has very limited flexibility and needs to be run alost at baseload (ramp-down only possible to 90\% of maximum load).
-Thus, for FT fuel synthesis sites in Northern Europe with more electricity from wind power are preferred to sites in Southern Europe, powered mostly by solar.
-Methanathion and Methanolisation both have higher flexibility, being able to ramp down to 50\% load during the night, for example.
-Thus, these fuel production sites prefer the best renewable resource sites in Southern Europe.
-And finally, as methanation produces synthetic methane that needs to be fed into the gas grid, methanation prefers those sites with good interconnection to the existing gas network.
-Since Italy possesses a large gas network with easy interconnection to Central Europe, methanation prefers sites in Southern Italy (and Greece, which would be connected to Southern Italy) over the Iberian Peninsula, where Methanolisation is more prevalent.
-
-% In summary, the implementation of a \carbon{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce low-cost hydrogen to make CU products (FT fuel, methanol, synthetic methane).
-DAC and electrolysis are deployed in the regions with the best renewable resources to meet carbonaceous fuel demands.
-% In the absence of a \carbon{} network and with a \hydrogengrid{} on the other hand, regions with access to offshore sequestration sites capture all carbon from low-cost point sources and then sequester it.
-The system deploys additional DAC facilities at the locations with sequestration access and low-cost renewables to capture the remaining necessary carbon for sequestration.
-Carbonaceous fuel demand is met at the locations which offer the best mix of low-cost carbon from point sources and low-cost hydrogen from renewables, with some additional electrolysis and DAC required to fully meet the demand.
-% Sequestered carbon is nearly always transported before
-
-% Therefore, one can conclude that a \carbon{} network favors a decentralized CC and a CU centralized at regions with abundant renewable resources.
-A \hydrogen{} network favors a centralized electrolysis system at regions with abundant renewable resources and a decentralized CU system.
-
-% Together with understanding the dynamics described above, we can highlight three cost advantages of the \hydrogen{} network in contrast to the \carbon{} network: Capturing carbon at the lowest-cost regions reduces 4 billion euros in spending on DAC.
-However, the \hydrogengrid{} model produces electrolytic hydrogen in the lowest-cost regions and in total spends 7 billion euros less on renewable energy (14 billion euros less on solar, and 7 billion euros more on wind) as well as 1 billion euros less on SMR.
-Furthermore, without a \hydrogen{} network, additional \hydrogen{} storage is needed in Spain at a cost of 5 billion euros to store hydrogen produced during the summer for utilization in FT fuel production during the winter.
-% And finally, the \carbongrid{} model spends 7 billion euros more on gas boilers, gas plants, gas infrastructure, and methanation.
-% TODO: why??
-% The costs of the respective networks, on the other hand, are almost equal: the \carbongrid{} model spends 8 billion euros on the \carbon{} network, while the \hydrogengrid{} model even spends 10 billion euros on the \hydrogen{} network.
-These factors add up to account for 14 billion euros of the difference of 20 billion euros in system costs between the two scenarios (785 billion euros for the \carbongrid{} and 765 billion euros for the \hydrogengrid{} model, see Fig.~\ref{fig:objective_heatmap}).
 
@@
 
-\begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
-    \caption{Sankey diagram of the optimal operation for a net-negative 10\% scenario.}
-    \label{fig:sankey_diagramm}
-\end{figure*}
+% \begin{figure*}
+%     \centering
+%     \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
+%     \caption{Sankey diagram of the optimal operation for a net-negative 10\% scenario.}
+%     \label{fig:sankey_diagramm}
+% \end{figure*}
 

### 2023-12-07 report/submissions/journal/paper/paper.tex
workflow: rename balance_map_single to balance_map; remove double balance_map

@@
         \centering
-        \includegraphics[width=\linewidth]{full/figures/90_nodes/balance_map_single_hydrogen.png}
+        \includegraphics[width=\linewidth]{full/figures/90_nodes/balance_map_hydrogen.png}
         \caption{\hydrogen{} Sector.}
@@
         \centering
-        \includegraphics[width=\linewidth]{full/figures/90_nodes/balance_map_single_carbon.png}
+        \includegraphics[width=\linewidth]{full/figures/90_nodes/balance_map_carbon.png}
         \caption{\carbon{} Sector.}

### 2023-12-07 workflow/scripts/plot_balance_map.py

@@ -12,12 +12,11 @@ import pandas as pd
 import matplotlib.pyplot as plt
 import cartopy.crs as ccrs
 from pypsa.plot import add_legend_circles, add_legend_patches, add_legend_lines
-from matplotlib.gridspec import GridSpec
+from pypsa.statistics import get_transmission_carriers
 from common import (
     import_network,
     mock_snakemake,
 )
-from pypsa.statistics import get_transmission_carriers
 import geopandas as gpd
 
 
@@ -32,31 +31,28 @@ if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     )
 
 sns.set_theme(**snakemake.params["theme"])
-plt.rc("patch", linewidth=0.05)
+plt.rc("patch", linewidth=0.1)
 
 n = import_network(snakemake.input.network, revert_dac=True)
 regions = gpd.read_file(snakemake.input.onshore_regions).set_index("name")
-which = "operation"
 config = snakemake.config
-run = snakemake.wildcards.run
 labels = config["labels"]
 
-alpha = 1
-region_alpha = 0.6
+alpha = config["plotting"]["balance_map"]["alpha"]
+region_alpha = config["plotting"]["balance_map"]["region_alpha"]
 
 s = n.statistics
 colors = n.carriers.set_index("nice_name").color
 
+run = snakemake.wildcards.run
+
 for kind, output in snakemake.output.items():
-    # Create a GridSpec object
-    gs = GridSpec(1, 2, width_ratios=[20, 20])
 
-    fig = plt.figure(figsize=snakemake.params.settings["figsize"]["double"])
-
-    # Create the first axes in the first column
-    ax1 = fig.add_subplot(gs[0, 0], projection=ccrs.EqualEarth())
-    ax2 = fig.add_subplot(gs[0, 1], projection=ccrs.EqualEarth())
-    axes = [ax1, ax2]
+    fig, ax = plt.subplots(
+        figsize=snakemake.params.settings["figsize"],
+        subplot_kw={"projection": ccrs.EqualEarth()},
+        layout="constrained",
+    )
 
     carriers = config["constants"]["carrier_to_buses"].get(kind, [kind])
 
@@ -72,139 +68,155 @@ for kind, output in snakemake.output.items():
     df = pd.concat([df.drop("Link"), sub])
     df = df.rename(lambda x: x.replace(" CC", ""), level=2)
[…]

### 2023-12-07 workflow/scripts/plot_balance_map_single.py

@@ -1,224 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Thu Nov 24 10:14:48 2022
-
-@author: fabian
-"""
-import os
-import warnings
-import seaborn as sns
-import pandas as pd
-import matplotlib.pyplot as plt
-import cartopy.crs as ccrs
-from pypsa.plot import add_legend_circles, add_legend_patches, add_legend_lines
-from pypsa.statistics import get_transmission_carriers
-from common import (
-    import_network,
-    mock_snakemake,
-)
-import geopandas as gpd
-
-
-warnings.filterwarnings("ignore", category=UserWarning)
-
-if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
-    snakemake = mock_snakemake(
-        "plot_balance_map_single",
-        run="full",
-        clusters=90,
-        ext="png",
-    )
-
-sns.set_theme(**snakemake.params["theme"])
-plt.rc("patch", linewidth=0.1)
-
-n = import_network(snakemake.input.network, revert_dac=True)
-regions = gpd.read_file(snakemake.input.onshore_regions).set_index("name")
-config = snakemake.config
-labels = config["labels"]
-
-alpha = config["plotting"]["balance_map"]["alpha"]
-region_alpha = config["plotting"]["balance_map"]["region_alpha"]
-
-s = n.statistics
-colors = n.carriers.set_index("nice_name").color
-
-run = snakemake.wildcards.run
-# %%
-for kind, output in snakemake.output.items():
-
-    figsize = snakemake.params.settings["figsize"]["single"]
-    fig, ax = plt.subplots(
-        figsize=figsize,
-        subplot_kw={"projection": ccrs.EqualEarth()},
-        layout="constrained",
-    )
-
-    carriers = config["constants"]["carrier_to_buses"].get(kind, [kind])
-
-    grouper = s.groupers.get_bus_and_carrier
-    df = s.dispatch(bus_carrier=carriers, groupby=grouper)
-    _ = get_transmission_carriers(n, bus_carrier=carriers)
-    transmission_carriers = _.set_levels(
-        n.carriers.nice_name[_.get_level_values(1)], level=1
-    )
-    sub = df.loc[["Link"]].drop(
-        transmission_carriers.unique(1), level=2, errors="ignore"
-    )
-    df = pd.concat([df.drop("Link"), sub])
-    df = df.rename(lambda x: x.replace(" CC", ""), level=2)
[…]

### 2023-12-08 workflow/scripts/plot_balance_map.py
revise figures; adjust according to comments; add graphics rule

@@ -159,6 +159,7 @@ for kind, output in snakemake.output.items():
 
     pad = 0.18
     carriers = n.carriers.set_index("nice_name")
+    carriers.loc["", "color"] = "None"
     prod_carriers = bus_sizes[bus_sizes > 0].index.unique("carrier").sort_values()
     cons_carriers = (
         bus_sizes[bus_sizes < 0]
@@ -166,6 +167,14 @@ for kind, output in snakemake.output.items():
         .difference(prod_carriers)
         .sort_values()
     )
+
+    if kind == "hydrogen":
+        # fix bug related to falsely clipped normalization
+        if kind == "hydrogen" and "H$_2$ For Industry" in prod_carriers:
+            prod_carriers = prod_carriers.difference(["H$_2$ For Industry"])
+            cons_carriers = cons_carriers.union(["H$_2$ For Industry"])
+        # cheat and add a dummy carrier to align legend with carbon balance
+        cons_carriers = list(cons_carriers) + ["", ""]
 
     add_legend_patches(
         ax,

### 2023-12-08 workflow/scripts/plot_balance_map_dedicated.py

@@ -53,7 +53,7 @@ fig, axes = plt.subplots(
 for n, axs, draw_legend in zip(networks, axes.T, [False, True]):
     s = n.statistics
     colors = n.carriers.set_index("nice_name").color
-    run = snakemake.wildcards.run
+    run = n.meta["wildcards"]["run"]
 
     for kind, ax in zip(["carbon", "hydrogen"], axs):
         carriers = config["constants"]["carrier_to_buses"].get(kind, [kind])
@@ -169,6 +169,11 @@ for n, axs, draw_legend in zip(networks, axes.T, [False, True]):
                 .sort_values()
             )
 
+            # fix bug related to falsely clipped normalization
+            if kind == "hydrogen" and "H$_2$ For Industry" in prod_carriers:
+                prod_carriers = prod_carriers.difference(["H$_2$ For Industry"])
+                cons_carriers = cons_carriers.union(["H$_2$ For Industry"])
+
             add_legend_patches(
                 ax,
                 colors[prod_carriers],
@@ -188,7 +193,7 @@ for n, axs, draw_legend in zip(networks, axes.T, [False, True]):
                 cons_carriers.map(wrapper.fill),
                 patch_kw={"alpha": alpha},
                 legend_kw={
-                    "bbox_to_anchor": (1, 0.4),
+                    "bbox_to_anchor": (1, 0.5),
                     "ncol": 1,
                     "title": "Consumption",
                     **legend_kwargs,
@@ -200,7 +205,10 @@ for n, axs, draw_legend in zip(networks, axes.T, [False, True]):
             ax,
             [s * bus_scale * 1e6 for s in legend_bus_sizes],
             [f"{s} {unit}" for s in legend_bus_sizes],
-            **legend_kwargs,
+            legend_kw={
+                "bbox_to_anchor": (0, 1),
+                **legend_kwargs,
+            },
         )
         legend_branch_sizes = specs["branch_sizes"]
         if legend_branch_sizes is not None:

### 2023-12-08 workflow/scripts/plot_cost_bar_transmission.py

@@ -20,7 +20,7 @@ if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
         "plot_cost_bar_transmission",
         ext="png",
         clusters=90,
-        comparison="emission-reduction-full",
+        comparison="subsidy",
     )
 
 sns.set_theme(**snakemake.params["theme"])
@@ -70,7 +70,7 @@ fig, ax = plt.subplots(
 )
 
 
-defaults = dict(kind="bar", stacked=True, lw=0, rot=0, alpha=0.8)
+defaults = dict(kind="bar", stacked=True, lw=0, rot=90, alpha=0.8)
 
 kwargs = {**defaults, **snakemake.params.settings.get("kwargs", {})}
 data.T.plot(ax=ax, color=colors.to_dict(), **kwargs)
@@ -78,7 +78,7 @@ for container in ax.containers:
     ax.bar_label(
         container,
         label_type="center",
-        fmt=lambda x: int(round(x, 0)),
+        fmt=lambda x: int(round(x, 0)) if x > 0 else "",
         fontsize=7,
         color="grey",
     )

### 2023-12-08 workflow/scripts/plot_demand_bar.py

@@ -11,11 +11,9 @@ import seaborn as sns
 from common import import_network
 import matplotlib.pyplot as plt
 
-sns.set_theme(style="white", context="paper")
+sns.set_theme(**snakemake.params["theme"])
 
-n = import_network(
-    "/home/fabian/papers/co2-network/results/baseline/networks/90_nodes.nc"
-)
+n = import_network(snakemake.input.network)
 s = n.statistics
 
 demand = s.withdrawal(groupby=s.groupers.get_carrier_and_bus_carrier).Load
@@ -61,6 +59,6 @@ ax.legend(title="", frameon=False, bbox_to_anchor=(0, 1.2), loc="upper left")
 sns.despine()
 
 fig.savefig(
-    "/home/fabian/papers/co2-network/results/baseline/figures/90_nodes/total-demand-bar.png",
+    snakemake.output.figure,
     bbox_inches="tight",
 )

### 2023-12-08 report/submissions/journal/paper/paper.tex
revise figures; adjust according to comments; add graphics rule
paper: update section 3.2 (still some repetitoions), consistent abbreviations for H2, CO2

@@
 % make abbreviation for co2
-\newcommand{\carbon}{CO$_2$}
-\newcommand{\CO}{CO$_2$}
-\newcommand{\hydrogen}{H$_2$}
+\newcommand{\COtwo}{CO$_2$}
 \newcommand{\Htwo}{H$_2$}
-\newcommand{\carbongrid}{\carbon{}--Grid}
-\newcommand{\hydrogengrid}{\hydrogen{}--Grid}
+\newcommand{\COgrid}{CO$_2$\=/Grid}
+\newcommand{\Hgrid}{H$_2$\=/Grid}
 \newcommand{\modBase}{Baseline model}
-\newcommand{\modCO}{CO$_2$-Grid model}
-\newcommand{\modH}{H$_2$-Grid model}
+\newcommand{\modCO}{CO$_2$\=/Grid model}
+\newcommand{\modH}{H$_2$\=/Grid model}
 \newcommand{\modHybrid}{Hybrid model}
@@
 
-\title{Assessing \hydrogen{} and \carbon{} Network Strategies in Europe}
+\title{Assessing \Htwo{} and \COtwo{} Network Strategies in Europe}
 % \title{Synthetic fuels in Europe: Transport Hydrogen to Carbon, or Carbon to Hydrogen?}
@@
 The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors.
-While management of both \carbon{} and \hydrogen{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructures effectively interact with comprehensive carbon management technologies, including carbon capture, transport, use, and storage.
+While management of both \COtwo{} and \Htwo{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructures effectively interact with comprehensive carbon management technologies, including carbon capture, transport, use, and storage.
 Hydrogen offers an efficient way to transport and store energy.
@@
 Additionally, it can be stored in geological formations, a process known as carbon sequestration (CS).
-In combination, \carbon{} and \hydrogen{} networks build the basis for climate-neutral fuels needed for aviation, shipping and industrial feedstocks, and are central to carbon utilization (CU) strategies.
+In combination, \COtwo{} and \Htwo{} networks build the basis for climate-neutral fuels needed for aviation, shipping and industrial feedstocks, and are central to carbon utilization (CU) strategies.
 
@@
 Initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen Infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier, and some gas pipelines are already repurposed to transport hydrogen~\cite{RohrFreiFuer}.
-At the same time, business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synfuels on the one hand and offer \carbon{} offtake on the other hand.
-The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a transport and sequestration capacity of 1.5 Mt \carbon{} per year  to be operative in 2024, expanding to a targeted scale of 5 Mt per year of sequestration by 2030.
+At the same time, business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synfuels on the one hand and offer \COtwo{} offtake on the other hand.
+The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a transport and sequestration capacity of 1.5 Mt \COtwo{} per year  to be operative in 2024, expanding to a targeted scale of 5 Mt per year of sequestration by 2030.
 % TODO: frame next two sentences to be more market oriented
@@
 (extend on hydrogen literature)
-The publication by Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \carbon{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
-This limited sectoral scope cannot capture important dynamics of carbon management, since it neglects the sectors that will need to handle most \carbon{} in the future.
-Another comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
-Hoewever, this limited spatial scope fails to consider other sequestration sites and co-benefits from connecting the \carbon{} network to neighboring countries.
-To this end, it is often argued that \carbon{} pipelines are a mature technology with an expected high learning rate, given the wide-spread installations in US and Canada for enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
+The publication by Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \COtwo{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
+This limited sectoral scope cannot capture important dynamics of carbon management, since it neglects the sectors that will need to handle most \COtwo{} in the future.
+Another comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \COtwo{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
+Hoewever, this limited spatial scope fails to consider other sequestration sites and co-benefits from connecting the \COtwo{} network to neighboring countries.
+To this end, it is often argued that \COtwo{} pipelines are a mature technology with an expected high learning rate, given the wide-spread installations in US and Canada for enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
 % However, the models are often limited with regard to both geographical scope and detail.
@@
 
-To our knowledge, no study has yet considered the co-optimization and comprehensive assessment of both \carbon{} and \hydrogen{} networks in a fully sector-coupled energy system.
+To our knowledge, no study has yet considered the co-optimization and comprehensive assessment of both \COtwo{} and \Htwo{} networks in a fully sector-coupled energy system.
 However, we argue that such an assessment is strongly needed to avoid suboptimal investments and to identify synergies between hydrogen and carbon management technologies.
@@
 The study is conducted using the PyPSA-Eur energy system model and encompasses all relevant energy sectors.
-We explore competition between \carbon{} and \hydrogen{} networks by adding them separately.
-Our evaluation focuses on the transport of \carbon{} and \hydrogen{} through their respective networks on the European continent.
+We explore competition between \COtwo{} and \Htwo{} networks by adding them separately.
+Our evaluation focuses on the transport of \COtwo{} and \Htwo{} through their respective networks on the European continent.
 We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available sequestration potential.
@@
 \begin{figure}
-    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/total-demand-bar.png}
+    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/demand_bar.png}
     \caption{Assumptions on exogenous demand, derived from~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023}.
@@
 We show the sum of all energy demands in Fig.~\ref{fig:total-demand-bar}.
-The system emits 633~Mt \carbon{} per year from industry, aviation, shipping and agriculture, 153~Mt of which are fossil-based process emissions.
+The system emits 633~Mt \COtwo{} per year from industry, aviation, shipping and agriculture, 153~Mt of which are fossil-based process emissions.
 Industrial energy demand and excess heat potentials are calculated per node on the basis of~\cite{hotmaps_industrial_db}.
@@
 In our model, hydrogen can be produced from electrolysis and steam methane reforming (SMR).
-Geological distribution and potentials for \hydrogen{} storage in salt caverns are derived from~\cite{caglayanTechnicalPotentialSalt2020}.
+Geological distribution and potentials for \Htwo{} storage in salt caverns are derived from~\cite{caglayanTechnicalPotentialSalt2020}.
 Re-electrification of hydrogen is possible via fuel cells.
@@
 
-The topology of the \carbon{}, \hydrogen{} and gas network is identical to the topology of the electricity network, connecting all neighboring regions.
+The topology of the \COtwo{}, \Htwo{} and gas network is identical to the topology of the electricity network, connecting all neighboring regions.
 Transport of liquid fuels like oil, methanol and Fischer-Tropsch (FT) fuels is ``copper-plated'', since transport costs per unit of energy are negligible due to their high energy density.
-Throughout this study, we focus on \carbon{} and \hydrogen{} networks because of their high relevance for infrastructure investments and thus public policy decisions.
+Throughout this study, we focus on \COtwo{} and \Htwo{} networks because of their high relevance for infrastructure investments and thus public policy decisions.
 The electricity network and gas network are both included in the model with full geographical detail, but not further analyzed.
@@
 Moving up the scale, gas used in industrial applications and biomass employed for industrial processes represent more costly yet viable options to capture carbon.
-DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand.
-Finally, based on our model results, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per marginal ton of captured \carbon{}.
+DAC, an emerging technology capable of extracting \COtwo{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand.
+Finally, based on our model results, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per marginal ton of captured \COtwo{}.
 Biogas input is a high-cost fuel which based on the endogenous modeling decisions is not used in large quantities.
 To the extent that biogas-to-gas facilities are used by the model to supply additional (carbon-neutral) gas as fuel, adding CC infrastructure incurs low costs (and thus ranges on the left-hand side of the merit order curve).
-However, the price of capturing additional \carbon{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the marginal cost of captured \carbon{}.
+However, the price of capturing additional \COtwo{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the marginal cost of captured \COtwo{}.
 This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.
-If we consider the spatial aspect of distributed carbon capture potentials, the \carbon{} network can exploit comparative cost advantages between CO2 bidding zones to utilize the lowest-cost CC facilities across the continent.
+If we consider the spatial aspect of distributed carbon capture potentials, the \COtwo{} network can exploit comparative cost advantages between CO2 bidding zones to utilize the lowest-cost CC facilities across the continent.
 We assume a capture rate of 90\% for CC on process emissions, SMR, biogas-to-gas, as well as gas and biomass used in industry, and 95\% for CC on biomass and gas CHPs.
@@
 Our estimates for carbon sequestration potentials are conservative, limiting total sequestration to 25~Mt per site and calculating annual storage availability over 25 years.
-Furthermore, we decide to cap the total sequestered \carbon{} to 200~Mt per year for our Net-Zero scenario.
+Furthermore, we decide to cap the total sequestered \COtwo{} to 200~Mt per year for our Net-Zero scenario.
 This is enough to offset the hardest-to-abate fossil process emissions and some limited slack of fossil fuels to be used where they can be most valuable.
@@
 
-Addressing climate targets, we define two \carbon{} target scenarios:
+Addressing climate targets, we define two \COtwo{} target scenarios:
 \begin{itemize}
-    \item[] \textit{Net--Zero}: aligning with the EU's 2050 emission targets. Also denoted as NZ.
-    \item[] \textit{Net--Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \carbon{} annually. Also denoted as NN.
+    \item[] \textit{Net\=/Zero}: aligning with the EU's 2050 emission targets. Also denoted as NZ.
+    \item[] \textit{Net\=/Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \COtwo{} annually. Also denoted as NN.
 \end{itemize}
@@
 
-Beyond \carbon{} targets, we consider four expansion scenarios, referred to as models:
+Beyond \COtwo{} targets, we consider four expansion scenarios, referred to as models:
 \begin{itemize}
-    \item[] \textit{Baseline:} Neither \carbon{} nor \hydrogen{} networks are constructed.
-    \item[] \textit{\carbongrid:} Only the \carbon{} network is developed.
-    \item[] \textit{\hydrogengrid:} Only the \hydrogen{} network is developed.
-    \item[] \textit{Hybrid:} Both \carbon{} and \hydrogen{} networks are developed.
+    \item[] \textit{Baseline:} Neither \COtwo{} nor \Htwo{} networks are constructed.
+    \item[] \textit{\COgrid{}:} Only the \COtwo{} network is developed.
+    \item[] \textit{\Hgrid{}:} Only the \Htwo{} network is developed.
+    \item[] \textit{Hybrid:} Both \COtwo{} and \Htwo{} networks are developed.
 \end{itemize}
@@
     \caption[short]{Total annual system cost subdivided into groups of technologies for the different models of the European energy system with a net-zero emission target.
-While the Baseline model has neither a \carbon{} nor a \hydrogen{} network, the Hybrid model is allowed to expand both.
-``Gas Infrastructure'' combines gas facilities for power and heat production, ``\carbon{} Infrastrucure'' combines transport, storage and sequestration, and ``H$_2$ Infrastructure'' combines transport and storage.
+While the \modBase{} has neither a \COtwo{} nor a \Htwo{} network, the \modHybrid{} is allowed to expand both.
+``Gas Infrastructure'' combines gas facilities for power and heat production, ``\COtwo{} Infrastructure'' combines transport, storage and sequestration, and ``\Htwo{} Infrastructure'' combines transport and storage.
 ``Carbon Capture at Point Sources'' combines all technologies with integrated carbon capture, including the cost of the main facility (e.g., CHP unit) and the carbon capture application.}
@@
 
-In all models, almost half of the system costs are spent on primary electricity production from wind and solar power (\label{}45\% or \label{}360 of \label{}799 billion euros total costs in the Baseline model), with another \label{}12\% (\label{}100 billion euros) spent on hydro, nuclear and biomass.
+In all models, almost half of the system costs are spent on primary electricity production from wind and solar power (\label{}45\% or \label{}360 of \label{}799 billion euros total costs in the \modBase{}), with another \label{}12\% (\label{}100 billion euros) spent on hydro, nuclear and biomass.
 Costs for energy transport networks for electricity, gas, hydrogen and carbon are below \label{}12\% (\label{}100 billion euros), with the electricity grid making up more of the system costs than the other three networks combined.
@@
 
-System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \carbon{} and \hydrogen{} networks, as well as carbon capture technologies (DAC and CC at point sources).
-As we show below, these are due to substantial, albeit fundamentally different, gains in flexibility and a shift in carbon utilization and storage strategies by the \hydrogen{} and \carbon{} networks.
+System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \COtwo{} and \Htwo{} networks, as well as carbon capture technologies (DAC and CC at point sources).
+As we show below, these are due to substantial, albeit fundamentally different, gains in flexibility and a shift in carbon utilization and storage strategies by the \Htwo{} and \COtwo{} networks.
 
-\subsection*{Transport Systems Unlock Point\=/Source Carbon Capture}
+\subsection*{Transport systems unlock point\=/source carbon capture}
 
@@
 By enabling the efficient transportation of carbon, hydrogen, or both, these models reduce the reliance on high-cost DAC, and source more carbon from low-cost point-source CC on biomass and gas in industrial applications (see Fig.~\ref{fig:balance_captured_carbon}).
-The \modH{} only captures \label{}200~Mt of \carbon{} from DAC compared to \label{}400~Mt in the \modBase{}.
+The \modH{} only captures \label{}200~Mt of \COtwo{} from DAC compared to \label{}400~Mt in the \modBase{}.
 Instead, the \modH{} captures more carbon from process emissions as well as gas and biomass for industry.
@@
 Positive values indicate carbon capture, negative values indicate carbon consumption.
-By integrating \hydrogen{} and \carbon{} networks, the predominant method for carbon removal shifts from Direct Air Capture (DAC) to a bioenergetic process that incorporates carbon capture.
+By integrating \Htwo{} and \COtwo{} networks, the predominant method for carbon removal shifts from Direct Air Capture (DAC) to a bioenergetic process that incorporates carbon capture.
 At the same time the reliance on methanation decreases.}
@@
 
-In summary, the implementation of \carbon{} and \hydrogen{} networks is instrumental in unlocking low-cost CC potentials at geographically fixed point sources across the continent.
-While the \carbon{} network facilitates the movement of carbon from low-cost sources to CU and CS sinks, reducing the reliance on high-cost DAC, the \hydrogen{} network enables the cost-effective delivery of hydrogen from regions rich in renewable energy to regions with low-cost CC potentials to produce CU.
-However, despite the \carbongrid{} model demonstrating higher absolute amounts of CC from point sources and higher CC shares across all point-source CC technologies, the \hydrogengrid{} model achieves lower total system costs (see Fig.~\ref{fig:cost_bar}).
+In summary, the implementation of \COtwo{} and \Htwo{} networks is instrumental in unlocking low-cost CC potentials at geographically fixed point sources across the continent.
+While the \COtwo{} network facilitates the movement of carbon from low-cost sources to CU and CS sinks, reducing the reliance on high-cost DAC, the \Htwo{} network enables the cost-effective delivery of hydrogen from regions rich in renewable energy to regions with low-cost CC potentials to produce CU.
+However, despite the \modCO{} demonstrating higher absolute amounts of CC from point sources and higher CC shares across all point-source CC technologies, the \modH{} achieves lower total system costs (see Fig.~\ref{fig:cost_bar}).
 In the following section, we will explore the reasons for this.
@@
 
-\subsection*{\hydrogen{}--Network enables decentral carbon utilization}
+% In the \modCO{}, the carbon network enables the system to capture carbon from low-cost point sources and transport it to the sequestration sites, trading off CC costs with transport costs.
+
+% In summary, the \COtwo{} network enables the transport of carbon from the lowest-cost CC sources to CU and CS carbon sinks, reducing the need for high-cost DAC.
+% The creation of a \Htwo{} network enables the cost-effective delivery of hydrogen from regions with a surplus of low-cost renewable energy to high-demand industrial centers with exogenous hydrogen demand and a surplus of low-cost CC potentials, reducing the need for DAC and for high-cost electrolysis at the industrial centers.
+% This highlights how both the implementation of a \COtwo{} or a \Htwo{} network unlocks low-cost carbon capture potentials at geographically fixed point sources across the continent.
+
+\subsection*{\Htwo{} network incentivizes distributed carbon utilization}\label{subsec:H2}
 
@@
     \includegraphics[width=\linewidth]{comparison/single-technologies/figures/90_nodes/balance_map_dedicated.png}
-    \caption{Optimal operation per sector for a net-zero energy system in Europe with average production on the left and average consumption on the right for both, (a) the \carbon{} sector in the \carbon{}-Grid model and (b) the \hydrogen{} sector in the \hydrogen-Grid model.}
+    \caption{Optimal operation per sector for a net-zero energy system in Europe with average production on the left and average consumption on the right for both, (a) the \COtwo{} sector in the \modCO{} and (b) the \Htwo{} sector in the \modH{}.}
     \label{fig:balance_map}
@@
 
-% Examining the cost-reductions across the different models raises an important question: why is the hydrogen network more cost-effective than its carbon counterpart, and how does it achieve similar cost reductions as the hybrid model?
-An initial hypothesis may be that the cost assumptions for \carbon{}
-%  and \hydrogen{} networks play a role.
-However, the amount of system costs spent on either network is substantially lower than the cost difference between the models (20 billion euros less total system costs in the \hydrogengrid{} model, while it spends 10 billion euros on the \hydrogen{} grid and the \carbongrid{} model spends 8 billion euros on the \carbon{} grid).
-Furthermore, we have conducted a sensitivity analysis with cost assumptions for the \carbon{} network reduced to 50\% of the default assumptions, and the results show that the \hydrogen{} network still outperforms the \carbon{} network in terms of total system costs.
-To find the reason for why the \hydrogen{} network achieves lower system costs than the \carbon{} network, it is crucial to understand the underlying mechanisms and geographical implications of the two transportation systems.
-Fig.~\ref{fig:balance_map} maps the average production, consumption and flow in the carbon (upper row) and hydrogen (lower row) sectors for the \carbongrid{} (left) and \hydrogen{} (right) model.
-Each map includes circles indicating the production (top half-circle) and consumption (bottom half-circle) for each region, with lines representing the import/export volumes and flow directions.
-The color of each region corresponds to the average price of the carrier.
+To find the reason for why the \Htwo{} network achieves lower system costs than the \COtwo{} network, it is crucial to understand the underlying mechanisms and geographical implications of the two transportation systems.
+Fig.~\ref{fig:balance_map} maps the average production, consumption and flow in the carbon (top row) and hydrogen (bottom row) sectors for the \modCO{} (left) and \modH{} (right).
+Each map displays circles indicating the production (top half-circle) and consumption (bottom half-circle) for each region, with lines and arrows representing the average import or export volumes and flow directions between regions.
+The color shading of each region corresponds to the average price of the carrier.
 
-The \carbongrid{} model captures large amounts of low-cost carbon from bioenergy and industrial activities, especially in Central and Eastern Europe (top left), almost none of which is used locally.
-A large part is transported from Western Europe to the Iberian Peninsula, the British Isles and Denmark, where it is used to supplement low-cost hydrogen in central plants for electrolysis, Fischer-Tropsch and methanolization (bottom left).
-A second part is transported to methanation plants such as those in Italy and Greece, from where the gas is fed into the gas grid, which also is set up for distribution from large gas import hubs.
-Finally, the third part is transported directly to sites in the North Sea, Baltic Sea and Mediterranean Sea where it is eventually enriched with \carbon{} from DAC to be finally sequestered.
-In absence of a \hydrogen{} network, the system distributes electrolysis where it is needed for industrial processes and fuel cell land transport and emphasizes gas imports into Central Europe.
+As we have also discussed above, the \modCO{} captures large amounts of low-cost carbon from bioenergy and industrial activities that were untapped in the \modBase{}, especially in Central and Eastern Europe, and the \COtwo{} network smoothes out price differences for captured carbon across regions (see Fig.~\ref{fig:balance_map}, top left).
+Hydrogen prices still vary strongly, with low price ``valleys'' in the regions with good renewable resources such as the Iberian Peninsula or Denmark, and high price ``peaks'' in Central Europe (see bottom left).
+The model optimally locates CU in the regions where it can minimize costs for its material inputs, carbon and hydrogen.
+Thus, when carbon prices are smoothed out in the \modCO{}, CU moves to the hydrogen price valleys in the periphery of the continent.
+The captured carbon is transported from Central Europe through several large pipeline corridors to these locations to produce carbonaceous fuels.
+Similarly to CU, the model optimally locates CS in regions where it can minimize costs for its only material input, carbon.
+Thus, CS moves to the carbon price valleys (which are not strongly pronounced due to the \COtwo{} network, but still prevalent) among the subset of regions with access to offshore sequestration sites.
+In the \modCO, these are sites  sites in the North Sea (Northern Germany, Norway, Southeastern England), Baltic Sea (Latvia, Eastern Denmark) and Mediterranean Sea (Italy, Greece).
+CU and CS compete for carbon sources and thus there is some interaction for their optimal placement.
+Finally, in the absence of an \Htwo{} network, the system places some electrolysis and SMR in Central Europe to supply exogenous hydrogen demand for industrial processes and fuel cell land transport (see bottom left).
+Due to the additional electricity demand for electrolysis, the model switches some winter heating from heat pumps to gas\=/fueled boilers and CHPs, and thus emphasizes overall more gas imports into Central Europe.
 
-The \hydrogengrid{} on the other hand, transports low-cost hydrogen from regions with abundant renewable resources to regions with high energy demand and abundant emissions (lower right).
-The main transportation routes go from Spain and the United Kingdom to Central Europe.
-In these regions, industries use the hydrogen for demand in industry and fuel cell land transport, and to produce synfuels with locally captured carbon from point sources, leading to a decentralized CU system across the continent.
-DAC and local emissions nearby the coast enable sequestration in the North Sea, Baltic Sea, Mediterranean Sea and, in particular, in Portugal (upper right).
-% Furthermore, the introduction of a hydrogen network enables to reduce both DAC production and FT fuels on the British Isles and Iberian Peninsula.
-In Great Britain an Ireland, the remaining local carbon is then sequestered, and all the hydrogen that previously went into FT fuel production is exported to Central Europe (see above).
-In Spain and Portugal, some DAC remains and the carbon sourced from DAC is used, along with locally captured carbon from point sources, to produce FT fuels (but less than without a hydrogen network).
-Surplus hydrogen from lower FT fuel production is exported to Central Europe.
+The \Htwo{} network on the other hand, smoothes out hydrogen prices across the entire continent (see bottom right).
+Thus, the model chooses to locate CU in the carbon price valleys across Western, Central and Eastern Europe (see top right).
+Low-cost hydrogen from regions with abundant renewable resources is transported to these regions with low-price carbon from abundant sources of fossil, biogenic and process emissions for point-source CC.
+The \Htwo{} network is also used to supply the exogenous \Htwo{} demand in these regions for industry and fuel cell land transport at much lower costs than local electrolysis or SMR.
+The model locates CS again in the regions that represent carbon price valleys among all those with access to offshore sequestration sites.
+In the case of the \modH{}, due to the lack of a \COtwo{} network all CC potentials from point sources are exhausted in all offshore sequestration regions, and the carbon is then sequestered.
+The additionally needed carbon to reach 200~Mt of CS is captured via DAC in the regions offshore sequestration regions with the lowest costs for DAC, with the largest share deployed in Portugal (see top right).
 
-The results illustrate the fundamentally different approaches of the grids: a \carbon{} grid prefers a decentralized CC system and a co-location of electrolysis and CU centrally in regions with abundant renewable resources; a \hydrogen{} grid prefers a centralized electrolysis system and a co-location of decentralized CC and CU systems across the continent.
-A \hydrogen{} network not only unlocks the potential of carbon capture from point sources, but also reduces the energy shortages in energy-intensive areas.
+The results illustrate the commonalities and differences between the two grids: both a \COtwo{} network and an \Htwo{} network prefer distributed CC at point sources and electrolysis in regions with abundant renewable resources, choosing the lowest-cost production sites for each while trading off transport costs.
+Nonetheless, the \COtwo{} network is forced to deploy some electrolysis in regions with poor renewables to supply their exogenous hydrogen demands for industry and fuel cell land transport.
 
-\subsection*{Hybrid configuration provides further flexibilities}
+The differences between the CO2-Grid and H2-Grid in the context of CU, CS and DAC are notable and highlight the unique operational characteristics of each grid type.
+CU always chooses the sites with the lowest costs for its hydrogen and carbon inputs.
+With the \COtwo{} network, \COtwo{} costs are levelized across the continent, and CU is thus located in the regions with the lowest \Htwo{} costs.
+With the \Htwo{} network, \Htwo{} costs are levelized across the continent, and CU is thus located in the regions with the lowest \COtwo{} costs.
+Since regions with low-cost \COtwo{} in the \modH{} are very spread out across the continent, the CU locations are also more distributed, while in the \modCO{} they cluster in a few pockets of low-cost \Htwo{}.
+CS co-locates in both models in the regions with the lowest \COtwo{} costs among those with offshore sequestration potential.
+However, due to the lack of \COtwo{} transport in the \modH{}, additional CS is located in the offshore locations where DAC, and therefore electricity, can be produced at the lowest costs.
+Finally, DAC in the \modCO{} is only used for CU, and is thus co-located in the regions with a mix of the lowest electricity costs (for running DAC to produce \COtwo{}, the first CU input material), and the lowest hydrogen costs for producing \Htwo{} (the second CU input material).
+These regions happen to co-incide with where the best renewable resources are located.
+In the \modH{}, DAC is used for both CU and CS and is located in the regions with the lowest costs for electricity to run the DAC, with the additional requirement of DAC for CS having to be placed in regions with offshore sequestration access.
+
+In summary, both network models provide largely similar distributions of electrolysis and CC sites, but very different siting of CU, CS and DAC due to their distinct regional price differentiations for \COtwo{} and \Htwo{}.
+Both systems unlock captured carbon potentials in Central Europe.
+The system cost advantage of the \modH{} stems from being able to supply the exogenous hydrogen demands in Central Europe with low-cost hydrogen imports, while the \modCO{} needs to deploy electrolysis powered by poor renewable resources as well as additional gas infrastructure in Central Europe.
+
+\subsection*{Hybrid configuration provides further flexibilities}\label{subsec:Hybrid}
 
@@
         \centering
+        \includegraphics[width=\linewidth]{full/figures/90_nodes/balance_map_carbon.png}
+        \label{fig:capacity_map_carbon_co2}
+    \end{subfigure}%
+    \begin{subfigure}{.49\textwidth}
+        \centering
         \includegraphics[width=\linewidth]{full/figures/90_nodes/balance_map_hydrogen.png}
-        \caption{\hydrogen{} Sector.}
         \label{fig:capacity_map_hydrogen_co2}
     \end{subfigure}
-    \hfill
-    \begin{subfigure}{.49\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{full/figures/90_nodes/balance_map_carbon.png}
-        \caption{\carbon{} Sector.}
-        \label{fig:capacity_map_carbon_co2}
-    \end{subfigure}%
-    \caption{Optimal production and transport capacities of the carbon and hydrogen sector in a net-zero energy system in Europe with both \carbon{} and \hydrogen{} network expansion (Hybrid).
+    \caption{Optimal production and transport capacities of the carbon and hydrogen sector in a net-zero energy system in Europe with both \COtwo{} and \Htwo{} network expansion (Hybrid).
     % Carbon network looks the same as in~\cite{morbeeOptimisedDeploymentEuropean2012}: two backbones, one in the nothern Europe other in south east.
@@
 
-The Hybrid model combines the advantages of hydrogen and carbon networks.
-While the topology of the \hydrogen{}-grid roughly corresponds to that of the \hydrogengrid{} model, the \carbon{}-grid plays more of a complementary role.
-Fig.~\ref{fig:capacity_maps} shows the average production, consumption and transport of the two grids in the hybrid model.
+The \modHybrid{} combines the advantages of hydrogen and carbon networks.
+While the topology of the \Htwo{} network roughly corresponds to that of the \modH{}, the \COtwo{} network plays more of a complementary role.
+Fig.~\ref{fig:capacity_maps} shows the average production, consumption and transport of the two grids in the \modHybrid{}.
 Note that despite similar overall capacities, the investment in the hydrogen infrastructure is four times the investment in the carbon infrastructure.
 
-Similar to the \hydrogengrid{} model, the Hybrid model takes advantage of a large hydrogen network that enables hydrogen to be transported from centralized production sites in western regions such as Spain to decentralized FT production sites across the continent.
-However, in the Hybrid model hydrogen in the UK is used locally to produce synfuels, and less hydrogen is transported to Central Europe.
+Similar to the \modH{}, the \modHybrid{} takes advantage of a large hydrogen network that enables hydrogen to be transported from centralized production sites in western regions such as Spain to decentralized FT production sites across the continent.
+However, in the \modHybrid{} hydrogen in the UK is used locally to produce synfuels, and less hydrogen is transported to Central Europe.
 The carbon network supplements this with smaller clusters of networks.
-In addition to a larger carbon grid cluster in Central Europe, which transports carbon to the North Sea, the system is building three large linear carbon routes.
+In addition to a larger \COtwo{} network cluster in Central Europe, which transports carbon to the North Sea, the system is building three large linear carbon routes.
 These transport carbon from Romania and Bulgaria to Greece, from northern Italy to central Italy and from Hungary, the Czech Republic and Poland to the Baltic Sea.
@@
 This reduces the need for DAC in the system and the sequestration sites are shifted to the end points of the carbon networks.
-Second-order effects can be seen in the shift in FT production compared to the \hydrogengrid{} model.
+Second-order effects can be seen in the shift in FT production compared to the \modH{}.
 Regions that have both good renewable resources and sequestration potential could now forego sequestration and use local process emissions for FT production, as can be observed in Ireland.
@@
 This finding is tightly linked to the fact of competing functionalities to offtake carbon emissions.
-A more detailed of differences between the \hydrogengrid{} and the Hybrid model can be found in Fig.~.
+A more detailed of differences between the \modH{} and the \modHybrid{} can be found in Fig.~.
 
-Despite only small decrease in cost in comparison to the \hydrogengrid{} model, hybrid approach brings several advantages.
+Despite only small decrease in cost in comparison to the \modH{}, hybrid approach brings several advantages.
 These include a broader range of technologies contributing to its robustness, reduced reliance on DAC, which may be costlier than anticipated in the cost projections, and less land use for wind and solar due to the decreased necessity for DAC.
-The increase used of biomass with carbon capture, transport and sequestration leads to slight cost advantages in comparison to the \hydrogengrid{} model.
+The increase used of biomass with carbon capture, transport and sequestration leads to slight cost advantages in comparison to the \modH{}.
 
-\subsection*{Net-negativity amplifies DAC and Bioenergy with CU and CS}
+\subsection*{Net-negativity amplifies DAC and Bioenergy with CU and CS}\label{subsec:NN}
 
@@
     \includegraphics[width=0.9\linewidth]{difference/comparison/emission-reduction-0.1/figures/90_nodes/cost_bar.png}
-    \caption[short]{Net change in investments when tightening the \carbon{} emission target from net-zero to net-negative 10\% of 1990s emissions.}
+    \caption[short]{Net change in investments when tightening the \COtwo{} emission target from net-zero to net-negative 10\% of 1990s emissions.}
     \label{fig:net-negative_cost_bar}
@@
 In transitioning from a net zero to a net emissions reduction target, a continuation of investment trends are perceived.
-Fig.~\ref{fig:net-negative_cost_bar} shows the net investment changes per technology group for the \hydrogengrid{} and Hybrid.
-The scenario imposes a net carbon removal of 460 Mt/a, with both models incurring a similar net cost increase - 95 billion €/a for \hydrogengrid{} and 96 billion €/a for Hybrid, balancing the earlier cost advantage of the Hybrid model.
+Fig.~\ref{fig:net-negative_cost_bar} shows the net investment changes per technology group for the \Hgrid{} and \modHybrid{}s.
+The scenario imposes a net carbon removal of 460 Mt/a, with both models incurring a similar net cost increase - 95 billion €/a for the \modH{} and 96 billion €/a for the \modHybrid{}, balancing the earlier cost advantage of the \modHybrid{}.
 Both models invest relatively evenly in additional DAC, wind and heat pump systems, with DAC accounting for more than a third of the additional system costs.
@@
 Reduced usage of gas in general lower the investments in gas infrastructure in both models.
-Fundamental difference in investment strategies occur in the carbon capture sector: In contrast to the \hydrogengrid{} model, the Hybrid model continues the strategy of sequestering carbon captured and transported from bio-energetic inputs, expanding the carbon transport routes from Central and Eastern Europe to the near shores.
-At the same time, carbonaceous fuel production in Central Europe partially moves to Spain and UK, where it uses carbon from new DAC facilities without the installation of a \carbon{} network.
+Fundamental difference in investment strategies occur in the carbon capture sector: In contrast to the \modH{}, the \modHybrid{} continues the strategy of sequestering carbon captured and transported from bio-energetic inputs, expanding the carbon transport routes from Central and Eastern Europe to the near shores.
+At the same time, carbonaceous fuel production in Central Europe partially moves to Spain and UK, where it uses carbon from new DAC facilities without the installation of a \COtwo{} network.
 At the same time, solar based electrolysis in France, Nothern Italy and Switzerland expands requiring more solar capacities but less hydrogen imports from the Iberian Peninsula.
-The \hydrogengrid{} model on the other hand, expands decentralized CU systems from the Net-Zero scenario, which are now increasingly supplied with carbon from biogas-to-gas facilities, showing up in both "Bioenergy" and "Carbon Capture at Point Sources" cost-contributions.
-The centralized DAC facilities at sequestration sites are more expanded than in the Hybrid model and are requiring additional heat from biomass CHPs.
+The \modH{} on the other hand, expands decentralized CU systems from the Net-Zero scenario, which are now increasingly supplied with carbon from biogas-to-gas facilities, showing up in both ``Bioenergy'' and ``Carbon Capture at Point Sources'' cost-contributions.
+The centralized DAC facilities at sequestration sites are more expanded than in the \modHybrid{} and are requiring additional heat from biomass CHPs.
 
-These findings highlight the continuation of trends of the two systemic approaches, which are interchangable at marginal cost-difference: While a stand-alone \hydrogen{} grid combines well with highly centralized DAC and decentralized CU with bioenergy as primary carbon source, the complementation with a \carbon{} grid promotes combining DAC and CU where good renewable resources are and collecting, transporting and sequestering carbon from decentral bioenergetic processes.
+These findings highlight the continuation of trends of the two systemic approaches, which are interchangable at marginal cost-difference: While a stand-alone \Htwo{} network combines well with highly centralized DAC and decentralized CU with bioenergy as primary carbon source, the complementation with a \COtwo{} network promotes combining DAC and CU where good renewable resources are and collecting, transporting and sequestering carbon from decentral bioenergetic processes.
 
@@
 %     \includegraphics[width=\linewidth]{difference/net-negative-0.1-h2-only-full/figures/90_nodes/cost_map.png}
-%     \caption[short]{Net difference in investments between the \hydrogengrid{} model and the Hybrid model for the Net-Negative scenario.}
+%     \caption[short]{Net difference in investments between the \modH{} and the \modHybrid{} for the Net-Negative scenario.}
 %     \label{fig:net-negative_cost_map}
@@
 These learning rates are derived from historical data, which may not necessarily be indicative of future trends.
-In particular, cost assumptions on DAC, electrolysis and \hydrogen{} pipeline costs, could have a strong impact in the model results.
+In particular, cost assumptions on DAC, electrolysis and \Htwo{} pipeline costs, could have a strong impact in the model results.
 % The modelling results are heavily driven by the demand and emissions from industrial clusters.
-Allowing the model to relocate these and/or incorporate flexibility measures, may lead to less dependency on both \carbon{} and \hydrogen{} networks.
+Allowing the model to relocate these and/or incorporate flexibility measures, may lead to less dependency on both \COtwo{} and \Htwo{} networks.
 % The technological flexibilities in the carbon sector might not be exploited to full extent.
-Therefore, the \carbon{} transport via truck or ship is not considered.
+Therefore, the \COtwo{} transport via truck or ship is not considered.
 Fischer-Tropsch facilities require to run on baseload with at least 90\% of their nominal capacity.
@@
     \includegraphics*[width=0.8\linewidth]{comparison/emission-reduction/figures/90_nodes/objective_heatmap.png}
-    \caption{Total annual system cost for the different scenarios and \carbon{} targets.
-While the `Baseline' scenario has neither a \carbon{} nor a \hydrogen{} network, `Hybrid' is allowed to expand both.}
+    \caption{Total annual system cost for the different network models and \COtwo{} targets.
+While the \modBase{} has neither a \COtwo{} nor a \Htwo{} network, the \modHybrid{} is allowed to expand both.}
     \label{fig:objective_heatmap}
@@
     \includegraphics[width=\linewidth]{difference/h2-only-full/figures/90_nodes/cost_map.png}
-    \caption{Difference in cost investments between \hydrogengrid{} and Hybrid model.
-The left subfigure shows higher spendings per technology and region and transport system for the \hydrogengrid{} model, the right shows higher spendings in the Hydrid model.}
+    \caption{Difference in cost investments between the \Hgrid{} and \modHybrid{}s.
+The left subfigure shows higher spendings per technology and region and transport system for the \modH{}, the right shows higher spendings in the \modHybrid{}.}
     \label{fig:cost_map_difference}
@@
 
+% [Dynamics \modCO{} vs \modH{} (old text blocks)]
+% There are four notable dynamics or differences between the two grid strategies:
+% First, the production of FT fuels always moves to where low-cost carbon and low-cost hydrogen can be brought together.
+In the \modCO{ }, carbon is moved from Western Europe to the Iberian Peninsula, the British Isles and to Denmark to produce FT fuels centrally in few regions.
+In the \modH{}, hydrogen is transported in the opposite direction to produce FT fuels decentrally across all regions in Central Europe with locally captured \COtwo{} from point sources.
+% Second, the \modCO{ } captures more low-cost carbon from biomass, especially in Central and Eastern Europe, and less high-cost carbon from DAC in Greece and Southern Italy.
+Instead, the \COtwo{} network transports carbon from Central and Eastern Europe to Greece and Italy, where it is sequestered, or feeds into CU together with locally produced, low-cost electrolytic hydrogen.
+The \modH{} directly operates electrolyzers and DAC in Greece and Southern Italy, with electricity from low-cost renewables, and thus avoids significant network buildouts in this region.
+% Third, sequestration in the \modH{} occurs more decentralized, spread out across several regions with access to offshore sequestration sites, making use of point-source CC potentials in each of these regions.
+In the \modCO{ }, sequestration is clustered at fewer sequestration sinks, each collecting low-cost \COtwo{} from neighboring regions.
+However, there is no centralization at only one or two sites, since all the sequestration sites offer the same costs and the system minimizes system-wide carbon transport distances to about eight to ten sequestration sites.
+% And fourth, how the three types of CU are spread across the continent is driven by the costs of carbon and hydrogen (the latter mostly driven by the costs of renewable electricity) as well as by renewables capacity factors: Clearly, all carbonaceous fuel production requires sites where both carbon and hydrogen are available, and they moreover prefer sites with low carbon and hydrogen prices.
+Based on our model assumptions, FT fuel synthesis has very limited flexibility and needs to be run alost at baseload (ramp-down only possible to 90\% of maximum load).
+Thus, for FT fuel synthesis sites in Northern Europe with more electricity from wind power are preferred to sites in Southern Europe, powered mostly by solar.
+Methanathion and Methanolisation both have higher flexibility, being able to ramp down to 50\% load during the night, for example.
+Thus, these fuel production sites prefer the best renewable resource sites in Southern Europe.
+And finally, as methanation produces synthetic methane that needs to be fed into the gas grid, methanation prefers those sites with good interconnection to the existing gas network.
+Since Italy possesses a large gas network with easy interconnection to Central Europe, methanation prefers sites in Southern Italy (and Greece, which would be connected to Southern Italy) over the Iberian Peninsula, where Methanolisation is more prevalent.
+
+% In summary, the implementation of a \COtwo{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce low-cost hydrogen to make CU products (FT fuel, methanol, synthetic methane).
+DAC and electrolysis are deployed in the regions with the best renewable resources to meet carbonaceous fuel demands.
+% In the absence of a \COtwo{} network and with an \Htwo{} network on the other hand, regions with access to offshore sequestration sites capture all carbon from low-cost point sources and then sequester it.
+The system deploys additional DAC facilities at the locations with sequestration access and low-cost renewables to capture the remaining necessary carbon for sequestration.
+Carbonaceous fuel demand is met at the locations which offer the best mix of low-cost carbon from point sources and low-cost hydrogen from renewables, with some additional electrolysis and DAC required to fully meet the demand.
+% Sequestered carbon is nearly always transported before
+
+% Therefore, one can conclude that a \COtwo{} network favors a decentralized CC and a CU centralized at regions with abundant renewable resources.
+A \Htwo{} network favors a centralized electrolysis system at regions with abundant renewable resources and a decentralized CU system.
+
+% Together with understanding the dynamics described above, we can highlight three cost advantages of the \Htwo{} network in contrast to the \COtwo{} network: Capturing carbon at the lowest-cost regions reduces 4 billion euros in spending on DAC.
+However, the \modH{} produces electrolytic hydrogen in the lowest-cost regions and in total spends 7 billion euros less on renewable energy (14 billion euros less on solar, and 7 billion euros more on wind) as well as 1 billion euros less on SMR.
+Furthermore, without a \Htwo{} network, additional \Htwo{} storage is needed in Spain at a cost of 5 billion euros to store hydrogen produced during the summer for utilization in FT fuel production during the winter.
+% And finally, the \modCO{ } spends 7 billion euros more on gas boilers, gas plants, gas infrastructure, and methanation.
+% TODO: why??
+% The costs of the respective networks, on the other hand, are almost equal: the \modCO{ } spends 8 billion euros on the \COtwo{} network, while the \modH{} even spends 10 billion euros on the \Htwo{} network.
+These factors add up to account for 14 billion euros of the difference of 20 billion euros in system costs between the two scenarios (785 billion euros for the \COgrid{} and 765 billion euros for the \modH{}, see Fig.~\ref{fig:objective_heatmap}).
+
 \end{document}

### 2023-12-08 report/submissions/journal/paper/paper.tex
paper appendix: add cost comparison for half-price co2 net

@@
     \caption[short]{Total annual system cost subdivided into groups of technologies for the different models of the European energy system with a net-zero emission target.
-While the \modBase{} has neither a \COtwo{} nor a \Htwo{} network, the \modHybrid{} is allowed to expand both.
+While the \modBase{} has neither a \COtwo{} nor an \Htwo{} network, the \modHybrid{} is allowed to expand both.
 ``Gas Infrastructure'' combines gas facilities for power and heat production, ``\COtwo{} Infrastructure'' combines transport, storage and sequestration, and ``\Htwo{} Infrastructure'' combines transport and storage.
@@
 % In summary, the \COtwo{} network enables the transport of carbon from the lowest-cost CC sources to CU and CS carbon sinks, reducing the need for high-cost DAC.
-% The creation of a \Htwo{} network enables the cost-effective delivery of hydrogen from regions with a surplus of low-cost renewable energy to high-demand industrial centers with exogenous hydrogen demand and a surplus of low-cost CC potentials, reducing the need for DAC and for high-cost electrolysis at the industrial centers.
-% This highlights how both the implementation of a \COtwo{} or a \Htwo{} network unlocks low-cost carbon capture potentials at geographically fixed point sources across the continent.
+% The creation of an \Htwo{} network enables the cost-effective delivery of hydrogen from regions with a surplus of low-cost renewable energy to high-demand industrial centers with exogenous hydrogen demand and a surplus of low-cost CC potentials, reducing the need for DAC and for high-cost electrolysis at the industrial centers.
+% This highlights how both the implementation of a \COtwo{} or an \Htwo{} network unlocks low-cost carbon capture potentials at geographically fixed point sources across the continent.
 
@@
 
-As we have also discussed above, the \modCO{} captures large amounts of low-cost carbon from bioenergy and industrial activities that were untapped in the \modBase{}, especially in Central and Eastern Europe, and the \COtwo{} network smoothes out price differences for captured carbon across regions (see Fig.~\ref{fig:balance_map}, top left).
-Hydrogen prices still vary strongly, with low price ``valleys'' in the regions with good renewable resources such as the Iberian Peninsula or Denmark, and high price ``peaks'' in Central Europe (see bottom left).
-The model optimally locates CU in the regions where it can minimize costs for its material inputs, carbon and hydrogen.
-Thus, when carbon prices are smoothed out in the \modCO{}, CU moves to the hydrogen price valleys in the periphery of the continent.
-The captured carbon is transported from Central Europe through several large pipeline corridors to these locations to produce carbonaceous fuels.
-Similarly to CU, the model optimally locates CS in regions where it can minimize costs for its only material input, carbon.
-Thus, CS moves to the carbon price valleys (which are not strongly pronounced due to the \COtwo{} network, but still prevalent) among the subset of regions with access to offshore sequestration sites.
-In the \modCO, these are sites  sites in the North Sea (Northern Germany, Norway, Southeastern England), Baltic Sea (Latvia, Eastern Denmark) and Mediterranean Sea (Italy, Greece).
-CU and CS compete for carbon sources and thus there is some interaction for their optimal placement.
-Finally, in the absence of an \Htwo{} network, the system places some electrolysis and SMR in Central Europe to supply exogenous hydrogen demand for industrial processes and fuel cell land transport (see bottom left).
-Due to the additional electricity demand for electrolysis, the model switches some winter heating from heat pumps to gas\=/fueled boilers and CHPs, and thus emphasizes overall more gas imports into Central Europe.
+As we have also discussed above, the \modCO{} captures large amounts of low-cost carbon from bioenergy and industrial activities that were untapped in the \modBase{}, especially in Central and Eastern Europe, and almost none of which is used locally (see Fig.~\ref{fig:balance_map}, top left).
+A large part of the captured carbon is transported from Western Europe to the Iberian Peninsula, the British Isles and Denmark, where it is used along with low-cost electrolytic hydrogen to produce Fischer-Tropsch fuels and methanol.
+A second part of captured carbon is transported to methanation plants in Italy and Greece, where the synthetic methane output is fed into the well-established gas grid since Italy serves as a large gas import hub for Europe.
+Finally, a third part of captured carbon is transported directly to sites in the North Sea, Baltic Sea and Mediterranean Sea where it is sequestered.
+In absence of an \Htwo{} network, the system places most electrolysis in regions with the best renewable resources to supply local CU, and smaller amounts of electrolyzers where hydrogen is needed for industrial processes across Central Europe.
+The model also emphasizes gas imports into Central Europe.
+Hydrogen prices still vary strongly across the continent, with low price ``valleys'' in regions with good renewable resources such as the Iberian Peninsula or Denmark, and high price ``peaks'' in Central Europe (see bottom left).
 
-The \Htwo{} network on the other hand, smoothes out hydrogen prices across the entire continent (see bottom right).
-Thus, the model chooses to locate CU in the carbon price valleys across Western, Central and Eastern Europe (see top right).
-Low-cost hydrogen from regions with abundant renewable resources is transported to these regions with low-price carbon from abundant sources of fossil, biogenic and process emissions for point-source CC.
-The \Htwo{} network is also used to supply the exogenous \Htwo{} demand in these regions for industry and fuel cell land transport at much lower costs than local electrolysis or SMR.
-The model locates CS again in the regions that represent carbon price valleys among all those with access to offshore sequestration sites.
-In the case of the \modH{}, due to the lack of a \COtwo{} network all CC potentials from point sources are exhausted in all offshore sequestration regions, and the carbon is then sequestered.
-The additionally needed carbon to reach 200~Mt of CS is captured via DAC in the regions offshore sequestration regions with the lowest costs for DAC, with the largest share deployed in Portugal (see top right).
+To understand the dynamics of the \modCO{}, we observe that the system locates CU in the regions where it can minimize costs for its material inputs, carbon and hydrogen.
+As the \COtwo{} network largely smoothes out carbon prices across the continent, CU is located in the hydrogen price valleys on the periphery of the continent.
+Similarly, the model locates CS in regions where it can minimize costs for its only material input, carbon.
+Thus, CS is located in the carbon price valleys among the subset of regions with access to offshore sequestration sites.
+Due to the \COtwo{} network these valleys are not strongly pronounced, but still prevalent and impact CS locations.
+DAC is only used for CU and thus co-located in regions with a mix of low electricity costs (for running DAC) and low hydrogen costs (where CU is located, see above).
 
-The results illustrate the commonalities and differences between the two grids: both a \COtwo{} network and an \Htwo{} network prefer distributed CC at point sources and electrolysis in regions with abundant renewable resources, choosing the lowest-cost production sites for each while trading off transport costs.
-Nonetheless, the \COtwo{} network is forced to deploy some electrolysis in regions with poor renewables to supply their exogenous hydrogen demands for industry and fuel cell land transport.
+The \Htwo{} network on the other hand, transports low-cost hydrogen from regions with abundant renewable resources to regions with low-price carbon from abundant sources of fossil, biogenic and process emissions (see bottom right).
+The main transportation routes go from Spain and the United Kingdom to Central Europe to produce synfuels with locally captured carbon from point sources.
+Also, the \Htwo{} network supplies exogenous \Htwo{} demand for industry in Central Europe at much lower costs than local electrolysis or SMR.
 
-The differences between the CO2-Grid and H2-Grid in the context of CU, CS and DAC are notable and highlight the unique operational characteristics of each grid type.
-CU always chooses the sites with the lowest costs for its hydrogen and carbon inputs.
-With the \COtwo{} network, \COtwo{} costs are levelized across the continent, and CU is thus located in the regions with the lowest \Htwo{} costs.
-With the \Htwo{} network, \Htwo{} costs are levelized across the continent, and CU is thus located in the regions with the lowest \COtwo{} costs.
-Since regions with low-cost \COtwo{} in the \modH{} are very spread out across the continent, the CU locations are also more distributed, while in the \modCO{} they cluster in a few pockets of low-cost \Htwo{}.
-CS co-locates in both models in the regions with the lowest \COtwo{} costs among those with offshore sequestration potential.
-However, due to the lack of \COtwo{} transport in the \modH{}, additional CS is located in the offshore locations where DAC, and therefore electricity, can be produced at the lowest costs.
-Finally, DAC in the \modCO{} is only used for CU, and is thus co-located in the regions with a mix of the lowest electricity costs (for running DAC to produce \COtwo{}, the first CU input material), and the lowest hydrogen costs for producing \Htwo{} (the second CU input material).
-These regions happen to co-incide with where the best renewable resources are located.
-In the \modH{}, DAC is used for both CU and CS and is located in the regions with the lowest costs for electricity to run the DAC, with the additional requirement of DAC for CS having to be placed in regions with offshore sequestration access.
+To explain the dynamics of the \modH{}, we observe that the model locates CU in the carbon price valleys across Western, Central and Eastern Europe (see top right), as the \Htwo{} network smoothes out hydrogen prices across the entire continent (see bottom right).
+CS is again located in the carbon price valleys among those regions with access to offshore sequestration sites.
+Due to the lack of a \COtwo{} network the full CC potentials from point sources are exhausted in all offshore sequestration regions, and the carbon is then sequestered.
+The additionally needed carbon to reach 200~Mt of CS is captured via DAC in the offshore sequestration regions with the lowest costs for DAC, with the largest share deployed in Portugal (see top right).
+DAC is used for both CU and CS in this model, and for both purposes DAC is located in the regions with the lowest costs for electricity.
 
-In summary, both network models provide largely similar distributions of electrolysis and CC sites, but very different siting of CU, CS and DAC due to their distinct regional price differentiations for \COtwo{} and \Htwo{}.
-Both systems unlock captured carbon potentials in Central Europe.
-The system cost advantage of the \modH{} stems from being able to supply the exogenous hydrogen demands in Central Europe with low-cost hydrogen imports, while the \modCO{} needs to deploy electrolysis powered by poor renewable resources as well as additional gas infrastructure in Central Europe.
+In summary, both network models prefer distributed CC at point sources and electrolysis in regions with abundant renewable resources, choosing the lowest-cost production sites for each while trading off transport costs.
+However, differences between the two models are notable for the placement of CU, CS and DAC due to their distinct \COtwo{} and \Htwo{} price patterns.
+CU locations are more distributed in the \modH{}, while they cluster in a few pockets of low-cost \Htwo{} in the \modCO{}.
+CS co-locates in the regions with the lowest \COtwo{} costs (among those with offshore sequestration potential), but these regions are very different between the models.
+DAC is placed in regions with the best renewable resources in both models, but it is used for only CU in the \modCO{} and for both CU and CS in the \modH{}.
+Finally, the system cost advantage of the \modH{} stems from being able to supply the exogenous hydrogen demands in Central Europe with low-cost hydrogen imports, while the \modCO{} is forced to deploy some electrolysis in regions with poor renewables resources as well as additional gas infrastructure in Central Europe.
 
@@
 While the topology of the \Htwo{} network roughly corresponds to that of the \modH{}, the \COtwo{} network plays more of a complementary role.
-Fig.~\ref{fig:capacity_maps} shows the average production, consumption and transport of the two grids in the \modHybrid{}.
+Fig.~\ref{fig:capacity_maps} shows the average production, consumption and transport of the two networks in the \modHybrid{}.
 Note that despite similar overall capacities, the investment in the hydrogen infrastructure is four times the investment in the carbon infrastructure.
@@
     \caption{Total annual system cost for the different network models and \COtwo{} targets.
-While the \modBase{} has neither a \COtwo{} nor a \Htwo{} network, the \modHybrid{} is allowed to expand both.}
+While the \modBase{} has neither a \COtwo{} nor an \Htwo{} network, the \modHybrid{} is allowed to expand both.}
     \label{fig:objective_heatmap}
@@
 
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=0.9\linewidth]{difference/comparison/subsidy/figures/90_nodes/cost_bar.png}
+    \caption[short]{Net change in system cost when reducing assumed costs for \carbon{} pipelines by 50\%.}
+    \label{fig:net-negative_cost_bar}
+\end{figure}
+
 % [Dynamics \modCO{} vs \modH{} (old text blocks)]
-% There are four notable dynamics or differences between the two grid strategies:
+% There are four notable dynamics or differences between the two network strategies:
 % First, the production of FT fuels always moves to where low-cost carbon and low-cost hydrogen can be brought together.
@@
 % Therefore, one can conclude that a \COtwo{} network favors a decentralized CC and a CU centralized at regions with abundant renewable resources.
-A \Htwo{} network favors a centralized electrolysis system at regions with abundant renewable resources and a decentralized CU system.
+An \Htwo{} network favors a centralized electrolysis system at regions with abundant renewable resources and a decentralized CU system.
 
@@
 However, the \modH{} produces electrolytic hydrogen in the lowest-cost regions and in total spends 7 billion euros less on renewable energy (14 billion euros less on solar, and 7 billion euros more on wind) as well as 1 billion euros less on SMR.
-Furthermore, without a \Htwo{} network, additional \Htwo{} storage is needed in Spain at a cost of 5 billion euros to store hydrogen produced during the summer for utilization in FT fuel production during the winter.
+Furthermore, without an \Htwo{} network, additional \Htwo{} storage is needed in Spain at a cost of 5 billion euros to store hydrogen produced during the summer for utilization in FT fuel production during the winter.
 % And finally, the \modCO{ } spends 7 billion euros more on gas boilers, gas plants, gas infrastructure, and methanation.

### 2023-12-11 report/submissions/journal/paper/paper.tex
paper: add descriptive paragraphs back into 3.2, remove repetitions
small fixes due to removed component location function allow to ignore gas store capex
move methods section to end
paper: incorporate initial feedback from LZ

@@
 We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available sequestration potential.
-
-\section{Methodology}
-\label{sec:methodology}
-
-The study is conducted on the basis of the open-source, capacity-expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
-The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, waste, agriculture, and transport sectors, including international aviation and shipping.
-
-\begin{figure}
-    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/demand_bar.png}
-    \caption{Assumptions on exogenous demand, derived from~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023}.
-The figure shows the total annual energy demand for each energy source which determine the model's endogenous investments and operation.
-Endogenous processes can lead to higher total production volumes of some energy carriers, e.g., the demand for methanol requires more hydrogen and carbon as secondary (energy) inputs, which are not considered here.
-The model are defined per region and time stamp.}
-    % TODO: adjust labels to show that these are exogenous assumptions
-    \label{fig:total-demand-bar}
-\end{figure}
-%
-In our configuration, the model's time horizon spans one year with a temporal resolution of 3 hours and a spatial resolution of 90 regions.
-Each of the regions consists of a complex subsystem with technologies for supplying, converting, storing and transporting energy.
-Exogenous assumptions on energy demand and non-abatable emissions are taken from various sources~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023}.
-The energy demand for electricity, transport, biomass, heat and gas are defined per region and time-step.
-Land transport demand is exogenously divided between electric vehicles (85\%) and fuel cell vehicles (15\%), the latter representing demand for heavy duty land transport.
-Demands for kerosene for aviation, methanol for shipping, and naphtha for industry are aggregated in the system scope and kept constant throughout all time-steps.
-Heat demand is regionally subdivided into shares of urban, rural and industrial sites.
-We show the sum of all energy demands in Fig.~\ref{fig:total-demand-bar}.
-The system emits 633~Mt \COtwo{} per year from industry, aviation, shipping and agriculture, 153~Mt of which are fossil-based process emissions.
-Industrial energy demand and excess heat potentials are calculated per node on the basis of~\cite{hotmaps_industrial_db}.
-% use https://pypsa-eur.readthedocs.io/en/latest/licenses.html
-
-Endogenous model results include the expansion of renewable energy sources, storage technologies, and transmission capacities.
-The model considers various energy carriers like electricity, hydrogen, methan, methanol, liquid hydrocarbons and biomass, together with conversions technologies.
-Carbon-neutral electricity is provided by wind, solar, biomass, hydro and nuclear power plants.
-Hydro and nuclear plant capacities cannot be extended.
-Weather-dependent power potentials for solar, wind and hydro are calculated from reanalysis and satellite data sets~\cite{hersbachERA5GlobalReanalysis2020,pfeifrothSurfaceRadiationData2017}  per region and time-stamp, using the open-source tool Atlite~\cite{hofmannAtliteLightweightPython2021}.
-Solar and wind power can be expanded in alignment with eligible land-use restrictions calculated on the basis of~\cite{eeaCorineLandCover2012,eeaNatura2000Data2016}.
-We restrict the electric transmission system's expansion to 20\% of its current capacity, acknowledging the challenges in inaugurating new transmission projects.
-For the use of biomass we consider only residual biomass products and no energy crops.
-We limit regional biomass use to the medium-level potentials from the ENPRESO database~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
-Inter-regional biomass transport is permitted with transport costs considered.
-
-In our model, hydrogen can be produced from electrolysis and steam methane reforming (SMR).
-Geological distribution and potentials for \Htwo{} storage in salt caverns are derived from~\cite{caglayanTechnicalPotentialSalt2020}.
-Re-electrification of hydrogen is possible via fuel cells.
-If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limit, considering costs for pipeline segments and compressors.
-Pipeline flows are modelled using net transfer capacities and without flow dynamics, pressure valves, or energy demand for compression.
-% No retrofitting of gas pipelines is considered, potentially overestimating hydrogen network cost.
-
-The topology of the \COtwo{}, \Htwo{} and gas network is identical to the topology of the electricity network, connecting all neighboring regions.
-Transport of liquid fuels like oil, methanol and Fischer-Tropsch (FT) fuels is ``copper-plated'', since transport costs per unit of energy are negligible due to their high energy density.
-Throughout this study, we focus on \COtwo{} and \Htwo{} networks because of their high relevance for infrastructure investments and thus public policy decisions.
-The electricity network and gas network are both included in the model with full geographical detail, but not further analyzed.
-Electricity networks are already in place, and we restrict further extension due to concerns of public acceptance.
-Gas networks also already exist, and will most likely experience decreased use in the future, removing any bottlenecks or constraints on the optimal system buildout or operation.
-
-Our model features three drop-in fuel production technologies for CU: methanation, methanolization, and Fischer-Tropsch synthesis.
-The processed fuels are not limited in their total quantity of use or production.
-Methane is transported through the gas network, while methanol and FT fuel can be transferred inter-regionally without additional costs, since dense fuels have negligible transport costs per unit of energy.
-Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial process heat.
-Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production, aviation kerosene, or agricultural machinery oil.
-
-To supply carbon needed for CS and CU, the system can choose to deploy carbon capture technology at various point sources (see below), or through DAC facilities.
-The concept of a merit order for captured carbon plays a pivotal role in optimizing and prioritizing the deployment of carbon capture technologies based on their economic feasibility.
-This merit order ranks all CC technologies according to their relative costs to capture a ton of carbon (see Fig.~X).
-At the lower end of the cost spectrum, CC technologies applied to process emissions, such as those from cement, offer a cost-effective starting point.
-Following this, biomass combined heat and power (CHP) systems provide a dual benefit of energy production and carbon capture.
-Moving up the scale, gas used in industrial applications and biomass employed for industrial processes represent more costly yet viable options to capture carbon.
-DAC, an emerging technology capable of extracting \COtwo{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand.
-Finally, based on our model results, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per marginal ton of captured \COtwo{}.
-Biogas input is a high-cost fuel which based on the endogenous modeling decisions is not used in large quantities.
-To the extent that biogas-to-gas facilities are used by the model to supply additional (carbon-neutral) gas as fuel, adding CC infrastructure incurs low costs (and thus ranges on the left-hand side of the merit order curve).
-However, the price of capturing additional \COtwo{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the marginal cost of captured \COtwo{}.
-This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.
-If we consider the spatial aspect of distributed carbon capture potentials, the \COtwo{} network can exploit comparative cost advantages between CO2 bidding zones to utilize the lowest-cost CC facilities across the continent.
-We assume a capture rate of 90\% for CC on process emissions, SMR, biogas-to-gas, as well as gas and biomass used in industry, and 95\% for CC on biomass and gas CHPs.
-
-To store carbon, we differentiate between short-term storage in steel tanks and long-term, irreversible sequestration in underground sequestration sites such as porous rock formations or depleted gas reservoirs.
-Costs for both options are included in the model.
-For carbon sequestration, we only consider offshore sites as potential sinks.
-We make this choice because offshore sequestration sites typically have larger capacity compared to onshore sequestration sites in saline aquifers and due to concerns over public safety for infrastructure near populated areas.
-Our estimates for carbon sequestration potentials are conservative, limiting total sequestration to 25~Mt per site and calculating annual storage availability over 25 years.
-Furthermore, we decide to cap the total sequestered \COtwo{} to 200~Mt per year for our Net-Zero scenario.
-This is enough to offset the hardest-to-abate fossil process emissions and some limited slack of fossil fuels to be used where they can be most valuable.
-This is in order to avoid offsetting fossil emissions with carbon removal if they can technologically be avoided in the first place.
-All technology cost assumptions are taken for the year 2040 and sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
-
-Addressing climate targets, we define two \COtwo{} target scenarios:
-\begin{itemize}
-    \item[] \textit{Net\=/Zero}: aligning with the EU's 2050 emission targets. Also denoted as NZ.
-    \item[] \textit{Net\=/Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \COtwo{} annually. Also denoted as NN.
-\end{itemize}
-Unless otherwise mentioned, our reference is the Net-Zero scenario. For the Net-Negative scenario, the cap of total carbon sequestration is adjusted from 200~Mt to 660~Mt per year accordingly.
-
-Beyond \COtwo{} targets, we consider four expansion scenarios, referred to as models:
-\begin{itemize}
-    \item[] \textit{Baseline:} Neither \COtwo{} nor \Htwo{} networks are constructed.
-    \item[] \textit{\COgrid{}:} Only the \COtwo{} network is developed.
-    \item[] \textit{\Hgrid{}:} Only the \Htwo{} network is developed.
-    \item[] \textit{Hybrid:} Both \COtwo{} and \Htwo{} networks are developed.
-\end{itemize}
 
@@
     \centering
-    \begin{subfigure}{.49\textwidth}
+    \begin{subfigure}{.51\textwidth}
         \centering
@@
 Fig.~\ref{fig:capacity_maps} shows the average production, consumption and transport of the two networks in the \modHybrid{}.
-Note that despite similar overall capacities, the investment in the hydrogen infrastructure is four times the investment in the carbon infrastructure.
+Note that despite similar overall capacities in terms of weight moved, the investment in the hydrogen infrastructure is four times the investment in the carbon infrastructure.
 
@@
 
-Despite only small decrease in cost in comparison to the \modH{}, hybrid approach brings several advantages.
+Despite only a further reduction of \label{}0.4 percentage points (\label{}3 billion euros) in total system costs compared to the \modH{}, the \modHybrid{} brings several advantages.
 These include a broader range of technologies contributing to its robustness, reduced reliance on DAC, which may be costlier than anticipated in the cost projections, and less land use for wind and solar due to the decreased necessity for DAC.
@@
 % The technology costs used in the model rely on cost projections that incorporate reductions based on learning rates.
-These learning rates are derived from historical data, which may not necessarily be indicative of future trends.
+These learning rates are derived from historical data, which may not necessarily be indicative of future trends, and they are based on projected global installed capacities of one technology, which can result in large uncertainties.
 In particular, cost assumptions on DAC, electrolysis and \Htwo{} pipeline costs, could have a strong impact in the model results.
@@
 It emphasizes the necessity of comprehensive infrastructure planning, integrating diverse energy carriers and technologies to achieve these objectives effectively.
+
+\section{Methodology}
+\label{sec:methodology}
+
+The study is conducted on the basis of the open-source, capacity-expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
+The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, waste, agriculture, and transport sectors, including international aviation and shipping.
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/demand_bar.png}
+    \caption{Assumptions on exogenous demand, derived from~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023}.
+The figure shows the total annual energy demand for each energy source which determine the model's endogenous investments and operation.
+Endogenous processes can lead to higher total production volumes of some energy carriers, e.g., the demand for methanol requires more hydrogen and carbon as secondary (energy) inputs, which are not considered here.
+The model are defined per region and time stamp.}
+    % TODO: adjust labels to show that these are exogenous assumptions
+    \label{fig:total-demand-bar}
+\end{figure}
+%
+In our configuration, the model's time horizon spans one year with a temporal resolution of 3 hours and a spatial resolution of 90 regions.
+Each of the regions consists of a complex subsystem with technologies for supplying, converting, storing and transporting energy.
+Exogenous assumptions on energy demand and non-abatable emissions are taken from various sources~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023}.
+The energy demand for electricity, transport, biomass, heat and gas are defined per region and time-step.
+Land transport demand is exogenously divided between electric vehicles (85\%) and fuel cell vehicles (15\%), the latter representing demand for heavy duty land transport.
+Demands for kerosene for aviation, methanol for shipping, and naphtha for industry are aggregated in the system scope and kept constant throughout all time-steps.
+Heat demand is regionally subdivided into shares of urban, rural and industrial sites.
+We show the sum of all energy demands in Fig.~\ref{fig:total-demand-bar}.
+The system emits 633~Mt \COtwo{} per year from industry, aviation, shipping and agriculture, 153~Mt of which are fossil-based process emissions.
+Industrial energy demand and excess heat potentials are calculated per node on the basis of~\cite{hotmaps_industrial_db}.
+% use https://pypsa-eur.readthedocs.io/en/latest/licenses.html
+
+Endogenous model results include the expansion of renewable energy sources, storage technologies, transmission capacities, heating technologies, peaking power plants, and the deployment of gray, blue or green hydrogen, among others.
+The model considers various energy carriers like electricity, hydrogen, methan, methanol, liquid hydrocarbons and biomass, together with conversions technologies.
+Carbon-neutral electricity is provided by wind, solar, biomass, hydro and nuclear power plants.
+Hydro and nuclear plant capacities cannot be extended.
+Weather-dependent power potentials for solar, wind and hydro are calculated from reanalysis and satellite data sets~\cite{hersbachERA5GlobalReanalysis2020,pfeifrothSurfaceRadiationData2017}  per region and time-stamp, using the open-source tool Atlite~\cite{hofmannAtliteLightweightPython2021}.
+Solar and wind power can be expanded in alignment with eligible land-use restrictions calculated on the basis of~\cite{eeaCorineLandCover2012,eeaNatura2000Data2016}.
+We restrict the electric transmission system's expansion to 20\% of its current capacity, acknowledging the challenges in inaugurating new transmission projects.
+For the use of biomass we consider only residual biomass products and no energy crops.
+We limit regional biomass use to the medium-level potentials from the ENPRESO database~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
+Inter-regional biomass transport is permitted with transport costs considered.
+
+In our model, we consider green, blue and gray hydrogen from electrolysis and steam methane reforming (SMR), the latter of which may be equipped with CC technology.
+Geological distribution and potentials for \Htwo{} storage in salt caverns are derived from~\cite{caglayanTechnicalPotentialSalt2020}.
+Re-electrification of hydrogen is possible via fuel cells.
+If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limit, considering costs for pipeline segments and compressors.
+Pipeline flows are modelled using net transfer capacities and without flow dynamics, pressure valves, or energy demand for compression.
+% No retrofitting of gas pipelines is considered, potentially overestimating hydrogen network cost.
+
+The topology of the \COtwo{}, \Htwo{} and gas network is identical to the topology of the electricity network, connecting all neighboring regions.
+Transport of liquid fuels like oil, methanol and Fischer-Tropsch (FT) fuels is ``copper-plated'', since transport costs per unit of energy are negligible due to their high energy density.
+Throughout this study, we focus on \COtwo{} and \Htwo{} networks because of their high relevance for infrastructure investments and thus public policy decisions.
+The electricity network and gas network are both included in the model with full geographical detail, but not further analyzed.
+Electricity networks are already in place, and we restrict further extension due to concerns of public acceptance.
+Gas networks also already exist, and will most likely experience decreased use in the future, removing any bottlenecks or constraints on the optimal system buildout or operation.
+
+Our model features three drop-in fuel production technologies for CU: methanation, methanolization, and Fischer-Tropsch synthesis.
+The processed fuels are not limited in their total quantity of use or production.
+Methane is transported through the gas network, while methanol and FT fuel can be transferred inter-regionally without additional costs, since dense fuels have negligible transport costs per unit of energy.
+Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial process heat.
+Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production, aviation kerosene, or agricultural machinery oil.
+
+To supply carbon needed for CS and CU, the system can choose to deploy carbon capture technology at various point sources (see below), or through DAC facilities.
+The concept of a merit order for captured carbon plays a pivotal role in optimizing the deployment of carbon capture technologies based on their economic feasibility.
+This merit order ranks all CC technologies according to their relative costs to capture an additional marginal ton\footnote[2]{The cost value for a technology depends on the exact system configuration and may vary by scenario and even by region, since many technologies produce other (primary) outputs such as electricity and heat.
+The costs to capture an additional marginal ton of carbon may also be different from the average cost of capture.} of carbon (see Fig.~X).
+At the lower end of the cost spectrum, CC technologies applied to process emissions, such as those from cement, offer a cost-effective starting point.
+Following this, biomass combined heat and power (CHP) systems provide a dual benefit of energy production and carbon capture.
+Moving up the scale, gas used in industrial applications and biomass employed for industrial processes represent more costly yet viable options to capture carbon.
+DAC, an emerging technology capable of extracting \COtwo{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand.
+Finally, based on our model results, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per marginal ton of captured \COtwo{}.
+Biogas input is a high-cost fuel which based on the endogenous modeling decisions is not used in large quantities.
+To the extent that biogas-to-gas facilities are used by the model to supply additional (carbon-neutral) gas as fuel, adding CC infrastructure incurs low costs (and thus ranges on the left-hand side of the merit order curve).
+However, the price of capturing additional \COtwo{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the marginal cost of captured \COtwo{}.
+This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.
+If we consider the spatial aspect of distributed carbon capture potentials, the \COtwo{} network can exploit comparative cost advantages between CO2 bidding zones to utilize the lowest-cost CC facilities across the continent.
+We assume a capture rate of 90\% for CC on process emissions, SMR, biogas-to-gas, as well as gas and biomass used in industry, and 95\% for CC on biomass and gas CHPs.
+
+To store carbon, we differentiate between short-term storage in steel tanks and long-term, irreversible sequestration in underground sequestration sites such as porous rock formations or depleted gas reservoirs.
+Costs for both options are included in the model.
+For carbon sequestration, we only consider offshore sites as potential sinks.
+We make this choice because offshore sequestration sites typically have larger capacity compared to onshore sequestration sites in saline aquifers and due to concerns over public safety for infrastructure near populated areas.
+Our estimates for carbon sequestration potentials are conservative, limiting total sequestration to 25~Mt per site and calculating annual storage availability over 25 years.
+Furthermore, we decide to cap the total sequestered \COtwo{} to 200~Mt per year for our Net-Zero scenario.
+This is enough to offset the hardest-to-abate fossil process emissions and some limited slack of fossil fuels to be used where they can be most valuable.
+This is in order to avoid offsetting fossil emissions with carbon removal if they can technologically be avoided in the first place.
+All technology cost assumptions are taken for the year 2040 and sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
+
+Addressing climate targets, we define two \COtwo{} target scenarios:
+\begin{itemize}
+    \item[] \textit{Net\=/Zero}: aligning with the EU's 2050 emission targets. Also denoted as NZ.
+    \item[] \textit{Net\=/Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \COtwo{} annually. Also denoted as NN.
+\end{itemize}
+Unless otherwise mentioned, our reference is the Net-Zero scenario. For the Net-Negative scenario, the cap of total carbon sequestration is adjusted from 200~Mt to 660~Mt per year accordingly.
+
+Beyond \COtwo{} targets, we consider four expansion scenarios, referred to as models:
+\begin{itemize}
+    \item[] \textit{Baseline:} Neither \COtwo{} nor \Htwo{} networks are constructed.
+    \item[] \textit{\COgrid{}:} Only the \COtwo{} network is developed.
+    \item[] \textit{\Hgrid{}:} Only the \Htwo{} network is developed.
+    \item[] \textit{Hybrid:} Both \COtwo{} and \Htwo{} networks are developed.
+\end{itemize}
 
@@
     \includegraphics[width=0.9\linewidth]{difference/comparison/subsidy/figures/90_nodes/cost_bar.png}
-    \caption[short]{Net change in system cost when reducing assumed costs for \carbon{} pipelines by 50\%.}
-    \label{fig:net-negative_cost_bar}
+    \caption[short]{Net change in system cost when reducing assumed costs for \COtwo{} pipelines by 50\%.}
+    \label{fig:half-CO2-price_cost_bar}
 \end{figure}

### 2023-12-11 report/submissions/journal/paper/paper.tex
further fine-tune plots; review all figure captions

@@
 \newcommand{\modHybrid}{Hybrid model}
+
+% alternative
+\newcommand{\carbon}{CO$_2$}
+\newcommand{\hydrogen}{H$_2$}
+\newcommand{\carbongrid}{CO$_2$\=/Grid}
+\newcommand{\hydrogengrid}{H$_2$\=/Grid}
+\newcommand{\baelinemodel}{Baseline model}
+\newcommand{\carbonmodel}{CO$_2$\=/Grid model}
+\newcommand{\hydrogenmodel}{H$_2$\=/Grid model}
+\newcommand{\hybridmodel}{Hybrid model}
 
@@
     \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/captureshare_line.png}
-    \caption{Share of facilities with integrated carbon capture for all models .}
+    \caption{Proportion of plants with integrated carbon capture for all models in the net zero scenario.
+The size of the dots corresponds to the average capacity factor of the respective technology.
+While peak load technologies such as gas-fired combined heat and power (CHP) plants, which are only in operation on a few days a year, are not expanded for carbon capture, baseload point sources such as biomass CHP and process emissions from industry are fully (partially) developed as soon as the transport of \COtwo{} (\Htwo{}) is permitted.}
     % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
@@
     \includegraphics[width=\linewidth]{comparison/single-technologies/figures/90_nodes/balance_map_dedicated.png}
-    \caption{Optimal operation per sector for a net-zero energy system in Europe with average production on the left and average consumption on the right for both, (a) the \COtwo{} sector in the \modCO{} and (b) the \Htwo{} sector in the \modH{}.}
+    \caption{Optimal operation, flows and prices of the carbon (top line) and hydrogen (bottom line) sectors for the \carbonmodel{} (left) and the \hydrogenmodel{} (right) in the net zero scenario.
+For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
+Lines and arrows show the interregional transportation.
+Both models transport \carbon{} and \hydrogen{} respectively from areas with low prices to areas with high prices to supply CU and, in the case of \carbonmodel{}, CS.
+    }
     \label{fig:balance_map}
@@
     \centering
-    \begin{subfigure}{.51\textwidth}
+    \begin{subfigure}{.5\textwidth}
         \centering
@@
     \end{subfigure}%
-    \begin{subfigure}{.49\textwidth}
+    \begin{subfigure}{.5\textwidth}
         \centering
@@
     \end{subfigure}
-    \caption{Optimal production and transport capacities of the carbon and hydrogen sector in a net-zero energy system in Europe with both \COtwo{} and \Htwo{} network expansion (Hybrid).
+    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hybridmodel{} in the net zero scenario.
+For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
+Lines and arrows show the interregional transportation.
+\carbon{} from point-source in the inland either supplies local CU with imported \hydrogen{} or facilitates sequestration in nearby offshore regions.
     % Carbon network looks the same as in~\cite{morbeeOptimisedDeploymentEuropean2012}: two backbones, one in the nothern Europe other in south east.
@@
 This finding is tightly linked to the fact of competing functionalities to offtake carbon emissions.
-A more detailed of differences between the \modH{} and the \modHybrid{} can be found in Fig.~.
+A more detailed of differences between the \modH{} and the \modHybrid{} can be found in Fig.~\ref{fig:cost_map_difference}.
 
@@
     \includegraphics[width=0.9\linewidth]{difference/comparison/emission-reduction-0.1/figures/90_nodes/cost_bar.png}
-    \caption[short]{Net change in investments when tightening the \COtwo{} emission target from net-zero to net-negative 10\% of 1990s emissions.}
+    \caption[short]{Net change in investments when tightening the \COtwo{} emission target from net-zero to net-negative 10\% of 1990s emissions.
+For all models, Direct Air Capture (DAC) contributes most to the additional \carbon{} removal, requiring further solar, wind and heat pump capacities for electricity and heat input.}
     \label{fig:net-negative_cost_bar}
@@
     \includegraphics[width=\linewidth]{difference/h2-only-full/figures/90_nodes/cost_map.png}
-    \caption{Difference in cost investments between the \Hgrid{} and \modHybrid{}s.
+    \caption{Difference in regional costs between the \Hgrid{} and \modHybrid{}s.
 The left subfigure shows higher spendings per technology and region and transport system for the \modH{}, the right shows higher spendings in the \modHybrid{}.}

### 2023-12-11 workflow/scripts/plot_balance_map.py

@@ -185,7 +185,7 @@ for kind, output in snakemake.output.items():
         sm,
         ax=ax,
         label=f"Average Marginal Price of {title} [{region_unit}]",
-        shrink=1,
+        shrink=0.95,
         pad=0.03,
         aspect=50,
         alpha=region_alpha,
@@ -264,4 +264,5 @@ for kind, output in snakemake.output.items():
     fig.savefig(
         output,
         dpi=300,
+        bbox_inches="tight",
     )

### 2023-12-11 report/submissions/journal/paper/paper.tex
maps: re-add sequestration to legend

@@
 Lines and arrows show the interregional transportation.
-Both models transport \carbon{} and \hydrogen{} respectively from areas with low prices to areas with high prices to supply CU and, in the case of \carbonmodel{}, CS.
+Carbon sequestration offshore are drawn as full circles.
+\carbon{} and \hydrogen{} transport goes from low price to high price areas to supply CU and, in the case of \carbonmodel{}, CS.
     }
@@
 For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
+Carbon Sequestration offshore are drawn as full circles.
 Lines and arrows show the interregional transportation.

### 2023-12-11 workflow/scripts/plot_balance_map.py

@@ -211,6 +211,8 @@ for kind, output in snakemake.output.items():
             cons_carriers = cons_carriers.union(["H$_2$ For Industry"])
         # cheat and add a dummy carrier to align legend with carbon balance
         cons_carriers = list(cons_carriers) + ["", ""]
+    if kind == "carbon":
+        cons_carriers = cons_carriers.union(["CO$_2$ Sequestration"])
 
     add_legend_patches(
         ax,

### 2023-12-11 workflow/scripts/plot_balance_map_dedicated.py

@@ -210,6 +210,8 @@ for n, axs, draw_legend in zip(networks, axes.T, [False, True]):
             if kind == "hydrogen" and "H$_2$ For Industry" in prod_carriers:
                 prod_carriers = prod_carriers.difference(["H$_2$ For Industry"])
                 cons_carriers = cons_carriers.union(["H$_2$ For Industry"])
+            elif kind == "carbon" and "CO$_2$ Sequestration" not in cons_carriers:
+                cons_carriers = cons_carriers.union(["CO$_2$ Sequestration"])
 
             add_legend_patches(
                 ax,

### 2023-12-11 report/submissions/journal/paper/paper.tex
implement to-dos in Introduction, add source for Euro-Calliope

@@
 \begin{abstract}
-    Hydrogen and carbon dioxide transportation are considered crucial in climate-neutral energy systems, with hydrogen enabling energy imports to high-demand areas and carbon transport aiding in emissions export from high-emission areas.
+    Hydrogen and carbon dioxide transportation are considered crucial in climate-neutral energy systems, with hydrogen transport enabling energy imports to high-demand areas and carbon transport aiding in emissions export from high-emission areas.
 Yet, possible synergies and competitions between the two systems are not fully understood.
@@
 The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors.
-While management of both \COtwo{} and \Htwo{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructures effectively interact with comprehensive carbon management technologies, including carbon capture, transport, use, and storage.
-Hydrogen offers an efficient way to transport and store energy.
-Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, and synthetic fuels through carbon capture (CC) techniques, or harvested from the atmosphere using direct air capture (DAC).
-Additionally, it can be stored in geological formations, a process known as carbon sequestration (CS).
-In combination, \COtwo{} and \Htwo{} networks build the basis for climate-neutral fuels needed for aviation, shipping and industrial feedstocks, and are central to carbon utilization (CU) strategies.
+While management of both \COtwo{} and \Htwo{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructure effectively interacts with carbon management technologies, including carbon capture, transport, storage, utilization, and sequestration.
+Hydrogen offers an efficient way to transport and store energy over long distances and timeframes, and can be produced from electrolysis or steam methane reforming (SMR).
+Carbon can be stored in geological formations to prevent it from entering the atmosphere, a process known as carbon sequestration (CS).
+Additionally, carbon is a crucial material input for carbonaceous fuels needed for aviation (synthetic kerosene from Fischer-Tropsch synthesis), shipping (synthetic methanol) and industrial feedstocks (e.g., synthetic methane).
+From hereon, we jointly refer to the production of these carbonaceous fuels as Carbon Utilization (CU).
+Carbon can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, or synthetic carbonaceous fuels through point-source carbon capture (CC) techniques, or harvested from the atmosphere using direct air capture (DAC).
+
+It is important to differentiate between the functions of carbon capture, carbon sequestration and carbon utilization.
+While both academic literature and common language mostly use the term ``carbon capture, utilization and storage'' (CCUS), or similar, to refer to all three functions, we argue that this is misleading.
+There is inherent benefit in separating these three functions geographically (linked through carbon transport), temporally (linked through carbon storage), and in terms of economic actors performing each function.
+CC should be performed at the place and with the technology that can capture carbon at the lowest cost (including a potential transport premium).
+CS should take place where sequestration potentials exist and carbon is available at the lowest cost, and CU where the combination of hydrogen and carbon inputs is available at lowest cost.
+By employing this separation in our model, we allow for a nuanced optimization and detailed analysis of the carbon management system.
 
@@
 
-However, up to this point, it remains unclear how the two transport systems of hydrogen and carbon may complement or replace each other, particularly when it comes to their transport systems.
-A hydrogen network facilitates the transport of energy from renewable sources to regions with geographically fixed hydrogen demand, such as for steelmaking, and also enables CU at the site of point-source capture.
-On the other hand, the carbon transport system allows for capturing and transporting carbon from areas with high emissions to regions with significant sequestration potential and abundant renewable sources, thereby enabling cost-effective CU.
-In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
+However, up to this point, it remains unclear how the two transport systems of hydrogen and carbon may complement or replace each other.
+Both can bridge the misalignment of sources (electrolysis, SMR, point-source CC, DAC) and sinks (CU, CS, industrial hydrogen demand) for carbon and hydrogen.
+A hydrogen network can supply regions with geographically fixed hydrogen demand, such as for steelmaking, with hydrogen from regions with the best renewable resources.
+A hydrogen network can also enable CU at the site of point-source CC.
+On the other hand, the carbon transport system allows for capturing and transporting carbon from areas with high emissions to regions with carbon sequestration potentials or high-quality renewable resources, the latter enabling cost-effective CU.
+In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008, ,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
 Such techno-economic models, in contrast for example to Integrated Assessment Models, can account for the spatial distribution of carbon sources and sinks which are crucial provide a holistic view of the energy system and its technological interactions.
@@
 (extend on hydrogen literature)
+
 The publication by Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \COtwo{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
@@
 Another comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \COtwo{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
-Hoewever, this limited spatial scope fails to consider other sequestration sites and co-benefits from connecting the \COtwo{} network to neighboring countries.
-To this end, it is often argued that \COtwo{} pipelines are a mature technology with an expected high learning rate, given the wide-spread installations in US and Canada for enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
+However, this limited spatial scope fails to consider other sequestration sites and co-benefits from connecting the \COtwo{} network to neighboring countries.
+Considering carbon pipeline as a technological option is a reasonable model assumption since \COtwo{} pipelines are a mature technology, given their wide-spread installations in the United States and Canada for enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
+Other high-resolution energy system models do not feature detailed CU technologies and \COtwo{} transport.
+Euro-Calliope, for example, does neither feature carbon nor hydrogen transport, and thus forces generic synthetic fuel production to rely on captive electrolysis and DAC in the same region \cite{pickeringDiversityOptionsEliminate2022}.
+In our model, we separate the functions of carbon capture (CC), carbon sequestration (CS) and utilization (CU).
 % However, the models are often limited with regard to both geographical scope and detail.

### 2023-12-11 report/submissions/journal/paper/paper.tex
methods section fneum WIP

--- 2023-12-11 report/submissions/journal/paper/paper.tex
@@
 \newcommand{\modHybrid}{Hybrid model}
-
-% alternative
-\newcommand{\carbon}{CO$_2$}
-\newcommand{\hydrogen}{H$_2$}
-\newcommand{\carbongrid}{CO$_2$\=/Grid}
-\newcommand{\hydrogengrid}{H$_2$\=/Grid}
-\newcommand{\baelinemodel}{Baseline model}
-\newcommand{\carbonmodel}{CO$_2$\=/Grid model}
-\newcommand{\hydrogenmodel}{H$_2$\=/Grid model}
-\newcommand{\hybridmodel}{Hybrid model}
 
@@
 
+Addressing climate targets, we define two \COtwo{} target scenarios:
+\begin{itemize}
+    \item[] \textit{Net\=/Zero}: aligning with the EU's 2050 emission targets. Also denoted as NZ.
+    \item[] \textit{Net\=/Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \COtwo{} annually. Also denoted as NN.
+\end{itemize}
+Unless otherwise mentioned, our reference is the Net-Zero scenario. For the Net-Negative scenario, the cap of total carbon sequestration is adjusted from 200~Mt to 660~Mt per year accordingly.
+
+Beyond \COtwo{} targets, we consider four expansion scenarios, referred to as models:
+\begin{itemize}
+    \item[] \textit{Baseline:} Neither \COtwo{} nor \Htwo{} networks are constructed.
+    \item[] \textit{\COgrid{}:} Only the \COtwo{} network is developed.
+    \item[] \textit{\Hgrid{}:} Only the \Htwo{} network is developed.
+    \item[] \textit{Hybrid:} Both \COtwo{} and \Htwo{} networks are developed.
+\end{itemize}
+
+% TODO add two sentences about the used methods for transition
+
 \section{Results}
@@
     \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/captureshare_line.png}
-    \caption{Proportion of plants with integrated carbon capture for all models in the net zero scenario.
-The size of the dots corresponds to the average capacity factor of the respective technology.
-While peak load technologies such as gas-fired combined heat and power (CHP) plants, which are only in operation on a few days a year, are not expanded for carbon capture, baseload point sources such as biomass CHP and process emissions from industry are fully (partially) developed as soon as the transport of \COtwo{} (\Htwo{}) is permitted.}
+    \caption{Share of facilities with integrated carbon capture for all models .}
     % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
@@
     \includegraphics[width=\linewidth]{comparison/single-technologies/figures/90_nodes/balance_map_dedicated.png}
-    \caption{Optimal operation, flows and prices of the carbon (top line) and hydrogen (bottom line) sectors for the \carbonmodel{} (left) and the \hydrogenmodel{} (right) in the net zero scenario.
-For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
-Lines and arrows show the interregional transportation.
-Both models transport \carbon{} and \hydrogen{} respectively from areas with low prices to areas with high prices to supply CU and, in the case of \carbonmodel{}, CS.
-    }
+    \caption{Optimal operation per sector for a net-zero energy system in Europe with average production on the left and average consumption on the right for both, (a) the \COtwo{} sector in the \modCO{} and (b) the \Htwo{} sector in the \modH{}.}
     \label{fig:balance_map}
@@
     \centering
-    \begin{subfigure}{.5\textwidth}
+    \begin{subfigure}{.51\textwidth}
         \centering
@@
     \end{subfigure}%
-    \begin{subfigure}{.5\textwidth}
+    \begin{subfigure}{.49\textwidth}
         \centering
@@
     \end{subfigure}
-    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hybridmodel{} in the net zero scenario.
-For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
-Lines and arrows show the interregional transportation.
-\carbon{} from point-source in the inland either supplies local CU with imported \hydrogen{} or facilitates sequestration in nearby offshore regions.
+    \caption{Optimal production and transport capacities of the carbon and hydrogen sector in a net-zero energy system in Europe with both \COtwo{} and \Htwo{} network expansion (Hybrid).
     % Carbon network looks the same as in~\cite{morbeeOptimisedDeploymentEuropean2012}: two backbones, one in the nothern Europe other in south east.
@@
 Fig.~\ref{fig:capacity_maps} shows the average production, consumption and transport of the two networks in the \modHybrid{}.
-Note that despite similar overall capacities in terms of weight moved, the investment in the hydrogen infrastructure is four times the investment in the carbon infrastructure.
+Note that despite similar overall capacities, the investment in the hydrogen infrastructure is four times the investment in the carbon infrastructure.
 
@@
 This finding is tightly linked to the fact of competing functionalities to offtake carbon emissions.
-A more detailed of differences between the \modH{} and the \modHybrid{} can be found in Fig.~\ref{fig:cost_map_difference}.
+A more detailed of differences between the \modH{} and the \modHybrid{} can be found in Fig.~.
 
-Despite only a further reduction of \label{}0.4 percentage points (\label{}3 billion euros) in total system costs compared to the \modH{}, the \modHybrid{} brings several advantages.
+Despite only small decrease in cost in comparison to the \modH{}, hybrid approach brings several advantages.
 These include a broader range of technologies contributing to its robustness, reduced reliance on DAC, which may be costlier than anticipated in the cost projections, and less land use for wind and solar due to the decreased necessity for DAC.
@@
     \includegraphics[width=0.9\linewidth]{difference/comparison/emission-reduction-0.1/figures/90_nodes/cost_bar.png}
-    \caption[short]{Net change in investments when tightening the \COtwo{} emission target from net-zero to net-negative 10\% of 1990s emissions.
-For all models, Direct Air Capture (DAC) contributes most to the additional \carbon{} removal, requiring further solar, wind and heat pump capacities for electricity and heat input.}
+    \caption[short]{Net change in investments when tightening the \COtwo{} emission target from net-zero to net-negative 10\% of 1990s emissions.}
     \label{fig:net-negative_cost_bar}
@@
 % The technology costs used in the model rely on cost projections that incorporate reductions based on learning rates.
-These learning rates are derived from historical data, which may not necessarily be indicative of future trends, and they are based on projected global installed capacities of one technology, which can result in large uncertainties.
+These learning rates are derived from historical data, which may not necessarily be indicative of future trends.
 In particular, cost assumptions on DAC, electrolysis and \Htwo{} pipeline costs, could have a strong impact in the model results.
@@
 
-\section{Methodology}
+\section*{Methods}
 \label{sec:methodology}
 
-The study is conducted on the basis of the open-source, capacity-expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
-The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, waste, agriculture, and transport sectors, including international aviation and shipping.
+The study is conducted on the basis of the open-source, capacity expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
+The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, agriculture, and transport sectors, including international aviation and shipping.
+All technology cost assumptions are taken for the year 2040 and sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
+Many of the assumptions contained therein are based on the technology catalogues published by the Danish Energy Agence~\cite{TODO}.
+Endogenous model results include the expansion of renewable energy sources, storage technologies, power-to-X conversion and transmission capacities.
+The model considers various energy carriers like electricity, heat, hydrogen, methane, methanol, liquid hydrocarbons and biomass, together with the corresponding conversions technologies.
 
@@
 \end{figure}
+
+% emits 633~Mt \COtwo{} per year from industry, aviation, shipping and agriculture, 153~Mt of which are fossil-based process emissions -> emissions from power and gas are flexible, so leave out?
+should also say how much is captured and sequestered
+% excess heat potentials -> are not used or taken from hotmaps; also energy demand comes from JRC-IDEES, Hotmaps only used for geographical distribution
+% more detail on heating technologies?
+% with transport costs considered? - how high are they?
+% CU: The processed fuels are not limited in their total quantity of use or production. -> implicitly limited through demand
+
 %
-In our configuration, the model's time horizon spans one year with a temporal resolution of 3 hours and a spatial resolution of 90 regions.
+In our configuration, the model's time horizon spans one year with a 3-hourly temporal resolution and a spatial resolution of 90 regions.
 Each of the regions consists of a complex subsystem with technologies for supplying, converting, storing and transporting energy.
@@
 The energy demand for electricity, transport, biomass, heat and gas are defined per region and time-step.
-Land transport demand is exogenously divided between electric vehicles (85\%) and fuel cell vehicles (15\%), the latter representing demand for heavy duty land transport.
-Demands for kerosene for aviation, methanol for shipping, and naphtha for industry are aggregated in the system scope and kept constant throughout all time-steps.
+Land transport demand is exogenously assumed to be fully electrified, including heavy-duty vehicles.
+Demands for kerosene for aviation, methanol for shipping, and naphtha for industry are not spatially resolved and assumed to be constant throughout all time-steps.
 Heat demand is regionally subdivided into shares of urban, rural and industrial sites.
@@
 Industrial energy demand and excess heat potentials are calculated per node on the basis of~\cite{hotmaps_industrial_db}.
-% use https://pypsa-eur.readthedocs.io/en/latest/licenses.html
 
-Endogenous model results include the expansion of renewable energy sources, storage technologies, transmission capacities, heating technologies, peaking power plants, and the deployment of gray, blue or green hydrogen, among others.
-The model considers various energy carriers like electricity, hydrogen, methan, methanol, liquid hydrocarbons and biomass, together with conversions technologies.
-Carbon-neutral electricity is provided by wind, solar, biomass, hydro and nuclear power plants.
-Hydro and nuclear plant capacities cannot be extended.
-Weather-dependent power potentials for solar, wind and hydro are calculated from reanalysis and satellite data sets~\cite{hersbachERA5GlobalReanalysis2020,pfeifrothSurfaceRadiationData2017}  per region and time-stamp, using the open-source tool Atlite~\cite{hofmannAtliteLightweightPython2021}.
-Solar and wind power can be expanded in alignment with eligible land-use restrictions calculated on the basis of~\cite{eeaCorineLandCover2012,eeaNatura2000Data2016}.
-We restrict the electric transmission system's expansion to 20\% of its current capacity, acknowledging the challenges in inaugurating new transmission projects.
+% Low-carbon electricity potentials
+
+Low-carbon electricity is provided by wind, solar, biomass, hydro-electricity and nuclear power plants.
+Hydro-electric and nuclear plants cannot be extended beyond currently installed capacities.
+Weather-dependent power potentials for solar, wind and hydro-electricity are calculated from the reanalysis and satellite data sets, ERA5 and SARAH-2,~\cite{hersbachERA5GlobalReanalysis2020,pfeifrothSurfaceRadiationData2017} per region and time-stamp, using the open-source tool Atlite~\cite{hofmannAtliteLightweightPython2021}.
+Solar and wind power can be expanded in alignment with land-use restrictions calculated, taking into account land usage classes and natural protection areas~\cite{eeaCorineLandCover2012,eeaNatura2000Data2016}.
+We restrict the total volume of power transmission expansion to 20\% of its current capacity, acknowledging the challenges in inaugurating new transmission projects.
 For the use of biomass we consider only residual biomass products and no energy crops.
-We limit regional biomass use to the medium-level potentials from the ENPRESO database~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
-Inter-regional biomass transport is permitted with transport costs considered.
+We limit regional biomass use to the medium-level potentials from the JRC-ENSPRESO database~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
+Inter-regional biomass transport is permitted with transport costs of XXX €/MWh/km considered.
 
-In our model, we consider green, blue and gray hydrogen from electrolysis and steam methane reforming (SMR), the latter of which may be equipped with CC technology.
-Geological distribution and potentials for \Htwo{} storage in salt caverns are derived from~\cite{caglayanTechnicalPotentialSalt2020}.
-Re-electrification of hydrogen is possible via fuel cells.
-If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limit, considering costs for pipeline segments and compressors.
-Pipeline flows are modelled using net transfer capacities and without flow dynamics, pressure valves, or energy demand for compression.
-% No retrofitting of gas pipelines is considered, potentially overestimating hydrogen network cost.
+% energy transport
 
 The topology of the \COtwo{}, \Htwo{} and gas network is identical to the topology of the electricity network, connecting all neighboring regions.
-Transport of liquid fuels like oil, methanol and Fischer-Tropsch (FT) fuels is ``copper-plated'', since transport costs per unit of energy are negligible due to their high energy density.
+Liquid fuels like oil, methanol and Fischer-Tropsch (FT) fuels are not spatially resolved, since transport costs per unit of energy are negligible due to their high energy density.
 Throughout this study, we focus on \COtwo{} and \Htwo{} networks because of their high relevance for infrastructure investments and thus public policy decisions.
@@
 Gas networks also already exist, and will most likely experience decreased use in the future, removing any bottlenecks or constraints on the optimal system buildout or operation.
+If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limit, considering costs for pipeline segments and compressors.
+Pipeline flows are modelled using net transfer capacities and without flow dynamics, pressure valves, or energy demand for compression.
+No retrofitting of gas pipelines is considered.
+% lossless linearised power flow approximation for electricity network
+% gas network data from SciGrid gas
+% power network data from ENTSO-E transparency map plus selected TYNDP projects as additional expansion options
+
+% modelling of hydrogen supply chain pathways
+
+In our model, hydrogen can be produced from electrolysis and steam methane reforming (SMR) with or without carbon capture.
+The geographical distribution of underground hydrogen storage potentials in salt caverns is derived from Caglayan et al.~\cite{caglayanTechnicalPotentialSalt2020}.
+Re-electrification of hydrogen is possible via fuel cells.
+
+% carbon utilisation
+
+% TODO add sentence distinguishing CU, CT, CS, CC
 
@@
 The processed fuels are not limited in their total quantity of use or production.
-Methane is transported through the gas network, while methanol and FT fuel can be transferred inter-regionally without additional costs, since dense fuels have negligible transport costs per unit of energy.
-Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial process heat.
-Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production, aviation kerosene, or agricultural machinery oil.
+Methane is transported through the gas network, while methanol and FT fuel can be transported between regions without additional costs or capacity constraints.
+Synthetic methane substitutes natural gas or upgraded biogas, serving combined heat and power plants, residential heating gas boilers, or industrial process heat.
+Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production for high-value chemicals, aviation kerosene, or agricultural machinery oil.
 
-To supply carbon needed for CS and CU, the system can choose to deploy carbon capture technology at various point sources (see below), or through DAC facilities.
-The concept of a merit order for captured carbon plays a pivotal role in optimizing the deployment of carbon capture technologies based on their economic feasibility.
-This merit order ranks all CC technologies according to their relative costs to capture an additional marginal ton\footnote[2]{The cost value for a technology depends on the exact system configuration and may vary by scenario and even by region, since many technologies produce other (primary) outputs such as electricity and heat.
-The costs to capture an additional marginal ton of carbon may also be different from the average cost of capture.} of carbon (see Fig.~X).
+% carbon supply
+
+To supply carbon needed for CS and CU, the system can choose to deploy carbon capture (CC) technologies at various point sources of fossil or biogenic origin (see below), or through DAC facilities.
+The concept of a merit order for captured carbon plays a pivotal role in optimizing and prioritizing the deployment of carbon capture technologies based on their economic feasibility.
+This merit order ranks all CC technologies according to their relative costs to capture a ton of carbon (see Fig.~X).
 At the lower end of the cost spectrum, CC technologies applied to process emissions, such as those from cement, offer a cost-effective starting point.
@@
 DAC, an emerging technology capable of extracting \COtwo{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand.
-Finally, based on our model results, biogas-to-gas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per marginal ton of captured \COtwo{}.
+Finally, based on our model results, biogas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per marginal ton of captured \COtwo{}.
 Biogas input is a high-cost fuel which based on the endogenous modeling decisions is not used in large quantities.
@@
 
-To store carbon, we differentiate between short-term storage in steel tanks and long-term, irreversible sequestration in underground sequestration sites such as porous rock formations or depleted gas reservoirs.
+% carbon storage and sequestration
+
+To store carbon, we differentiate between short-term storage in steel tanks without permanent containment and long-term, irreversible sequestration in underground sequestration sites such as porous rock formations or depleted gas reservoirs.
 Costs for both options are included in the model.
 For carbon sequestration, we only consider offshore sites as potential sinks.
-We make this choice because offshore sequestration sites typically have larger capacity compared to onshore sequestration sites in saline aquifers and due to concerns over public safety for infrastructure near populated areas.
-Our estimates for carbon sequestration potentials are conservative, limiting total sequestration to 25~Mt per site and calculating annual storage availability over 25 years.
-Furthermore, we decide to cap the total sequestered \COtwo{} to 200~Mt per year for our Net-Zero scenario.
-This is enough to offset the hardest-to-abate fossil process emissions and some limited slack of fossil fuels to be used where they can be most valuable.
+We make this choice because offshore sequestration sites typically have larger capacity compared to onshore sequestration sites in saline aquifers and due to concerns over public safety for infrastructure near populated areas~\cite{TODO}.
+Our estimates for carbon sequestration potentials are restrictive, limiting the annual sequestration potential to 25~Mt per region and constraining it further such that this amount could be  25 years.
+Furthermore, we decide to limit the total amount of sequestered \COtwo{} to 200~Mt per year for our Net-Zero scenario and XXX~Mt per year for our Net-Negative scenario.
+This is enough to offset hard-to-abate fossil process emissions (e.g.
+calcination of limestone in cement manufacturing) and some limited slack of fossil fuels to be used where they can be most valuable.
 This is in order to avoid offsetting fossil emissions with carbon removal if they can technologically be avoided in the first place.
-All technology cost assumptions are taken for the year 2040 and sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
-
-Addressing climate targets, we define two \COtwo{} target scenarios:
-\begin{itemize}
-    \item[] \textit{Net\=/Zero}: aligning with the EU's 2050 emission targets. Also denoted as NZ.
-    \item[] \textit{Net\=/Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \COtwo{} annually. Also denoted as NN.
-\end{itemize}
-Unless otherwise mentioned, our reference is the Net-Zero scenario. For the Net-Negative scenario, the cap of total carbon sequestration is adjusted from 200~Mt to 660~Mt per year accordingly.
-
-Beyond \COtwo{} targets, we consider four expansion scenarios, referred to as models:
-\begin{itemize}
-    \item[] \textit{Baseline:} Neither \COtwo{} nor \Htwo{} networks are constructed.
-    \item[] \textit{\COgrid{}:} Only the \COtwo{} network is developed.
-    \item[] \textit{\Hgrid{}:} Only the \Htwo{} network is developed.
-    \item[] \textit{Hybrid:} Both \COtwo{} and \Htwo{} networks are developed.
-\end{itemize}
 
@@
     \includegraphics[width=\linewidth]{difference/h2-only-full/figures/90_nodes/cost_map.png}
-    \caption{Difference in regional costs between the \Hgrid{} and \modHybrid{}s.
+    \caption{Difference in cost investments between the \Hgrid{} and \modHybrid{}s.
 The left subfigure shows higher spendings per technology and region and transport system for the \modH{}, the right shows higher spendings in the \modHybrid{}.}
@@
     \caption[short]{Net change in system cost when reducing assumed costs for \COtwo{} pipelines by 50\%.}
-    \label{fig:half-CO2-price_cost_bar}
+    \label{fig:net-negative_cost_bar}
 \end{figure}

### 2023-12-11 report/submissions/journal/paper/paper.tex

--- 2023-12-11 report/submissions/journal/paper/paper.tex
@@
 We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available sequestration potential.
+
+Addressing climate targets, we define two \COtwo{} target scenarios:
+\begin{itemize}
+    \item[] \textit{Net\=/Zero}: aligning with the EU's 2050 emission targets. Also denoted as NZ.
+    \item[] \textit{Net\=/Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \COtwo{} annually. Also denoted as NN.
+\end{itemize}
+Unless otherwise mentioned, our reference is the Net-Zero scenario. For the Net-Negative scenario, the cap of total carbon sequestration is adjusted from 200~Mt to 660~Mt per year accordingly.
+
+Beyond \COtwo{} targets, we consider four expansion scenarios, referred to as models:
+\begin{itemize}
+    \item[] \textit{Baseline:} Neither \COtwo{} nor \Htwo{} networks are constructed.
+    \item[] \textit{\COgrid{}:} Only the \COtwo{} network is developed.
+    \item[] \textit{\Hgrid{}:} Only the \Htwo{} network is developed.
+    \item[] \textit{Hybrid:} Both \COtwo{} and \Htwo{} networks are developed.
+\end{itemize}
+
+% TODO add two sentences about the used methods for transition
 
@@
 
-\section{Methodology}
+\section*{Methods}
 \label{sec:methodology}
 
-The study is conducted on the basis of the open-source, capacity-expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
-The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, waste, agriculture, and transport sectors, including international aviation and shipping.
+The study is conducted on the basis of the open-source, capacity expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
+The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, agriculture, and transport sectors, including international aviation and shipping.
+All technology cost assumptions are taken for the year 2040 and sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
+Many of the assumptions contained therein are based on the technology catalogues published by the Danish Energy Agence~\cite{TODO}.
+Endogenous model results include the expansion of renewable energy sources, storage technologies, power-to-X conversion and transmission capacities, heating technologies, peaking power plants, and the deployment of gray, blue or green hydrogen, among others.
+The model considers various energy carriers like electricity, heat, hydrogen, methane, methanol, liquid hydrocarbons and biomass, together with the corresponding conversions technologies.
 
@@
 Gas networks also already exist, and will most likely experience decreased use in the future, removing any bottlenecks or constraints on the optimal system buildout or operation.
+If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limit, considering costs for pipeline segments and compressors.
+Pipeline flows are modelled using net transfer capacities and without flow dynamics, pressure valves, or energy demand for compression.
+No retrofitting of gas pipelines is considered.
+% lossless linearised power flow approximation for electricity network
+% gas network data from SciGrid gas
+% power network data from ENTSO-E transparency map plus selected TYNDP projects as additional expansion options
+
+% modelling of hydrogen supply chain pathways
+
+In our model, we consider green, blue and gray hydrogen from electrolysis and steam methane reforming (SMR), the latter of which may be equipped with CC technology.
+The geographical distribution of underground hydrogen storage potentials in salt caverns is derived from Caglayan et al.~\cite{caglayanTechnicalPotentialSalt2020}.
+Re-electrification of hydrogen is possible via fuel cells.
+
+% carbon utilisation
+
+% TODO add sentence distinguishing CU, CT, CS, CC
 
@@
 The processed fuels are not limited in their total quantity of use or production.
-Methane is transported through the gas network, while methanol and FT fuel can be transferred inter-regionally without additional costs, since dense fuels have negligible transport costs per unit of energy.
-Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial process heat.
-Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production, aviation kerosene, or agricultural machinery oil.
+Methane is transported through the gas network, while methanol and FT fuel can be transported between regions without additional costs or capacity constraints.
+Synthetic methane substitutes natural gas or upgraded biogas, serving combined heat and power plants, residential heating gas boilers, or industrial process heat.
+Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production for high-value chemicals, aviation kerosene, or agricultural machinery oil.
 
-To supply carbon needed for CS and CU, the system can choose to deploy carbon capture technology at various point sources (see below), or through DAC facilities.
+% carbon supply
+
+To supply carbon needed for CS and CU, the system can choose to deploy carbon capture (CC) technologies at various point sources of fossil or biogenic origin (see below), or through DAC facilities.
 The concept of a merit order for captured carbon plays a pivotal role in optimizing the deployment of carbon capture technologies based on their economic feasibility.
 This merit order ranks all CC technologies according to their relative costs to capture an additional marginal ton\footnote[2]{The cost value for a technology depends on the exact system configuration and may vary by scenario and even by region, since many technologies produce other (primary) outputs such as electricity and heat.
-The costs to capture an additional marginal ton of carbon may also be different from the average cost of capture.} of carbon (see Fig.~X).
+The costs to capture an additional marginal ton of carbon may also be different from the average cost of capture.}
 At the lower end of the cost spectrum, CC technologies applied to process emissions, such as those from cement, offer a cost-effective starting point.

### 2023-12-12 report/submissions/journal/paper/paper.tex
paper: add discussed changes in Intro

@@
 The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors.
-While management of both \COtwo{} and \Htwo{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructure effectively interacts with carbon management technologies, including carbon capture, transport, storage, utilization, and sequestration.
-Hydrogen offers an efficient way to transport and store energy over long distances and timeframes, and can be produced from electrolysis or steam methane reforming (SMR).
+While management of both \COtwo{} and \Htwo{}, also refered to as carbon and hydrogen, is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructure effectively interacts with carbon management technologies, including carbon capture, transport, storage, utilization, and sequestration.
+Hydrogen offers an efficient way to transport and store energy over long distances and timeframes, and can be produced from electrolysis or steam methane reforming (SMR), for example.
 Carbon can be stored in geological formations to prevent it from entering the atmosphere, a process known as carbon sequestration (CS).
-Additionally, carbon is a crucial material input for carbonaceous fuels needed for aviation (synthetic kerosene from Fischer-Tropsch synthesis), shipping (synthetic methanol) and industrial feedstocks (e.g., synthetic methane).
+Additionally, carbon can be combined with hydrogen to produce carbonaceous fuels needed for aviation (synthetic kerosene from Fischer-Tropsch synthesis), shipping (synthetic methanol) and industrial feedstocks (e.g., synthetic methane).
 From hereon, we jointly refer to the production of these carbonaceous fuels as Carbon Utilization (CU).
 Carbon can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, or synthetic carbonaceous fuels through point-source carbon capture (CC) techniques, or harvested from the atmosphere using direct air capture (DAC).
+In combination, \COtwo{} and \Htwo{} networks are crucial to enable these carbon management strategies necessary for net-zero economies.
 
@@
 The study is conducted using the PyPSA-Eur energy system model and encompasses all relevant energy sectors.
-We explore competition between \COtwo{} and \Htwo{} networks by adding them separately.
-Our evaluation focuses on the transport of \COtwo{} and \Htwo{} through their respective networks on the European continent.
-We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available sequestration potential.
+We investigate the competitive transport dynamics between \COtwo{} and \Htwo{} networks.
+This is done by contrasting models that feature only one of these networks, neither, or both combined.
+We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon and hydrogen networks varies given net-zero or net-negative emissions targets.
 

### 2023-12-13 report/submissions/journal/paper/paper.tex
paper intro gather remaining todos

@@
 Yet, possible synergies and competitions between the two systems are not fully understood.
-Our study employs optimization techniques to develop a cost-optimal European energy system, integrating transport, storage, and sequestration of both carbon dioxide and hydrogen, along with renewable energy sources.
-Results indicate that a hydrogen network is more cost-effective than a carbon network, facilitating low-cost hydrogen imports to Central Europe and local carbon capture and utilization.
+Our study employs optimization techniques to develop a cost-optimal European energy system, integrating production, transport and storage of both carbon dioxide and hydrogen, along with renewable energy sources.
+Results indicate that, in direct comparison, a hydrogen network is more cost-effective than a carbon network, facilitating low-cost hydrogen transport to Central Europe and local carbon capture and utilization.
 However, in a hybrid scenario, the carbon network effectively complements the hydrogen network, promoting carbon capture from biomass and reducing reliance on direct air capture.
@@
 Hydrogen offers an efficient way to transport and store energy over long distances and timeframes, and can be produced from electrolysis or steam methane reforming (SMR), for example.
-Carbon can be stored in geological formations to prevent it from entering the atmosphere, a process known as carbon sequestration (CS).
+Carbon can be stored in geological formations to permanently extract it from the atmosphere, a process known as carbon sequestration (CS).
 Additionally, carbon can be combined with hydrogen to produce carbonaceous fuels needed for aviation (synthetic kerosene from Fischer-Tropsch synthesis), shipping (synthetic methanol) and industrial feedstocks (e.g., synthetic methane).
@@
 Carbon can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, or synthetic carbonaceous fuels through point-source carbon capture (CC) techniques, or harvested from the atmosphere using direct air capture (DAC).
-In combination, \COtwo{} and \Htwo{} networks are crucial to enable these carbon management strategies necessary for net-zero economies.
+In combination, \COtwo{} and \Htwo{} networks are crucial to enable these carbon management strategies for net-zero economies.
+% TODO: integrate definition of Carbon Management (CM) into last sentence.
+% Cx: CU, CS, CT, CC, Cstorage --> CO2 + H2 networks are crucial to carry out CM strategies
+% Suggestion:
 
+% TODO: check definition of carbonaceous fuels for "missing" (relevant) fuel types
+
+% TODO: boil this down to one sentences and integrate it to the previous paragraph or move it two the literature part.
+% Suggestion: We argue that it is crucial to consider CC, CU and CS as separate technologies, rather than grouping them under the often-used but misleading umbrella term "CCUS", to allow for more accurate optimization and analysis of each function in the carbon management system.
+% Suggestion: We argue that it is crucial to consider CC, CU and CS as separate technologies, rather than grouping them under the often-used but misleading umbrella term ``carbon capture, utilization and storage'' (CCUS), to allow for independent optimization and adequate representation of each function in the carbon management system.
 It is important to differentiate between the functions of carbon capture, carbon sequestration and carbon utilization.
@@
 
-Recently, policymakers and industry in Europe have started developing carbon management strategies~\cite{GermanyDevelopingStrategy2023,CarbonManagementStrategie}, planning infrastructure components~\cite{CONetz}, and committing on the first carbon utilization projects~\cite{EFuelsPilotPlant2022,OrstedAssumesFull,GROUNDBREAKINGEFUELPRODUCTION,DLREfuelsDLR}.
+Recently, policymakers and industry in Europe have started developing hydrogen and carbon management strategies~\cite{GermanyDevelopingStrategy2023,CarbonManagementStrategie}, planning infrastructure components~\cite{CONetz}, and committing on the first carbon utilization projects~\cite{EFuelsPilotPlant2022,OrstedAssumesFull,GROUNDBREAKINGEFUELPRODUCTION,DLREfuelsDLR}.
 With the European Union's goal of achieving climate neutrality by 2050, a wide range of programs, funding models and initiatives have been established in this regard~\cite{eu2023netzero,europeangreendeal,europeaninnovationfund}.
-Initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen Infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier, and some gas pipelines are already repurposed to transport hydrogen~\cite{RohrFreiFuer}.
-At the same time, business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synfuels on the one hand and offer \COtwo{} offtake on the other hand.
-The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a transport and sequestration capacity of 1.5 Mt \COtwo{} per year  to be operative in 2024, expanding to a targeted scale of 5 Mt per year of sequestration by 2030.
+% TODO: mention policy ambition for carbon and hydrogen strategies separately in two sentences.
+Collaborative industry initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen Infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier, and some gas pipelines are already repurposed to transport hydrogen~\cite{RohrFreiFuer}.
+Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synfuels on the one hand and offer \COtwo{} offtake on the other hand.
+The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a transport and sequestration capacity of 1.5~Mt \COtwo{} per year to be operative in 2024, expanding to a targeted scale of 5~Mt per year of sequestration by 2030.
 % TODO: frame next two sentences to be more market oriented
-For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.7~Gt of carbon capture from point sources per year, which represent roughly 50\% of all emissions.
-In combination with large sequestration potentials as stated in~\cite{weiProposedGlobalLayout2021}, this highlights the vast potential for decarbonization.
+% TODO: rather cite europeancommissionEuropeanCO2Storage
+For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.7~Gt of carbon capture from point sources per year, which represent roughly 50\% of all current emissions.
+In combination with large sequestration potentials as stated in~\cite{weiProposedGlobalLayout2021}, this highlights the vast potential for carbon capture.
+% TODO: optionally find better place for this
 To this end, the Clean Air Task Force underlines the importance of a carbon transport system in Europe to facilitate the carbon economy~\cite{lockwoodEuropeanStrategyCarbon}.
+Considering carbon pipeline as a technological option is a reasonable model assumption since \COtwo{} pipelines are a mature technology, given their wide-spread installations in the United States and Canada for enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
 
+% TODO: explain why misalignment takes place -> exogenous factors renewable potentials, industrial clusters with h2 demand, co2 point sources, sequestration sites,
 However, up to this point, it remains unclear how the two transport systems of hydrogen and carbon may complement or replace each other.
 Both can bridge the misalignment of sources (electrolysis, SMR, point-source CC, DAC) and sinks (CU, CS, industrial hydrogen demand) for carbon and hydrogen.
-A hydrogen network can supply regions with geographically fixed hydrogen demand, such as for steelmaking, with hydrogen from regions with the best renewable resources.
-A hydrogen network can also enable CU at the site of point-source CC.
-On the other hand, the carbon transport system allows for capturing and transporting carbon from areas with high emissions to regions with carbon sequestration potentials or high-quality renewable resources, the latter enabling cost-effective CU.
-In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008, ,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}.
-Such techno-economic models, in contrast for example to Integrated Assessment Models, can account for the spatial distribution of carbon sources and sinks which are crucial provide a holistic view of the energy system and its technological interactions.
-The work in~\cite{neumannBenefitsHydrogenNetwork2022} examines the effect of a hydrogen network in Europe...
-(extend on hydrogen literature)
-
+A hydrogen network can supply regions with geographically fixed hydrogen demand, such as for steelmaking, with hydrogen from regions with the best renewable resources as well as enable CU at the site of point-source CC.
+On the other hand, the carbon transport system allows for transporting captured carbon to regions with sequestration potentials or high-quality renewable resources, the latter enabling cost-effective CU.
+In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022,neumannBenefitsHydrogenNetwork2022}.
+Such techno-economic models, in contrast for example to Integrated Assessment Models, can account for the spatial distribution of carbon sources and sinks which are crucial to provide a holistic view of the energy system and its technological interactions.
+A remarkable example was developed by Neumann et al.~\cite{neumannBenefitsHydrogenNetwork2022}, which comprehensively demonstrates the interchangeability of hydrogen and electricity grid expansion in a climate-neutral European energy system at the expense of higher investments.
+The underlying, highly-resolved model encompasses most relevant sectors, however neglects the option to transport carbon across regions.
 The publication by Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \COtwo{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
 This limited sectoral scope cannot capture important dynamics of carbon management, since it neglects the sectors that will need to handle most \COtwo{} in the future.
-Another comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \COtwo{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
+A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \COtwo{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
 However, this limited spatial scope fails to consider other sequestration sites and co-benefits from connecting the \COtwo{} network to neighboring countries.
-Considering carbon pipeline as a technological option is a reasonable model assumption since \COtwo{} pipelines are a mature technology, given their wide-spread installations in the United States and Canada for enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
 Other high-resolution energy system models do not feature detailed CU technologies and \COtwo{} transport.
-Euro-Calliope, for example, does neither feature carbon nor hydrogen transport, and thus forces generic synthetic fuel production to rely on captive electrolysis and DAC in the same region \cite{pickeringDiversityOptionsEliminate2022}.
-In our model, we separate the functions of carbon capture (CC), carbon sequestration (CS) and utilization (CU).
-% However, the models are often limited with regard to both geographical scope and detail.
-While representing a single country with spatial resolution may neglect synergies of international cooperation, a coarse grained representation of multiple countries may neglect important geographical properties.
+Euro-Calliope, for example, does neither feature carbon nor hydrogen transport, and thus forces generic synthetic fuel production to rely on captive electrolysis and DAC in the same region~\cite{pickeringDiversityOptionsEliminate2022}.
 
 To our knowledge, no study has yet considered the co-optimization and comprehensive assessment of both \COtwo{} and \Htwo{} networks in a fully sector-coupled energy system.
-However, we argue that such an assessment is strongly needed to avoid suboptimal investments and to identify synergies between hydrogen and carbon management technologies.
+However, we argue that such an assessment is strongly needed to identify realms of competition and synergy between hydrogen and carbon management technologies.
 In this paper, we present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies.
@@
 
-Addressing climate targets, we define two \COtwo{} target scenarios:
+We define two \COtwo{} target scenarios:
 \begin{itemize}
-    \item[] \textit{Net\=/Zero}: aligning with the EU's 2050 emission targets. Also denoted as NZ.
-    \item[] \textit{Net\=/Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \COtwo{} annually. Also denoted as NN.
+    \item[] \textit{Net\=/Zero}: aligning with the EU's 2050 emission targets. Limiting carbon sequestration to 200~Mt/a. Also denoted as NZ.
+    \item[] \textit{Net\=/Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \COtwo{} annually. Limiting carbon sequestration to 660~Mt/a. Also denoted as NN.
 \end{itemize}
-Unless otherwise mentioned, our reference is the Net-Zero scenario. For the Net-Negative scenario, the cap of total carbon sequestration is adjusted from 200~Mt to 660~Mt per year accordingly.
+Unless otherwise mentioned, our reference is the Net-Zero scenario.
 
@@
 
+% TODO: avoid to long paranthesis
 In all models, almost half of the system costs are spent on primary electricity production from wind and solar power (\label{}45\% or \label{}360 of \label{}799 billion euros total costs in the \modBase{}), with another \label{}12\% (\label{}100 billion euros) spent on hydro, nuclear and biomass.
@@
 System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \COtwo{} and \Htwo{} networks, as well as carbon capture technologies (DAC and CC at point sources).
-As we show below, these are due to substantial, albeit fundamentally different, gains in flexibility and a shift in carbon utilization and storage strategies by the \Htwo{} and \COtwo{} networks.
+As we show below, these are due to substantial, albeit fundamentally different, gains in flexibility and a shift in carbon utilization and sequestration strategies by the \Htwo{} and \COtwo{} networks.
 
@@
 In comparison to the other three models, the \modBase{} reveals significant inefficiencies: On the one hand, regions with no access to sequestration sites (i.e., ``landlocked'' regions) underutilize CC potentials from process emissions, fossil gas and biomass.
-On the other hand, sequestration sites rely on costly DAC to capture the carbon for CS from the air.
-Furthermore, hydrogen and power prices show high regional differences, indicating system inefficiencies due to transport constraints caused by non-existent hydrogen or insufficient power transmission.
+On the other hand, sequestration regions rely on costly DAC on-site.
+Furthermore, hydrogen and power prices show high regional differences, indicating system inefficiencies due to transport constraints caused by non\=/existent hydrogen or insufficient power transmission.
 Prices are significantly higher in Central Europe, a region with high population density and high concentration of energy-intensive industries.
@@
 Another knock-on effect of the reduced reliance on DAC is reduced endogenous demand for renewables to power DAC facilities.
+% TODO: make sure FT and methanolisation is properly explained in methods section as almost exogenously fixed products
 
@@
 
-The increase in captured carbon from point sources in the network models is mostly due to CC infrastructure being deployed on industrial or power plant facilities that are present in all models (as opposed to the deployment of additional electricity generating capacity for the purpose of carbon capture).
+% TODO: either explain what network models mean or use another framing
+% TODO: extend first sentence
+The increase in captured carbon from point sources in the network models is mostly due to CC infrastructure being deployed on industrial or power plant facilities that are present in all models, as opposed to the deployment of additional electricity generating capacity for the purpose of carbon capture.
 This is evident when analyzing the capture share of the different CC technologies (see Fig.~\ref{fig:captureshare_line}).
@@
 In the \modBase{}, the capture share is highest for biogas-to-gas facilities, followed by process emissions, biomass for industry, biomass CHPs, gas for industry, SMR and gas CHPs.
-For the network models, we can observe a correlation between CC share and capacity factors of the underlying technologies: Process emissions as well as gas and biomass for industry have high capacity factors (above 80\%), and CC shares close to 100\% (the only exception being gas for industry in the \modH{}).
+For the network models, we can observe a correlation between CC share and capacity factors of the underlying technologies: Process emissions as well as gas and biomass for industry have high capacity factors (above 80\%), and CC shares close to 100\%, the only exception being gas for industry in the \modH{}.
 Biomass CHPs have a capacity factor of around 40\%, and CC shares of 50\% in the \modH{} and 95\% in the \modCO{}.
 SMR and gas CHPs have capacity factors below 20\%, and CC shares around 50\% and 0\%, respectively, across all network models.
-Gas CHPs serve as peak load electricity production technology, and with very few operating hours, the high investment costs into CC applications are not efficient.\footnote[1]{Similarly, SMR serves as ``peak-load'' hydrogen production technology in regions with poor renewable resources for electrolysis.}
+Gas CHPs serve as peak load electricity production technology, and with very few operating hours, thus the high investment costs into CC applications are not efficient.
+Similarly, SMR serves as ``peak-load'' hydrogen production technology in regions with poor renewable resources for electrolysis.
 
+% TODO: change opener for next section: co2gird has less DAC and more CC at point sources, however h2grid is cheaper
 In summary, the implementation of \COtwo{} and \Htwo{} networks is instrumental in unlocking low-cost CC potentials at geographically fixed point sources across the continent.
-While the \COtwo{} network facilitates the movement of carbon from low-cost sources to CU and CS sinks, reducing the reliance on high-cost DAC, the \Htwo{} network enables the cost-effective delivery of hydrogen from regions rich in renewable energy to regions with low-cost CC potentials to produce CU.
 However, despite the \modCO{} demonstrating higher absolute amounts of CC from point sources and higher CC shares across all point-source CC technologies, the \modH{} achieves lower total system costs (see Fig.~\ref{fig:cost_bar}).
@@
 Furthermore, we decide to limit the total amount of sequestered \COtwo{} to 200~Mt per year for our Net-Zero scenario and XXX~Mt per year for our Net-Negative scenario.
-This is enough to offset hard-to-abate fossil process emissions (e.g.
-calcination of limestone in cement manufacturing) and some limited slack of fossil fuels to be used where they can be most valuable.
+This is enough to offset hard-to-abate fossil process emissions (e.g., calcination of limestone in cement manufacturing) and some limited slack of fossil fuels to be used where they can be most valuable.
 This is in order to avoid offsetting fossil emissions with carbon removal if they can technologically be avoided in the first place.

### 2023-12-13 workflow/scripts/plot_balance_bar.py
workflow: write out plotted data to csv files

@@ -42,6 +42,8 @@ df = pd.concat(df, axis=1).fillna(0)
 df = df.rename(lambda k: k.replace(" CC", ""), level=1).groupby(level=[0, 1, 2]).sum()
 
 for kind, output in snakemake.output.items():
+    if kind.startswith("table"):
+        continue
     kind_to_carrier = {"hydrogen": "H2", "carbon": "co2 stored", "electricity": "AC"}
     carrier = kind_to_carrier.get(kind, kind)
     label = config["labels"].get(kind, kind.title())
@@ -107,3 +109,4 @@ for kind, output in snakemake.output.items():
 
     sns.despine()
     fig.savefig(output, dpi=300)
+    ds.round(3).to_csv(snakemake.output[f"table-{kind}"])

### 2023-12-13 workflow/scripts/plot_captureshare_line.py

@@ -120,7 +120,8 @@ plt.xticks(
 
 sns.despine()
 
-fig.savefig(snakemake.output[0], dpi=300)
+fig.savefig(snakemake.output.figure, dpi=300)
+data.to_csv(snakemake.output.table)
 
 # ALTERATIVE PLOT
 # fig, ax = plt.subplots(

### 2023-12-13 workflow/scripts/plot_cost_bar.py

@@ -114,3 +114,4 @@ ax.legend(
 
 sns.despine()
 fig.savefig(snakemake.output[0], dpi=300)
+grouped.fillna(0).round(3).to_csv(snakemake.output.table)

### 2023-12-13 workflow/scripts/plot_cost_bar_compare_diff.py

@@ -99,4 +99,5 @@ ax.legend(
 )
 
 sns.despine()
-fig.savefig(snakemake.output[0], dpi=300)
+fig.savefig(snakemake.output.figure, dpi=300)
+grouped.round(3).to_csv(snakemake.output.table)

### 2023-12-13 report/submissions/journal/paper/paper.tex
workflow: write out plotted data to csv files
update numbers to new results adjust legend in capture line plot

@@
 
-Total annual system costs in the \modBase{} are highest, totalling \label{}803 billion euros.
-The models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs, with around \label{}2\% cost reduction (\label{}781 billion euros) in the \modCO{}, and around \label{}4\% cost reduction in the \modH{} and the \modHybrid{} (\label{}765 and \label{}764 billion euros, respectively).
+Total annual system costs in the \modBase{} are highest, totalling \label{}779 billion euros.
+The models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs, with \label{}3.1\% cost reduction (\label{}754~bn€) in the \modCO{}, and \label{}4.8\% (\label{}741~bn€) in the \modH{}, and \label{}5.2\% (\label{}738~bn€) in the the \modHybrid{}.
 
 % TODO: avoid to long paranthesis
-In all models, almost half of the system costs are spent on primary electricity production from wind and solar power (\label{}45\% or \label{}360 of \label{}799 billion euros total costs in the \modBase{}), with another \label{}12\% (\label{}100 billion euros) spent on hydro, nuclear and biomass.
-Costs for energy transport networks for electricity, gas, hydrogen and carbon are below \label{}12\% (\label{}100 billion euros), with the electricity grid making up more of the system costs than the other three networks combined.
-This is the case for all four models.
-Capital expenditures on carbonaceous fuel production are around \label{}3\% (\label{}25 billion euros) in each model, while electrolysis makes up around \label{}5\% (\label{}40 billion euros).
-Heat pump installations account for around \label{}8\% (60 billion euros) of total system costs.
+In all models, almost half of the system costs are spent on primary electricity production from wind and solar power (\label{}47\% or \label{}371~ billion euros in the \modBase{}), with another \label{}10\% (\label{}79 billion euros) spent on hydro, nuclear and biomass.
+Costs for energy transport networks for electricity, gas, hydrogen and carbon are below \label{}4\% (\label{}25 billion euros), with the electricity grid making up more of the system costs than the other three networks combined.
+This is not the case for the other models, where combined pipeline costs slightly surpass electricity transmission system costs (up to 20\% for the \hybridmodel).
+Capital expenditures on carbonaceous fuel production are around \label{}4\% (\label{}30 billion euros) in each model, while electrolysis makes up around \label{}4.7\% (\label{}37 billion euros).
+Heat pump installations account for around \label{}8\% (63 billion euros) of total system costs.
 
@@
 \subsection*{Transport systems unlock point\=/source carbon capture}
-
+i
 In comparison to the other three models, the \modBase{} reveals significant inefficiencies: On the one hand, regions with no access to sequestration sites (i.e., ``landlocked'' regions) underutilize CC potentials from process emissions, fossil gas and biomass.
@@
 By enabling the efficient transportation of carbon, hydrogen, or both, these models reduce the reliance on high-cost DAC, and source more carbon from low-cost point-source CC on biomass and gas in industrial applications (see Fig.~\ref{fig:balance_captured_carbon}).
-The \modH{} only captures \label{}200~Mt of \COtwo{} from DAC compared to \label{}400~Mt in the \modBase{}.
+The \modH{} only captures \label{}193~Mt of \COtwo{} from DAC compared to \label{}392~Mt in the \modBase{}.
 Instead, the \modH{} captures more carbon from process emissions as well as gas and biomass for industry.
-The \modCO{} uses only \label{}150~Mt from DAC, and instead also captures more carbon than the \modH{} from gas for industry and biomass CHPs.
-And finally, the \modHybrid{} captures only \label{}100~Mt from DAC, but slightly more carbon from all other CC technologies.
+The \modCO{} uses only \label{}154~Mt from DAC, and instead also captures more carbon than the \modH{} from gas for industry and biomass CHPs.
+And finally, the \modHybrid{} captures only \label{}112~Mt from DAC, but slightly more carbon from all other CC technologies.
 Also, the networks reduce the reliance on methanation in the model, reducing the overall demand for CC volumes.
@@
 For the network models, we can observe a correlation between CC share and capacity factors of the underlying technologies: Process emissions as well as gas and biomass for industry have high capacity factors (above 80\%), and CC shares close to 100\%, the only exception being gas for industry in the \modH{}.
-Biomass CHPs have a capacity factor of around 40\%, and CC shares of 50\% in the \modH{} and 95\% in the \modCO{}.
-SMR and gas CHPs have capacity factors below 20\%, and CC shares around 50\% and 0\%, respectively, across all network models.
+Biomass CHPs have a capacity factor of around 50\%, and CC shares of 50\% in the \modH{} and 97\% in the \modCO{}.
+SMR and gas CHPs have capacity factors around 15\%, and CC shares at nearly 0\% for the \baelinemodel.
+The capacity factor and CC share of SMR increases to around 20\% and 50-65\%, respectively, in the other models.
 Gas CHPs serve as peak load electricity production technology, and with very few operating hours, thus the high investment costs into CC applications are not efficient.
@@
 In transitioning from a net zero to a net emissions reduction target, a continuation of investment trends are perceived.
-Fig.~\ref{fig:net-negative_cost_bar} shows the net investment changes per technology group for the \Hgrid{} and \modHybrid{}s.
-The scenario imposes a net carbon removal of 460 Mt/a, with both models incurring a similar net cost increase - 95 billion €/a for the \modH{} and 96 billion €/a for the \modHybrid{}, balancing the earlier cost advantage of the \modHybrid{}.
+Fig.~\ref{fig:net-negative_cost_bar} shows the net investment changes per technology group for the \carbonmodel, the \Hgrid{} and \modHybrid{}s.
+The scenario imposes a net carbon removal of 460 Mt/a, with all models incurring a roughly similar net cost increase - 77 billion €/a for the \modCO{}, 81 billion €/a for the \modCO{}, and 78 billion €/a for the \modHybrid{}, balancing the earlier cost advantage of the \modHybrid{}.
 Both models invest relatively evenly in additional DAC, wind and heat pump systems, with DAC accounting for more than a third of the additional system costs.

### 2023-12-13 workflow/scripts/plot_captureshare_line.py

@@ -1,5 +1,6 @@
 import os
 import re
+import textwrap
 import numpy as np
 from pathlib import Path
 from common import (
@@ -97,15 +98,21 @@ handles, labels = ax.get_legend_handles_labels()
 labels = [l + " %" if re.match(r"^\d+$", l) else l for l in labels]
 
 pos = labels.index("Capacity Factor")
-handles.insert(pos, Patch(facecolor="none", edgecolor="none", alpha=0))
-labels.insert(pos, "")
+# add gap between legend parts
+# handles.insert(pos, Patch(facecolor="none", edgecolor="none", alpha=0))
+# labels.insert(pos, "")
+
+# remove 40 and 60 from legend
+for label in ["60 %", "40 %"]:
+    handles.pop(labels.index(label))
+    labels.remove(label)
 
 # Add the invisible handles and labels to the legend
 ax.legend(
     handles,
-    labels,
-    loc="center left",
-    bbox_to_anchor=(1.0, 0.5),
+    [textwrap.fill(label, 22) for label in labels],
+    loc="upper left",
+    bbox_to_anchor=(1, 1),
     frameon=False,
     ncol=1,
 )
@@ -121,7 +128,7 @@ plt.xticks(
 sns.despine()
 
 fig.savefig(snakemake.output.figure, dpi=300)
-data.to_csv(snakemake.output.table)
+data.round(3).to_csv(snakemake.output.table)
 
 # ALTERATIVE PLOT
 # fig, ax = plt.subplots(

### 2023-12-13 workflow/scripts/plot_cost_bar.py

@@ -14,7 +14,7 @@ region_alpha = 0.8
 
 if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
     snakemake = mock_snakemake(
-        "plot_cost_bar", ext="png", clusters=90, comparison="emission-reduction"
+        "plot_cost_bar", ext="png", clusters=90, comparison="default"
     )
 
 sns.set_theme(**snakemake.params["theme"])

### 2023-12-13 report/submissions/journal/paper/paper.tex
gather todos with christoph

@@
 \subsection*{Transport systems unlock point\=/source carbon capture}
-i
+
 In comparison to the other three models, the \modBase{} reveals significant inefficiencies: On the one hand, regions with no access to sequestration sites (i.e., ``landlocked'' regions) underutilize CC potentials from process emissions, fossil gas and biomass.
@@
 
+% TODO: smoother transition
 To find the reason for why the \Htwo{} network achieves lower system costs than the \COtwo{} network, it is crucial to understand the underlying mechanisms and geographical implications of the two transportation systems.
@@
 
+% TODO: split the  first sentence
+% TODO: add numbers: electrolyzers GW (Spain/Iberia, Germany), hydrogen prices (valleys, peaks)
 As we have also discussed above, the \modCO{} captures large amounts of low-cost carbon from bioenergy and industrial activities that were untapped in the \modBase{}, especially in Central and Eastern Europe, and almost none of which is used locally (see Fig.~\ref{fig:balance_map}, top left).
 A large part of the captured carbon is transported from Western Europe to the Iberian Peninsula, the British Isles and Denmark, where it is used along with low-cost electrolytic hydrogen to produce Fischer-Tropsch fuels and methanol.
-A second part of captured carbon is transported to methanation plants in Italy and Greece, where the synthetic methane output is fed into the well-established gas grid since Italy serves as a large gas import hub for Europe.
+A second part of the captured carbon is transported to methanation plants in Italy and Greece, where the synthetic methane produced is fed into the well-developed gas grid, as Italy serves as a major gas import hub for Europe.
 Finally, a third part of captured carbon is transported directly to sites in the North Sea, Baltic Sea and Mediterranean Sea where it is sequestered.
@@
 The model also emphasizes gas imports into Central Europe.
-Hydrogen prices still vary strongly across the continent, with low price ``valleys'' in regions with good renewable resources such as the Iberian Peninsula or Denmark, and high price ``peaks'' in Central Europe (see bottom left).
+Hydrogen prices vary strongly across the continent, with low price ``valleys'' in regions with good renewable resources such as the Iberian Peninsula or Denmark, and high price ``peaks'' in Central Europe (see bottom left).
 
 To understand the dynamics of the \modCO{}, we observe that the system locates CU in the regions where it can minimize costs for its material inputs, carbon and hydrogen.
-As the \COtwo{} network largely smoothes out carbon prices across the continent, CU is located in the hydrogen price valleys on the periphery of the continent.
-Similarly, the model locates CS in regions where it can minimize costs for its only material input, carbon.
+As the \COtwo{} network largely smoothes out carbon prices across regions, CU is located in the hydrogen price valleys on the periphery of the continent.
+Similarly, the model locates CS in regions where it can minimize the costs of the carbon to be sequestered.
 Thus, CS is located in the carbon price valleys among the subset of regions with access to offshore sequestration sites.
@@
 
-The \Htwo{} network on the other hand, transports low-cost hydrogen from regions with abundant renewable resources to regions with low-price carbon from abundant sources of fossil, biogenic and process emissions (see bottom right).
-The main transportation routes go from Spain and the United Kingdom to Central Europe to produce synfuels with locally captured carbon from point sources.
+The \Htwo{} network in the \modH{} on the other hand, transports low-cost hydrogen from regions with abundant renewable resources to regions with low-price carbon from abundant sources of fossil, biogenic and process emissions (see bottom right).
+The main transportation routes go from Spain and the United Kingdom to Central Europe to produce carbonaceous fuels with locally captured carbon from point sources.
 Also, the \Htwo{} network supplies exogenous \Htwo{} demand for industry in Central Europe at much lower costs than local electrolysis or SMR.
 
+% TODO: implement "onshore sequestration hub" (somewhere above) for "region with access to offshore sequestration site"
 To explain the dynamics of the \modH{}, we observe that the model locates CU in the carbon price valleys across Western, Central and Eastern Europe (see top right), as the \Htwo{} network smoothes out hydrogen prices across the entire continent (see bottom right).
@@
 Due to the lack of a \COtwo{} network the full CC potentials from point sources are exhausted in all offshore sequestration regions, and the carbon is then sequestered.
-The additionally needed carbon to reach 200~Mt of CS is captured via DAC in the offshore sequestration regions with the lowest costs for DAC, with the largest share deployed in Portugal (see top right).
-DAC is used for both CU and CS in this model, and for both purposes DAC is located in the regions with the lowest costs for electricity.
+The additionally needed carbon to reach 200~Mt of CS is captured via DAC in the regions with access to offshore sequestration sites with the lowest costs for DAC, with the largest share deployed in Portugal (see top right).
+DAC is also used for CU in this model, and for both purposes DAC is located in the regions with the lowest costs for electricity.
 
+% TODO: write more densely (less repetitions) --> main findings: electrolysis, point-source CC even; CU different; H2 Network: co-location CU/CC; CO2 network: co-location CU/ely
 In summary, both network models prefer distributed CC at point sources and electrolysis in regions with abundant renewable resources, choosing the lowest-cost production sites for each while trading off transport costs.
 However, differences between the two models are notable for the placement of CU, CS and DAC due to their distinct \COtwo{} and \Htwo{} price patterns.
-CU locations are more distributed in the \modH{}, while they cluster in a few pockets of low-cost \Htwo{} in the \modCO{}.
+CU locations are more distributed in the \modH{}, while they cluster in a few valleys of low-cost \Htwo{} in the \modCO{}.
 CS co-locates in the regions with the lowest \COtwo{} costs (among those with offshore sequestration potential), but these regions are very different between the models.
 DAC is placed in regions with the best renewable resources in both models, but it is used for only CU in the \modCO{} and for both CU and CS in the \modH{}.
-Finally, the system cost advantage of the \modH{} stems from being able to supply the exogenous hydrogen demands in Central Europe with low-cost hydrogen imports, while the \modCO{} is forced to deploy some electrolysis in regions with poor renewables resources as well as additional gas infrastructure in Central Europe.
+Finally, the system cost advantage of the \modH{} stems from being able to supply the exogenous hydrogen demands in Central Europe with low-cost hydrogen imports, while the \modCO{} is forced to deploy electrolysis in regions with poor renewables resources as well as additional gas infrastructure in Central Europe.
 
@@
 
+% TODO: smoother transition, incl. end of 3.2
+% TODO: inlcude in opener: both networks, lower system costs; motivation: what is the advantage vs both others? other advantages (more resilient)/disadvantages (not focusing on one grid)
 The \modHybrid{} combines the advantages of hydrogen and carbon networks.
+Fig.~\ref{fig:capacity_maps} shows the average production, consumption and transport of the two networks in the \modHybrid{}.
+
 While the topology of the \Htwo{} network roughly corresponds to that of the \modH{}, the \COtwo{} network plays more of a complementary role.
-Fig.~\ref{fig:capacity_maps} shows the average production, consumption and transport of the two networks in the \modHybrid{}.
-Note that despite similar overall capacities in terms of weight moved, the investment in the hydrogen infrastructure is four times the investment in the carbon infrastructure.
 
-Similar to the \modH{}, the \modHybrid{} takes advantage of a large hydrogen network that enables hydrogen to be transported from centralized production sites in western regions such as Spain to decentralized FT production sites across the continent.
-However, in the \modHybrid{} hydrogen in the UK is used locally to produce synfuels, and less hydrogen is transported to Central Europe.
+% TODO: mention investment difference in network expansion with appendix plot (The investment in the hydrogen network is four times the investment in the carbon network.)
+% TODO: rewrite: mostly short (1-2 regions) linear (no cluster/mashed network) routes towards sequestration sites (for purpose of seq.); short 1-region transports to bridge local carbon price diffs, for purpose of CU.
+Similar to the \modH{}, the \modHybrid{} takes advantage of a large hydrogen network that enables hydrogen to be transported from centralized production sites in Western regions such as Spain to decentralized CU production sites close to CC point sources across the continent.
 The carbon network supplements this with smaller clusters of networks.
@@
 A more detailed of differences between the \modH{} and the \modHybrid{} can be found in Fig.~\ref{fig:cost_map_difference}.
+
+% TODO: add interpretative paragraph: no CU where low-price ely, H2 is imported.
+% Both networks available to smooth prices; CS draws in CC from adjacent regions; CU everywhere locally, except the CS sites; DAC centered on (very few) regions with optimal RES (and Italy/Greece for methanation)
 

### 2023-12-14 report/submissions/journal/paper/paper.tex
follow up
paper: address todos in intro plots: minor changes

@@
 \title{Assessing \Htwo{} and \COtwo{} Network Strategies in Europe}
+\title{Competition and Synergies of \Htwo{} and \COtwo{} Networks in Europe}
+
 % \title{Synthetic fuels in Europe: Transport Hydrogen to Carbon, or Carbon to Hydrogen?}
@@
 The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors.
-While management of both \COtwo{} and \Htwo{}, also refered to as carbon and hydrogen, is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructure effectively interacts with carbon management technologies, including carbon capture, transport, storage, utilization, and sequestration.
+While management of both \COtwo{} and \Htwo{}, also refered to as carbon and hydrogen, is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructure effectively interacts with carbon management technologies, including of carbon capture, transport, storage, utilization, and sequestration.
 Hydrogen offers an efficient way to transport and store energy over long distances and timeframes, and can be produced from electrolysis or steam methane reforming (SMR), for example.
 Carbon can be stored in geological formations to permanently extract it from the atmosphere, a process known as carbon sequestration (CS).
-Additionally, carbon can be combined with hydrogen to produce carbonaceous fuels needed for aviation (synthetic kerosene from Fischer-Tropsch synthesis), shipping (synthetic methanol) and industrial feedstocks (e.g., synthetic methane).
+Additionally, carbon can be combined with hydrogen to produce carbonaceous fuels as for example needed for aviation in form of synthetic kerosene from Fischer-Tropsch synthesis, for shipping in form of synthetic methanol and for industrial feedstocks through synthetic methane.
 From hereon, we jointly refer to the production of these carbonaceous fuels as Carbon Utilization (CU).
 Carbon can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, or synthetic carbonaceous fuels through point-source carbon capture (CC) techniques, or harvested from the atmosphere using direct air capture (DAC).
-In combination, \COtwo{} and \Htwo{} networks are crucial to enable these carbon management strategies for net-zero economies.
-% TODO: integrate definition of Carbon Management (CM) into last sentence.
-% Cx: CU, CS, CT, CC, Cstorage --> CO2 + H2 networks are crucial to carry out CM strategies
-% Suggestion:
+Both \COtwo{} and \Htwo{} networks are likely to play a crucial role in the cost-effective integration of these technologies to enable net-zero economies.
+% [TODO: integrate definition of Carbon Management (CM) into last sentence.]. COMMENT: CM was already mention earlier, don't repeat it instead changed framing of last sentence
 
-% TODO: check definition of carbonaceous fuels for "missing" (relevant) fuel types
-
-% TODO: boil this down to one sentences and integrate it to the previous paragraph or move it two the literature part.
-% Suggestion: We argue that it is crucial to consider CC, CU and CS as separate technologies, rather than grouping them under the often-used but misleading umbrella term "CCUS", to allow for more accurate optimization and analysis of each function in the carbon management system.
-% Suggestion: We argue that it is crucial to consider CC, CU and CS as separate technologies, rather than grouping them under the often-used but misleading umbrella term ``carbon capture, utilization and storage'' (CCUS), to allow for independent optimization and adequate representation of each function in the carbon management system.
-It is important to differentiate between the functions of carbon capture, carbon sequestration and carbon utilization.
-While both academic literature and common language mostly use the term ``carbon capture, utilization and storage'' (CCUS), or similar, to refer to all three functions, we argue that this is misleading.
-There is inherent benefit in separating these three functions geographically (linked through carbon transport), temporally (linked through carbon storage), and in terms of economic actors performing each function.
-CC should be performed at the place and with the technology that can capture carbon at the lowest cost (including a potential transport premium).
-CS should take place where sequestration potentials exist and carbon is available at the lowest cost, and CU where the combination of hydrogen and carbon inputs is available at lowest cost.
-By employing this separation in our model, we allow for a nuanced optimization and detailed analysis of the carbon management system.
+% TODO: Move to better place in paper
+We argue that it is crucial to consider CC, CU and CS as separate technologies, rather than grouping them under the often-used but misleading umbrella term ``carbon capture, utilization and storage'' (CCUS), to allow for independent optimization and adequate representation of each function in the carbon management system.
 
 Recently, policymakers and industry in Europe have started developing hydrogen and carbon management strategies~\cite{GermanyDevelopingStrategy2023,CarbonManagementStrategie}, planning infrastructure components~\cite{CONetz}, and committing on the first carbon utilization projects~\cite{EFuelsPilotPlant2022,OrstedAssumesFull,GROUNDBREAKINGEFUELPRODUCTION,DLREfuelsDLR}.
-With the European Union's goal of achieving climate neutrality by 2050, a wide range of programs, funding models and initiatives have been established in this regard~\cite{eu2023netzero,europeangreendeal,europeaninnovationfund}.
-% TODO: mention policy ambition for carbon and hydrogen strategies separately in two sentences.
-Collaborative industry initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen Infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier, and some gas pipelines are already repurposed to transport hydrogen~\cite{RohrFreiFuer}.
+With the European Union's goal of achieving climate neutrality by 2050, a wide range of programs, funding models and initiatives have been established with dedicated instruments supporting both the ramp-up of the hydrogen and carbon economy~\cite{eu2023netzero,europeangreendeal,europeaninnovationfund}.
+Collaborative industry initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen Infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier.
+Some natural gas pipelines are already repurposed to transport hydrogen~\cite{RohrFreiFuer}.
 Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synfuels on the one hand and offer \COtwo{} offtake on the other hand.
 The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a transport and sequestration capacity of 1.5~Mt \COtwo{} per year to be operative in 2024, expanding to a targeted scale of 5~Mt per year of sequestration by 2030.
-% TODO: frame next two sentences to be more market oriented
-% TODO: rather cite europeancommissionEuropeanCO2Storage
-For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.7~Gt of carbon capture from point sources per year, which represent roughly 50\% of all current emissions.
-In combination with large sequestration potentials as stated in~\cite{weiProposedGlobalLayout2021}, this highlights the vast potential for carbon capture.
-% TODO: optionally find better place for this
-To this end, the Clean Air Task Force underlines the importance of a carbon transport system in Europe to facilitate the carbon economy~\cite{lockwoodEuropeanStrategyCarbon}.
-Considering carbon pipeline as a technological option is a reasonable model assumption since \COtwo{} pipelines are a mature technology, given their wide-spread installations in the United States and Canada for enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
+The market potential for carbon capture and sequestration in Europe is underscored by the Capture Map's estimated potential~\cite{ToolsGreenTransition} of 1.7 Gt carbon annually from point sources, accounting for about half of the continent's emissions, as well as estimated sequestration capacity potentials of up to 3~Gt/a~\cite{europeancommissionEuropeanCO2Storage}.
+However, it's important to note that carbon sequestration technology, unlike carbon capture, is not yet fully mature and its full potential and implications remain to be explored in depth.
+To advance the carbon economy, the Clean Air Task Force emphasizes the importance of developing a carbon transport system in Europe alongside a hydrogen network~\cite{lockwoodEuropeanStrategyCarbon}.
+Carbon pipelines, considered a mature technology, have seen widespread installations in the United States and Canada, primarily for enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
 
-% TODO: explain why misalignment takes place -> exogenous factors renewable potentials, industrial clusters with h2 demand, co2 point sources, sequestration sites,
 However, up to this point, it remains unclear how the two transport systems of hydrogen and carbon may complement or replace each other.
-Both can bridge the misalignment of sources (electrolysis, SMR, point-source CC, DAC) and sinks (CU, CS, industrial hydrogen demand) for carbon and hydrogen.
+Both can bridge the misalignment of sources and sinks for carbon and hydrogen which may well be determined by exogenous factors.
 A hydrogen network can supply regions with geographically fixed hydrogen demand, such as for steelmaking, with hydrogen from regions with the best renewable resources as well as enable CU at the site of point-source CC.
@@
 We investigate the competitive transport dynamics between \COtwo{} and \Htwo{} networks.
-This is done by contrasting models that feature only one of these networks, neither, or both combined.
-We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon and hydrogen networks varies given net-zero or net-negative emissions targets.
+This is done by contrasting the following four models that feature neither one of these networks, only one, or both combined:
 
-We define two \COtwo{} target scenarios:
-\begin{itemize}
-    \item[] \textit{Net\=/Zero}: aligning with the EU's 2050 emission targets. Limiting carbon sequestration to 200~Mt/a. Also denoted as NZ.
-    \item[] \textit{Net\=/Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \COtwo{} annually. Limiting carbon sequestration to 660~Mt/a. Also denoted as NN.
-\end{itemize}
-Unless otherwise mentioned, our reference is the Net-Zero scenario.
-
-Beyond \COtwo{} targets, we consider four expansion scenarios, referred to as models:
 \begin{itemize}
@@
 
-% TODO add two sentences about the used methods for transition
+Anticipating with tight \COtwo{} emission targets and limited sequestration assumptions, the models are exposed to the following two scenarios:
+\begin{itemize}
+    \item[] \textit{Net\=/Zero}: aligning with the EU's 2050 emission targets. Limiting carbon sequestration to 200~Mt/a. Also denoted as NZ.
+    \item[] \textit{Net\=/Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \COtwo{} annually. Limiting carbon sequestration to 660~Mt/a. Also denoted as NN.
+\end{itemize}
+
+We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon and hydrogen networks varies given net-zero or net-negative emissions targets.
+
+In the following we show how a separate investment into no or only one of the transport technologies leads to inefficient market mechanisms.
+We do so, by first focusing on the Net-Zero scenario and a detailed description of the different systemic implications of the model.
+A analysis of the .
 
@@
 While the \modBase{} has neither a \COtwo{} nor an \Htwo{} network, the \modHybrid{} is allowed to expand both.
-``Gas Infrastructure'' combines gas facilities for power and heat production, ``\COtwo{} Infrastructure'' combines transport, storage and sequestration, and ``\Htwo{} Infrastructure'' combines transport and storage.
+``Gas Infrastructure'' combines gas facilities for power and heat production, ``\COtwo{} Infrastructure'' and ``\Htwo{} Infrastructure'' combine transport and storage for each carrier.
 ``Carbon Capture at Point Sources'' combines all technologies with integrated carbon capture, including the cost of the main facility (e.g., CHP unit) and the carbon capture application.}
@@
 Total annual system costs in the \modBase{} are highest, totalling \label{}779 billion euros.
-The models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs, with \label{}3.1\% cost reduction (\label{}754~bn€) in the \modCO{}, and \label{}4.8\% (\label{}741~bn€) in the \modH{}, and \label{}5.2\% (\label{}738~bn€) in the the \modHybrid{}.
+The models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs, with \label{}3.1\% cost reduction (\label{}754~bn€) in the \modCO{}, and \label{}4.8\% (\label{}742~bn€) in the \modH{}, and \label{}5.2\% (\label{}738~bn€) in the the \modHybrid{}.
 
@@
 % Both networks available to smooth prices; CS draws in CC from adjacent regions; CU everywhere locally, except the CS sites; DAC centered on (very few) regions with optimal RES (and Italy/Greece for methanation)
+% compare to H2 Grid model: küstennahe regionen mit longest distance to good RES(?) anzapfen, dort kein CU, CU jewtzt im Hinterland
 
@@
 
-In transitioning from a net zero to a net emissions reduction target, a continuation of investment trends are perceived.
+% TODO: start with explanation Fig 6: similarities b/w all models: sequestration, more DAC (=+wind, solar, heat pump, battery), point-source CC (+bioenergy); diff: level of DAC vs level of point-source CC, co2 network, h2 network
+
+% carbon:
+% - fundamentally new network layout
+% - CU in spain is now supplied by carbon from DAC
+% - carbon in Central Europe is mainly transported to seq sites
+% - stranded investment in carbon network of 3bn
+
+% hydrogen:
+% - no significant changes in strategies
+
+% hybrid:
+% - more CU in spain
+% - less cu in central europe
+% - more sequestration in central europe and Portugal
+% - less h2 import from spain to central european
+% - mostly additive investments, except for some CU plants near the coast (of 4.3bn) and h2 network from Spain to Central Europe (2bn)
+
+% TODO: add operation plots into appendix all models (hydrogen and carbon)
+
+\ac{}In transitioning from a net zero to a net emissions reduction target, a continuation of investment trends are perceived.
 Fig.~\ref{fig:net-negative_cost_bar} shows the net investment changes per technology group for the \carbonmodel, the \Hgrid{} and \modHybrid{}s.
-The scenario imposes a net carbon removal of 460 Mt/a, with all models incurring a roughly similar net cost increase - 77 billion €/a for the \modCO{}, 81 billion €/a for the \modCO{}, and 78 billion €/a for the \modHybrid{}, balancing the earlier cost advantage of the \modHybrid{}.
+The scenario imposes a net carbon removal of 460 Mt/a, with all models incurring a roughly similar net cost increase - 77 billion €/a for the \modCO{}, 81 billion €/a for the \modH{}, and 78 billion €/a for the \modHybrid{}.
 Both models invest relatively evenly in additional DAC, wind and heat pump systems, with DAC accounting for more than a third of the additional system costs.
 These remove carbon directly at sequestration sites at the coast, in particular in Portugal and UK, together with heat and power from new heat pumps at site and new solar and wind installation in the vicinity.
-More heat pumps, solar and wind installations across all Central and Eastern European countries as well as biogas-to-gas installations compensate for the reduction of fossil gas based heating and powering.
-Reduced usage of gas in general lower the investments in gas infrastructure in both models.
+
 Fundamental difference in investment strategies occur in the carbon capture sector: In contrast to the \modH{}, the \modHybrid{} continues the strategy of sequestering carbon captured and transported from bio-energetic inputs, expanding the carbon transport routes from Central and Eastern Europe to the near shores.
-At the same time, carbonaceous fuel production in Central Europe partially moves to Spain and UK, where it uses carbon from new DAC facilities without the installation of a \COtwo{} network.
-At the same time, solar based electrolysis in France, Nothern Italy and Switzerland expands requiring more solar capacities but less hydrogen imports from the Iberian Peninsula.
+At the same time, carbonaceous fuel production in Central Europe partially moves to Spain, where it uses carbon from new DAC facilities without the installation of a \COtwo{} network.
+At the same time, less hydrogen imports from the Iberian Peninsula to Central Europe is needed due to less CU.
 The \modH{} on the other hand, expands decentralized CU systems from the Net-Zero scenario, which are now increasingly supplied with carbon from biogas-to-gas facilities, showing up in both ``Bioenergy'' and ``Carbon Capture at Point Sources'' cost-contributions.

### 2023-12-14 workflow/scripts/plot_cost_bar_compare_diff.py

@@ -53,6 +53,8 @@ unit = "bn€/a"
 
 groups = snakemake.config["plotting"]["technology_groups"]
 groups = {v: groups[k] for k, v in n.carriers.nice_name.items()}
+# keep sequestration separate
+groups["CO$_2$ Sequestration"] = "CO$_2$ Sequestration"
 grouped = diff.mul(-1).groupby(groups).sum().div(norm)
 # grouped = grouped[grouped.max(axis=1) > 5e4]
 colors = snakemake.config["plotting"]["technology_group_colors"]

### 2023-12-14 report/submissions/journal/paper/paper.tex
paper: rewrite section 3.3 (Hybrid model)

@@
 \newcommand{\hydrogengrid}{H$_2$\=/Grid}
-\newcommand{\baelinemodel}{Baseline model}
+\newcommand{\baselinemodel}{Baseline model}
 \newcommand{\carbonmodel}{CO$_2$\=/Grid model}
@@
 Some natural gas pipelines are already repurposed to transport hydrogen~\cite{RohrFreiFuer}.
-Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synfuels on the one hand and offer \COtwo{} offtake on the other hand.
+Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synthetic fuels on the one hand and offer \COtwo{} offtake on the other hand.
 The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a transport and sequestration capacity of 1.5~Mt \COtwo{} per year to be operative in 2024, expanding to a targeted scale of 5~Mt per year of sequestration by 2030.
@@
 In this paper, we present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies.
-The study is conducted using the PyPSA-Eur energy system model and encompasses all relevant energy sectors.
-We investigate the competitive transport dynamics between \COtwo{} and \Htwo{} networks.
-This is done by contrasting the following four models that feature neither one of these networks, only one, or both combined:
-
+The study is conducted using the PyPSA-Eur energy system model which optimizes operation and investments in all relevant energy sectors.
+To investigate the competing transport dynamics between \COtwo{} and \Htwo{} networks, we contrast the following four models that have none of these networks, only one, or both combined:
+%
 \begin{itemize}
@@
 \end{itemize}
-
-Anticipating with tight \COtwo{} emission targets and limited sequestration assumptions, the models are exposed to the following two scenarios:
+%
+In anticipation of the future \carbon{} emission targets, the models are subjected to the following two scenarios:
 \begin{itemize}
-    \item[] \textit{Net\=/Zero}: aligning with the EU's 2050 emission targets. Limiting carbon sequestration to 200~Mt/a. Also denoted as NZ.
+    \item[] \textit{Net\=/Zero}: aligning with the EU's 2050 emission targets. Limiting carbon sequestration to 200~Mt/a. This scenario is also denoted as NZ.
     \item[] \textit{Net\=/Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \COtwo{} annually. Limiting carbon sequestration to 660~Mt/a. Also denoted as NN.
@@
 
-We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon and hydrogen networks varies given net-zero or net-negative emissions targets.
-
-In the following we show how a separate investment into no or only one of the transport technologies leads to inefficient market mechanisms.
-We do so, by first focusing on the Net-Zero scenario and a detailed description of the different systemic implications of the model.
-A analysis of the .
+We analyze how the deployment of carbon and hydrogen networks affects different decarbonization strategies in the models and scenarios and how investments in none or only one of the transport technologies lead to inefficient market layouts.
+We do this by first focusing on the net-zero scenario and a detailed description of the different technological and geographical impacts of the transportation systems.
+Subsequently our analysis demonstrates the extent to which different deployment strategies remain robust under a tightened emissions target in the Net-Negative scenario.
 
@@
 
-Total annual system costs in the \modBase{} are highest, totalling \label{}779 billion euros.
+Total annual system costs in the \modBase{} are highest, totalling \label{}779~bn€.
 The models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs, with \label{}3.1\% cost reduction (\label{}754~bn€) in the \modCO{}, and \label{}4.8\% (\label{}742~bn€) in the \modH{}, and \label{}5.2\% (\label{}738~bn€) in the the \modHybrid{}.
 
-% TODO: avoid to long paranthesis
-In all models, almost half of the system costs are spent on primary electricity production from wind and solar power (\label{}47\% or \label{}371~ billion euros in the \modBase{}), with another \label{}10\% (\label{}79 billion euros) spent on hydro, nuclear and biomass.
-Costs for energy transport networks for electricity, gas, hydrogen and carbon are below \label{}4\% (\label{}25 billion euros), with the electricity grid making up more of the system costs than the other three networks combined.
+In all models, almost half of the system costs are spent on primary electricity production from wind and solar power, with around another \label{}10\% spent on hydro, nuclear and biomass.
+For the \baselinemodel, transmission system costs for electricity, gas, hydrogen and carbon are below \label{}4\% (\label{}25~bn€), with the electricity grid making up more of the system costs than the other three networks combined.
 This is not the case for the other models, where combined pipeline costs slightly surpass electricity transmission system costs (up to 20\% for the \hybridmodel).
-Capital expenditures on carbonaceous fuel production are around \label{}4\% (\label{}30 billion euros) in each model, while electrolysis makes up around \label{}4.7\% (\label{}37 billion euros).
-Heat pump installations account for around \label{}8\% (63 billion euros) of total system costs.
+Capital expenditures on carbonaceous fuel production are around \label{}4\% (\label{}30~bn€) in each model, while electrolysis makes up around \label{}4.7\% (\label{}37~bn€).
+Heat pump installations account for around \label{}8\% of total system costs (64~bn€ in the \baselinemodel).
 
@@
 And finally, the \modHybrid{} captures only \label{}112~Mt from DAC, but slightly more carbon from all other CC technologies.
-Also, the networks reduce the reliance on methanation in the model, reducing the overall demand for CC volumes.
-Another knock-on effect of the reduced reliance on DAC is reduced endogenous demand for renewables to power DAC facilities.
-% TODO: make sure FT and methanolisation is properly explained in methods section as almost exogenously fixed products
+Also, the transport of \carbon{} and \hydrogen{} reduce the reliance on methanation and therefore the overall demand for CC volumes in the model.
+Another knock-on effect of less reliance on DAC is its reduced endogenous demand for dedicated renewables as well as heat pumps and biomass CHPs to supply power and heat.
 
@@
 
-% TODO: either explain what network models mean or use another framing
-% TODO: extend first sentence
-The increase in captured carbon from point sources in the network models is mostly due to CC infrastructure being deployed on industrial or power plant facilities that are present in all models, as opposed to the deployment of additional electricity generating capacity for the purpose of carbon capture.
+The increase in captured carbon from point sources in the models with \carbon{} and/or \hydrogen{} transport is more related to an upgrading of existing industrial or power plant facilities to capture carbon rather than a shift in technologies.
 This is evident when analyzing the capture share of the different CC technologies (see Fig.~\ref{fig:captureshare_line}).
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
 In the \modBase{}, the capture share is highest for biogas-to-gas facilities, followed by process emissions, biomass for industry, biomass CHPs, gas for industry, SMR and gas CHPs.
-For the network models, we can observe a correlation between CC share and capacity factors of the underlying technologies: Process emissions as well as gas and biomass for industry have high capacity factors (above 80\%), and CC shares close to 100\%, the only exception being gas for industry in the \modH{}.
+For models with additional transport systems, we observe a correlation between CC share and capacity factors of the underlying technologies: Process emissions as well as gas and biomass for industry have high capacity factors (above 80\%), and CC shares close to 100\%, the only exception being gas for industry in the \modH{}.
 Biomass CHPs have a capacity factor of around 50\%, and CC shares of 50\% in the \modH{} and 97\% in the \modCO{}.
-SMR and gas CHPs have capacity factors around 15\%, and CC shares at nearly 0\% for the \baelinemodel.
+SMR and gas CHPs have capacity factors around 15\%, and CC shares at nearly 0\% for the \baselinemodel.
 The capacity factor and CC share of SMR increases to around 20\% and 50-65\%, respectively, in the other models.
@@
 
-% TODO: change opener for next section: co2gird has less DAC and more CC at point sources, however h2grid is cheaper
-In summary, the implementation of \COtwo{} and \Htwo{} networks is instrumental in unlocking low-cost CC potentials at geographically fixed point sources across the continent.
-However, despite the \modCO{} demonstrating higher absolute amounts of CC from point sources and higher CC shares across all point-source CC technologies, the \modH{} achieves lower total system costs (see Fig.~\ref{fig:cost_bar}).
-In the following section, we will explore the reasons for this.
+In summary, the implementation of \COtwo{} and \Htwo{} networks is instrumental in unlocking CC potentials and reducing system costs.
+However, despite the \modCO{} demonstrating higher amounts and shares of CC from point sources and therefore less DAC deployment, the \modH{} achieves even  lower total system costs (see Fig.~\ref{fig:cost_bar}).
 
@@
 
-% TODO: smoother transition
 To find the reason for why the \Htwo{} network achieves lower system costs than the \COtwo{} network, it is crucial to understand the underlying mechanisms and geographical implications of the two transportation systems.
@@
 
-% TODO: split the  first sentence
 % TODO: add numbers: electrolyzers GW (Spain/Iberia, Germany), hydrogen prices (valleys, peaks)
-As we have also discussed above, the \modCO{} captures large amounts of low-cost carbon from bioenergy and industrial activities that were untapped in the \modBase{}, especially in Central and Eastern Europe, and almost none of which is used locally (see Fig.~\ref{fig:balance_map}, top left).
-A large part of the captured carbon is transported from Western Europe to the Iberian Peninsula, the British Isles and Denmark, where it is used along with low-cost electrolytic hydrogen to produce Fischer-Tropsch fuels and methanol.
-A second part of the captured carbon is transported to methanation plants in Italy and Greece, where the synthetic methane produced is fed into the well-developed gas grid, as Italy serves as a major gas import hub for Europe.
+As we have also discussed above, the \modCO{} captures large amounts of carbon from bioenergy and industrial activities that were untapped in the \modBase{}.
+In particular, the carbon capture takes place in Central and Eastern Europe at relatively low cost, and almost none of the carbon is used locally (see Fig.~\ref{fig:balance_map}, top left).
+A large part of the captured carbon is transported from Western Europe to the Iberian Peninsula (70~Mt/h), the British Isles (66~Mt/h) and Denmark (25~Mt/h), where it is used along with low-cost electrolytic hydrogen to produce Fischer-Tropsch fuels and methanol.
+A second part of the captured carbon is transported to methanation plants in Italy and Greece (25~Mt/h), where the synthetic methane produced is fed into the well-developed gas grid, as Italy serves as a major gas import hub for Europe.
 Finally, a third part of captured carbon is transported directly to sites in the North Sea, Baltic Sea and Mediterranean Sea where it is sequestered.
-In absence of an \Htwo{} network, the system places most electrolysis in regions with the best renewable resources to supply local CU, and smaller amounts of electrolyzers where hydrogen is needed for industrial processes across Central Europe.
+In absence of an \Htwo{} network, the system places most electrolysis in regions with the best renewable resources to supply local CU ($\sim$1100~GW$_\text{el}$), and smaller amounts of electrolyzers where hydrogen is needed for industrial processes across Central Europe ($\sim$75~GW$_\text{el}$).
 The model also emphasizes gas imports into Central Europe.
-Hydrogen prices vary strongly across the continent, with low price ``valleys'' in regions with good renewable resources such as the Iberian Peninsula or Denmark, and high price ``peaks'' in Central Europe (see bottom left).
+Hydrogen prices vary strongly across the continent, with low price ``valleys'' in regions with good renewable resources such as the Iberian Peninsula or Denmark with around 60~€/MWh, and high price ``peaks'' in Central Europe with around 95~€/MWh (see bottom left).
 
@@
 Thus, CS is located in the carbon price valleys among the subset of regions with access to offshore sequestration sites.
-Due to the \COtwo{} network these valleys are not strongly pronounced, but still prevalent and impact CS locations.
-DAC is only used for CU and thus co-located in regions with a mix of low electricity costs (for running DAC) and low hydrogen costs (where CU is located, see above).
+Due to the \COtwo{} network these low-price regions are not strongly pronounced, but still prevalent and impact CS locations.
+DAC is only used for CU and thus co-located in regions with a mix of low electricity costs (for running DAC) and low hydrogen costs.
 
 The \Htwo{} network in the \modH{} on the other hand, transports low-cost hydrogen from regions with abundant renewable resources to regions with low-price carbon from abundant sources of fossil, biogenic and process emissions (see bottom right).
-The main transportation routes go from Spain and the United Kingdom to Central Europe to produce carbonaceous fuels with locally captured carbon from point sources.
-Also, the \Htwo{} network supplies exogenous \Htwo{} demand for industry in Central Europe at much lower costs than local electrolysis or SMR.
+The main transportation routes go from Spain (430~GW) and the United Kingdom (155~GW) to Central Europe to produce carbonaceous fuels with locally captured carbon from point sources.
+Also, the \Htwo{} network supplies exogenous \Htwo{} demand for industry in Central Europe (22~GW in total) at much lower costs than local electrolysis or SMR.
 
-% TODO: implement "onshore sequestration hub" (somewhere above) for "region with access to offshore sequestration site"
 To explain the dynamics of the \modH{}, we observe that the model locates CU in the carbon price valleys across Western, Central and Eastern Europe (see top right), as the \Htwo{} network smoothes out hydrogen prices across the entire continent (see bottom right).
@@
 
-% TODO: write more densely (less repetitions) --> main findings: electrolysis, point-source CC even; CU different; H2 Network: co-location CU/CC; CO2 network: co-location CU/ely
-In summary, both network models prefer distributed CC at point sources and electrolysis in regions with abundant renewable resources, choosing the lowest-cost production sites for each while trading off transport costs.
-However, differences between the two models are notable for the placement of CU, CS and DAC due to their distinct \COtwo{} and \Htwo{} price patterns.
+In summary, both network models prefer CC from point sources distributed across the continent as well as electrolysis concentrated in regions with abundant renewable resources.
+The two network approaches reveal fundamentally different strategies to integrate these low-cost resources into the system: The \carbonmodel{} transports and sequesters low-cost carbon from distributed point sources and co-locates CU with low-cost electrolysis and DAC.
+The \hydrogenmodel{} distributes CU alongside with carbon point sources to which low-cost hydrogen from regions with abundant renewable resources is transported.
+%
+% However, differences between the two models are notable for the placement of CU, CS and DAC due to their distinct \COtwo{} and \Htwo{} price patterns.
 CU locations are more distributed in the \modH{}, while they cluster in a few valleys of low-cost \Htwo{} in the \modCO{}.
@@
 % TODO: smoother transition, incl. end of 3.2
-% TODO: inlcude in opener: both networks, lower system costs; motivation: what is the advantage vs both others? other advantages (more resilient)/disadvantages (not focusing on one grid)
+% TODO: add network investment costs plot to Appendix. Mention investment H2 > CO2. (The investment in the hydrogen network is four times the investment in the carbon network.)
 The \modHybrid{} combines the advantages of hydrogen and carbon networks.
-Fig.~\ref{fig:capacity_maps} shows the average production, consumption and transport of the two networks in the \modHybrid{}.
+Fig.~\ref{fig:cost_bar} shows that the \modHybrid{} achieves a cost reduction of \label{}5.2\% (\label{}41 billion euros) compared to the \modBase{}, which is lower than the cost reductions of the \COgrid{} and \modH{}.
+Fig.~\ref{fig:capacity_maps} shows the average production, consumption and transport of carbon and hydrogen in the \modHybrid{}.
+While the topology of the \Htwo{} network roughly corresponds to that of the \modH{}, the \COtwo{} network plays more of a complementary role.
+In this section, we explore the advantages and disadvantages compared to the other models.
 
-While the topology of the \Htwo{} network roughly corresponds to that of the \modH{}, the \COtwo{} network plays more of a complementary role.
+Similar to the \modH{}, the \modHybrid{} transports hydrogen from concentrated electrolysis sites in Western regions (Iberian Peninsula and British Isles) to distributed CU production sites across the continent.
+For the \COtwo{} network, this model employs short and direct pipeline routes, typically spanning one or two segments, to transport carbon from inland CC point sources to perform CS at nearby sequestration sites (e.g., Northern Germany, Italy and Greece).
+Additionally, the \COtwo{} network utilizes single-segment network sections to equalize local variations in carbon demand and supply to produce CU, thereby also further converging carbon prices.
 
-% TODO: mention investment difference in network expansion with appendix plot (The investment in the hydrogen network is four times the investment in the carbon network.)
-% TODO: rewrite: mostly short (1-2 regions) linear (no cluster/mashed network) routes towards sequestration sites (for purpose of seq.); short 1-region transports to bridge local carbon price diffs, for purpose of CU.
-Similar to the \modH{}, the \modHybrid{} takes advantage of a large hydrogen network that enables hydrogen to be transported from centralized production sites in Western regions such as Spain to decentralized CU production sites close to CC point sources across the continent.
-The carbon network supplements this with smaller clusters of networks.
-In addition to a larger \COtwo{} network cluster in Central Europe, which transports carbon to the North Sea, the system is building three large linear carbon routes.
-These transport carbon from Romania and Bulgaria to Greece, from northern Italy to central Italy and from Hungary, the Czech Republic and Poland to the Baltic Sea.
-These enable the system to transport carbon available at low cost from inland biomass sources and sequester it (see Fig.~\ref{fig:captureshare_line}).
-This reduces the need for DAC in the system and the sequestration sites are shifted to the end points of the carbon networks.
-Second-order effects can be seen in the shift in FT production compared to the \modH{}.
-Regions that have both good renewable resources and sequestration potential could now forego sequestration and use local process emissions for FT production, as can be observed in Ireland.
-A notable aspect of the hybrid model is the lack of overlap in the hydrogen and carbon network topologies.
-This finding is tightly linked to the fact of competing functionalities to offtake carbon emissions.
-A more detailed of differences between the \modH{} and the \modHybrid{} can be found in Fig.~\ref{fig:cost_map_difference}.
+To understand the dynamics of the \modHybrid{} model, we observe that the model locates CU everywere, except for a few coastal regions that perform CS, or that supply their locally captured carbon to a neigbouring sequestration region.
+The model uses the \COtwo{} network to transport carbon over short distances to the CS sites, and the \Htwo{} network to transport hydrogen over longer distances from the lowest-cost hydrogen regions to CU sites across the continent.
+DAC is deployed to supply carbon for the remaining CU needed by the model (mostly in Southern Spain).
 
-% TODO: add interpretative paragraph: no CU where low-price ely, H2 is imported.
-% Both networks available to smooth prices; CS draws in CC from adjacent regions; CU everywhere locally, except the CS sites; DAC centered on (very few) regions with optimal RES (and Italy/Greece for methanation)
-% compare to H2 Grid model: küstennahe regionen mit longest distance to good RES(?) anzapfen, dort kein CU, CU jewtzt im Hinterland
-
-Despite only a further reduction of \label{}0.4 percentage points (\label{}3 billion euros) in total system costs compared to the \modH{}, the \modHybrid{} brings several advantages.
-These include a broader range of technologies contributing to its robustness, reduced reliance on DAC, which may be costlier than anticipated in the cost projections, and less land use for wind and solar due to the decreased necessity for DAC.
-The increase used of biomass with carbon capture, transport and sequestration leads to slight cost advantages in comparison to the \modH{}.
+Despite only a further reduction of \label{}0.4 percentage points (\label{}3~bn€) in total system costs compared to the \modH{}, the \modHybrid{} brings several advantages: (1) The \modHybrid{} uses both \COtwo{} and \Htwo{} networks to build a more robust energy system; (2) it relies even less on DAC than the other models, which as a technology may be costlier than anticipated in the cost projections; and (3) it requires a smaller land use footprint for wind and solar power due to lower DAC deployment.
+On the other hand, the \modHybrid{} needs slightly more biomass, equipped with carbon capture, to achieve its cost advantage in comparison to the \modH{}.
 
@@
 
-\ac{}In transitioning from a net zero to a net emissions reduction target, a continuation of investment trends are perceived.
+In transitioning from a net zero to a net emissions reduction target, a continuation of investment trends are perceived.
 Fig.~\ref{fig:net-negative_cost_bar} shows the net investment changes per technology group for the \carbonmodel, the \Hgrid{} and \modHybrid{}s.
-The scenario imposes a net carbon removal of 460 Mt/a, with all models incurring a roughly similar net cost increase - 77 billion €/a for the \modCO{}, 81 billion €/a for the \modH{}, and 78 billion €/a for the \modHybrid{}.
+The scenario imposes a net carbon removal of 460 Mt/a, with all models incurring a roughly similar net annual cost increase - 77~bn€ for the \modCO{}, 81~bn€ for the \modH{}, and 78~bn€ for the \modHybrid{}.
 Both models invest relatively evenly in additional DAC, wind and heat pump systems, with DAC accounting for more than a third of the additional system costs.
@@
 The concept of a merit order for captured carbon plays a pivotal role in optimizing the deployment of carbon capture technologies based on their economic feasibility.
-This merit order ranks all CC technologies according to their relative costs to capture an additional marginal ton\footnote[2]{The cost value for a technology depends on the exact system configuration and may vary by scenario and even by region, since many technologies produce other (primary) outputs such as electricity and heat.
-The costs to capture an additional marginal ton of carbon may also be different from the average cost of capture.}
+This merit order ranks all CC technologies according to their relative costs to capture an additional marginal ton.
+The cost value for a technology depends on the exact system configuration and may vary by scenario and even by region, since many technologies produce other (primary) outputs such as electricity and heat.
+The costs to capture an additional marginal ton of carbon may also be different from the average cost of capture.
 At the lower end of the cost spectrum, CC technologies applied to process emissions, such as those from cement, offer a cost-effective starting point.
@@
     \centering
-    \includegraphics[width=\linewidth]{comparison/emission-reduction-full/figures/90_nodes/cost_bar_transmission.png}
-    \caption{Annual transmission system cost as a function of the net carbon removal scenarios considered in the study.}
+    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/cost_bar_transmission.png}
+    \caption{Annual transmission system cost for the different models in the net-zero emission scenario.}
     \label{fig:cost_bar_transmission}
+\end{figure}
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=\linewidth]{comparison/net-negative-0.1/figures/90_nodes/cost_bar_transmission.png}
+    \caption{Annual transmission system cost for the different models in the net-negative emissions scenario.}
+    \label{fig:cost_bar_transmission_nn}
 \end{figure}
@@
 \end{figure}
+
+\begin{figure*}[ht!]
+    \centering
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{net-negative-0.1/co2-only/figures/90_nodes/balance_map_carbon.png}
+        \label{fig:capacity_map_carbon_co2_nn}
+    \end{subfigure}%
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{net-negative-0.1/co2-only/figures/90_nodes/balance_map_hydrogen.png}
+        \label{fig:capacity_map_hydrogen_co2_nn}
+    \end{subfigure}
+    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \carbonmodel{} in the net negative scenario.
+For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
+Carbon Sequestration offshore are drawn as full circles.
+Lines and arrows show the interregional transportation.
+The \carbongrid{} reveals a fundamentally different layout than in the Net-Zero scenario.
+    }
+    \label{fig:capacity_maps_co2_nn}
+\end{figure*}
+
+\begin{figure*}[ht!]
+    \centering
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{net-negative-0.1/h2-only/figures/90_nodes/balance_map_carbon.png}
+        \label{fig:capacity_map_carbon_h2_nn}
+    \end{subfigure}%
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{net-negative-0.1/h2-only/figures/90_nodes/balance_map_hydrogen.png}
+        \label{fig:capacity_map_hydrogen_h2_nn}
+    \end{subfigure}
+    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hydrogenmodel{} in the net negative scenario.
+For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
+Carbon Sequestration offshore are drawn as full circles.
+Lines and arrows show the interregional transportation.
+With the tightened emission target, the \hydrogengrid{} expands the \hydrogengrid{} layout of the Net-Zero scenario model results while increasing DAC facilities at the coast.
+    }
+    \label{fig:capacity_maps_h2_nn}
+\end{figure*}
+
+\begin{figure*}[ht!]
+    \centering
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/balance_map_carbon.png}
+        \label{fig:capacity_map_carbon_full_nn}
+    \end{subfigure}%
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/balance_map_hydrogen.png}
+        \label{fig:capacity_map_hydrogen_full_nn}
+    \end{subfigure}
+    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hydrogenmodel{} in the net negative scenario.
+For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
+Carbon Sequestration offshore are drawn as full circles.
+Lines and arrows show the interregional transportation.
+With the tightened emission target, the \hybridmodel{} decreases the \hydrogengrid{} layout in comparison to the Net-Zero scenario and increases \carbon{} transport from inland point sources to the sequestration sites.
+    }
+    \label{fig:capacity_maps_full_nn}
+\end{figure*}
 
@@
 
-% Together with understanding the dynamics described above, we can highlight three cost advantages of the \Htwo{} network in contrast to the \COtwo{} network: Capturing carbon at the lowest-cost regions reduces 4 billion euros in spending on DAC.
-However, the \modH{} produces electrolytic hydrogen in the lowest-cost regions and in total spends 7 billion euros less on renewable energy (14 billion euros less on solar, and 7 billion euros more on wind) as well as 1 billion euros less on SMR.
-Furthermore, without an \Htwo{} network, additional \Htwo{} storage is needed in Spain at a cost of 5 billion euros to store hydrogen produced during the summer for utilization in FT fuel production during the winter.
-% And finally, the \modCO{ } spends 7 billion euros more on gas boilers, gas plants, gas infrastructure, and methanation.
+% Together with understanding the dynamics described above, we can highlight three cost advantages of the \Htwo{} network in contrast to the \COtwo{} network: Capturing carbon at the lowest-cost regions reduces 4~bn€ in spending on DAC.
+However, the \modH{} produces electrolytic hydrogen in the lowest-cost regions and in total spends 7~bn€ less on renewable energy (14~bn€ less on solar, and 7~bn€ more on wind) as well as 1~bn€ less on SMR.
+Furthermore, without an \Htwo{} network, additional \Htwo{} storage is needed in Spain at a cost of 5~bn€ to store hydrogen produced during the summer for utilization in FT fuel production during the winter.
+% And finally, the \modCO{ } spends 7~bn€ more on gas boilers, gas plants, gas infrastructure, and methanation.
 % TODO: why??
-% The costs of the respective networks, on the other hand, are almost equal: the \modCO{ } spends 8 billion euros on the \COtwo{} network, while the \modH{} even spends 10 billion euros on the \Htwo{} network.
-These factors add up to account for 14 billion euros of the difference of 20 billion euros in system costs between the two scenarios (785 billion euros for the \COgrid{} and 765 billion euros for the \modH{}, see Fig.~\ref{fig:objective_heatmap}).
+% The costs of the respective networks, on the other hand, are almost equal: the \modCO{ } spends 8~bn€ on the \COtwo{} network, while the \modH{} even spends 10~bn€ on the \Htwo{} network.
+These factors add up to account for 14~bn€ of the difference of 20~bn€ in system costs between the two scenarios (785~bn€ for the \COgrid{} and 765~bn€ for the \modH{}, see Fig.~\ref{fig:objective_heatmap}).
 

### 2023-12-14 report/submissions/journal/paper/paper.tex
address TODOs until 3.3

@@
 Some natural gas pipelines are already repurposed to transport hydrogen~\cite{RohrFreiFuer}.
-Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synthetic fuels on the one hand and offer \COtwo{} offtake on the other hand.
+Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synfuels on the one hand and offer \COtwo{} offtake on the other hand.
 The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a transport and sequestration capacity of 1.5~Mt \COtwo{} per year to be operative in 2024, expanding to a targeted scale of 5~Mt per year of sequestration by 2030.
@@
 In summary, both network models prefer CC from point sources distributed across the continent as well as electrolysis concentrated in regions with abundant renewable resources.
-The two network approaches reveal fundamentally different strategies to integrate these low-cost resources into the system: The \carbonmodel{} transports and sequesters low-cost carbon from distributed point sources and co-locates CU with low-cost electrolysis and DAC.
-The \hydrogenmodel{} distributes CU alongside with carbon point sources to which low-cost hydrogen from regions with abundant renewable resources is transported.
+The two network approaches reveal fundamentally different strategies to integrate these low-cost resources into the system: The \carbongridmodel{} transports and sequesters low-cost carbon from distributed point sources and co-locates CU with low-cost electrolysis and DAC.
+The \hydrogengridmodel{} distributes CU alongside with carbon point sources to which low-cost hydrogen from regions with abundant renewable resources is transported.
 %
@@
 % TODO: smoother transition, incl. end of 3.2
-% TODO: add network investment costs plot to Appendix. Mention investment H2 > CO2. (The investment in the hydrogen network is four times the investment in the carbon network.)
+% TODO: inlcude in opener: both networks, lower system costs; motivation: what is the advantage vs both others? other advantages (more resilient)/disadvantages (not focusing on one grid)
 The \modHybrid{} combines the advantages of hydrogen and carbon networks.
-Fig.~\ref{fig:cost_bar} shows that the \modHybrid{} achieves a cost reduction of \label{}5.2\% (\label{}41 billion euros) compared to the \modBase{}, which is lower than the cost reductions of the \COgrid{} and \modH{}.
-Fig.~\ref{fig:capacity_maps} shows the average production, consumption and transport of carbon and hydrogen in the \modHybrid{}.
+Fig.~\ref{fig:capacity_maps} shows the average production, consumption and transport of the two networks in the \modHybrid{}.
+
 While the topology of the \Htwo{} network roughly corresponds to that of the \modH{}, the \COtwo{} network plays more of a complementary role.
-In this section, we explore the advantages and disadvantages compared to the other models.
 
-Similar to the \modH{}, the \modHybrid{} transports hydrogen from concentrated electrolysis sites in Western regions (Iberian Peninsula and British Isles) to distributed CU production sites across the continent.
-For the \COtwo{} network, this model employs short and direct pipeline routes, typically spanning one or two segments, to transport carbon from inland CC point sources to perform CS at nearby sequestration sites (e.g., Northern Germany, Italy and Greece).
-Additionally, the \COtwo{} network utilizes single-segment network sections to equalize local variations in carbon demand and supply to produce CU, thereby also further converging carbon prices.
+% TODO: mention investment difference in network expansion with appendix plot (The investment in the hydrogen network is four times the investment in the carbon network.)
+% TODO: rewrite: mostly short (1-2 regions) linear (no cluster/mashed network) routes towards sequestration sites (for purpose of seq.); short 1-region transports to bridge local carbon price diffs, for purpose of CU.
+Similar to the \modH{}, the \modHybrid{} takes advantage of a large hydrogen network that enables hydrogen to be transported from centralized production sites in Western regions such as Spain to decentralized CU production sites close to CC point sources across the continent.
+The carbon network supplements this with smaller clusters of networks.
+In addition to a larger \COtwo{} network cluster in Central Europe, which transports carbon to the North Sea, the system is building three large linear carbon routes.
+These transport carbon from Romania and Bulgaria to Greece, from northern Italy to central Italy and from Hungary, the Czech Republic and Poland to the Baltic Sea.
+These enable the system to transport carbon available at low cost from inland biomass sources and sequester it (see Fig.~\ref{fig:captureshare_line}).
+This reduces the need for DAC in the system and the sequestration sites are shifted to the end points of the carbon networks.
+Second-order effects can be seen in the shift in FT production compared to the \modH{}.
+Regions that have both good renewable resources and sequestration potential could now forego sequestration and use local process emissions for FT production, as can be observed in Ireland.
+A notable aspect of the hybrid model is the lack of overlap in the hydrogen and carbon network topologies.
+This finding is tightly linked to the fact of competing functionalities to offtake carbon emissions.
+A more detailed of differences between the \modH{} and the \modHybrid{} can be found in Fig.~\ref{fig:cost_map_difference}.
 
-To understand the dynamics of the \modHybrid{} model, we observe that the model locates CU everywere, except for a few coastal regions that perform CS, or that supply their locally captured carbon to a neigbouring sequestration region.
-The model uses the \COtwo{} network to transport carbon over short distances to the CS sites, and the \Htwo{} network to transport hydrogen over longer distances from the lowest-cost hydrogen regions to CU sites across the continent.
-DAC is deployed to supply carbon for the remaining CU needed by the model (mostly in Southern Spain).
+% TODO: add interpretative paragraph: no CU where low-price ely, H2 is imported.
+% Both networks available to smooth prices; CS draws in CC from adjacent regions; CU everywhere locally, except the CS sites; DAC centered on (very few) regions with optimal RES (and Italy/Greece for methanation)
+% compare to H2 Grid model: küstennahe regionen mit longest distance to good RES(?) anzapfen, dort kein CU, CU jewtzt im Hinterland
 
-Despite only a further reduction of \label{}0.4 percentage points (\label{}3~bn€) in total system costs compared to the \modH{}, the \modHybrid{} brings several advantages: (1) The \modHybrid{} uses both \COtwo{} and \Htwo{} networks to build a more robust energy system; (2) it relies even less on DAC than the other models, which as a technology may be costlier than anticipated in the cost projections; and (3) it requires a smaller land use footprint for wind and solar power due to lower DAC deployment.
-On the other hand, the \modHybrid{} needs slightly more biomass, equipped with carbon capture, to achieve its cost advantage in comparison to the \modH{}.
+Despite only a further reduction of \label{}0.4 percentage points (\label{}3~bn€) in total system costs compared to the \modH{}, the \modHybrid{} brings several advantages.
+These include a broader range of technologies contributing to its robustness, reduced reliance on DAC, which may be costlier than anticipated in the cost projections, and less land use for wind and solar due to the decreased necessity for DAC.
+The increase used of biomass with carbon capture, transport and sequestration leads to slight cost advantages in comparison to the \modH{}.
 
@@
 The concept of a merit order for captured carbon plays a pivotal role in optimizing the deployment of carbon capture technologies based on their economic feasibility.
-This merit order ranks all CC technologies according to their relative costs to capture an additional marginal ton.
-The cost value for a technology depends on the exact system configuration and may vary by scenario and even by region, since many technologies produce other (primary) outputs such as electricity and heat.
-The costs to capture an additional marginal ton of carbon may also be different from the average cost of capture.
+This merit order ranks all CC technologies according to their relative costs to capture an additional marginal ton\footnote[2]{The cost value for a technology depends on the exact system configuration and may vary by scenario and even by region, since many technologies produce other (primary) outputs such as electricity and heat.
+The costs to capture an additional marginal ton of carbon may also be different from the average cost of capture.}
 At the lower end of the cost spectrum, CC technologies applied to process emissions, such as those from cement, offer a cost-effective starting point.
@@
     \centering
-    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/cost_bar_transmission.png}
-    \caption{Annual transmission system cost for the different models in the net-zero emission scenario.}
+    \includegraphics[width=\linewidth]{comparison/emission-reduction-full/figures/90_nodes/cost_bar_transmission.png}
+    \caption{Annual transmission system cost as a function of the net carbon removal scenarios considered in the study.}
     \label{fig:cost_bar_transmission}
-\end{figure}
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=\linewidth]{comparison/net-negative-0.1/figures/90_nodes/cost_bar_transmission.png}
-    \caption{Annual transmission system cost for the different models in the net-negative emissions scenario.}
-    \label{fig:cost_bar_transmission_nn}
 \end{figure}
@@
 \end{figure}
-
-\begin{figure*}[ht!]
-    \centering
-    \begin{subfigure}{.5\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{net-negative-0.1/co2-only/figures/90_nodes/balance_map_carbon.png}
-        \label{fig:capacity_map_carbon_co2_nn}
-    \end{subfigure}%
-    \begin{subfigure}{.5\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{net-negative-0.1/co2-only/figures/90_nodes/balance_map_hydrogen.png}
-        \label{fig:capacity_map_hydrogen_co2_nn}
-    \end{subfigure}
-    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \carbonmodel{} in the net negative scenario.
-For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
-Carbon Sequestration offshore are drawn as full circles.
-Lines and arrows show the interregional transportation.
-The \carbongrid{} reveals a fundamentally different layout than in the Net-Zero scenario.
-    }
-    \label{fig:capacity_maps_co2_nn}
-\end{figure*}
-
-\begin{figure*}[ht!]
-    \centering
-    \begin{subfigure}{.5\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{net-negative-0.1/h2-only/figures/90_nodes/balance_map_carbon.png}
-        \label{fig:capacity_map_carbon_h2_nn}
-    \end{subfigure}%
-    \begin{subfigure}{.5\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{net-negative-0.1/h2-only/figures/90_nodes/balance_map_hydrogen.png}
-        \label{fig:capacity_map_hydrogen_h2_nn}
-    \end{subfigure}
-    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hydrogenmodel{} in the net negative scenario.
-For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
-Carbon Sequestration offshore are drawn as full circles.
-Lines and arrows show the interregional transportation.
-With the tightened emission target, the \hydrogengrid{} expands the \hydrogengrid{} layout of the Net-Zero scenario model results while increasing DAC facilities at the coast.
-    }
-    \label{fig:capacity_maps_h2_nn}
-\end{figure*}
-
-\begin{figure*}[ht!]
-    \centering
-    \begin{subfigure}{.5\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/balance_map_carbon.png}
-        \label{fig:capacity_map_carbon_full_nn}
-    \end{subfigure}%
-    \begin{subfigure}{.5\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/balance_map_hydrogen.png}
-        \label{fig:capacity_map_hydrogen_full_nn}
-    \end{subfigure}
-    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hydrogenmodel{} in the net negative scenario.
-For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
-Carbon Sequestration offshore are drawn as full circles.
-Lines and arrows show the interregional transportation.
-With the tightened emission target, the \hybridmodel{} decreases the \hydrogengrid{} layout in comparison to the Net-Zero scenario and increases \carbon{} transport from inland point sources to the sequestration sites.
-    }
-    \label{fig:capacity_maps_full_nn}
-\end{figure*}
 

### 2023-12-15 report/submissions/journal/paper/paper.tex
paper: remove footnote, add more figures to appendix

@@
 In summary, both network models prefer CC from point sources distributed across the continent as well as electrolysis concentrated in regions with abundant renewable resources.
-The two network approaches reveal fundamentally different strategies to integrate these low-cost resources into the system: The \carbongridmodel{} transports and sequesters low-cost carbon from distributed point sources and co-locates CU with low-cost electrolysis and DAC.
-The \hydrogengridmodel{} distributes CU alongside with carbon point sources to which low-cost hydrogen from regions with abundant renewable resources is transported.
+The two network approaches reveal fundamentally different strategies to integrate these low-cost resources into the system: The \carbonmodel{} transports and sequesters low-cost carbon from distributed point sources and co-locates CU with low-cost electrolysis and DAC.
+The \hydrogenmodel{} distributes CU alongside with carbon point sources to which low-cost hydrogen from regions with abundant renewable resources is transported.
 %

### 2023-12-15 report/submissions/journal/paper/paper.tex
make large map tighter; add appendix

--- 2023-12-14 report/submissions/journal/paper/paper.tex
@@
     \centering
-    \includegraphics[width=\linewidth]{comparison/single-technologies/figures/90_nodes/balance_map_dedicated.png}
+    \includegraphics[width=0.98\linewidth]{comparison/single-technologies/figures/90_nodes/balance_map_dedicated.png}
     \caption{Optimal operation, flows and prices of the carbon (top line) and hydrogen (bottom line) sectors for the \carbonmodel{} (left) and the \hydrogenmodel{} (right) in the net zero scenario.
@@
 
-% TODO: add numbers: electrolyzers GW (Spain/Iberia, Germany), hydrogen prices (valleys, peaks)
 As we have also discussed above, the \modCO{} captures large amounts of carbon from bioenergy and industrial activities that were untapped in the \modBase{}.
@@
 
+\newpage
 \appendix
+\setcounter{section}{0}
+\renewcommand{\thesection}{\Alph{section}}
+\renewcommand{\thefigure}{\Alph{section}.\arabic{figure}} % Add this line to apply the new counter to figures
+
+\onecolumn % Add this line to adjust the layout to single column
+
+\section*{Appendix}
 
@@
 
-\begin{figure}
+\subsection{Cost comparison across models}
+
+The following figures show further comparisons of total system costs and cost contributions of single technologies across the models and scenarios.
+
+\begin{figure}[h!]
     \centering
-    \includegraphics*[width=0.8\linewidth]{comparison/emission-reduction/figures/90_nodes/objective_heatmap.png}
+    \includegraphics*[width=0.4\linewidth]{comparison/emission-reduction/figures/90_nodes/objective_heatmap.png}
     \caption{Total annual system cost for the different network models and \COtwo{} targets.
@@
     \centering
+    \begin{subfigure}{.5\textwidth}
+        \centering
         \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/cost_bar_transmission.png}
-    \caption{Annual transmission system cost for the different models in the net-zero emission scenario.}
+        \caption{}
         \label{fig:cost_bar_transmission}
+    \end{subfigure}%
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{comparison/net-negative-0.1/figures/90_nodes/cost_bar_transmission.png}
+        \caption{}
+        \label{fig:cost_bar_transmission_nn}
+    \end{subfigure}
+    \caption{Annual transmission system costs for different models and scenarios in the Net-Zero scenario (a) and Net-Negative scenario (b).}
 \end{figure}
 
-\begin{figure}[ht]
+\begin{figure}[ht!]
     \centering
-    \includegraphics[width=\linewidth]{comparison/net-negative-0.1/figures/90_nodes/cost_bar_transmission.png}
-    \caption{Annual transmission system cost for the different models in the net-negative emissions scenario.}
-    \label{fig:cost_bar_transmission_nn}
+    \includegraphics[width=\linewidth]{difference/h2-only-full/figures/90_nodes/cost_map.png}
+    \caption{Difference in regional costs between the \Hgrid{} and \modHybrid{}s in the Net-Zero scenario.
+The left figure shows higher spendings per technology and region and transport system for the \modH{}, the right shows higher spendings in the \modHybrid{}.}
+    \label{fig:cost_map_difference}
 \end{figure}
 
-\begin{figure*}[ht]
+\begin{figure}[ht!]
     \centering
-    \includegraphics[width=\linewidth]{difference/h2-only-full/figures/90_nodes/cost_map.png}
-    \caption{Difference in regional costs between the \Hgrid{} and \modHybrid{}s.
-The left subfigure shows higher spendings per technology and region and transport system for the \modH{}, the right shows higher spendings in the \modHybrid{}.}
-    \label{fig:cost_map_difference}
-\end{figure*}
-
-\begin{figure}[htb!]
-    \centering
-    \includegraphics[width=0.9\linewidth]{difference/comparison/subsidy/figures/90_nodes/cost_bar.png}
-    \caption[short]{Net change in system cost when reducing assumed costs for \COtwo{} pipelines by 50\%.}
-    \label{fig:half-CO2-price_cost_bar}
+    \includegraphics[width=\linewidth]{difference/net-negative-0.1-h2-only-full/figures/90_nodes/cost_map.png}
+    \caption{Difference in regional costs between the \Hgrid{} and \modHybrid{}s in the Net-Negative scenario.
+The left figure shows higher spendings per technology and region and transport system for the \modH{}, the right shows higher spendings in the \modHybrid{}.}
+    \label{fig:cost_map_difference_nn}
 \end{figure}
 
-\begin{figure*}[ht!]
+\clearpage
+\subsection{Subsidizing \carbon{} pipelines}
+
+The effect of a 50\% subsidy on \carbon{} pipelines impacts the optimal technology deployment to small extent.
+Fig.~\ref{fig:cost_bar_subsidy} shows the corresponding total system cost of all models in the Net-Zero scenario.
+Fig.~\ref{fig:cost_bar_diff_subsidy} shows the net change in system cost between non-subsidized and subsidized models, showing only the \carbonmodel{} and \hybridmodel{} where changes occur.
+
+\begin{figure}[ht!]
+    \centering
+    \begin{subfigure}{.5\textwidth}
+    \includegraphics[width=\linewidth]{comparison/subsidy/figures/90_nodes/cost_bar.png}
+    \caption{}
+    \label{fig:cost_bar_subsidy}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+    \centering
+    \includegraphics[width=\linewidth]{difference/comparison/subsidy/figures/90_nodes/cost_bar.png}
+    \caption{}
+    \label{fig:cost_bar_diff_subsidy}
+\end{subfigure}
+\caption{System costs for the all models in the net-zero emission scenario with 50\% subsidy of \carbon{} pipelines (a) and change in system cost for the \carbonmodel{} and \hybridmodel{} when subsidizing \carbon{} pipelines by 50\% (b).}
+\end{figure}
+
+\clearpage
+\subsection{Operation in Net-Negative Scenarios}
+The following figures display the optimal operation of the hydrogen and carbon sector for all models in the Net-Negative scenario.
+
+\begin{figure}[ht!]
     \centering
@@
     \label{fig:capacity_maps_co2_nn}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}[ht!]
+\begin{figure}[ht!]
     \centering
@@
     \label{fig:capacity_maps_h2_nn}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}[ht!]
+\begin{figure}[ht!]
     \centering
@@
     \label{fig:capacity_maps_full_nn}
-\end{figure*}
+\end{figure}
 

### 2023-12-15 workflow/scripts/plot_balance_map_dedicated.py

@@ -53,12 +53,15 @@ region_alpha = config["plotting"]["balance_map"]["region_alpha"]
 fig, axes = plt.subplots(
     2,
     2,
-    figsize=(8, 10),
+    figsize=(8, 8),
     subplot_kw={"projection": ccrs.EqualEarth()},
     layout="constrained",
 )
 
-for n, axs, draw_legend in zip(networks, axes.T, [False, True]):
+bounds = snakemake.config["plotting"]["extent"]
+bounds = [-10, 28, 36, 70]  # adjust for better alignment with latex document
+
+for n, axs, right_subplot in zip(networks, axes.T, [False, True]):
     s = n.statistics
     colors = n.carriers.set_index("nice_name").color
     run = n.meta["wildcards"]["run"]
@@ -85,8 +88,7 @@ for n, axs, draw_legend in zip(networks, axes.T, [False, True]):
         branch_scale = float(specs["branch_scale"])
         flow_scale = float(specs["flow_scale"])
         legend_kwargs = {
-            "loc": "upper left",
-            "frameon": True,
+            "frameon": False,
             "framealpha": 1,
             "edgecolor": "None",
         }
@@ -114,7 +116,7 @@ for n, axs, draw_legend in zip(networks, axes.T, [False, True]):
                 ax=ax,
                 color_geomap=False,
                 geomap=True,
-                boundaries=snakemake.config["plotting"]["extent"],
+                boundaries=bounds,
             )
             offregions.plot(
                 ax=ax,
@@ -146,7 +148,7 @@ for n, axs, draw_legend in zip(networks, axes.T, [False, True]):
             margin=0.2,
             color_geomap={"border": "darkgrey", "coastline": "darkgrey"},
             geomap="10m",
-            boundaries=snakemake.config["plotting"]["extent"],
+            boundaries=bounds,
         )
 
         buses = n.buses.query("carrier in @carriers").index
@@ -178,7 +180,7 @@ for n, axs, draw_legend in zip(networks, axes.T, [False, True]):
             aspect="equal",
         )
 
-        if draw_legend:
+        if right_subplot:
             sm = plt.cm.ScalarMappable(
                 cmap=cmap, norm=plt.Normalize(vmin=vmin, vmax=vmax)
             )
@@ -187,7 +189,7 @@ for n, axs, draw_legend in zip(networks, axes.T, [False, True]):
                 sm,
                 ax=axes[0] if kind == "carbon" else axes[1],
[…]

### 2023-12-15 report/submissions/journal/paper/paper.tex
paper: finish 3.4

@@
 % TODO: add network investment costs plot to Appendix. Mention investment H2 > CO2. (The investment in the hydrogen network is four times the investment in the carbon network.)
-The \modHybrid{} combines the advantages of hydrogen and carbon networks.
-Fig.~\ref{fig:cost_bar} shows that the \modHybrid{} achieves a cost reduction of \label{}5.2\% (\label{}41 billion euros) compared to the \modBase{}, which is lower than the cost reductions of the \COgrid{} and \modH{}.
-Fig.~\ref{fig:capacity_maps} shows the average production, consumption and transport of carbon and hydrogen in the \modHybrid{}.
+The \modHybrid{} combines the advantages of hydrogen and carbon networks, resulting in a cost reduction of \label{}5.2\% (\label{}41 billion euros) compared to the \modBase{}, which is lower than the cost reductions of the \COgrid{} and \modH{} (see Fig.~\ref{fig:cost_bar}).
+In particular, production, consumption and transport in the carbon and hydrogen sectors in the \modHybrid{} combine elements of the \carbonmodel{} and \hydrogenmodel{} (see Fig.~\ref{fig:capacity_maps}).
 While the topology of the \Htwo{} network roughly corresponds to that of the \modH{}, the \COtwo{} network plays more of a complementary role.
@@
 Additionally, the \COtwo{} network utilizes single-segment network sections to equalize local variations in carbon demand and supply to produce CU, thereby also further converging carbon prices.
-
-To understand the dynamics of the \modHybrid{} model, we observe that the model locates CU everywere, except for a few coastal regions that perform CS, or that supply their locally captured carbon to a neigbouring sequestration region.
-The model uses the \COtwo{} network to transport carbon over short distances to the CS sites, and the \Htwo{} network to transport hydrogen over longer distances from the lowest-cost hydrogen regions to CU sites across the continent.
+The model locates CU everywhere, except for a few coastal regions that perform CS, or that supply their locally captured carbon to a neigbouring sequestration region.
 DAC is deployed to supply carbon for the remaining CU needed by the model (mostly in Southern Spain).
 
-Despite only a further reduction of \label{}0.4 percentage points (\label{}3~bn€) in total system costs compared to the \modH{}, the \modHybrid{} brings several advantages: (1) The \modHybrid{} uses both \COtwo{} and \Htwo{} networks to build a more robust energy system; (2) it relies even less on DAC than the other models, which as a technology may be costlier than anticipated in the cost projections; and (3) it requires a smaller land use footprint for wind and solar power due to lower DAC deployment.
-On the other hand, the \modHybrid{} needs slightly more biomass, equipped with carbon capture, to achieve its cost advantage in comparison to the \modH{}.
+Despite only a further reduction of \label{}0.4 percentage points (\label{}3~bn€) in total system costs compared to the \modH{}, the \modHybrid{} brings several advantages: (1) The \modHybrid{} relies even less on DAC than the other models, which as a technology may be costlier than anticipated in the cost projections, by making use of carbon point sources; (2) it requires a smaller land use footprint for wind and solar power due to lower DAC deployment.
+On the other hand, the \modHybrid{} needs slightly more biomass, equipped with carbon capture, to achieve its cost advantage in comparison to the \modH{}; (3) it uses both \COtwo{} and \Htwo{} networks to flatten price distribution across the continent and to build a more robust energy system.
 
-\subsection*{Net-negativity amplifies DAC and Bioenergy with CU and CS}\label{subsec:NN}
+\subsection*{Net-negativity amplifies DAC and \carbon{} transport}\label{subsec:NN}
 
@@
     \includegraphics[width=0.9\linewidth]{difference/comparison/emission-reduction-0.1/figures/90_nodes/cost_bar.png}
-    \caption[short]{Net change in investments when tightening the \COtwo{} emission target from net-zero to net-negative 10\% of 1990s emissions.
+    \caption[short]{Net change in system cost when tightening the \COtwo{} emission target from net-zero to net-negative 10\% of 1990s emissions.
+Each cost bar is split into contribution of the same technology groups, except for \carbon{} sequestration cost which are now diplayed separately.
 For all models, Direct Air Capture (DAC) contributes most to the additional \carbon{} removal, requiring further solar, wind and heat pump capacities for electricity and heat input.}
@@
 
-% TODO: start with explanation Fig 6: similarities b/w all models: sequestration, more DAC (=+wind, solar, heat pump, battery), point-source CC (+bioenergy); diff: level of DAC vs level of point-source CC, co2 network, h2 network
+In the following, we analyze how investment trends and network strategies alter when transitioning from a net zero to a net emissions reduction target.
+Therefore, we subject the models to the Net-Negative scenario with a net carbon removal of 460~Mt/a and an increased total sequestration limit of 660~Mt/a.
+In comparison to the Net-Zero scenario, all models with transport options incur a roughly similar net annual cost increase - 77~bn€ for the \modCO{}, 81~bn€ for the \modH{}, and 78~bn€ for the \modHybrid{} (see Fig.~\ref{fig:net-negative_cost_bar}), resulting in a higher cost-benefit of 6~bn€ for the \hybridmodel{} in comparison to the \hydrogenmodel{}.
 
-% carbon:
-% - fundamentally new network layout
-% - CU in spain is now supplied by carbon from DAC
-% - carbon in Central Europe is mainly transported to seq sites
-% - stranded investment in carbon network of 3bn
+The models equally invest in the expansion of sequestration capacities up to the global limit (5~bn€).
+The investments in additional DAC (36-41~bn€), wind (6-7~bn€) and heat pumps (6-7~bn€) are relatively evenly for all models, with DAC accounting for roughly half of the additional system costs.
+Fundamental differences in investment strategies occur in the deployment of solar, carbon capture and bio-energy technologies: While the \carbonmodel{} expands CC at point sources the most, the \hydrogenmodel{} and the \hybridmodel{} invest relatively more in new solar plants.
+The \carbonmodel{} invests three time more in new bio-energy sources (6~bn€) than the other models.
+All models only reveal small net downscaling of technologies.
+However, as we discuss below, the geographic impact can vary depending on the model, sometimes leading to disruptive changes in optimal network strategies and stranded assets.
 
-% hydrogen:
-% - no significant changes in strategies
+In contrast to its layout in the Net-Zero scenario, the \carbonmodel{} fundamentally reorganizes the carbon transport pathways (see Fig.~\ref{fig:balance_maps_co2_nn}).
+In Spain, carbon utilization (CU) is now predominantly supplied by carbon extracted from DAC.
+Central Europe sees a major transformation, with carbon mainly being transported to sequestration sites in the north and south-east.
+This reorganization leads to stranded investment in the carbon network infrastructure, amounting to approximately 3~bn€.
+%
+The \modH{} on the other hand, primarily leaves the network topology unchanged and only expands DAC facilities at regions with access to sequestration sites as well as additional solar, wind and heat pumps in the vicinity to supply power and heat (see Fig.~\ref{fig:balance_maps_h2_nn}).
+%
+Finally, the \modHybrid{} expands the sequestration with carbon captured from bio-energetic inputs and transported from Central and Eastern Europe to the near shores (see Fig.~\ref{fig:balance_maps_full_nn}), leading to an additional investment in the \carbongrid{} (1.2~bn€, see Fig.~\ref{fig:cost_bar_transmission_nn}).
+At the same time, carbonaceous fuel production in Central Europe partially moves to Spain, where it uses carbon from new DAC facilities without the installation of a \COtwo{} network.
+As a trade-off, the model transports less hydrogen from the Iberian Peninsula to Central Europe, resulting in a dismantling of 2~bn€ in hydrogen network costs compared to the Net-Zero scenario.
 
-% hybrid:
-% - more CU in spain
-% - less cu in central europe
-% - more sequestration in central europe and Portugal
-% - less h2 import from spain to central european
-% - mostly additive investments, except for some CU plants near the coast (of 4.3bn) and h2 network from Spain to Central Europe (2bn)
-
-% TODO: add operation plots into appendix all models (hydrogen and carbon)
-
-In transitioning from a net zero to a net emissions reduction target, a continuation of investment trends are perceived.
-Fig.~\ref{fig:net-negative_cost_bar} shows the net investment changes per technology group for the \carbonmodel, the \Hgrid{} and \modHybrid{}s.
-The scenario imposes a net carbon removal of 460 Mt/a, with all models incurring a roughly similar net annual cost increase - 77~bn€ for the \modCO{}, 81~bn€ for the \modH{}, and 78~bn€ for the \modHybrid{}.
-Both models invest relatively evenly in additional DAC, wind and heat pump systems, with DAC accounting for more than a third of the additional system costs.
-These remove carbon directly at sequestration sites at the coast, in particular in Portugal and UK, together with heat and power from new heat pumps at site and new solar and wind installation in the vicinity.
-
-Fundamental difference in investment strategies occur in the carbon capture sector: In contrast to the \modH{}, the \modHybrid{} continues the strategy of sequestering carbon captured and transported from bio-energetic inputs, expanding the carbon transport routes from Central and Eastern Europe to the near shores.
-At the same time, carbonaceous fuel production in Central Europe partially moves to Spain, where it uses carbon from new DAC facilities without the installation of a \COtwo{} network.
-At the same time, less hydrogen imports from the Iberian Peninsula to Central Europe is needed due to less CU.
-The \modH{} on the other hand, expands decentralized CU systems from the Net-Zero scenario, which are now increasingly supplied with carbon from biogas-to-gas facilities, showing up in both ``Bioenergy'' and ``Carbon Capture at Point Sources'' cost-contributions.
-The centralized DAC facilities at sequestration sites are more expanded than in the \modHybrid{} and are requiring additional heat from biomass CHPs.
-
-These findings highlight the continuation of trends of the two systemic approaches, which are interchangable at marginal cost-difference: While a stand-alone \Htwo{} network combines well with highly centralized DAC and decentralized CU with bioenergy as primary carbon source, the complementation with a \COtwo{} network promotes combining DAC and CU where good renewable resources are and collecting, transporting and sequestering carbon from decentral bioenergetic processes.
-
-% The difference in absolute investment changes between the two models, regionally displayed in Fig.~\ref{fig:net-negative_cost_map}.
-% \begin{figure*}[htb!]
-%     \centering
-%     \includegraphics[width=\linewidth]{difference/net-negative-0.1-h2-only-full/figures/90_nodes/cost_map.png}
-%     \caption[short]{Net difference in investments between the \modH{} and the \modHybrid{} for the Net-Negative scenario.}
-%     \label{fig:net-negative_cost_map}
-% \end{figure*}
+These findings highlight the continued trends of the two systemic approaches which are combined in the \hybridmodel.
+With an increased demand for sequestration, the \carbongrid{} relatively gains more importance, leading to higher cost-benefits by enabling to combine increased need DAC with a more dominant collection and transportation scheme of carbon from distributed point sources.
 
@@
 
-% \begin{figure*}
-%     \centering
-%     \includegraphics[width=\linewidth]{full/figures/90_nodes/sankey_diagramm.png}
-%     \caption{Sankey diagram of the optimal operation for a net-zero scenario.}
-%     \label{fig:sankey_diagramm}
-% \end{figure*}
+\subsection{Model Inputs}
 
-% \begin{figure*}
-%     \centering
-%     \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
-%     \caption{Sankey diagram of the optimal operation for a net-negative 10\% scenario.}
-%     \label{fig:sankey_diagramm}
-% \end{figure*}
+The following figures extend on the model assumptions and inputs.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/sequestration_map.png}
+    \caption{Maximal sequestration potential per offshore region used as input for all models.
+Note that in most model runs the regional sequestration potentials are not exploited due the the global limit on sequestration.}
+\end{figure}
 
@@
 \clearpage
+\subsection{Operation of the \hybridmodel{} in Net-Zero Scenarios}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{full/figures/90_nodes/sankey_diagramm.png}
+    \caption{Sankey diagram of the carbon flows in the \hybridmodel{} in the Net-Zero scenario.}
+    \label{fig:sankey_diagramm}
+\end{figure}
+
+\clearpage
 \subsection{Operation in Net-Negative Scenarios}
@@
         \includegraphics[width=\linewidth]{net-negative-0.1/co2-only/figures/90_nodes/balance_map_carbon.png}
-        \label{fig:capacity_map_carbon_co2_nn}
+        \label{fig:balance_map_carbon_co2_nn}
     \end{subfigure}%
@@
         \includegraphics[width=\linewidth]{net-negative-0.1/co2-only/figures/90_nodes/balance_map_hydrogen.png}
-        \label{fig:capacity_map_hydrogen_co2_nn}
+        \label{fig:balance_map_hydrogen_co2_nn}
     \end{subfigure}
@@
     }
-    \label{fig:capacity_maps_co2_nn}
+    \label{fig:balance_maps_co2_nn}
 \end{figure}
@@
         \includegraphics[width=\linewidth]{net-negative-0.1/h2-only/figures/90_nodes/balance_map_carbon.png}
-        \label{fig:capacity_map_carbon_h2_nn}
+        \label{fig:balance_map_carbon_h2_nn}
     \end{subfigure}%
@@
         \includegraphics[width=\linewidth]{net-negative-0.1/h2-only/figures/90_nodes/balance_map_hydrogen.png}
-        \label{fig:capacity_map_hydrogen_h2_nn}
+        \label{fig:balance_map_hydrogen_h2_nn}
     \end{subfigure}
@@
     }
-    \label{fig:capacity_maps_h2_nn}
+    \label{fig:balance_maps_h2_nn}
 \end{figure}
@@
         \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/balance_map_carbon.png}
-        \label{fig:capacity_map_carbon_full_nn}
+        \label{fig:balance_map_carbon_full_nn}
     \end{subfigure}%
@@
         \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/balance_map_hydrogen.png}
-        \label{fig:capacity_map_hydrogen_full_nn}
+        \label{fig:balance_map_hydrogen_full_nn}
     \end{subfigure}
@@
     }
-    \label{fig:capacity_maps_full_nn}
+    \label{fig:balance_maps_full_nn}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
+    \caption{Sankey diagram of the carbon flows in the \hybridmodel{} in the Net-Negative scenario.}
+    \label{fig:sankey_diagramm}
 \end{figure}

### 2023-12-15 report/submissions/journal/paper/paper.tex
review methods section; adjust conclusion

@@
 
-\title{Assessing \Htwo{} and \COtwo{} Network Strategies in Europe}
+% \title{Assessing \Htwo{} and \COtwo{} Network Strategies in Europe}
 \title{Competition and Synergies of \Htwo{} and \COtwo{} Networks in Europe}
@@
 Both \COtwo{} and \Htwo{} networks are likely to play a crucial role in the cost-effective integration of these technologies to enable net-zero economies.
-% [TODO: integrate definition of Carbon Management (CM) into last sentence.]. COMMENT: CM was already mention earlier, don't repeat it instead changed framing of last sentence
-
-% TODO: Move to better place in paper
-We argue that it is crucial to consider CC, CU and CS as separate technologies, rather than grouping them under the often-used but misleading umbrella term ``carbon capture, utilization and storage'' (CCUS), to allow for independent optimization and adequate representation of each function in the carbon management system.
 
@@
 
-% \section{Limitations}
-% \label{sec:limitations}
-% Despite the detailed model representation, there are some limitations to the validity of the results.
-The model itself is based on a linearized optimization with perfect foresight for the entire modeling year.
-In reality, long-term energy demand and renewable supply can only be roughly estimated, while short-term predictions still entail some uncertainty.
-The model's perfect foresight may lead to non-reproducible behavior, such as precisely aligning energy storage with future energy shortages at a specific point in time.
-% The technology costs used in the model rely on cost projections that incorporate reductions based on learning rates.
-These learning rates are derived from historical data, which may not necessarily be indicative of future trends, and they are based on projected global installed capacities of one technology, which can result in large uncertainties.
-In particular, cost assumptions on DAC, electrolysis and \Htwo{} pipeline costs, could have a strong impact in the model results.
-% The modelling results are heavily driven by the demand and emissions from industrial clusters.
-Allowing the model to relocate these and/or incorporate flexibility measures, may lead to less dependency on both \COtwo{} and \Htwo{} networks.
-% The technological flexibilities in the carbon sector might not be exploited to full extent.
-Therefore, the \COtwo{} transport via truck or ship is not considered.
-Fischer-Tropsch facilities require to run on baseload with at least 90\% of their nominal capacity.
-% Furthermore, the model assumes a fixed rate for industrial process emissions, which cannot be altered through investments.
-This simplification may not accurately represent real-world scenarios, where many industries are considering adopting low-carbon processes and technologies.
-The transport of fuel (FT, gas, oil) between the regions is not unlimited which may overestimate the flexibility provided by these commodities.
-
 \section{Conclusion}
@@
 
-This study offers a detailed exploration of Europe's energy system in 2050, focusing on the integration and optimization of hydrogen and carbon networks to achieve climate neutrality.
-The analysis reveals significant efficiencies gained through these networks, particularly in terms of leveraging low-cost carbon sources and facilitating a decentralized approach to carbon utilization.
+Our study presents a comprehensive assessment of the roles of hydrogen (\Htwo{}) and carbon dioxide (\COtwo{}) networks in the future European energy system.
+By comparing models with and without these networks under net-zero and net-negative scenarios, we have elucidated the benefits and trade-offs associated with each approach.
+The Baseline model, lacking both \Htwo{} and \COtwo{} networks, shows significant inefficiencies and higher system costs.
+The \COgrid{} model, with its \COtwo{} network, effectively integrates low-cost carbon from distributed point sources, primarily for carbon utilization and sequestration.
+Conversely, the \Hgrid{} model, with its \Htwo{} network, facilitates low-cost hydrogen delivery to high-demand areas, thus enabling efficient carbon utilization alongside point-source carbon capture.
+The Hybrid model, combining both \COtwo{} and \Htwo{} networks, leverages the strengths of each, resulting in the lowest system costs and a robust energy system.
+This model effectively utilizes \COtwo{} network for short-range carbon transport for sequestration and complements this with the \Htwo{} network for long-range hydrogen transport to support distributed carbon utilization.
+Under the net-negative scenario, the importance of Direct Air Capture (DAC) increases across all models, highlighting the growing significance of DAC in achieving deeper decarbonization goals.
 
-The hydrogen network demonstrates superior cost-effectiveness, primarily due to its capacity to transport low-cost hydrogen to high-demand regions and promote decentralized carbon capture and utilization.
-This network reduces reliance on direct air capture (DAC), offering a shift towards more economically viable and sustainable solutions.
-The combination of hydrogen and carbon networks in the hybrid model introduces additional flexibilities.
-This model not only achieves further cost reductions but also shows resilience against potential increases in DAC costs and reduces land use for renewable energy installations.
-Under net-negative emission scenarios, the importance of DAC and bioenergy with carbon capture and storage (BECCS) is amplified.
-Both the hydrogen and hybrid models demonstrate adaptability to these more stringent climate targets, highlighting the robustness of the networked approaches.
+Despite the comprehensive nature of the model, there are some limitations that affect the validity of the results, as well as important discussion points that need to be put into context.
+In reality, energy demand and renewable supply can only be estimated for the next few hours/days.
+The model's perfect foresight may lead to non-reproducible technology dispatch which neglects uncertainties from long-term predictions and necessary reserve margins.
+The modelling results are heavily driven by the demand and emissions from industrial clusters.
+Allowing the model to relocate these, incorporate flexibility measures or alter underlying processes, may lead to less dependency on both \COtwo{} and \Htwo{} networks.
+The transport of fuel (FT, gas, oil) between the regions is not limited which may overestimate the flexibility provided by these commodities.
+However, the distributed CU systems in the \hydrogenmodel{} and \hybridmodel{} align well with today's distributed locations for oil-refineries, providing kerosene for aviation and naphtha for industry.
+The assumption to run Fischer-Tropsch facilities on baseload with at least 90\% of their nominal capacity, is likely shaping the deployment of CU, underestimating potential important flexibilities.
+The assumed cost projections on technologies are subject to uncertainties.
+As shown in \ref{sec:subsidy}, the effect on a 50\% cost reduction on \carbon{} pipelines alter total system costs to small extent (below 1\%).
+However, a different set of cost on DAC and electrolysis may-well impact in the model results.
+In combination with a net-zero or net-negative emission target, a relaxation of the global sequestration limit drives system costs down by introducing fossil fuels, as shown in~\cite{hofmannDesigningCO2Network2023}.
+Allowing the model to sequester onshore, would likely reduce the dominant co-location of CU with the point sources due to easier access to sequestration sites.
+Finally, the large deployment of DAC is subject to conservative biomass potentials, as shown in~\cite{lauerCrucialRoleBioenergy2023}.
 
-% TODO: distribution of current refineries might shape the CU network. These are quite decentralized and might fit to the cost-optimal layout.
-
-The study conclusively demonstrates the pivotal role of strategically integrated hydrogen and carbon networks in steering Europe towards a sustainable and climate-neutral energy future.
-It emphasizes the necessity of comprehensive infrastructure planning, integrating diverse energy carriers and technologies to achieve these objectives effectively.
+In summary, our study underscores the need for strategic planning and integration of both \COtwo{} and \Htwo{} networks in the European energy landscape.
+The flexibility and cost-effectiveness provided by these networks are pivotal for realizing the EU's climate neutrality ambitions.
+By harnessing the synergies between \Htwo{} and \COtwo{} transportation and management, Europe can effectively navigate the complexities of a transition to a sustainable, low-carbon energy future.
 
@@
 All technology cost assumptions are taken for the year 2040 and sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
-Many of the assumptions contained therein are based on the technology catalogues published by the Danish Energy Agence~\cite{TODO}.
+Many of the assumptions contained therein are based on the technology catalogues published by the Danish Energy Agency~\cite{danishenergyagencyTechnologyDataGeneration2019,thedanishenergyagencyTechnologyDataCarbon2023}.
 Endogenous model results include the expansion of renewable energy sources, storage technologies, power-to-X conversion and transmission capacities, heating technologies, peaking power plants, and the deployment of gray, blue or green hydrogen, among others.
@@
 
-% emits 633~Mt \COtwo{} per year from industry, aviation, shipping and agriculture, 153~Mt of which are fossil-based process emissions -> emissions from power and gas are flexible, so leave out?
-should also say how much is captured and sequestered
-% excess heat potentials -> are not used or taken from hotmaps; also energy demand comes from JRC-IDEES, Hotmaps only used for geographical distribution
 % more detail on heating technologies?
@@
 We show the sum of all energy demands in Fig.~\ref{fig:total-demand-bar}.
-The system emits 633~Mt \COtwo{} per year from industry, aviation, shipping and agriculture, 153~Mt of which are fossil-based process emissions.
-Industrial energy demand and excess heat potentials are calculated per node on the basis of~\cite{hotmaps_industrial_db}.
+The system exogenously produces 633~Mt \COtwo{} per year from industry, aviation, shipping and agriculture, 153~Mt of which are fossil-based process emissions.
+Location for industrial clusters are taken from~\cite{hotmaps_industrial_db}. Energy demand for industries are calculated from~\cite{mantzosJRCIDEES20152018}.
 
@@
 We limit regional biomass use to the medium-level potentials from the JRC-ENSPRESO database~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
-Inter-regional biomass transport is permitted with transport costs of XXX €/MWh/km considered.
+Inter-regional biomass transport is permitted with transport costs of $\sim$0.1~€/MWh/km considered.
 
@@
 
+The topology and capacities of the electricity transmission system are taken from the ENTSO-E transparency map~\cite{wiegmansGridkitExtractEntsoE2016} and selected Ten Year Development Plan (TYNDP) projects.
+The power flow is based on the lossless linearized power flow approximation~.
 The topology of the \COtwo{}, \Htwo{} and gas network is identical to the topology of the electricity network, connecting all neighboring regions.
@@
 No retrofitting of gas pipelines is considered.
-% lossless linearised power flow approximation for electricity network
 % gas network data from SciGrid gas
-% power network data from ENTSO-E transparency map plus selected TYNDP projects as additional expansion options
 
@@
 
-% TODO add sentence distinguishing CU, CT, CS, CC
-
+We argue that it is crucial for a proper modelling to consider CC, CU and CS as separate technologies, rather than grouping them under the often-used but misleading umbrella term ``carbon capture, utilization and storage'' (CCUS), to allow for independent optimization and adequate representation of each function in the carbon management system.
 Our model features three drop-in fuel production technologies for CU: methanation, methanolization, and Fischer-Tropsch synthesis.
@@
 % carbon supply
-
 To supply carbon needed for CS and CU, the system can choose to deploy carbon capture (CC) technologies at various point sources of fossil or biogenic origin (see below), or through DAC facilities.
@@
 For carbon sequestration, we only consider offshore sites as potential sinks.
-We make this choice because offshore sequestration sites typically have larger capacity compared to onshore sequestration sites in saline aquifers and due to concerns over public safety for infrastructure near populated areas~\cite{TODO}.
+We make this choice because offshore sequestration sites typically have larger capacity compared to onshore sequestration sites in saline aquifers and due to concerns over public safety for infrastructure near populated areas.
 Our estimates for carbon sequestration potentials are restrictive, limiting the annual sequestration potential to 25~Mt per region and constraining it further such that this amount could be  25 years.
@@
 \subsection{Subsidizing \carbon{} pipelines}
+\label{sec:subsidy}
 

### 2023-12-15 report/submissions/journal/paper/paper.tex
paper: slight improvements to appendix and title

@@
 
-% \title{Assessing \Htwo{} and \COtwo{} Network Strategies in Europe}
-\title{Competition and Synergies of \Htwo{} and \COtwo{} Networks in Europe}
+\title{Evaluating \Htwo{} and \COtwo{} Network Strategies for the European Energy System}
+% \title{Competition and Synergies of \Htwo{} and \COtwo{} Networks in Europe}
 
@@
 
-\begin{figure}
+\begin{figure}[h!]
     \centering
-    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/sequestration_map.png}
+    \includegraphics[width=.5\linewidth]{baseline/figures/90_nodes/sequestration_map.png}
     \caption{Maximal sequestration potential per offshore region used as input for all models.
@@
 
+\clearpage
 \subsection{Cost comparison across models}
@@
 
-\begin{figure}
+\begin{figure}[h!]
     \centering
-    \includegraphics[width=\linewidth]{full/figures/90_nodes/sankey_diagramm.png}
+    \includegraphics[width=.8\linewidth]{full/figures/90_nodes/sankey_diagramm.png}
     \caption{Sankey diagram of the carbon flows in the \hybridmodel{} in the Net-Zero scenario.}
@@
     \centering
-    \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
+    \includegraphics[width=.8\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
     \caption{Sankey diagram of the carbon flows in the \hybridmodel{} in the Net-Negative scenario.}

### 2023-12-18 report/submissions/journal/paper/paper.tex
strip unneeded comments and text

@@
 The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors.
-While management of both \COtwo{} and \Htwo{}, also refered to as carbon and hydrogen, is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructure effectively interacts with carbon management technologies, including of carbon capture, transport, storage, utilization, and sequestration.
+While management of both \COtwo{} and \Htwo{}, also referred to as carbon and hydrogen, is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructure effectively interacts with carbon management technologies, including of carbon capture, transport, storage, utilization, and sequestration.
 Hydrogen offers an efficient way to transport and store energy over long distances and timeframes, and can be produced from electrolysis or steam methane reforming (SMR), for example.
@@
 
-% [potentially relevant text blobs from old 3.1]
-% CU is preferably deployed in the regions where carbon, hydrogen as well as the inputs and by-product outputs (e.g., waste heat) of the synthesis processes have the lowest cost (or highest value).
-
-% Now we turn to how carbon for CS is sourced by the different models.
-In the \modBase{}, because of the absence of a carbon network, all carbon that is sequestered (200 Mt, the same in each model) needs to be captured in the regions with access to an offshore CS site.
-The \modBase{} deploys all available low-cost point-source CC in these regions, and adds DAC in the sequestration locations with the lowest-cost electricity.
-In other words, the \modBase{} uses the atmosphere as a substitute carbon transport pathway, allowing \CO{} from inland point sources to escape into the atmosphere, and extracting it again from the atmosphere with DAC at the offshore locations.
-
-% In the \modCO{}, the carbon network enables the system to capture carbon from low-cost point sources and transport it to the sequestration sites, trading off CC costs with transport costs.
-
-% In summary, the \COtwo{} network enables the transport of carbon from the lowest-cost CC sources to CU and CS carbon sinks, reducing the need for high-cost DAC.
-% The creation of an \Htwo{} network enables the cost-effective delivery of hydrogen from regions with a surplus of low-cost renewable energy to high-demand industrial centers with exogenous hydrogen demand and a surplus of low-cost CC potentials, reducing the need for DAC and for high-cost electrolysis at the industrial centers.
-% This highlights how both the implementation of a \COtwo{} or an \Htwo{} network unlocks low-cost carbon capture potentials at geographically fixed point sources across the continent.
-
 \subsection*{\Htwo{} network incentivizes distributed carbon utilization}\label{subsec:H2}
@@
     \includegraphics[width=0.98\linewidth]{comparison/single-technologies/figures/90_nodes/balance_map_dedicated.png}
-    \caption{Optimal operation, flows and prices of the carbon (top line) and hydrogen (bottom line) sectors for the \carbonmodel{} (left) and the \hydrogenmodel{} (right) in the net zero scenario.
+    \caption{Optimal production, consumption, flows and prices of the carbon (top line) and hydrogen (bottom line) sectors for the \carbonmodel{} (left) and the \hydrogenmodel{} (right) in the net zero scenario.
 For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
@@
 To find the reason for why the \Htwo{} network achieves lower system costs than the \COtwo{} network, it is crucial to understand the underlying mechanisms and geographical implications of the two transportation systems.
-Fig.~\ref{fig:balance_map} maps the average production, consumption and flow in the carbon (top row) and hydrogen (bottom row) sectors for the \modCO{} (left) and \modH{} (right).
-Each map displays circles indicating the production (top half-circle) and consumption (bottom half-circle) for each region, with lines and arrows representing the average import or export volumes and flow directions between regions.
-The color shading of each region corresponds to the average price of the carrier.
-
 As we have also discussed above, the \modCO{} captures large amounts of carbon from bioenergy and industrial activities that were untapped in the \modBase{}.
@@
 
-% TODO: smoother transition, incl. end of 3.2
-% TODO: add network investment costs plot to Appendix. Mention investment H2 > CO2. (The investment in the hydrogen network is four times the investment in the carbon network.)
 The \modHybrid{} combines the advantages of hydrogen and carbon networks, resulting in a cost reduction of \label{}5.2\% (\label{}41 billion euros) compared to the \modBase{}, which is lower than the cost reductions of the \COgrid{} and \modH{} (see Fig.~\ref{fig:cost_bar}).
 In particular, production, consumption and transport in the carbon and hydrogen sectors in the \modHybrid{} combine elements of the \carbonmodel{} and \hydrogenmodel{} (see Fig.~\ref{fig:capacity_maps}).
-While the topology of the \Htwo{} network roughly corresponds to that of the \modH{}, the \COtwo{} network plays more of a complementary role.
+While the topology of the \Htwo{} network (8~bn€ investments) roughly corresponds to that of the \modH{}, the \COtwo{} network plays more of a complementary role (3~bn€ investments, see Fig.~\ref{fig:cost_bar_transmission}).
 In this section, we explore the advantages and disadvantages compared to the other models.
@@
 Each of the regions consists of a complex subsystem with technologies for supplying, converting, storing and transporting energy.
-Exogenous assumptions on energy demand and non-abatable emissions are taken from various sources~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023}.
+Exogenous assumptions on energy demand and non-abatable emissions are taken from various sources~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023} (see Fig.~\ref{fig:total-demand-bar}).
 The energy demand for electricity, transport, biomass, heat and gas are defined per region and time-step.
@@
 Heat demand is regionally subdivided into shares of urban, rural and industrial sites.
-We show the sum of all energy demands in Fig.~\ref{fig:total-demand-bar}.
 The system exogenously produces 633~Mt \COtwo{} per year from industry, aviation, shipping and agriculture, 153~Mt of which are fossil-based process emissions.

### 2023-12-20 report/submissions/journal/paper/paper.tex
paper: tidy up
incorp reviews, restructure figures

@@
 \newcommand{\hybridmodel}{Hybrid model}
+\newcommand{\baselinescenario}{Baseline scenario}
+\newcommand{\carbonscenario}{CO$_2$\=/Grid scenario}
+\newcommand{\hydrogenscenario}{H$_2$\=/Grid scenario}
+\newcommand{\hybridscenario}{Hybrid scenario}
 
@@
 \author{
-    Fabian Hofmann, Christoph Tries, Fabian Neumann, Lisa Zeyen, Tom Brown \\
+    Fabian Hofmann, Christoph Tries, Fabian Neumann, Elisabeth Zeyen, Tom Brown \\
     \textit{Institute of Energy Technology} \\
@@
 \begin{abstract}
-    Hydrogen and carbon dioxide transportation are considered crucial in climate-neutral energy systems, with hydrogen transport enabling energy imports to high-demand areas and carbon transport aiding in emissions export from high-emission areas.
-Yet, possible synergies and competitions between the two systems are not fully understood.
-Our study employs optimization techniques to develop a cost-optimal European energy system, integrating production, transport and storage of both carbon dioxide and hydrogen, along with renewable energy sources.
-Results indicate that, in direct comparison, a hydrogen network is more cost-effective than a carbon network, facilitating low-cost hydrogen transport to Central Europe and local carbon capture and utilization.
-However, in a hybrid scenario, the carbon network effectively complements the hydrogen network, promoting carbon capture from biomass and reducing reliance on direct air capture.
-This research demonstrates the effectiveness of combining hydrogen and carbon transport networks with power grid expansion in achieving climate neutrality in Europe.
+    Hydrogen and carbon dioxide transport can both play an important role in climate-neutral energy systems.
+Hydrogen networks help serve regions with high energy demand, while excess emissions are transported away in carbon dioxide networks.
+When it comes to the synthesis of carbonaceous fuels, it is less clear which input should be transported: hydrogen to carbon point sources or carbon to low-cost hydrogen.
+We explore the potential synergies and competition of both networks in a cost-optimal carbon-neutral European energy system.
+In direct comparison, a hydrogen network is more cost-effective than a carbon network, since it serves to transport hydrogen to demand and to point source of carbon for utilization.
+However, in a hybrid scenario where both networks are present, the carbon network effectively complements the hydrogen network, promoting carbon capture from biomass and reducing reliance on direct air capture.
+This study shows how the European energy system in future net-zero and net-negative emissions scenarios can benefits from integrating both networks.
 \end{abstract}
@@
 
-The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors.
-While management of both \COtwo{} and \Htwo{}, also referred to as carbon and hydrogen, is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructure effectively interacts with carbon management technologies, including of carbon capture, transport, storage, utilization, and sequestration.
-Hydrogen offers an efficient way to transport and store energy over long distances and timeframes, and can be produced from electrolysis or steam methane reforming (SMR), for example.
-Carbon can be stored in geological formations to permanently extract it from the atmosphere, a process known as carbon sequestration (CS).
-Additionally, carbon can be combined with hydrogen to produce carbonaceous fuels as for example needed for aviation in form of synthetic kerosene from Fischer-Tropsch synthesis, for shipping in form of synthetic methanol and for industrial feedstocks through synthetic methane.
+The transition to a climate-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors.
+While management of both carbon dioxide (\COtwo{}) and hydrogen (\Htwo{}) is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructure effectively interacts with carbon management technologies, including carbon capture, transport, storage, utilization, and sequestration.
+Hydrogen is consider in several sectors in industry not amenable to electrification, and offers an efficient way to transport and store energy over long distances and timeframes, and can be produced from electrolysis or steam methane reforming (SMR), for example.
+Captured carbon can be stored in geological formations to permanently extract it from the atmosphere, a process known as carbon sequestration (CS).
+Additionally, carbon can be combined with hydrogen to produce carbonaceous fuels to supply for example synthetic kerosene for aviation, synthetic methanol for shipping or synthetic methane for industrial feedstocks.
 From hereon, we jointly refer to the production of these carbonaceous fuels as Carbon Utilization (CU).
@@
 
-Recently, policymakers and industry in Europe have started developing hydrogen and carbon management strategies~\cite{GermanyDevelopingStrategy2023,CarbonManagementStrategie}, planning infrastructure components~\cite{CONetz}, and committing on the first carbon utilization projects~\cite{EFuelsPilotPlant2022,OrstedAssumesFull,GROUNDBREAKINGEFUELPRODUCTION,DLREfuelsDLR}.
+Recently, policymakers and industry in Europe have started developing hydrogen and carbon management strategies~\cite{GermanyDevelopingStrategy2023,CarbonManagementStrategie}, planning infrastructure components~\cite{CONetz}, and committing to the first carbon utilization projects~\cite{EFuelsPilotPlant2022,OrstedAssumesFull,GROUNDBREAKINGEFUELPRODUCTION,DLREfuelsDLR}.
 With the European Union's goal of achieving climate neutrality by 2050, a wide range of programs, funding models and initiatives have been established with dedicated instruments supporting both the ramp-up of the hydrogen and carbon economy~\cite{eu2023netzero,europeangreendeal,europeaninnovationfund}.
 Collaborative industry initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen Infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier.
-Some natural gas pipelines are already repurposed to transport hydrogen~\cite{RohrFreiFuer}.
+Some natural gas pipelines have already been repurposed to transport hydrogen~\cite{RohrFreiFuer}.
 Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synthetic fuels on the one hand and offer \COtwo{} offtake on the other hand.
@@
 To advance the carbon economy, the Clean Air Task Force emphasizes the importance of developing a carbon transport system in Europe alongside a hydrogen network~\cite{lockwoodEuropeanStrategyCarbon}.
-Carbon pipelines, considered a mature technology, have seen widespread installations in the United States and Canada, primarily for enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
+Carbon pipelines, considered a mature technology, have seen widespread installations in the United States and Canada, primarily to supply \carbon{} to enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
 
-However, up to this point, it remains unclear how the two transport systems of hydrogen and carbon may complement or replace each other.
-Both can bridge the misalignment of sources and sinks for carbon and hydrogen which may well be determined by exogenous factors.
+Up to this point, it remains unclear how the two transport systems of hydrogen and carbon may complement or even compete with each other.
+Both can bridge the misalignment of sources and sinks for carbon and hydrogen.
 A hydrogen network can supply regions with geographically fixed hydrogen demand, such as for steelmaking, with hydrogen from regions with the best renewable resources as well as enable CU at the site of point-source CC.
 On the other hand, the carbon transport system allows for transporting captured carbon to regions with sequestration potentials or high-quality renewable resources, the latter enabling cost-effective CU.
-In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022,neumannBenefitsHydrogenNetwork2022}.
+
+In the literature, the two network approaches and underlying technologies have been discussed in several of publications, all of which, however, dealt with the isolated aspects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022,neumannBenefitsHydrogenNetwork2022}.
 Such techno-economic models, in contrast for example to Integrated Assessment Models, can account for the spatial distribution of carbon sources and sinks which are crucial to provide a holistic view of the energy system and its technological interactions.
-A remarkable example was developed by Neumann et al.~\cite{neumannBenefitsHydrogenNetwork2022}, which comprehensively demonstrates the interchangeability of hydrogen and electricity grid expansion in a climate-neutral European energy system at the expense of higher investments.
-The underlying, highly-resolved model encompasses most relevant sectors, however neglects the option to transport carbon across regions.
-The publication by Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \COtwo{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
-This limited sectoral scope cannot capture important dynamics of carbon management, since it neglects the sectors that will need to handle most \COtwo{} in the future.
+Neumann et al.~\cite{neumannBenefitsHydrogenNetwork2022} demonstrated the interchangeability of hydrogen and electricity grid expansion in a climate-neutral European energy system at the expense of higher investments.
+The underlying, highly-resolved model encompasses most relevant sectors, however it neglects the option to transport carbon across regions.
+Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \COtwo{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
+This limited sectoral scope cannot capture important dynamics of carbon management, since it neglects the sectors like industry that will need to handle most \COtwo{} in the future.
 A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \COtwo{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
@@
 
-To our knowledge, no study has yet considered the co-optimization and comprehensive assessment of both \COtwo{} and \Htwo{} networks in a fully sector-coupled energy system.
-However, we argue that such an assessment is strongly needed to identify realms of competition and synergy between hydrogen and carbon management technologies.
+No study has yet considered the co-optimization and comprehensive assessment of both \COtwo{} and \Htwo{} networks in a fully sector-coupled energy system.
+However, such an assessment is strongly needed to identify realms of competition and synergy between hydrogen and carbon management technologies.
 In this paper, we present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies.
-The study is conducted using the PyPSA-Eur energy system model which optimizes operation and investments in all relevant energy sectors.
-To investigate the competing transport dynamics between \COtwo{} and \Htwo{} networks, we contrast the following four models that have none of these networks, only one, or both combined:
+The study is conducted using the PyPSA-Eur energy system model which optimizes operation and investments in all relevant energy sectors to supply projected energy demands for 2050 (see Fig.~\ref{fig:total-demand-bar}).
+It is assumed that Europe is self-sufficient in energy and does not import any fuels.
+This drives carbonaceous fuels production to be located at sites where both \carbon{} and \hydrogen{} can be provided at low cost.
+
+\begin{figure}[h!]
+    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/demand_bar.png}
+    \caption{Assumptions on exogenous demand, derived from~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023}.
+The figure shows the total annual energy demand for each energy source which determine the model's endogenous investments and operation.
+Endogenous processes can lead to higher total production volumes of some energy carriers, e.g., the demand for methanol requires more hydrogen and carbon as secondary (energy) inputs, which are not considered here.
+The model are defined per region and time stamp.}
+    % TODO: adjust labels to show that these are exogenous assumptions
+    \label{fig:total-demand-bar}
+\end{figure}
+
+% TODO: This drives all CCU....
+% TODO: The demands should be mentioned properly introduced
+
+To investigate the competing transport dynamics between \COtwo{} and \Htwo{} networks, we contrast the following four model scenarios:
 %
@@
 %
-In anticipation of the future \carbon{} emission targets, the models are subjected to the following two scenarios:
-\begin{itemize}
-    \item[] \textit{Net\=/Zero}: aligning with the EU's 2050 emission targets. Limiting carbon sequestration to 200~Mt/a. This scenario is also denoted as NZ.
-    \item[] \textit{Net\=/Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \COtwo{} annually. Limiting carbon sequestration to 660~Mt/a. Also denoted as NN.
-\end{itemize}
+% In anticipation of the future \carbon{} emission targets, the models are subjected to the following two scenarios:
+% \begin{itemize}
+%     \item[] \textit{Net\=/Zero}: aligning with the EU's 2050 emission targets. Limiting carbon sequestration to 200~Mt/a. This scenario is also denoted as NZ.
+%     \item[] \textit{Net\=/Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \COtwo{} annually. Limiting carbon sequestration to 660~Mt/a. Also denoted as NN.
+% \end{itemize}
 
 We analyze how the deployment of carbon and hydrogen networks affects different decarbonization strategies in the models and scenarios and how investments in none or only one of the transport technologies lead to inefficient market layouts.
-We do this by first focusing on the net-zero scenario and a detailed description of the different technological and geographical impacts of the transportation systems.
-Subsequently our analysis demonstrates the extent to which different deployment strategies remain robust under a tightened emissions target in the Net-Negative scenario.
+We do this by focusing first on a net-zero \carbon{} emissions target, while limiting carbon sequestration to a total 200~Mt/a.
+We give a detailed description of the different technological and geographical impacts of the transportation systems.
+Subsequently, our analysis demonstrates the extent to which different deployment strategies remain robust under a tightened net-negative \carbon{} emissions target.
 
@@
 
-Focusing first on the Net-Zero scenario, we can highlight notable characteristics of the energy system buildouts shared by all models, as well as significant variations in system costs and technology deployment between the models (see Fig.~\ref{fig:cost_bar}).
+All scenarios share key energy system characteristics but differ in system costs and technology use (Fig.~\ref{fig:cost_bar}).
+The \modBase{} has the highest annual cost at 779~bn€/a.
+Models with \carbon{} and \hydrogen{} transport options are more cost-effective, reducing costs by 3.1\% in \modCO{}, 4.8\% in \modH{}, and 5.2\% in \modHybrid{}.
+% in system costs: 779bn€ (baseline), 754bn€ (carbon), 742bn€ (hydrogen), 738bn€ (hybrid)
 
@@
 
-Total annual system costs in the \modBase{} are highest, totalling \label{}779~bn€.
-The models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs, with \label{}3.1\% cost reduction (\label{}754~bn€) in the \modCO{}, and \label{}4.8\% (\label{}742~bn€) in the \modH{}, and \label{}5.2\% (\label{}738~bn€) in the the \modHybrid{}.
+% In every model, nearly half the costs are attributable to wind and solar electricity production, and about 10\% to hydro, nuclear, and biomass.
+In the \baselinemodel, transmission costs (electricity, gas, hydrogen, carbon) are below 4\% (25 bn€) and are dominated by electricity grid expenses.
+In contrast, the other models have higher combined pipeline costs, peaking at 20\% in the \hybridmodel.
+Both carbonaceous fuel production and electrolysis account for about 4\% (30 bn€) and 4.7\% (37 bn€) of costs respectively, with heat pump installations contributing around 8\% (64 bn€ in the \baselinemodel).
 
-In all models, almost half of the system costs are spent on primary electricity production from wind and solar power, with around another \label{}10\% spent on hydro, nuclear and biomass.
-For the \baselinemodel, transmission system costs for electricity, gas, hydrogen and carbon are below \label{}4\% (\label{}25~bn€), with the electricity grid making up more of the system costs than the other three networks combined.
-This is not the case for the other models, where combined pipeline costs slightly surpass electricity transmission system costs (up to 20\% for the \hybridmodel).
-Capital expenditures on carbonaceous fuel production are around \label{}4\% (\label{}30~bn€) in each model, while electrolysis makes up around \label{}4.7\% (\label{}37~bn€).
-Heat pump installations account for around \label{}8\% of total system costs (64~bn€ in the \baselinemodel).
-
-System cost differences between the models are driven by renewable energy resources (solar, wind, and bioenergy), the \COtwo{} and \Htwo{} networks, as well as carbon capture technologies (DAC and CC at point sources).
-As we show below, these are due to substantial, albeit fundamentally different, gains in flexibility and a shift in carbon utilization and sequestration strategies by the \Htwo{} and \COtwo{} networks.
-
-\subsection*{Transport systems unlock point\=/source carbon capture}
-
-In comparison to the other three models, the \modBase{} reveals significant inefficiencies: On the one hand, regions with no access to sequestration sites (i.e., ``landlocked'' regions) underutilize CC potentials from process emissions, fossil gas and biomass.
-On the other hand, sequestration regions rely on costly DAC on-site.
-Furthermore, hydrogen and power prices show high regional differences, indicating system inefficiencies due to transport constraints caused by non\=/existent hydrogen or insufficient power transmission.
-Prices are significantly higher in Central Europe, a region with high population density and high concentration of energy-intensive industries.
-
-By enabling the efficient transportation of carbon, hydrogen, or both, these models reduce the reliance on high-cost DAC, and source more carbon from low-cost point-source CC on biomass and gas in industrial applications (see Fig.~\ref{fig:balance_captured_carbon}).
-The \modH{} only captures \label{}193~Mt of \COtwo{} from DAC compared to \label{}392~Mt in the \modBase{}.
-Instead, the \modH{} captures more carbon from process emissions as well as gas and biomass for industry.
-The \modCO{} uses only \label{}154~Mt from DAC, and instead also captures more carbon than the \modH{} from gas for industry and biomass CHPs.
-And finally, the \modHybrid{} captures only \label{}112~Mt from DAC, but slightly more carbon from all other CC technologies.
-Also, the transport of \carbon{} and \hydrogen{} reduce the reliance on methanation and therefore the overall demand for CC volumes in the model.
+System cost differences between the models are driven by wind and solar as well as carbon capture technologies.
+In scenarios with \carbon{} and \hydrogen{} transportation, the model primarily reduces the reliance on high-cost DAC from 392~Mt/a carbon removal in the \baselinescenario{} to 112~Mt/a in the \hybridscenario{} (see Fig.~\ref{fig:balance_captured_carbon}).
+At the same time, carbon capture from biomass combustion, gas-based industry processes and biogas-to-gas upgrading increase.
+The increased usage of upgraded biogas leads to a lower synthetic methane production.
+% At the same time, the transport of \carbon{} and \hydrogen{} reduce the reliance on methanation and therefore the overall demand for CC volumes in the model.
 Another knock-on effect of less reliance on DAC is its reduced endogenous demand for dedicated renewables as well as heat pumps and biomass CHPs to supply power and heat.
-
 \begin{figure}[ht!]
@@
 
-The increase in captured carbon from point sources in the models with \carbon{} and/or \hydrogen{} transport is more related to an upgrading of existing industrial or power plant facilities to capture carbon rather than a shift in technologies.
+In the \baselinescenario{}, where there is no carbon and hydrogen transport, hydrogen must be produced where it is consumed.
+For CU, either \hydrogen{} must be produced where \carbon{} is cheap at point sources, or DAC must be used at sites with low-cost \hydrogen{}.
+Where \carbon{} from point sources is not able to be sequestered, it has to be compensated elsewhere by negative emissions from DAC or biomass capture with sequestration.
+All of these options are sub-optimal.
+
+\subsection*{Networks unlock low-cost \hydrogen{} and \carbon{}}
+
+Both a carbon network and a hydrogen network can help alleviate these inefficiencies.
+A carbon network allows \carbon{} from point sources to be transported to low-cost hydrogen as well as sequestration sites.
+A hydrogen network on the other hand allows low-cost hydrogen to be transported from regions with good renewable resources to point sources for CU.
+This results in fundamentally different network flows in each scenario (see Fig.~\ref{fig:balance_map}).
+
+\begin{figure*}[ht!]
+    \centering
+    \includegraphics[width=0.98\linewidth]{comparison/single-technologies/figures/90_nodes/balance_map_dedicated.png}
+    \caption{Average production, consumption, flows and prices of the carbon (top line) and hydrogen (bottom line) sectors in the \carbongrid{} (left) and the \hydrogenscenario{} (right).
+For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
+Lines and arrows show the interregional transportation.
+Carbon sequestration offshore are drawn as full circles.
+    }
+    \label{fig:balance_map}
+\end{figure*}
+
+In the \carbonscenario, there are three major transport routes of the low-cost carbon captured in Central and Eastern Europe (see Fig.~\ref{fig:balance_map}, top left).
+A large part of the carbon is transported from Western Europe to the Iberian Peninsula (8~kt/h), the British Isles (7.5~kt/h) and Denmark (2.8~kt/h), where it is used along with low-cost electrolytic hydrogen to produce Fischer-Tropsch fuels and methanol.
+A second part is transported to methanation plants in Italy and Greece (2.8~kt/h), where the synthetic methane produced is fed into the gas grid in the vicinity of major gas import hubs.
+Finally, a third part of captured carbon is transported directly to sites in the North Sea, Baltic Sea and Mediterranean Sea where it is sequestered.
+Almost none of the captured carbon is used locally at the point-source.
+In addition to electrolysis in regions with the good renewable resources to supply CU ($\sim$125~GW$_\text{el}$), the model places smaller amounts of electrolyzers where hydrogen is needed for industrial processes across Central Europe ($\sim$8.5~GW$_\text{el}$) (see bottom left).
+Hydrogen prices vary strongly across the continent, with low price ``valleys'' where the CU locates with around 60~€/MWh, and high price ``peaks'' in Central Europe with around 95~€/MWh.
+Despite the transport system, carbon prices still vary from 110~€/t in Central Europe to 160~€/t at sequestration sites.
+
+In the \hydrogenscenario{}, the main hydrogen transportation routes go from Spain (50~GW) and the United Kingdom (18~GW) to Central Europe to supply local carbonaceous fuel production with captured carbon from point sources (see Fig~\ref{fig:balance_map}, bottom right).
+Also, the \Htwo{} network supplies spatially-fixed \Htwo{} demand for industry in Central Europe (2.5~GW in total) at much lower costs than local electrolysis or SMR.
+Carbon sequestration primarily locates in Portugal where large DAC facilities operate (14~kt/h, see Fig~\ref{fig:balance_map}, top right).
+\carbon{} prices are low in Central Europe (110-120~€/t) and highest at sequestration regions (160~€/t).
+Hydrogen prices have flattened out at a relatively low level of €60~/MWh in all regions.
+
+In both cases, CU is placed where material inputs, \carbon{} and \hydrogen{}, are provided at minimal cost, i.e., in price valleys of hydrogen and carbon in the \carbongrid{} and \hydrogenscenario{} respectively.
+In the \carbonscenario{}, CU co-locates with low-cost hydrogen production and uses transported carbon from the inland.
+In the \hydrogenscenario{}, CU co-locates with carbon point sources and uses transported low-cost hydrogen.
+Both scenarios effectively combine low-cost renewable hydrogen production and carbon management for combustion and spatially-fixed industrial processes.
+
+However, both transport systems provide individual cost-benefits to the system.
+While the carbon grid makes it possible to transport captured carbon for sequestration at low cost, the dependence on DAC near the coast causes higher costs in the hydrogen scenario{}.
+On the other hand, the \hydrogengrid{} is able to supply the spatially-fixed hydrogen demands in Central Europe with low-cost hydrogen imports, while the \modCO{} is forced to deploy electrolysis in regions with poor renewable resources as well as additional gas infrastructure in Central Europe.
+The latter aspect dominates the first, leading to lower system cost in the \hydrogenscenario{}.
+
+% To understand the dynamics of the \modCO{}, we observe that the system locates CU in the regions where it can minimize costs for its material inputs, carbon and hydrogen.
+As the \COtwo{} network largely smoothes out carbon prices across regions, CU is located in the hydrogen price valleys on the periphery of the continent.
+Similarly, the model locates CS in regions where it can minimize the costs of the carbon to be sequestered.
+Thus, CS is located in the carbon price valleys among the subset of regions with access to offshore sequestration sites.
+Due to the \COtwo{} network these low-price regions are not strongly pronounced, but still prevalent and impact CS locations.
+DAC is only used for CU and thus co-located in regions with a mix of low electricity costs (for running DAC) and low hydrogen costs.
+
+% To explain the dynamics of the \modH{}, we observe that the model locates CU in the carbon price valleys across Western, Central and Eastern Europe (see top right), as the \Htwo{} network smoothes out hydrogen prices across the entire continent (see bottom right).
+CS is again located in the carbon price valleys among those regions with access to offshore sequestration sites.
+Due to the lack of a \COtwo{} network the full CC potentials from point sources are exhausted in all offshore sequestration regions, and the carbon is then sequestered.
+The additionally needed carbon to reach 200~Mt of CS is captured via DAC in the regions with access to offshore sequestration sites with the lowest costs for DAC, with the largest share deployed in Portugal (see top right).
+DAC is also used for CU in this model, and for both purposes DAC is located in the regions with the lowest costs for electricity.
+
+% The increase in captured carbon from point sources in the models with \carbon{} and/or \hydrogen{} transport is more related to an upgrading of existing industrial or power plant facilities to capture carbon rather than a shift in technologies.
 This is evident when analyzing the capture share of the different CC technologies (see Fig.~\ref{fig:captureshare_line}).
+
+% In summary, the implementation of \COtwo{} and \Htwo{} networks is instrumental in unlocking CC potentials and reducing system costs.
+However, despite the \modCO{} demonstrating higher amounts and shares of CC from point sources and therefore less DAC deployment, the \modH{} achieves even  lower total system costs (see Fig.~\ref{fig:cost_bar}).
+
+\subsection*{Hybrid configuration provides further flexibilities}\label{subsec:Hybrid}
+
+\begin{figure*}[ht!]
+    \centering
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{full/figures/90_nodes/balance_map_carbon.png}
+        \label{fig:balance_map_carbon_full}
+    \end{subfigure}%
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{full/figures/90_nodes/balance_map_hydrogen.png}
+        \label{fig:balance_map_hydrogen_full}
+    \end{subfigure}
+    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hybridmodel{} in the net zero scenario.
+For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
+Carbon Sequestration offshore are drawn as full circles.
+Lines and arrows show the interregional transportation.
+\carbon{} from point-source in the inland either supplies local CU with imported \hydrogen{} or facilitates sequestration in nearby offshore regions.
+    % Carbon network looks the same as in~\cite{morbeeOptimisedDeploymentEuropean2012}: two backbones, one in the nothern Europe other in south east.
+    }
+    \label{fig:balance_map_full}
+\end{figure*}
+
+The \hybridscenario{}, the model combines the advantages of hydrogen and carbon networks, resulting in the highest system cost reduction of \label{}5.2\% (\label{}41~bn€/a) compared to the \modBase{}.
+As in the \hydrogenscenario{}, the model transports low-cost hydrogen to the point sources in Central Europe to supply CU, but also transports low-cost carbon from point sources close the shore to sequestration sites (see Fig.~\ref{fig:balance_map_full}).
+This reduces reliance on high-cost DAC and provides low-cost hydrogen across the regions.
+
+For the \carbon{} network, the model employs short and direct pipeline routes to transport carbon from inland point sources to nearby sequestration sites, as for example in Northern Germany, Italy and Greece (see Fig.~\ref{fig:balance_map_full}, left).
+Additionally, the \COtwo{} network balances out small local variations in carbon demand and supply to produce CU, as for example in the middle of France.
+The hydrogen routes are similar to those in the \hydrogenscenario{} with transports from Western regions (Iberian Peninsula and British Isles) to distributed CU production sites across the continent (see Fig.~\ref{fig:balance_map_full}, right).
+While the topology of the \Htwo{} network (8~bn€ investments) roughly corresponds to that of the \modH{} primarily driven by CU, the \COtwo{} network plays more of a complementary role primarily driven by CS (3~bn€ investments, see Fig.~\ref{fig:cost_bar_transmission}).
 
@@
 
-In the \modBase{}, the capture share is highest for biogas-to-gas facilities, followed by process emissions, biomass for industry, biomass CHPs, gas for industry, SMR and gas CHPs.
-For models with additional transport systems, we observe a correlation between CC share and capacity factors of the underlying technologies: Process emissions as well as gas and biomass for industry have high capacity factors (above 80\%), and CC shares close to 100\%, the only exception being gas for industry in the \modH{}.
-Biomass CHPs have a capacity factor of around 50\%, and CC shares of 50\% in the \modH{} and 97\% in the \modCO{}.
-SMR and gas CHPs have capacity factors around 15\%, and CC shares at nearly 0\% for the \baselinemodel.
-The capacity factor and CC share of SMR increases to around 20\% and 50-65\%, respectively, in the other models.
-Gas CHPs serve as peak load electricity production technology, and with very few operating hours, thus the high investment costs into CC applications are not efficient.
-Similarly, SMR serves as ``peak-load'' hydrogen production technology in regions with poor renewable resources for electrolysis.
+In this combination, the networks enable full carbon capture upgrades for all technologies that run on a high average capacity factor (above 70\%, see Fig.~\ref{fig:captureshare_line}).
+Except for SMR and gas-fired Combined Heat and Power Plants (CHP) which operate a few weeks annually, all emissions from carbon-emitting technologies are captured.
+% In comparison with the \hydrogenscenario, Biomass CHP supplies more flexible heat, power and relatively low-cost carbon.
+Thus, only a few Direct Air Capture (DAC) installations are installed, such as those in southern Spain for methanolization plants.
+This leads to reduced power consumption and, therefore, less required wind and solar power deployment.
 
-In summary, the implementation of \COtwo{} and \Htwo{} networks is instrumental in unlocking CC potentials and reducing system costs.
-However, despite the \modCO{} demonstrating higher amounts and shares of CC from point sources and therefore less DAC deployment, the \modH{} achieves even  lower total system costs (see Fig.~\ref{fig:cost_bar}).
-
-\subsection*{\Htwo{} network incentivizes distributed carbon utilization}\label{subsec:H2}
-
-\begin{figure*}[ht!]
-    \centering
-    \includegraphics[width=0.98\linewidth]{comparison/single-technologies/figures/90_nodes/balance_map_dedicated.png}
-    \caption{Optimal production, consumption, flows and prices of the carbon (top line) and hydrogen (bottom line) sectors for the \carbonmodel{} (left) and the \hydrogenmodel{} (right) in the net zero scenario.
-For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
-Lines and arrows show the interregional transportation.
-Carbon sequestration offshore are drawn as full circles.
-\carbon{} and \hydrogen{} transport goes from low price to high price areas to supply CU and, in the case of \carbonmodel{}, CS.
-    }
-    \label{fig:balance_map}
-\end{figure*}
-
-To find the reason for why the \Htwo{} network achieves lower system costs than the \COtwo{} network, it is crucial to understand the underlying mechanisms and geographical implications of the two transportation systems.
-As we have also discussed above, the \modCO{} captures large amounts of carbon from bioenergy and industrial activities that were untapped in the \modBase{}.
-In particular, the carbon capture takes place in Central and Eastern Europe at relatively low cost, and almost none of the carbon is used locally (see Fig.~\ref{fig:balance_map}, top left).
-A large part of the captured carbon is transported from Western Europe to the Iberian Peninsula (70~Mt/h), the British Isles (66~Mt/h) and Denmark (25~Mt/h), where it is used along with low-cost electrolytic hydrogen to produce Fischer-Tropsch fuels and methanol.
-A second part of the captured carbon is transported to methanation plants in Italy and Greece (25~Mt/h), where the synthetic methane produced is fed into the well-developed gas grid, as Italy serves as a major gas import hub for Europe.
-Finally, a third part of captured carbon is transported directly to sites in the North Sea, Baltic Sea and Mediterranean Sea where it is sequestered.
-In absence of an \Htwo{} network, the system places most electrolysis in regions with the best renewable resources to supply local CU ($\sim$1100~GW$_\text{el}$), and smaller amounts of electrolyzers where hydrogen is needed for industrial processes across Central Europe ($\sim$75~GW$_\text{el}$).
-The model also emphasizes gas imports into Central Europe.
-Hydrogen prices vary strongly across the continent, with low price ``valleys'' in regions with good renewable resources such as the Iberian Peninsula or Denmark with around 60~€/MWh, and high price ``peaks'' in Central Europe with around 95~€/MWh (see bottom left).
-
-To understand the dynamics of the \modCO{}, we observe that the system locates CU in the regions where it can minimize costs for its material inputs, carbon and hydrogen.
-As the \COtwo{} network largely smoothes out carbon prices across regions, CU is located in the hydrogen price valleys on the periphery of the continent.
-Similarly, the model locates CS in regions where it can minimize the costs of the carbon to be sequestered.
-Thus, CS is located in the carbon price valleys among the subset of regions with access to offshore sequestration sites.
-Due to the \COtwo{} network these low-price regions are not strongly pronounced, but still prevalent and impact CS locations.
-DAC is only used for CU and thus co-located in regions with a mix of low electricity costs (for running DAC) and low hydrogen costs.
-
-The \Htwo{} network in the \modH{} on the other hand, transports low-cost hydrogen from regions with abundant renewable resources to regions with low-price carbon from abundant sources of fossil, biogenic and process emissions (see bottom right).
-The main transportation routes go from Spain (430~GW) and the United Kingdom (155~GW) to Central Europe to produce carbonaceous fuels with locally captured carbon from point sources.
-Also, the \Htwo{} network supplies exogenous \Htwo{} demand for industry in Central Europe (22~GW in total) at much lower costs than local electrolysis or SMR.
-
-To explain the dynamics of the \modH{}, we observe that the model locates CU in the carbon price valleys across Western, Central and Eastern Europe (see top right), as the \Htwo{} network smoothes out hydrogen prices across the entire continent (see bottom right).
-CS is again located in the carbon price valleys among those regions with access to offshore sequestration sites.
-Due to the lack of a \COtwo{} network the full CC potentials from point sources are exhausted in all offshore sequestration regions, and the carbon is then sequestered.
-The additionally needed carbon to reach 200~Mt of CS is captured via DAC in the regions with access to offshore sequestration sites with the lowest costs for DAC, with the largest share deployed in Portugal (see top right).
-DAC is also used for CU in this model, and for both purposes DAC is located in the regions with the lowest costs for electricity.
-
-In summary, both network models prefer CC from point sources distributed across the continent as well as electrolysis concentrated in regions with abundant renewable resources.
-The two network approaches reveal fundamentally different strategies to integrate these low-cost resources into the system: The \carbonmodel{} transports and sequesters low-cost carbon from distributed point sources and co-locates CU with low-cost electrolysis and DAC.
-The \hydrogenmodel{} distributes CU alongside with carbon point sources to which low-cost hydrogen from regions with abundant renewable resources is transported.
-%
-% However, differences between the two models are notable for the placement of CU, CS and DAC due to their distinct \COtwo{} and \Htwo{} price patterns.
-CU locations are more distributed in the \modH{}, while they cluster in a few valleys of low-cost \Htwo{} in the \modCO{}.
-CS co-locates in the regions with the lowest \COtwo{} costs (among those with offshore sequestration potential), but these regions are very different between the models.
-DAC is placed in regions with the best renewable resources in both models, but it is used for only CU in the \modCO{} and for both CU and CS in the \modH{}.
-Finally, the system cost advantage of the \modH{} stems from being able to supply the exogenous hydrogen demands in Central Europe with low-cost hydrogen imports, while the \modCO{} is forced to deploy electrolysis in regions with poor renewables resources as well as additional gas infrastructure in Central Europe.
-
-\subsection*{Hybrid configuration provides further flexibilities}\label{subsec:Hybrid}
-
-\begin{figure*}[ht!]
-    \centering
-    \begin{subfigure}{.5\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{full/figures/90_nodes/balance_map_carbon.png}
-        \label{fig:capacity_map_carbon_co2}
-    \end{subfigure}%
-    \begin{subfigure}{.5\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{full/figures/90_nodes/balance_map_hydrogen.png}
-        \label{fig:capacity_map_hydrogen_co2}
-    \end{subfigure}
-    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hybridmodel{} in the net zero scenario.
-For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
-Carbon Sequestration offshore are drawn as full circles.
-Lines and arrows show the interregional transportation.
-\carbon{} from point-source in the inland either supplies local CU with imported \hydrogen{} or facilitates sequestration in nearby offshore regions.
-    % Carbon network looks the same as in~\cite{morbeeOptimisedDeploymentEuropean2012}: two backbones, one in the nothern Europe other in south east.
-    }
-    \label{fig:capacity_maps}
-\end{figure*}
-
-The \modHybrid{} combines the advantages of hydrogen and carbon networks, resulting in a cost reduction of \label{}5.2\% (\label{}41 billion euros) compared to the \modBase{}, which is lower than the cost reductions of the \COgrid{} and \modH{} (see Fig.~\ref{fig:cost_bar}).
-In particular, production, consumption and transport in the carbon and hydrogen sectors in the \modHybrid{} combine elements of the \carbonmodel{} and \hydrogenmodel{} (see Fig.~\ref{fig:capacity_maps}).
-While the topology of the \Htwo{} network (8~bn€ investments) roughly corresponds to that of the \modH{}, the \COtwo{} network plays more of a complementary role (3~bn€ investments, see Fig.~\ref{fig:cost_bar_transmission}).
-In this section, we explore the advantages and disadvantages compared to the other models.
-
-Similar to the \modH{}, the \modHybrid{} transports hydrogen from concentrated electrolysis sites in Western regions (Iberian Peninsula and British Isles) to distributed CU production sites across the continent.
-For the \COtwo{} network, this model employs short and direct pipeline routes, typically spanning one or two segments, to transport carbon from inland CC point sources to perform CS at nearby sequestration sites (e.g., Northern Germany, Italy and Greece).
-Additionally, the \COtwo{} network utilizes single-segment network sections to equalize local variations in carbon demand and supply to produce CU, thereby also further converging carbon prices.
-The model locates CU everywhere, except for a few coastal regions that perform CS, or that supply their locally captured carbon to a neigbouring sequestration region.
-DAC is deployed to supply carbon for the remaining CU needed by the model (mostly in Southern Spain).
-
-Despite only a further reduction of \label{}0.4 percentage points (\label{}3~bn€) in total system costs compared to the \modH{}, the \modHybrid{} brings several advantages: (1) The \modHybrid{} relies even less on DAC than the other models, which as a technology may be costlier than anticipated in the cost projections, by making use of carbon point sources; (2) it requires a smaller land use footprint for wind and solar power due to lower DAC deployment.
-On the other hand, the \modHybrid{} needs slightly more biomass, equipped with carbon capture, to achieve its cost advantage in comparison to the \modH{}; (3) it uses both \COtwo{} and \Htwo{} networks to flatten price distribution across the continent and to build a more robust energy system.
-
-\subsection*{Net-negativity amplifies DAC and \carbon{} transport}\label{subsec:NN}
+The carbon prices are relatively evenly distributed across the continent at high level 115-155~€/t. This indicates
 
@@
 
-Our study presents a comprehensive assessment of the roles of hydrogen (\Htwo{}) and carbon dioxide (\COtwo{}) networks in the future European energy system.
-By comparing models with and without these networks under net-zero and net-negative scenarios, we have elucidated the benefits and trade-offs associated with each approach.
+Our study assesses the roles of \Htwo{} and \COtwo{} networks in Europe's future energy system, revealing the trade-offs and benefits under various scenarios.
 The Baseline model, lacking both \Htwo{} and \COtwo{} networks, shows significant inefficiencies and higher system costs.
@@
 The model considers various energy carriers like electricity, heat, hydrogen, methane, methanol, liquid hydrocarbons and biomass, together with the corresponding conversions technologies.
-
-\begin{figure}
-    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/demand_bar.png}
-    \caption{Assumptions on exogenous demand, derived from~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023}.
-The figure shows the total annual energy demand for each energy source which determine the model's endogenous investments and operation.
-Endogenous processes can lead to higher total production volumes of some energy carriers, e.g., the demand for methanol requires more hydrogen and carbon as secondary (energy) inputs, which are not considered here.
-The model are defined per region and time stamp.}
-    % TODO: adjust labels to show that these are exogenous assumptions
-    \label{fig:total-demand-bar}
-\end{figure}
 
@@
 Our estimates for carbon sequestration potentials are restrictive, limiting the annual sequestration potential to 25~Mt per region and constraining it further such that this amount could be  25 years.
-Furthermore, we decide to limit the total amount of sequestered \COtwo{} to 200~Mt per year for our Net-Zero scenario and XXX~Mt per year for our Net-Negative scenario.
+Furthermore, we decide to limit the total amount of sequestered \COtwo{} to 200~Mt per year for our net-zero \carbon{} emission target and 660~Mt per year for our net-negative emission target.
 This is enough to offset hard-to-abate fossil process emissions (e.g., calcination of limestone in cement manufacturing) and some limited slack of fossil fuels to be used where they can be most valuable.
@@
 \end{figure}
+
+\subsection{Technologies with Carbon Capture}
+
+In the \modBase{}, the capture share is highest for biogas-to-gas facilities, followed by process emissions, biomass for industry, biomass CHPs, gas for industry, SMR and gas CHPs.
+For models with additional transport systems, we observe a correlation between CC share and capacity factors of the underlying technologies: Process emissions as well as gas and biomass for industry have high capacity factors (above 80\%), and CC shares close to 100\%, the only exception being gas for industry in the \modH{}.
+Biomass CHPs have a capacity factor of around 50\%, and CC shares of 50\% in the \modH{} and 97\% in the \modCO{}.
+SMR and gas CHPs have capacity factors around 15\%, and CC shares at nearly 0\% for the \baselinemodel.
+The capacity factor and CC share of SMR increases to around 20\% and 50-65\%, respectively, in the other models.
+Gas CHPs serve as peak load electricity production technology, and with very few operating hours, thus the high investment costs into CC applications are not efficient.
+Similarly, SMR serves as ``peak-load'' hydrogen production technology in regions with poor renewable resources for electrolysis.
 
@@
 
+\end{document}
+
+% LEFT OVER
+% =========
+%
+%
 % [Dynamics \modCO{} vs \modH{} (old text blocks)]
@@
 These factors add up to account for 14~bn€ of the difference of 20~bn€ in system costs between the two scenarios (785~bn€ for the \COgrid{} and 765~bn€ for the \modH{}, see Fig.~\ref{fig:objective_heatmap}).
-
-\end{document}

### 2023-12-20 report/submissions/journal/paper/word-count.tex
incorp reviews, restructure figures

@@
-Hydrogen and carbon dioxide transportation are considered crucial in climate-neutral energy systems, with hydrogen enabling energy imports to high-demand areas and carbon transport aiding in emissions export from high-emission zones.
-Yet, the synergy or competition between these systems is not fully understood.
-Our study employs optimization techniques to develop a cost-optimal European energy system, integrating transport, storage, and sequestration of both carbon dioxide and hydrogen, along with renewable energy sources.
-Results indicate that a hydrogen network is more cost-effective than a carbon network, facilitating low-cost hydrogen imports and local carbon capture.
-However, in a hybrid scenario, the carbon network effectively complements the hydrogen network, promoting carbon capture from biomass and reducing reliance on direct air capture.
-This research demonstrates the effectiveness of combining hydrogen and carbon transport networks with power grid expansion in achieving climate neutrality in Europe.
+Hydrogen and carbon dioxide transport can both play an important role in climate-neutral energy systems.
+Hydrogen networks help serve regions with high energy demand, while excess emissions are transported away in carbon dioxide networks.
+When it comes to the synthesis of carbonaceous fuels, it is less clear which input should be transported: hydrogen to carbon point sources or carbon to low-cost hydrogen.
+We explore the potential synergies and competition of both networks in a cost-optimal carbon-neutral European energy system.
+In direct comparison, a hydrogen network is more cost-effective than a carbon network, since it serves to transport hydrogen to demand and to point source of carbon for utilization.
+However, in a hybrid scenario where both networks are present, the carbon network effectively complements the hydrogen network, promoting carbon capture from biomass and reducing reliance on direct air capture.
+This study shows how the European energy system in a future net-zero and net-negative emissions scenarios may benefits from integrating both networks.

### 2023-12-20 report/submissions/journal/paper/paper.tex
finish restructering

@@
 However, in a hybrid scenario where both networks are present, the carbon network effectively complements the hydrogen network, promoting carbon capture from biomass and reducing reliance on direct air capture.
-This study shows how the European energy system in future net-zero and net-negative emissions scenarios can benefits from integrating both networks.
+This study shows how the European energy system in future net-zero and net-negative carbon dioxide emissions scenarios can benefits from integrating both networks.
 \end{abstract}
@@
 Endogenous processes can lead to higher total production volumes of some energy carriers, e.g., the demand for methanol requires more hydrogen and carbon as secondary (energy) inputs, which are not considered here.
-The model are defined per region and time stamp.}
+In the model, demands are defined per region and time stamp.}
     % TODO: adjust labels to show that these are exogenous assumptions
@@
 \end{itemize}
-%
-% In anticipation of the future \carbon{} emission targets, the models are subjected to the following two scenarios:
-% \begin{itemize}
-%     \item[] \textit{Net\=/Zero}: aligning with the EU's 2050 emission targets. Limiting carbon sequestration to 200~Mt/a. This scenario is also denoted as NZ.
-%     \item[] \textit{Net\=/Negative}: 10\% net-negative emissions relative to 1990 levels, equaling 460~Mt \COtwo{} annually. Limiting carbon sequestration to 660~Mt/a. Also denoted as NN.
-% \end{itemize}
 
@@
 We give a detailed description of the different technological and geographical impacts of the transportation systems.
-Subsequently, our analysis demonstrates the extent to which different deployment strategies remain robust under a tightened net-negative \carbon{} emissions target.
+Subsequently, our analysis demonstrates the extent to which different deployment strategies remain robust under a tightened net \carbon{} removal target.
 
@@
 Models with \carbon{} and \hydrogen{} transport options are more cost-effective, reducing costs by 3.1\% in \modCO{}, 4.8\% in \modH{}, and 5.2\% in \modHybrid{}.
-% in system costs: 779bn€ (baseline), 754bn€ (carbon), 742bn€ (hydrogen), 738bn€ (hybrid)
 
@@
 
-% To understand the dynamics of the \modCO{}, we observe that the system locates CU in the regions where it can minimize costs for its material inputs, carbon and hydrogen.
-As the \COtwo{} network largely smoothes out carbon prices across regions, CU is located in the hydrogen price valleys on the periphery of the continent.
-Similarly, the model locates CS in regions where it can minimize the costs of the carbon to be sequestered.
-Thus, CS is located in the carbon price valleys among the subset of regions with access to offshore sequestration sites.
-Due to the \COtwo{} network these low-price regions are not strongly pronounced, but still prevalent and impact CS locations.
-DAC is only used for CU and thus co-located in regions with a mix of low electricity costs (for running DAC) and low hydrogen costs.
-
-% To explain the dynamics of the \modH{}, we observe that the model locates CU in the carbon price valleys across Western, Central and Eastern Europe (see top right), as the \Htwo{} network smoothes out hydrogen prices across the entire continent (see bottom right).
-CS is again located in the carbon price valleys among those regions with access to offshore sequestration sites.
-Due to the lack of a \COtwo{} network the full CC potentials from point sources are exhausted in all offshore sequestration regions, and the carbon is then sequestered.
-The additionally needed carbon to reach 200~Mt of CS is captured via DAC in the regions with access to offshore sequestration sites with the lowest costs for DAC, with the largest share deployed in Portugal (see top right).
-DAC is also used for CU in this model, and for both purposes DAC is located in the regions with the lowest costs for electricity.
-
-% The increase in captured carbon from point sources in the models with \carbon{} and/or \hydrogen{} transport is more related to an upgrading of existing industrial or power plant facilities to capture carbon rather than a shift in technologies.
-This is evident when analyzing the capture share of the different CC technologies (see Fig.~\ref{fig:captureshare_line}).
-
-% In summary, the implementation of \COtwo{} and \Htwo{} networks is instrumental in unlocking CC potentials and reducing system costs.
-However, despite the \modCO{} demonstrating higher amounts and shares of CC from point sources and therefore less DAC deployment, the \modH{} achieves even  lower total system costs (see Fig.~\ref{fig:cost_bar}).
-
 \subsection*{Hybrid configuration provides further flexibilities}\label{subsec:Hybrid}
@@
 
-In this combination, the networks enable full carbon capture upgrades for all technologies that run on a high average capacity factor (above 70\%, see Fig.~\ref{fig:captureshare_line}).
+In this combination, the networks enable all technologies that run on a high average, above 70\%, capacity factor to capture carbon (see Fig.~\ref{fig:captureshare_line}).
 Except for SMR and gas-fired Combined Heat and Power Plants (CHP) which operate a few weeks annually, all emissions from carbon-emitting technologies are captured.
@@
 
-The carbon prices are relatively evenly distributed across the continent at high level 115-155~€/t. This indicates
-
 \begin{figure}[htb!]
@@
     \caption[short]{Net change in system cost when tightening the \COtwo{} emission target from net-zero to net-negative 10\% of 1990s emissions.
-Each cost bar is split into contribution of the same technology groups, except for \carbon{} sequestration cost which are now diplayed separately.
+Each cost bar is split into contribution of the same technology groups, except for \carbon{} sequestration cost, which are now displayed separately.
 For all models, Direct Air Capture (DAC) contributes most to the additional \carbon{} removal, requiring further solar, wind and heat pump capacities for electricity and heat input.}
@@
 
-In the following, we analyze how investment trends and network strategies alter when transitioning from a net zero to a net emissions reduction target.
-Therefore, we subject the models to the Net-Negative scenario with a net carbon removal of 460~Mt/a and an increased total sequestration limit of 660~Mt/a.
-In comparison to the Net-Zero scenario, all models with transport options incur a roughly similar net annual cost increase - 77~bn€ for the \modCO{}, 81~bn€ for the \modH{}, and 78~bn€ for the \modHybrid{} (see Fig.~\ref{fig:net-negative_cost_bar}), resulting in a higher cost-benefit of 6~bn€ for the \hybridmodel{} in comparison to the \hydrogenmodel{}.
+However, the exploitation of carbon point sources is close to its feasible limit.
+When applying a stricter \carbon{} emissions target and force the model to remove 460~Mt \carbon{} over one year (equivalent to 10\% of \carbon{} emissions in 1990) while sequestering maximally 660~Mt, investment in DAC and associated solar and wind power and heat supply cover a large share of additional system cost in all scenarios (see Fig.~\ref{fig:net-negative_cost_bar}).
+In the \hybridscenario{}, this share is at 80\% while new carbon capture at point sources only make up 6.5\% of new system costs.
+The additional sequestration is mainly supplied by large DAC facilities near the shore.
+To a small extend it is supplied by additional carbon captured and transported from bio-energetic sources in the inland that were used to supply CU before (compare Fig.~\ref{fig:balance_maps_full_nn}).
+At the same time, some of the CU in Central Europe moves to Spain, where it uses carbon from new DAC facilities without the installation of a \COtwo{} network.
+This slight shift in CU deployment leads to less investments in the hydrogen transport route from the Iberian Peninsula to Central Europe (2~bn€/a) and to an additional investment in the \carbongrid{} from Central and Eastern Europe to the shores (1.2~bn€/a, compare Fig.~\ref{fig:cost_bar_transmission_nn}).
 
-The models equally invest in the expansion of sequestration capacities up to the global limit (5~bn€).
-The investments in additional DAC (36-41~bn€), wind (6-7~bn€) and heat pumps (6-7~bn€) are relatively evenly for all models, with DAC accounting for roughly half of the additional system costs.
-Fundamental differences in investment strategies occur in the deployment of solar, carbon capture and bio-energy technologies: While the \carbonmodel{} expands CC at point sources the most, the \hydrogenmodel{} and the \hybridmodel{} invest relatively more in new solar plants.
-The \carbonmodel{} invests three time more in new bio-energy sources (6~bn€) than the other models.
-All models only reveal small net downscaling of technologies.
-However, as we discuss below, the geographic impact can vary depending on the model, sometimes leading to disruptive changes in optimal network strategies and stranded assets.
-
-In contrast to its layout in the Net-Zero scenario, the \carbonmodel{} fundamentally reorganizes the carbon transport pathways (see Fig.~\ref{fig:balance_maps_co2_nn}).
-In Spain, carbon utilization (CU) is now predominantly supplied by carbon extracted from DAC.
-Central Europe sees a major transformation, with carbon mainly being transported to sequestration sites in the north and south-east.
-This reorganization leads to stranded investment in the carbon network infrastructure, amounting to approximately 3~bn€.
-%
-The \modH{} on the other hand, primarily leaves the network topology unchanged and only expands DAC facilities at regions with access to sequestration sites as well as additional solar, wind and heat pumps in the vicinity to supply power and heat (see Fig.~\ref{fig:balance_maps_h2_nn}).
-%
-Finally, the \modHybrid{} expands the sequestration with carbon captured from bio-energetic inputs and transported from Central and Eastern Europe to the near shores (see Fig.~\ref{fig:balance_maps_full_nn}), leading to an additional investment in the \carbongrid{} (1.2~bn€, see Fig.~\ref{fig:cost_bar_transmission_nn}).
-At the same time, carbonaceous fuel production in Central Europe partially moves to Spain, where it uses carbon from new DAC facilities without the installation of a \COtwo{} network.
-As a trade-off, the model transports less hydrogen from the Iberian Peninsula to Central Europe, resulting in a dismantling of 2~bn€ in hydrogen network costs compared to the Net-Zero scenario.
-
-These findings highlight the continued trends of the two systemic approaches which are combined in the \hybridmodel.
-With an increased demand for sequestration, the \carbongrid{} relatively gains more importance, leading to higher cost-benefits by enabling to combine increased need DAC with a more dominant collection and transportation scheme of carbon from distributed point sources.
+Overall, the net cost increase by 77~bn€ in the \carbonscenario{}, by 81~bn€ in the \hydrogenscenario{}, and by 78~bn€ in the \hybridscenario{}, resulting in a cost-benefit of 6~bn€ for the \hybridscenario{} in comparison to the other scenarios.
 
@@
 
-Our study assesses the roles of \Htwo{} and \COtwo{} networks in Europe's future energy system, revealing the trade-offs and benefits under various scenarios.
-The Baseline model, lacking both \Htwo{} and \COtwo{} networks, shows significant inefficiencies and higher system costs.
-The \COgrid{} model, with its \COtwo{} network, effectively integrates low-cost carbon from distributed point sources, primarily for carbon utilization and sequestration.
-Conversely, the \Hgrid{} model, with its \Htwo{} network, facilitates low-cost hydrogen delivery to high-demand areas, thus enabling efficient carbon utilization alongside point-source carbon capture.
-The Hybrid model, combining both \COtwo{} and \Htwo{} networks, leverages the strengths of each, resulting in the lowest system costs and a robust energy system.
-This model effectively utilizes \COtwo{} network for short-range carbon transport for sequestration and complements this with the \Htwo{} network for long-range hydrogen transport to support distributed carbon utilization.
-Under the net-negative scenario, the importance of Direct Air Capture (DAC) increases across all models, highlighting the growing significance of DAC in achieving deeper decarbonization goals.
+Our study assesses the roles of hydrogen (\Htwo{}) and carbon dioxide (\COtwo{}) networks in Europe's future energy system under net-neutrality and net-negativity targets for \carbon{} emissions.
+Both networks have strong impact of the optimal deployment of carbon capture, carbon utilization and sequestration.
+A \carbon{} network effectively reduces costs by transporting low-cost carbon from distributed point sources to sequestration sites and centralized carbon utilization at sites with low-cost hydrogen production.
+A hydrogen network provides greater cost savings as low-cost hydrogen can be transported to regions with high demand, allowing carbon utilization at point-sources.
+A combination of both networks emphasize the strengths of each infrastructure.
+In the hybrid configuration, the \hydrogen{} network supplies low-cost hydrogen to spatially-fixed hydrogen demand and carbon utilization at point-sources across regions.
+At the same time the \carbon{} network transports low-cost carbon captured from point-sources close to the coast to sequestration sites, reducing the overall reliance on high-cost Direct Air Capture.
+This results not only in the lowest system cost, but also in a more robust energy system.
 
-Despite the comprehensive nature of the model, there are some limitations that affect the validity of the results, as well as important discussion points that need to be put into context.
-In reality, energy demand and renewable supply can only be estimated for the next few hours/days.
-The model's perfect foresight may lead to non-reproducible technology dispatch which neglects uncertainties from long-term predictions and necessary reserve margins.
-The modelling results are heavily driven by the demand and emissions from industrial clusters.
-Allowing the model to relocate these, incorporate flexibility measures or alter underlying processes, may lead to less dependency on both \COtwo{} and \Htwo{} networks.
-The transport of fuel (FT, gas, oil) between the regions is not limited which may overestimate the flexibility provided by these commodities.
-However, the distributed CU systems in the \hydrogenmodel{} and \hybridmodel{} align well with today's distributed locations for oil-refineries, providing kerosene for aviation and naphtha for industry.
-The assumption to run Fischer-Tropsch facilities on baseload with at least 90\% of their nominal capacity, is likely shaping the deployment of CU, underestimating potential important flexibilities.
-The assumed cost projections on technologies are subject to uncertainties.
-As shown in \ref{sec:subsidy}, the effect on a 50\% cost reduction on \carbon{} pipelines alter total system costs to small extent (below 1\%).
-However, a different set of cost on DAC and electrolysis may-well impact in the model results.
-In combination with a net-zero or net-negative emission target, a relaxation of the global sequestration limit drives system costs down by introducing fossil fuels, as shown in~\cite{hofmannDesigningCO2Network2023}.
-Allowing the model to sequester onshore, would likely reduce the dominant co-location of CU with the point sources due to easier access to sequestration sites.
-Finally, the large deployment of DAC is subject to conservative biomass potentials, as shown in~\cite{lauerCrucialRoleBioenergy2023}.
+We discuss limitations of the model in section~\ref{sec:limitations} and sensitivities to \carbon{} pipeline costs in section~\ref{sec:subsidy}.
+Further research should explore the impacts of synthetic fuel imports and biomass potentials.
 
 In summary, our study underscores the need for strategic planning and integration of both \COtwo{} and \Htwo{} networks in the European energy landscape.
-The flexibility and cost-effectiveness provided by these networks are pivotal for realizing the EU's climate neutrality ambitions.
-By harnessing the synergies between \Htwo{} and \COtwo{} transportation and management, Europe can effectively navigate the complexities of a transition to a sustainable, low-carbon energy future.
+The flexibility and cost-effectiveness provided by these networks are pivotal for realizing the EU's climate targets.
+By harnessing the synergies between \Htwo{} and \COtwo{} transportation and management, Europe can effectively reduce system costs while ensuring a robust and sustainable energy future.
 
@@
 
-The study is conducted on the basis of the open-source, capacity expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
+The study is conducted based on the open-source, capacity expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
 The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, agriculture, and transport sectors, including international aviation and shipping.
@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=.5\linewidth]{baseline/figures/90_nodes/sequestration_map.png}
+    \caption{Maximal sequestration potential per offshore region used as input for all models.
+Note that in most model runs the regional sequestration potentials are not exploited due the the global limit on sequestration.}
+    \label{fig:sequestration_map}
+\end{figure}
+
 To store carbon, we differentiate between short-term storage in steel tanks without permanent containment and long-term, irreversible sequestration in underground sequestration sites such as porous rock formations or depleted gas reservoirs.
 Costs for both options are included in the model.
-For carbon sequestration, we only consider offshore sites as potential sinks.
+For carbon sequestration, we only consider offshore sites as potential sinks (see Fig.~\ref{fig:sequestration_map}).
 We make this choice because offshore sequestration sites typically have larger capacity compared to onshore sequestration sites in saline aquifers and due to concerns over public safety for infrastructure near populated areas.
@@
 
-\subsection{Model Inputs}
-
-The following figures extend on the model assumptions and inputs.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=.5\linewidth]{baseline/figures/90_nodes/sequestration_map.png}
-    \caption{Maximal sequestration potential per offshore region used as input for all models.
-Note that in most model runs the regional sequestration potentials are not exploited due the the global limit on sequestration.}
-\end{figure}
+\subsection{Limitations}
+\label{sec:limitations}
+Despite the comprehensive nature of the model, some limitations affect the validity of the results, as well as important discussion points that need to be put into context.
+In reality, energy demand and renewable supply can only be estimated for the next few hours/days.
+The model's perfect foresight may lead to non-reproducible technology dispatch which neglects uncertainties from long-term predictions and necessary reserve margins.
+Demand and emissions from industrial clusters heavily drive the modelling results.
+Allowing the model to relocate these, incorporate flexibility measures or alter underlying processes, may lead to less dependency on \COtwo{} and \Htwo{} networks.
+The transport of fuel (FT, gas, oil) between the regions is not limited which may overestimate their flexibility.
+However, the distributed CU systems, as enabled through the hydrogen network, align well with today's distributed locations for oil-refineries, providing kerosene for aviation and naphtha for industry.
+The assumption to run Fischer-Tropsch facilities on baseload with at least 90\% of their nominal capacity, is likely shaping the deployment of CU, underestimating potential important flexibilities.
+The assumed cost projections on technologies are subject to uncertainties.
+As shown in \ref{sec:subsidy}, the effect on a 50\% cost reduction on \carbon{} pipelines alter total system costs to small extent (below 1\%).
+However, a different set of cost on DAC and electrolysis may-well impact in the model results.
+In combination with a net \carbon{} neutrality or reduction target, a relaxation of the global sequestration limit drives system costs down by introducing fossil fuels, as shown in~\cite{hofmannDesigningCO2Network2023}.
+Allowing the model to sequester onshore, would likely reduce the dominant co-location of CU with the point sources due to easier access to sequestration sites.
+Finally, DAC deployment could be in parts replaced if larger sustainable biomass sources are assumed, as shown in ~\cite{lauerCrucialRoleBioenergy2023}.
 
@@
     \end{subfigure}
-    \caption{Annual transmission system costs for different models and scenarios in the Net-Zero scenario (a) and Net-Negative scenario (b).}
+    \caption{Annual transmission system costs for different models and scenarios for net \carbon{} neutrality (a) and net \carbon{} removal scenario (b).}
 \end{figure}
@@
     \includegraphics[width=\linewidth]{difference/net-negative-0.1-h2-only-full/figures/90_nodes/cost_map.png}
-    \caption{Difference in regional costs between the \Hgrid{} and \modHybrid{}s in the Net-Negative scenario.
+    \caption{Difference in regional costs between the \Hgrid{} and \modHybrid{}s in the net \carbon{} removal scenario.
 The left figure shows higher spendings per technology and region and transport system for the \modH{}, the right shows higher spendings in the \modHybrid{}.}
@@
 \clearpage
-\subsection{Operation in Net-Negative Scenarios}
-The following figures display the optimal operation of the hydrogen and carbon sector for all models in the Net-Negative scenario.
+\subsection{Operation in Net \carbon{} Removal Scenarios}
+The following figures display the optimal operation of the hydrogen and carbon sector for all models in the net \carbon{} removal scenario.
 
@@
     \end{subfigure}
-    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hydrogenmodel{} in the net negative scenario.
+    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hydrogenmodel{} in the net \carbon{} removal scenario.
 For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
@@
 % =========
+
+% To understand the dynamics of the \modCO{}, we observe that the system locates CU in the regions where it can minimize costs for its material inputs, carbon and hydrogen.
+As the \COtwo{} network largely smoothes out carbon prices across regions, CU is located in the hydrogen price valleys on the periphery of the continent.
+Similarly, the model locates CS in regions where it can minimize the costs of the carbon to be sequestered.
+Thus, CS is located in the carbon price valleys among the subset of regions with access to offshore sequestration sites.
+Due to the \COtwo{} network these low-price regions are not strongly pronounced, but still prevalent and impact CS locations.
+DAC is only used for CU and thus co-located in regions with a mix of low electricity costs (for running DAC) and low hydrogen costs.
+
+% To explain the dynamics of the \modH{}, we observe that the model locates CU in the carbon price valleys across Western, Central and Eastern Europe (see top right), as the \Htwo{} network smoothes out hydrogen prices across the entire continent (see bottom right).
+CS is again located in the carbon price valleys among those regions with access to offshore sequestration sites.
+Due to the lack of a \COtwo{} network the full CC potentials from point sources are exhausted in all offshore sequestration regions, and the carbon is then sequestered.
+The additionally needed carbon to reach 200~Mt of CS is captured via DAC in the regions with access to offshore sequestration sites with the lowest costs for DAC, with the largest share deployed in Portugal (see top right).
+DAC is also used for CU in this model, and for both purposes DAC is located in the regions with the lowest costs for electricity.
+
+% The increase in captured carbon from point sources in the models with \carbon{} and/or \hydrogen{} transport is more related to an upgrading of existing industrial or power plant facilities to capture carbon rather than a shift in technologies.
+This is evident when analyzing the capture share of the different CC technologies (see Fig.~\ref{fig:captureshare_line}).
+
+% In summary, the implementation of \COtwo{} and \Htwo{} networks is instrumental in unlocking CC potentials and reducing system costs.
+However, despite the \modCO{} demonstrating higher amounts and shares of CC from point sources and therefore less DAC deployment, the \modH{} achieves even  lower total system costs (see Fig.~\ref{fig:cost_bar}).
 %
+%
+% Comparison carbon and hydrogen network
 %
@@
 These factors add up to account for 14~bn€ of the difference of 20~bn€ in system costs between the two scenarios (785~bn€ for the \COgrid{} and 765~bn€ for the \modH{}, see Fig.~\ref{fig:objective_heatmap}).
+
+% Net-negative
+
+% Fundamental differences in investment strategies occur in the deployment of solar, carbon capture and bio-energy technologies: While the \carbonmodel{} expands CC at point sources the most, the \hydrogenmodel{} and the \hybridmodel{} invest relatively more in new solar plants.
+The \carbonmodel{} invests three time more in new bio-energy sources (6~bn€) than the other models.
+% All models only reveal small net downscaling of technologies.
+However, as we discuss below, the geographic impact can vary depending on the model, sometimes leading to disruptive changes in optimal network strategies and stranded assets.
+
+% In contrast to its layout in the Net-Zero scenario, the \carbonmodel{} fundamentally reorganizes the carbon transport pathways (see Fig.~\ref{fig:balance_maps_co2_nn}).
+% In Spain, carbon utilization (CU) is now predominantly supplied by carbon extracted from DAC.
+Central Europe sees a major transformation, with carbon mainly being transported to sequestration sites in the north and south-east.
+This reorganization leads to stranded investment in the carbon network infrastructure, amounting to approximately 3~bn€.
+
+% The \modH{} on the other hand, primarily leaves the network topology unchanged and only expands DAC facilities at regions with access to sequestration sites as well as additional solar, wind and heat pumps in the vicinity to supply power and heat (see Fig.~\ref{fig:balance_maps_h2_nn}).
+
+% Finally, the \modHybrid{} expands the sequestration with carbon captured from bio-energetic inputs and transported from Central and Eastern Europe to the near shores (see Fig.~\ref{fig:balance_maps_full_nn}), leading to an additional investment in the \carbongrid{} (1.2~bn€, see Fig.~\ref{fig:cost_bar_transmission_nn}).
+At the same time, carbonaceous fuel production in Central Europe partially moves to Spain, where it uses carbon from new DAC facilities without the installation of a \COtwo{} network.
+As a trade-off, the model transports less hydrogen from the Iberian Peninsula to Central Europe, resulting in a dismantling of 2~bn€ in hydrogen network costs compared to the Net-Zero scenario.
+
+% These findings highlight the continued trends of the two systemic approaches which are combined in the \hybridmodel.
+With an increased demand for sequestration, the \carbongrid{} relatively gains more importance, leading to higher cost-benefits by enabling to combine increased need DAC with a more dominant collection and transportation scheme of carbon from distributed point sources.

### 2023-12-20 report/submissions/journal/paper/paper.tex
convert "models" into "scenarios"

@@
 % make abbreviation for co2
-\newcommand{\COtwo}{CO$_2$}
-\newcommand{\Htwo}{H$_2$}
-\newcommand{\COgrid}{CO$_2$\=/Grid}
-\newcommand{\Hgrid}{H$_2$\=/Grid}
-\newcommand{\modBase}{Baseline model}
-\newcommand{\modCO}{CO$_2$\=/Grid model}
-\newcommand{\modH}{H$_2$\=/Grid model}
-\newcommand{\modHybrid}{Hybrid model}
-
-% alternative
 \newcommand{\carbon}{CO$_2$}
@@
 \newcommand{\hydrogengrid}{H$_2$\=/Grid}
-\newcommand{\baselinemodel}{Baseline model}
-\newcommand{\carbonmodel}{CO$_2$\=/Grid model}
-\newcommand{\hydrogenmodel}{H$_2$\=/Grid model}
-\newcommand{\hybridmodel}{Hybrid model}
 \newcommand{\baselinescenario}{Baseline scenario}
@@
 
-\title{Evaluating \Htwo{} and \COtwo{} Network Strategies for the European Energy System}
-% \title{Competition and Synergies of \Htwo{} and \COtwo{} Networks in Europe}
+\title{Evaluating Hydrogen and Carbon Network Strategies for the European Energy System}
+% \title{Competition and Synergies of \hydrogen{} and \carbon{} Networks in Europe}
 
@@
 Hydrogen networks help serve regions with high energy demand, while excess emissions are transported away in carbon dioxide networks.
-When it comes to the synthesis of carbonaceous fuels, it is less clear which input should be transported: hydrogen to carbon point sources or carbon to low-cost hydrogen.
+Regarding the synthesis of carbonaceous fuels, it is less clear which input should be transported: hydrogen to carbon point sources or carbon to low-cost hydrogen.
 We explore the potential synergies and competition of both networks in a cost-optimal carbon-neutral European energy system.
@@
 However, in a hybrid scenario where both networks are present, the carbon network effectively complements the hydrogen network, promoting carbon capture from biomass and reducing reliance on direct air capture.
-This study shows how the European energy system in future net-zero and net-negative carbon dioxide emissions scenarios can benefits from integrating both networks.
+This study shows how the European energy system in future net-zero and net-negative carbon dioxide emissions scenarios benefits from integrating both networks.
 \end{abstract}
@@
 The transition to a climate-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors.
-While management of both carbon dioxide (\COtwo{}) and hydrogen (\Htwo{}) is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructure effectively interacts with carbon management technologies, including carbon capture, transport, storage, utilization, and sequestration.
-Hydrogen is consider in several sectors in industry not amenable to electrification, and offers an efficient way to transport and store energy over long distances and timeframes, and can be produced from electrolysis or steam methane reforming (SMR), for example.
+While management of both carbon dioxide (\carbon{}) and hydrogen (\hydrogen{}) is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructure effectively interacts with carbon management technologies, including carbon capture, transport, storage, utilization, and sequestration.
+Hydrogen is considered in several sectors in industry not amenable to electrification, and offers an efficient way to transport and store energy over long distances and periods, and can be produced from electrolysis or steam methane reforming (SMR), for example.
 Captured carbon can be stored in geological formations to permanently extract it from the atmosphere, a process known as carbon sequestration (CS).
-Additionally, carbon can be combined with hydrogen to produce carbonaceous fuels to supply for example synthetic kerosene for aviation, synthetic methanol for shipping or synthetic methane for industrial feedstocks.
+Additionally, carbon can be combined with hydrogen to produce carbonaceous fuels to supply for example synthetic kerosene for aviation, synthetic methanol for shipping or synthetic methane for industrial feedstock.
 From hereon, we jointly refer to the production of these carbonaceous fuels as Carbon Utilization (CU).
 Carbon can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, or synthetic carbonaceous fuels through point-source carbon capture (CC) techniques, or harvested from the atmosphere using direct air capture (DAC).
-Both \COtwo{} and \Htwo{} networks are likely to play a crucial role in the cost-effective integration of these technologies to enable net-zero economies.
+Both \carbon{} and \hydrogen{} networks are likely to play a crucial role in the cost-effective integration of these technologies to enable net-zero economies.
 
@@
 Some natural gas pipelines have already been repurposed to transport hydrogen~\cite{RohrFreiFuer}.
-Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synthetic fuels on the one hand and offer \COtwo{} offtake on the other hand.
-The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a transport and sequestration capacity of 1.5~Mt \COtwo{} per year to be operative in 2024, expanding to a targeted scale of 5~Mt per year of sequestration by 2030.
+Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synthetic fuels on the one hand and offer to purchase \carbon{} on the other hand.
+The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a transport and sequestration capacity of 1.5~Mt \carbon{} per year to be operative in 2024, expanding to a targeted scale of 5~Mt per year of sequestration by 2030.
 The market potential for carbon capture and sequestration in Europe is underscored by the Capture Map's estimated potential~\cite{ToolsGreenTransition} of 1.7 Gt carbon annually from point sources, accounting for about half of the continent's emissions, as well as estimated sequestration capacity potentials of up to 3~Gt/a~\cite{europeancommissionEuropeanCO2Storage}.
@@
 Both can bridge the misalignment of sources and sinks for carbon and hydrogen.
-A hydrogen network can supply regions with geographically fixed hydrogen demand, such as for steelmaking, with hydrogen from regions with the best renewable resources as well as enable CU at the site of point-source CC.
+A hydrogen network can supply regions with geographically fixed hydrogen demand, such as for steel production, with hydrogen from regions with the best renewable resources as well as enable CU at the site of point-source CC.
 On the other hand, the carbon transport system allows for transporting captured carbon to regions with sequestration potentials or high-quality renewable resources, the latter enabling cost-effective CU.
@@
 The underlying, highly-resolved model encompasses most relevant sectors, however it neglects the option to transport carbon across regions.
-Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \COtwo{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
-This limited sectoral scope cannot capture important dynamics of carbon management, since it neglects the sectors like industry that will need to handle most \COtwo{} in the future.
-A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \COtwo{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
-However, this limited spatial scope fails to consider other sequestration sites and co-benefits from connecting the \COtwo{} network to neighboring countries.
-Other high-resolution energy system models do not feature detailed CU technologies and \COtwo{} transport.
+Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \carbon{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
+This limited sectoral scope cannot capture important dynamics of carbon management, since it neglects the sectors like industry that will need to handle most \carbon{} in the future.
+A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
+However, this limited spatial scope fails to consider other sequestration sites and co-benefits from connecting the \carbon{} network to neighboring countries.
+Other high-resolution energy system models do not feature detailed CU technologies and \carbon{} transport.
 Euro-Calliope, for example, does neither feature carbon nor hydrogen transport, and thus forces generic synthetic fuel production to rely on captive electrolysis and DAC in the same region~\cite{pickeringDiversityOptionsEliminate2022}.
 
-No study has yet considered the co-optimization and comprehensive assessment of both \COtwo{} and \Htwo{} networks in a fully sector-coupled energy system.
+No study has yet considered the co-optimization and comprehensive assessment of both \carbon{} and \hydrogen{} networks in a fully sector-coupled energy system.
 However, such an assessment is strongly needed to identify realms of competition and synergy between hydrogen and carbon management technologies.
@@
 
-To investigate the competing transport dynamics between \COtwo{} and \Htwo{} networks, we contrast the following four model scenarios:
+To investigate the competing transport dynamics between \carbon{} and \hydrogen{} networks, we contrast the following four model scenarios:
 %
 \begin{itemize}
-    \item[] \textit{Baseline:} Neither \COtwo{} nor \Htwo{} networks are constructed.
-    \item[] \textit{\COgrid{}:} Only the \COtwo{} network is developed.
-    \item[] \textit{\Hgrid{}:} Only the \Htwo{} network is developed.
-    \item[] \textit{Hybrid:} Both \COtwo{} and \Htwo{} networks are developed.
+    \item[] \textit{Baseline:} Neither \carbon{} nor \hydrogen{} networks are constructed.
+    \item[] \textit{\carbongrid{}:} Only the \carbon{} network is developed.
+    \item[] \textit{\hydrogengrid{}:} Only the \hydrogen{} network is developed.
+    \item[] \textit{Hybrid:} Both \carbon{} and \hydrogen{} networks are developed.
 \end{itemize}
@@
 All scenarios share key energy system characteristics but differ in system costs and technology use (Fig.~\ref{fig:cost_bar}).
-The \modBase{} has the highest annual cost at 779~bn€/a.
-Models with \carbon{} and \hydrogen{} transport options are more cost-effective, reducing costs by 3.1\% in \modCO{}, 4.8\% in \modH{}, and 5.2\% in \modHybrid{}.
+Costs in the \baselinescenario{} are the highest at 779~bn€/a.
+Models with \carbon{} and \hydrogen{} transport options are more cost-effective, reducing costs by 3.1\% in the \carbongrid{}, 4.8\% in \hydrogengrid{}, and 5.2\% in \hybridscenario{}.
 
@@
     \caption[short]{Total annual system cost subdivided into groups of technologies for the different models of the European energy system with a net-zero emission target.
-While the \modBase{} has neither a \COtwo{} nor an \Htwo{} network, the \modHybrid{} is allowed to expand both.
-``Gas Infrastructure'' combines gas facilities for power and heat production, ``\COtwo{} Infrastructure'' and ``\Htwo{} Infrastructure'' combine transport and storage for each carrier.
+While in the \baselinescenario{}, the model has neither a \carbon{} nor an \hydrogen{} network, it can expand both in the \hydrogenscenario{}.
+``Gas Infrastructure'' combines gas facilities for power and heat production, ``\carbon{} Infrastructure'' and ``\hydrogen{} Infrastructure'' combine transport and storage for each carrier.
 ``Carbon Capture at Point Sources'' combines all technologies with integrated carbon capture, including the cost of the main facility (e.g., CHP unit) and the carbon capture application.}
@@
 The increased usage of upgraded biogas leads to a lower synthetic methane production.
-% At the same time, the transport of \carbon{} and \hydrogen{} reduce the reliance on methanation and therefore the overall demand for CC volumes in the model.
-Another knock-on effect of less reliance on DAC is its reduced endogenous demand for dedicated renewables as well as heat pumps and biomass CHPs to supply power and heat.
+% Another knock-on effect of less reliance on DAC is its reduced endogenous demand for dedicated renewables as well as heat pumps and biomass CHPs to supply power and heat.
+
 \begin{figure}[ht!]
@@
 Positive values indicate carbon capture, negative values indicate carbon consumption.
-By integrating \Htwo{} and \COtwo{} networks, the predominant method for carbon removal shifts from Direct Air Capture (DAC) to a bioenergetic process that incorporates carbon capture.
+By integrating \hydrogen{} and \carbon{} networks, the predominant method for carbon removal shifts from Direct Air Capture (DAC) to bio-energetic processes with capture.
 At the same time the reliance on methanation decreases.}
@@
 In the \hydrogenscenario{}, the main hydrogen transportation routes go from Spain (50~GW) and the United Kingdom (18~GW) to Central Europe to supply local carbonaceous fuel production with captured carbon from point sources (see Fig~\ref{fig:balance_map}, bottom right).
-Also, the \Htwo{} network supplies spatially-fixed \Htwo{} demand for industry in Central Europe (2.5~GW in total) at much lower costs than local electrolysis or SMR.
+Also, the \hydrogen{} network supplies spatially-fixed \hydrogen{} demand for industry in Central Europe (2.5~GW in total) at much lower costs than local electrolysis or SMR.
 Carbon sequestration primarily locates in Portugal where large DAC facilities operate (14~kt/h, see Fig~\ref{fig:balance_map}, top right).
@@
 While the carbon grid makes it possible to transport captured carbon for sequestration at low cost, the dependence on DAC near the coast causes higher costs in the hydrogen scenario{}.
-On the other hand, the \hydrogengrid{} is able to supply the spatially-fixed hydrogen demands in Central Europe with low-cost hydrogen imports, while the \modCO{} is forced to deploy electrolysis in regions with poor renewable resources as well as additional gas infrastructure in Central Europe.
+On the other hand, the \hydrogengrid{} is able to supply the spatially-fixed hydrogen demands in Central Europe with low-cost hydrogen imports, while the model in the \carbonscenario{} is forced to deploy electrolysis in regions with poor renewable resources as well as additional gas infrastructure in Central Europe.
 The latter aspect dominates the first, leading to lower system cost in the \hydrogenscenario{}.
@@
     \end{subfigure}
-    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hybridmodel{} in the net zero scenario.
+    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors in the \hybridscenario{} assuming net \carbon{} neutrality.
 For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
@@
 
-The \hybridscenario{}, the model combines the advantages of hydrogen and carbon networks, resulting in the highest system cost reduction of \label{}5.2\% (\label{}41~bn€/a) compared to the \modBase{}.
-As in the \hydrogenscenario{}, the model transports low-cost hydrogen to the point sources in Central Europe to supply CU, but also transports low-cost carbon from point sources close the shore to sequestration sites (see Fig.~\ref{fig:balance_map_full}).
+The \hybridscenario{}, the model combines the advantages of hydrogen and carbon networks, resulting in the highest system cost reduction of 5.2\% (41~bn€/a) compared to the \baselinescenario{}.
+Again, the model transports low-cost hydrogen to the point sources in Central Europe to supply CU, but also transports low-cost carbon from point sources close the shore to sequestration sites (see Fig.~\ref{fig:balance_map_full}).
 This reduces reliance on high-cost DAC and provides low-cost hydrogen across the regions.
@@
 For the \carbon{} network, the model employs short and direct pipeline routes to transport carbon from inland point sources to nearby sequestration sites, as for example in Northern Germany, Italy and Greece (see Fig.~\ref{fig:balance_map_full}, left).
-Additionally, the \COtwo{} network balances out small local variations in carbon demand and supply to produce CU, as for example in the middle of France.
+Additionally, the \carbon{} network balances out small local variations in carbon demand and supply to produce CU, as for example in the middle of France.
 The hydrogen routes are similar to those in the \hydrogenscenario{} with transports from Western regions (Iberian Peninsula and British Isles) to distributed CU production sites across the continent (see Fig.~\ref{fig:balance_map_full}, right).
-While the topology of the \Htwo{} network (8~bn€ investments) roughly corresponds to that of the \modH{} primarily driven by CU, the \COtwo{} network plays more of a complementary role primarily driven by CS (3~bn€ investments, see Fig.~\ref{fig:cost_bar_transmission}).
+While the topology of the \hydrogen{} network (8~bn€ investments) roughly corresponds to that of the \hydrogenscenario{} primarily driven by CU, the \carbon{} network plays more of a complementary role primarily driven by CS (3~bn€ investments, see Fig.~\ref{fig:cost_bar_transmission}).
 
@@
 The size of the dots corresponds to the average capacity factor of the respective technology.
-While peak load technologies such as gas-fired combined heat and power (CHP) plants, which are only in operation on a few days a year, are not expanded for carbon capture, baseload point sources such as biomass CHP and process emissions from industry are fully (partially) developed as soon as the transport of \COtwo{} (\Htwo{}) is permitted.}
+While peak load technologies such as gas-fired combined heat and power (CHP) plants, which are only in operation on a few days a year, are not expanded for carbon capture, baseload point sources such as biomass CHP and process emissions from industry are fully (partially) developed as soon as the transport of \carbon{} (\hydrogen{}) is permitted.}
     % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
@@
 Except for SMR and gas-fired Combined Heat and Power Plants (CHP) which operate a few weeks annually, all emissions from carbon-emitting technologies are captured.
-% In comparison with the \hydrogenscenario, Biomass CHP supplies more flexible heat, power and relatively low-cost carbon.
 Thus, only a few Direct Air Capture (DAC) installations are installed, such as those in southern Spain for methanolization plants.
@@
     \includegraphics[width=0.9\linewidth]{difference/comparison/emission-reduction-0.1/figures/90_nodes/cost_bar.png}
-    \caption[short]{Net change in system cost when tightening the \COtwo{} emission target from net-zero to net-negative 10\% of 1990s emissions.
+    \caption[short]{Net change in system cost when tightening the \carbon{} emission target from net-zero to net-negative 10\% of 1990s emissions.
 Each cost bar is split into contribution of the same technology groups, except for \carbon{} sequestration cost, which are now displayed separately.
@@
 To a small extend it is supplied by additional carbon captured and transported from bio-energetic sources in the inland that were used to supply CU before (compare Fig.~\ref{fig:balance_maps_full_nn}).
-At the same time, some of the CU in Central Europe moves to Spain, where it uses carbon from new DAC facilities without the installation of a \COtwo{} network.
+At the same time, some of the CU in Central Europe moves to Spain, where it uses carbon from new DAC facilities without the installation of a \carbon{} network.
 This slight shift in CU deployment leads to less investments in the hydrogen transport route from the Iberian Peninsula to Central Europe (2~bn€/a) and to an additional investment in the \carbongrid{} from Central and Eastern Europe to the shores (1.2~bn€/a, compare Fig.~\ref{fig:cost_bar_transmission_nn}).
@@
 
-Our study assesses the roles of hydrogen (\Htwo{}) and carbon dioxide (\COtwo{}) networks in Europe's future energy system under net-neutrality and net-negativity targets for \carbon{} emissions.
+Our study assesses the roles of hydrogen (\hydrogen{}) and carbon dioxide (\carbon{}) networks in Europe's future energy system under net-neutrality and net-negativity targets for \carbon{} emissions.
 Both networks have strong impact of the optimal deployment of carbon capture, carbon utilization and sequestration.
@@
 
-In summary, our study underscores the need for strategic planning and integration of both \COtwo{} and \Htwo{} networks in the European energy landscape.
-The flexibility and cost-effectiveness provided by these networks are pivotal for realizing the EU's climate targets.
-By harnessing the synergies between \Htwo{} and \COtwo{} transportation and management, Europe can effectively reduce system costs while ensuring a robust and sustainable energy future.
+In summary, our study highlights the need for strategic planning and integration of both \carbon{} and \hydrogen{} networks in the European energy landscape.
+The flexibility and cost-effectiveness provided by these networks are essential to achieve EU's climate targets.
+By harnessing the synergies between \hydrogen{} and \carbon{} transportation and management, Europe can effectively reduce system costs while ensuring a robust and sustainable energy system.
 
@@
 % with transport costs considered? - how high are they?
-% CU: The processed fuels are not limited in their total quantity of use or production. -> implicitly limited through demand
 
@@
 Heat demand is regionally subdivided into shares of urban, rural and industrial sites.
-The system exogenously produces 633~Mt \COtwo{} per year from industry, aviation, shipping and agriculture, 153~Mt of which are fossil-based process emissions.
+The system exogenously produces 633~Mt \carbon{} per year from industry, aviation, shipping and agriculture, 153~Mt of which are fossil-based process emissions.
 Location for industrial clusters are taken from~\cite{hotmaps_industrial_db}. Energy demand for industries are calculated from~\cite{mantzosJRCIDEES20152018}.
@@
 The power flow is based on the lossless linearized power flow approximation~.
-The topology of the \COtwo{}, \Htwo{} and gas network is identical to the topology of the electricity network, connecting all neighboring regions.
+The topology of the \carbon{}, \hydrogen{} and gas network is identical to the topology of the electricity network, connecting all neighboring regions.
 Liquid fuels like oil, methanol and Fischer-Tropsch (FT) fuels are not spatially resolved, since transport costs per unit of energy are negligible due to their high energy density.
-Throughout this study, we focus on \COtwo{} and \Htwo{} networks because of their high relevance for infrastructure investments and thus public policy decisions.
+Throughout this study, we focus on \carbon{} and \hydrogen{} networks because of their high relevance for infrastructure investments and thus public policy decisions.
 The electricity network and gas network are both included in the model with full geographical detail, but not further analyzed.
@@
 Moving up the scale, gas used in industrial applications and biomass employed for industrial processes represent more costly yet viable options to capture carbon.
-DAC, an emerging technology capable of extracting \COtwo{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand.
-Finally, based on our model results, biogas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per marginal ton of captured \COtwo{}.
+DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand.
+Finally, based on our model results, biogas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per marginal ton of captured \carbon{}.
 Biogas input is a high-cost fuel which based on the endogenous modeling decisions is not used in large quantities.
 To the extent that biogas-to-gas facilities are used by the model to supply additional (carbon-neutral) gas as fuel, adding CC infrastructure incurs low costs (and thus ranges on the left-hand side of the merit order curve).
-However, the price of capturing additional \COtwo{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the marginal cost of captured \COtwo{}.
+However, the price of capturing additional \carbon{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the marginal cost of captured \carbon{}.
 This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.
-If we consider the spatial aspect of distributed carbon capture potentials, the \COtwo{} network can exploit comparative cost advantages between CO2 bidding zones to utilize the lowest-cost CC facilities across the continent.
+If we consider the spatial aspect of distributed carbon capture potentials, the \carbon{} network can exploit comparative cost advantages between CO2 bidding zones to utilize the lowest-cost CC facilities across the continent.
 We assume a capture rate of 90\% for CC on process emissions, SMR, biogas-to-gas, as well as gas and biomass used in industry, and 95\% for CC on biomass and gas CHPs.
@@
 Our estimates for carbon sequestration potentials are restrictive, limiting the annual sequestration potential to 25~Mt per region and constraining it further such that this amount could be  25 years.
-Furthermore, we decide to limit the total amount of sequestered \COtwo{} to 200~Mt per year for our net-zero \carbon{} emission target and 660~Mt per year for our net-negative emission target.
+Furthermore, we decide to limit the total amount of sequestered \carbon{} to 200~Mt per year for our net-zero \carbon{} emission target and 660~Mt per year for our net-negative emission target.
 This is enough to offset hard-to-abate fossil process emissions (e.g., calcination of limestone in cement manufacturing) and some limited slack of fossil fuels to be used where they can be most valuable.
@@
 Demand and emissions from industrial clusters heavily drive the modelling results.
-Allowing the model to relocate these, incorporate flexibility measures or alter underlying processes, may lead to less dependency on \COtwo{} and \Htwo{} networks.
+Allowing the model to relocate these, incorporate flexibility measures or alter underlying processes, may lead to less dependency on \carbon{} and \hydrogen{} networks.
 The transport of fuel (FT, gas, oil) between the regions is not limited which may overestimate their flexibility.
@@
     \includegraphics*[width=0.4\linewidth]{comparison/emission-reduction/figures/90_nodes/objective_heatmap.png}
-    \caption{Total annual system cost for the different network models and \COtwo{} targets.
-While the \modBase{} has neither a \COtwo{} nor an \Htwo{} network, the \modHybrid{} is allowed to expand both.}
+    \caption{Total annual system cost for the different network models and \carbon{} targets.
+While in the \baselinescenario{}, the model has neither a \carbon{} nor an \hydrogen{} network, it can expand both in the \hybridscenario{}.}
     \label{fig:objective_heatmap}
@@
     \includegraphics[width=\linewidth]{difference/h2-only-full/figures/90_nodes/cost_map.png}
-    \caption{Difference in regional costs between the \Hgrid{} and \modHybrid{}s in the Net-Zero scenario.
-The left figure shows higher spendings per technology and region and transport system for the \modH{}, the right shows higher spendings in the \modHybrid{}.}
+    \caption{Difference in regional costs between the \hydrogengrid{} and \hybridscenario{} assuming net \carbon{} neutrality.
+The left figure shows higher spendings per technology and region and transport system in the \hydrogenscenario{}, the right shows higher spendings in the \hybridscenario{}.}
     \label{fig:cost_map_difference}
@@
     \includegraphics[width=\linewidth]{difference/net-negative-0.1-h2-only-full/figures/90_nodes/cost_map.png}
-    \caption{Difference in regional costs between the \Hgrid{} and \modHybrid{}s in the net \carbon{} removal scenario.
-The left figure shows higher spendings per technology and region and transport system for the \modH{}, the right shows higher spendings in the \modHybrid{}.}
+    \caption{Difference in regional costs between the \hydrogengrid{} and \hybridscenario{} in the net \carbon{} removal scenario.
+The left figure shows higher spendings per technology and region and transport system for the \hydrogenscenario{}, the right shows higher spendings in the \hybridscenario{}.}
     \label{fig:cost_map_difference_nn}
@@
 
-In the \modBase{}, the capture share is highest for biogas-to-gas facilities, followed by process emissions, biomass for industry, biomass CHPs, gas for industry, SMR and gas CHPs.
-For models with additional transport systems, we observe a correlation between CC share and capacity factors of the underlying technologies: Process emissions as well as gas and biomass for industry have high capacity factors (above 80\%), and CC shares close to 100\%, the only exception being gas for industry in the \modH{}.
-Biomass CHPs have a capacity factor of around 50\%, and CC shares of 50\% in the \modH{} and 97\% in the \modCO{}.
-SMR and gas CHPs have capacity factors around 15\%, and CC shares at nearly 0\% for the \baselinemodel.
+In the \baselinescenario{}, the capture share is highest for biogas-to-gas facilities, followed by process emissions, biomass for industry, biomass CHPs, gas for industry, SMR and gas CHPs.
+For models with additional transport systems, we observe a correlation between CC share and capacity factors of the underlying technologies: Process emissions as well as gas and biomass for industry have high capacity factors (above 80\%), and CC shares close to 100\%, the only exception being gas for industry in the \hydrogenscenario{}.
+Biomass CHPs have a capacity factor of around 50\%, and CC shares of 50\% in the \hydrogenscenario{} and 97\% in the \carbonscenario{}.
+SMR and gas CHPs have capacity factors around 15\%, and CC shares at nearly 0\% for the \baselinescenario{}.
 The capacity factor and CC share of SMR increases to around 20\% and 50-65\%, respectively, in the other models.
@@
 Fig.~\ref{fig:cost_bar_subsidy} shows the corresponding total system cost of all models in the Net-Zero scenario.
-Fig.~\ref{fig:cost_bar_diff_subsidy} shows the net change in system cost between non-subsidized and subsidized models, showing only the \carbonmodel{} and \hybridmodel{} where changes occur.
+Fig.~\ref{fig:cost_bar_diff_subsidy} shows the net change in system cost between non-subsidized and subsidized models, showing only the \carbongrid{} and \hybridscenario{} where changes occur.
 
@@
 \end{subfigure}
-\caption{System costs for the all models in the net-zero emission scenario with 50\% subsidy of \carbon{} pipelines (a) and change in system cost for the \carbonmodel{} and \hybridmodel{} when subsidizing \carbon{} pipelines by 50\% (b).}
+\caption{System costs for the all models in the net-zero emission scenario with 50\% subsidy of \carbon{} pipelines (a) and change in system cost for the \carbongrid{} and \hybridscenario{} when subsidizing \carbon{} pipelines by 50\% (b).}
 \end{figure}
@@
 \clearpage
-\subsection{Operation of the \hybridmodel{} in Net-Zero Scenarios}
+\subsection{Operation in the \hybridscenario{}}
 
@@
     \includegraphics[width=.8\linewidth]{full/figures/90_nodes/sankey_diagramm.png}
-    \caption{Sankey diagram of the carbon flows in the \hybridmodel{} in the Net-Zero scenario.}
+    \caption{Sankey diagram of the carbon flows in the \hybridscenario{} assuming net \carbon{} neutrality.}
     \label{fig:sankey_diagramm}
@@
     \end{subfigure}
-    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \carbonmodel{} in the net negative scenario.
+    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors in the \carbonscenario{} in the net negative scenario.
 For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
@@
     \end{subfigure}
-    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hydrogenmodel{} in the net negative scenario.
+    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors in the \hydrogenscenario{} in the net negative scenario.
 For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
@@
     \end{subfigure}
-    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hydrogenmodel{} in the net \carbon{} removal scenario.
+    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hydrogenscenario{} in the net \carbon{} removal scenario.
 For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
@@
 Lines and arrows show the interregional transportation.
-With the tightened emission target, the \hybridmodel{} decreases the \hydrogengrid{} layout in comparison to the Net-Zero scenario and increases \carbon{} transport from inland point sources to the sequestration sites.
+With the tightened emission target, the model decreases the \hydrogengrid{} layout in comparison to the Net-Zero scenario and increases \carbon{} transport from inland point sources to the sequestration sites.
     }
@@
     \includegraphics[width=.8\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
-    \caption{Sankey diagram of the carbon flows in the \hybridmodel{} in the Net-Negative scenario.}
-    \label{fig:sankey_diagramm}
+    \caption{Sankey diagram of the carbon flows in the \hybridscenario{} in the Net-Negative scenario.}
+    \label{fig:sankey_diagramm_nn}
 \end{figure}
@@
 \end{document}
-
-% LEFT OVER
-% =========
-
-% To understand the dynamics of the \modCO{}, we observe that the system locates CU in the regions where it can minimize costs for its material inputs, carbon and hydrogen.
-As the \COtwo{} network largely smoothes out carbon prices across regions, CU is located in the hydrogen price valleys on the periphery of the continent.
-Similarly, the model locates CS in regions where it can minimize the costs of the carbon to be sequestered.
-Thus, CS is located in the carbon price valleys among the subset of regions with access to offshore sequestration sites.
-Due to the \COtwo{} network these low-price regions are not strongly pronounced, but still prevalent and impact CS locations.
-DAC is only used for CU and thus co-located in regions with a mix of low electricity costs (for running DAC) and low hydrogen costs.
-
-% To explain the dynamics of the \modH{}, we observe that the model locates CU in the carbon price valleys across Western, Central and Eastern Europe (see top right), as the \Htwo{} network smoothes out hydrogen prices across the entire continent (see bottom right).
-CS is again located in the carbon price valleys among those regions with access to offshore sequestration sites.
-Due to the lack of a \COtwo{} network the full CC potentials from point sources are exhausted in all offshore sequestration regions, and the carbon is then sequestered.
-The additionally needed carbon to reach 200~Mt of CS is captured via DAC in the regions with access to offshore sequestration sites with the lowest costs for DAC, with the largest share deployed in Portugal (see top right).
-DAC is also used for CU in this model, and for both purposes DAC is located in the regions with the lowest costs for electricity.
-
-% The increase in captured carbon from point sources in the models with \carbon{} and/or \hydrogen{} transport is more related to an upgrading of existing industrial or power plant facilities to capture carbon rather than a shift in technologies.
-This is evident when analyzing the capture share of the different CC technologies (see Fig.~\ref{fig:captureshare_line}).
-
-% In summary, the implementation of \COtwo{} and \Htwo{} networks is instrumental in unlocking CC potentials and reducing system costs.
-However, despite the \modCO{} demonstrating higher amounts and shares of CC from point sources and therefore less DAC deployment, the \modH{} achieves even  lower total system costs (see Fig.~\ref{fig:cost_bar}).
-%
-%
-% Comparison carbon and hydrogen network
-%
-% [Dynamics \modCO{} vs \modH{} (old text blocks)]
-% There are four notable dynamics or differences between the two network strategies:
-% First, the production of FT fuels always moves to where low-cost carbon and low-cost hydrogen can be brought together.
-In the \modCO{ }, carbon is moved from Western Europe to the Iberian Peninsula, the British Isles and to Denmark to produce FT fuels centrally in few regions.
-In the \modH{}, hydrogen is transported in the opposite direction to produce FT fuels decentrally across all regions in Central Europe with locally captured \COtwo{} from point sources.
-% Second, the \modCO{ } captures more low-cost carbon from biomass, especially in Central and Eastern Europe, and less high-cost carbon from DAC in Greece and Southern Italy.
-Instead, the \COtwo{} network transports carbon from Central and Eastern Europe to Greece and Italy, where it is sequestered, or feeds into CU together with locally produced, low-cost electrolytic hydrogen.
-The \modH{} directly operates electrolyzers and DAC in Greece and Southern Italy, with electricity from low-cost renewables, and thus avoids significant network buildouts in this region.
-% Third, sequestration in the \modH{} occurs more decentralized, spread out across several regions with access to offshore sequestration sites, making use of point-source CC potentials in each of these regions.
-In the \modCO{ }, sequestration is clustered at fewer sequestration sinks, each collecting low-cost \COtwo{} from neighboring regions.
-However, there is no centralization at only one or two sites, since all the sequestration sites offer the same costs and the system minimizes system-wide carbon transport distances to about eight to ten sequestration sites.
-% And fourth, how the three types of CU are spread across the continent is driven by the costs of carbon and hydrogen (the latter mostly driven by the costs of renewable electricity) as well as by renewables capacity factors: Clearly, all carbonaceous fuel production requires sites where both carbon and hydrogen are available, and they moreover prefer sites with low carbon and hydrogen prices.
-Based on our model assumptions, FT fuel synthesis has very limited flexibility and needs to be run alost at baseload (ramp-down only possible to 90\% of maximum load).
-Thus, for FT fuel synthesis sites in Northern Europe with more electricity from wind power are preferred to sites in Southern Europe, powered mostly by solar.
-Methanathion and Methanolisation both have higher flexibility, being able to ramp down to 50\% load during the night, for example.
-Thus, these fuel production sites prefer the best renewable resource sites in Southern Europe.
-And finally, as methanation produces synthetic methane that needs to be fed into the gas grid, methanation prefers those sites with good interconnection to the existing gas network.
-Since Italy possesses a large gas network with easy interconnection to Central Europe, methanation prefers sites in Southern Italy (and Greece, which would be connected to Southern Italy) over the Iberian Peninsula, where Methanolisation is more prevalent.
-
-% In summary, the implementation of a \COtwo{} network allows for connecting regions with abundant, low-price carbon (from point sources) to regions that can sequester the carbon or that can produce low-cost hydrogen to make CU products (FT fuel, methanol, synthetic methane).
-DAC and electrolysis are deployed in the regions with the best renewable resources to meet carbonaceous fuel demands.
-% In the absence of a \COtwo{} network and with an \Htwo{} network on the other hand, regions with access to offshore sequestration sites capture all carbon from low-cost point sources and then sequester it.
-The system deploys additional DAC facilities at the locations with sequestration access and low-cost renewables to capture the remaining necessary carbon for sequestration.
-Carbonaceous fuel demand is met at the locations which offer the best mix of low-cost carbon from point sources and low-cost hydrogen from renewables, with some additional electrolysis and DAC required to fully meet the demand.
-% Sequestered carbon is nearly always transported before
-
-% Therefore, one can conclude that a \COtwo{} network favors a decentralized CC and a CU centralized at regions with abundant renewable resources.
-An \Htwo{} network favors a centralized electrolysis system at regions with abundant renewable resources and a decentralized CU system.
-
-% Together with understanding the dynamics described above, we can highlight three cost advantages of the \Htwo{} network in contrast to the \COtwo{} network: Capturing carbon at the lowest-cost regions reduces 4~bn€ in spending on DAC.
-However, the \modH{} produces electrolytic hydrogen in the lowest-cost regions and in total spends 7~bn€ less on renewable energy (14~bn€ less on solar, and 7~bn€ more on wind) as well as 1~bn€ less on SMR.
-Furthermore, without an \Htwo{} network, additional \Htwo{} storage is needed in Spain at a cost of 5~bn€ to store hydrogen produced during the summer for utilization in FT fuel production during the winter.
-% And finally, the \modCO{ } spends 7~bn€ more on gas boilers, gas plants, gas infrastructure, and methanation.
-% TODO: why??
-% The costs of the respective networks, on the other hand, are almost equal: the \modCO{ } spends 8~bn€ on the \COtwo{} network, while the \modH{} even spends 10~bn€ on the \Htwo{} network.
-These factors add up to account for 14~bn€ of the difference of 20~bn€ in system costs between the two scenarios (785~bn€ for the \COgrid{} and 765~bn€ for the \modH{}, see Fig.~\ref{fig:objective_heatmap}).
-
-% Net-negative
-
-% Fundamental differences in investment strategies occur in the deployment of solar, carbon capture and bio-energy technologies: While the \carbonmodel{} expands CC at point sources the most, the \hydrogenmodel{} and the \hybridmodel{} invest relatively more in new solar plants.
-The \carbonmodel{} invests three time more in new bio-energy sources (6~bn€) than the other models.
-% All models only reveal small net downscaling of technologies.
-However, as we discuss below, the geographic impact can vary depending on the model, sometimes leading to disruptive changes in optimal network strategies and stranded assets.
-
-% In contrast to its layout in the Net-Zero scenario, the \carbonmodel{} fundamentally reorganizes the carbon transport pathways (see Fig.~\ref{fig:balance_maps_co2_nn}).
-% In Spain, carbon utilization (CU) is now predominantly supplied by carbon extracted from DAC.
-Central Europe sees a major transformation, with carbon mainly being transported to sequestration sites in the north and south-east.
-This reorganization leads to stranded investment in the carbon network infrastructure, amounting to approximately 3~bn€.
-
-% The \modH{} on the other hand, primarily leaves the network topology unchanged and only expands DAC facilities at regions with access to sequestration sites as well as additional solar, wind and heat pumps in the vicinity to supply power and heat (see Fig.~\ref{fig:balance_maps_h2_nn}).
-
-% Finally, the \modHybrid{} expands the sequestration with carbon captured from bio-energetic inputs and transported from Central and Eastern Europe to the near shores (see Fig.~\ref{fig:balance_maps_full_nn}), leading to an additional investment in the \carbongrid{} (1.2~bn€, see Fig.~\ref{fig:cost_bar_transmission_nn}).
-At the same time, carbonaceous fuel production in Central Europe partially moves to Spain, where it uses carbon from new DAC facilities without the installation of a \COtwo{} network.
-As a trade-off, the model transports less hydrogen from the Iberian Peninsula to Central Europe, resulting in a dismantling of 2~bn€ in hydrogen network costs compared to the Net-Zero scenario.
-
-% These findings highlight the continued trends of the two systemic approaches which are combined in the \hybridmodel.
-With an increased demand for sequestration, the \carbongrid{} relatively gains more importance, leading to higher cost-benefits by enabling to combine increased need DAC with a more dominant collection and transportation scheme of carbon from distributed point sources.

### 2023-12-21 report/submissions/journal/cover_letter/letter.tex
prefinal internal reivew

% FDUletter_example.tex - an example latex file to illustrate FDUletter.cls
%
% Template by Brian Wood (brian.wood@oregonstate.edu).  Please feel free to send suggestions for changes; this template/cls is not exactly elegantly done!
% Modified by Huang Weiran (huangweiran1998@outlook.com) to fit the need of FDU students.
% Based on SCU version and delete the watermark
% SHU version modified by Zhendong Li (zhendong.li2001@outlook.com) to fit SHU students.

\documentclass[12pt]{SHUletter}
\usepackage{tikz}
\usepackage{eurosym}
\usepackage{xcolor}
	\definecolor{tubred}{RGB}{153,0,0}
\usepackage{lipsum}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{eso-pic}
%
% This section is just a bunch of busywork so that the second and following pages read ``Page X of Y''
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\rhead{Page \thepage \hspace{1pt} of \pageref{LastPage}}
%
%
% Set custom font here. Comment this line out if you do not have a Cambria font (originally included with this template) installed; computer modern (or whatever your current default font is) will be substituted.
%
%\setmainfont{[Cambria.ttf]}[BoldFont  = [CambriaBold.ttf], ItalicFont  = [CambriaItalic.ttf], BoldItalicFont = [CambriaBoldItalic.ttf] ]

\newcommand{\watermark}[3]{\AddToShipoutPictureBG{
\parbox[b][\paperheight]{\paperwidth}{
\vfill%
\centering%
\begin{tikzpicture}
    \path (0,0) -- (\paperwidth,\paperheight);
    \node[opacity=.06] at (current page.center)
    {\includegraphics[width=0.55\textwidth]{tublogo.pdf.pdf}};   % sculogo watermark
    \end{tikzpicture}
\vfill}}}

% The material below is a whole big dang thing whose purpose is just to set up a fixed coordinate system for \tikz so that you can put the Department or School address in the upper right-hand side without it moving all around every time you change something in the page.  I think it works.
% Defining a new coordinate system for the page:
%
% --------------------------
% |(-1,1)    (0,1)    (1,1)|
% |                        |
% |(-1,0)    (0,0)    (1,0)|
% |                        |
% |(-1,-1)   (0,-1)  (1,-1)|
% --------------------------
\makeatletter
\def\parsecomma#1,#2\endparsecomma{\def\page@x{#1}\def\page@y{#2}}
\tikzdeclarecoordinatesystem{page}{
	\parsecomma#1\endparsecomma
	\pgfpointanchor{current page}{north east}
	% Save the upper right corner
	\pgf@xc=\pgf@x%
	\pgf@yc=\pgf@y%
	% save the lower left corner
	\pgfpointanchor{current page}{south west}
	\pgf@xb=\pgf@x%
	\pgf@yb=\pgf@y%
	% Transform to the correct placement
	\pgfmathparse{(\pgf@xc-\pgf@xb)/2.*\page@x+(\pgf@xc+\pgf@xb)/2.}
	\expandafter\pgf@x\expandafter=\pgfmathresult pt
	\pgfmathparse{(\pgf@yc-\pgf@yb)/2.*\page@y+(\pgf@yc+\pgf@yb)/2.}
	\expandafter\pgf@y\expandafter=\pgfmathresult pt
}
\makeatother

%%------------------------------------------------------------------------------%%
%

%%%%%%%%%%% Put Recommender(your mentor) Information Here %%%%%%%%%%%
%
\def\name{Philipp Glaum \& Fabian Hofmann
}
%
% List your degree(s), licences, etc. here if you like.
%\def\What{, Your degrees, etc.}
%
% Set the name of your Department or School here
% I honestly don't know why the negative spacing is necessary to get the alignment of the first line correct.  This must be a ``\tikz'' thing.
%%%%%%%%%%%%%%%%%%  School or Department %%%%%%%%%%%%%%%
\def\Where{\hspace{-1.2mm}\textbf{\color{tubred}
	Department of Digital\\Transformation in Energy Systems,\\ TU Berlin
}}

%%%%%%%%%%%%  Additional Contact Information %%%%%%%%%%%
%
% Set your preferred primary contact address here.
\def\Address{Straße des 17. Juni 135
}
%
\def\CityZip{Berlin, 10623 \\
	Germany}
%
% Set your e-mail here
\def\Email{\textbf{\color{tubred}E-mail}: p.glaum@tu-berlin.de}
%
% Set your preferred contact number here
%\def\TEL{\textbf{\color{tubred}Tel/Fax}: +86 28 85412720}
%
% Set your department or personal website here
\def\URL{\textbf{\color{tubred}URL}: {https://www.ensys.tu-berlin.de}}
%

%%%%%%%%%%%%%%%%  Footer information  %%%%%%%%%%%%%%%%%%
%
%  The next line is for your college, used as a footer.  If you prefer not to have this, just comment out these lines in favor of the line labeled "[[Alternate]]" below
\def\school{\small{
		TU Berlin $\cdot$
		~Department of Digital\\ Transformation in Energy Systems $\cdot$
		~25, Einsteinufer  $\cdot$
		~Berlin, Germany} }
% \def\school{~}  % [[Alternate]]
%

%%%%%%%%%%%%%%%%%%%%%  Signature line  %%%%%%%%%%%%%%%%%%%%%
%
% Set your signature line here.
% One can add a signature image in a PDF file using the following code; this requires a file called "signature_block.pdf" to be installed in the same folder as the .tex file.  The vertical spacing (\vspace) and the scaling will have to be adjusted to get things to look correct for your particular signature image. Alternatively, comment out the following line in favor of the one labeled "[[Alternate]]" if you want to sign a paper copy of the letter.
%
\signature{
	\name
}
%\signature{\name}  % [[Alternate]]
%%------------------------------------------------------------------------------%%

% This block sets up the address on the right-hand side of the header.
%
% The following lines just compile the information you set up into the LaTex letter variable "address" for later use.
%
%The following command "clears out" the default address so that it can be better set using \tikz
\address{}

\def\newaddress{
	\Where\\
	\Address\\
	\CityZip\\
	%\TEL\\
	\Email\\
	\URL
}
%
%%%%%%%%%%%  DATE  %%%%%%%%%%%%%%%%%%%%%%%%%
% If you want a date different from the current date, comment out the next line in favor of the line labeled "[[Alternate]]".  The ``\vspace{10mm}'' just moves the date down a tiny bit so it doesn't interfere with the header.  This can be adjusted to your preference.
%
\date{\vspace{11mm} \today}
%\date{\vspace{10mm} 20 September 2020}  %[[Alternate]]
%
%%%%%%%%%%% Set the subject here if there is one  %%%%
%\subject{Stuff} % optional subject line

%%-----------------------------------------------------------------------------------------%%
%
\begin{document}
	%
	%
	%%%%%%%%  The "To" address goes here.
	%
   %Also, you can use this
	% \begin{letter}{
	% 		Target University\\
	% 		Some Addresss\\
	% 		SomeTown, SomeState 					       				  		 ~~SomeZip
	% 	}
 \begin{letter}{Dear Editor,}

		% This line sets up the return address to the right-side of the OSU logo.  The location is set with absolute node addresses using ``\tikz''.  It can still be a bit fussy, and you may need to alter this a little to get things to look right.  The bit that changes the position are the numbers in parentheses ``at (14.2,2.7)''
		%
		\begin{tikzpicture}[remember picture,overlay,,every node/.style={anchor=center}]
		\node[text width=7cm] at (page cs:0.5,0.73){\small \newaddress};
		\end{tikzpicture}

		%%%%%%  The ``opening'' is just the method of address you would like to use at the start of the letter.
		%
		\opening{}

		%%%%%%%%%% Body of letter   %%%%%%%%%%%%%%
		% Remove it if you do not want watermark
		%\watermark{}{}{}

		%%%%%%%%%%% 	Text   %%%%%%%%%%%%%%
		Thank you very much for the opportunity to submit our manuscript to Applied Energy. Hereby we submit the paper “Leveraging the Existing German Transmission Grid with Dynamic Line Rating”.

		In our work, we present the first large-scale investment model that evaluates the benefits of Dynamic Line Rating (DLR) for renewable energy integration.
		Taking Germany as an example, the model includes a high temporal and spatial resolution as well as a high level of detail to enable a realistic representation of the system dynamics.
		The paper appears at an important time as  politicians and Transmission System Operators (TSOs) have recognized the problem of lacking transmission capacities when moving towards high renewable shares in Europe and all over the world.
		% In many official reports DLR is considered as a promising measure to relief the transmission grid.

		Our analysis not only confirms and quantitatively describes transmission system benefits but also shows how the implementation of DLR affects the optimal system design for different shares of renewable power production. By better integrating wind power resources, especially offshore, DLR leads to significant cost savings of 5\% of the system cost, mainly due to the reduced need for cost-driving energy storage. This key finding can be generalized to all countries with normal to high wind resources and is therefore of particular importance for international network operators and the research community. Furthermore, we provide an open-source workflow which can be used to reproduce the results and extend the model to other countries.

		The paper was proofread by Prof. Tom Brown, an energy system expert and native tongue speaker, as well as from Rena Kuwahata from Ampacimon, an expert on transmission lines and DLR applications.

		We want to thank the editor for taking the time to consider the publication of our paper in Applied Energy.

		% What is the novelty of this work? Is the paper appealing to a popular or scientific audience?
		% Why the authors think the paper is important and why the journal should publish it? Has the article
		% been checked by a native tongue speaker with expertise in the field? Are you available as a reviewer
		% for at least three other articles for Applied Energy during the current year? In addition to answering
		% those five questions, the authors should also describe in one or two paragraphs the significance of
		% their work and what new information is described in the manuscript.

		%%%%%%% ``closing'' sets the sign-off line.
		\closing{Sincerely,}
		% Comment out/in the lines below as necessary
		%\encl{If an enclosure is provided, let them know what it is.}
		%\ps{A postscript if that is a thing you do.}
		%\cc{Someone Who Cares (and is copied).}
	\end{letter}

\end{document}
%
%%-----------------------------------------------------------------------------------------%%

### 2023-12-21 report/submissions/journal/paper/paper.tex
prefinal internal reivew

@@
 
-\title{Evaluating Hydrogen and Carbon Network Strategies for the European Energy System}
+\title{\hydrogen{} and \carbon{} Network Strategies for the European Energy System}
 % \title{Competition and Synergies of \hydrogen{} and \carbon{} Networks in Europe}
@@
 \begin{abstract}
-    Hydrogen and carbon dioxide transport can both play an important role in climate-neutral energy systems.
+    Hydrogen and carbon dioxide transport can both play an essential role in climate-neutral energy systems.
 Hydrogen networks help serve regions with high energy demand, while excess emissions are transported away in carbon dioxide networks.
-Regarding the synthesis of carbonaceous fuels, it is less clear which input should be transported: hydrogen to carbon point sources or carbon to low-cost hydrogen.
-We explore the potential synergies and competition of both networks in a cost-optimal carbon-neutral European energy system.
+For the synthesis of carbonaceous fuels, it is less clear which input should be transported: hydrogen to carbon point sources or carbon to low-cost hydrogen.
+We explore both networks' potential synergies and competition in a cost-optimal carbon-neutral European energy system.
 In direct comparison, a hydrogen network is more cost-effective than a carbon network, since it serves to transport hydrogen to demand and to point source of carbon for utilization.
 However, in a hybrid scenario where both networks are present, the carbon network effectively complements the hydrogen network, promoting carbon capture from biomass and reducing reliance on direct air capture.
-This study shows how the European energy system in future net-zero and net-negative carbon dioxide emissions scenarios benefits from integrating both networks.
+Our analysis suggests integrating hydrogen and carbon dioxide networks into European energy policy for a robust, carbon-neutral or carbon-negative future energy system.
 \end{abstract}
@@
 While management of both carbon dioxide (\carbon{}) and hydrogen (\hydrogen{}) is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructure effectively interacts with carbon management technologies, including carbon capture, transport, storage, utilization, and sequestration.
-Hydrogen is considered in several sectors in industry not amenable to electrification, and offers an efficient way to transport and store energy over long distances and periods, and can be produced from electrolysis or steam methane reforming (SMR), for example.
-Captured carbon can be stored in geological formations to permanently extract it from the atmosphere, a process known as carbon sequestration (CS).
-Additionally, carbon can be combined with hydrogen to produce carbonaceous fuels to supply for example synthetic kerosene for aviation, synthetic methanol for shipping or synthetic methane for industrial feedstock.
-From hereon, we jointly refer to the production of these carbonaceous fuels as Carbon Utilization (CU).
+Hydrogen is being considered in several industries that are not eligible for electrification.
+It offers an efficient way to transport and store energy over long distances and periods, and can be produced from electrolysis or steam methane reforming (SMR), for example.
+Captured carbon can be stored in geological formations to extract it from the atmosphere permanently, a process known as carbon sequestration (CS).
+Additionally, carbon can be combined with hydrogen to produce carbonaceous fuels, such as synthetic kerosene for aviation, synthetic methanol for shipping or synthetic methane for industrial feedstock.
+We jointly refer to the production of these carbonaceous fuels as Carbon Utilization (CU).
 Carbon can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, or synthetic carbonaceous fuels through point-source carbon capture (CC) techniques, or harvested from the atmosphere using direct air capture (DAC).
@@
 
-Recently, policymakers and industry in Europe have started developing hydrogen and carbon management strategies~\cite{GermanyDevelopingStrategy2023,CarbonManagementStrategie}, planning infrastructure components~\cite{CONetz}, and committing to the first carbon utilization projects~\cite{EFuelsPilotPlant2022,OrstedAssumesFull,GROUNDBREAKINGEFUELPRODUCTION,DLREfuelsDLR}.
+Recently, policymakers in Europe have started developing hydrogen and carbon management strategies~\cite{GermanyDevelopingStrategy2023,CarbonManagementStrategie}, planning infrastructure components~\cite{CONetz}, and committing to the first carbon utilization projects~\cite{EFuelsPilotPlant2022,OrstedAssumesFull,GROUNDBREAKINGEFUELPRODUCTION,DLREfuelsDLR}.
 With the European Union's goal of achieving climate neutrality by 2050, a wide range of programs, funding models and initiatives have been established with dedicated instruments supporting both the ramp-up of the hydrogen and carbon economy~\cite{eu2023netzero,europeangreendeal,europeaninnovationfund}.
+% TODO: split into hydrogen and carbon?
+% What about NZIA "proposes a CO2 injection target of 50 million tonnes per year by 2030, which will be stored within the European Union (EU)"
+% What about RED III targets, 10 MtH2/a domestic in 2030? Name a few targets rather than hiding everything in references - readers won't go through all references..
+
 Collaborative industry initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen Infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier.
@@
 A hydrogen network can supply regions with geographically fixed hydrogen demand, such as for steel production, with hydrogen from regions with the best renewable resources as well as enable CU at the site of point-source CC.
-On the other hand, the carbon transport system allows for transporting captured carbon to regions with sequestration potentials or high-quality renewable resources, the latter enabling cost-effective CU.
+On the other hand, the carbon transport system allows transporting captured carbon to regions with sequestration potentials or high-quality renewable resources, the latter enabling cost-effective CU.
 
-In the literature, the two network approaches and underlying technologies have been discussed in several of publications, all of which, however, dealt with the isolated aspects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022,neumannBenefitsHydrogenNetwork2022}.
-Such techno-economic models, in contrast for example to Integrated Assessment Models, can account for the spatial distribution of carbon sources and sinks which are crucial to provide a holistic view of the energy system and its technological interactions.
+In the literature, the two network approaches and underlying technologies have been discussed in several publications, all of which, however, dealt with the isolated aspects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022,neumannBenefitsHydrogenNetwork2022}.
+Such techno-economic models, in contrast for example to Integrated Assessment Models, can account for the spatial distribution of carbon sources and sinks, which are crucial to provide a holistic view of the energy system and its technological interactions.
 Neumann et al.~\cite{neumannBenefitsHydrogenNetwork2022} demonstrated the interchangeability of hydrogen and electricity grid expansion in a climate-neutral European energy system at the expense of higher investments.
-The underlying, highly-resolved model encompasses most relevant sectors, however it neglects the option to transport carbon across regions.
+The underlying, highly-resolved model encompasses the most relevant sectors, however it neglects the option to transport carbon across regions.
 Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \carbon{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
-This limited sectoral scope cannot capture important dynamics of carbon management, since it neglects the sectors like industry that will need to handle most \carbon{} in the future.
-A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
-However, this limited spatial scope fails to consider other sequestration sites and co-benefits from connecting the \carbon{} network to neighboring countries.
+This limited sectoral scope cannot capture important dynamics of carbon management, since it neglects sectors like industry that will need to handle most \carbon{} in the future.
+A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connected to a remote sequestration site in Norway.
+However, this limited spatial scope fails to consider other sequestration sites and the co-benefits from connecting the \carbon{} network to neighboring countries.
 Other high-resolution energy system models do not feature detailed CU technologies and \carbon{} transport.
-Euro-Calliope, for example, does neither feature carbon nor hydrogen transport, and thus forces generic synthetic fuel production to rely on captive electrolysis and DAC in the same region~\cite{pickeringDiversityOptionsEliminate2022}.
+Euro-Calliope, for example, does not feature carbon nor hydrogen transport, and thus forces generic synthetic fuel production to rely on captive electrolysis and DAC in the same region~\cite{pickeringDiversityOptionsEliminate2022}.
 
@@
 In this paper, we present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies.
-The study is conducted using the PyPSA-Eur energy system model which optimizes operation and investments in all relevant energy sectors to supply projected energy demands for 2050 (see Fig.~\ref{fig:total-demand-bar}).
+The study is conducted using the PyPSA-Eur energy system model~\cite{brownPyPSAEurOpenSectorcoupled2023} which optimizes operations and investments in all relevant energy sectors to supply projected energy demands for 2050 (see Fig.~\ref{fig:total-demand-bar}).
 It is assumed that Europe is self-sufficient in energy and does not import any fuels.
-This drives carbonaceous fuels production to be located at sites where both \carbon{} and \hydrogen{} can be provided at low cost.
+This drives carbonaceous fuel production to be located at sites where both \carbon{} and \hydrogen{} can be provided at low cost.
 
@@
     \caption{Assumptions on exogenous demand, derived from~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023}.
-The figure shows the total annual energy demand for each energy source which determine the model's endogenous investments and operation.
+The figure shows total annual energy demands for each energy source, which determine the model's endogenous investments and operation.
 Endogenous processes can lead to higher total production volumes of some energy carriers, e.g., the demand for methanol requires more hydrogen and carbon as secondary (energy) inputs, which are not considered here.
@@
 
-To investigate the competing transport dynamics between \carbon{} and \hydrogen{} networks, we contrast the following four model scenarios:
+To investigate the competing transport dynamics between \carbon{} and \hydrogen{} networks, we contrast the following four scenarios:
 %
@@
 
-We analyze how the deployment of carbon and hydrogen networks affects different decarbonization strategies in the models and scenarios and how investments in none or only one of the transport technologies lead to inefficient market layouts.
-We do this by focusing first on a net-zero \carbon{} emissions target, while limiting carbon sequestration to a total 200~Mt/a.
+We analyze how the deployment of carbon and hydrogen networks affects different decarbonization strategies in the scenarios and how investments in none or only one of the transport technologies lead to inefficient market layouts.
+We do this by focusing first on a net-zero \carbon{} emissions target, while limiting carbon sequestration to 200~Mt/a.
 We give a detailed description of the different technological and geographical impacts of the transportation systems.
@@
 Costs in the \baselinescenario{} are the highest at 779~bn€/a.
-Models with \carbon{} and \hydrogen{} transport options are more cost-effective, reducing costs by 3.1\% in the \carbongrid{}, 4.8\% in \hydrogengrid{}, and 5.2\% in \hybridscenario{}.
+Scenarios with carbon and hydrogen networks are more cost-effective, reducing costs by 3.1\% in the \carbongrid{}, 4.8\% in the \hydrogengrid{}, and 5.2\% in the \hybridscenario{}.
 
@@
     \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/cost_bar.png}
-    \caption[short]{Total annual system cost subdivided into groups of technologies for the different models of the European energy system with a net-zero emission target.
-While in the \baselinescenario{}, the model has neither a \carbon{} nor an \hydrogen{} network, it can expand both in the \hydrogenscenario{}.
+    \caption[short]{Total annual system cost, subdivided into groups of technologies for the different models of the European energy system with a net-zero emission target.
+While in the \baselinescenario{}, the model has neither a carbon nor a hydrogen network, it can expand both in the \hybridscenario{}.
 ``Gas Infrastructure'' combines gas facilities for power and heat production, ``\carbon{} Infrastructure'' and ``\hydrogen{} Infrastructure'' combine transport and storage for each carrier.
@@
 In contrast, the other models have higher combined pipeline costs, peaking at 20\% in the \hybridmodel.
-Both carbonaceous fuel production and electrolysis account for about 4\% (30 bn€) and 4.7\% (37 bn€) of costs respectively, with heat pump installations contributing around 8\% (64 bn€ in the \baselinemodel).
+Both carbonaceous fuel production and electrolysis account for about 4\% (30 bn€) and 4.7\% (37 bn€) of costs respectively, with heat pump installations contributing around 8\% (64 bn€/a in the \baselinemodel).
 
-System cost differences between the models are driven by wind and solar as well as carbon capture technologies.
-In scenarios with \carbon{} and \hydrogen{} transportation, the model primarily reduces the reliance on high-cost DAC from 392~Mt/a carbon removal in the \baselinescenario{} to 112~Mt/a in the \hybridscenario{} (see Fig.~\ref{fig:balance_captured_carbon}).
-At the same time, carbon capture from biomass combustion, gas-based industry processes and biogas-to-gas upgrading increase.
-The increased usage of upgraded biogas leads to a lower synthetic methane production.
+System cost differences between the scenarios are driven by wind and solar, as well as carbon capture technologies.
+In scenarios with carbon and hydrogen networks, the model primarily reduces the reliance on high-cost DAC, from 392~Mt/a carbon removal in the \baselinescenario{} to 112~Mt/a in the \hybridscenario{} (see Fig.~\ref{fig:balance_captured_carbon}).
+At the same time, carbon capture from biomass combustion, gas-based industry processes and biogas-to-gas upgrading increase, the latter lowering the need for synthetic methane production.
 % Another knock-on effect of less reliance on DAC is its reduced endogenous demand for dedicated renewables as well as heat pumps and biomass CHPs to supply power and heat.
@@
     \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/balance_bar_carbon.png}
-    \caption{Balance of captured carbon for all models in the Net-Zero scenario.
+    \caption{Balance of captured carbon for all scenarios assuming net \carbon{} neutrality.
 Positive values indicate carbon capture, negative values indicate carbon consumption.
-By integrating \hydrogen{} and \carbon{} networks, the predominant method for carbon removal shifts from Direct Air Capture (DAC) to bio-energetic processes with capture.
-At the same time the reliance on methanation decreases.}
+By integrating hydrogen and carbon networks, the predominant method for carbon removal shifts from Direct Air Capture (DAC) to bio-energetic processes with capture.
+At the same time, the reliance on methanation decreases.}
     \label{fig:balance_captured_carbon}
@@
 In the \baselinescenario{}, where there is no carbon and hydrogen transport, hydrogen must be produced where it is consumed.
-For CU, either \hydrogen{} must be produced where \carbon{} is cheap at point sources, or DAC must be used at sites with low-cost \hydrogen{}.
-Where \carbon{} from point sources is not able to be sequestered, it has to be compensated elsewhere by negative emissions from DAC or biomass capture with sequestration.
-All of these options are sub-optimal.
+For CU, either hydrogen must be produced where carbon is cheap at point sources, or DAC must be used at sites with low-cost hydrogen.
+Where carbon from point sources cannot be sequestered, it has to be compensated elsewhere by negative emissions from DAC or biomass capture with sequestration.
+All of these options are sub-optimal from a cost-perspective.
 
@@
 Both a carbon network and a hydrogen network can help alleviate these inefficiencies.
-A carbon network allows \carbon{} from point sources to be transported to low-cost hydrogen as well as sequestration sites.
-A hydrogen network on the other hand allows low-cost hydrogen to be transported from regions with good renewable resources to point sources for CU.
+On the one hand, a carbon network allows carbon from point sources to be transported to low-cost hydrogen as well as to sequestration sites.
+On the other hand, a hydrogen network allows low-cost hydrogen to be transported from regions with good renewable resources to point sources for CU.
 This results in fundamentally different network flows in each scenario (see Fig.~\ref{fig:balance_map}).
@@
     \centering
-    \includegraphics[width=0.98\linewidth]{comparison/single-technologies/figures/90_nodes/balance_map_dedicated.png}
+    \includegraphics[width=\linewidth]{comparison/single-technologies/figures/90_nodes/balance_map_dedicated.png}
     \caption{Average production, consumption, flows and prices of the carbon (top line) and hydrogen (bottom line) sectors in the \carbongrid{} (left) and the \hydrogenscenario{} (right).
@@
 Lines and arrows show the interregional transportation.
-Carbon sequestration offshore are drawn as full circles.
+Carbon sequestration offshore is drawn in full circles.
     }
@@
 
-In the \carbonscenario, there are three major transport routes of the low-cost carbon captured in Central and Eastern Europe (see Fig.~\ref{fig:balance_map}, top left).
-A large part of the carbon is transported from Western Europe to the Iberian Peninsula (8~kt/h), the British Isles (7.5~kt/h) and Denmark (2.8~kt/h), where it is used along with low-cost electrolytic hydrogen to produce Fischer-Tropsch fuels and methanol.
-A second part is transported to methanation plants in Italy and Greece (2.8~kt/h), where the synthetic methane produced is fed into the gas grid in the vicinity of major gas import hubs.
+In the \carbonscenario, there are three main transport purposes of the low-cost carbon captured in Central and Eastern Europe (see Fig.~\ref{fig:balance_map}, top left).
+A large part of the carbon is transported to the Iberian Peninsula (8~kt/h), the British Isles (7.5~kt/h) and Denmark (2.8~kt/h), where it is used along with low-cost electrolytic hydrogen to produce Fischer-Tropsch fuels and methanol.
+A second part is transported to methanation plants in Italy and Greece (2.8~kt/h), where the synthetic methane is fed into the gas grid near large, geographically defined terminals where fossil gas is imported from outside Europe.
 Finally, a third part of captured carbon is transported directly to sites in the North Sea, Baltic Sea and Mediterranean Sea where it is sequestered.
 Almost none of the captured carbon is used locally at the point-source.
-In addition to electrolysis in regions with the good renewable resources to supply CU ($\sim$125~GW$_\text{el}$), the model places smaller amounts of electrolyzers where hydrogen is needed for industrial processes across Central Europe ($\sim$8.5~GW$_\text{el}$) (see bottom left).
-Hydrogen prices vary strongly across the continent, with low price ``valleys'' where the CU locates with around 60~€/MWh, and high price ``peaks'' in Central Europe with around 95~€/MWh.
+In addition to electrolysis in regions with good renewable resources to supply CU ($\sim$125~GW$_\text{el}$), the model places smaller amounts of electrolyzers where hydrogen is needed for industrial processes across Central Europe ($\sim$8.5~GW$_\text{el}$) (see bottom left).
+Hydrogen prices vary widely across the continent, with low price ``valleys'' at around 60~€/MWh, where the CU is located, and high price ``peaks'' in Central Europe at around 95~€/MWh.
 Despite the transport system, carbon prices still vary from 110~€/t in Central Europe to 160~€/t at sequestration sites.
@@
 In the \hydrogenscenario{}, the main hydrogen transportation routes go from Spain (50~GW) and the United Kingdom (18~GW) to Central Europe to supply local carbonaceous fuel production with captured carbon from point sources (see Fig~\ref{fig:balance_map}, bottom right).
-Also, the \hydrogen{} network supplies spatially-fixed \hydrogen{} demand for industry in Central Europe (2.5~GW in total) at much lower costs than local electrolysis or SMR.
-Carbon sequestration primarily locates in Portugal where large DAC facilities operate (14~kt/h, see Fig~\ref{fig:balance_map}, top right).
-\carbon{} prices are low in Central Europe (110-120~€/t) and highest at sequestration regions (160~€/t).
+Also, the hydrogen network supplies spatially-fixed hydrogen demand for industry in Central Europe (2.5~GW in total) at much lower costs than local electrolysis or SMR.
+Carbon sequestration is primarily located in Portugal where large DAC facilities operate (14~kt/h, see Fig~\ref{fig:balance_map}, top right).
+Carbon prices are low in Central Europe (110-120~€/t) and highest in sequestration regions (160~€/t).
 Hydrogen prices have flattened out at a relatively low level of €60~/MWh in all regions.
 
-In both cases, CU is placed where material inputs, \carbon{} and \hydrogen{}, are provided at minimal cost, i.e., in price valleys of hydrogen and carbon in the \carbongrid{} and \hydrogenscenario{} respectively.
+In both scenarios, CU is placed where material inputs, \carbon{} and \hydrogen{}, are provided at minimal cost, i.e., in price valleys of hydrogen and carbon in the \carbongrid{} and \hydrogenscenario{} respectively.
 In the \carbonscenario{}, CU co-locates with low-cost hydrogen production and uses transported carbon from the inland.
 In the \hydrogenscenario{}, CU co-locates with carbon point sources and uses transported low-cost hydrogen.
-Both scenarios effectively combine low-cost renewable hydrogen production and carbon management for combustion and spatially-fixed industrial processes.
+Thus, both scenarios effectively combine low-cost renewable hydrogen production and carbon management for combustion and spatially-fixed industrial processes.
 
-However, both transport systems provide individual cost-benefits to the system.
-While the carbon grid makes it possible to transport captured carbon for sequestration at low cost, the dependence on DAC near the coast causes higher costs in the hydrogen scenario{}.
+However, there are individual cost-benefits.
+On the one hand, the carbon network enables transport of low-cost captured carbon to sequestration sites, reducing the dependence on high-cost DAC near the coast which are present in the \hydrogenscenario{}.
 On the other hand, the \hydrogengrid{} is able to supply the spatially-fixed hydrogen demands in Central Europe with low-cost hydrogen imports, while the model in the \carbonscenario{} is forced to deploy electrolysis in regions with poor renewable resources as well as additional gas infrastructure in Central Europe.
-The latter aspect dominates the first, leading to lower system cost in the \hydrogenscenario{}.
+The latter aspect finally leads to lower system costs in the \hydrogenscenario{} than in the \carbonscenario{}.
 
-\subsection*{Hybrid configuration provides further flexibilities}\label{subsec:Hybrid}
+\subsection*{Hybrid configuration combines advantages}\label{subsec:Hybrid}
 
@@
     \end{subfigure}
-    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors in the \hybridscenario{} assuming net \carbon{} neutrality.
+    \caption{Average operation, flows and prices of the carbon (left) and hydrogen (right) sectors in the \hybridscenario{} assuming net \carbon{} neutrality.
 For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
-Carbon Sequestration offshore are drawn as full circles.
-Lines and arrows show the interregional transportation.
-\carbon{} from point-source in the inland either supplies local CU with imported \hydrogen{} or facilitates sequestration in nearby offshore regions.
+Carbon Sequestration offshore is drawn in full circles.
+Lines and arrows show interregional transportation.
+Carbon from point-source in the inland either supplies local CU with imported hydrogen or facilitates sequestration in nearby offshore regions.
     % Carbon network looks the same as in~\cite{morbeeOptimisedDeploymentEuropean2012}: two backbones, one in the nothern Europe other in south east.
@@
 
-The \hybridscenario{}, the model combines the advantages of hydrogen and carbon networks, resulting in the highest system cost reduction of 5.2\% (41~bn€/a) compared to the \baselinescenario{}.
-Again, the model transports low-cost hydrogen to the point sources in Central Europe to supply CU, but also transports low-cost carbon from point sources close the shore to sequestration sites (see Fig.~\ref{fig:balance_map_full}).
-This reduces reliance on high-cost DAC and provides low-cost hydrogen across the regions.
+The \hybridscenario{} combines the advantages of hydrogen and carbon networks, resulting in the highest system cost reduction of 5.2\% (41~bn€/a) compared to the \baselinescenario{}.
+Again, the model transports low-cost hydrogen to the point sources in Central Europe to supply CU, but also transports low-cost carbon from point sources close to the shore to sequestration sites (see Fig.~\ref{fig:balance_map_full}).
+This provides low-cost hydrogen across the regions and reduces reliance on high-cost DAC.
 
-For the \carbon{} network, the model employs short and direct pipeline routes to transport carbon from inland point sources to nearby sequestration sites, as for example in Northern Germany, Italy and Greece (see Fig.~\ref{fig:balance_map_full}, left).
-Additionally, the \carbon{} network balances out small local variations in carbon demand and supply to produce CU, as for example in the middle of France.
+For the carbon network, the model employs short and direct pipeline routes to transport carbon from inland point sources to nearby sequestration sites, as for example in Northern Germany, Italy and Greece (see Fig.~\ref{fig:balance_map_full}, left).
+Additionally, the carbon network balances out small local variations in carbon demand and supply to produce CU, for example in the centre of France.
 The hydrogen routes are similar to those in the \hydrogenscenario{} with transports from Western regions (Iberian Peninsula and British Isles) to distributed CU production sites across the continent (see Fig.~\ref{fig:balance_map_full}, right).
-While the topology of the \hydrogen{} network (8~bn€ investments) roughly corresponds to that of the \hydrogenscenario{} primarily driven by CU, the \carbon{} network plays more of a complementary role primarily driven by CS (3~bn€ investments, see Fig.~\ref{fig:cost_bar_transmission}).
+While the topology of the hydrogen network (8~bn€/a investments) roughly corresponds to that of the \hydrogenscenario{} and is primarily driven by CU, the carbon network plays more of a complementary role and is primarily driven by CS (3~bn€/a investments, compare Fig.~\ref{fig:cost_bar_transmission}).
 
@@
     \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/captureshare_line.png}
-    \caption{Proportion of plants with integrated carbon capture for all models in the net zero scenario.
+    \caption{Proportion of plants with integrated carbon capture for scenarios.
 The size of the dots corresponds to the average capacity factor of the respective technology.
-While peak load technologies such as gas-fired combined heat and power (CHP) plants, which are only in operation on a few days a year, are not expanded for carbon capture, baseload point sources such as biomass CHP and process emissions from industry are fully (partially) developed as soon as the transport of \carbon{} (\hydrogen{}) is permitted.}
+While peak load technologies such as gas-fired combined heat and power (CHP) plants, which are only in operation a few weeks a year, are not expanded for carbon capture, point sources with base load characteristics such as biomass CHP and process emissions from industry are fully (partially) developed as soon as the transport of carbon (hydrogen) is permitted.}
     % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
@@
 
-In this combination, the networks enable all technologies that run on a high average, above 70\%, capacity factor to capture carbon (see Fig.~\ref{fig:captureshare_line}).
-Except for SMR and gas-fired Combined Heat and Power Plants (CHP) which operate a few weeks annually, all emissions from carbon-emitting technologies are captured.
-Thus, only a few Direct Air Capture (DAC) installations are installed, such as those in southern Spain for methanolization plants.
-This leads to reduced power consumption and, therefore, less required wind and solar power deployment.
+In this combination, the networks enable all technologies that run on a high average capacity factor, i.e., above 70\%, to capture carbon to the full extent (see Fig.~\ref{fig:captureshare_line}).
+Except for SMR and gas-fired CHP which operate a few weeks annually, all emissions from carbon-emitting technologies are captured.
+Thus, only a few DAC facilities are installed, such as those in southern Spain for methanolization plants.
+In comparison to the other scenarios, this leads to reduced power consumption and, therefore, less required wind and solar power deployment.
 
+\subsection*{Stricter emission targets increase \carbon{} network benefits}
 \begin{figure}[htb!]
@@
     \caption[short]{Net change in system cost when tightening the \carbon{} emission target from net-zero to net-negative 10\% of 1990s emissions.
-Each cost bar is split into contribution of the same technology groups, except for \carbon{} sequestration cost, which are now displayed separately.
-For all models, Direct Air Capture (DAC) contributes most to the additional \carbon{} removal, requiring further solar, wind and heat pump capacities for electricity and heat input.}
+Each cost bar is split into contributions of the same technology groups as used in Fig.~\ref{fig:cost_bar}, except for carbon sequestration costs, which are now displayed separately.
+For all models, Direct Air Capture (DAC) contributes the most to additional \carbon{} removal, requiring further solar, wind and heat pump capacities for electricity and heat input.}
     \label{fig:net-negative_cost_bar}
@@
 
-However, the exploitation of carbon point sources is close to its feasible limit.
-When applying a stricter \carbon{} emissions target and force the model to remove 460~Mt \carbon{} over one year (equivalent to 10\% of \carbon{} emissions in 1990) while sequestering maximally 660~Mt, investment in DAC and associated solar and wind power and heat supply cover a large share of additional system cost in all scenarios (see Fig.~\ref{fig:net-negative_cost_bar}).
-In the \hybridscenario{}, this share is at 80\% while new carbon capture at point sources only make up 6.5\% of new system costs.
-The additional sequestration is mainly supplied by large DAC facilities near the shore.
-To a small extend it is supplied by additional carbon captured and transported from bio-energetic sources in the inland that were used to supply CU before (compare Fig.~\ref{fig:balance_maps_full_nn}).
-At the same time, some of the CU in Central Europe moves to Spain, where it uses carbon from new DAC facilities without the installation of a \carbon{} network.
-This slight shift in CU deployment leads to less investments in the hydrogen transport route from the Iberian Peninsula to Central Europe (2~bn€/a) and to an additional investment in the \carbongrid{} from Central and Eastern Europe to the shores (1.2~bn€/a, compare Fig.~\ref{fig:cost_bar_transmission_nn}).
+When applying a stricter \carbon{} emissions target and forcing the model to remove 460~Mt \carbon{} over one year (equivalent to 10\% of \carbon{} emissions in 1990) while sequestering maximally 660~Mt, investment in DAC and associated solar and wind power and heat supply covers a large share of additional system costs in all scenarios (see Fig.~\ref{fig:net-negative_cost_bar}).
+In the \hybridscenario{}, this share is at 80\% while new carbon capture at point sources only makes up 6.5\% of new system costs.
+This shows that the exploitation of carbon point sources is close to its feasible limit.
+The additional sequestration is primarily supplied by large DAC facilities near the shore.
+To a much smaller extent, carbon captured from bio-energetic sources in the inland, that was used to supply CU, before is now transported and sequestered (compare Fig.~\ref{fig:balance_maps_full_nn}).
+Some of the CU in Central Europe moves to Spain, where it uses carbon from new DAC facilities.
 
-Overall, the net cost increase by 77~bn€ in the \carbonscenario{}, by 81~bn€ in the \hydrogenscenario{}, and by 78~bn€ in the \hybridscenario{}, resulting in a cost-benefit of 6~bn€ for the \hybridscenario{} in comparison to the other scenarios.
+In the network layouts, the slight shift in CU deployment results in partial dismantling of hydrogen pipelines, especially from the Iberian Peninsula to Central Europe (0.5~bn€/a).
+This adjustment, however, leaves 80\% of the hydrogen topology intact.
+Simultaneously, the \carbon{} network expands routes from Central and Eastern Europe to the shores (1.2~bn€/a).
+The benefits of \carbon{} transport and sequestration with tighter emission targets are also evident when comparing the other scenarios.
+The carbon network in the \carbonscenario{} expands to transport more carbon from point sources to sequestration sites rather than to CU facilities.
+In the \hydrogenscenario{}, where all carbon for CU and CS must come from DAC, the hydrogen transport topology remain mainly untouched (for detailed cost-contributions, refer to sections \ref{sec:cost_comparison} and \ref{sec:operation_nn} in the Appendix).
+
+Overall, this leads to the net cost increases of 77~bn€/a in the \carbonscenario{}, by 81~bn€/a in the \hydrogenscenario{}, and by 78~bn€/a in the \hybridscenario{}.
+The resulting cost-benefit of the \hybridscenario{} in comparison to the other scenarios is 6~bn€/a.
 
@@
 
-Our study assesses the roles of hydrogen (\hydrogen{}) and carbon dioxide (\carbon{}) networks in Europe's future energy system under net-neutrality and net-negativity targets for \carbon{} emissions.
-Both networks have strong impact of the optimal deployment of carbon capture, carbon utilization and sequestration.
-A \carbon{} network effectively reduces costs by transporting low-cost carbon from distributed point sources to sequestration sites and centralized carbon utilization at sites with low-cost hydrogen production.
+Our study assesses the roles of hydrogen (\hydrogen{}) and carbon dioxide (\carbon{}) networks in Europe's future energy system under net carbon neutrality and net-removal targets.
+We show that both networks strongly impact the optimal deployment of carbon capture, carbon utilization and sequestration.
+A \carbon{} network effectively reduces costs by transporting low-cost carbon from distributed point sources to sequestration sites and sites with low-cost hydrogen production to produce carbonaceous fuels.
 A hydrogen network provides greater cost savings as low-cost hydrogen can be transported to regions with high demand, allowing carbon utilization at point-sources.
-A combination of both networks emphasize the strengths of each infrastructure.
+A combination of both networks emphasizes the strengths of each infrastructure.
 In the hybrid configuration, the \hydrogen{} network supplies low-cost hydrogen to spatially-fixed hydrogen demand and carbon utilization at point-sources across regions.
-At the same time the \carbon{} network transports low-cost carbon captured from point-sources close to the coast to sequestration sites, reducing the overall reliance on high-cost Direct Air Capture.
+At the same time, the \carbon{} network transports low-cost carbon captured from point-sources close to the coast to sequestration sites, reducing the overall reliance on high-cost Direct Air Capture.
 This results not only in the lowest system cost, but also in a more robust energy system.
 
-We discuss limitations of the model in section~\ref{sec:limitations} and sensitivities to \carbon{} pipeline costs in section~\ref{sec:subsidy}.
-Further research should explore the impacts of synthetic fuel imports and biomass potentials.
+Limitations of the study, such as the neglect of synthetic fuel imports and assumptions on biomass potentials, are discussed in the Appendix in Section ~\ref{sec:limitations}.
+A sensitivity analysis of the costs of \carbon{} pipelines, which show negligible effects on the results, can be found in the Appendix in Section ~\ref{sec:subsidy}.
 
-In summary, our study highlights the need for strategic planning and integration of both \carbon{} and \hydrogen{} networks in the European energy landscape.
-The flexibility and cost-effectiveness provided by these networks are essential to achieve EU's climate targets.
-By harnessing the synergies between \hydrogen{} and \carbon{} transportation and management, Europe can effectively reduce system costs while ensuring a robust and sustainable energy system.
+In summary, our study highlights the benefits for strategic planning and integration of both \carbon{} and \hydrogen{} networks in the European energy landscape.
+The flexibility and cost-efficiency offered by these networks can make a significant contribution to achieving the EU's climate targets, in particular by reducing dependence on technologies that are still in the early market phase, such as Direct Air Capture.
+Leveraging the synergies between \hydrogen{} and \carbon{} transportation and management, Europe can effectively reduce system costs while ensuring a robust and sustainable energy system.
 
@@
 The study is conducted based on the open-source, capacity expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
+% TODO: cite pypsa and linopy
 The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, agriculture, and transport sectors, including international aviation and shipping.
@@
 Endogenous model results include the expansion of renewable energy sources, storage technologies, power-to-X conversion and transmission capacities, heating technologies, peaking power plants, and the deployment of gray, blue or green hydrogen, among others.
-The model considers various energy carriers like electricity, heat, hydrogen, methane, methanol, liquid hydrocarbons and biomass, together with the corresponding conversions technologies.
+The model considers various energy carriers like electricity, heat, hydrogen, methane, methanol, liquid hydrocarbons and biomass, together with the corresponding conversion technologies.
 
@@
 %
-In our configuration, the model's time horizon spans one year with a 3-hourly temporal resolution and a spatial resolution of 90 regions.
+In our configuration, the model's time horizon spans one year, with a 3-hourly temporal resolution and a spatial resolution of 90 regions.
 Each of the regions consists of a complex subsystem with technologies for supplying, converting, storing and transporting energy.
 Exogenous assumptions on energy demand and non-abatable emissions are taken from various sources~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023} (see Fig.~\ref{fig:total-demand-bar}).
-The energy demand for electricity, transport, biomass, heat and gas are defined per region and time-step.
+The energy demand for electricity, transport, biomass, heat and gas is defined per region and time-step.
 Land transport demand is exogenously assumed to be fully electrified, including heavy-duty vehicles.
@@
 The system exogenously produces 633~Mt \carbon{} per year from industry, aviation, shipping and agriculture, 153~Mt of which are fossil-based process emissions.
-Location for industrial clusters are taken from~\cite{hotmaps_industrial_db}. Energy demand for industries are calculated from~\cite{mantzosJRCIDEES20152018}.
+Locations for industrial clusters are taken from~\cite{hotmaps_industrial_db}. Energy demand for industries is calculated from~\cite{mantzosJRCIDEES20152018}.
 
@@
 Low-carbon electricity is provided by wind, solar, biomass, hydro-electricity and nuclear power plants.
-Hydro-electric and nuclear plants cannot be extended beyond currently installed capacities.
+Hydro-electric and nuclear plants cannot be extended beyond their currently installed capacities.
 Weather-dependent power potentials for solar, wind and hydro-electricity are calculated from the reanalysis and satellite data sets, ERA5 and SARAH-2,~\cite{hersbachERA5GlobalReanalysis2020,pfeifrothSurfaceRadiationData2017} per region and time-stamp, using the open-source tool Atlite~\cite{hofmannAtliteLightweightPython2021}.
@@
 We limit regional biomass use to the medium-level potentials from the JRC-ENSPRESO database~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
-Inter-regional biomass transport is permitted with transport costs of $\sim$0.1~€/MWh/km considered.
+Inter-regional biomass transport is permitted, with transport costs of $\sim$0.1~€/MWh/km considered.
 
@@
 Liquid fuels like oil, methanol and Fischer-Tropsch (FT) fuels are not spatially resolved, since transport costs per unit of energy are negligible due to their high energy density.
-Throughout this study, we focus on \carbon{} and \hydrogen{} networks because of their high relevance for infrastructure investments and thus public policy decisions.
+Throughout this study, we focus on \carbon{} and \hydrogen{} networks because of their high relevance to infrastructure investments and thus public policy decisions.
 The electricity network and gas network are both included in the model with full geographical detail, but not further analyzed.
@@
 Gas networks also already exist, and will most likely experience decreased use in the future, removing any bottlenecks or constraints on the optimal system buildout or operation.
-If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limit, considering costs for pipeline segments and compressors.
+If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limits, considering costs for pipeline segments and compressors.
 Pipeline flows are modelled using net transfer capacities and without flow dynamics, pressure valves, or energy demand for compression.
@@
 To supply carbon needed for CS and CU, the system can choose to deploy carbon capture (CC) technologies at various point sources of fossil or biogenic origin (see below), or through DAC facilities.
-The concept of a merit order for captured carbon plays a pivotal role in optimizing the deployment of carbon capture technologies based on their economic feasibility.
+The concept of a merit order for capturing carbon plays a pivotal role in optimizing the deployment of carbon capture technologies based on their economic feasibility.
 This merit order ranks all CC technologies according to their relative costs to capture an additional marginal ton.
@@
 At the lower end of the cost spectrum, CC technologies applied to process emissions, such as those from cement, offer a cost-effective starting point.
-Following this, biomass combined heat and power (CHP) systems provide a dual benefit of energy production and carbon capture.
-Moving up the scale, gas used in industrial applications and biomass employed for industrial processes represent more costly yet viable options to capture carbon.
-DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand.
+Following this, biomass combined heat and power (CHP) systems provide the dual benefit of energy production and carbon capture.
+Moving up the scale, gas used in industrial applications and biomass employed in industrial processes represent more costly yet viable options to capture carbon.
+DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a high cost level due to its current high capital costs and energy demand.
 Finally, based on our model results, biogas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per marginal ton of captured \carbon{}.
-Biogas input is a high-cost fuel which based on the endogenous modeling decisions is not used in large quantities.
+Biogas input is a high-cost fuel which based on endogenous modeling decisions is not used in large quantities.
 To the extent that biogas-to-gas facilities are used by the model to supply additional (carbon-neutral) gas as fuel, adding CC infrastructure incurs low costs (and thus ranges on the left-hand side of the merit order curve).
@@
     \centering
-    \includegraphics[width=.5\linewidth]{baseline/figures/90_nodes/sequestration_map.png}
+    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/sequestration_map.png}
     \caption{Maximal sequestration potential per offshore region used as input for all models.
@@
 Our estimates for carbon sequestration potentials are restrictive, limiting the annual sequestration potential to 25~Mt per region and constraining it further such that this amount could be  25 years.
-Furthermore, we decide to limit the total amount of sequestered \carbon{} to 200~Mt per year for our net-zero \carbon{} emission target and 660~Mt per year for our net-negative emission target.
-This is enough to offset hard-to-abate fossil process emissions (e.g., calcination of limestone in cement manufacturing) and some limited slack of fossil fuels to be used where they can be most valuable.
+Furthermore, we decided to limit the total amount of sequestered \carbon{} to 200~Mt per year for our net-zero \carbon{} emission target and 660~Mt per year for our net-negative emission target.
+This is enough to offset hard-to-abate fossil process emissions (e.g., calcination of limestone in cement manufacturing) and some limited slack for fossil fuels to be used where they can be most valuable.
 This is in order to avoid offsetting fossil emissions with carbon removal if they can technologically be avoided in the first place.
@@
 \renewcommand{\thesection}{\Alph{section}}
-\renewcommand{\thefigure}{\Alph{section}.\arabic{figure}} % Add this line to apply the new counter to figures
+\renewcommand{\thefigure}{\Alph{section}.\arabic{figure}}
 
@@
 
-\section*{Appendix}
+\section{Appendix}
 
@@
 
-\clearpage
-\subsection{Cost comparison across models}
-
-The following figures show further comparisons of total system costs and cost contributions of single technologies across the models and scenarios.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics*[width=0.4\linewidth]{comparison/emission-reduction/figures/90_nodes/objective_heatmap.png}
-    \caption{Total annual system cost for the different network models and \carbon{} targets.
-While in the \baselinescenario{}, the model has neither a \carbon{} nor an \hydrogen{} network, it can expand both in the \hybridscenario{}.}
-    \label{fig:objective_heatmap}
-\end{figure}
-
-\begin{figure}[ht]
-    \centering
-    \begin{subfigure}{.5\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/cost_bar_transmission.png}
-        \caption{}
-        \label{fig:cost_bar_transmission}
-    \end{subfigure}%
-    \begin{subfigure}{.5\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{comparison/net-negative-0.1/figures/90_nodes/cost_bar_transmission.png}
-        \caption{}
-        \label{fig:cost_bar_transmission_nn}
-    \end{subfigure}
-    \caption{Annual transmission system costs for different models and scenarios for net \carbon{} neutrality (a) and net \carbon{} removal scenario (b).}
-\end{figure}
-
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=\linewidth]{difference/h2-only-full/figures/90_nodes/cost_map.png}
-    \caption{Difference in regional costs between the \hydrogengrid{} and \hybridscenario{} assuming net \carbon{} neutrality.
-The left figure shows higher spendings per technology and region and transport system in the \hydrogenscenario{}, the right shows higher spendings in the \hybridscenario{}.}
-    \label{fig:cost_map_difference}
-\end{figure}
-
-\begin{figure}[ht!]
-    \centering
-    \includegraphics[width=\linewidth]{difference/net-negative-0.1-h2-only-full/figures/90_nodes/cost_map.png}
-    \caption{Difference in regional costs between the \hydrogengrid{} and \hybridscenario{} in the net \carbon{} removal scenario.
-The left figure shows higher spendings per technology and region and transport system for the \hydrogenscenario{}, the right shows higher spendings in the \hybridscenario{}.}
-    \label{fig:cost_map_difference_nn}
-\end{figure}
-
 \subsection{Technologies with Carbon Capture}
 
+In the following, we give a more detailed description of technologies which are upgraded to integrating carbon capture.
 In the \baselinescenario{}, the capture share is highest for biogas-to-gas facilities, followed by process emissions, biomass for industry, biomass CHPs, gas for industry, SMR and gas CHPs.
@@
 
-\clearpage
 \subsection{Subsidizing \carbon{} pipelines}
@@
 The effect of a 50\% subsidy on \carbon{} pipelines impacts the optimal technology deployment to small extent.
-Fig.~\ref{fig:cost_bar_subsidy} shows the corresponding total system cost of all models in the Net-Zero scenario.
+Fig.~\ref{fig:cost_bar_subsidy} shows the corresponding total system cost of all scenarios with a net carbon neutrality target.
 Fig.~\ref{fig:cost_bar_diff_subsidy} shows the net change in system cost between non-subsidized and subsidized models, showing only the \carbongrid{} and \hybridscenario{} where changes occur.
@@
 \clearpage
-\subsection{Operation in the \hybridscenario{}}
+\subsection{Cost comparison across scenarios}
+\label{sec:cost_comparison}
+
+The following figures show further comparisons of total system costs and cost contributions of single technologies across the models and scenarios.
 
@@
     \centering
-    \includegraphics[width=.8\linewidth]{full/figures/90_nodes/sankey_diagramm.png}
-    \caption{Sankey diagram of the carbon flows in the \hybridscenario{} assuming net \carbon{} neutrality.}
-    \label{fig:sankey_diagramm}
+    \includegraphics*[width=0.4\linewidth]{comparison/emission-reduction/figures/90_nodes/objective_heatmap.png}
+    \caption{Total annual system cost for the different network models and \carbon{} targets, net-zero (NZ) left, net-negative (NN) right.}
+    \label{fig:objective_heatmap}
 \end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/cost_bar_transmission.png}
+        \caption{}
+        \label{fig:cost_bar_transmission}
+    \end{subfigure}%
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{comparison/net-negative-0.1/figures/90_nodes/cost_bar_transmission.png}
+        \caption{}
+        \label{fig:cost_bar_transmission_nn}
+    \end{subfigure}
+    \caption{Annual transmission system costs for different models and scenarios for net \carbon{} neutrality (a) and net \carbon{} negative (NN) scenario (b).}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=\linewidth]{difference/emission-reduction-0.1-full/figures/90_nodes/cost_map.png}
+    \caption{Difference in regional costs between when tightening the emission target in the \hybridscenario{} from \carbon{} net-neutrality to net-removal.
+The left figure shows higher spendings per technology and region and transport system net-neutrality case, the right shows higher spendings net-negative (NN) case.}
+    \label{fig:cost_map_difference_full_nn}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=\linewidth]{difference/emission-reduction-0.1-co2-only/figures/90_nodes/cost_map.png}
+    \caption{Difference in regional costs between when tightening the emission target in the \carbonscenario{} from \carbon{} net-neutrality to net-removal.
+The left figure shows higher spendings per technology and region and transport system net-neutrality case, the right shows higher spendings net-negative (NN) case.}
+    \label{fig:cost_map_difference_co2_nn}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=\linewidth]{difference/emission-reduction-0.1-h2-only/figures/90_nodes/cost_map.png}
+    \caption{Difference in regional costs between when tightening the emission target in the \hydrogenscenario{} from \carbon{} net-neutrality to net-removal.
+The left figure shows higher spendings per technology and region and transport system net-neutrality case, the right shows higher spendings net-negative (NN) case.}
+    \label{fig:cost_map_difference_h2_nn}
+\end{figure}
+
+% \begin{figure}[ht!]
+%     \centering
+%     \includegraphics[width=\linewidth]{difference/emission-reduction-0.1-co2-only/figures/90_nodes/cost_map.png}
+%     \caption{Difference in regional costs between the \hydrogengrid{} and \hybridscenario{} assuming net \carbon{} neutrality.
+The left figure shows higher spendings per technology and region and transport system in the \hydrogenscenario{}, the right shows higher spendings in the \hybridscenario{}.}
+%     \label{fig:cost_map_difference}
+% \end{figure}
+
+% \begin{figure}[ht!]
+%     \centering
+%     \includegraphics[width=\linewidth]{difference/net-negative-0.1-h2-only-full/figures/90_nodes/cost_map.png}
+%     \caption{Difference in regional costs between the \hydrogengrid{} and \hybridscenario{} in the net \carbon{} negative (NN) scenario.
+The left figure shows higher spendings per technology and region and transport system for the \hydrogenscenario{}, the right shows higher spendings in the \hybridscenario{}.}
+%     \label{fig:cost_map_difference_nn}
+% \end{figure}
 
@@
 \subsection{Operation in Net \carbon{} Removal Scenarios}
+\label{sec:operation_nn}
 The following figures display the optimal operation of the hydrogen and carbon sector for all models in the net \carbon{} removal scenario.
+
+\begin{figure}[ht!]
+    \centering
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/balance_map_carbon.png}
+        \label{fig:balance_map_carbon_full_nn}
+    \end{subfigure}%
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/balance_map_hydrogen.png}
+        \label{fig:balance_map_hydrogen_full_nn}
+    \end{subfigure}
+    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hydrogenscenario{} in the net \carbon{} removal scenario.
+For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
+Carbon Sequestration offshore are drawn as full circles.
+Lines and arrows show the interregional transportation.
+With the tightened emission target, the model decreases the \hydrogengrid{} layout in comparison to the net carbon neutrality scenario and increases \carbon{} transport from inland point sources to the sequestration sites.
+    }
+    \label{fig:balance_maps_full_nn}
+\end{figure}
 
@@
 Lines and arrows show the interregional transportation.
-The \carbongrid{} reveals a fundamentally different layout than in the Net-Zero scenario.
+The \carbongrid{} reveals a fundamentally different layout than with net carbon neutrality target.
     }
@@
 Lines and arrows show the interregional transportation.
-With the tightened emission target, the \hydrogengrid{} expands the \hydrogengrid{} layout of the Net-Zero scenario model results while increasing DAC facilities at the coast.
+With the tightened emission target, the \hydrogengrid{} expands the \hydrogengrid{} layout from the net carbon neutrality target results while increasing DAC facilities at the coast.
     }
@@
 
-\begin{figure}[ht!]
+\clearpage
+\subsection{Inter-sectoral \carbon{} flows}
+
+The following figure show Sankey Diagrams of the carbon flows between the sectors.
+
+\begin{figure}[h!]
     \centering
-    \begin{subfigure}{.5\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/balance_map_carbon.png}
-        \label{fig:balance_map_carbon_full_nn}
-    \end{subfigure}%
-    \begin{subfigure}{.5\textwidth}
-        \centering
-        \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/balance_map_hydrogen.png}
-        \label{fig:balance_map_hydrogen_full_nn}
-    \end{subfigure}
-    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hydrogenscenario{} in the net \carbon{} removal scenario.
-For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
-Carbon Sequestration offshore are drawn as full circles.
-Lines and arrows show the interregional transportation.
-With the tightened emission target, the model decreases the \hydrogengrid{} layout in comparison to the Net-Zero scenario and increases \carbon{} transport from inland point sources to the sequestration sites.
-    }
-    \label{fig:balance_maps_full_nn}
+    \includegraphics[width=.8\linewidth]{full/figures/90_nodes/sankey_diagramm.png}
+    \caption{Sankey diagram of the carbon flows in the \hybridscenario{} assuming net \carbon{} neutrality.}
+    \label{fig:sankey_diagramm}
 \end{figure}
@@
     \includegraphics[width=.8\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
-    \caption{Sankey diagram of the carbon flows in the \hybridscenario{} in the Net-Negative scenario.}
+    \caption{Sankey diagram of the carbon flows in the \hybridscenario{} with a net \carbon{} removal target.}
     \label{fig:sankey_diagramm_nn}

### 2023-12-21 workflow/scripts/plot_demand_bar.py

@@ -32,7 +32,7 @@ rename_bus_carrier = {
     "low voltage": "electricity",
     "heat": "low-t heat",
     "Li ion": "electricity",
-    "oil": "carbonaceous\nfuel",
+    "oil": "liquid\nhydrocarbons",
 }
 rename_carrier = {
     "Heat": "Residential and Services Heat",

### 2023-12-21 report/submissions/journal/paper/paper.tex
adjust intro on policy and industry

@@
 
-Recently, policymakers in Europe have started developing hydrogen and carbon management strategies~\cite{GermanyDevelopingStrategy2023,CarbonManagementStrategie}, planning infrastructure components~\cite{CONetz}, and committing to the first carbon utilization projects~\cite{EFuelsPilotPlant2022,OrstedAssumesFull,GROUNDBREAKINGEFUELPRODUCTION,DLREfuelsDLR}.
-With the European Union's goal of achieving climate neutrality by 2050, a wide range of programs, funding models and initiatives have been established with dedicated instruments supporting both the ramp-up of the hydrogen and carbon economy~\cite{eu2023netzero,europeangreendeal,europeaninnovationfund}.
-% TODO: split into hydrogen and carbon?
-% What about NZIA "proposes a CO2 injection target of 50 million tonnes per year by 2030, which will be stored within the European Union (EU)"
-% What about RED III targets, 10 MtH2/a domestic in 2030? Name a few targets rather than hiding everything in references - readers won't go through all references..
+% Recently, policymakers in Europe have started developing hydrogen and carbon management strategies,
+% ~\cite{GermanyDevelopingStrategy2023,CarbonManagementStrategie}, planning infrastructure components~\cite{CONetz}, and committing to the first carbon utilization projects~\cite{EFuelsPilotPlant2022,OrstedAssumesFull,GROUNDBREAKINGEFUELPRODUCTION,DLREfuelsDLR}.
+In line the climate neutrality target by 2050 of the European Union (EU) under the European Green Deal~\cite{europeangreendeal}, numerous programs, financial models, and initiatives have been launched to support the ramp-up of the hydrogen and carbon economy.
+The Renewable Energy Directive III~\cite{DirectiveEU20232023} aims to produce 10~Mt hydrogen from renewable energy sources domestically in the EU.
+The Net-Zero Industry Act~\cite{eu2023netzero} proposes a \carbon{} injection target of 50~Mt per year by 2030, which will be sequestered within the European Union.
+The European Innovation Fund~\cite{europeaninnovationfund} supports carbon management technologies to enhance the global competitiveness of European industries.
 
-Collaborative industry initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen Infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier.
-Some natural gas pipelines have already been repurposed to transport hydrogen~\cite{RohrFreiFuer}.
+At the same time, collaborative industry initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen Infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier.
+% Some natural gas pipelines have already been repurposed to transport hydrogen~\cite{RohrFreiFuer}.
 Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synthetic fuels on the one hand and offer to purchase \carbon{} on the other hand.
@@
 
-Limitations of the study, such as the neglect of synthetic fuel imports and assumptions on biomass potentials, are discussed in the Appendix in Section ~\ref{sec:limitations}.
-A sensitivity analysis of the costs of \carbon{} pipelines, which show negligible effects on the results, can be found in the Appendix in Section ~\ref{sec:subsidy}.
+Limitations of the study, such as the neglect of synthetic fuel imports and assumptions on biomass potentials, are discussed in the Appendix in Section~\ref{sec:limitations}.
+A sensitivity analysis of the costs of \carbon{} pipelines, which show negligible effects on the results, can be found in the Appendix in Section~\ref{sec:subsidy}.
 
@@
 The study is conducted based on the open-source, capacity expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
-% TODO: cite pypsa and linopy
+The model uses the open-source software tools PyPSA~\cite{brownPyPSAPythonPower2018} and Linopy~\cite{hofmannLinopyLinearOptimization2023} for modelling and cost-optimization.
 The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, agriculture, and transport sectors, including international aviation and shipping.

### 2023-12-21 report/submissions/journal/paper/paper.tex
add pypsa and linopy refs and BE acknowledgement

@@
 The study is conducted based on the open-source, capacity expansion model PyPSA-Eur~\cite{horschPyPSAEurOpenOptimisation2018,brownSynergiesSectorCoupling2018,PyPSAEurSecSectorCoupledOpen2023}.
-The model uses the open-source software tools PyPSA~\cite{brownPyPSAPythonPower2018} and Linopy~\cite{hofmannLinopyLinearOptimization2023} for modelling and cost-optimization.
+The model uses the open-source software tools PyPSA~\cite{brownPyPSAPythonPower2018} and Linopy~\cite{hofmannLinopyLinearOptimization2023}
+for modelling and cost-optimization.
 The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, agriculture, and transport sectors, including international aviation and shipping.
@@
 Many of the assumptions contained therein are based on the technology catalogues published by the Danish Energy Agency~\cite{danishenergyagencyTechnologyDataGeneration2019,thedanishenergyagencyTechnologyDataCarbon2023}.
-Endogenous model results include the expansion of renewable energy sources, storage technologies, power-to-X conversion and transmission capacities, heating technologies, peaking power plants, and the deployment of gray, blue or green hydrogen, among others.
+Endogenous model results include the expansion of renewable energy sources, storage technologies, Power-To-X conversion and transmission capacities, heating technologies, peaking power plants, and the deployment of gray, blue or green hydrogen, among others.
 The model considers various energy carriers like electricity, heat, hydrogen, methane, methanol, liquid hydrocarbons and biomass, together with the corresponding conversion technologies.
@@
 This is in order to avoid offsetting fossil emissions with carbon removal if they can technologically be avoided in the first place.
+
+\subsection*{Acknowledgements}
+
+Fabian Hofmann and Christoph Tries are funded by the BreakthroughEnergy Initiative~\cite{BreakthroughEnergy2023}.
 

### 2023-12-29 report/submissions/journal/paper/paper.tex
integrate final internal review suggestions

@@
 \usepackage{tabularx, multirow}
+\usepackage{hyperref} % Add this line to enable hyperlinks
 \usepackage{fancyhdr,lipsum}
@@
 \bibliography{../../../references.bib}
+
+\hypersetup{
+colorlinks=true,
+    linkcolor=black, % Change the link color to blue
+    urlcolor=blue, % Change the URL color to blue
+    citecolor=black, % Change the citation color to blue
+    filecolor=black % Change the file color to blue
+}
 
@@
 The market potential for carbon capture and sequestration in Europe is underscored by the Capture Map's estimated potential~\cite{ToolsGreenTransition} of 1.7 Gt carbon annually from point sources, accounting for about half of the continent's emissions, as well as estimated sequestration capacity potentials of up to 3~Gt/a~\cite{europeancommissionEuropeanCO2Storage}.
-However, it's important to note that carbon sequestration technology, unlike carbon capture, is not yet fully mature and its full potential and implications remain to be explored in depth.
+However, it is important to note that carbon sequestration technology, unlike carbon capture, is not yet fully mature and its full potential and implications remain to be explored in depth.
 To advance the carbon economy, the Clean Air Task Force emphasizes the importance of developing a carbon transport system in Europe alongside a hydrogen network~\cite{lockwoodEuropeanStrategyCarbon}.
@@
 
-In the network layouts, the slight shift in CU deployment results in partial dismantling of hydrogen pipelines, especially from the Iberian Peninsula to Central Europe (0.5~bn€/a).
+In the network layouts, the slight shift in CU deployment results in a partial dismantling of hydrogen pipelines, especially from the Iberian Peninsula to Central Europe (0.5~bn€/a).
 This adjustment, however, leaves 80\% of the hydrogen topology intact.
@@
 The carbon network in the \carbonscenario{} expands to transport more carbon from point sources to sequestration sites rather than to CU facilities.
-In the \hydrogenscenario{}, where all carbon for CU and CS must come from DAC, the hydrogen transport topology remain mainly untouched (for detailed cost-contributions, refer to sections \ref{sec:cost_comparison} and \ref{sec:operation_nn} in the Appendix).
+In the \hydrogenscenario{}, where all carbon for CU and CS must come from DAC, the hydrogen transport topology remain mainly untouched (for detailed cost contributions, refer to sections \ref{sec:cost_comparison} and \ref{sec:operation_nn} in the Appendix).
 
@@
 
-Fabian Hofmann and Christoph Tries are funded by the BreakthroughEnergy Initiative~\cite{BreakthroughEnergy2023}.
+Fabian Hofmann and Christoph Tries are funded by the Breakthrough Energy Initiative (\href{https://breakthroughenergy.org/}{https://breakthroughenergy.org/}).
 
@@
 
-\clearpage
-\subsection{Inter-sectoral \carbon{} flows}
+% \clearpage
+% \subsection{Inter-sectoral \carbon{} flows}
 
-The following figure show Sankey Diagrams of the carbon flows between the sectors.
+% The following figure show Sankey Diagrams of the carbon flows between the sectors.
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=.8\linewidth]{full/figures/90_nodes/sankey_diagramm.png}
-    \caption{Sankey diagram of the carbon flows in the \hybridscenario{} assuming net \carbon{} neutrality.}
-    \label{fig:sankey_diagramm}
-\end{figure}
+% \begin{figure}[h!]
+%     \centering
+%     \includegraphics[width=.8\linewidth]{full/figures/90_nodes/sankey_diagramm.png}
+%     \caption{Sankey diagram of the carbon flows in the \hybridscenario{} assuming net \carbon{} neutrality.}
+%     \label{fig:sankey_diagramm}
+% \end{figure}
 
-\begin{figure}
-    \centering
-    \includegraphics[width=.8\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
-    \caption{Sankey diagram of the carbon flows in the \hybridscenario{} with a net \carbon{} removal target.}
-    \label{fig:sankey_diagramm_nn}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=.8\linewidth]{net-negative-0.1/full/figures/90_nodes/sankey_diagramm.png}
+%     \caption{Sankey diagram of the carbon flows in the \hybridscenario{} with a net \carbon{} removal target.}
+%     \label{fig:sankey_diagramm_nn}
+% \end{figure}
 

### 2023-12-29 report/submissions/journal/paper/paper.tex
add reproducibility link

@@
 
-\subsection*{Acknowledgements}
+\subsubsection*{Reproducibility}
 
-Fabian Hofmann and Christoph Tries are funded by the Breakthrough Energy Initiative (\href{https://breakthroughenergy.org/}{https://breakthroughenergy.org/}).
+The workflow to reproduce all figures and numbers is available at \href{https://github.com/FabianHofmann/carbon-networks}{github.com/FabianHofmann/carbon-networks}.
+
+\subsubsection*{Acknowledgements}
+
+Fabian Hofmann and Christoph Tries are funded by the Breakthrough Energy Initiative (\href{https://breakthroughenergy.org/}{breakthroughenergy.org}).
 

### 2023-12-30 report/submissions/journal/cover_letter/letter.tex
add cover letter

@@
 %
-\def\name{Philipp Glaum \& Fabian Hofmann
+\def\name{Fabian Hofmann, Christoph Tries, Fabian Neumann, Elisabeth Zeyen, Tom Brown.
 }
@@
 % Set your e-mail here
-\def\Email{\textbf{\color{tubred}E-mail}: p.glaum@tu-berlin.de}
+\def\Email{\textbf{\color{tubred}E-mail}: m.hofmann@tu-berlin.de}
 %
@@
 
+\newcommand{\carbon}{CO$_2$}
+\newcommand{\hydrogen}{H$_2$}
+
 %%-----------------------------------------------------------------------------------------%%
@@
 		%%%%%%%%%%% 	Text   %%%%%%%%%%%%%%
-		Thank you very much for the opportunity to submit our manuscript to Applied Energy. Hereby we submit the paper “Leveraging the Existing German Transmission Grid with Dynamic Line Rating”.
+		Thank you for the opportunity to submit our manuscript to Nature Energy. Hereby we submit the paper “\hydrogen{} and \carbon{} Network Strategies for the European Energy System”.
 
-		In our work, we present the first large-scale investment model that evaluates the benefits of Dynamic Line Rating (DLR) for renewable energy integration.
-		Taking Germany as an example, the model includes a high temporal and spatial resolution as well as a high level of detail to enable a realistic representation of the system dynamics.
-		The paper appears at an important time as  politicians and Transmission System Operators (TSOs) have recognized the problem of lacking transmission capacities when moving towards high renewable shares in Europe and all over the world.
-		% In many official reports DLR is considered as a promising measure to relief the transmission grid.
+		In our study, we introduce the first large-scale investment model that assesses the co-benefits of a carbon-dioxide and a hydrogen transport system in Europe.
+The model optimizes the production, transport, and storage of various carriers, including electricity, hydrogen, methane, and carbon-dioxide, across all relevant sectors.
 
-		Our analysis not only confirms and quantitatively describes transmission system benefits but also shows how the implementation of DLR affects the optimal system design for different shares of renewable power production.
-By better integrating wind power resources, especially offshore, DLR leads to significant cost savings of 5\% of the system cost, mainly due to the reduced need for cost-driving energy storage.
-This key finding can be generalized to all countries with normal to high wind resources and is therefore of particular importance for international network operators and the research community.
-Furthermore, we provide an open-source workflow which can be used to reproduce the results and extend the model to other countries.
+		The paper appears at an important time, with policy makers and industry leaders concretizing plans and investments for hydrogen and carbon-dioxide management systems.
+In particular, it remains unclear how and to what extent synthetic fuel production and permanent carbon storage (carbon sequestration) should be integrated into the future energy system.
 
-		The paper was proofread by Prof. Tom Brown, an energy system expert and native tongue speaker, as well as from Rena Kuwahata from Ampacimon, an expert on transmission lines and DLR applications.
+		Our analysis not only confirms the crucial flexibility that \hydrogen{} and \carbon{} transport systems bring to \carbon{} emission neutral energy systems, but also illustrates how their combination further enhances energy efficiency and social welfare.
+We show that the hydrogen grid optimally transports low-cost hydrogen from areas with good renewable resources to regionally-distributed carbon sources such as industrial clusters and biomass combustion plants to enable synthetic fuel production at site.
+A carbon-dioxide network, on the other hand, is most cost-beneficial to transport low-cost carbon from point sources near the coast to sequestration sites to permanently store it.
+In such a setup, the system unlocks most carbon capture potentials from biomass and reduces the reliance on cost-intensive direct air capture.
+Moreover, the benefits of deploying both networks increase when enforcing a net carbon-removal target, leading to system-cost reductions of 6 billion Euros per year.
 
-		We want to thank the editor for taking the time to consider the publication of our paper in Applied Energy.
+		Our analysis underlines the importance of incorporating hydrogen and carbon dioxide networks into European energy policy for a resilient, carbon-neutral or negative future energy system.
 
-		% What is the novelty of this work? Is the paper appealing to a popular or scientific audience?
-		% Why the authors think the paper is important and why the journal should publish it? Has the article
-		% been checked by a native tongue speaker with expertise in the field? Are you available as a reviewer
-		% for at least three other articles for Applied Energy during the current year? In addition to answering
-		% those five questions, the authors should also describe in one or two paragraphs the significance of
-		% their work and what new information is described in the manuscript.
+		We did not have any prior discussions with a Nature Energy editor about this manuscript. The manuscript has not been published and is not under consideration for publication elsewhere.
+
+		We believe that our findings are in line with the scope of Nature Energy and will be of interest to your readership.
+We want to thank the editor for taking the time to consider the publication of our paper in Nature Energy.
 

### 2023-12-30 report/submissions/journal/cover_letter/letter.tex
further work on cover letter

@@
 %
-\def\name{Fabian Hofmann, Christoph Tries, Fabian Neumann, Elisabeth Zeyen, Tom Brown.
+\def\name{Dr. Fabian Hofmann,\\Christoph Tries,\\Dr. Fabian Neumann,\\Elisabeth Zeyen,\\Prof. Dr. Tom Brown.
 }
@@
 		%%%%%%%%%%% 	Text   %%%%%%%%%%%%%%
-		Thank you for the opportunity to submit our manuscript to Nature Energy. Hereby we submit the paper “\hydrogen{} and \carbon{} Network Strategies for the European Energy System”.
+		Thank you for the opportunity to submit our manuscript to Nature Energy.
+We are pleased to present our paper titled “\hydrogen{} and \carbon{} Network Strategies for the European Energy System”.
+This collaborative effort from researchers at the Institute of Energy Technology, Technical University of Berlin, contributes to the critical discourse on climate-neutral energy systems in Europe.
 
-		In our study, we introduce the first large-scale investment model that assesses the co-benefits of a carbon-dioxide and a hydrogen transport system in Europe.
-The model optimizes the production, transport, and storage of various carriers, including electricity, hydrogen, methane, and carbon-dioxide, across all relevant sectors.
+		In our study, we introduce the first large-scale investment model to analyze potential synergies and competition between hydrogen and carbon dioxide transport networks within a carbon-neutral European energy system.
+Our model represents an advancement over existing models by optimizing a comprehensive range of technologies for the production, transportation, and storage of various energy carriers across all relevant sectors at high temporal and spatial resolution.
 
-		The paper appears at an important time, with policy makers and industry leaders concretizing plans and investments for hydrogen and carbon-dioxide management systems.
-In particular, it remains unclear how and to what extent synthetic fuel production and permanent carbon storage (carbon sequestration) should be integrated into the future energy system.
+		The publication of this paper is timely, considering that policymakers and industry leaders are currently formulating plans and investments for hydrogen and carbon dioxide management systems.
+Our study offers critical insights and practical implications for the integration of synthetic fuel production and permanent carbon storage (carbon sequestration) into future energy systems.
 
-		Our analysis not only confirms the crucial flexibility that \hydrogen{} and \carbon{} transport systems bring to \carbon{} emission neutral energy systems, but also illustrates how their combination further enhances energy efficiency and social welfare.
-We show that the hydrogen grid optimally transports low-cost hydrogen from areas with good renewable resources to regionally-distributed carbon sources such as industrial clusters and biomass combustion plants to enable synthetic fuel production at site.
-A carbon-dioxide network, on the other hand, is most cost-beneficial to transport low-cost carbon from point sources near the coast to sequestration sites to permanently store it.
-In such a setup, the system unlocks most carbon capture potentials from biomass and reduces the reliance on cost-intensive direct air capture.
-Moreover, the benefits of deploying both networks increase when enforcing a net carbon-removal target, leading to system-cost reductions of 6 billion Euros per year.
+		Our analysis not only confirms the role of hydrogen and carbon dioxide transport systems in providing crucial flexibilities to carbon-neutral energy systems.
+It also illustrates how their combination enhances energy efficiency and social welfare.
+We demonstrate that a hydrogen grid is optimally used to transport low-cost hydrogen from regions with abundant renewable resources to distributed carbon sources, such as industrial clusters and biomass combustion plants, to enable on-site synthetic fuel production.
+Conversely, a carbon dioxide network efficiently transports low-cost carbon from point sources near the coast to sequestration sites for permanent storage.
+Such an approach increase carbon capture rates from biomass combustion and reduces reliance on cost-intensive direct air capture.
+Furthermore, the benefits of deploying both networks are amplified under a net carbon-removal target.
+		As underscored by our analysis, the integration of hydrogen and carbon dioxide networks into European energy policy is vital for a resilient, carbon-neutral, or negative future energy system.
 
-		Our analysis underlines the importance of incorporating hydrogen and carbon dioxide networks into European energy policy for a resilient, carbon-neutral or negative future energy system.
+		This manuscript has not been published nor is it under consideration elsewhere. It has not been previously discussed with any Nature Energy editor.
 
-		We did not have any prior discussions with a Nature Energy editor about this manuscript. The manuscript has not been published and is not under consideration for publication elsewhere.
-
-		We believe that our findings are in line with the scope of Nature Energy and will be of interest to your readership.
-We want to thank the editor for taking the time to consider the publication of our paper in Nature Energy.
+		We believe that our findings align with the scope of Nature Energy and will interest your readership. We appreciate the editor's time in considering our manuscript for publication.
 

### 2023-12-30 report/submissions/journal/cover_letter/letter.tex
review cover letter

@@
 		Thank you for the opportunity to submit our manuscript to Nature Energy.
-We are pleased to present our paper titled “\hydrogen{} and \carbon{} Network Strategies for the European Energy System”.
-This collaborative effort from researchers at the Institute of Energy Technology, Technical University of Berlin, contributes to the critical discourse on climate-neutral energy systems in Europe.
+We are pleased to present our paper titled “\hydrogen{} and \carbon{} Network Strategies for the European Energy System.” This collaborative effort from researchers at the Institute of Energy Technology, Technical University of Berlin, contributes significantly to the critical discourse on climate-neutral energy systems in Europe.
 
@@
 
-		Our analysis not only confirms the role of hydrogen and carbon dioxide transport systems in providing crucial flexibilities to carbon-neutral energy systems.
-It also illustrates how their combination enhances energy efficiency and social welfare.
+		Our analysis not only confirms the role of hydrogen and carbon dioxide transport systems in providing crucial flexibilities to carbon-neutral energy systems, but it also illustrates how their combination enhances energy efficiency and social welfare.
 We demonstrate that a hydrogen grid is optimally used to transport low-cost hydrogen from regions with abundant renewable resources to distributed carbon sources, such as industrial clusters and biomass combustion plants, to enable on-site synthetic fuel production.
 Conversely, a carbon dioxide network efficiently transports low-cost carbon from point sources near the coast to sequestration sites for permanent storage.
-Such an approach increase carbon capture rates from biomass combustion and reduces reliance on cost-intensive direct air capture.
+Such an approach increases carbon capture rates from biomass combustion and reduces reliance on cost-intensive direct air capture.
 Furthermore, the benefits of deploying both networks are amplified under a net carbon-removal target.
+
 		As underscored by our analysis, the integration of hydrogen and carbon dioxide networks into European energy policy is vital for a resilient, carbon-neutral, or negative future energy system.

### 2024-01-01 report/submissions/journal/cover_letter/letter.tex
cover letter: correct webiste
incorp Tom review of the cover letter

@@
 % Set your department or personal website here
-\def\URL{\textbf{\color{tubred}URL}: {https://www.ensys.tu-berlin.de}}
+\def\URL{\textbf{\color{tubred}URL}: {https://www.tu.berlin/ensys}}
 %
@@
 		%%%%%%%%%%% 	Text   %%%%%%%%%%%%%%
-		Thank you for the opportunity to submit our manuscript to Nature Energy.
 		We are pleased to present our paper titled “\hydrogen{} and \carbon{} Network Strategies for the European Energy System.” This collaborative effort from researchers at the Institute of Energy Technology, Technical University of Berlin, contributes significantly to the critical discourse on climate-neutral energy systems in Europe.
@@
 		In our study, we introduce the first large-scale investment model to analyze potential synergies and competition between hydrogen and carbon dioxide transport networks within a carbon-neutral European energy system.
-Our model represents an advancement over existing models by optimizing a comprehensive range of technologies for the production, transportation, and storage of various energy carriers across all relevant sectors at high temporal and spatial resolution.
+Our model represents an advancement over existing models by optimizing a comprehensive range of technologies for the production, transportation, and storage of of hydrogen and carbon dioxide, along with other energy carriers, across all relevant sectors at high temporal and spatial resolution.
 

### 2024-01-01 report/submissions/journal/cover_letter/letter.tex
follow up

@@
 		%%%%%%%%%%% 	Text   %%%%%%%%%%%%%%
-		We are pleased to present our paper titled “\hydrogen{} and \carbon{} Network Strategies for the European Energy System.” This collaborative effort from researchers at the Institute of Energy Technology, Technical University of Berlin, contributes significantly to the critical discourse on climate-neutral energy systems in Europe.
+		We are pleased to present our paper titled “\hydrogen{} and \carbon{} Network Strategies for the European Energy System.” This collaborative effort from researchers at the Institute of Energy Technology, Technical University of Berlin, contributes to the critical discourse on climate-neutral energy systems in Europe.
 
 		In our study, we introduce the first large-scale investment model to analyze potential synergies and competition between hydrogen and carbon dioxide transport networks within a carbon-neutral European energy system.
-Our model represents an advancement over existing models by optimizing a comprehensive range of technologies for the production, transportation, and storage of of hydrogen and carbon dioxide, along with other energy carriers, across all relevant sectors at high temporal and spatial resolution.
+Our model represents an advancement over existing models by optimizing a comprehensive range of technologies for the production, transportation, and storage of hydrogen and carbon dioxide, along with other energy carriers, across all relevant sectors at high temporal and spatial resolution.
 
@@
 
-		As underscored by our analysis, the integration of hydrogen and carbon dioxide networks into European energy policy is vital for a resilient, carbon-neutral, or negative future energy system.
+		As underscored by our analysis, the integration of both hydrogen and carbon dioxide networks into European energy policy is vital for a resilient, carbon-neutral, or negative future energy system.
 

### 2024-01-23 report/submissions/journal/paper/paper.tex
paper: add data and code availability statement

@@
 Hydrogen is being considered in several industries that are not eligible for electrification.
-It offers an efficient way to transport and store energy over long distances and periods, and can be produced from electrolysis or steam methane reforming (SMR), for example.
+It offers an efficient way to transport and store energy over long distances and periods and can be produced from electrolysis or steam methane reforming (SMR), for example.
 Captured carbon can be stored in geological formations to extract it from the atmosphere permanently, a process known as carbon sequestration (CS).
@@
 % ~\cite{GermanyDevelopingStrategy2023,CarbonManagementStrategie}, planning infrastructure components~\cite{CONetz}, and committing to the first carbon utilization projects~\cite{EFuelsPilotPlant2022,OrstedAssumesFull,GROUNDBREAKINGEFUELPRODUCTION,DLREfuelsDLR}.
-In line the climate neutrality target by 2050 of the European Union (EU) under the European Green Deal~\cite{europeangreendeal}, numerous programs, financial models, and initiatives have been launched to support the ramp-up of the hydrogen and carbon economy.
+In line with the climate neutrality target by 2050 of the European Union (EU) under the European Green Deal~\cite{europeangreendeal}, numerous programs, financial models, and initiatives have been launched to support the ramp-up of the hydrogen and carbon economy.
 The Renewable Energy Directive III~\cite{DirectiveEU20232023} aims to produce 10~Mt hydrogen from renewable energy sources domestically in the EU.
@@
 % TODO: This drives all CCU....
-% TODO: The demands should be mentioned properly introduced
+% TODO: The demands should be mentioned properly and introduced
 
@@
 We analyze how the deployment of carbon and hydrogen networks affects different decarbonization strategies in the scenarios and how investments in none or only one of the transport technologies lead to inefficient market layouts.
-We do this by focusing first on a net-zero \carbon{} emissions target, while limiting carbon sequestration to 200~Mt/a.
+We do this by focusing first on a net-zero \carbon{} emissions target while limiting carbon sequestration to 200~Mt/a.
 We give a detailed description of the different technological and geographical impacts of the transportation systems.
@@
     \caption{Balance of captured carbon for all scenarios assuming net \carbon{} neutrality.
-Positive values indicate carbon capture, negative values indicate carbon consumption.
+Positive values indicate carbon capture and negative values indicate carbon consumption.
 By integrating hydrogen and carbon networks, the predominant method for carbon removal shifts from Direct Air Capture (DAC) to bio-energetic processes with capture.
@@
 Where carbon from point sources cannot be sequestered, it has to be compensated elsewhere by negative emissions from DAC or biomass capture with sequestration.
-All of these options are sub-optimal from a cost-perspective.
+All of these options are sub-optimal from a cost perspective.
 
@@
 Finally, a third part of captured carbon is transported directly to sites in the North Sea, Baltic Sea and Mediterranean Sea where it is sequestered.
-Almost none of the captured carbon is used locally at the point-source.
+Almost none of the captured carbon is used locally at the point source.
 In addition to electrolysis in regions with good renewable resources to supply CU ($\sim$125~GW$_\text{el}$), the model places smaller amounts of electrolyzers where hydrogen is needed for industrial processes across Central Europe ($\sim$8.5~GW$_\text{el}$) (see bottom left).
@@
 In the \hydrogenscenario{}, the main hydrogen transportation routes go from Spain (50~GW) and the United Kingdom (18~GW) to Central Europe to supply local carbonaceous fuel production with captured carbon from point sources (see Fig~\ref{fig:balance_map}, bottom right).
-Also, the hydrogen network supplies spatially-fixed hydrogen demand for industry in Central Europe (2.5~GW in total) at much lower costs than local electrolysis or SMR.
+Also, the hydrogen network supplies spatially fixed hydrogen demand for industry in Central Europe (2.5~GW in total) at much lower costs than local electrolysis or SMR.
 Carbon sequestration is primarily located in Portugal where large DAC facilities operate (14~kt/h, see Fig~\ref{fig:balance_map}, top right).
@@
 However, there are individual cost-benefits.
-On the one hand, the carbon network enables transport of low-cost captured carbon to sequestration sites, reducing the dependence on high-cost DAC near the coast which are present in the \hydrogenscenario{}.
+On the one hand, the carbon network enables the transport of low-cost captured carbon to sequestration sites, reducing the dependence on high-cost DAC near the coast which is present in the \hydrogenscenario{}.
 On the other hand, the \hydrogengrid{} is able to supply the spatially-fixed hydrogen demands in Central Europe with low-cost hydrogen imports, while the model in the \carbonscenario{} is forced to deploy electrolysis in regions with poor renewable resources as well as additional gas infrastructure in Central Europe.
@@
 Carbon from point-source in the inland either supplies local CU with imported hydrogen or facilitates sequestration in nearby offshore regions.
-    % Carbon network looks the same as in~\cite{morbeeOptimisedDeploymentEuropean2012}: two backbones, one in the nothern Europe other in south east.
+    % Carbon network looks the same as in~\cite{morbeeOptimisedDeploymentEuropean2012}: two backbones, one in northern Europe other in south east.
     }
@@
 
-For the carbon network, the model employs short and direct pipeline routes to transport carbon from inland point sources to nearby sequestration sites, as for example in Northern Germany, Italy and Greece (see Fig.~\ref{fig:balance_map_full}, left).
-Additionally, the carbon network balances out small local variations in carbon demand and supply to produce CU, for example in the centre of France.
-The hydrogen routes are similar to those in the \hydrogenscenario{} with transports from Western regions (Iberian Peninsula and British Isles) to distributed CU production sites across the continent (see Fig.~\ref{fig:balance_map_full}, right).
+For the carbon network, the model employs short and direct pipeline routes to transport carbon from inland point sources to nearby sequestration sites, for example in Northern Germany, Italy and Greece (see Fig.~\ref{fig:balance_map_full}, left).
+Additionally, the carbon network balances out small local variations in carbon demand and supply to produce CU, for example in the center of France.
+The hydrogen routes are similar to those in the \hydrogenscenario{} with transports from Western regions (the Iberian Peninsula and the British Isles) to distributed CU production sites across the continent (see Fig.~\ref{fig:balance_map_full}, right).
 While the topology of the hydrogen network (8~bn€/a investments) roughly corresponds to that of the \hydrogenscenario{} and is primarily driven by CU, the carbon network plays more of a complementary role and is primarily driven by CS (3~bn€/a investments, compare Fig.~\ref{fig:cost_bar_transmission}).
@@
 While peak load technologies such as gas-fired combined heat and power (CHP) plants, which are only in operation a few weeks a year, are not expanded for carbon capture, point sources with base load characteristics such as biomass CHP and process emissions from industry are fully (partially) developed as soon as the transport of carbon (hydrogen) is permitted.}
-    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
+    % A CO2 network unlocks BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
     \label{fig:captureshare_line}
@@
 The carbon network in the \carbonscenario{} expands to transport more carbon from point sources to sequestration sites rather than to CU facilities.
-In the \hydrogenscenario{}, where all carbon for CU and CS must come from DAC, the hydrogen transport topology remain mainly untouched (for detailed cost contributions, refer to sections \ref{sec:cost_comparison} and \ref{sec:operation_nn} in the Appendix).
+In the \hydrogenscenario{}, where all carbon for CU and CS must come from DAC, the hydrogen transport topology remains mainly untouched (for detailed cost contributions, refer to sections \ref{sec:cost_comparison} and \ref{sec:operation_nn} in the Appendix).
 
@@
 A \carbon{} network effectively reduces costs by transporting low-cost carbon from distributed point sources to sequestration sites and sites with low-cost hydrogen production to produce carbonaceous fuels.
-A hydrogen network provides greater cost savings as low-cost hydrogen can be transported to regions with high demand, allowing carbon utilization at point-sources.
+A hydrogen network provides greater cost savings as low-cost hydrogen can be transported to regions with high demand, allowing carbon utilization at point sources.
 A combination of both networks emphasizes the strengths of each infrastructure.
-In the hybrid configuration, the \hydrogen{} network supplies low-cost hydrogen to spatially-fixed hydrogen demand and carbon utilization at point-sources across regions.
-At the same time, the \carbon{} network transports low-cost carbon captured from point-sources close to the coast to sequestration sites, reducing the overall reliance on high-cost Direct Air Capture.
-This results not only in the lowest system cost, but also in a more robust energy system.
+In the hybrid configuration, the \hydrogen{} network supplies low-cost hydrogen to spatially fixed hydrogen demand and carbon utilization at point sources across regions.
+At the same time, the \carbon{} network transports low-cost carbon captured from point sources close to the coast to sequestration sites, reducing the overall reliance on high-cost Direct Air Capture.
+This results not only in the lowest system cost but also in a more robust energy system.
 
@@
 
-In summary, our study highlights the benefits for strategic planning and integration of both \carbon{} and \hydrogen{} networks in the European energy landscape.
+In summary, our study highlights the benefits of strategic planning and integration of both \carbon{} and \hydrogen{} networks in the European energy landscape.
 The flexibility and cost-efficiency offered by these networks can make a significant contribution to achieving the EU's climate targets, in particular by reducing dependence on technologies that are still in the early market phase, such as Direct Air Capture.
@@
 The model uses the open-source software tools PyPSA~\cite{brownPyPSAPythonPower2018} and Linopy~\cite{hofmannLinopyLinearOptimization2023}
-for modelling and cost-optimization.
+for modeling and cost optimization.
 The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, agriculture, and transport sectors, including international aviation and shipping.
 All technology cost assumptions are taken for the year 2040 and sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
-Many of the assumptions contained therein are based on the technology catalogues published by the Danish Energy Agency~\cite{danishenergyagencyTechnologyDataGeneration2019,thedanishenergyagencyTechnologyDataCarbon2023}.
+Many of the assumptions contained therein are based on the technology catalogs published by the Danish Energy Agency~\cite{danishenergyagencyTechnologyDataGeneration2019,thedanishenergyagencyTechnologyDataCarbon2023}.
 Endogenous model results include the expansion of renewable energy sources, storage technologies, Power-To-X conversion and transmission capacities, heating technologies, peaking power plants, and the deployment of gray, blue or green hydrogen, among others.
@@
 %
-In our configuration, the model's time horizon spans one year, with a 3-hourly temporal resolution and a spatial resolution of 90 regions.
+In our configuration, the model's time horizon spans one year, with a 3-hour temporal resolution and a spatial resolution of 90 regions.
 Each of the regions consists of a complex subsystem with technologies for supplying, converting, storing and transporting energy.
@@
 Land transport demand is exogenously assumed to be fully electrified, including heavy-duty vehicles.
-Demands for kerosene for aviation, methanol for shipping, and naphtha for industry are not spatially resolved and assumed to be constant throughout all time-steps.
+Demands for kerosene for aviation, methanol for shipping, and naphtha for industry are not spatially resolved and assumed to be constant throughout all time steps.
 Heat demand is regionally subdivided into shares of urban, rural and industrial sites.
@@
 Low-carbon electricity is provided by wind, solar, biomass, hydro-electricity and nuclear power plants.
-Hydro-electric and nuclear plants cannot be extended beyond their currently installed capacities.
+Hydroelectric and nuclear plants cannot be extended beyond their currently installed capacities.
 Weather-dependent power potentials for solar, wind and hydro-electricity are calculated from the reanalysis and satellite data sets, ERA5 and SARAH-2,~\cite{hersbachERA5GlobalReanalysis2020,pfeifrothSurfaceRadiationData2017} per region and time-stamp, using the open-source tool Atlite~\cite{hofmannAtliteLightweightPython2021}.
@@
 We restrict the total volume of power transmission expansion to 20\% of its current capacity, acknowledging the challenges in inaugurating new transmission projects.
-For the use of biomass we consider only residual biomass products and no energy crops.
+For the use of biomass, we consider only residual biomass products and no energy crops.
 We limit regional biomass use to the medium-level potentials from the JRC-ENSPRESO database~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
@@
 The electricity network and gas network are both included in the model with full geographical detail, but not further analyzed.
-Electricity networks are already in place, and we restrict further extension due to concerns of public acceptance.
+Electricity networks are already in place, and we restrict further extension due to concerns about public acceptance.
 Gas networks also already exist, and will most likely experience decreased use in the future, removing any bottlenecks or constraints on the optimal system buildout or operation.
 If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limits, considering costs for pipeline segments and compressors.
-Pipeline flows are modelled using net transfer capacities and without flow dynamics, pressure valves, or energy demand for compression.
+Pipeline flows are modeled using net transfer capacities and without flow dynamics, pressure valves, or energy demand for compression.
 No retrofitting of gas pipelines is considered.
@@
 
-% modelling of hydrogen supply chain pathways
+% Modelling of hydrogen supply chain pathways
 
@@
 
-% carbon utilisation
+% carbon utilization
 
-We argue that it is crucial for a proper modelling to consider CC, CU and CS as separate technologies, rather than grouping them under the often-used but misleading umbrella term ``carbon capture, utilization and storage'' (CCUS), to allow for independent optimization and adequate representation of each function in the carbon management system.
+We argue that it is crucial for proper modeling to consider CC, CU and CS as separate technologies, rather than grouping them under the often-used but misleading umbrella term ``carbon capture, utilization and storage'' (CCUS), to allow for independent optimization and adequate representation of each function in the carbon management system.
 Our model features three drop-in fuel production technologies for CU: methanation, methanolization, and Fischer-Tropsch synthesis.
@@
 This merit order ranks all CC technologies according to their relative costs to capture an additional marginal ton.
-The cost value for a technology depends on the exact system configuration and may vary by scenario and even by region, since many technologies produce other (primary) outputs such as electricity and heat.
+The cost value for a technology depends on the exact system configuration and may vary by scenario and even by region since many technologies produce other (primary) outputs such as electricity and heat.
 The costs to capture an additional marginal ton of carbon may also be different from the average cost of capture.
@@
 Moving up the scale, gas used in industrial applications and biomass employed in industrial processes represent more costly yet viable options to capture carbon.
-DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a high cost level due to its current high capital costs and energy demand.
+DAC, an emerging technology capable of extracting \carbon{} directly from the atmosphere, stands at a high-cost level due to its current high capital costs and energy demand.
 Finally, based on our model results, biogas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per marginal ton of captured \carbon{}.
-Biogas input is a high-cost fuel which based on endogenous modeling decisions is not used in large quantities.
+Biogas input is a high-cost fuel that based on endogenous modeling decisions is not used in large quantities.
 To the extent that biogas-to-gas facilities are used by the model to supply additional (carbon-neutral) gas as fuel, adding CC infrastructure incurs low costs (and thus ranges on the left-hand side of the merit order curve).
@@
 For carbon sequestration, we only consider offshore sites as potential sinks (see Fig.~\ref{fig:sequestration_map}).
-We make this choice because offshore sequestration sites typically have larger capacity compared to onshore sequestration sites in saline aquifers and due to concerns over public safety for infrastructure near populated areas.
+We make this choice because offshore sequestration sites typically have a larger capacity compared to onshore sequestration sites in saline aquifers and due to concerns over public safety for infrastructure near populated areas.
 Our estimates for carbon sequestration potentials are restrictive, limiting the annual sequestration potential to 25~Mt per region and constraining it further such that this amount could be  25 years.
@@
 
-\subsubsection*{Reproducibility}
+\subsection*{Data availability}
 
-The workflow to reproduce all figures and numbers is available at \href{https://github.com/FabianHofmann/carbon-networks}{github.com/FabianHofmann/carbon-networks}.
+All input, processed and resulting data is available or can be generated via the workflow \href{https://github.com/FabianHofmann/carbon-networks}{github.com/FabianHofmann/carbon-networks} and the underlying workflow \href{https://github.com/PyPSA/pypsa-eur}{github.com/pypsa/PyPSA-Eur}.
+Data on the optimized energy system models analyzed throughout the study was made publicly available under \href{https://doi.org/10.6084/m9.figshare.24923028}{10.6084/m9.figshare.24923028}.
+
+\subsubsection*{Code availability}
+
+The code to reproduce all figures and numbers is available at \href{https://github.com/FabianHofmann/carbon-networks}{github.com/FabianHofmann/carbon-networks}.
 
@@
 
-\onecolumn % Add this line to adjust the layout to single column
+\onecolumn % Add this line to adjust the layout to a single column
 
@@
 The model's perfect foresight may lead to non-reproducible technology dispatch which neglects uncertainties from long-term predictions and necessary reserve margins.
-Demand and emissions from industrial clusters heavily drive the modelling results.
+Demand and emissions from industrial clusters heavily drive the modeling results.
 Allowing the model to relocate these, incorporate flexibility measures or alter underlying processes, may lead to less dependency on \carbon{} and \hydrogen{} networks.
 The transport of fuel (FT, gas, oil) between the regions is not limited which may overestimate their flexibility.
-However, the distributed CU systems, as enabled through the hydrogen network, align well with today's distributed locations for oil-refineries, providing kerosene for aviation and naphtha for industry.
-The assumption to run Fischer-Tropsch facilities on baseload with at least 90\% of their nominal capacity, is likely shaping the deployment of CU, underestimating potential important flexibilities.
+However, the distributed CU systems, as enabled through the hydrogen network, align well with today's distributed locations for oil refineries, providing kerosene for aviation and naphtha for industry.
+The assumption to run Fischer-Tropsch facilities on baseload with at least 90\% of their nominal capacity is likely shaping the deployment of CU, underestimating potential important flexibilities.
 The assumed cost projections on technologies are subject to uncertainties.
-As shown in \ref{sec:subsidy}, the effect on a 50\% cost reduction on \carbon{} pipelines alter total system costs to small extent (below 1\%).
-However, a different set of cost on DAC and electrolysis may-well impact in the model results.
+As shown in \ref{sec:subsidy}, the effect of a 50\% cost reduction on \carbon{} pipelines alters total system costs to a small extent (below 1\%).
+However, a different set of costs on DAC and electrolysis may well impact the model results.
 In combination with a net \carbon{} neutrality or reduction target, a relaxation of the global sequestration limit drives system costs down by introducing fossil fuels, as shown in~\cite{hofmannDesigningCO2Network2023}.
@@
 
-In the following, we give a more detailed description of technologies which are upgraded to integrating carbon capture.
+In the following, we give a more detailed description of technologies that are upgraded to integrate carbon capture.
 In the \baselinescenario{}, the capture share is highest for biogas-to-gas facilities, followed by process emissions, biomass for industry, biomass CHPs, gas for industry, SMR and gas CHPs.
@@
 SMR and gas CHPs have capacity factors around 15\%, and CC shares at nearly 0\% for the \baselinescenario{}.
-The capacity factor and CC share of SMR increases to around 20\% and 50-65\%, respectively, in the other models.
+The capacity factor and CC share of SMR increase to around 20\% and 50-65\%, respectively, in the other models.
 Gas CHPs serve as peak load electricity production technology, and with very few operating hours, thus the high investment costs into CC applications are not efficient.
-Similarly, SMR serves as ``peak-load'' hydrogen production technology in regions with poor renewable resources for electrolysis.
+Similarly, SMR serves as a ``peak-load'' hydrogen production technology in regions with poor renewable resources for electrolysis.
 
@@
 
-The effect of a 50\% subsidy on \carbon{} pipelines impacts the optimal technology deployment to small extent.
+The effect of a 50\% subsidy on \carbon{} pipelines impacts the optimal technology deployment to a small extent.
 Fig.~\ref{fig:cost_bar_subsidy} shows the corresponding total system cost of all scenarios with a net carbon neutrality target.
@@
 \end{subfigure}
-\caption{System costs for the all models in the net-zero emission scenario with 50\% subsidy of \carbon{} pipelines (a) and change in system cost for the \carbongrid{} and \hybridscenario{} when subsidizing \carbon{} pipelines by 50\% (b).}
+\caption{System costs for all models in the net-zero emission scenario with 50\% subsidy of \carbon{} pipelines (a) and change in system cost for the \carbongrid{} and \hybridscenario{} when subsidizing \carbon{} pipelines by 50\% (b).}
 \end{figure}
@@
     \caption{Difference in regional costs between when tightening the emission target in the \hybridscenario{} from \carbon{} net-neutrality to net-removal.
-The left figure shows higher spendings per technology and region and transport system net-neutrality case, the right shows higher spendings net-negative (NN) case.}
+The left figure shows higher spending per technology and region and transport system net-neutrality case, the right shows higher spending net-negative (NN) case.}
     \label{fig:cost_map_difference_full_nn}
@@
     \caption{Difference in regional costs between when tightening the emission target in the \carbonscenario{} from \carbon{} net-neutrality to net-removal.
-The left figure shows higher spendings per technology and region and transport system net-neutrality case, the right shows higher spendings net-negative (NN) case.}
+The left figure shows higher spending per technology and region and transport system net-neutrality case, the right shows higher spending net-negative (NN) case.}
     \label{fig:cost_map_difference_co2_nn}
@@
     \caption{Difference in regional costs between when tightening the emission target in the \hydrogenscenario{} from \carbon{} net-neutrality to net-removal.
-The left figure shows higher spendings per technology and region and transport system net-neutrality case, the right shows higher spendings net-negative (NN) case.}
+The left figure shows higher spending per technology and region and transport system net-neutrality case, the right shows higher spending net-negative (NN) case.}
     \label{fig:cost_map_difference_h2_nn}
@@
 %     \caption{Difference in regional costs between the \hydrogengrid{} and \hybridscenario{} assuming net \carbon{} neutrality.
-The left figure shows higher spendings per technology and region and transport system in the \hydrogenscenario{}, the right shows higher spendings in the \hybridscenario{}.}
+The left figure shows higher spending per technology and region and transport system in the \hydrogenscenario{}, and the right shows higher spending in the \hybridscenario{}.}
 %     \label{fig:cost_map_difference}
@@
 %     \caption{Difference in regional costs between the \hydrogengrid{} and \hybridscenario{} in the net \carbon{} negative (NN) scenario.
-The left figure shows higher spendings per technology and region and transport system for the \hydrogenscenario{}, the right shows higher spendings in the \hybridscenario{}.}
+The left figure shows higher spending per technology and region and transport system for the \hydrogenscenario{}, and the right shows higher spending in the \hybridscenario{}.}
 %     \label{fig:cost_map_difference_nn}
@@
 For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
-Carbon Sequestration offshore are drawn as full circles.
+Carbon Sequestration offshore is drawn in full circles.
 Lines and arrows show the interregional transportation.
@@
 For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
-Carbon Sequestration offshore are drawn as full circles.
+Carbon Sequestration offshore is drawn in full circles.
 Lines and arrows show the interregional transportation.
-The \carbongrid{} reveals a fundamentally different layout than with net carbon neutrality target.
+The \carbongrid{} reveals a fundamentally different layout than with the net carbon neutrality target.
     }
@@
 
-% The following figure show Sankey Diagrams of the carbon flows between the sectors.
+% The following figure shows Sankey Diagrams of the carbon flows between the sectors.
 

### 2024-08-12 report/submissions/journal/paper/paper.tex
paper: change subsection to subsubsection
start sequestration limit discussion

@@
 
+\subsection*{Higher sequestration limits reintroduce fossil fuels}
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\linewidth]{comparison/sequestration/figures/90_nodes/cost_bar.png}
+    \caption[short]{Total system cost as a function of the sequestration limit.}
+    \label{fig:sequestration_cost_bar}
+\end{figure}
+
 \subsection*{Stricter emission targets increase \carbon{} network benefits}
@@
 The carbon network in the \carbonscenario{} expands to transport more carbon from point sources to sequestration sites rather than to CU facilities.
-In the \hydrogenscenario{}, where all carbon for CU and CS must come from DAC, the hydrogen transport topology remains mainly untouched (for detailed cost contributions, refer to sections \ref{sec:cost_comparison} and \ref{sec:operation_nn} in the Appendix).
+In the \hydrogenscenario{}, where all carbon for CU and CS must come from DAC, the hydrogen transport topology remains mainly untouched (for detailed cost contributions, refer to sections~\ref{sec:cost_comparison} and~\ref{sec:operation_nn} in the Appendix).
 
@@
 \label{sec:conclusion}
-
 Our study assesses the roles of hydrogen (\hydrogen{}) and carbon dioxide (\carbon{}) networks in Europe's future energy system under net carbon neutrality and net-removal targets.
@@
 The topology and capacities of the electricity transmission system are taken from the ENTSO-E transparency map~\cite{wiegmansGridkitExtractEntsoE2016} and selected Ten Year Development Plan (TYNDP) projects.
-The power flow is based on the lossless linearized power flow approximation~.
-The topology of the \carbon{}, \hydrogen{} and gas network is identical to the topology of the electricity network, connecting all neighboring regions.
+The power flow is based on the linearized power flow approximation assuming a linear transport efficieny of 95.75\% per 1000~km accounting for resistive losses on the power lines.
+%TODO find reference
+The topology of the \hydrogen{}, \carbon{} and gas network is identical to the topology of the electricity network, connecting all neighboring regions.
+For hydrogen pipelines an average loss of 1.2\% per 1000~km, aligning with industry estimates that account for hydrogen's high permeability and small molecular size, which can lead to higher losses compared to other gases.
+The electrical demand for compression stations is assumed to be 1.8\% of the transported energy.
+For \carbon{} pipelines an average loss rate of 0.8\% per 1000~km is reflecting a generally low leakage rate observed in supercritical or dense phase CO2 transport~\cite{liuExperimentalStudyLeakage2023,vitaliRisksSafetyCO22021}.
+The compression energy needed is set to 250~KWh per 1000~km and tonne \carbon{} to maintain a supercritical state for efficient transport.
+For natural gas pipelines, the model assumes a 1\% loss rate per 1000~km, accounting for typical leakage and operational losses in well-maintained natural gas infrastructure as reported in~\cite{NaturalGasTransmission2021}.
+The electrical demand for compression is set at 1\% of the transported energy, which falls within the common range for natural gas pipeline operations~\cite{mcvayreneeMethaneEmissionsGas2023}.
+%
 Liquid fuels like oil, methanol and Fischer-Tropsch (FT) fuels are not spatially resolved, since transport costs per unit of energy are negligible due to their high energy density.
@@
 
-\subsection*{Data availability}
+\subsubsection*{Data availability}
 

### 2024-08-12 workflow/scripts/plot_cost_bar_compare_diff.py

@@ -67,7 +67,8 @@ fig, ax = plt.subplots(
     1, 1, figsize=snakemake.params.settings["figsize"], layout="constrained"
 )
 
-grouped = grouped[grouped.round(0).ne(0).any(axis=1)]
+if not snakemake.config["configs"]["test"]:
+    grouped = grouped[grouped.round(0).ne(0).any(axis=1)]
 grouped.T.plot(kind="bar", stacked=True, ax=ax, color=colors, legend=True, alpha=0.9)
 for container in ax.containers:
     if abs(container.datavalues).sum() > grouped.abs().sum().sum() / 20:

### 2024-08-12 workflow/scripts/plot_cost_map_diff.py

@@ -29,8 +29,13 @@ plt.rc("patch", linewidth=0.1)
 config = snakemake.config
 labels = config["labels"]
 specs = config["plotting"]["cost_map_diff"]
-cutoff_bus = specs["cutoff_bus"]
-cutoff_branch = specs["cutoff_branch"]
+
+if config["configs"]["test"]:
+    cutoff_bus = 0
+    cutoff_branch = 0
+else:
+    cutoff_bus = specs["cutoff_bus"]
+    cutoff_branch = specs["cutoff_branch"]
 
 networks = [
     import_network(path, remove_gas_store_capital_cost=True)
@@ -56,7 +61,7 @@ for n in networks:
     df_bus[key] = costs
 
     branches = get_transmission_branches(n)
-    df_branch[key] = s.capex(groupby=False).loc[branches]
+    df_branch[key] = s.capex(groupby=False).reindex(branches, fill_value=0)
 
     carriers.append(n.carriers)
 

### 2024-08-18 report/submissions/journal/paper/paper.tex
answer reviewer questions and make corresponding small text amendments

@@
 The additional sequestration is primarily supplied by large DAC facilities near the shore.
-To a much smaller extent, carbon captured from bio-energetic sources in the inland, that was used to supply CU, before is now transported and sequestered (compare Fig.~\ref{fig:balance_maps_full_nn}).
-Some of the CU in Central Europe moves to Spain, where it uses carbon from new DAC facilities.
+To a much smaller extent, carbon captured from bio-energetic sources in the inland, that under a net-zero emissions assumption supplies CU, is instead transported and sequestered (compare Fig.~\ref{fig:balance_maps_full_nn}).
+Some of the CU which in net-zero scenarios locates in Central Europe instead locates in Spain, where it uses carbon from new DAC facilities.
 
-In the network layouts, the slight shift in CU deployment results in a partial dismantling of hydrogen pipelines, especially from the Iberian Peninsula to Central Europe (0.5~bn€/a).
-This adjustment, however, leaves 80\% of the hydrogen topology intact.
-Simultaneously, the \carbon{} network expands routes from Central and Eastern Europe to the shores (1.2~bn€/a).
+In the \carbon{} and \hydrogen{} scenarios, this slight shift in CU deployment results in a lower buildout of hydrogen pipelines, especially from the Iberian Peninsula to Central Europe (0.5~bn€/a).
+Nonetheless, 80\% of the hydrogen topology which is built under net-zero emissions is also built under the assumption of net-negative emissions.
+Simultaneously, under net-negative emissions the \carbon{} network also includes routes from Central and Eastern Europe to the shores (1.2~bn€/a).
 The benefits of \carbon{} transport and sequestration with tighter emission targets are also evident when comparing the other scenarios.
-The carbon network in the \carbonscenario{} expands to transport more carbon from point sources to sequestration sites rather than to CU facilities.
-In the \hydrogenscenario{}, where all carbon for CU and CS must come from DAC, the hydrogen transport topology remains mainly untouched (for detailed cost contributions, refer to sections~\ref{sec:cost_comparison} and~\ref{sec:operation_nn} in the Appendix).
+The carbon network under the net-negative \carbonscenario{} is larger than under the net-zero \carbonscenario{} and transports more carbon from point sources to sequestration sites rather than to CU facilities.
+In the \hydrogenscenario{}, where all carbon for CU and CS must come from DAC, the hydrogen transport topology remains almost identical to the net-zero scenarios (for detailed cost contributions, refer to sections~\ref{sec:cost_comparison} and~\ref{sec:operation_nn} in the Appendix).
 
@@
 In the hybrid configuration, the \hydrogen{} network supplies low-cost hydrogen to spatially fixed hydrogen demand and carbon utilization at point sources across regions.
-At the same time, the \carbon{} network transports low-cost carbon captured from point sources close to the coast to sequestration sites, reducing the overall reliance on high-cost Direct Air Capture.
+At the same time, the \carbon{} network transports low-cost carbon captured from point sources near the coast to sequestration sites, reducing the reliance on Direct Air Capture.
 This results not only in the lowest system cost but also in a more robust energy system.
@@
 All technology cost assumptions are taken for the year 2040 and sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
+The database applies a discount rate of 7\% for all major technologies, accounting for the weighted average costs of capital (WACC) for investments.
 Many of the assumptions contained therein are based on the technology catalogs published by the Danish Energy Agency~\cite{danishenergyagencyTechnologyDataGeneration2019,thedanishenergyagencyTechnologyDataCarbon2023}.

### 2024-08-26 report/submissions/journal/paper/paper.tex
adjust text and plots until new subsection

@@
 All scenarios share key energy system characteristics but differ in system costs and technology use (Fig.~\ref{fig:cost_bar}).
-Costs in the \baselinescenario{} are the highest at 779~bn€/a.
-Scenarios with carbon and hydrogen networks are more cost-effective, reducing costs by 3.1\% in the \carbongrid{}, 4.8\% in the \hydrogengrid{}, and 5.2\% in the \hybridscenario{}.
+Costs in the \baselinescenario{} are the highest at 764~bn€/a.
+Scenarios with carbon and hydrogen networks are more cost-effective, reducing costs by 3.1\% in the \carbongrid{}, 4.7\% in the \hydrogengrid{}, and 5.3\% in the \hybridscenario{}.
 
@@
 System cost differences between the scenarios are driven by wind and solar, as well as carbon capture technologies.
-In scenarios with carbon and hydrogen networks, the model primarily reduces the reliance on high-cost DAC, from 392~Mt/a carbon removal in the \baselinescenario{} to 112~Mt/a in the \hybridscenario{} (see Fig.~\ref{fig:balance_captured_carbon}).
+In scenarios with carbon and hydrogen networks, the model primarily reduces the reliance on high-cost DAC, from 365~Mt/a carbon removal in the \baselinescenario{} to 113~Mt/a in the \hybridscenario{} (see Fig.~\ref{fig:balance_captured_carbon}).
 At the same time, carbon capture from biomass combustion, gas-based industry processes and biogas-to-gas upgrading increase, the latter lowering the need for synthetic methane production.
@@
 In the \carbonscenario, there are three main transport purposes of the low-cost carbon captured in Central and Eastern Europe (see Fig.~\ref{fig:balance_map}, top left).
-A large part of the carbon is transported to the Iberian Peninsula (8~kt/h), the British Isles (7.5~kt/h) and Denmark (2.8~kt/h), where it is used along with low-cost electrolytic hydrogen to produce Fischer-Tropsch fuels and methanol.
-A second part is transported to methanation plants in Italy and Greece (2.8~kt/h), where the synthetic methane is fed into the gas grid near large, geographically defined terminals where fossil gas is imported from outside Europe.
+A large part of the carbon is transported to the Iberian Peninsula (9.8~kt/h) and the British Isles (11.8~kt/h), where it is used along with low-cost electrolytic hydrogen to produce Fischer-Tropsch fuels and methanol.
+In Italy and Greece, a portion of the carbon (3.3~kt/h) is directed to methanation facilities where the synthetic methane is fed into the gas grid at key terminals that also receive fossil gas imports.
 Finally, a third part of captured carbon is transported directly to sites in the North Sea, Baltic Sea and Mediterranean Sea where it is sequestered.
 Almost none of the captured carbon is used locally at the point source.
-In addition to electrolysis in regions with good renewable resources to supply CU ($\sim$125~GW$_\text{el}$), the model places smaller amounts of electrolyzers where hydrogen is needed for industrial processes across Central Europe ($\sim$8.5~GW$_\text{el}$) (see bottom left).
+In addition to electrolysis in regions with good renewable resources to supply CU ($\sim$125~GW$_\text{el}$), the model places smaller amounts of electrolyzers where hydrogen is needed for industrial processes across Central Europe ($\sim$8~GW$_\text{el}$) (see bottom left).
 Hydrogen prices vary widely across the continent, with low price ``valleys'' at around 60~€/MWh, where the CU is located, and high price ``peaks'' in Central Europe at around 95~€/MWh.
-Despite the transport system, carbon prices still vary from 110~€/t in Central Europe to 160~€/t at sequestration sites.
+Despite the transport system, carbon prices still vary from 105~€/t in Central Europe to 160~€/t at sequestration sites.
 
-In the \hydrogenscenario{}, the main hydrogen transportation routes go from Spain (50~GW) and the United Kingdom (18~GW) to Central Europe to supply local carbonaceous fuel production with captured carbon from point sources (see Fig~\ref{fig:balance_map}, bottom right).
+In the \hydrogenscenario{}, the main hydrogen transportation route goes from Spain (66~GW) to Central Europe to supply local carbonaceous fuel production with captured carbon from point sources (see Fig~\ref{fig:balance_map}, bottom right).
 Also, the hydrogen network supplies spatially fixed hydrogen demand for industry in Central Europe (2.5~GW in total) at much lower costs than local electrolysis or SMR.
-Carbon sequestration is primarily located in Portugal where large DAC facilities operate (14~kt/h, see Fig~\ref{fig:balance_map}, top right).
+Carbon sequestration takes place at several coastal sites with local DAC facilities.
+The largest carbon sink is located in Portugal (3.5~kt/h, see Fig~\ref{fig:balance_map}, top right).
 Carbon prices are low in Central Europe (110-120~€/t) and highest in sequestration regions (160~€/t).
@@
 In the \hydrogenscenario{}, CU co-locates with carbon point sources and uses transported low-cost hydrogen.
-Thus, both scenarios effectively combine low-cost renewable hydrogen production and carbon management for combustion and spatially-fixed industrial processes.
+% Both scenarios effectively combine low-cost renewable hydrogen production and carbon management for combustion and spatially-fixed industrial processes.
 
-However, there are individual cost-benefits.
+We conclude, there are individual cost-benefits of the transport networks.
 On the one hand, the carbon network enables the transport of low-cost captured carbon to sequestration sites, reducing the dependence on high-cost DAC near the coast which is present in the \hydrogenscenario{}.
@@
 
-The \hybridscenario{} combines the advantages of hydrogen and carbon networks, resulting in the highest system cost reduction of 5.2\% (41~bn€/a) compared to the \baselinescenario{}.
+The \hybridscenario{} combines the advantages of hydrogen and carbon networks, resulting in the highest system cost reduction of 5.3\% (41~bn€/a) compared to the \baselinescenario{}.
 Again, the model transports low-cost hydrogen to the point sources in Central Europe to supply CU, but also transports low-cost carbon from point sources close to the shore to sequestration sites (see Fig.~\ref{fig:balance_map_full}).
@@
 For the carbon network, the model employs short and direct pipeline routes to transport carbon from inland point sources to nearby sequestration sites, for example in Northern Germany, Italy and Greece (see Fig.~\ref{fig:balance_map_full}, left).
-Additionally, the carbon network balances out small local variations in carbon demand and supply to produce CU, for example in the center of France.
+In addition, pipelines transport rather small amounts of carbon from spatially fixed point sources to CU facilities.
+A notable outlier is the route from the Netherlands via the UK and France to large methanolisation and Fischer-Tropsch plants in northern Spain.
 The hydrogen routes are similar to those in the \hydrogenscenario{} with transports from Western regions (the Iberian Peninsula and the British Isles) to distributed CU production sites across the continent (see Fig.~\ref{fig:balance_map_full}, right).
-While the topology of the hydrogen network (8~bn€/a investments) roughly corresponds to that of the \hydrogenscenario{} and is primarily driven by CU, the carbon network plays more of a complementary role and is primarily driven by CS (3~bn€/a investments, compare Fig.~\ref{fig:cost_bar_transmission}).
+While the topology of the hydrogen network (1.2~bn€/a) roughly corresponds to that of the \hydrogenscenario{} and is primarily driven by CU, the carbon network (0.6~bn€/a) plays a secondary role and is driven by CS (compare Fig.~\ref{fig:cost_bar_transmission}).
 
@@
 
-In this combination, the networks enable all technologies that run on a high average capacity factor, i.e., above 70\%, to capture carbon to the full extent (see Fig.~\ref{fig:captureshare_line}).
-Except for SMR and gas-fired CHP which operate a few weeks annually, all emissions from carbon-emitting technologies are captured.
-Thus, only a few DAC facilities are installed, such as those in southern Spain for methanolization plants.
-In comparison to the other scenarios, this leads to reduced power consumption and, therefore, less required wind and solar power deployment.
+In this setup, the networks enable all carbon-emitting technologies that operate at a high average capacity factor, i.e., above 70\%, to capture most of their carbon (see Fig.~\ref{fig:captureshare_line}).
+Except for SMR and gas-fired CHP which operate a few weeks annually, more than 90\% emissions are captured.
+Due to the better integration of point sources, only a few DAC facilities are installed, such as those used for methanolisation in southern Spain.
+This, in turn, leads to reduced power consumption and less required wind and solar power deployment than in the other scenarios.
 
 \subsection*{Higher sequestration limits reintroduce fossil fuels}
+\label{sec:higher_sequestration}
+Carbon management and hard-to-abate sectors of the optimized system are strongly driven by the model's limited sequestration potential of 200~Mt/a.
+In the Hybrid scenario, this limitations imposes a shadow-price of 225€/t to the system, meaning the system would gain 225€ from an additional tonne of sequestration potential.
+As soon as the sequestration limit is increased, the model starts to reintroduce fossil fuels as a primary energy source in the the aviation and industrial sectors.
 

### 2024-08-26 workflow/scripts/plot_cost_bar_diff.py

@@ -26,7 +26,7 @@ else:
 df = {}
 carriers = []
 for path in snakemake.input.networks:
-    n = import_network(path, remove_gas_store_capital_cost=True)
+    n = import_network(path, remove_gas_store_capital_cost=True, revert_dac=False)
     capex = n.statistics.capex()
     opex = n.statistics.opex(aggregate_time="sum")
 

### 2024-08-28 report/submissions/journal/paper/paper.tex
add policy conclusions and brief references to recent COM and JRC
CMS work

@@
 The Net-Zero Industry Act~\cite{eu2023netzero} proposes a \carbon{} injection target of 50~Mt per year by 2030, which will be sequestered within the European Union.
+The European Commission's 2024 industrial carbon management strategy~\cite{comcms} plans for capture of 450~Mt/a by 2050 in the EU, of which almost 250~Mt/a is sequestered.
 The European Innovation Fund~\cite{europeaninnovationfund} supports carbon management technologies to enhance the global competitiveness of European industries.
@@
 This limited sectoral scope cannot capture important dynamics of carbon management, since it neglects sectors like industry that will need to handle most \carbon{} in the future.
+The European Commission's Joint Research Center analysed the possible topology of a \carbon{} network in 2024 \cite{jrc2024}, but didn't include the interaction of  \carbon{} sources and sinks with the rest of the energy system.
 A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connected to a remote sequestration site in Norway.
@@
 A sensitivity analysis of the costs of \carbon{} pipelines, which show negligible effects on the results, can be found in the Appendix in Section~\ref{sec:subsidy}.
+
+For policy makers these results show the need for coordinated planning across sectors like carbon, hydrogen and synthetic fuels since they are strongly interacting, as well as across borders because of the localized nature of good renewable resources and sequestration potentials.
+Deploying multiple networks offers some robustness should unforeseen problems arise with one of them, and our results show that the system can even cope with neither a \carbon{} nor \hydrogen{} network for a cost penalty.
+Many of the technologies have not yet been deployed at scale before and will need support for financing, regulation and gaining the acceptance of local populations.
 

### 2024-08-31 report/submissions/journal/paper/paper.tex
add appendix section about higher sequestration potentials

@@
 This limited sectoral scope cannot capture important dynamics of carbon management, since it neglects sectors like industry that will need to handle most \carbon{} in the future.
-The European Commission's Joint Research Center analysed the possible topology of a \carbon{} network in 2024 \cite{jrc2024}, but didn't include the interaction of  \carbon{} sources and sinks with the rest of the energy system.
+The European Commission's Joint Research Center analysed the possible topology of a \carbon{} network in 2024~\cite{jrc2024}, but didn't include the interaction of  \carbon{} sources and sinks with the rest of the energy system.
 A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connected to a remote sequestration site in Norway.
@@
 To investigate the competing transport dynamics between \carbon{} and \hydrogen{} networks, we contrast the following four scenarios:
-%
-\begin{itemize}
-    \item[] \textit{Baseline:} Neither \carbon{} nor \hydrogen{} networks are constructed.
-    \item[] \textit{\carbongrid{}:} Only the \carbon{} network is developed.
-    \item[] \textit{\hydrogengrid{}:} Only the \hydrogen{} network is developed.
-    \item[] \textit{Hybrid:} Both \carbon{} and \hydrogen{} networks are developed.
-\end{itemize}
+\begin{table}[]
+    \resizebox{\textwidth}{!}{%
+        \begin{tabular}{@{}lcc@{}}
+            \toprule
+            \textbf{Scenario} & H$_2$ Network & CO$_2$ Network \\ \midrule
+            \textbf{Baseline} & \xmark          & \xmark           \\
+            \textbf{CO$_2$-Grid} & \xmark          & \checkmark            \\
+            \textbf{H$_2$-Grid}  & \checkmark          & \xmark           \\
+            \textbf{Hybrid}   & \checkmark           & \checkmark            \\ \bottomrule
+        \end{tabular}%
+    }
+\caption{Overview of the four different scenarios.}
+\label{tab:overview_scenarios}
+\end{table}
 
@@
 
-\subsection*{Higher sequestration limits reintroduce fossil fuels}
-\label{sec:higher_sequestration}
-Carbon management and hard-to-abate sectors of the optimized system are strongly driven by the model's limited sequestration potential of 200~Mt/a.
-In the Hybrid scenario, this limitations imposes a shadow-price of 225€/t to the system, meaning the system would gain 225€ from an additional tonne of sequestration potential.
-As soon as the sequestration limit is increased, the model starts to reintroduce fossil fuels as a primary energy source in the the aviation and industrial sectors.
-
-\begin{figure}[htb!]
-    \centering
-    \includegraphics[width=\linewidth]{comparison/sequestration/figures/90_nodes/cost_bar.png}
-    \caption[short]{Total system cost as a function of the sequestration limit.}
-    \label{fig:sequestration_cost_bar}
-\end{figure}
-
 \subsection*{Stricter emission targets increase \carbon{} network benefits}
@@
 Nonetheless, 80\% of the hydrogen topology which is built under net-zero emissions is also built under the assumption of net-negative emissions.
-Simultaneously, under net-negative emissions the \carbon{} network also includes routes from Central and Eastern Europe to the shores (1.2~bn€/a).
+Under net-negative emissions the \carbon{} network also includes routes from Central and Eastern Europe to the shores (1.2~bn€/a).
 The benefits of \carbon{} transport and sequestration with tighter emission targets are also evident when comparing the other scenarios.
@@
 Limitations of the study, such as the neglect of synthetic fuel imports and assumptions on biomass potentials, are discussed in the Appendix in Section~\ref{sec:limitations}.
-A sensitivity analysis of the costs of \carbon{} pipelines, which show negligible effects on the results, can be found in the Appendix in Section~\ref{sec:subsidy}.
+A sensitivity analysis of the costs of \carbon{} pipelines and costs of electrolyzers, which show relatively small effects on the results, can be found in the Appendix in Section~\ref{sec:subsidy}.
+As the cost optimum is affected by the sequestration potential limits, a more detailed analysis of a relaxed sequestration limit is presented in the Appendix in Section~\ref{sec:higher_sequestration}.
 
@@
 
+\subsection*{Higher sequestration limits}
+\label{sec:higher_sequestration}
+Carbon management in the optimised energy system, particularly in the hard to abate sectors, is heavily influenced by the model's sequestration potential, which is capped at 200~Mt/year.
+Lifting the sequestration limit leads to cost benefits and a shift in primary energy supply (see Fig.~\ref{fig:sequestration_cost_bar}).
+Fossil oil and gas increasingly provide critical energy services in the aviation, industrial and heating sectors.
+Their emissions are either directly capture at the point source and transported to be sequestered or offset large DAC facilities close to coastal sequestration sites.
+At the same time, the role of the \hydrogen{} network is reduced from 57~TWkm to 20~TWkm due to the less CU required.
+At 800~Mt/a this already reduces wind and solar investments by a third and total costs by 9.1\%, roughly aligning with the findings in~\cite{hofmannDesigningCO2Network2023}.
+
+As the sequestration capacity increases, the cost savings from additional sequestration begin to plateau.
+At 200Mt/a, the shadow price of the sequestration constraint shows that each additional tonne saves 225€/t, but at 800Mt/a, this saving drops to 51€/t.
+Extrapolating this behaviour, saturation of sequestration can be expected at around 1200~Mt/a.
+
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=\linewidth]{comparison/sequestration/figures/90_nodes/cost_bar.png}
+    \caption[short]{Total system cost as a function of the sequestration limit.}
+    \label{fig:sequestration_cost_bar}
+\end{figure}
+
 \clearpage

### 2024-08-31 workflow/scripts/plot_capacity_bar_transmission.py

@@ -1,0 +1,98 @@
+import os
+import seaborn as sns
+import matplotlib.pyplot as plt
+import pandas as pd
+from common import (
+    get_ordered_handles_labels,
+    get_transmission_links,
+    import_network,
+    mock_snakemake,
+    sort_rows_by_relative_diff,
+)
+
+alpha = 1
+region_alpha = 0.8
+
+rename = {"DC": "DC Line", "AC": "AC Line", "Gas Pipeline New": "Gas Pipeline"}
+
+if os.path.dirname(os.path.abspath(__file__)) == os.getcwd():
+    snakemake = mock_snakemake(
+        "plot_cost_bar_transmission",
+        ext="png",
+        clusters=90,
+        comparison="subsidy",
+    )
+
+sns.set_theme(**snakemake.params["theme"])
+
+
+df = {}
+carriers = []
+for path in snakemake.input.networks:
+    n = import_network(path)
+
+    capacity = n.links.p_nom_opt * n.links.length
+    cap = (
+        capacity.groupby(n.links.carrier.map(n.carriers.nice_name))
+        .sum()
+        .round(2)[lambda x: x > 0]
+    )  # GW*km / Mt*km
+
+    key = snakemake.params.labels[n.meta["wildcards"]["run"]]
+
+    df[key] = cap
+    carriers.append(n.carriers)
+
+data = pd.concat(df, axis=1)
+if "emission-reduction" in snakemake.wildcards.comparison and len(data.columns) == 3:
+    data = data.set_axis(["0", "230", "460"], axis=1)
+    data = data.rename_axis(columns="Net Carbon Removal [Mt/a]")
+data = data.rename(rename, axis=0).groupby(level=0).sum()
+
+carriers = pd.concat(carriers).drop_duplicates()
+colors = carriers.set_index("nice_name").color
+colors = colors.rename(rename, axis=0)
+
+# use plural for labels
+data = data.rename(lambda x: x + "s", axis=0)
+colors = colors.rename(lambda x: x + "s", axis=0)
+
+
+norm = 1e3
+data = sort_rows_by_relative_diff(data).div(norm)
+
+fig, ax = plt.subplots(
+    1, 1, figsize=snakemake.params.settings["figsize"], layout="constrained"
+)
+
+
+defaults = dict(kind="bar", stacked=True, lw=0, rot=90, alpha=0.8)
+
[…]

### 2024-08-31 workflow/scripts/plot_cost_bar.py

@@ -79,32 +79,31 @@ for container in ax.containers:
             color="grey",
         )
 
-if snakemake.wildcards.comparison == "default":
-    pad = 5
-    total = grouped.sum()
-    baseline = total["Baseline"]
-    decrease = 1 - total / baseline
-    for i, (val, y) in enumerate(zip(decrease, total)):
-        if val == 0:
-            ax.text(
-                i,
-                y + pad,
-                f"{baseline:.0f}",
-                ha="center",
-                va="bottom",
-                fontsize=7,
-                color="gray",
-            )
-        else:
-            ax.text(
-                i,
-                y + pad,
-                f"{-val:.1%}",
-                ha="center",
-                va="bottom",
-                fontsize=7,
-                color="gray",
-            )
+pad = 5
+total = grouped.sum()
+baseline = total[total.index[0]]
+decrease = 1 - total / baseline
+for i, (val, y) in enumerate(zip(decrease, total)):
+    if val == 0:
+        ax.text(
+            i,
+            y + pad,
+            f"{baseline:.0f}",
+            ha="center",
+            va="bottom",
+            fontsize=7,
+            color="gray",
+        )
+    else:
+        ax.text(
+            i,
+            y + pad,
+            f"{-val:.1%}",
+            ha="center",
+            va="bottom",
+            fontsize=7,
+            color="gray",
+        )
 
 
 ax.axhline(0, color="k", lw=1)

### 2024-09-06 report/submissions/journal/paper/paper.tex
update scenarios, continue with review comments

--- 2024-08-28 report/submissions/journal/paper/paper.tex
@@
 We analyze how the deployment of carbon and hydrogen networks affects different decarbonization strategies in the scenarios and how investments in none or only one of the transport technologies lead to inefficient market layouts.
-We do this by focusing first on a net-zero \carbon{} emissions target while limiting carbon sequestration to 200~Mt/a.
+Initially, we focus on a net-zero carbon emissions target, limiting carbon sequestration to 200~Mt/a.
+This constraint ensures that carbon removal is reserved for the hardest-to-abate emissions, while avoiding reliance on sequestration where other mitigation options are feasible.
 We give a detailed description of the different technological and geographical impacts of the transportation systems.
@@
 All scenarios share key energy system characteristics but differ in system costs and technology use (Fig.~\ref{fig:cost_bar}).
-Costs in the \baselinescenario{} are the highest at 764~bn€/a.
+Costs in the \baselinescenario{} are the highest at 764~bn€a.
 Scenarios with carbon and hydrogen networks are more cost-effective, reducing costs by 3.1\% in the \carbongrid{}, 4.7\% in the \hydrogengrid{}, and 5.3\% in the \hybridscenario{}.
@@
 In the \baselinemodel, transmission costs (electricity, gas, hydrogen, carbon) are below 4\% (25 bn€) and are dominated by electricity grid expenses.
-In contrast, the other models have higher combined pipeline costs, peaking at 20\% in the \hybridmodel.
+In contrast, the other models have higher combined pipeline costs, peaking at 20\% in the \hybridscenario.
 Both carbonaceous fuel production and electrolysis account for about 4\% (30 bn€) and 4.7\% (37 bn€) of costs respectively, with heat pump installations contributing around 8\% (64 bn€/a in the \baselinemodel).
@@
 Limitations of the study, such as the neglect of synthetic fuel imports and assumptions on biomass potentials, are discussed in the Appendix in Section~\ref{sec:limitations}.
-A sensitivity analysis of the costs of \carbon{} pipelines, which show negligible effects on the results, can be found in the Appendix in Section~\ref{sec:subsidy}.
+The Appendix in Section~\ref{sec:sensitivities} provides an analysis of the model's sensitivities.
+Notably, relaxing the sequestration limit has a significant impact, reducing system costs by 9.1\% at 800~Mt/a in the hybrid scenario.
+This reduction is achieved by opting for fossil fuels with carbon capture, transport, and sequestration instead of carbon utilization.
+Therefore, investments in renewable power assets which serve the synthetic fuel production for the aviation, shipping and industry sectors drop by roughly a third.
+Less effects can be seen from reducing capital costs for \carbon{} pipelines and costs of electrolyzers, on one hand indicate robustness against price changes but also little effect of subsidizing these technologies.
+As the cost optimum is affected by the sequestration potential limits, a more detailed analysis of a relaxed sequestration limit is presented in the Appendix in Section~\ref{sec:higher_sequestration}.
 
@@
 Furthermore, we decided to limit the total amount of sequestered \carbon{} to 200~Mt per year for our net-zero \carbon{} emission target and 660~Mt per year for our net-negative emission target.
+This constraint is deliberately implemented to avoid over-reliance on carbon capture and sequestration as a backstop technology, which could otherwise offset fossil emissions instead of prioritizing the reduction of fossil fuel use where technologically feasible.
+By capping the sequestration at 200 Mt/year, we aim to promote a balanced energy transition, where sequestration is used to offset hard-to-abate fossil process emissions (e.g., calcination of limestone in cement manufacturing) but does not become a substitute for more sustainable mitigation measures.
+The implications of this choice are discussed in the Appendix in Section~\ref{sec:higher_sequestration}.
+
 This is enough to offset hard-to-abate fossil process emissions (e.g., calcination of limestone in cement manufacturing) and some limited slack for fossil fuels to be used where they can be most valuable.
@@
 
-\subsection{Subsidizing \carbon{} pipelines}
+\clearpage
+\subsection{Sensitivities}
+\label{sec:sensitivities}
+
+\subsubsection*{Higher sequestration limits}
+\label{sec:higher_sequestration}
+Carbon management in the optimized energy system, particularly in the hard-to-abate sectors, is heavily influenced by the model's sequestration potential, which is capped at 200~Mt/year.
+Lifting the sequestration limit leads to cost benefits and a shift in primary energy supply and its implications (see Fig.~\ref{fig:sequestration_cost_bar}).
+With higher sequestration rates, more fossil fuels are used in the aviation, heating, and industry sectors in land-locked regions where hydrogen is expensive.
+This reduces the reliance on carbon utilization (CU) and, therefore, the production and transport of hydrogen.
+At the same time, the carbon network is expanded to transport more carbon from inland point sources to sequestration sites.
+Sequestration is further supported by large DAC facilities near coastal areas to offset additional emissions from fossil fuels.
+These effects lead to significant changes in investments when increasing sequestration from 200~Mt/a to 800~Mt/a.
+Due to reduced hydrogen production, investments in wind and solar drop by a third.
+Total system costs are reduced by 9.1\%, primarily due to lower renewable energy investments.
+Regarding transport networks, hydrogen network capacity is halved, while the carbon network quadruples (see Fig.~\ref{fig:sequestration_cost_bar_transmission}).
+These findings are consistent with the effects discussed in~\cite{hofmannDesigningCO2Network2023}.
+
+\begin{figure}[ht!]
+    \centering
+    \begin{subfigure}{.5\textwidth}
+        \includegraphics[width=\linewidth]{comparison/sequestration/figures/90_nodes/cost_bar.png}
+        \caption[short]{}
+        \label{fig:sequestration_cost_bar}
+    \end{subfigure}%
+    \begin{subfigure}{.5\textwidth}
+        \includegraphics[width=\linewidth]{comparison/sequestration/figures/90_nodes/cost_bar_transmission.png}
+        \caption[short]{}
+        \label{fig:sequestration_cost_bar_transmission}
+    \end{subfigure}
+    \caption{System costs for the \hybridscenario{} as a function of the sequestration limit (a). The transmission system cost for the \hybridscenario{} as a function of the sequestration limit (b).}
+\end{figure}
+
+As sequestration capacity increases, the cost savings from additional sequestration begin to plateau.
+At 200~Mt/a, the shadow price of the sequestration constraint shows that each additional tonne saves 225~€/t, but at 800~Mt/a, this saving drops to 51~€/t.
+Extrapolating this behavior, saturation of sequestration can be expected around 1200~Mt/a.
+
+While increasing sequestration capacity offers economic benefits, it poses potential risks.
+Over-reliance on sequestration may delay necessary investments in cleaner technologies and renewable energy.
+Additionally, the large-scale expansion of carbon storage infrastructure may encounter public opposition due to concerns about the long-term stability of CO$_2$ storage sites and possible leakage risks.
+Furthermore, the focus on sequestration could perpetuate the use of fossil fuels in sectors where decarbonization alternatives, such as electrification or green hydrogen, may offer more sustainable solutions.
+Thus, careful consideration must be given to avoid locking in fossil fuel usage and undermining the broader energy transition goals.
+
+\subsubsection{Lower \carbon{} pipelines costs}
 \label{sec:subsidy}
 
-The effect of a 50\% subsidy on \carbon{} pipelines impacts the optimal technology deployment to a small extent.
+The effect of a 50\% cost-reduction on \carbon{} pipelines, such as through subsidies, impacts the optimal technology deployment to a small extent.
 Fig.~\ref{fig:cost_bar_subsidy} shows the corresponding total system cost of all scenarios with a net carbon neutrality target.
@@
 \clearpage
-\subsection{Cost comparison across scenarios}
+\subsection{Other sectors in the energy system}
+
+The following maps show the optimal supply, demand, transport and prices of energy carriers in the \hybridscenario{} not covered in the main text.
+
+\begin{figure}[ht!]
+    \centering
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\linewidth]{full/figures/90_nodes/balance_map_electricity.png}
+        \caption{Optimal supply, consumption, transport and prices of electricity in the \hybridscenario{}.}
+        \label{fig:balance_map_electricity}
+    \end{subfigure}%
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\linewidth]{full/figures/90_nodes/balance_map_heat.png}
+        \caption{Optimal supply, consumption, transport and prices of heat in the \hybridscenario{}.}
+        \label{fig:balance_map_heat}
+    \end{subfigure}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\linewidth]{full/figures/90_nodes/balance_map_gas.png}
+        \caption{Optimal supply, consumption, transport and prices of gas in the \hybridscenario{}.}
+        \label{fig:balance_map_gas}
+    \end{subfigure}%
+    \begin{subfigure}{0.5\textwidth}
+        \includegraphics[width=\linewidth]{full/figures/90_nodes/balance_map_biogas.png}
+        \caption{Optimal supply, consumption, transport and prices of biogas in the \hybridscenario{}.}
+        \label{fig:balance_map_biogas}
+    \end{subfigure}
+\end{figure}
+
+\clearpage
+\subsection{Cost comparison across net-zero scenarios}
 \label{sec:cost_comparison}

### 2024-09-06 workflow/scripts/plot_captureshare_line.py

@@ -30,7 +30,7 @@ which = "capacity"
 share = {}
 cf = {}
 for path in snakemake.input.networks:
-    n = import_network(path)
+    n = import_network(path, secondary_color=True)
     key = snakemake.params.labels[n.meta["wildcards"]["run"]]
 
     s = n.statistics

### 2024-09-06 report/submissions/journal/paper/paper.tex
finish addressing review comments

@@
 \usepackage{algorithmic}
+\usepackage{booktabs}
 \usepackage{graphicx}
@@
 
-To investigate the competing transport dynamics between \carbon{} and \hydrogen{} networks, we contrast the following four scenarios:
+To investigate the competing transport dynamics between \carbon{} and \hydrogen{} networks, we contrast the following four scenarios with different sets of networks present:
 %
-\begin{itemize}
-    \item[] \textit{Baseline:} Neither \carbon{} nor \hydrogen{} networks are constructed.
-    \item[] \textit{\carbongrid{}:} Only the \carbon{} network is developed.
-    \item[] \textit{\hydrogengrid{}:} Only the \hydrogen{} network is developed.
-    \item[] \textit{Hybrid:} Both \carbon{} and \hydrogen{} networks are developed.
-\end{itemize}
+
+\begin{table}[ht!]
+    \centering
+    \begin{tabular}{c|c|c}
+        \toprule
+        \textbf{Scenario} & \textbf{\carbon{} Network} & \textbf{\hydrogen{} Network} \\
+        \midrule
+        Baseline & -- & -- \\
+        \carbongrid{} & \checkmark & -- \\
+        \hydrogengrid{} & -- & \checkmark \\
+        Hybrid & \checkmark & \checkmark \\
+        \bottomrule
+    \end{tabular}
+    \caption{Set of transportation networks in the developed scenarios.}
+    \label{tab:scenarios}
+\end{table}
 
@@
 
-\subsection*{Stricter emission targets increase \carbon{} network benefits}
+\subsection*{Network layouts are robust againg tighter emission targets}
 \begin{figure}[htb!]
@@
 
-When applying a stricter \carbon{} emissions target and forcing the model to remove 460~Mt \carbon{} over one year (equivalent to 10\% of \carbon{} emissions in 1990) while sequestering maximally 660~Mt, investment in DAC and associated solar and wind power and heat supply covers a large share of additional system costs in all scenarios (see Fig.~\ref{fig:net-negative_cost_bar}).
-In the \hybridscenario{}, this share is at 80\% while new carbon capture at point sources only makes up 6.5\% of new system costs.
-This shows that the exploitation of carbon point sources is close to its feasible limit.
-The additional sequestration is primarily supplied by large DAC facilities near the shore.
-To a much smaller extent, carbon captured from bio-energetic sources in the inland, that under a net-zero emissions assumption supplies CU, is instead transported and sequestered (compare Fig.~\ref{fig:balance_maps_full_nn}).
-Some of the CU which in net-zero scenarios locates in Central Europe instead locates in Spain, where it uses carbon from new DAC facilities.
+Applying a stricter \carbon{} emissions target and forcing the model to remove 460~Mt of \carbon{} in one year (equivalent to 10\% of 1990 \carbon{} emissions) while sequestering a maximum of 660~Mt, investment in DAC and associated solar and wind power and heat supply covers a large proportion of additional system costs in all scenarios (see Fig.~\ref{fig:net-negative_cost_bar}).
+In the \hybridscenario{}, this share is 78\%, while new carbon capture at point sources accounts for only 4\% of incremental system costs.
+This shows that point source carbon capture is closing the gap to its feasible limit.
+The additional sequestration is mainly provided by large DAC facilities near the coast.
+To a much lesser extent, carbon captured from inland bioenergy sources, which would provide CU under a net-zero assumption, is instead transported and sequestered (see Fig.~\ref{fig:balance_maps_full_nn}).
+Some of the CU located in Central Europe in net-zero scenarios is instead located in Spain, where it uses carbon from new DAC facilities.
 
-In the \carbon{} and \hydrogen{} scenarios, this slight shift in CU deployment results in a lower buildout of hydrogen pipelines, especially from the Iberian Peninsula to Central Europe (0.5~bn€/a).
-Nonetheless, 80\% of the hydrogen topology which is built under net-zero emissions is also built under the assumption of net-negative emissions.
-Under net-negative emissions the \carbon{} network also includes routes from Central and Eastern Europe to the shores (1.2~bn€/a).
-The benefits of \carbon{} transport and sequestration with tighter emission targets are also evident when comparing the other scenarios.
-The carbon network under the net-negative \carbonscenario{} is larger than under the net-zero \carbonscenario{} and transports more carbon from point sources to sequestration sites rather than to CU facilities.
-In the \hydrogenscenario{}, where all carbon for CU and CS must come from DAC, the hydrogen transport topology remains almost identical to the net-zero scenarios (for detailed cost contributions, refer to sections~\ref{sec:cost_comparison} and~\ref{sec:operation_nn} in the Appendix).
+These changes have little effect on the layout of the \carbon{} and \hydrogen{} networks.
+The slight shift in CU deployment results in less hydrogen pipeline construction in some parts of the system.
+Nevertheless, 80\% of the hydrogen topology built under the net-zero emissions assumption is also built under the net-negative emissions assumption.
+The \carbon{} network partially extends routes from Central and Eastern Europe to the coasts.
+The relative benefits of \carbon{} transport and sequestration with tighter emissions targets are evident when compared with the other scenarios.
+The carbon network in the net-negative \carbonscenario{} is larger than in the net-zero \carbonscenario{} and transports more carbon from point sources to sequestration sites than to CU facilities.
+In the \hydrogenscenario{}, where all carbon for CU and CS must come from DAC, the hydrogen transport topology remains almost identical to the net-zero scenarios (for detailed cost contributions see sections~\ref{sec:cost_comparison} and~\ref{sec:operation_nn} in the Appendix).
 
-Overall, this leads to the net cost increases of 77~bn€/a in the \carbonscenario{}, by 81~bn€/a in the \hydrogenscenario{}, and by 78~bn€/a in the \hybridscenario{}.
+Overall, this leads to the net cost increases of 78~bn€/a in the \carbonscenario{}, by 80~bn€/a in the \hydrogenscenario{}, and by 78~bn€/a in the \hybridscenario{}.
 The resulting cost-benefit of the \hybridscenario{} in comparison to the other scenarios is 6~bn€/a.
@@
 However, a different set of costs on DAC and electrolysis may well impact the model results.
-In combination with a net \carbon{} neutrality or reduction target, a relaxation of the global sequestration limit drives system costs down by introducing fossil fuels, as shown in~\cite{hofmannDesigningCO2Network2023}.
-Allowing the model to sequester onshore, would likely reduce the dominant co-location of CU with the point sources due to easier access to sequestration sites.
-Finally, DAC deployment could be in parts replaced if larger sustainable biomass sources are assumed, as shown in ~\cite{lauerCrucialRoleBioenergy2023}.
+Considering different strategies for carbon removal, DAC deployment could be in parts replaced if larger sustainable biomass sources are assumed, as shown in~\cite{lauerCrucialRoleBioenergy2023}.
+
+Finally, our energy system model assumes total autarky of the European system and neglects potential imports from elsewhere.
+Depending on the type of energy carrier and whether it is \carbon{} neutral or not, allowing imports can change the optimal design and operation of the system quite significantly, as discussed in~\cite{neumannEnergyImportsInfrastructure2024}.
+In our model, we see that when higher sequestration rates are allowed, the model imports more fossil oil, which reduces the use of CU in Central Europe and thus the hydrogen transport from South to Central Europe (see section~\ref{sec:higher_sequestration}).
+While this substitution reduces the need for Fischer-Tropsch fuel production, it also requires significant carbon sequestration within Europe to offset the emissions from fossil oil use.
+If the imported oil were carbon neutral or 'green', there would be no need for dedicated sequestration infrastructure as the emissions would already be accounted for.
+In such a case, the overall energy infrastructure would remain largely the same, minus the additional carbon sequestration capacity needed to offset the emissions from fossil imports.
+Thus both the \hydrogen{} and \carbon{} networks would be less relevant.
+In the context of green hydrogen imports, as highlighted by~\cite{neumannEnergyImportsInfrastructure2024}, the model is expected to reduce the use of electrolysis and methanolisation in southern European countries, as well as hydrogen transport routes from there to central Europe.
+However, the deployment of CU near point sources of \carbon{} would still be as in our scenario.
+The same applies to the deployment of a carbon network driven by fixed sequestration sites.
 

### 2024-09-09 report/submissions/journal/review-responses/review-responses-first-round.md
address internal review

We thank both reviewers for their helpful comments. We have rerun all scenarios in the paper with improved assumptions as suggested by the authors (improved accounting for leakage and compression losses) as well as new scenarios requested by the reviewers for a sensitivity to sequestration rates.

## Reviewer #1 (Remarks to the Author):

@FabianHofmann
> The article presents an interesting analysis of the foreseen H2 and CO2 transport infrastructures.
> Literature is not accurate and there are many references cited without going in their details.
> The study lacks of realistic elements such as leakage issues, location of compression stations in relation with the capture location.

Thank you for your comments. We appreciate your feedback and will address your concerns in the revised manuscript.

We have expanded the literature review section to provide more context for the study, as well as discussing recent policy measures for carbon management in the EU. Since the study touches a wide range of topics in the political and technological landscape, we have tried to find a good balance between providing a comprehensive overview and keeping the text concise. (#TODO)

To simulate a more realistic transport system, the following assumptions have been integrated into the energy system model for transport efficiencies and compression requirements. For hydrogen pipelines an average loss of 1.2% per 1000 km, aligning with industry estimates that account for hydrogen's high permeability and small molecular size, which can lead to higher losses compared to other gases​. The compression energy required is assumed to be 1.8% of the transported energy, which is consistent with the significant energy demands for compressing hydrogen due to its low energy density​ [Environmental Defense Fund​](https://www.edf.org/sites/default/files/documents/Pipeline%20Methane%20Leaks%20Report.pdf). For natural gas pipelines, the model assumes a 1% loss rate per 1000 km, accounting for typical leakage and operational losses in well-maintained natural gas infrastructure​ as reported in [Global Energy Monitor](https://www.gem.wiki/Natural_gas_transmission_leakage_rates). The compression requirement is set at 1%, which falls within the common range for natural gas pipeline operations​ [Environmental Defense Fund​](https://www.edf.org/sites/default/files/documents/Pipeline%20Methane%20Leaks%20Report.pdf)​. For CO2 pipelines an average loss rate of 0.8% per 1000 km is reflecting a generally low leakage rate observed in supercritical or dense phase CO2 transport​ [SpringerLink](https://link.springer.com/article/10.1007/s11356-023-27289-3) [MDPI](https://www.mdpi.com/1996-1073/14/15/4601)​. The compression energy needed is set to 250 kWh per 1000km and tonne CO2 reflecting the energy required to compress and maintain CO2 in a supercritical state for efficient transport​.

---
@FabianHofmann
> There are several graphs that make the reader confused rather than explain the amount of work done by the authors.

We have revised the layout of the figures to make them more accessible to the reader and have adjusted the color scheme where colors tended to be quite similar. However, we would appreciate more specific feedback on which figures you find confusing so that we can address these issues more effectively. Most of the figures contain a lot of information that we feel is necessary to support the argument. We have tried to make them as clear as possible.

---

@nworbmot
> The results are not understandable in terms of transferring into policies.

We have written a short policy outlook in the conclusion as well as linking in the introduction to the recently-announced industrial carbon management strategy of the European Commission. Our addition to the conclusion reads:

"For policy makers these results show the need for coordinated planning across sectors like carbon, hydrogen and synthetic fuels since they are strongly interacting, as well as across borders because of the localized nature of good renewable resources and sequestration potentials. Deploying multiple networks offers some robustness should unforeseen problems arise with one of them, and our results show that the system can even cope with neither a CO2 nor H2 network for a cost penalty.  Many of the technologies have not yet been deployed at scale before and will need support for financing, regulation and gaining the acceptance of local populations."

---

## Reviewer #2 (Remarks to the Author):
@FabianHofmann

> Overall, a very impressive paper, with a very clear and concise structure, easy to follow with many interesting insights. The research gap and necessity of the paper clearly are demonstrated, the topic is higly relevant and the chosen methodology are well suited for the task.

> There are a couple of aspects I would like to understand better, i.e. aspects in which I think additional information would be beneficial.

> 1: CCS is a challenge for scenario analyses like yours, since it low cost makes it a backstop technology. The model will try to solve many things with CCS if one does not prevent somehow. Your approach is to limit CS. “This is in order to avoid offsetting fossil emissions with carbon removal if they can technologically be avoided in the first place.” I was wondering about this throughout the paper. So, first of all I could be helpful to point out the importance and weight of this setting early in the text. I think the implications of this should be discussed more, because I believe this to be crucial: At your point of operation, your model is able to provide and sequester CO2 for roughly160 €/t if I’m correct. That basically means that all mitigation measures costlier than that are only part of the solution because of the CS limit. It would be interesting to discuss the dual value/shadow price of tis cap, to understand how much economic pressure could be on relying more on CCS. The fossil lock-in your CS cap prevents could be an economically very attractive. So, if compensating fossil fuel usage is cheaper than the chosen path of FT and other PtX options, I think this would deserve a discussion in the paper. Especially because the compensation would not need to take place in Europe but could be done in regions with even lower DACS/BECCS costs.

Thank you for your insightful comment. In response, we have added an explanation for our choice to the introduction and have added new modelling runs in a new subchapter in the Appendix "Higher Sequestration Limits" which we refer to in the main text. The added chapter discusses new model results for higher seqestration limits and their implications.

Our analysis shows that increasing the limit from 200 Mt/year to 800 Mt/year reduces system costs quite significantly, by up to 9.1%. This reduction is primarily due to a shift towards greater reliance on fossil fuels either with CCS or unabated and compensated with CDR, which in turn reduces the need for CU and renewable investments by about a third. At the same time, the hydrogen network loses importance whereas the carbon network quadruples its capacity. Point-sources of carbon in the inland where fossil fuels are used, like gas for heating or naphtha for industry, are nearly all captured and tranpsorted to be sequesterd.

Another interesting finding is the drop of the shadow price from €225/t to €51/t when increasing the sequestration limit. Additional sequestration capacity provides economic benefits, these benefits, however, diminish as the limit increases​.

However, this also introduces the risk of reinforcing fossil fuel dependence, potentially undermining long-term decarbonization efforts. The discussion in the appendix now highlights these trade-offs, emphasizing the importance of carefully balancing sequestration limits to avoid economic over-reliance on CCS​.

---

@chrstphtrs
> 2. The H2 provision in the Baseline case confuses me a bit, namely that it works surprisingly well. How is a constant supply possible in regions without geological storage potential? Does this mean that in hours of low/no RES power plants in the regions with storages produce electricity, which is then brought to H2 sinks and turned into H2 again? In our scenarios this led to much bigger costs and almost impossible necessities for electricity network expansions. Is this because your H2 demand is relatively low, and you transform it into other carriers? Or do you have some other form of storage option at the consumption sites. Overall, I would just like to understand better how you supply H2 steadily without a network.

Thank you for the comment, this is indeed a very interesting question.

The spatially-fixed hydrogen demand from industry (primarily for primary steel but also for ammonia) is comparatively low, while the bulk of the H2 demand (FT, methanolisation, methanation) can be moved by the model to locations with low-cost renewables and geological H2 storage. This avoids over-burdening the electricity grid.

In addition, to provide hydrogen in some regions with poor renewable resources the model resorts to SMR with either green or fossil methane transported in the gas grid.

---
@fneum
> 3: In the introduction: “It is assumed that Europe is self-sufficient in energy and does not import any fuels.” I would like to see the consequences of this setting discussed, especially regarding the large liquid quantities it entails. Just a suggestion, but a variation in which half of the quantities are imported discussed briefliy in the Annex would be very valuable. Much of the infrastructure you model and discuss discuss hinges on this, and importing a major share is clearly an realistic and attractive option in reality.

This is indeed an important point that we have not discussed in detail in the paper. The assumption of Europe's self-sufficiency in energy was made to simplify the system design and focus on internal infrastructure. It also has the advantage of making good use of Europe's sustainable biomass sources for sourcing CO2, and avoiding the difficulty of regulating sustainable CO2 provision for carbonaceous fuels synthesised outside of Europe. However, we agree that importing a significant share of energy carriers, particularly liquid fuels, is a realistic and potentially cost-beneficial option.

We have expanded the discussion of the limitations in the appendix to explore the potential consequences of allowing imports. In the higher sequestration scenario, for example, our model shows that increased imports of fossil oil significantly reduce the need for domestic Fischer-Tropsch (FT) fuel production. While this reduces hydrogen transport and carbon utilization requirements in Central Europe, it also necessitates substantial carbon sequestration within Europe to offset emissions from fossil-based imports. If these imports were carbon-neutral, the need for dedicated sequestration infrastructure would be significantly reduced, as the emissions would already be compensated.
Thus, the carbon neutrality of imported fuels has a profound effect on the optimal infrastructure layout, especially for hydrogen transport and carbon sequestration. A scenario that allows for partial energy imports could lead to lower investments in the carbon and hydrogen networks as well as direct air capture.

However, a critical political and strategic consideration is the control over the CO2 supply chain. There may not be sufficient sustainable CO2 available abroad, and Europe may prefer to maintain control of the CO2 supply chain to ensure security and reliability. Relying on external sources for carbon-neutral fuels could introduce geopolitical risks, similar to those faced with fossil fuel imports.

---
@chrstphtrs
> Your H2 prices seem comparatively low to similar exercises. What interest rate/ WACC did you use? I think that should always be made explicit, as it has huge a impact on many results.

We use technology cost assumptions from the Danish Energy Agency referenced in the Methods section. The database applies a discount rate of 7% to derive annualized investment costs for all major technologies, including hydrogen electrolyzers. We have added this information to the Methods section.

---
@FabianHofmann
> Maybe this would be to much, but while seeing your H2 prices I was very interested in seeing your electricity prices. If a map/maps would be too much, maybe state the ranges somewhere for reference?

We have added a chapter to the Appendix that showes maps for other energy carrier (electricity, heat, gas, biomass) in the hybrid scenario. These show the average production, consumption, transport and prices per carrier. We hope this will provide the information you are looking for.

---
@chrstphtrs
> I think an overview table with differences of the scenario settings in the beginning would be helpful, and maybe with some key results in the end.

We have added an overview table in the introduction that shows the main differences between the scenarios. Since the difference of the scenarios is important on the qualitative level, we have not added a table with key results in the end. We hope that the reader can follow the argumentation and the results are clear enough.

---
@FabianHofmann
> The colour coding in Figure 6 could be better, the greens are almost indistinguishable and too many similar browns.

Thank you for your feedback. We have revised the color scheme in Figure 6 to make it more accessible to the reader. We hope that the new color scheme will make the figure easier to read.

---
@chrstphtrs
> In the end of section 2: “In the network layouts, the slight shift in CU deployment results in a partial dismantling of hydrogen pipelines, especially from the Iberian Peninsula to Central Europe (0.5 bn€/a).” Is “dismantling” really the right word here? The model simply does not build those pipelines in this scenario, and you cannot dismantle what you never built. Or maybe you can, and this is a tricky koan.

You are absolutely correct. We rephrased this passage, and similar ones, that the the net-zero and net-negative scenarios are modelled as different futures, not as chronological events.

---
@chrstphtrs
> “High-cost Direct Air Capture”: I may be nitpicking, but I was wondering whether this expression is really fair in combination with the CS cap. I think you force the model to pursue many even more expensive option like FT, this is just the most expensive carbon option for sequestration. Again, really just a thought.

You are correct, the relative statement of low and high cost only applies to elementary carbon dioxide needed for sequestration. We amended the statement to make it more clear.

### 2024-09-09 report/submissions/journal/paper/paper.tex
address internal review

@@
 In direct comparison, a hydrogen network is more cost-effective than a carbon network, since it serves to transport hydrogen to demand and to point source of carbon for utilization.
-However, in a hybrid scenario where both networks are present, the carbon network effectively complements the hydrogen network, promoting carbon capture from biomass and reducing reliance on direct air capture.
-Our analysis suggests integrating hydrogen and carbon dioxide networks into European energy policy for a robust, carbon-neutral or carbon-negative future energy system.
+However, in a hybrid scenario where both networks are present, the carbon network effectively complements the hydrogen network, promoting carbon capture from distributed biomass and reducing reliance on direct air capture.
+The layouts of the hydrogen and carbon dioxide networks are robust if the climate target is tightened to be net negative.
 \end{abstract}
@@
 It offers an efficient way to transport and store energy over long distances and periods and can be produced from electrolysis or steam methane reforming (SMR), for example.
-Captured carbon can be stored in geological formations to extract it from the atmosphere permanently, a process known as carbon sequestration (CS).
-Additionally, carbon can be combined with hydrogen to produce carbonaceous fuels, such as synthetic kerosene for aviation, synthetic methanol for shipping or synthetic methane for industrial feedstock.
-We jointly refer to the production of these carbonaceous fuels as Carbon Utilization (CU).
+Captured carbon can be stored in geological formations permanently, a process known as carbon sequestration (CS).
+Additionally, carbon can be combined with hydrogen to produce carbonaceous fuels, such as synthetic kerosene for aviation, synthetic methanol for shipping or synthetic methane for industrial feedstock or synthetic naphtha as a feedstock for high value chemicals.
+The production of these carbonaceous fuels is jointly referred to as Carbon Utilization (CU).
 Carbon can be effectively captured from industrial processes and the combustion of biomass, fossil fuels, or synthetic carbonaceous fuels through point-source carbon capture (CC) techniques, or harvested from the atmosphere using direct air capture (DAC).
-Both \carbon{} and \hydrogen{} networks are likely to play a crucial role in the cost-effective integration of these technologies to enable net-zero economies.
+Both \carbon{} and \hydrogen{} networks are likely to play a crucial role in the cost-effective integration of these carbon management technologies to enable net-zero economies.
 
@@
 Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synthetic fuels on the one hand and offer to purchase \carbon{} on the other hand.
-The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a transport and sequestration capacity of 1.5~Mt \carbon{} per year to be operative in 2024, expanding to a targeted scale of 5~Mt per year of sequestration by 2030.
-The market potential for carbon capture and sequestration in Europe is underscored by the Capture Map's estimated potential~\cite{ToolsGreenTransition} of 1.7 Gt carbon annually from point sources, accounting for about half of the continent's emissions, as well as estimated sequestration capacity potentials of up to 3~Gt/a~\cite{europeancommissionEuropeanCO2Storage}.
+The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning a transport and sequestration capacity of 1.5~Mt \carbon{} per year to be operative in 2024, expanding to a targeted scale of 5~Mt per year of sequestration by 2030.
+The market potential for carbon capture and sequestration in Europe is underscored by the Capture Map's estimated potential of 1.7 Gt carbon annually from point sources~\cite{ToolsGreenTransition}, accounting for about half of the continent's emissions, as well as estimated sequestration capacity potentials of up to 3~Gt/a~\cite{europeancommissionEuropeanCO2Storage}.
 However, it is important to note that carbon sequestration technology, unlike carbon capture, is not yet fully mature and its full potential and implications remain to be explored in depth.
 To advance the carbon economy, the Clean Air Task Force emphasizes the importance of developing a carbon transport system in Europe alongside a hydrogen network~\cite{lockwoodEuropeanStrategyCarbon}.
-Carbon pipelines, considered a mature technology, have seen widespread installations in the United States and Canada, primarily to supply \carbon{} to enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
+Carbon pipelines, considered a mature technology, have seen widespread installations in the United States and Canada, primarily to supply \carbon{} to enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmann2020net}.
 
-Up to this point, it remains unclear how the two transport systems of hydrogen and carbon may complement or even compete with each other.
+Despite the abundance of policies and initiatives, up to this point, it remains unclear how the two transport systems of hydrogen and carbon may complement or even compete with each other.
 Both can bridge the misalignment of sources and sinks for carbon and hydrogen.
@@
 
-In the literature, the two network approaches and underlying technologies have been discussed in several publications, all of which, however, dealt with the isolated aspects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022,neumannBenefitsHydrogenNetwork2022}.
-Such techno-economic models, in contrast for example to Integrated Assessment Models, can account for the spatial distribution of carbon sources and sinks, which are crucial to provide a holistic view of the energy system and its technological interactions.
-Neumann et al.~\cite{neumannBenefitsHydrogenNetwork2022} demonstrated the interchangeability of hydrogen and electricity grid expansion in a climate-neutral European energy system at the expense of higher investments.
-The underlying, highly-resolved model encompasses the most relevant sectors, however it neglects the option to transport carbon across regions.
+In the literature, the two network approaches and underlying technologies have been discussed in several publications~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022,neumannBenefitsHydrogenNetwork2022}.
+Such techno-economic models, in contrast for example to integrated Assessment Models, can account for the spatial distribution of carbon sources and sinks, which are crucial to provide a holistic view of the energy system and its technological interactions.
+However, all of the techno-economic studies have only dealt with isolated aspects of hydrogen and carbon networks.
+Neumann et al.~\cite{neumannBenefitsHydrogenNetwork2022} demonstrated the extent to which hydrogen and electricity grid expansion is interchangeable in a climate-neutral European energy system at the expense of higher investments.
+The underlying, highly-resolved model encompasses the most relevant sectors, however it neglects the infrastructure required to transport carbon accross regions.
 Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \carbon{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
 This limited sectoral scope cannot capture important dynamics of carbon management, since it neglects sectors like industry that will need to handle most \carbon{} in the future.
-The European Commission's Joint Research Center analysed the possible topology of a \carbon{} network in 2024~\cite{jrc2024}, but didn't include the interaction of  \carbon{} sources and sinks with the rest of the energy system.
-A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connected to a remote sequestration site in Norway.
+The European Commission's Joint Research Center analysed the possible topology of a \carbon{} network in 2024~\cite{jrc2024}, but did not include the interaction of  \carbon{} sources and sinks with the rest of the energy system.
+Becattini et al~\cite{becattiniCarbonDioxideCapture2022} present a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connected to a remote sequestration site in Norway.
 However, this limited spatial scope fails to consider other sequestration sites and the co-benefits from connecting the \carbon{} network to neighboring countries.
 Other high-resolution energy system models do not feature detailed CU technologies and \carbon{} transport.
-Euro-Calliope, for example, does not feature carbon nor hydrogen transport, and thus forces generic synthetic fuel production to rely on captive electrolysis and DAC in the same region~\cite{pickeringDiversityOptionsEliminate2022}.
+In Pickering et al.~\cite{pickeringDiversityOptionsEliminate2022}, the Euro-Calliope model, for example, features neither carbon nor hydrogen transport, and thus forces generic synthetic fuel production to rely on locale electrolysis and DAC in the same region.
 
 No study has yet considered the co-optimization and comprehensive assessment of both \carbon{} and \hydrogen{} networks in a fully sector-coupled energy system.
-However, such an assessment is strongly needed to identify realms of competition and synergy between hydrogen and carbon management technologies.
+Such an assessment is strongly needed to identify realms of competition and synergy between hydrogen and carbon management technologies.
 In this paper, we present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies.
@@
 
-We analyze how the deployment of carbon and hydrogen networks affects different decarbonization strategies in the scenarios and how investments in none or only one of the transport technologies lead to inefficient market layouts.
+We analyze how the option to deploy carbon and hydrogen networks affects different decarbonization strategies in the scenarios and how investments in none or only one of the transport technologies lead to inefficient infrastructure layouts.
 Initially, we focus on a net-zero carbon emissions target, limiting carbon sequestration to 200~Mt/a.
-This constraint ensures that carbon removal is reserved for the hardest-to-abate emissions, while avoiding reliance on sequestration where other mitigation options are feasible.
+This constraint ensures that carbon sequestration is reserved for the hardest-to-abate emissions, while avoiding reliance on sequestration where other mitigation options are feasible.
 We give a detailed description of the different technological and geographical impacts of the transportation systems.
-Subsequently, our analysis demonstrates the extent to which different deployment strategies remain robust under a tightened net \carbon{} removal target.
+Subsequently, our analysis demonstrates the extent to which different deployment strategies remain robust under a tightened target with net \carbon{} removal.
 
@@
 
-All scenarios share key energy system characteristics but differ in system costs and technology use (Fig.~\ref{fig:cost_bar}).
-Costs in the \baselinescenario{} are the highest at 764~bn€a.
+All scenarios share key characteristics but differ in system costs and technology use (Fig.~\ref{fig:cost_bar}).
+Costs in the \baselinescenario{} are the highest at 764~bn€/a.
 Scenarios with carbon and hydrogen networks are more cost-effective, reducing costs by 3.1\% in the \carbongrid{}, 4.7\% in the \hydrogengrid{}, and 5.3\% in the \hybridscenario{}.
@@
 While in the \baselinescenario{}, the model has neither a carbon nor a hydrogen network, it can expand both in the \hybridscenario{}.
-``Gas Infrastructure'' combines gas facilities for power and heat production, ``\carbon{} Infrastructure'' and ``\hydrogen{} Infrastructure'' combine transport and storage for each carrier.
-``Carbon Capture at Point Sources'' combines all technologies with integrated carbon capture, including the cost of the main facility (e.g., CHP unit) and the carbon capture application.}
+``Gas Infrastructure'' combines gas facilities for transport, power and heat production, ``\carbon{} Infrastructure'' and ``\hydrogen{} Infrastructure'' combine transport and storage for each carrier.
+``Carbon Capture at Point Sources'' combines all technologies with integrated carbon capture, including the cost of the main facility (e.g., CHP units) and the carbon capture application.}
     \label{fig:cost_bar}
@@
 
-System cost differences between the scenarios are driven by wind and solar, as well as carbon capture technologies.
-In scenarios with carbon and hydrogen networks, the model primarily reduces the reliance on high-cost DAC, from 365~Mt/a carbon removal in the \baselinescenario{} to 113~Mt/a in the \hybridscenario{} (see Fig.~\ref{fig:balance_captured_carbon}).
-At the same time, carbon capture from biomass combustion, gas-based industry processes and biogas-to-gas upgrading increase, the latter lowering the need for synthetic methane production.
+System cost differences between the scenarios are driven by investments in wind and solar, as well as carbon capture technologies.
+In scenarios with carbon and hydrogen networks, the model primarily reduces the reliance on costly DAC, from 365~Mt/a carbon capture in the \baselinescenario{} to 113~Mt/a in the \hybridscenario{} (see Fig.~\ref{fig:balance_captured_carbon}).
+At the same time, carbon capture from biomass combustion, gas-based industry processes and biogas upgrading increase.
+The latter lowers the need for synthetic methane production.
 % Another knock-on effect of less reliance on DAC is its reduced endogenous demand for dedicated renewables as well as heat pumps and biomass CHPs to supply power and heat.
@@
 For CU, either hydrogen must be produced where carbon is cheap at point sources, or DAC must be used at sites with low-cost hydrogen.
-Where carbon from point sources cannot be sequestered, it has to be compensated elsewhere by negative emissions from DAC or biomass capture with sequestration.
-All of these options are sub-optimal from a cost perspective.
+Where fossil carbon from point sources cannot be sequestered, it has to be compensated elsewhere by negative emissions from DAC or biomass capture with sequestration.
+To provide hydrogen in some regions with poor renewable resources the model resorts to SMR with either green or fossil methane transported in the gas grid.All of these options are sub-optimal from a cost perspective.
 
@@
 In the \carbonscenario, there are three main transport purposes of the low-cost carbon captured in Central and Eastern Europe (see Fig.~\ref{fig:balance_map}, top left).
-A large part of the carbon is transported to the Iberian Peninsula (9.8~kt/h) and the British Isles (11.8~kt/h), where it is used along with low-cost electrolytic hydrogen to produce Fischer-Tropsch fuels and methanol.
-In Italy and Greece, a portion of the carbon (3.3~kt/h) is directed to methanation facilities where the synthetic methane is fed into the gas grid at key terminals that also receive fossil gas imports.
+A large part of the carbon is transported to the Iberian Peninsula (avg.
+9.8~kt/h) and the British Isles (avg.
+11.8~kt/h), where it is used along with low-cost electrolytic hydrogen to produce Fischer-Tropsch fuels and methanol.
+In Italy and Greece, a portion of the carbon (avg.
+3.3~kt/h) is directed to methanation facilities where the synthetic methane is fed into the gas grid at key terminals that also receive fossil gas imports.
 Finally, a third part of captured carbon is transported directly to sites in the North Sea, Baltic Sea and Mediterranean Sea where it is sequestered.
@@
 Hydrogen prices vary widely across the continent, with low price ``valleys'' at around 60~€/MWh, where the CU is located, and high price ``peaks'' in Central Europe at around 95~€/MWh.
-Despite the transport system, carbon prices still vary from 105~€/t in Central Europe to 160~€/t at sequestration sites.
+Despite the transport system, marginal cost of carbon capture still vary from 105~€/t in Central Europe to 160~€/t at sequestration sites.
 
-In the \hydrogenscenario{}, the main hydrogen transportation route goes from Spain (66~GW) to Central Europe to supply local carbonaceous fuel production with captured carbon from point sources (see Fig~\ref{fig:balance_map}, bottom right).
+In the \hydrogenscenario{}, the main hydrogen transportation route goes from Spain (66~GW) to Central Europe to supply local carbonaceous fuel production with captured carbon from point sources, since \carbon{} can no longer be transported (see Fig~\ref{fig:balance_map}, bottom right).
 Also, the hydrogen network supplies spatially fixed hydrogen demand for industry in Central Europe (2.5~GW in total) at much lower costs than local electrolysis or SMR.
@@
 The largest carbon sink is located in Portugal (3.5~kt/h, see Fig~\ref{fig:balance_map}, top right).
-Carbon prices are low in Central Europe (110-120~€/t) and highest in sequestration regions (160~€/t).
+Marginal costs of carbon capture are low in Central Europe (110-120~€/t) and highest in sequestration regions (160~€/t).
 Hydrogen prices have flattened out at a relatively low level of €60~/MWh in all regions.
@@
 In the \hydrogenscenario{}, CU co-locates with carbon point sources and uses transported low-cost hydrogen.
+
+% CU also has an incentive to co-locate in Central Europe with district heating demand, which can use the waste heat from the Fischer-Tropsch process.
+
 % Both scenarios effectively combine low-cost renewable hydrogen production and carbon management for combustion and spatially-fixed industrial processes.
@@
 On the one hand, the carbon network enables the transport of low-cost captured carbon to sequestration sites, reducing the dependence on high-cost DAC near the coast which is present in the \hydrogenscenario{}.
-On the other hand, the \hydrogengrid{} is able to supply the spatially-fixed hydrogen demands in Central Europe with low-cost hydrogen imports, while the model in the \carbonscenario{} is forced to deploy electrolysis in regions with poor renewable resources as well as additional gas infrastructure in Central Europe.
+On the other hand, the \hydrogengrid{} is able to supply the spatially-fixed hydrogen demands in Central Europe with low-cost hydrogen imports, while the model in the \carbonscenario{} is forced to deploy electrolysis in regions with poor renewable resources.
 The latter aspect finally leads to lower system costs in the \hydrogenscenario{} than in the \carbonscenario{}.
@@
 Carbon Sequestration offshore is drawn in full circles.
-Lines and arrows show interregional transportation.
+Lines and arrows show interregional transportation volume.
 Carbon from point-source in the inland either supplies local CU with imported hydrogen or facilitates sequestration in nearby offshore regions.
@@
 Again, the model transports low-cost hydrogen to the point sources in Central Europe to supply CU, but also transports low-cost carbon from point sources close to the shore to sequestration sites (see Fig.~\ref{fig:balance_map_full}).
-This provides low-cost hydrogen across the regions and reduces reliance on high-cost DAC.
+This provides low-cost hydrogen across the regions and reduces reliance on more expensive DAC.
 
@@
 In this setup, the networks enable all carbon-emitting technologies that operate at a high average capacity factor, i.e., above 70\%, to capture most of their carbon (see Fig.~\ref{fig:captureshare_line}).
-Except for SMR and gas-fired CHP which operate a few weeks annually, more than 90\% emissions are captured.
-Due to the better integration of point sources, only a few DAC facilities are installed, such as those used for methanolisation in southern Spain.
-This, in turn, leads to reduced power consumption and less required wind and solar power deployment than in the other scenarios.
+More than 90\% of emissions are captured for most technologies.
+The exceptions are SMR and gas-fired CHP plants, which operate only a few weeks annually, making carbon capture units economically unfeasible for these applications.
+Due to the better integration of point sources, only a few DAC facilities are installed, such as those used for methanolisation in Spain.
+This, in turn, leads to reduced power consumption and less wind and solar power deployment than in the other scenarios.
 
-\subsection*{Network layouts are robust againg tighter emission targets}
+\subsection*{Network layouts are robust against tighter emission targets}
 \begin{figure}[htb!]
@@
 
-Applying a stricter \carbon{} emissions target and forcing the model to remove 460~Mt of \carbon{} in one year (equivalent to 10\% of 1990 \carbon{} emissions) while sequestering a maximum of 660~Mt, investment in DAC and associated solar and wind power and heat supply covers a large proportion of additional system costs in all scenarios (see Fig.~\ref{fig:net-negative_cost_bar}).
+In a further set of scenarios, we apply a stricter \carbon{} emissions target and force the model to remove 460~Mt of \carbon{} in one year (equivalent to 10\% of 1990 \carbon{} emissions) while sequestering a maximum of 660~Mt.
+There are additional system costs of 78--82~bn€/a in all scenarios, of which the largest proportion is the investment in DAC and associated solar and wind power and heat supply (see Fig.~\ref{fig:net-negative_cost_bar}).
 In the \hybridscenario{}, this share is 78\%, while new carbon capture at point sources accounts for only 4\% of incremental system costs.
-This shows that point source carbon capture is closing the gap to its feasible limit.
+This shows that point source carbon capture is already almost exhausted in the net-zero case.
 The additional sequestration is mainly provided by large DAC facilities near the coast.
 To a much lesser extent, carbon captured from inland bioenergy sources, which would provide CU under a net-zero assumption, is instead transported and sequestered (see Fig.~\ref{fig:balance_maps_full_nn}).
-Some of the CU located in Central Europe in net-zero scenarios is instead located in Spain, where it uses carbon from new DAC facilities.
+Some of the CU located in Central Europe in net-zero scenarios is instead diverted to Spain, where it uses carbon from new DAC facilities.
 
-These changes have little effect on the layout of the \carbon{} and \hydrogen{} networks.
+These changes have little effect on the layout of the \carbon{} and \hydrogen{} networks (see Fig.~\ref{fig:cost_map_difference_full_nn} and Fig.~\ref{fig:balance_maps_full_nn}).
 The slight shift in CU deployment results in less hydrogen pipeline construction in some parts of the system.
-Nevertheless, 80\% of the hydrogen topology built under the net-zero emissions assumption is also built under the net-negative emissions assumption.
+Nevertheless, 80\% of the hydrogen topology built under the net-zero emissions assumption is also built under the net-negative target.
 The \carbon{} network partially extends routes from Central and Eastern Europe to the coasts.
@@
 
-\section{Conclusion}
+\section{Discussion}
 \label{sec:conclusion}
@@
 We show that both networks strongly impact the optimal deployment of carbon capture, carbon utilization and sequestration.
-A \carbon{} network effectively reduces costs by transporting low-cost carbon from distributed point sources to sequestration sites and sites with low-cost hydrogen production to produce carbonaceous fuels.
+A \carbon{} network reduces costs by transporting low-cost carbon from distributed point sources to sequestration sites and sites with low-cost hydrogen production to produce carbonaceous fuels.
 A hydrogen network provides greater cost savings as low-cost hydrogen can be transported to regions with high demand, allowing carbon utilization at point sources.
@@
 In the hybrid configuration, the \hydrogen{} network supplies low-cost hydrogen to spatially fixed hydrogen demand and carbon utilization at point sources across regions.
-At the same time, the \carbon{} network transports low-cost carbon captured from point sources near the coast to sequestration sites, reducing the reliance on Direct Air Capture.
-This results not only in the lowest system cost but also in a more robust energy system.
+At the same time, the \carbon{} network transports low-cost carbon captured from point sources near the coast to sequestration sites, reducing the reliance on direct air capture.
+This results not only in the lowest system cost but also in a system robust to tightening the target to net-negative emissions.
 
@@
 The Appendix in Section~\ref{sec:sensitivities} provides an analysis of the model's sensitivities.
-Notably, relaxing the sequestration limit has a significant impact, reducing system costs by 9.1\% at 800~Mt/a in the hybrid scenario.
-This reduction is achieved by opting for fossil fuels with carbon capture, transport, and sequestration instead of carbon utilization.
-Therefore, investments in renewable power assets which serve the synthetic fuel production for the aviation, shipping and industry sectors drop by roughly a third.
-Less effects can be seen from reducing capital costs for \carbon{} pipelines and costs of electrolyzers, on one hand indicate robustness against price changes but also little effect of subsidizing these technologies.
-As the cost optimum is affected by the sequestration potential limits, a more detailed analysis of a relaxed sequestration limit is presented in the Appendix in Section~\ref{sec:higher_sequestration}.
+Notably, relaxing the limit from 200 to 800~Mt/a reduces system costs by 9.1\% in the hybrid scenario.
+This reduction is achieved by replacing the production of synthetic e-fuels with fossil fuels that are either captured and sequestered, or left unabated and compensated by carbon dioxide removal.
+Investments in renewable power assets and hydrogen infrastructure which serve the synthetic fuel production for the aviation, shipping and industry sectors drop by roughly a third.
+A similar effect on synthetic fuel production in Europe would be expected if synthetic fuels were instead imported from outside Europe.
+Smaller effects can be seen from reducing capital costs for \carbon{} pipelines and costs of electrolyzers.
 
@@
 The model optimizes the design and operation of the European energy system, encompassing the power, heat, industry, agriculture, and transport sectors, including international aviation and shipping.
-All technology cost assumptions are taken for the year 2040 and sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.
-The database applies a discount rate of 7\% for all major technologies, accounting for the weighted average costs of capital (WACC) for investments.
+All technology cost assumptions are taken for the year 2040 and sourced from an open-source database~\cite{zeyenPyPSATechnologydataTechnology2023}.
+The database applies a discount rate of 7\% for all major technologies, accounting for the weighted average costs of capital (WACC) of investments.
 Many of the assumptions contained therein are based on the technology catalogs published by the Danish Energy Agency~\cite{danishenergyagencyTechnologyDataGeneration2019,thedanishenergyagencyTechnologyDataCarbon2023}.
-Endogenous model results include the expansion of renewable energy sources, storage technologies, Power-To-X conversion and transmission capacities, heating technologies, peaking power plants, and the deployment of gray, blue or green hydrogen, among others.
+Endogenous model results include the expansion of renewable energy sources, storage technologies, Power-To-X conversion and transmission capacities, heating technologies, peaking power plants, and the deployment of gray, blue or green hydrogen production facilitates, among others.
 The model considers various energy carriers like electricity, heat, hydrogen, methane, methanol, liquid hydrocarbons and biomass, together with the corresponding conversion technologies.
@@
 %
-In our configuration, the model's time horizon spans one year, with a 3-hour temporal resolution and a spatial resolution of 90 regions.
+In our configuration, the model's time horizon spans one year, with a 3-hourly temporal resolution and a spatial resolution of 90 regions.
 Each of the regions consists of a complex subsystem with technologies for supplying, converting, storing and transporting energy.
@@
 
-The topology and capacities of the electricity transmission system are taken from the ENTSO-E transparency map~\cite{wiegmansGridkitExtractEntsoE2016} and selected Ten Year Development Plan (TYNDP) projects.
+The topology and capacities of the electricity transmission system are taken from the ENTSO-E transparency map~\cite{wiegmansGridkitExtractEntsoE2016} and selected Ten Year Network Development Plan (TYNDP) projects.
 The power flow is based on the linearized power flow approximation assuming a linear transport efficieny of 95.75\% per 1000~km accounting for resistive losses on the power lines.
@@
 For hydrogen pipelines an average loss of 1.2\% per 1000~km, aligning with industry estimates that account for hydrogen's high permeability and small molecular size, which can lead to higher losses compared to other gases.
-The electrical demand for compression stations is assumed to be 1.8\% of the transported energy.
-For \carbon{} pipelines an average loss rate of 0.8\% per 1000~km is reflecting a generally low leakage rate observed in supercritical or dense phase CO2 transport~\cite{liuExperimentalStudyLeakage2023,vitaliRisksSafetyCO22021}.
-The compression energy needed is set to 250~KWh per 1000~km and tonne \carbon{} to maintain a supercritical state for efficient transport.
+The electrical demand for compression stations is assumed to be 1.8\% per 1000~km of the transported energy.
+For \carbon{} pipelines an average loss rate of 0.8\% per 1000~km is reflecting a generally low leakage rate observed in supercritical or dense phase \carbon{} transport~\cite{liuExperimentalStudyLeakage2023,vitaliRisksSafetyCO22021}.
+The compression energy needed is set to 250~kWh per 1000~km and tonne \carbon{} to maintain a supercritical state for efficient transport.
 For natural gas pipelines, the model assumes a 1\% loss rate per 1000~km, accounting for typical leakage and operational losses in well-maintained natural gas infrastructure as reported in~\cite{NaturalGasTransmission2021}.
-The electrical demand for compression is set at 1\% of the transported energy, which falls within the common range for natural gas pipeline operations~\cite{mcvayreneeMethaneEmissionsGas2023}.
+The electrical demand for compression is set at 1\% per 1000~km of the transported energy, which falls within the common range for natural gas pipeline operations~\cite{mcvayreneeMethaneEmissionsGas2023}.
 %
@@
 Throughout this study, we focus on \carbon{} and \hydrogen{} networks because of their high relevance to infrastructure investments and thus public policy decisions.
-The electricity network and gas network are both included in the model with full geographical detail, but not further analyzed.
-Electricity networks are already in place, and we restrict further extension due to concerns about public acceptance.
+The electricity network and gas network are both included in the model with full geographical detail, but are not further analyzed.
+Electricity networks are already in place, and we restrict further extension to 25\% due to concerns about public acceptance.
 Gas networks also already exist, and will most likely experience decreased use in the future, removing any bottlenecks or constraints on the optimal system buildout or operation.
@@
 This merit order ranks all CC technologies according to their relative costs to capture an additional marginal ton.
-The cost value for a technology depends on the exact system configuration and may vary by scenario and even by region since many technologies produce other (primary) outputs such as electricity and heat.
+The cost value for a technology depends on the exact system configuration and operation and may vary by scenario and even by region since many technologies produce other (primary) outputs such as electricity and heat.
 The costs to capture an additional marginal ton of carbon may also be different from the average cost of capture.
@@
 This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.
-If we consider the spatial aspect of distributed carbon capture potentials, the \carbon{} network can exploit comparative cost advantages between CO2 bidding zones to utilize the lowest-cost CC facilities across the continent.
+If we consider the spatial aspect of distributed carbon capture potentials, the \carbon{} network can exploit comparative cost advantages between \carbon{} bidding zones to utilize the lowest-cost CC facilities across the continent.
 We assume a capture rate of 90\% for CC on process emissions, SMR, biogas-to-gas, as well as gas and biomass used in industry, and 95\% for CC on biomass and gas CHPs.
@@
 We make this choice because offshore sequestration sites typically have a larger capacity compared to onshore sequestration sites in saline aquifers and due to concerns over public safety for infrastructure near populated areas.
-Our estimates for carbon sequestration potentials are restrictive, limiting the annual sequestration potential to 25~Mt per region and constraining it further such that this amount could be  25 years.
+Our estimates for carbon sequestration potentials are restrictive, limiting the annual sequestration potential to 25~Mt per region and constraining it further such that it provides enough volume to allow for 25 years of sequestration.
 Furthermore, we decided to limit the total amount of sequestered \carbon{} to 200~Mt per year for our net-zero \carbon{} emission target and 660~Mt per year for our net-negative emission target.
 This constraint is deliberately implemented to avoid over-reliance on carbon capture and sequestration as a backstop technology, which could otherwise offset fossil emissions instead of prioritizing the reduction of fossil fuel use where technologically feasible.
-By capping the sequestration at 200 Mt/year, we aim to promote a balanced energy transition, where sequestration is used to offset hard-to-abate fossil process emissions (e.g., calcination of limestone in cement manufacturing) but does not become a substitute for more sustainable mitigation measures.
+By capping the sequestration at 200 Mt/a, we aim to promote a balanced energy transition, where sequestration is used to offset hard-to-abate fossil process emissions (e.g., calcination of limestone in cement manufacturing) but does not become a substitute for more sustainable mitigation measures.
 The implications of this choice are discussed in the Appendix in Section~\ref{sec:higher_sequestration}.
-
-This is enough to offset hard-to-abate fossil process emissions (e.g., calcination of limestone in cement manufacturing) and some limited slack for fossil fuels to be used where they can be most valuable.
-This is in order to avoid offsetting fossil emissions with carbon removal if they can technologically be avoided in the first place.
 
@@
 Demand and emissions from industrial clusters heavily drive the modeling results.
-Allowing the model to relocate these, incorporate flexibility measures or alter underlying processes, may lead to less dependency on \carbon{} and \hydrogen{} networks.
+Allowing the model to relocate these, incorporate flexibility measures or alter underlying processes, may lead to less dependency on \carbon{} and \hydrogen{} networks than with the static industry locations assumed in this study.
 The transport of fuel (FT, gas, oil) between the regions is not limited which may overestimate their flexibility.
 However, the distributed CU systems, as enabled through the hydrogen network, align well with today's distributed locations for oil refineries, providing kerosene for aviation and naphtha for industry.
-The assumption to run Fischer-Tropsch facilities on baseload with at least 90\% of their nominal capacity is likely shaping the deployment of CU, underestimating potential important flexibilities.
+Our assumption to run Fischer-Tropsch facilities on baseload with at least 90\% of their nominal capacity is likely shaping the deployment of CU, underestimating potential important flexibilities.
 The assumed cost projections on technologies are subject to uncertainties.
@@
 However, a different set of costs on DAC and electrolysis may well impact the model results.
-Considering different strategies for carbon removal, DAC deployment could be in parts replaced if larger sustainable biomass sources are assumed, as shown in~\cite{lauerCrucialRoleBioenergy2023}.
+Considering different strategies for carbon removal, DAC deployment could be in parts replaced if larger sustainable biomass sources are assumed or imported, as shown in~\cite{lauerCrucialRoleBioenergy2023} and~\cite{millingerDiversityBiomassUsage2023} respectively.
 
@@
 Depending on the type of energy carrier and whether it is \carbon{} neutral or not, allowing imports can change the optimal design and operation of the system quite significantly, as discussed in~\cite{neumannEnergyImportsInfrastructure2024}.
-In our model, we see that when higher sequestration rates are allowed, the model imports more fossil oil, which reduces the use of CU in Central Europe and thus the hydrogen transport from South to Central Europe (see section~\ref{sec:higher_sequestration}).
+In our model, we see that when higher sequestration rates are allowed, the model imports more fossil oil, which reduces the use of CU in Central Europe and thus the hydrogen transport from South to Central Europe (see Section~\ref{sec:higher_sequestration}).
 While this substitution reduces the need for Fischer-Tropsch fuel production, it also requires significant carbon sequestration within Europe to offset the emissions from fossil oil use.
-If the imported oil were carbon neutral or 'green', there would be no need for dedicated sequestration infrastructure as the emissions would already be accounted for.
+If the imported oil were carbon neutral or green, there would be no need for dedicated sequestration infrastructure as the emissions would already be accounted for.
 In such a case, the overall energy infrastructure would remain largely the same, minus the additional carbon sequestration capacity needed to offset the emissions from fossil imports.
@@
 The same applies to the deployment of a carbon network driven by fixed sequestration sites.
+A disadvantage of importing synthetic fuels would be that the significant \carbon{} available in sustainable biomass sources in Europe would be underutilised, and there would be an additional burden of regulating the sourcing of sustainable \carbon{} outside of Europe.
 
@@
 SMR and gas CHPs have capacity factors around 15\%, and CC shares at nearly 0\% for the \baselinescenario{}.
-The capacity factor and CC share of SMR increase to around 20\% and 50-65\%, respectively, in the other models.
+The capacity factor and CC share of SMR increase to around 20\% and 50ß-65\%, respectively, in the other models.
 Gas CHPs serve as peak load electricity production technology, and with very few operating hours, thus the high investment costs into CC applications are not efficient.
@@
 \label{sec:higher_sequestration}
-Carbon management in the optimized energy system, particularly in the hard-to-abate sectors, is heavily influenced by the model's sequestration potential, which is capped at 200~Mt/year.
+Carbon management in the optimized energy system, particularly in the hard-to-abate sectors, is heavily influenced by the model's sequestration potential, which is capped at 200~Mt/a.
 Lifting the sequestration limit leads to cost benefits and a shift in primary energy supply and its implications (see Fig.~\ref{fig:sequestration_cost_bar}).
@@
 Furthermore, the focus on sequestration could perpetuate the use of fossil fuels in sectors where decarbonization alternatives, such as electrification or green hydrogen, may offer more sustainable solutions.
+In particular, if sequestration sites are overused for compensating continuing unabated fossil emissions, they become unavailable for net-negative targets necessary to compensate historical emissions.
 Thus, careful consideration must be given to avoid locking in fossil fuel usage and undermining the broader energy transition goals.

### 2024-09-10 report/submissions/journal/review-responses/review-responses-reviewer-1.md
update colors; insert remaining internal review comments; separate review responses

--- 2024-09-09 report/submissions/journal/review-responses/review-responses-first-round.md
@@
-We thank both reviewers for their helpful comments.
-We have rerun all scenarios in the paper with improved assumptions as suggested by the authors (improved accounting for leakage and compression losses) as well as new scenarios requested by the reviewers for a sensitivity to sequestration rates.
+# Response to Reviewer 1
 
-## Reviewer #1 (Remarks to the Author):
+We thank the reviewers for their helpful comments.
+We have rerun all scenarios in the paper with improved assumptions as suggested by the authors (improved accounting for leakage and compression losses) as well as new scenarios requested by the reviewers for a sensitivity to sequestration rates and technology costs.
 
-@FabianHofmann
+Please let us reply to the reviewers' comments below.
+
+Comment:
+
 > The article presents an interesting analysis of the foreseen H2 and CO2 transport infrastructures.
@@
 
-Thank you for your comments. We appreciate your feedback and will address your concerns in the revised manuscript.
+Thank you for your comments. We appreciate your feedback and have addressed your concerns in the revised manuscript.
 
-We have expanded the literature review section to provide more context for the study, as well as discussing recent policy measures for carbon management in the EU.
+We have expanded the referenced literature to provide more context for the study, as well as discussing recent policy measures for carbon management in the EU.
 Since the study touches a wide range of topics in the political and technological landscape, we have tried to find a good balance between providing a comprehensive overview and keeping the text concise.
-(#TODO)
 
@@
 
----
-@FabianHofmann
+Comment:
+
 > There are several graphs that make the reader confused rather than explain the amount of work done by the authors.
@@
 
----
+Comment:
 
-@nworbmot
 > The results are not understandable in terms of transferring into policies.
@@
 Many of the technologies have not yet been deployed at scale before and will need support for financing, regulation and gaining the acceptance of local populations."
-
----
-
-## Reviewer #2 (Remarks to the Author):
-@FabianHofmann
-
-> Overall, a very impressive paper, with a very clear and concise structure, easy to follow with many interesting insights.
-The research gap and necessity of the paper clearly are demonstrated, the topic is higly relevant and the chosen methodology are well suited for the task.
-
-> There are a couple of aspects I would like to understand better, i.e. aspects in which I think additional information would be beneficial.
-
-> 1: CCS is a challenge for scenario analyses like yours, since it low cost makes it a backstop technology.
-The model will try to solve many things with CCS if one does not prevent somehow.
-Your approach is to limit CS.
-“This is in order to avoid offsetting fossil emissions with carbon removal if they can technologically be avoided in the first place.” I was wondering about this throughout the paper.
-So, first of all I could be helpful to point out the importance and weight of this setting early in the text.
-I think the implications of this should be discussed more, because I believe this to be crucial: At your point of operation, your model is able to provide and sequester CO2 for roughly160 €/t if I’m correct.
-That basically means that all mitigation measures costlier than that are only part of the solution because of the CS limit.
-It would be interesting to discuss the dual value/shadow price of tis cap, to understand how much economic pressure could be on relying more on CCS.
-The fossil lock-in your CS cap prevents could be an economically very attractive.
-So, if compensating fossil fuel usage is cheaper than the chosen path of FT and other PtX options, I think this would deserve a discussion in the paper.
-Especially because the compensation would not need to take place in Europe but could be done in regions with even lower DACS/BECCS costs.
-
-Thank you for your insightful comment.
-In response, we have added an explanation for our choice to the introduction and have added new modelling runs in a new subchapter in the Appendix "Higher Sequestration Limits" which we refer to in the main text.
-The added chapter discusses new model results for higher seqestration limits and their implications.
-
-Our analysis shows that increasing the limit from 200 Mt/year to 800 Mt/year reduces system costs quite significantly, by up to 9.1%.
-This reduction is primarily due to a shift towards greater reliance on fossil fuels either with CCS or unabated and compensated with CDR, which in turn reduces the need for CU and renewable investments by about a third.
-At the same time, the hydrogen network loses importance whereas the carbon network quadruples its capacity.
-Point-sources of carbon in the inland where fossil fuels are used, like gas for heating or naphtha for industry, are nearly all captured and tranpsorted to be sequesterd.
-
-Another interesting finding is the drop of the shadow price from €225/t to €51/t when increasing the sequestration limit.
-Additional sequestration capacity provides economic benefits, these benefits, however, diminish as the limit increases​.
-
-However, this also introduces the risk of reinforcing fossil fuel dependence, potentially undermining long-term decarbonization efforts.
-The discussion in the appendix now highlights these trade-offs, emphasizing the importance of carefully balancing sequestration limits to avoid economic over-reliance on CCS​.
-
----
-
-@chrstphtrs
-> 2.
-The H2 provision in the Baseline case confuses me a bit, namely that it works surprisingly well.
-How is a constant supply possible in regions without geological storage potential?
-Does this mean that in hours of low/no RES power plants in the regions with storages produce electricity, which is then brought to H2 sinks and turned into H2 again?
-In our scenarios this led to much bigger costs and almost impossible necessities for electricity network expansions.
-Is this because your H2 demand is relatively low, and you transform it into other carriers?
-Or do you have some other form of storage option at the consumption sites.
-Overall, I would just like to understand better how you supply H2 steadily without a network.
-
-Thank you for the comment, this is indeed a very interesting question.
-
-The spatially-fixed hydrogen demand from industry (primarily for primary steel but also for ammonia) is comparatively low, while the bulk of the H2 demand (FT, methanolisation, methanation) can be moved by the model to locations with low-cost renewables and geological H2 storage.
-This avoids over-burdening the electricity grid.
-
-In addition, to provide hydrogen in some regions with poor renewable resources the model resorts to SMR with either green or fossil methane transported in the gas grid.
-
----
-@fneum
-> 3: In the introduction: “It is assumed that Europe is self-sufficient in energy and does not import any fuels.” I would like to see the consequences of this setting discussed, especially regarding the large liquid quantities it entails.
-Just a suggestion, but a variation in which half of the quantities are imported discussed briefliy in the Annex would be very valuable.
-Much of the infrastructure you model and discuss discuss hinges on this, and importing a major share is clearly an realistic and attractive option in reality.
-
-This is indeed an important point that we have not discussed in detail in the paper.
-The assumption of Europe's self-sufficiency in energy was made to simplify the system design and focus on internal infrastructure.
-It also has the advantage of making good use of Europe's sustainable biomass sources for sourcing CO2, and avoiding the difficulty of regulating sustainable CO2 provision for carbonaceous fuels synthesised outside of Europe.
-However, we agree that importing a significant share of energy carriers, particularly liquid fuels, is a realistic and potentially cost-beneficial option.
-
-We have expanded the discussion of the limitations in the appendix to explore the potential consequences of allowing imports.
-In the higher sequestration scenario, for example, our model shows that increased imports of fossil oil significantly reduce the need for domestic Fischer-Tropsch (FT) fuel production.
-While this reduces hydrogen transport and carbon utilization requirements in Central Europe, it also necessitates substantial carbon sequestration within Europe to offset emissions from fossil-based imports.
-If these imports were carbon-neutral, the need for dedicated sequestration infrastructure would be significantly reduced, as the emissions would already be compensated.
-Thus, the carbon neutrality of imported fuels has a profound effect on the optimal infrastructure layout, especially for hydrogen transport and carbon sequestration.
-A scenario that allows for partial energy imports could lead to lower investments in the carbon and hydrogen networks as well as direct air capture.
-
-However, a critical political and strategic consideration is the control over the CO2 supply chain.
-There may not be sufficient sustainable CO2 available abroad, and Europe may prefer to maintain control of the CO2 supply chain to ensure security and reliability.
-Relying on external sources for carbon-neutral fuels could introduce geopolitical risks, similar to those faced with fossil fuel imports.
-
----
-@chrstphtrs
-> Your H2 prices seem comparatively low to similar exercises. What interest rate/ WACC did you use? I think that should always be made explicit, as it has huge a impact on many results.
-
-We use technology cost assumptions from the Danish Energy Agency referenced in the Methods section.
-The database applies a discount rate of 7% to derive annualized investment costs for all major technologies, including hydrogen electrolyzers.
-We have added this information to the Methods section.
-
----
-@FabianHofmann
-> Maybe this would be to much, but while seeing your H2 prices I was very interested in seeing your electricity prices. If a map/maps would be too much, maybe state the ranges somewhere for reference?
-
-We have added a chapter to the Appendix that showes maps for other energy carrier (electricity, heat, gas, biomass) in the hybrid scenario.
-These show the average production, consumption, transport and prices per carrier.
-We hope this will provide the information you are looking for.
-
----
-@chrstphtrs
-> I think an overview table with differences of the scenario settings in the beginning would be helpful, and maybe with some key results in the end.
-
-We have added an overview table in the introduction that shows the main differences between the scenarios.
-Since the difference of the scenarios is important on the qualitative level, we have not added a table with key results in the end.
-We hope that the reader can follow the argumentation and the results are clear enough.
-
----
-@FabianHofmann
-> The colour coding in Figure 6 could be better, the greens are almost indistinguishable and too many similar browns.
-
-Thank you for your feedback. We have revised the color scheme in Figure 6 to make it more accessible to the reader. We hope that the new color scheme will make the figure easier to read.
-
----
-@chrstphtrs
-> In the end of section 2: “In the network layouts, the slight shift in CU deployment results in a partial dismantling of hydrogen pipelines, especially from the Iberian Peninsula to Central Europe (0.5 bn€/a).” Is “dismantling” really the right word here?
-The model simply does not build those pipelines in this scenario, and you cannot dismantle what you never built.
-Or maybe you can, and this is a tricky koan.
-
-You are absolutely correct. We rephrased this passage, and similar ones, that the the net-zero and net-negative scenarios are modelled as different futures, not as chronological events.
-
----
-@chrstphtrs
-> “High-cost Direct Air Capture”: I may be nitpicking, but I was wondering whether this expression is really fair in combination with the CS cap.
-I think you force the model to pursue many even more expensive option like FT, this is just the most expensive carbon option for sequestration.
-Again, really just a thought.
-
-You are correct, the relative statement of low and high cost only applies to elementary carbon dioxide needed for sequestration. We amended the statement to make it more clear.

### 2024-09-10 report/submissions/journal/review-responses/review-responses-reviewer-2.md
update colors; insert remaining internal review comments; separate review responses

--- 2024-09-09 report/submissions/journal/review-responses/review-responses-first-round.md
@@
+# Response to Reviewer 2
+
 We thank both reviewers for their helpful comments.
@@
 
-## Reviewer #1 (Remarks to the Author):
+Please let us reply to the reviewers' comments below.
 
-@FabianHofmann
-> The article presents an interesting analysis of the foreseen H2 and CO2 transport infrastructures.
-> Literature is not accurate and there are many references cited without going in their details.
-> The study lacks of realistic elements such as leakage issues, location of compression stations in relation with the capture location.
-
-Thank you for your comments. We appreciate your feedback and will address your concerns in the revised manuscript.
-
-We have expanded the literature review section to provide more context for the study, as well as discussing recent policy measures for carbon management in the EU.
-Since the study touches a wide range of topics in the political and technological landscape, we have tried to find a good balance between providing a comprehensive overview and keeping the text concise.
-(#TODO)
-
-To simulate a more realistic transport system, the following assumptions have been integrated into the energy system model for transport efficiencies and compression requirements.
-For hydrogen pipelines an average loss of 1.2% per 1000 km, aligning with industry estimates that account for hydrogen's high permeability and small molecular size, which can lead to higher losses compared to other gases​.
-The compression energy required is assumed to be 1.8% of the transported energy, which is consistent with the significant energy demands for compressing hydrogen due to its low energy density​ [Environmental Defense Fund​](https://www.edf.org/sites/default/files/documents/Pipeline%20Methane%20Leaks%20Report.pdf).
-For natural gas pipelines, the model assumes a 1% loss rate per 1000 km, accounting for typical leakage and operational losses in well-maintained natural gas infrastructure​ as reported in [Global Energy Monitor](https://www.gem.wiki/Natural_gas_transmission_leakage_rates).
-The compression requirement is set at 1%, which falls within the common range for natural gas pipeline operations​ [Environmental Defense Fund​](https://www.edf.org/sites/default/files/documents/Pipeline%20Methane%20Leaks%20Report.pdf)​.
-For CO2 pipelines an average loss rate of 0.8% per 1000 km is reflecting a generally low leakage rate observed in supercritical or dense phase CO2 transport​ [SpringerLink](https://link.springer.com/article/10.1007/s11356-023-27289-3) [MDPI](https://www.mdpi.com/1996-1073/14/15/4601)​.
-The compression energy needed is set to 250 kWh per 1000km and tonne CO2 reflecting the energy required to compress and maintain CO2 in a supercritical state for efficient transport​.
-
----
-@FabianHofmann
-> There are several graphs that make the reader confused rather than explain the amount of work done by the authors.
-
-We have revised the layout of the figures to make them more accessible to the reader and have adjusted the color scheme where colors tended to be quite similar.
-However, we would appreciate more specific feedback on which figures you find confusing so that we can address these issues more effectively.
-Most of the figures contain a lot of information that we feel is necessary to support the argument.
-We have tried to make them as clear as possible.
-
----
-
-@nworbmot
-> The results are not understandable in terms of transferring into policies.
-
-We have written a short policy outlook in the conclusion as well as linking in the introduction to the recently-announced industrial carbon management strategy of the European Commission.
-Our addition to the conclusion reads:
-
-"For policy makers these results show the need for coordinated planning across sectors like carbon, hydrogen and synthetic fuels since they are strongly interacting, as well as across borders because of the localized nature of good renewable resources and sequestration potentials.
-Deploying multiple networks offers some robustness should unforeseen problems arise with one of them, and our results show that the system can even cope with neither a CO2 nor H2 network for a cost penalty.
-Many of the technologies have not yet been deployed at scale before and will need support for financing, regulation and gaining the acceptance of local populations."
-
----
-
-## Reviewer #2 (Remarks to the Author):
-@FabianHofmann
+Comment:
 
@@
 
----
+Comment:
 
-@chrstphtrs
 > 2.
@@
 
----
-@fneum
+Comment:
+
 > 3: In the introduction: “It is assumed that Europe is self-sufficient in energy and does not import any fuels.” I would like to see the consequences of this setting discussed, especially regarding the large liquid quantities it entails.
@@
 
----
-@chrstphtrs
+Comment:
+
 > Your H2 prices seem comparatively low to similar exercises. What interest rate/ WACC did you use? I think that should always be made explicit, as it has huge a impact on many results.
@@
 
----
-@FabianHofmann
+Comment:
+
 > Maybe this would be to much, but while seeing your H2 prices I was very interested in seeing your electricity prices. If a map/maps would be too much, maybe state the ranges somewhere for reference?
@@
 
----
-@chrstphtrs
+Comment:
+
 > I think an overview table with differences of the scenario settings in the beginning would be helpful, and maybe with some key results in the end.
@@
 
----
-@FabianHofmann
+Comment:
+
 > The colour coding in Figure 6 could be better, the greens are almost indistinguishable and too many similar browns.
@@
 
----
-@chrstphtrs
+Comment:
+
 > In the end of section 2: “In the network layouts, the slight shift in CU deployment results in a partial dismantling of hydrogen pipelines, especially from the Iberian Peninsula to Central Europe (0.5 bn€/a).” Is “dismantling” really the right word here?
@@
 
----
-@chrstphtrs
+Comment:
+
 > “High-cost Direct Air Capture”: I may be nitpicking, but I was wondering whether this expression is really fair in combination with the CS cap.

### 2024-09-10 report/submissions/journal/paper/paper.tex
update colors; insert remaining internal review comments; separate review responses

@@
 We explore both networks' potential synergies and competition in a cost-optimal carbon-neutral European energy system.
-In direct comparison, a hydrogen network is more cost-effective than a carbon network, since it serves to transport hydrogen to demand and to point source of carbon for utilization.
+In a direct comparison, a hydrogen network is more cost-effective than a carbon network, as it serves to transport hydrogen to demand and to point source of carbon for utilization.
 However, in a hybrid scenario where both networks are present, the carbon network effectively complements the hydrogen network, promoting carbon capture from distributed biomass and reducing reliance on direct air capture.
-The layouts of the hydrogen and carbon dioxide networks are robust if the climate target is tightened to be net negative.
+The layouts of the hydrogen and carbon dioxide networks are robust if the climate target is tightened to be net-negative.
 \end{abstract}
@@
 The Renewable Energy Directive III~\cite{DirectiveEU20232023} aims to produce 10~Mt hydrogen from renewable energy sources domestically in the EU.
-The Net-Zero Industry Act~\cite{eu2023netzero} proposes a \carbon{} injection target of 50~Mt per year by 2030, which will be sequestered within the European Union.
+The Net-Zero Industry Act~\cite{eu2023netzero} proposes a \carbon{} injection target of 50~Mt/a by 2030, which will be sequestered within the European Union.
 The European Commission's 2024 industrial carbon management strategy~\cite{comcms} plans for capture of 450~Mt/a by 2050 in the EU, of which almost 250~Mt/a is sequestered.
@@
 Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, and synthetic fuels on the one hand and offer to purchase \carbon{} on the other hand.
-The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning a transport and sequestration capacity of 1.5~Mt \carbon{} per year to be operative in 2024, expanding to a targeted scale of 5~Mt per year of sequestration by 2030.
+The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning a transport and sequestration capacity of 1.5~Mt/a \carbon{} to be operative in 2024, expanding to a targeted scale of 5~Mt/a of sequestration by 2030.
 The market potential for carbon capture and sequestration in Europe is underscored by the Capture Map's estimated potential of 1.7 Gt carbon annually from point sources~\cite{ToolsGreenTransition}, accounting for about half of the continent's emissions, as well as estimated sequestration capacity potentials of up to 3~Gt/a~\cite{europeancommissionEuropeanCO2Storage}.
@@
 A hydrogen network can supply regions with geographically fixed hydrogen demand, such as for steel production, with hydrogen from regions with the best renewable resources as well as enable CU at the site of point-source CC.
-On the other hand, the carbon transport system allows transporting captured carbon to regions with sequestration potentials or high-quality renewable resources, the latter enabling cost-effective CU.
+On the other hand, the carbon transport system allows transporting captured carbon to regions with sequestration potentials or high-quality renewable resources, the latter enabling cost-effective CU.\@
 
@@
 Such techno-economic models, in contrast for example to integrated Assessment Models, can account for the spatial distribution of carbon sources and sinks, which are crucial to provide a holistic view of the energy system and its technological interactions.
-However, all of the techno-economic studies have only dealt with isolated aspects of hydrogen and carbon networks.
+However, all the techno-economic studies have only dealt with isolated aspects of hydrogen and carbon networks.
 Neumann et al.~\cite{neumannBenefitsHydrogenNetwork2022} demonstrated the extent to which hydrogen and electricity grid expansion is interchangeable in a climate-neutral European energy system at the expense of higher investments.
-The underlying, highly-resolved model encompasses the most relevant sectors, however it neglects the infrastructure required to transport carbon accross regions.
+The underlying, highly-resolved model encompasses the most relevant sectors, however it neglects the infrastructure required to transport carbon across regions.
 Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \carbon{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
 This limited sectoral scope cannot capture important dynamics of carbon management, since it neglects sectors like industry that will need to handle most \carbon{} in the future.
-The European Commission's Joint Research Center analysed the possible topology of a \carbon{} network in 2024~\cite{jrc2024}, but did not include the interaction of  \carbon{} sources and sinks with the rest of the energy system.
-Becattini et al~\cite{becattiniCarbonDioxideCapture2022} present a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connected to a remote sequestration site in Norway.
+The European Commission's Joint Research Center analyzed the possible topology of a \carbon{} network in 2024~\cite{jrc2024}, but did not include the interaction of  \carbon{} sources and sinks with the rest of the energy system.
+Becattini et al.~\cite{becattiniCarbonDioxideCapture2022} present a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connected to a remote sequestration site in Norway.
 However, this limited spatial scope fails to consider other sequestration sites and the co-benefits from connecting the \carbon{} network to neighboring countries.
@@
 ``Gas Infrastructure'' combines gas facilities for transport, power and heat production, ``\carbon{} Infrastructure'' and ``\hydrogen{} Infrastructure'' combine transport and storage for each carrier.
-``Carbon Capture at Point Sources'' combines all technologies with integrated carbon capture, including the cost of the main facility (e.g., CHP units) and the carbon capture application.}
+``Carbon Capture at Point Sources'' combines all technologies with integrated carbon capture, including the cost of the main facility (e.g., Combined Heat and Power units) and the carbon capture application.}
     \label{fig:cost_bar}
@@
 Where fossil carbon from point sources cannot be sequestered, it has to be compensated elsewhere by negative emissions from DAC or biomass capture with sequestration.
-To provide hydrogen in some regions with poor renewable resources the model resorts to SMR with either green or fossil methane transported in the gas grid.All of these options are sub-optimal from a cost perspective.
+To provide hydrogen in some regions with poor renewable resources the model resorts to SMR with either green or fossil methane transported in the gas grid.
+All of these options are suboptimal from a cost perspective.
 
@@
 In addition, pipelines transport rather small amounts of carbon from spatially fixed point sources to CU facilities.
-A notable outlier is the route from the Netherlands via the UK and France to large methanolisation and Fischer-Tropsch plants in northern Spain.
+A notable outlier is the route from the Netherlands via the UK and France to large methanolization and Fischer-Tropsch plants in northern Spain.
 The hydrogen routes are similar to those in the \hydrogenscenario{} with transports from Western regions (the Iberian Peninsula and the British Isles) to distributed CU production sites across the continent (see Fig.~\ref{fig:balance_map_full}, right).
@@
 The exceptions are SMR and gas-fired CHP plants, which operate only a few weeks annually, making carbon capture units economically unfeasible for these applications.
-Due to the better integration of point sources, only a few DAC facilities are installed, such as those used for methanolisation in Spain.
+Due to the better integration of point sources, only a few DAC facilities are installed, such as those used for methanolization in Spain.
 This, in turn, leads to reduced power consumption and less wind and solar power deployment than in the other scenarios.
@@
 
-For policy makers these results show the need for coordinated planning across sectors like carbon, hydrogen and synthetic fuels since they are strongly interacting, as well as across borders because of the localized nature of good renewable resources and sequestration potentials.
+For policymakers these results show the need for coordinated planning across sectors like carbon, hydrogen and synthetic fuels since they are strongly interacting, as well as across borders because of the localized nature of good renewable resources and sequestration potentials.
 Deploying multiple networks offers some robustness should unforeseen problems arise with one of them, and our results show that the system can even cope with neither a \carbon{} nor \hydrogen{} network for a cost penalty.
@@
 Many of the assumptions contained therein are based on the technology catalogs published by the Danish Energy Agency~\cite{danishenergyagencyTechnologyDataGeneration2019,thedanishenergyagencyTechnologyDataCarbon2023}.
-Endogenous model results include the expansion of renewable energy sources, storage technologies, Power-To-X conversion and transmission capacities, heating technologies, peaking power plants, and the deployment of gray, blue or green hydrogen production facilitates, among others.
-The model considers various energy carriers like electricity, heat, hydrogen, methane, methanol, liquid hydrocarbons and biomass, together with the corresponding conversion technologies.
+Endogenous model results include expansion of renewable energy sources, storage technologies, Power-To-X conversion and transmission capacities, heating technologies, peaking power plants, and the deployment of gray, blue or green hydrogen production facilitates, among others.
+The model considers various energy carriers like electricity, heat, hydrogen, methane, methanol, liquid hydrocarbons, and biomass, together with the corresponding conversion technologies.
 
@@
 Heat demand is regionally subdivided into shares of urban, rural and industrial sites.
-The system exogenously produces 633~Mt \carbon{} per year from industry, aviation, shipping and agriculture, 153~Mt of which are fossil-based process emissions.
+The system exogenously produces 633~Mt/a \carbon{} from industry, aviation, shipping and agriculture, 153~Mt of which are fossil-based process emissions.
 Locations for industrial clusters are taken from~\cite{hotmaps_industrial_db}. Energy demand for industries is calculated from~\cite{mantzosJRCIDEES20152018}.
@@
 
-Low-carbon electricity is provided by wind, solar, biomass, hydro-electricity and nuclear power plants.
+Low-carbon electricity is provided by wind, solar, biomass, hydroelectricity and nuclear power plants.
 Hydroelectric and nuclear plants cannot be extended beyond their currently installed capacities.
@@
 We limit regional biomass use to the medium-level potentials from the JRC-ENSPRESO database~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}.
-Inter-regional biomass transport is permitted, with transport costs of $\sim$0.1~€/MWh/km considered.
+Interregional biomass transport is permitted, with transport costs of $\sim$0.1~€/MWh/km considered.
 
@@
 The topology and capacities of the electricity transmission system are taken from the ENTSO-E transparency map~\cite{wiegmansGridkitExtractEntsoE2016} and selected Ten Year Network Development Plan (TYNDP) projects.
-The power flow is based on the linearized power flow approximation assuming a linear transport efficieny of 95.75\% per 1000~km accounting for resistive losses on the power lines.
+The power flow is based on the linearized power flow approximation assuming a linear transport efficiency of 95.75\% per 1000~km accounting for resistive losses on the power lines.
 %TODO find reference
@@
 Our estimates for carbon sequestration potentials are restrictive, limiting the annual sequestration potential to 25~Mt per region and constraining it further such that it provides enough volume to allow for 25 years of sequestration.
-Furthermore, we decided to limit the total amount of sequestered \carbon{} to 200~Mt per year for our net-zero \carbon{} emission target and 660~Mt per year for our net-negative emission target.
+Furthermore, we decided to limit the total amount of sequestered \carbon{} to 200~Mt/a for our net-zero \carbon{} emission target and 660~Mt/a for our net-negative emission target.
 This constraint is deliberately implemented to avoid over-reliance on carbon capture and sequestration as a backstop technology, which could otherwise offset fossil emissions instead of prioritizing the reduction of fossil fuel use where technologically feasible.
@@
 \label{sec:limitations}
-Despite the comprehensive nature of the model, some limitations affect the validity of the results, as well as important discussion points that need to be put into context.
+While the model is comprehensive, it has certain limitations that affect the validity of its results.
+Additionally, there are important aspects of the findings that require further context and discussion.
 In reality, energy demand and renewable supply can only be estimated for the next few hours/days.
@@
 While this substitution reduces the need for Fischer-Tropsch fuel production, it also requires significant carbon sequestration within Europe to offset the emissions from fossil oil use.
-If the imported oil were carbon neutral or green, there would be no need for dedicated sequestration infrastructure as the emissions would already be accounted for.
+If the imported oil were carbon-neutral or green, there would be no need for dedicated sequestration infrastructure as the emissions would already be accounted for.
 In such a case, the overall energy infrastructure would remain largely the same, minus the additional carbon sequestration capacity needed to offset the emissions from fossil imports.
-Thus both the \hydrogen{} and \carbon{} networks would be less relevant.
-In the context of green hydrogen imports, as highlighted by~\cite{neumannEnergyImportsInfrastructure2024}, the model is expected to reduce the use of electrolysis and methanolisation in southern European countries, as well as hydrogen transport routes from there to central Europe.
+Thus, both the \hydrogen{} and \carbon{} networks would be less relevant.
+In the context of green hydrogen imports, as highlighted by~\cite{neumannEnergyImportsInfrastructure2024}, the model is expected to reduce the use of electrolysis and methanolization in southern European countries, as well as hydrogen transport routes from there to central Europe.
 However, the deployment of CU near point sources of \carbon{} would still be as in our scenario.
 The same applies to the deployment of a carbon network driven by fixed sequestration sites.
-A disadvantage of importing synthetic fuels would be that the significant \carbon{} available in sustainable biomass sources in Europe would be underutilised, and there would be an additional burden of regulating the sourcing of sustainable \carbon{} outside of Europe.
+A disadvantage of importing synthetic fuels would be that the significant \carbon{} available in sustainable biomass sources in Europe would be underutilized, and there would be an additional burden of regulating the sourcing of sustainable \carbon{} outside of Europe.
 
@@
     \caption{System costs for all models in the net-zero emission scenario with 50\% subsidy of \carbon{} pipelines (a) and change in system cost for the \carbongrid{} and \hybridscenario{} when subsidizing \carbon{} pipelines by 50\% (b).}
+\end{figure}
+
+\subsubsection{Higher Electrolysis Costs}
+
+The effect of a 50\% increase in electrolysis investment costs on the optimal technology deployment is shown in Fig.~\ref{fig:cost_bar_electrolysis_co2} for the \carbonscenario{}, Fig.~\ref{fig:cost_bar_electrolysis_h2} for the \hydrogenscenario{}, and Fig.~\ref{fig:cost_bar_electrolysis_full} for the \hybridscenario{}.
+The total system costs increase by 1.7\% in the \carbonscenario{}, by 1.8\% in the \hydrogenscenario{} and the \hybridscenario{}.
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=0.5\linewidth]{comparison/updated-electrolisys-capex-co2-only/figures/90_nodes/cost_bar.png}
+    \caption{System costs for the \carbonscenario{} when increasing electrolysis investment costs by 50\%.}
+    \label{fig:cost_bar_electrolysis_co2}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=0.5\linewidth]{comparison/updated-electrolisys-capex-h2-only/figures/90_nodes/cost_bar.png}
+    \caption{System costs for the \hydrogenscenario{} when increasing electrolysis investment costs by 50\%.}
+    \label{fig:cost_bar_electrolysis_h2}
+\end{figure}
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=0.5\linewidth]{comparison/updated-electrolisys-capex-full/figures/90_nodes/cost_bar.png}
+    \caption{System costs for the \hybridscenario{} when increasing electrolysis investment costs by 50\%.}
+    \label{fig:cost_bar_electrolysis_full}
 \end{figure}
@@
     \end{subfigure}
-    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors in the \carbonscenario{} in the net negative scenario.
+    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors in the \carbonscenario{} for net-negative emission targets.
 For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.
@@
     \end{subfigure}
-    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors in the \hydrogenscenario{} in the net negative scenario.
+    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors in the \hydrogenscenario{} for net-negative emission targets.
 For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices.

### 2024-09-10 workflow/scripts/plot_cost_bar.py

@@ -98,7 +98,7 @@ for i, (val, y) in enumerate(zip(decrease, total)):
         ax.text(
             i,
             y + pad,
-            f"{-val:.1%}",
+            f"{-val:+.1%}",
             ha="center",
             va="bottom",
             fontsize=7,

### 2024-09-10 workflow/scripts/plot_cost_bar_diff.py

@@ -63,7 +63,7 @@ for container in ax.containers:
             label_type="center",
             fontsize=7,
             color="grey",
-            labels=diff.round(0).astype(int),
+            labels=f"{diff:+.0g}",
         )
 
 ax.vlines(0, -0.5, len(diff) - 0.5, color="k", lw=0.5)

### 2024-09-10 report/submissions/journal/review-responses/review-responses-reviewer-1.md
TAG: RESUBMISSION (10.09.24) / update reviewer response style

@@
-# Response to Reviewer 1
+---
+title: "Response to Reviewer 1"
+author: "Fabian Hofmann"
+date: "September 10, 2024"
+geometry: margin=1in
+fontsize: 11pt
+mainfont: Helvetica
+monofont: Menlo
+header-includes:
+  - \usepackage{fancyhdr}
+  - \pagestyle{fancy}
+  - \fancyhead[L]{Response to Reviewer 1}
+  - \fancyhead[R]{\thepage}
+  - \fancyfoot[C]{}
+---
 
@@
 
-Comment:
+**Comment:**
 
-> The article presents an interesting analysis of the foreseen H2 and CO2 transport infrastructures.
-> Literature is not accurate and there are many references cited without going in their details.
-> The study lacks of realistic elements such as leakage issues, location of compression stations in relation with the capture location.
+> _The article presents an interesting analysis of the foreseen H2 and CO2 transport infrastructures.
+Literature is not accurate and there are many references cited without going in their details.
+The study lacks of realistic elements such as leakage issues, location of compression stations in relation with the capture location._
 
@@
 
-Comment:
+**Comment:**
 
-> There are several graphs that make the reader confused rather than explain the amount of work done by the authors.
+> *There are several graphs that make the reader confused rather than explain the amount of work done by the authors.*
 
@@
 
-Comment:
+**Comment:**
 
-> The results are not understandable in terms of transferring into policies.
+> *The results are not understandable in terms of transferring into policies.*
 

### 2024-09-10 report/submissions/journal/review-responses/review-responses-reviewer-2.md
TAG: RESUBMISSION (10.09.24) / update reviewer response style

@@
-# Response to Reviewer 2
+---
+title: "Response to Reviewer 2"
+author: "Fabian Hofmann"
+date: "September 10, 2024"
+geometry: margin=1in
+fontsize: 11pt
+mainfont: Helvetica
+monofont: Menlo
+header-includes:
+  - \usepackage{fancyhdr}
+  - \pagestyle{fancy}
+  - \fancyhead[L]{Response to Reviewer 2}
+  - \fancyhead[R]{\thepage}
+  - \fancyfoot[C]{}
+---
 
@@
 
-Comment:
+**Comment:**
 
-> Overall, a very impressive paper, with a very clear and concise structure, easy to follow with many interesting insights.
-The research gap and necessity of the paper clearly are demonstrated, the topic is higly relevant and the chosen methodology are well suited for the task.
+> *Overall, a very impressive paper, with a very clear and concise structure, easy to follow with many interesting insights.
+The research gap and necessity of the paper clearly are demonstrated, the topic is higly relevant and the chosen methodology are well suited for the task.*
 
-> There are a couple of aspects I would like to understand better, i.e. aspects in which I think additional information would be beneficial.
+> *There are a couple of aspects I would like to understand better, i.e. aspects in which I think additional information would be beneficial.*
 
-> 1: CCS is a challenge for scenario analyses like yours, since it low cost makes it a backstop technology.
+> *1: CCS is a challenge for scenario analyses like yours, since it low cost makes it a backstop technology.
 The model will try to solve many things with CCS if one does not prevent somehow.
@@
 So, if compensating fossil fuel usage is cheaper than the chosen path of FT and other PtX options, I think this would deserve a discussion in the paper.
-Especially because the compensation would not need to take place in Europe but could be done in regions with even lower DACS/BECCS costs.
+Especially because the compensation would not need to take place in Europe but could be done in regions with even lower DACS/BECCS costs.*
 
@@
 
-Comment:
+**Comment:**
 
-> 2.
+> *2.
 The H2 provision in the Baseline case confuses me a bit, namely that it works surprisingly well.
@@
 Or do you have some other form of storage option at the consumption sites.
-Overall, I would just like to understand better how you supply H2 steadily without a network.
+Overall, I would just like to understand better how you supply H2 steadily without a network.*
 
@@
 
-Comment:
+**Comment:**
 
-> 3: In the introduction: “It is assumed that Europe is self-sufficient in energy and does not import any fuels.” I would like to see the consequences of this setting discussed, especially regarding the large liquid quantities it entails.
+> *3: In the introduction: “It is assumed that Europe is self-sufficient in energy and does not import any fuels.” I would like to see the consequences of this setting discussed, especially regarding the large liquid quantities it entails.
 Just a suggestion, but a variation in which half of the quantities are imported discussed briefliy in the Annex would be very valuable.
-Much of the infrastructure you model and discuss discuss hinges on this, and importing a major share is clearly an realistic and attractive option in reality.
+Much of the infrastructure you model and discuss discuss hinges on this, and importing a major share is clearly an realistic and attractive option in reality.*
 
@@
 
-Comment:
+**Comment:**
 
-> Your H2 prices seem comparatively low to similar exercises. What interest rate/ WACC did you use? I think that should always be made explicit, as it has huge a impact on many results.
+> *Your H2 prices seem comparatively low to similar exercises. What interest rate/ WACC did you use? I think that should always be made explicit, as it has huge a impact on many results.*
 
@@
 
-Comment:
+**Comment:**
 
-> Maybe this would be to much, but while seeing your H2 prices I was very interested in seeing your electricity prices. If a map/maps would be too much, maybe state the ranges somewhere for reference?
+> *Maybe this would be to much, but while seeing your H2 prices I was very interested in seeing your electricity prices.
+If a map/maps would be too much, maybe state the ranges somewhere for reference?*
 
@@
 
-Comment:
+**Comment:**
 
-> I think an overview table with differences of the scenario settings in the beginning would be helpful, and maybe with some key results in the end.
+> *I think an overview table with differences of the scenario settings in the beginning would be helpful, and maybe with some key results in the end.*
 
@@
 
-Comment:
+**Comment:**
 
-> The colour coding in Figure 6 could be better, the greens are almost indistinguishable and too many similar browns.
+> *The colour coding in Figure 6 could be better, the greens are almost indistinguishable and too many similar browns.*
 
@@
 
-Comment:
+**Comment:**
 
-> In the end of section 2: “In the network layouts, the slight shift in CU deployment results in a partial dismantling of hydrogen pipelines, especially from the Iberian Peninsula to Central Europe (0.5 bn€/a).” Is “dismantling” really the right word here?
+> *In the end of section 2: “In the network layouts, the slight shift in CU deployment results in a partial dismantling of hydrogen pipelines, especially from the Iberian Peninsula to Central Europe (0.5 bn€/a).” Is “dismantling” really the right word here?
 The model simply does not build those pipelines in this scenario, and you cannot dismantle what you never built.
-Or maybe you can, and this is a tricky koan.
+Or maybe you can, and this is a tricky koan.*
 
@@
 
-Comment:
+**Comment:**
 
-> “High-cost Direct Air Capture”: I may be nitpicking, but I was wondering whether this expression is really fair in combination with the CS cap.
+> *“High-cost Direct Air Capture”: I may be nitpicking, but I was wondering whether this expression is really fair in combination with the CS cap.
 I think you force the model to pursue many even more expensive option like FT, this is just the most expensive carbon option for sequestration.
-Again, really just a thought.
+Again, really just a thought.*
 
